### 2012-03-26 paper.tex
Added existing files.

\documentclass[a4paper,10pt]{article}

\usepackage{mathbf-abbrevs}
\input{defs}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

\usepackage[square,comma,numbers,sort&compress]{natbib}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Carrier phase and amplitude estimation using pilots and data}
\author{Robby McKilliam}

\begin{document}

\maketitle

In this document we describe estimators of carrier phase and signal amplitude of a transmitted communications signal that contains both pilots, known to the receiver, and data, unknown to the receiver.  We focus on signaling constellations that have symbols lying on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and M-PSK.

In this setting, the transmitted symbols take the form,
\[
s_i = \exp\left( j\frac{2\pi}{M} u_i \right),
\]
where $j = \sqrt{-1}$ and $u_i$ is an integer in the set $\{0, 1, \dots, M-1\}$ and $M$ is the size of the constellation, i.e. we have assumed an $M$-PSK constellation.  We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.  So,
\[
s_i = \begin{cases}
p_i & i \in P, \\
d_i & i \in D, \\
0 & otherwise,
\end{cases}
\]
where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.  The sets $P$ and $D$ are disjoint, i.e. $P \cap D = \emptyset$, and the union $P \cup D$ contains all those indices where the transmitter is active.  Outside $P \cup D$ the transmitter is silent (not active).

We assume that time offset estimation (symbol timing) and matched filtering have already been performed.  The recieved symbols are then given by,
\begin{equation}\label{eq:sigmod}
y_i = a_0 s_i + w_i, \qquad i \in P \cup D,
\end{equation}
where $w_i$ is noise and $a_0$ is a complex number representing both carrier phase and signal amplitude.  Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.  Complicating matters is that the data symbols $d_i$ are not known to the reciever.

Estimation problems of this type have undergone extensive prior study~\cite{Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.  The existing literature mostly considers what is called \emph{noncoherent detection} where no pilots symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).  In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.  Arguably, the literature culminates in the paper of Mackenthun~\cite{Mackenthun1994} who described a least squares estimator (maximum likelihood under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian) requiring only $O(L \log L)$ arithmetic operations, where $L = \abs{P \cup D}$ is the total number of symbols transmitted.  Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.

Here, we will describe a new algorithm for computing the least squares estimator that requires only $O(L)$ operations.  We also consider a slightly more general model than is usual, one that allows the utilisation of pilot symbols, i.e., \emph{coherent detection}.  Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.

In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and the complex gain $a_0$ is a nuisance parameter.  The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.  Whilst estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite approach here.  Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.  This is motivated by the fact that in our intended application (narrowband satellite communications) and in most modern communication systems the data symbols are \emph{coded}.  For this reason raw symbol error rate is not of specific interest at this stage.  Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.  The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.  Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $E\abs{a_0 - \hat{a}}^2$ where $E$ denotes the expected value.

\section{Mackentun's least squares estimator}\label{sec:least-squar-estim}

In this section we derive Mackentun's~\cite{Mackenthun1994} least squares estimator for the coherent case.  Mackenthun specifically consider the noncoherent case, but only minor modifications are required to meaninfully include the pilot symbols, and the fundamental ideas in this section come from~\cite{Mackenthun1994}.  For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.  In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.  This complexity arises from the need to sort a list of $\abs{D}$ elements.  In Section~\ref{sec:line-time-algor} we will show that a full sort is not neccesary, and that the least squares estimator can be implemented in $O(L)$ operations.

Define the sum of squares function
\begin{align}
SS(a, \{d_i, i \in D\}) &= \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
&= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*. \label{eq:SS}
\end{align}
Fixing the $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find that the least squares estimator of $a_0$, satisfies
\begin{equation}\label{eq:hata}
\hat{a} = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{Y}{L}
\end{equation}
where $L = \abs{P \cup D}$ is the total number of symbols transmitted, and to simplify our notation we put 
\[
Y = \sum_{i \in P \cup D} y_i s_i^* = \sum_{i \in P } y_i p_i^* + \sum_{i \in D } y_i d_i^*
\]  
Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to surpress the argument $(\{ d_i, i \in D\})$ for notational clarity.  Substituting $\hat{a} = \frac{Y}{L}$ into~\eqref{eq:SS} we obtain the sum of squares function, conditioned on minimisation with respect to $a$,
\begin{equation}\label{eq:SSdatasymbols}
SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y}^2,
\end{equation}
where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.  Observe that given a candidate value for the data symbols, we can compute the corresponding sum of squares function $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.  We will show that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.  %When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.

To see this, let $a = \rho e^{j\theta}$.  Now the sum of squares function can be written as
\begin{align*}
SS(\rho, \theta, \{d_i, i \in D\}) &= \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \\
&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2.
\end{align*}
For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
\begin{equation}\label{eq:hatdfinxtheta}
\hat{d}_i(\theta) = \exp\left( j\frac{2\pi}{M}\hat{u}_i(\theta) \right) \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\frac{M}{2\pi}\angle( e^{-j\theta}y_i)},
\end{equation}
where $\round{\cdot}$ rounds its argument to the nearest integer\footnote{The direction of rounding for half-integers is not important so long as it is consistent.  The authors have chosen to round up half-integers in their own implementation.}.  Note that $\hat{d}_i(\theta)$ does not depend on the amplitude $\rho$.  As defined, $\hat{u}_i(\theta)$ is not strictly inside the set of integers $\{0, 1, \dots, M-1\}$, but this is not of consequence, as we intend its value to be considered modulo $M$.  With this in mind let $\phi_i = \angle y_i$, then,
\[
\hat{u}_i(\theta) = \round{\frac{M}{2\pi}( \phi_i - \theta )}
\]
which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $M$.

We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Define the integer $b_i = \hat{d}_i(0)$ and the real number 
\[
z_i = \phi_i - \frac{2\pi}{M}\hat{u}_i(0) = \phi_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}\phi_i},
\]
then,
\begin{equation}\label{eq:uicombos}
\hat{d}_i(\theta) = 
\begin{cases}
b_i, &  0 \leq \theta < z_i + \frac{\pi}{M} \\
b_i e^{-j2\pi/M}, & z_i + \frac{\pi}{M} \leq \theta < z_i + \frac{3\pi}{M} \\ 
\vdots & \\
b_i e^{-j2\pi k /M}, & z_i + \frac{\pi(2k - 1)}{M} \leq \theta < z_i + \frac{\pi(2k + 1)}{M}  \\ 
\vdots & \\
b_i e^{-j2\pi}, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
\end{cases}
\end{equation}

Let $f(\theta)$ be the sequence $\{ \hat{d}_i(\theta), i \in D \}$. So, $f(\theta)$ is a function mapping the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.  Observe that $f(\theta)$ is piecewise continuous.  The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains contant are determined by the values of $\{z_i, i \in D\}$.

Let $S$ be the set of all $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$, that is,
\[
S = \{ f(\theta) \mid \theta \in [0, 2 \pi) \}.
\]
It is clear from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i, i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., the minimiser of~\eqref{eq:SSdatasymbols}.  Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.

The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indicies in $D$ such that $z_{\sigma(i)}$ are in ascending order, i.e.
\begin{equation}\label{eq:sigmasortind}
z_{\sigma(i)} \leq z_{\sigma(k)}
\end{equation}
whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}-1\}$.  It is convenient to define the indices into $\sigma$ to be taken modulo $\abs{D}$, that is, if $m$ is an integer not from $\{0, 1, \dots, \abs{D}-1\}$ then we define $\sigma(m) = \sigma(k)$ where $k \equiv m \mod \abs{D}$ and $k \in  \{0, 1, \dots, \abs{D}-1\}$.  The first sequence in $S$ is 
\[
f_0 = f(0) = \{ b_i, i \in D \}
\]  
The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/M}$.  Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.  Using this notation,  
\[
f_1 = f_0 e_{\sigma(0)}.
\] 
The next sequence in $S$ is correspondingly 
\[
f_2 = f_0 e_{\sigma(0)} e_{\sigma(1)} = f_1 e_{\sigma(1)},
\]
and the $k$th sequence is
\begin{equation}\label{eq:fkrec}
f_{k+1} = f_{k} e_{\sigma(k)},
\end{equation}
In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively computed.

We want to find the sequence $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).  A naive approach would be to compute $SS(f_k)$ for each sequence in $S$ and return the minimiser.  Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetics operations.  So, this naive approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.  Mackenthun~\cite{Mackenthun1994} showed how $SS(f_k)$ can be computed recursively.

Let,
\begin{equation}\label{eq:SSfk}
SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
\end{equation}
where, 
\begin{align*}
Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
&= B + \sum_{i \in D}g_{ki},
\end{align*}
where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in the sequence $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.  Letting $g_{k} =\{ y_i f_{ki}^*, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
\[
g_{k+1} = g_{k} e_{\sigma(k)}^*,
\]
where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/M}$.  Now,
\[
Y_0 = B + \sum_{i \in D} g_{0i}
\] 
can be computed in $O(L)$ operations, and
\begin{align*}
Y_1 &= B + \sum_{i \in D} g_{1i} \\
&= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
&= Y_0 + (e^{j2\pi/M} - 1) g_{0\sigma(0)},
\end{align*}
In general we have,
\[
Y_{k+1} = Y_k + (e^{j2\pi/M} - 1) g_{k\sigma(k)}.
\]
So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a constant number of operations.  Given $Y_k$, the value of $SS(f_k)$ can be computed in constant number of operations using~\eqref{eq:SSfk}.  Let $\hat{k}$ be integer the sequence that minimises $SS(f_k)$.  The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata}, as 
\[
\hat{a} = \frac{1}{L} Y_{\hat{k}}.
\]
Psuedocode is given in Algorithm~\ref{alg:loglinear}.  This descibes a practical implementation of Mackenthun's algorithms.  Line~\ref{alg_return} contains the function $\operatorname{sortindicies}$ that, given the sequence $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ of the indices in $D$ as desrcibed in~\eqref{eq:sigmasortind}.  Computing the $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.  This requires $O(L \log L)$ operations.  The $\operatorname{sortindicies}$ is the bottleneck in this algorithm.  All the other lines require only $O(L)$ operations to execute.  In the next sections we will show how to the sortinces function can be avoided.  This leads to an algorithm that requires only $O(L)$ operations.

\begin{algorithm} \label{alg:loglinear}
\SetAlCapFnt{\small}
\SetAlTitleFnt{}
\caption{Mackenthun's least squares estimator}
\dontprintsemicolon
\KwIn{$\{y_i, i \in P \cup D \}$}
\For{$i \in D$}{
$\phi = \angle{y_i}$ \;
$u = \frac{2\pi}{M}\round{\frac{M}{2\pi}\phi} $ \;
$z_i = \phi -  u $ \;
$g_i = y_i e^{-j u}$ \;
}
$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_return} \;
$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \;
$\hat{a} = \frac{1}{L} Y$ \;
$\hat{Q} = \frac{1}{L}\abs{Y}^2$\;
\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$}{
$Y = Y + (e^{j2\pi/M} - 1) g_{\sigma(k)}$ \;
$g_{\sigma(k)} = g_{\sigma(k)} e^{j2\pi/M} $\;
$Q = \frac{1}{L}\abs{Y}^2$\;
\If{$Q > \hat{Q}$}{
 	$\hat{Q} = Q$ \;
 	$\hat{a} =  \frac{1}{L} Y$ \;
 }
}
\Return{$\hat{a}$ \nllabel{alg_return}}
\end{algorithm}

%Let $f$ be a function maping $\reals$ to $\reals^N$ defined so that the $i$th element of $f(\theta)$ is $\round{\frac{M}{2\pi}( \phi_i - \theta )}$.  Let $\phi$ be the vector of length $N$ with element $[\phi_1, \dots, \phi_N]$ and define the $\round{\cdots}$ to operate elementwise on vectors, then
%\[
%f(\theta) = \round{\frac{M}{2\pi}( \phi - \theta )}.
%\]
%The function $f(\theta)$ is piecewise continuous.

\section{A linear-time algorithm}\label{sec:line-time-algor}

\section{Simulations}\label{sec:simulations}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/plot-1.mps}
		\caption{Amplitude error}
		\label{fig:plotamp}
\end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/plot-2.mps}
		\caption{Phase error}
		\label{fig:plotamp}
\end{figure*}

\begin{figure*}[tp]
	\centering
		\includegraphics{code/plot-3.mps}
		\caption{Complex gain error}
		\label{fig:plotamp}
\end{figure*}

\appendix

% \section{Matched filtering: samples to symbols}

% The continuous time transmitted signal takes the form,
% \[
% x(t) = \sum_{i \in \ints} s_i g(t - iT),
% \]
% where $g(t)$ is the transmission pulse shape, $T$ is the symbol period.  Assume that the received signal undergoes time and phase offset, is attenuated by some amount and is also subject to additive noise.  The model for the analogue received signal is,
% \begin{align*}
% r(t) &=  w(t) + \alpha_0 x(t - \tau_0) \\
% &= v(t) + \sum_{i \in \ints} s_i g(t - iT - \tau_0),
% \end{align*}
% where $\alpha_0 \in \complex$ is a complex number representing both attenuation and phase offset, $\tau_0$ is the time offset and $w(t)$ is a continuous noise process.  Assume that an accurate estimate of the time offset $\tau_0$ is available.  We would like to process the signal $r(t)$ so as to obtain a signal in the form of~\eqref{eq:sigmod}.  This is the task of \emph{matched filtering}.

% Consider the inner product
% \begin{align*}
% \dotprod{r(t)}{g(T_sn - Ti -  \tau_0)} &= \dotprod{v(t) + \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)}  \\
% &= w_i +  \dotprod{\alpha_0 \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
% &= w_i +  \alpha_0 \sum_{k \in \ints} s_k \dotprod{ g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
% &= w_i + \alpha_0s_i = y_i
% \end{align*}
% where we put $w_i = \dotprod{v(t)}{g(T_sn - Ti - \tau)}$, and we have required that the transmission pulse $g(t)$ and it's shifts $g(t - Ti)$ are orthogonal, i.e.
% \begin{equation}
% \dotprod{g(t - iT)}{g(t - kT)} = \int_{\infty}^{\infty} g(t - iT)g^*(t-kT) \, dt = \begin{cases} 1 & i=k \\ 0 & i \neq k, \end{cases} \label{eq:contorthpulse}
% \end{equation}
% So, the recieved symbols can be obtained by computing the inner product of the recieved sequence signal with the pulse.

% Thus far, we have assumed continuous processing of a continuous recieved signal.  However, typically in modern communications, only a sampled version of $r(t)$ is available at the reciever, i.e.
% \[
% r_n = r(nT_s) = v_n + \alpha_0 x(T_sn - \tau_0),
% \] 
% where $n$  is an integer and $T_s$ is the sampling period and the sequence $\{v_n, n \in \ints \}$ is a discrete noise process.  In this case, we replace the continuous time inner product, with the discrete time inner product, i.e. put
% \[
% \dotprod{r(t)}{g(t - iT - \tau)} = \sum_{n\in\ints}r(nT_s)g^*(T_sn - Ti - \tau_0) = \sum_{n\in\ints} r_n g^*(T_sn - Ti - \tau_0)
% \]
% The inner product above is an infinite sum and probably cannot be computed in practice.  However, in practical systems the number of samples observed is finite.  In what follows we assume that $N$ consecutive samples $r_1, r_2, \dots, r_N$ are observed and we set $r_n = 0$ whenever $n \notin \{1, 2, \dots, N\}$.  Now, the inner products are always finite sums
% \[
% \dotprod{r(t)}{g(t - iT - \tau_0)} =  \sum_{n=1}^N r_n g^*(T_sn - Ti - \tau_0).
% \]
% These finite inner products are computable in practice.  Computing the inner products directly using the summation above is not necessarily the most efficient approach.  Instead, the fast Fourier transfrom can be used, as we now show.

% We assume that $T_s$ is rationally related to $T$, i.e. $T_s = \frac{p}{q}T$ for some relatively prime integers $p$ and $q$.  This property is a common assumption in the literature.  In fact, it is common to assume the stronger condition, that $T$ is a multiple of $T_s$~\cite{Oerder_synch_square_circstat_2008}.  At the very least, this assumption can be made an accurate approximation by an appropriate choice of $p$ and $q$~\cite{Mordell_Diophantine_equations_1969,Clarkson_thesis}.

% Let 
% \[
% g_n = g^*(-\tfrac{T_s}{p} n - \tau_0).
% \]
% Then
% \[
% g_{k-n} = g^*(-\tfrac{T_s}{p}(k - n) - \tau_0) = g^*(n\tfrac{T_s}{p} - k\tfrac{T_s}{p} - \tau_0) = g^*(n\tfrac{T_s}{p} - \tfrac{k}{q}T - \tau_0)
% \]
% Let
% \begin{equation} \label{eq:zzerofill}
% z_n = \begin{cases} 
% r_{n/c} & n/p \in \{1, \dots, N \} \\
% 0 & \text{otherwise}
% \end{cases}
% \end{equation}
% be a zero filled version of $r_n$.  Now
% \begin{align*}
% h_k &= \dotprod{r(t)}{g(t - \tfrac{k}{q}T - \tau_0)} \\
% &=  \sum_{n\in\ints} r_n g^*(T_sn - \tfrac{k}{q}T - \tau_0) \\
% &= \sum_{n\in\ints} z_n g^*(\tfrac{T_s}{p}n - \tfrac{k}{q}T - \tau_0) \\
% &= \sum_{n\in\ints} z_n g_{k-n}.
% \end{align*}
% So $h_k$ is the convolution between the sequence $\{z_n\}$ and the sequence $\{g_n\}$.  The convoultion can be computed using the fast Fourier transform.  The length of the transform required is
% \[
% BLERG.
% \]  
% It only remains to observe that $y_i = h_{iq}$.

\bibliography{../../bib/bib}

\end{document}

### 2012-03-26 code/scala/test/cam/psk/MackenthunTest.scala

@@ -1,0 +1,47 @@
+
+package cam.psk
+
+import org.junit.Test
+import org.junit.Assert._
+
+import numbers.finite.Complex
+import numbers.finite.PolarComplex
+import numbers.finite.RectComplex
+
+/** 
+ * @author Robby McKilliam
+ */
+class MackenthunTest {
+
+  val rand = new scala.util.Random
+  
+  val a0 = new PolarComplex(1,rand.nextDouble*scala.math.Pi*2)
+  val L = 100
+  val M = 4 //QPSK
+  val s = (1 to L).map(m => new PolarComplex(1, 2*scala.math.Pi*rand.nextInt(M)/M))
+    
+  //pilot position setup
+  val numpilots = 10
+  val P = 0 until numpilots //pilots at the front
+  val D = numpilots until L //data at the back
+  
+  val est = new Mackenthun(M,P,D,s)
+  
+  @Test
+  def testEstimate() {   
+    val iters = 100
+    var mse = 0.0
+    for(i <- 1 to iters) {
+      val std = 0.01
+      val y = s.map(si => a0*si + new RectComplex(std*rand.nextGaussian, std*rand.nextGaussian) ) 
+      val ahat = est.estimate(y)
+      //println(ahat + " and " + a0)
+      val err = (ahat - a0).mag2
+      mse = mse + err
+      assertTrue(err < 0.01)
+    }
+    println(mse/iters)
+  }
+  
+  
+}

### 2012-03-28 paper.tex
Added IEEEbst and fixed other minor stuff.
added ant build script.
Added some refs

@@
 
-In this section we derive Mackentun's~\cite{Mackenthun1994} least squares estimator for the coherent case.
-Mackenthun specifically consider the noncoherent case, but only minor modifications are required to meaninfully include the pilot symbols, and the fundamental ideas in this section come from~\cite{Mackenthun1994}.
+In this section we derive Mackentun's~\cite{Mackenthun1994} least squares estimator.
+Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to meaninfully include the pilot symbols, so we also consider the coherent setting here.
 For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 \end{align}
-Fixing the $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find that the least squares estimator of $a_0$, satisfies
+Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find that the least squares estimator of $a_0$, satisfies
 \begin{equation}\label{eq:hata}
-\hat{a} = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{Y}{L}
+\hat{a} = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y
 \end{equation}
@@
 Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to surpress the argument $(\{ d_i, i \in D\})$ for notational clarity.
-Substituting $\hat{a} = \frac{Y}{L}$ into~\eqref{eq:SS} we obtain the sum of squares function, conditioned on minimisation with respect to $a$,
+Substituting $\hat{a} = \frac{1}{L}Y$ into~\eqref{eq:SS} we obtain the sum of squares function, conditioned on minimisation with respect to $a$,
 \begin{equation}\label{eq:SSdatasymbols}
@@
 \end{equation}
-where $\round{\cdot}$ rounds its argument to the nearest integer\footnote{The direction of rounding for half-integers is not important so long as it is consistent.
+and where $\round{\cdot}$ rounds its argument to the nearest integer\footnote{The direction of rounding for half-integers is not important so long as it is consistent.
 The authors have chosen to round up half-integers in their own implementation.}.
@@
 
+\begin{lemma}
+Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ so that $\hat{\theta}$ is the least squares estimator of the carrier phase and $\hat{\rho}$ is the least squares estimator of the amplitude.  Let 
+\[
+\lambda_i = \phi_i - \hat{\theta} - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\phi_i - \hat{\theta})}.
+\]
+Then 
+\[
+\abs{\lambda_i} \leq \frac{1}{2M} - \frac{1}{2LM BLERG}
+\]
+for all $i \in P \cup D$.
+\end{lemma}
+\begin{proof}
+\end{proof}
+
 \section{Simulations}\label{sec:simulations}
@@
 		\caption{Phase error}
-		\label{fig:plotamp}
+		\label{fig:plotphase}
 \end{figure*}
@@
 		\caption{Complex gain error}
-		\label{fig:plotamp}
+		\label{fig:plotcamp}
 \end{figure*}
@@
 
-\bibliography{../../bib/bib}
+\bibliography{bib}
 

### 2012-03-28 build.sh
moved data so that it sits in it's own file

@@ -1,6 +1,6 @@
-cd code
+cd code/data/
 mpost plot.mp
-cd ..
+cd ../..
 
 pdflatex paper.tex
 bibtex paper

### 2012-03-29 paper.tex
moved data so that it sits in it's own file
Added noncoherent plots.  ViterbiViterbi works very similarly to the Mackenthun algorithm.  It's not really clear how pilots should be included into ViterbiViterbi.  Need to run a simulations with a larger number of symbols to see if this biasing effect continues.
minor mods

@@
 
-The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indicies in $D$ such that $z_{\sigma(i)}$ are in ascending order, i.e.
+The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indicies in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
 \begin{equation}\label{eq:sigmasortind}
@@
 \end{equation}
-In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively computed.
+In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively enumerated.
 
@@
 A naive approach would be to compute $SS(f_k)$ for each sequence in $S$ and return the minimiser.
-Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetics operations.
+Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
 So, this naive approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
@@
 &= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
-&= Y_0 + (e^{j2\pi/M} - 1) g_{0\sigma(0)},
+&= Y_0 + \eta g_{0\sigma(0)},
 \end{align*}
-In general we have,
+where $\eta = (e^{j2\pi/M} - 1)$.  In general we have,
 \[
-Y_{k+1} = Y_k + (e^{j2\pi/M} - 1) g_{k\sigma(k)}.
+Y_{k+1} = Y_k + \eta g_{k\sigma(k)}.
 \]
@@
 Psuedocode is given in Algorithm~\ref{alg:loglinear}.
-This descibes a practical implementation of Mackenthun's algorithms.
-Line~\ref{alg_return} contains the function $\operatorname{sortindicies}$ that, given the sequence $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ of the indices in $D$ as desrcibed in~\eqref{eq:sigmasortind}.
+This descibes a practical implementation of Mackenthun's algorithm.
+Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given the sequence $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ of the indices in $D$ as described in~\eqref{eq:sigmasortind}.
 Computing the $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
 This requires $O(L \log L)$ operations.
-The $\operatorname{sortindicies}$ is the bottleneck in this algorithm.
-All the other lines require only $O(L)$ operations to execute.
+The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm.
+The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all requirse $O(L)$ or less operations.
 In the next sections we will show how to the sortinces function can be avoided.
@@
 \KwIn{$\{y_i, i \in P \cup D \}$}
-\For{$i \in D$}{
+\For{$i \in D$ \nllabel{alg_loop_setup}}{
 $\phi = \angle{y_i}$ \;
@@
 }
-$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_return} \;
-$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \;
+$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
+$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
 $\hat{a} = \frac{1}{L} Y$ \;
-$\hat{Q} = \frac{1}{L}\abs{Y}^2$\;
-\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$}{
+$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
+\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
 $Y = Y + (e^{j2\pi/M} - 1) g_{\sigma(k)}$ \;
@@
 \begin{proof}
+Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
+The proof now proceeds by contradiction.
+Assume that $\lambda_i > \frac{1}{2M} - \frac{1}{2LM BLERG}$ for some $k$.
+Put $r_i = \hat{d}_i e_k$, then
+\begin{align*}
+Y(\{ r_i, i \in D\}) &= \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) \\
+&= \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
+\end{align*}
+where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
+\begin{align*}
+SS(\{r_i, i \in D\}) &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
+&= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
+&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
+&= SS(\{\hat{d}_i, i \in D\}) - R
+\end{align*}
+where 
+\[
+R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
+\]
+We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+
+To see that $R$ is positive write
+\begin{align*}
+R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
+&= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
+\end{align*}
+so it is sufficient to show that
+\[
+\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
+\]
+Unfinished.
+This proof will work, but the problem is that $\abs{y_k}$ can be really small.
+The would appear to neccesitate a very large number of buckets.
+In some sense, this is no problem, but you have to be able to allocate memory for all of these?
+Atleast you need a way of ordering them, up to a certain resolution, without sorting.
+I'm not convinced this will ever be possible.
 \end{proof}

### 2012-03-29 code/runsim.sh
added to data etc

@@ -1,2 +1,3 @@
 #export JAVA_OPTS="-Xprof"
 scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar montecarlo.scala
+scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar montecarlonc.scala

### 2012-04-05 paper.tex
added to data etc
minor edits to paper.

@@
 
+\section{Statistical properties}
+
 %Let $f$ be a function maping $\reals$ to $\reals^N$ defined so that the $i$th element of $f(\theta)$ is $\round{\frac{M}{2\pi}( \phi_i - \theta )}$.
@@
 
-\section{A linear-time algorithm}\label{sec:line-time-algor}
+% \section{A linear-time algorithm}\label{sec:line-time-algor}
 
-\begin{lemma}
-Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ so that $\hat{\theta}$ is the least squares estimator of the carrier phase and $\hat{\rho}$ is the least squares estimator of the amplitude.  Let 
-\[
-\lambda_i = \phi_i - \hat{\theta} - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\phi_i - \hat{\theta})}.
-\]
-Then 
-\[
-\abs{\lambda_i} \leq \frac{1}{2M} - \frac{1}{2LM BLERG}
-\]
-for all $i \in P \cup D$.
-\end{lemma}
-\begin{proof}
-Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
+% \begin{lemma}
+% Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ so that $\hat{\theta}$ is the least squares estimator of the carrier phase and $\hat{\rho}$ is the least squares estimator of the amplitude.  Let 
+% \[
+% \lambda_i = \phi_i - \hat{\theta} - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\phi_i - \hat{\theta})}.
+% \]
+% Then 
+% \[
+% \abs{\lambda_i} \leq \frac{1}{2M} - \frac{1}{2LM BLERG}
+% \]
+% for all $i \in P \cup D$.
+% \end{lemma}
+% \begin{proof}
+% Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
 The proof now proceeds by contradiction.
@@
 Put $r_i = \hat{d}_i e_k$, then
-\begin{align*}
-Y(\{ r_i, i \in D\}) &= \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) \\
-&= \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
-\end{align*}
-where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
-\begin{align*}
-SS(\{r_i, i \in D\}) &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
-&= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
-&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
-&= SS(\{\hat{d}_i, i \in D\}) - R
-\end{align*}
-where 
-\[
-R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
-\]
-We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+% \begin{align*}
+% Y(\{ r_i, i \in D\}) &= \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) \\
+% &= \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
+% \end{align*}
+% where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
+% \begin{align*}
+% SS(\{r_i, i \in D\}) &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
+% &= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
+% &= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
+% &= SS(\{\hat{d}_i, i \in D\}) - R
+% \end{align*}
+% where 
+% \[
+% R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
+% \]
+% We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
 
-To see that $R$ is positive write
-\begin{align*}
-R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
-&= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
-\end{align*}
-so it is sufficient to show that
-\[
-\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
-\]
-Unfinished.
+% To see that $R$ is positive write
+% \begin{align*}
+% R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
+% &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
+% \end{align*}
+% so it is sufficient to show that
+% \[
+% \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
+% \]
+% Unfinished.
 This proof will work, but the problem is that $\abs{y_k}$ can be really small.
@@
 I'm not convinced this will ever be possible.
-\end{proof}
+% \end{proof}
 
@@
 
+\subsection{The coherent setting}
+
+Our first set of simulations considers the case where there are a number of pilot symbols.
+
+%BPSK plots
 \begin{figure*}[tp]
 	\centering
-		\includegraphics{code/plot-1.mps}
-		\caption{Amplitude error}
+		\includegraphics{code/data/plotM2-1.mps}
+		\caption{Amplitude error for BPSK}
 		\label{fig:plotamp}
@@
 	\centering
-		\includegraphics{code/plot-2.mps}
-		\caption{Phase error}
+		\includegraphics{code/data/plotM2-2.mps}
+		\caption{Phase error for BPSK}
 		\label{fig:plotphase}
+\end{figure*}
+
+% \begin{figure*}[tp]
+% 	\centering
+%         \includegraphics{code/data/plotM2-3.mps}
+% 		\caption{Complex gain error for BPSK}
+% 		\label{fig:plotcamp}
+% \end{figure*}
+
+%QPSK plots
+\begin{figure*}[tp]
+	\centering
+		\includegraphics{code/data/plotM4-1.mps}
+		\caption{Amplitude error for QPSK}
+		\label{fig:plotamp}
 \end{figure*}
@@
 	\centering
-		\includegraphics{code/plot-3.mps}
-		\caption{Complex gain error}
-		\label{fig:plotcamp}
+		\includegraphics{code/data/plotM4-2.mps}
+		\caption{Phase error for QPSK}
+		\label{fig:plotphase}
 \end{figure*}
+
+% \begin{figure*}[tp]
+% 	\centering
+%         \includegraphics{code/data/plotM4-3.mps}
+% 		\caption{Complex gain error for QPSK}
+% 		\label{fig:plotcamp}
+% \end{figure*}
+
+%8PSK plots
+\begin{figure*}[tp]
+	\centering
+		\includegraphics{code/data/plotM8-1.mps}
+		\caption{Amplitude error for 8PSK}
+		\label{fig:plotamp}
+\end{figure*}
+
+\begin{figure*}[tp]
+	\centering
+		\includegraphics{code/data/plotM8-2.mps}
+		\caption{Phase error for 8PSK}
+		\label{fig:plotphase}
+\end{figure*}
+
+% \begin{figure*}[tp]
+% 	\centering
+%         \includegraphics{code/data/plotM8-3.mps}
+% 		\caption{Complex gain error for 8PSK}
+% 		\label{fig:plotcamp}
+% \end{figure*}
+
+\subsection{The noncoherent setting}
+
+%BPSK plots
+ \begin{figure*}[tp]
+ 	\centering
+ 		\includegraphics{code/data/plotM2-1.mps}
+ 		\caption{Amplitude error for BPSK}
+ 		\label{fig:plotamp}
+ \end{figure*}
+
+\begin{figure*}[tp]
+	\centering
+		\includegraphics{code/data/plotncM2-2.mps}
+		\caption{Phase error for BPSK}
+		\label{fig:plotphase}
+\end{figure*}
+
+% \begin{figure*}[tp]
+% 	\centering
+%         \includegraphics{code/data/plotM2-3.mps}
+% 		\caption{Complex gain error for BPSK}
+% 		\label{fig:plotcamp}
+% \end{figure*}
+
+%QPSK plots
+ \begin{figure*}[tp]
+ 	\centering
+ 		\includegraphics{code/data/plotM4-1.mps}
+ 		\caption{Amplitude error for QPSK}
+ 		\label{fig:plotamp}
+ \end{figure*}
+
+\begin{figure*}[tp]
+	\centering
+		\includegraphics{code/data/plotncM4-2.mps}
+		\caption{Phase error for QPSK}
+		\label{fig:plotphase}
+\end{figure*}
+
+ % \begin{figure*}[tp]
+ % 	\centering
+ %         \includegraphics{code/data/plotM4-3.mps}
+ % 		\caption{Complex gain error for QPSK}
+ % 		\label{fig:plotcamp}
+ % \end{figure*}
+
+%8PSK plots
+ \begin{figure*}[tp]
+ 	\centering
+ 		\includegraphics{code/data/plotM8-1.mps}
+ 		\caption{Amplitude error for 8PSK}
+ 		\label{fig:plotamp}
+ \end{figure*}
+
+\begin{figure*}[tp]
+	\centering
+		\includegraphics{code/data/plotncM8-2.mps}
+		\caption{Phase error for 8PSK}
+		\label{fig:plotphase}
+\end{figure*}
+
+% \begin{figure*}[tp]
+% 	\centering
+%         \includegraphics{code/data/plotM8-3.mps}
+% 		\caption{Complex gain error for 8PSK}
+% 		\label{fig:plotcamp}
+% \end{figure*}
 
@@
 
-% \section{Matched filtering: samples to symbols}
+\section{Matched filtering: samples to symbols}
 
-% The continuous time transmitted signal takes the form,
-% \[
-% x(t) = \sum_{i \in \ints} s_i g(t - iT),
-% \]
-% where $g(t)$ is the transmission pulse shape, $T$ is the symbol period.
+The continuous time transmitted signal takes the form,
+\[
+x(t) = \sum_{i \in \ints} s_i g(t - iT),
+\]
+where $g(t)$ is the transmission pulse shape, $T$ is the symbol period.
 Assume that the received signal undergoes time and phase offset, is attenuated by some amount and is also subject to additive noise.
 The model for the analogue received signal is,
-% \begin{align*}
-% r(t) &=  w(t) + \alpha_0 x(t - \tau_0) \\
-% &= v(t) + \sum_{i \in \ints} s_i g(t - iT - \tau_0),
-% \end{align*}
-% where $\alpha_0 \in \complex$ is a complex number representing both attenuation and phase offset, $\tau_0$ is the time offset and $w(t)$ is a continuous noise process.
+\begin{align*}
+r(t) &=  w(t) + \alpha_0 x(t - \tau_0) \\
+&= v(t) + \sum_{i \in \ints} s_i g(t - iT - \tau_0),
+\end{align*}
+where $\alpha_0 \in \complex$ is a complex number representing both attenuation and phase offset, $\tau_0$ is the time offset and $w(t)$ is a continuous noise process.
 Assume that an accurate estimate of the time offset $\tau_0$ is available.
@@
 
-% Consider the inner product
-% \begin{align*}
-% \dotprod{r(t)}{g(T_sn - Ti -  \tau_0)} &= \dotprod{v(t) + \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)}  \\
-% &= w_i +  \dotprod{\alpha_0 \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
-% &= w_i +  \alpha_0 \sum_{k \in \ints} s_k \dotprod{ g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
-% &= w_i + \alpha_0s_i = y_i
-% \end{align*}
-% where we put $w_i = \dotprod{v(t)}{g(T_sn - Ti - \tau)}$, and we have required that the transmission pulse $g(t)$ and it's shifts $g(t - Ti)$ are orthogonal, i.e.
-% \begin{equation}
-% \dotprod{g(t - iT)}{g(t - kT)} = \int_{\infty}^{\infty} g(t - iT)g^*(t-kT) \, dt = \begin{cases} 1 & i=k \\ 0 & i \neq k, \end{cases} \label{eq:contorthpulse}
-% \end{equation}
-% So, the recieved symbols can be obtained by computing the inner product of the recieved sequence signal with the pulse.
+Consider the inner product
+\begin{align*}
+\dotprod{r(t)}{g(T_sn - Ti -  \tau_0)} &= \dotprod{v(t) + \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)}  \\
+&= w_i +  \dotprod{\alpha_0 \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
+&= w_i +  \alpha_0 \sum_{k \in \ints} s_k \dotprod{ g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
+&= w_i + \alpha_0s_i = y_i
+\end{align*}
+where we put $w_i = \dotprod{v(t)}{g(T_sn - Ti - \tau)}$, and we have required that the transmission pulse $g(t)$ and it's shifts $g(t - Ti)$ are orthogonal, i.e.
+\begin{equation}
+\dotprod{g(t - iT)}{g(t - kT)} = \int_{\infty}^{\infty} g(t - iT)g^*(t-kT) \, dt = \begin{cases} 1 & i=k \\ 0 & i \neq k, \end{cases} \label{eq:contorthpulse}
+\end{equation}
+So, the recieved symbols can be obtained by computing the inner product of the recieved sequence signal with the pulse.
 
-% Thus far, we have assumed continuous processing of a continuous recieved signal.  However, typically in modern communications, only a sampled version of $r(t)$ is available at the reciever, i.e.
-% \[
-% r_n = r(nT_s) = v_n + \alpha_0 x(T_sn - \tau_0),
-% \] 
-% where $n$  is an integer and $T_s$ is the sampling period and the sequence $\{v_n, n \in \ints \}$ is a discrete noise process.
+Thus far, we have assumed continuous processing of a continuous recieved signal.  However, typically in modern communications, only a sampled version of $r(t)$ is available at the reciever, i.e.
+\[
+r_n = r(nT_s) = v_n + \alpha_0 x(T_sn - \tau_0),
+\] 
+where $n$  is an integer and $T_s$ is the sampling period and the sequence $\{v_n, n \in \ints \}$ is a discrete noise process.
 In this case, we replace the continuous time inner product, with the discrete time inner product, i.e.
 put
-% \[
-% \dotprod{r(t)}{g(t - iT - \tau)} = \sum_{n\in\ints}r(nT_s)g^*(T_sn - Ti - \tau_0) = \sum_{n\in\ints} r_n g^*(T_sn - Ti - \tau_0)
-% \]
-% The inner product above is an infinite sum and probably cannot be computed in practice.
+\[
+\dotprod{r(t)}{g(t - iT - \tau)} = \sum_{n\in\ints}r(nT_s)g^*(T_sn - Ti - \tau_0) = \sum_{n\in\ints} r_n g^*(T_sn - Ti - \tau_0)
+\]
+The inner product above is an infinite sum and probably cannot be computed in practice.
 However, in practical systems the number of samples observed is finite.
@@
 Now, the inner products are always finite sums
-% \[
-% \dotprod{r(t)}{g(t - iT - \tau_0)} =  \sum_{n=1}^N r_n g^*(T_sn - Ti - \tau_0).
-% \]
-% These finite inner products are computable in practice.
+\[
+\dotprod{r(t)}{g(t - iT - \tau_0)} =  \sum_{n=1}^N r_n g^*(T_sn - Ti - \tau_0).
+\]
+These finite inner products are computable in practice.
 Computing the inner products directly using the summation above is not necessarily the most efficient approach.
@@
 
-% We assume that $T_s$ is rationally related to $T$, i.e.
+We assume that $T_s$ is rationally related to $T$, i.e.
 $T_s = \frac{p}{q}T$ for some relatively prime integers $p$ and $q$.
@@
 
-% Let 
-% \[
-% g_n = g^*(-\tfrac{T_s}{p} n - \tau_0).
-% \]
-% Then
-% \[
-% g_{k-n} = g^*(-\tfrac{T_s}{p}(k - n) - \tau_0) = g^*(n\tfrac{T_s}{p} - k\tfrac{T_s}{p} - \tau_0) = g^*(n\tfrac{T_s}{p} - \tfrac{k}{q}T - \tau_0)
-% \]
-% Let
-% \begin{equation} \label{eq:zzerofill}
-% z_n = \begin{cases} 
-% r_{n/c} & n/p \in \{1, \dots, N \} \\
-% 0 & \text{otherwise}
-% \end{cases}
-% \end{equation}
-% be a zero filled version of $r_n$.  Now
-% \begin{align*}
-% h_k &= \dotprod{r(t)}{g(t - \tfrac{k}{q}T - \tau_0)} \\
-% &=  \sum_{n\in\ints} r_n g^*(T_sn - \tfrac{k}{q}T - \tau_0) \\
-% &= \sum_{n\in\ints} z_n g^*(\tfrac{T_s}{p}n - \tfrac{k}{q}T - \tau_0) \\
-% &= \sum_{n\in\ints} z_n g_{k-n}.
-% \end{align*}
-% So $h_k$ is the convolution between the sequence $\{z_n\}$ and the sequence $\{g_n\}$.  The convoultion can be computed using the fast Fourier transform.  The length of the transform required is
-% \[
-% BLERG.
-% \]  
-% It only remains to observe that $y_i = h_{iq}$.
+Let 
+\[
+g_n = g^*(-\tfrac{T_s}{p} n - \tau_0).
+\]
+Then
+\[
+g_{k-n} = g^*(-\tfrac{T_s}{p}(k - n) - \tau_0) = g^*(n\tfrac{T_s}{p} - k\tfrac{T_s}{p} - \tau_0) = g^*(n\tfrac{T_s}{p} - \tfrac{k}{q}T - \tau_0)
+\]
+Let
+\begin{equation} \label{eq:zzerofill}
+z_n = \begin{cases} 
+r_{n/c} & n/p \in \{1, \dots, N \} \\
+0 & \text{otherwise}
+\end{cases}
+\end{equation}
+be a zero filled version of $r_n$.  Now
+\begin{align*}
+h_k &= \dotprod{r(t)}{g(t - \tfrac{k}{q}T - \tau_0)} \\
+&=  \sum_{n\in\ints} r_n g^*(T_sn - \tfrac{k}{q}T - \tau_0) \\
+&= \sum_{n\in\ints} z_n g^*(\tfrac{T_s}{p}n - \tfrac{k}{q}T - \tau_0) \\
+&= \sum_{n\in\ints} z_n g_{k-n}.
+\end{align*}
+So $h_k$ is the convolution between the sequence $\{z_n\}$ and the sequence $\{g_n\}$.  The convoultion can be computed using the fast Fourier transform.  The length of the transform required is
+\[
+BLERG.
+\]  
+It only remains to observe that $y_i = h_{iq}$.
 

### 2012-05-04 paper.tex
Added to complex amp paper.  Started added the statistics

@@
 
-Here, we will describe a new algorithm for computing the least squares estimator that requires only $O(L)$ operations.
-We also consider a slightly more general model than is usual, one that allows the utilisation of pilot symbols, i.e., \emph{coherent detection}.
+Here, consider a slightly more general model than is usual, one that allows the utilisation of pilot symbols, i.e., \emph{coherent detection}.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
@@
 This complexity arises from the need to sort a list of $\abs{D}$ elements.
-In Section~\ref{sec:line-time-algor} we will show that a full sort is not neccesary, and that the least squares estimator can be implemented in $O(L)$ operations.
+%In Section~\ref{sec:line-time-algor} we will show that a full sort is not neccesary, and that the least squares estimator can be implemented in $O(L)$ operations.
 
@@
 \[
-Y = \sum_{i \in P \cup D} y_i s_i^* = \sum_{i \in P } y_i p_i^* + \sum_{i \in D } y_i d_i^*
+Y = \sum_{i \in P \cup D} y_i s_i^* = \sum_{i \in P } y_i p_i^* + \sum_{i \in D } y_i d_i^*.
 \]  
@@
 So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a constant number of operations.
-Given $Y_k$, the value of $SS(f_k)$ can be computed in constant number of operations using~\eqref{eq:SSfk}.
-Let $\hat{k}$ be integer the sequence that minimises $SS(f_k)$.
+Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.
+Let $\hat{k}$ be the integer that minimises $SS(f_k)$, i.e.
+$\hat{k} = \arg\min SS(f_k)$.
 The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata}, as
@@
 Psuedocode is given in Algorithm~\ref{alg:loglinear}.
-This descibes a practical implementation of Mackenthun's algorithm.
 Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given the sequence $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ of the indices in $D$ as described in~\eqref{eq:sigmasortind}.
@@
 The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all requirse $O(L)$ or less operations.
-In the next sections we will show how to the sortinces function can be avoided.
+%In the next sections we will show how to the sortinces function can be avoided.
 This leads to an algorithm that requires only $O(L)$ operations.
@@
 \caption{Mackenthun's least squares estimator}
-\dontprintsemicolon
+\DontPrintSemicolon
 \KwIn{$\{y_i, i \in P \cup D \}$}
@@
 
-%Let $f$ be a function maping $\reals$ to $\reals^N$ defined so that the $i$th element of $f(\theta)$ is $\round{\frac{M}{2\pi}( \phi_i - \theta )}$.
-Let $\phi$ be the vector of length $N$ with element $[\phi_1, \dots, \phi_N]$ and define the $\round{\cdots}$ to operate elementwise on vectors, then
-%\[
-%f(\theta) = \round{\frac{M}{2\pi}( \phi - \theta )}.
-%\]
-%The function $f(\theta)$ is piecewise continuous.
+BLERG $\hat{a}$ is a biased estimator, particularly when the SNR is low, we show firs that $\hat{\theta}$ is strongly consistent, we are then able to analyse the behaviour of $\hat{a}$.
+
+Again let $a = \rho e^{j\theta}$ and write the sum of squares function  
+\[
+SS(\rho, \theta, \{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{y_i}^2 - \rho y_i e^{-j\theta} s_i^* - \rho y_i e^{-\theta} s_i + \rho^2.
+\]
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives
+\[
+\rho = \frac{Z + Z^*}{2L}
+\]
+where $Z = \sum_{i\in P \cup D}y_i e^{-j\theta} s_i^*$.
+Observe that $Z$ is really a function of $\theta$ and $\{d_i, i \in D\}$, but we have supressed the arguments for notational clarity.
+We can now write
+\[
+SS(\rho, \theta, \{d_i, i \in D\}) = A - \rho Z - \rho Z^* + L \rho^2
+\]
+and substituting the above expression for $\rho$ into this gives the sum of squares function conditioned upon minimisation with respect to $\rho$,
+\[
+SS(\theta, \{d_i, i \in D\}) = A - \frac{1}{L}\abs{Z}^2.
+\]
+Substituting~(\ref{eq:hatdfinxtheta}) into this gives the sum of squares function conditioned upon minimisation with respect to $\rho$ and the data sequence $\{d_i, i \in D\}$,
+\[
+SS(\theta) = A - \frac{1}{L}\abs{Z(\theta)}^2
+\]
+where
+\[
+Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta).
+\]
+We have reused $Z$ here, but this should not cause confusion as $Z(\theta)$ and $Z(\theta, \{d_i, i \in D\})$ are easily told apart from there arguments.
+From now on, when we right $Z$ without arguments, we will always be refering to $Z(\theta)$.
+
+We are interested in analysing $\hat{\theta}$ the maximiser of $\abs{Z(\theta)}^2$.
+Put $\lambda_L = \fracpart{\theta_0 - \hat{\theta}}$.
+We want to show that $\lambda_L$ converges to zero as the number of transmitted symbols $L$ increases to $\infty$ where $\fracpart{x}$ denotes $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
+\[
+\fracpart{x} = x - 2\pi\round{\frac{x}{2\pi}}.
+\]
+Under the assumption that the noise $\{w_i\}$ is circularly symetric then
+\[
+y_i = a_0 s_i + w_i = \rho_j e^{j ( \theta_0 + \Phi_i + k_i) }
+\] 
+where $k_i = \angle s_i$ and the $\Phi_i$ are identically distributed.
+The $\Phi_i$ can be identified as \emph{circular random variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.
+Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
+\[
+\hat{u}_i(\theta) = \round{\frac{M}{2\pi}(\angle{y_i} - \theta)} = \round{\frac{M}{2\pi}(\lambda + \Phi_i + k_i)}
+\]
+where $\lambda = \fracpart{\theta_0 - \theta}$.  So,
+\begin{align*}
+ y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_j \exp{j \left(\lambda + \Phi_i + k_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\lambda + \Phi_i + k_i)}\right)} \\
+&= \rho_j \exp{j \left(\lambda + \Phi_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\lambda + \Phi_i) }\right)}
+\end{align*}
+when $i \in D$ and
+\[
+y_i e^{-j\theta} p_i^* = \rho_j \exp j \left(\lambda + \Phi_i )
+\]
+when $i \in P$.  Now we can write
+\[
+Z(\theta) = \sum_{i \in P} \rho_j e^{j(\lambda + \Phi_i)} + \sum_{i \in D}  \rho_j \exp{j \left( \lambda + \Phi_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\lambda + \Phi_i)}\right)}.
+\]
+Put $S_L(\lambda) = \frac{1}{L}\abs{Z(\theta)}^2$ and put $\lambda_L = \fracpart{\theta_0 - \hat{\theta}}$.
+Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.
+We will show that $\lambda_L$ converges almost surely to zero as $L \rightarrow \infty$.
+From this, the proof of strong consistency follows.
+
+BLERG  Show $E S_L$ is maximised uniquely at 0, proof follows.
+
+We now prove the asymptotics normality of the estimator.
+At the same time we are able to describe the distribution of the amplitude estimator $\hat{\rho}$.
+BLERG, first show that it is differentiable at the truth.
 

### 2012-05-05 paper.tex
Added to statitics part of this paper.  Starting to look quite good.

@@
 \[
-s_i = \exp\left( j\frac{2\pi}{M} u_i \right),
+s_i = e^{j u_i},
 \]
-where $j = \sqrt{-1}$ and $u_i$ is an integer in the set $\{0, 1, \dots, M-1\}$ and $M$ is the size of the constellation, i.e.
+where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, (M-1)\tfrac{2\pi}{M}\}$ and $M$ is the size of the constellation, i.e.
 we have assumed an $M$-PSK constellation.
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-
-Here, consider a slightly more general model than is usual, one that allows the utilisation of pilot symbols, i.e., \emph{coherent detection}.
+Here, we consider a slightly more general model than is usual, one that allows the utilisation of pilot symbols, i.e., \emph{coherent detection}.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
+By making a minor modification to Mackenthun's estimator, it is possible to meanifully include pilot symbols.
 
@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $E\abs{a_0 - \hat{a}}^2$ where $E$ denotes the expected value.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator and $\abs{x}$ denotes the magnitude of the complex number $x$.
+It will actually be informative to consider the magnitude and complex argument (or phase) of the estimator separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ and $a_0 = \rho_0e^{j\theta_0}$ where both $\rho_0$ and $\hat{\rho}$ are real, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
+The function $\fracpart{\cdot}$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
+It will become apparent why $\prob\vert\langle\theta_0 - \hat{\theta}\rangle\vert^2$ rather than $\prob\vert\theta_0 - \hat{\theta}\vert^2$ is the appropriate measure of error for the phase parameter.
+
+It is worth commenting on our use of $\prob$ rather than the more usual $\expect$ or $E$ to denote the expected value operator.
+The notation comes from Pollard~\cite[Ch 1]{Pollard_users_guide_prob_2002} and is originally due to de Finetti.
+The notation is good because it removes unecessary distinction between `probability' and expectation.
+Given a random variable $X$ with cumulative density $X$, the probability of an event, say $X \in S$, where $S$ is some subset of the range of $X$, is
+\[
+\prob \indicator \{X \in S\} = \int \indicator \{X \in S\}(x) dF(x) = \int_{S} dF(x)
+\]
+where $\indicator \{X \in S\}$ is the indicator function of the set $S$, i.e $\indicator \{X \in S\}(x) = 1$ when the argument $x \in S$ and zero otherwise.
+We will usually drop the $\onebf$ and simply write $\prob \{ X > \delta \}$ to mean $\prob \onebf\{ X > \delta \}$.
+As an example of the utility of this notation, Markov's inequality is plainly
+\[
+\prob \{ \abs{X} > \delta \}  \leq \prob \frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \} \leq \frac{1}{\delta}\prob\abs{X}.
+\]
+where $\frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \}(x)$ is the function equal to $\abs{x}/\delta$ when the argument $x > \delta$ and zero otherwise.
 
@@
 \end{equation}
-where $L = \abs{P \cup D}$ is the total number of symbols transmitted, and to simplify our notation we put 
+where $L = \abs{P \cup D}$ is the total number of symbols transmitted, and to simplify our notation we have put 
 \[
@@
 Observe that given a candidate value for the data symbols, we can compute the corresponding sum of squares function $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-We will show that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
+It turns out that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
 %When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
@@
 \begin{equation}\label{eq:hatdfinxtheta}
-\hat{d}_i(\theta) = \exp\left( j\frac{2\pi}{M}\hat{u}_i(\theta) \right) \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\frac{M}{2\pi}\angle( e^{-j\theta}y_i)},
+\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
 \end{equation}
-and where $\round{\cdot}$ rounds its argument to the nearest integer\footnote{The direction of rounding for half-integers is not important so long as it is consistent.
+and where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
+A word of caution here, it is often the case that $\round{\cdot}$ denotes rounding to the nearest \emph{integer}\footnote{The direction of rounding for half-integers is not important so long as it is consistent.
 The authors have chosen to round up half-integers in their own implementation.}.
+This is not the case here.
+If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer then, here,
+\[
+\round{x} = \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
+\] 
 Note that $\hat{d}_i(\theta)$ does not depend on the amplitude $\rho$.
-As defined, $\hat{u}_i(\theta)$ is not strictly inside the set of integers $\{0, 1, \dots, M-1\}$, but this is not of consequence, as we intend its value to be considered modulo $M$.
-With this in mind let $\phi_i = \angle y_i$, then,
+As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, (M-1)\tfrac{2\pi}{M}\}$, but this is not of consequence, as we intend its value to be considered modulo $2\pi$.
+With this in mind,
 \[
-\hat{u}_i(\theta) = \round{\frac{M}{2\pi}( \phi_i - \theta )}
+\hat{u}_i(\theta) = \round{\angle{y_i} - \theta }
 \]
-which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $M$.
+which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $2\pi$.
 
@@
 \[
-z_i = \phi_i - \frac{2\pi}{M}\hat{u}_i(0) = \phi_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}\phi_i},
+z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}},
 \]
@@
 Let $f(\theta)$ be the sequence $\{ \hat{d}_i(\theta), i \in D \}$.
-So, $f(\theta)$ is a function mapping the interval $[-\nicefrac{1}{2}, \nicefrac{1}{2})$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
+So, $f(\theta)$ is a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
 Observe that $f(\theta)$ is piecewise continuous.
@@
 The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata}, as
-\[
+\begin{equation}\label{eq:ahatYhat}
 \hat{a} = \frac{1}{L} Y_{\hat{k}}.
-\]
+\end{equation}
 Psuedocode is given in Algorithm~\ref{alg:loglinear}.
@@
 We are interested in analysing $\hat{\theta}$ the maximiser of $\abs{Z(\theta)}^2$.
-Put $\lambda_L = \fracpart{\theta_0 - \hat{\theta}}$.
-We want to show that $\lambda_L$ converges to zero as the number of transmitted symbols $L$ increases to $\infty$ where $\fracpart{x}$ denotes $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
+Put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
+We want to show that $\lambda_L$ converges to zero as the number of transmitted symbols $L$ increases to $\infty$ where $\sfracpart{x}$ denotes $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
 \[
-\fracpart{x} = x - 2\pi\round{\frac{x}{2\pi}}.
+\fracpart{x} = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
 \]
@@
 Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
-\[
-\hat{u}_i(\theta) = \round{\frac{M}{2\pi}(\angle{y_i} - \theta)} = \round{\frac{M}{2\pi}(\lambda + \Phi_i + k_i)}
-\]
+\begin{align*}
+\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
+&= \round{\theta_0 + \Phi_i + k_i - \theta} \\
+&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + k_i} \pmod M \\
+&= \round{ \lambda + \Phi_i + k_i } \\
+\end{align*}
 where $\lambda = \fracpart{\theta_0 - \theta}$.  So,
 \begin{align*}
- y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_j \exp{j \left(\lambda + \Phi_i + k_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\lambda + \Phi_i + k_i)}\right)} \\
-&= \rho_j \exp{j \left(\lambda + \Phi_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\lambda + \Phi_i) }\right)}
+ y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_i e^{j(\lambda + \Phi_i + k_i - \round{\lambda + \Phi_i + k_i})} \\
+&= \rho_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})}
 \end{align*}
@@
 \[
-y_i e^{-j\theta} p_i^* = \rho_j \exp j \left(\lambda + \Phi_i )
+y_i e^{-j\theta} p_i^* = \rho_i e^{j(\lambda + \Phi_i)}
 \]
@@
 \[
-Z(\theta) = \sum_{i \in P} \rho_j e^{j(\lambda + \Phi_i)} + \sum_{i \in D}  \rho_j \exp{j \left( \lambda + \Phi_i - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\lambda + \Phi_i)}\right)}.
+Z(\theta) = \sum_{i \in P} \rho_i e^{j(\lambda + \Phi_i)} + \sum_{i \in D}  \rho_i e^{j( \lambda + \Phi_i - \round{\lambda + \Phi_i})}.
 \]
-Put $S_L(\lambda) = \frac{1}{L}\abs{Z(\theta)}^2$ and put $\lambda_L = \fracpart{\theta_0 - \hat{\theta}}$.
+Put $S_L(\lambda) = \frac{1}{L}\abs{Z(\theta)}^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
 Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.
@@
 
-We now prove the asymptotics normality of the estimator.
-At the same time we are able to describe the distribution of the amplitude estimator $\hat{\rho}$.
-BLERG, first show that it is differentiable at the truth.
+We now prove the asymptotic normality of the estimator.
+At the same time we are able to describe the distribution of the magnitude estimator $\hat{\rho}$.
+Observe that $Z(\theta) = e^{-j\theta}Y(\{\hat{d}_i(\hat{\theta}), i \in D\})$.
+From~\eqref{eq:ahatYhat} we obtain
+\begin{align*}\label{eq:hata}
+\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \\
+&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \\
+&= \frac{1}{L}  Z(\hat{\theta}) \\
+&= \frac{1}{L} \sum_{i \in P} \rho_i e^{j(\lambda_L + \Phi_i)} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j( \lambda_L + \Phi_i - \round{\lambda_L + \Phi_i})}
+\end{align*}
+from which it follows that
+\begin{align*}\label{eq:hatrhoworking}
+\hat{\rho} &=  \frac{1}{L} \sum_{i \in P} \rho_i e^{j\Phi_i} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j(\Phi_i - \round{\lambda_L + \Phi_i})} \\
+ &=  \frac{1}{L} \sum_{i \in P} (\rho_0 + w_i) + \frac{1}{L}  \sum_{i \in D}  (\rho_0 + w_i)e^{-j\round{\lambda_L + \Phi_i}} 
+\end{align*}
 

### 2012-05-05 paper.tex
Added a bit more to stats

@@
 \]
-where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, (M-1)\tfrac{2\pi}{M}\}$ and $M$ is the size of the constellation, i.e.
+where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation, i.e.
 we have assumed an $M$-PSK constellation.
@@
 &= \round{\theta_0 + \Phi_i + k_i - \theta} \\
-&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + k_i} \pmod M \\
+&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + k_i} \pmod 2\pi \\
 &= \round{ \lambda + \Phi_i + k_i } \\
@@
 \end{align*}
-from which it follows that
+and muliplying both sides of this equation by $e^{-j\lambda_L}$,
 \begin{align*}\label{eq:hatrhoworking}

### 2012-05-06 paper.tex
Added theorem statement.  Still needs some work, but is probably on the right track.

@@
 \section{Statistical properties}
+
+\begin{theorem}
+Let $\{w_i, i \in \ints \}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables with $w_i = \rho_i e^{j\Phi_i}$ such that $h = \prob e^{j\round{\Phi_i}} = \prob e^{j\round{\Phi_1}}$ is a positive real number.
+Given $\{y_i, i \in P \cup D\}$ from~\ref{eq:sigmod} let $\hat{a}$ be the least squares estimator of $a_0$.
+Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}$ and $a_0 = \rho_0 e^{j\theta_0}$.
+Put $L = \abs{P \cup D}$ and assume that $\abs{P}/L \rightarrow p$ and $\abs{D}/L \rightarrow d$ as $L \rightarrow \infty$.
+Then:
+\begin{enumerate}
+\item $\fracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item $\hat{\rho} \rightarrow \rho_0( p + d h )$ almost surely as $L \rightarrow \infty$,
+\end{enumerate}
+\end{theorem}
 

### 2012-05-06 paper.tex
Added more to paper.

--- 2012-05-05 paper.tex
@@
 
+\begin{theorem}
+Let $\{w_i\}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables and let $\{\Phi_i\}$ be the sequence of idenpendent and identically distributed random variables given by $\Phi_i = \angle{(\rho_0 + w_i)}$.
+Put
+\[
+h_1 = \prob e^{j\round{\Phi_1}} \qquad \text{and} \qquad h_2 = \prob w_1 e^{j\round{\Phi_1}}.
+\]
+Given the observed noisy data symbols $\{y_i, i \in P \cup D\}$ from~\eqref{eq:sigmod} let $\hat{a}$ be the least squares estimator of $a_0$.
+Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}$ and $a_0 = \rho_0 e^{j\theta_0}$.
+Put $L = \abs{P \cup D}$.
+If
+\[
+\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d
+\] 
+as $L \rightarrow \infty$ for positive real numbers $p$ and $d$, and if both $h_1$ and $h_2$ are positive real numbers, then:
+\begin{enumerate}
+\item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item $\hat{\rho} \rightarrow \rho_0( p + d h_1 ) + h_2$ almost surely as $L \rightarrow \infty$,
+\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $m_L = \hat{\rho} - \rho_0( p + d h_1 ) + h_2$.
+Then the vector $[\lambda_L, m_L]$ converges to the bivariate normal distribution with zero mean and covariance
+\[
+C = ?
+\]
+as $L \rightarrow \infty$.
+\end{enumerate}
+\end{theorem}
+
 BLERG $\hat{a}$ is a biased estimator, particularly when the SNR is low, we show firs that $\hat{\theta}$ is strongly consistent, we are then able to analyse the behaviour of $\hat{a}$.
@@
 We have reused $Z$ here, but this should not cause confusion as $Z(\theta)$ and $Z(\theta, \{d_i, i \in D\})$ are easily told apart from there arguments.
-From now on, when we right $Z$ without arguments, we will always be refering to $Z(\theta)$.
+%From now on, when we write $Z$ without arguments, we will always be refering to $Z(\theta)$.
 
@@
 &\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + k_i} \pmod 2\pi \\
-&= \round{ \lambda + \Phi_i + k_i } \\
+&= \round{ \lambda + \Phi_i + k_i },
 \end{align*}
@@
 
-\appendix
-
-\section{Matched filtering: samples to symbols}
-
-The continuous time transmitted signal takes the form,
-\[
-x(t) = \sum_{i \in \ints} s_i g(t - iT),
-\]
-where $g(t)$ is the transmission pulse shape, $T$ is the symbol period.
-Assume that the received signal undergoes time and phase offset, is attenuated by some amount and is also subject to additive noise.
-The model for the analogue received signal is,
-\begin{align*}
-r(t) &=  w(t) + \alpha_0 x(t - \tau_0) \\
-&= v(t) + \sum_{i \in \ints} s_i g(t - iT - \tau_0),
-\end{align*}
-where $\alpha_0 \in \complex$ is a complex number representing both attenuation and phase offset, $\tau_0$ is the time offset and $w(t)$ is a continuous noise process.
-Assume that an accurate estimate of the time offset $\tau_0$ is available.
-We would like to process the signal $r(t)$ so as to obtain a signal in the form of~\eqref{eq:sigmod}.
-This is the task of \emph{matched filtering}.
-
-Consider the inner product
-\begin{align*}
-\dotprod{r(t)}{g(T_sn - Ti -  \tau_0)} &= \dotprod{v(t) + \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)}  \\
-&= w_i +  \dotprod{\alpha_0 \sum_{k \in \ints} s_k g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
-&= w_i +  \alpha_0 \sum_{k \in \ints} s_k \dotprod{ g(t - kT - \tau_0)}{g(T_sn - Ti - \tau)} \\
-&= w_i + \alpha_0s_i = y_i
-\end{align*}
-where we put $w_i = \dotprod{v(t)}{g(T_sn - Ti - \tau)}$, and we have required that the transmission pulse $g(t)$ and it's shifts $g(t - Ti)$ are orthogonal, i.e.
-\begin{equation}
-\dotprod{g(t - iT)}{g(t - kT)} = \int_{\infty}^{\infty} g(t - iT)g^*(t-kT) \, dt = \begin{cases} 1 & i=k \\ 0 & i \neq k, \end{cases} \label{eq:contorthpulse}
-\end{equation}
-So, the recieved symbols can be obtained by computing the inner product of the recieved sequence signal with the pulse.
-
-Thus far, we have assumed continuous processing of a continuous recieved signal.  However, typically in modern communications, only a sampled version of $r(t)$ is available at the reciever, i.e.
-\[
-r_n = r(nT_s) = v_n + \alpha_0 x(T_sn - \tau_0),
-\] 
-where $n$  is an integer and $T_s$ is the sampling period and the sequence $\{v_n, n \in \ints \}$ is a discrete noise process.
-In this case, we replace the continuous time inner product, with the discrete time inner product, i.e.
-put
-\[
-\dotprod{r(t)}{g(t - iT - \tau)} = \sum_{n\in\ints}r(nT_s)g^*(T_sn - Ti - \tau_0) = \sum_{n\in\ints} r_n g^*(T_sn - Ti - \tau_0)
-\]
-The inner product above is an infinite sum and probably cannot be computed in practice.
-However, in practical systems the number of samples observed is finite.
-In what follows we assume that $N$ consecutive samples $r_1, r_2, \dots, r_N$ are observed and we set $r_n = 0$ whenever $n \notin \{1, 2, \dots, N\}$.
-Now, the inner products are always finite sums
-\[
-\dotprod{r(t)}{g(t - iT - \tau_0)} =  \sum_{n=1}^N r_n g^*(T_sn - Ti - \tau_0).
-\]
-These finite inner products are computable in practice.
-Computing the inner products directly using the summation above is not necessarily the most efficient approach.
-Instead, the fast Fourier transfrom can be used, as we now show.
-
-We assume that $T_s$ is rationally related to $T$, i.e.
-$T_s = \frac{p}{q}T$ for some relatively prime integers $p$ and $q$.
-This property is a common assumption in the literature.
-In fact, it is common to assume the stronger condition, that $T$ is a multiple of $T_s$~\cite{Oerder_synch_square_circstat_2008}.
-At the very least, this assumption can be made an accurate approximation by an appropriate choice of $p$ and $q$~\cite{Mordell_Diophantine_equations_1969,Clarkson_thesis}.
-
-Let 
-\[
-g_n = g^*(-\tfrac{T_s}{p} n - \tau_0).
-\]
-Then
-\[
-g_{k-n} = g^*(-\tfrac{T_s}{p}(k - n) - \tau_0) = g^*(n\tfrac{T_s}{p} - k\tfrac{T_s}{p} - \tau_0) = g^*(n\tfrac{T_s}{p} - \tfrac{k}{q}T - \tau_0)
-\]
-Let
-\begin{equation} \label{eq:zzerofill}
-z_n = \begin{cases} 
-r_{n/c} & n/p \in \{1, \dots, N \} \\
-0 & \text{otherwise}
-\end{cases}
-\end{equation}
-be a zero filled version of $r_n$.  Now
-\begin{align*}
-h_k &= \dotprod{r(t)}{g(t - \tfrac{k}{q}T - \tau_0)} \\
-&=  \sum_{n\in\ints} r_n g^*(T_sn - \tfrac{k}{q}T - \tau_0) \\
-&= \sum_{n\in\ints} z_n g^*(\tfrac{T_s}{p}n - \tfrac{k}{q}T - \tau_0) \\
-&= \sum_{n\in\ints} z_n g_{k-n}.
-\end{align*}
-So $h_k$ is the convolution between the sequence $\{z_n\}$ and the sequence $\{g_n\}$.  The convoultion can be computed using the fast Fourier transform.  The length of the transform required is
-\[
-BLERG.
-\]  
-It only remains to observe that $y_i = h_{iq}$.
-
 \bibliography{bib}

### 2012-05-06 paper.tex
Added to paper.

@@
 \[
-h_1 = \prob e^{j\round{\Phi_1}} \qquad \text{and} \qquad h_2 = \prob w_1 e^{j\round{\Phi_1}}.
+h_1 = \prob e^{-j\round{\Phi_1}} \qquad \text{and} \qquad h_2 = \prob w_1 e^{-j\round{\Phi_1}}.
 \]
@@
 \item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $m_L = \hat{\rho} - \rho_0( p + d h_1 ) + h_2$.
-Then the vector $[\lambda_L, m_L]$ converges to the bivariate normal distribution with zero mean and covariance
+Then the vector $[\sqrt{N}\lambda_L, \sqrt{N}m_L]$ converges to the bivariate normal distribution with zero mean and covariance
 \[
@@
 From~\eqref{eq:ahatYhat} we obtain
-\begin{align*}\label{eq:hata}
-\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \\
-&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \\
-&= \frac{1}{L}  Z(\hat{\theta}) \\
-&= \frac{1}{L} \sum_{i \in P} \rho_i e^{j(\lambda_L + \Phi_i)} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j( \lambda_L + \Phi_i - \round{\lambda_L + \Phi_i})}
-\end{align*}
+\begin{align}\label{eq:hata}
+\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \nonumber \\
+&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \nonumber \\
+&= \frac{1}{L}  Z(\hat{\theta}) \nonumber \\
+&= \frac{1}{L} \sum_{i \in P} \rho_i e^{j(\lambda_L + \Phi_i)} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j( \lambda_L + \Phi_i - \round{\lambda_L + \Phi_i})} \label{eq:ahatfornorm}
+\end{align}
 and muliplying both sides of this equation by $e^{-j\lambda_L}$,
@@
 \hat{\rho} &=  \frac{1}{L} \sum_{i \in P} \rho_i e^{j\Phi_i} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j(\Phi_i - \round{\lambda_L + \Phi_i})} \\
- &=  \frac{1}{L} \sum_{i \in P} (\rho_0 + w_i) + \frac{1}{L}  \sum_{i \in D}  (\rho_0 + w_i)e^{-j\round{\lambda_L + \Phi_i}} 
+ &=  \frac{1}{L} \sum_{i \in P} (\rho_0 + w_i) + \frac{1}{L}  \sum_{i \in D}  (\rho_0 + w_i)e^{-j\round{\lambda_L + \Phi_i}}.
 \end{align*}
+By the strong law of large numbers, and since $\lambda_L \rightarrow 0$ (BLERG needs more care than this, but it's true),
+\begin{align*}
+\frac{1}{L}\sum_{i \in P}w_i \rightarrow 0 \\
+\frac{1}{L}\sum_{i \in D}\rho_0e^{-j\round{\lambda_L + \Phi_i}} \rightarrow  \rho_0 h_1 \\
+\frac{1}{L}\sum_{i \in P}w_i e^{-j\round{\lambda_L + \Phi_i}} \rightarrow h_2 \\
+\end{align*}
+and it immediately follows that $\hat{\rho} \rightarrow \rho_0( p + d h_1 ) + h_2$.  This proves statement 2 of our theorem.
+
+We have
+\begin{align*}
+\sqrt{L} m_L =  A_L + B_L +  C_L
+\end{align*}
+where
+\[
+A_L = \frac{1}{\sqrt{L}} \sum_{i \in P}w_i,
+\]
+and
+\[
+B_L = \frac{\rho_0}{\sqrt{L}} \sum_{i \in D}(e^{-j\round{\lambda_L + \Phi_i}} - h_1),
+\]
+and
+\[
+C_L = \frac{1}{\sqrt{L}} \sum_{i \in D}(w_i e^{-j\round{\lambda_L + \Phi_i}} - h_2).
+\]
+Now $A_L$ converges in distribution to the complex normal with zero mean and variance $p\sigma^2$.
 

### 2012-05-07 paper.tex
minor edits.  Added IEEE headers
Added to paper.

--- 2012-05-06 paper.tex
@@
-\documentclass[a4paper,10pt]{article}
+%\documentclass[a4paper,10pt]{article}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 
@@
 \maketitle
+
+\begin{abstract}
+BLERG
+\end{abstract}
+\begin{IEEEkeywords}
+BLERG
+\end{IEEEkeywords}
 
@@
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-By making a minor modification to Mackenthun's estimator, it is possible to meanifully include pilot symbols.
+We show in Section~\ref{sec:least-squar-estim}~ that by making a minor modification to Mackenthun's estimator, it is possible to meanifully include pilot symbols.
 
@@
 Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator and $\abs{x}$ denotes the magnitude of the complex number $x$.
-It will actually be informative to consider the magnitude and complex argument (or phase) of the estimator separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ and $a_0 = \rho_0e^{j\theta_0}$ where both $\rho_0$ and $\hat{\rho}$ are real, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
+It will actually be informative to consider the magnitude and complex argument (or phase) of the estimator separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ and $a_0 = \rho_0e^{j\theta_0}$ where both $\rho_0$ and $\hat{\rho}$ are positive real numbers, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
 The function $\fracpart{\cdot}$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
@@
 The notation is good because it removes unecessary distinction between `probability' and expectation.
-Given a random variable $X$ with cumulative density $X$, the probability of an event, say $X \in S$, where $S$ is some subset of the range of $X$, is
+Given a random variable $X$ with cumulative density function $F$, the probability of an event, say $X \in S$, where $S$ is some subset of the range of $X$, is
 \[
@@
 
-In this section we derive Mackentun's~\cite{Mackenthun1994} least squares estimator.
+In this section we derive Mackentun's least squares estimator~\cite{Mackenthun1994}.
 Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to meaninfully include the pilot symbols, so we also consider the coherent setting here.
@@
 
-To see this, let $a = \rho e^{j\theta}$.  Now the sum of squares function can be written as
+To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a positive real.  Now the sum of squares function can be written as
 \begin{align*}
@@
 and where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
-A word of caution here, it is often the case that $\round{\cdot}$ denotes rounding to the nearest \emph{integer}\footnote{The direction of rounding for half-integers is not important so long as it is consistent.
-The authors have chosen to round up half-integers in their own implementation.}.
+A word of caution here, it is often the case that $\round{\cdot}$ denotes rounding to the nearest \emph{integer}.
 This is not the case here.
-If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer then, here,
+If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer then,
 \[
@@
 Note that $\hat{d}_i(\theta)$ does not depend on the amplitude $\rho$.
-As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, (M-1)\tfrac{2\pi}{M}\}$, but this is not of consequence, as we intend its value to be considered modulo $2\pi$.
+As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered modulo $2\pi$.
 With this in mind,
@@
 
-We only require to consider $\theta$ in the interval $[0, 2\pi)$.
-Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.
-Define the integer $b_i = \hat{d}_i(0)$ and the real number
+We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Define $b_i = \hat{d}_i(0)$ and the real number 
 \[
@@
 \[
-f_0 = f(0) = \{ b_i, i \in D \}
+f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}
 \]  
@@
 \begin{equation}\label{eq:fkrec}
-f_{k+1} = f_{k} e_{\sigma(k)},
+f_{k+1} = f_{k} e_{\sigma(k)}.
 \end{equation}
@@
 The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm.
-The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all requirse $O(L)$ or less operations.
+The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
 %In the next sections we will show how to the sortinces function can be avoided.
@@
 
+We use $\sfracpart{x}$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
+\[
+\fracpart{x} = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
+\]
+where $\round(\cdot)$ takes its argument to the nearest integer.
+The direction of rounding for half-integers is not important so long as it is consistent.
+We have chosen to round up half-integers here.
+
 \begin{theorem}
-Let $\{w_i\}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables and let $\{\Phi_i\}$ be the sequence of idenpendent and identically distributed random variables given by $\Phi_i = \angle{(\rho_0 + w_i)}$.
-Put
+Let $\{w_i\}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables.
+Put $X_i = \rho_0 + w_i e^{-j\theta_0} s_i^*$ and $\Phi_i = \angle{X_i}$ and
 \[
-h_1 = \prob e^{j\round{\Phi_1}} \qquad \text{and} \qquad h_2 = \prob w_1 e^{j\round{\Phi_1}}.
+h = \prob X_i e^{-j\round{\Phi_i}} = \prob X_1 e^{-j\round{\Phi_1}}.
 \]
@@
 \] 
-as $L \rightarrow \infty$ for positive real numbers $p$ and $d$, and if both $h_1$ and $h_2$ are positive real numbers, then:
+as $L \rightarrow \infty$ and each of $p$ and $d$ and $h$ are positive real numbers, then:
 \begin{enumerate}
 \item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0( p + d h_1 ) + h_2$ almost surely as $L \rightarrow \infty$,
-\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $m_L = \hat{\rho} - \rho_0( p + d h_1 ) + h_2$.
-Then the vector $[\lambda_L, m_L]$ converges to the bivariate normal distribution with zero mean and covariance
+\item $\hat{\rho} \rightarrow \rho_0 p + d h$ almost surely as $L \rightarrow \infty$,
+\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $m_L = \hat{\rho} - \rho_0(p + dh)$.
+Then the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal distribution with zero mean and covariance
 \[
@@
 We are interested in analysing $\hat{\theta}$ the maximiser of $\abs{Z(\theta)}^2$.
-Put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
-We want to show that $\lambda_L$ converges to zero as the number of transmitted symbols $L$ increases to $\infty$ where $\sfracpart{x}$ denotes $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
-\[
-\fracpart{x} = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
-\]
 Under the assumption that the noise $\{w_i\}$ is circularly symetric then
 \[
-y_i = a_0 s_i + w_i = \rho_j e^{j ( \theta_0 + \Phi_i + k_i) }
+y_i = a_0 s_i + w_i = s_i e^{j \theta_0}( \rho_0 + w_i s_i^* e^{-j \theta_0}) = X_i e^{j ( \theta_0 + \angle{s_i}) }
 \] 
-where $k_i = \angle s_i$ and the $\Phi_i$ are identically distributed.
 The $\Phi_i$ can be identified as \emph{circular random variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.
@@
 \hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
-&= \round{\theta_0 + \Phi_i + k_i - \theta} \\
-&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + k_i} \pmod 2\pi \\
-&= \round{ \lambda + \Phi_i + k_i },
+&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
+&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
+&= \round{ \lambda + \Phi_i + \angle{s_i} },
 \end{align*}
-where $\lambda = \fracpart{\theta_0 - \theta}$.  So,
+where $\lambda = \fracpart{\theta_0 - \theta}$.  When $i \in D$,
 \begin{align*}
- y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_i e^{j(\lambda + \Phi_i + k_i - \round{\lambda + \Phi_i + k_i})} \\
-&= \rho_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})}
+ y_i e^{-j\theta} \hat{d}_i(\theta)^* &= X_i e^{j(\lambda + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
+&= X_i e^{j(\lambda - \round{\lambda + \Phi_i})},
 \end{align*}
-when $i \in D$ and
+since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  When $i \in P$,  
 \[
-y_i e^{-j\theta} p_i^* = \rho_i e^{j(\lambda + \Phi_i)}
+y_i e^{-j\theta} p_i^* = X_i e^{j\lambda}
 \]
-when $i \in P$.  Now we can write
+Now we can write
+\begin{align*}
+Z(\theta) &= \sum_{i \in P} X_i e^{j\lambda} + \sum_{i \in D} X_i e^{j(\lambda - \round{\lambda + \Phi_i})} \\
+&= e^{j\lambda} \left( \sum_{i \in P} X_i + \sum_{i \in D} X_i e^{-j\round{\lambda + \Phi_i}} \right)
+\end{align*}
+Put 
 \[
-Z(\theta) = \sum_{i \in P} \rho_i e^{j(\lambda + \Phi_i)} + \sum_{i \in D}  \rho_i e^{j( \lambda + \Phi_i - \round{\lambda + \Phi_i})}.
+S_L(\lambda) = \frac{1}{L^2}\abs{Z(\theta)}^2 = \abs{\frac{1}{L}\sum_{i \in P} X_i + \frac{1}{L}\sum_{i \in D} X_i e^{-j\round{\lambda + \Phi_i}}}^2
 \] 
-Put $S_L(\lambda) = \frac{1}{L}\abs{Z(\theta)}^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
+and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
 Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.
@@
 
-We now prove the asymptotic normality of the estimator.
-At the same time we are able to describe the distribution of the magnitude estimator $\hat{\rho}$.
-Observe that $Z(\theta) = e^{-j\theta}Y(\{\hat{d}_i(\hat{\theta}), i \in D\})$.
-From~\eqref{eq:ahatYhat} we obtain
-\begin{align*}\label{eq:hata}
-\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \\
-&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \\
-&= \frac{1}{L}  Z(\hat{\theta}) \\
-&= \frac{1}{L} \sum_{i \in P} \rho_i e^{j(\lambda_L + \Phi_i)} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j( \lambda_L + \Phi_i - \round{\lambda_L + \Phi_i})}
+Observe that $Z(\theta) = e^{-j\theta}Y(\{\hat{d}_i(\hat{\theta}), i \in D\})$.  From~\eqref{eq:ahatYhat} we obtain
+\begin{align}\label{eq:hata}
+\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \nonumber \\
+&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \nonumber \\
+&= \frac{1}{L}  Z(\hat{\theta}) \nonumber \\
+&= \frac{1}{L} \sum_{i \in P} X_i e^{j\lambda_L} + \frac{1}{L} \sum_{i \in D}X_i e^{j( \lambda_L  - \round{\lambda_L + \Phi_i})} \label{eq:ahatfornorm}
+\end{align}
+and muliplying both sides of this equation by $e^{-j\lambda_L}$,
+\[
+\hat{\rho} =  \frac{1}{L}\sum_{i \in P} X_i + \frac{1}{L} \sum_{i \in D} X_i e^{-j\round{\lambda_L + \Phi_i}}
+\]
+By the strong law of large numbers, and since $\lambda_L \rightarrow 0$ (BLERG needs more care than this, but it's true),
+\begin{align*}
+\frac{1}{L}\sum_{i \in P}X_i \rightarrow p \rho_0 \\
+\frac{1}{L} \sum_{i \in D} X_i e^{-j\round{\lambda_L + \Phi_i}} \rightarrow  d h 
 \end{align*}
-and muliplying both sides of this equation by $e^{-j\lambda_L}$,
-\begin{align*}\label{eq:hatrhoworking}
-\hat{\rho} &=  \frac{1}{L} \sum_{i \in P} \rho_i e^{j\Phi_i} + \frac{1}{L}  \sum_{i \in D}  \rho_i e^{j(\Phi_i - \round{\lambda_L + \Phi_i})} \\
- &=  \frac{1}{L} \sum_{i \in P} (\rho_0 + w_i) + \frac{1}{L}  \sum_{i \in D}  (\rho_0 + w_i)e^{-j\round{\lambda_L + \Phi_i}} 
+and it immediately follows that $\hat{\rho} \rightarrow \rho_0 p + d h$.  This proves statement 2 of our theorem.
+
+We have
+\begin{align*}
+\sqrt{L} m_L =  A_L + B_L
 \end{align*}
+where
+\[
+A_L = \frac{1}{\sqrt{L}} \sum_{i \in P}w_i,
+\]
+and
+\[
+B_L = \frac{1}{\sqrt{L}} \sum_{i \in D}X_i e^{-j\round{\lambda_L + \Phi_i}} - h.
+\]
+Now $A_L$ converges in distribution to the complex normal with zero mean and variance $p\sigma^2$ and $B_L$ converges in probability to
+\[
+\lambda_L \sqrt{L} \sum_{k=-M}^{M-1} g(\psi_k) e^{j\psi_k}
+\]
+where $\psi_k = \frac{2\pi}{M}k + \frac{\pi}{M}$ and where
+\[
+g(\phi) = \frac{\rho_0}{2\pi} \int_{0}^\infty \frac{r^2f_W(\rho_0^{-1}\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}} dr
+\]
 

### 2012-05-07 paper.tex
Added to paper.  Fixed a few silly things.

@@
 
-We use $\sfracpart{x}$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
+We use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
 \[
-\fracpart{x} = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
+\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
 \]
-where $\round(\cdot)$ takes its argument to the nearest integer.
+where $\operatorname{round}(\cdot)$ takes its argument to the nearest integer.
 The direction of rounding for half-integers is not important so long as it is consistent.
 We have chosen to round up half-integers here.
+Similarly we use $\sfracpart{x}$ to denote $x$ taken `modulo $\tfrac{2\pi}{M}$' into the interval $\left[-\tfrac{\pi}{M}, \tfrac{\pi}{M}\right)$, that is
+\[
+\fracpart{x} = x - \frac{2\pi}{M}\operatorname{round}\left(\frac{Mx}{2\pi}\right).
+\]
 
@@
 \begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
 \item $\hat{\rho} \rightarrow \rho_0 p + d h$ almost surely as $L \rightarrow \infty$,
-\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $m_L = \hat{\rho} - \rho_0(p + dh)$.
+\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0(p + dh)$.
 Then the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal distribution with zero mean and covariance
@@
 \[
-\rho = \frac{Z + Z^*}{2L}
+\rho = \frac{Z + Z^*}{2L} = \frac{1}{L}\Re(Z)
 \]
-where $Z = \sum_{i\in P \cup D}y_i e^{-j\theta} s_i^*$.
+where $Z = \sum_{i\in P \cup D}y_i e^{-j\theta} s_i^*$ and $\Re(\cdot)$ denotes the real part.
 Observe that $Z$ is really a function of $\theta$ and $\{d_i, i \in D\}$, but we have supressed the arguments for notational clarity.
@@
 \[
-SS(\theta, \{d_i, i \in D\}) = A - \frac{1}{L}\abs{Z}^2.
+SS(\theta, \{d_i, i \in D\}) = A - \frac{1}{L}\Re(Z)^2.
 \]
@@
 \[
-SS(\theta) = A - \frac{1}{L}\abs{Z(\theta)}^2
+SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2
 \]
@@
 
-We are interested in analysing $\hat{\theta}$ the maximiser of $\abs{Z(\theta)}^2$.
-Under the assumption that the noise $\{w_i\}$ is circularly symetric then
+We are interested in analysing $\hat{\theta}$, the maximiser of $\abs{Z(\theta)}^2$.
+Because the noise $\{w_i\}$ is circularly symetric then
 \[
-y_i = a_0 s_i + w_i = s_i e^{j \theta_0}( \rho_0 + w_i s_i^* e^{-j \theta_0}) = X_i e^{j ( \theta_0 + \angle{s_i}) }
+y_i = a_0 s_i + w_i = s_i e^{j \theta_0}( \rho_0 + w_i s_i^* e^{-j \theta_0}) = R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }
 \] 
-The $\Phi_i$ can be identified as \emph{circular random variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.
+The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.
 Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
@@
 \end{align*}
-where $\lambda = \fracpart{\theta_0 - \theta}$.  When $i \in D$,
+where $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  When $i \in D$,
 \begin{align*}
- y_i e^{-j\theta} \hat{d}_i(\theta)^* &= X_i e^{j(\lambda + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
-&= X_i e^{j(\lambda - \round{\lambda + \Phi_i})},
+ y_i e^{-j\theta} \hat{d}_i(\theta)^* &= R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
+&= R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
+&= R_i e^{j\sfracpart{\lambda + \Phi_i}}
 \end{align*}
@@
 \[
-y_i e^{-j\theta} p_i^* = X_i e^{j\lambda}
+y_i e^{-j\theta} p_i^* = R_i e^{j(\lambda + \Phi_i)}
 \]
 Now we can write
-\begin{align*}
-Z(\theta) &= \sum_{i \in P} X_i e^{j\lambda} + \sum_{i \in D} X_i e^{j(\lambda - \round{\lambda + \Phi_i})} \\
-&= e^{j\lambda} \left( \sum_{i \in P} X_i + \sum_{i \in D} X_i e^{-j\round{\lambda + \Phi_i}} \right)
-\end{align*}
-Put 
+\begin{align}
+Z(\theta) &= \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
+\end{align}
+and
 \[
-S_L(\lambda) = \frac{1}{L^2}\abs{Z(\theta)}^2 = \abs{\frac{1}{L}\sum_{i \in P} X_i + \frac{1}{L}\sum_{i \in D} X_i e^{-j\round{\lambda + \Phi_i}}}^2
+\Re(Z(\theta)) = \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \sum_{i \in D} R_i \cos(\sfracpart{\lambda + \Phi_i})
 \] 
-and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
+Let $S_L(\lambda) = \frac{1}{L^2}\Re(Z)^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
 Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.

### 2012-05-08 paper.tex
Added a bit more to paper.

@@
 \begin{theorem}
-Let $\{w_i\}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables.
-Put $X_i = \rho_0 + w_i e^{-j\theta_0} s_i^*$ and $\Phi_i = \angle{X_i}$ and
+Let $\{w_i\}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables.  Let $R_i > 0$ and $\Phi_i \in [-\pi,\pi)$ be random variables satisfying
 \[
-h = \prob X_i e^{-j\round{\Phi_i}} = \prob X_1 e^{-j\round{\Phi_1}}.
+R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*.
+\]
+and define the functions
+\[
+h_1(x) = \prob R_1 \cos(x + \Phi_1) \;\;\; \text{and} \;\;\; h_2(x) =  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
 \]
@@
 \] 
-as $L \rightarrow \infty$ and each of $p$ and $d$ and $h$ are positive real numbers, then:
+as $L \rightarrow \infty$ for positive real numbers $p$ and $d$ and if $h_2(x)$ is uniquely maximised at $x = 0$ over the interval $[\tfrac{-\pi}{M},\tfrac{-\pi}{M})$, then:
 \begin{enumerate}
 \item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0 p + d h$ almost surely as $L \rightarrow \infty$,
-\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0(p + dh)$.
+\item $\hat{\rho} \rightarrow \rho_0( p h_1(0) + d h_2(0) )$ almost surely as $L \rightarrow \infty$,
+\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0(p h_1(0) + d h_2(0) )$.
 Then the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal distribution with zero mean and covariance
@@
 BLERG  Show $E S_L$ is maximised uniquely at 0, proof follows.
-
-Observe that $Z(\theta) = e^{-j\theta}Y(\{\hat{d}_i(\hat{\theta}), i \in D\})$.  From~\eqref{eq:ahatYhat} we obtain
-\begin{align}\label{eq:hata}
-\hat{a} e^{-j\theta} &= \hat{\rho} e^{j(\hat{\theta} - \theta)} = \hat{\rho} e^{j\lambda_L} \nonumber \\
-&= \frac{1}{L} e^{-j\theta} Y(\{\hat{d}_i(\hat{\theta}), i \in D\}) \nonumber \\
-&= \frac{1}{L}  Z(\hat{\theta}) \nonumber \\
-&= \frac{1}{L} \sum_{i \in P} X_i e^{j\lambda_L} + \frac{1}{L} \sum_{i \in D}X_i e^{j( \lambda_L  - \round{\lambda_L + \Phi_i})} \label{eq:ahatfornorm}
-\end{align}
-and muliplying both sides of this equation by $e^{-j\lambda_L}$,
-\[
-\hat{\rho} =  \frac{1}{L}\sum_{i \in P} X_i + \frac{1}{L} \sum_{i \in D} X_i e^{-j\round{\lambda_L + \Phi_i}}
-\]
-By the strong law of large numbers, and since $\lambda_L \rightarrow 0$ (BLERG needs more care than this, but it's true),
-\begin{align*}
-\frac{1}{L}\sum_{i \in P}X_i \rightarrow p \rho_0 \\
-\frac{1}{L} \sum_{i \in D} X_i e^{-j\round{\lambda_L + \Phi_i}} \rightarrow  d h 
-\end{align*}
-and it immediately follows that $\hat{\rho} \rightarrow \rho_0 p + d h$.  This proves statement 2 of our theorem.
-
-We have
-\begin{align*}
-\sqrt{L} m_L =  A_L + B_L
-\end{align*}
-where
-\[
-A_L = \frac{1}{\sqrt{L}} \sum_{i \in P}w_i,
-\]
-and
-\[
-B_L = \frac{1}{\sqrt{L}} \sum_{i \in D}X_i e^{-j\round{\lambda_L + \Phi_i}} - h.
-\]
-Now $A_L$ converges in distribution to the complex normal with zero mean and variance $p\sigma^2$ and $B_L$ converges in probability to
-\[
-\lambda_L \sqrt{L} \sum_{k=-M}^{M-1} g(\psi_k) e^{j\psi_k}
-\]
-where $\psi_k = \frac{2\pi}{M}k + \frac{\pi}{M}$ and where
-\[
-g(\phi) = \frac{\rho_0}{2\pi} \int_{0}^\infty \frac{r^2f_W(\rho_0^{-1}\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}} dr
-\]
 

### 2012-05-08 paper.tex
Added more to paper.

@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:consistency}
 Let $\{w_i\}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables.  Let $R_i > 0$ and $\Phi_i \in [-\pi,\pi)$ be random variables satisfying
@@
 and define the functions
-\[
-h_1(x) = \prob R_1 \cos(x + \Phi_1) \;\;\; \text{and} \;\;\; h_2(x) =  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
-\]
+\begin{align*}
+h_1(x) &= \prob R_1 \cos(x + \Phi_1) \;\;\; \text{and}, \\
+h_2(x) &=  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
+\end{align*}
 Given the observed noisy data symbols $\{y_i, i \in P \cup D\}$ from~\eqref{eq:sigmod} let $\hat{a}$ be the least squares estimator of $a_0$.
@@
 \item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0( p h_1(0) + d h_2(0) )$ almost surely as $L \rightarrow \infty$,
-\item Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0(p h_1(0) + d h_2(0) )$.
-Then the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal distribution with zero mean and covariance
+\item $\hat{\rho} \rightarrow \rho_0 \big( p h_1(0) + d h_2(0) \big)$ almost surely as $L \rightarrow \infty$.
+\end{enumerate}
+\end{theorem}
+
+BLERG discuss assumptions.
+We have that $h_1$ is uniquely maximised at $0$ over the interval $[-\pi,\pi)$ since $\{w_i\}$ has zero mean.
+The requirement that $h_2$ is uniquely maximised at 0 is stronger that we need.
+We actually only require that the sum $p h_1(x) + dh_2(x)$ is maximised uniquely at 0 over the interval $[-\pi, \pi)$.
+However, we have stated the theorem in this stronger form as it makes the proof in the noncoherent case (i.e.
+when there are no pilot symbols) follow more transparently.
+
+\begin{theorem}\label{thm:normality}
+Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
+\[
+g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
+\]
+Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0\big(p h_1(0) + d h_2(0) \big)$.
+If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at multiples of $\tfrac{\pi}{M}$, then the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal distribution with zero mean and covariance
 \[
@@
 as $L \rightarrow \infty$.
-\end{enumerate}
 \end{theorem}
 
-BLERG $\hat{a}$ is a biased estimator, particularly when the SNR is low, we show firs that $\hat{\theta}$ is strongly consistent, we are then able to analyse the behaviour of $\hat{a}$.
+%BLERG $\hat{a}$ is a biased estimator, particularly when the SNR is low, we show firs that $\hat{\theta}$ is strongly consistent, we are then able to analyse the behaviour of $\hat{a}$.
 
@@
 Because the noise $\{w_i\}$ is circularly symetric then
-\[
-y_i = a_0 s_i + w_i = s_i e^{j \theta_0}( \rho_0 + w_i s_i^* e^{-j \theta_0}) = R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }
-\] 
-The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.
+\begin{align*}
+y_i &= a_0 s_i + w_i \\
+&= \rho_0 s_i e^{j \theta_0}\left( 1 + \frac{1}{\rho_0} w_i s_i^* e^{-j \theta_0} \right) \\
+&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
+\end{align*}
+%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
 Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
@@
 \begin{align*}
- y_i e^{-j\theta} \hat{d}_i(\theta)^* &= R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
-&= R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
-&= R_i e^{j\sfracpart{\lambda + \Phi_i}}
+ y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
+&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
+&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}}
 \end{align*}
@@
 \[
-y_i e^{-j\theta} p_i^* = R_i e^{j(\lambda + \Phi_i)}
+y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
 \]
 Now we can write
-\begin{align}
-Z(\theta) &= \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
-\end{align}
+\[
+Z(\theta) &= \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
+\]
 and
 \[
-\Re(Z(\theta)) = \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \sum_{i \in D} R_i \cos(\sfracpart{\lambda + \Phi_i})
+\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
 \] 
-Let $S_L(\lambda) = \frac{1}{L^2}\Re(Z)^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
+Let $S_L(\lambda) = L^{-2}\rho_0^{-2}\Re(Z(\theta))^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
 Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.
@@
 
-BLERG  Show $E S_L$ is maximised uniquely at 0, proof follows.
+Put $G_L(\lambda) = \Re(Z(\theta))$ and put
+\[
+G(\theta) = \lim_{L \rightarrow\infty} \prob G_L(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
+\]
+Lemma~\ref{lem:Zconunif} shows that,
+\[
+\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - \prob G_L(\lambda)} \rightarrow 0
+\]
+as $N\rightarrow 0$.  As a result of the theorem statement that $G(\lambda)$ is uniquely maximised at $0$ and therefore 
+\[
+0 \leq G(\lambda_L) - G(0).
+\]
+Since $\lambda_L$ is the maximiser of $G_L$ then
+\[
+0 \leq G_L(0) - G_L(\lambda_L).
+\]
+Adding both inequalities and taking appropriate absolute values gives
+\begin{align*}
+0 &\leq G(\lambda_L) - G(0) \\ 
+& \leq G(\lambda_L) - G(0) + G_L(0) - G_L(\lambda_L) \\
+& \leq |G(\lambda_L) - G_L(\lambda_L)| + |G_L(0) - G(0)|,
+\end{align*}
+and because of~\eqref{eq:supELNLN} the last line converges almost surely to zero as $L \rightarrow \infty$ and therefore $G(\lambda_L) \rightarrow G(0)$ almost surely.
+To complete the proof of strong consistency we use the device of Jenrich~\cite{Jenrich_asymp_nonlin_ulln_1969}.
+We use the
+subscript $N$ to emphasise that $\hat{\lambda}$ depends on $N.$ Suppose that the sequence
+$\left\{ \hat{\lambda}_{N}\right\} $ does not converge almost surely to $0$ as $N\rightarrow\infty,$ and let $A$ be the subset of the sample space on which $\left\{  \hat{\lambda}_{N}\right\}  $ does not converge to 0 as $N\rightarrow\infty$.
+Then $P\left(  A\right)  >0$.
+For any point in $A$, there is an infinite subsequence $\left\{  \hat{\lambda}_{N_{j}}\right\}$ which converges to some $\lambda^{\prime}\neq0$ in the compact interval $[-\nicefrac{1}{2}, \nicefrac{1}{2}]$.
+Thus $Q\left(\hat{\lambda}_{N_{j}}\right)$ converges to $Q\left(\lambda^{\prime}\right)$, since $Q$ is continuous in $\lambda$.
+However, $Q\left(\lambda^{\prime}\right) > Q\left(  0\right)$, contradicting the fact that $Q\left( \hat{\lambda}_{N_{j}}\right)  $ converges to $Q(0)$.
+It follows that $\left\{\hat{\lambda}_{N}\right\}$ converges almost
+surely to $0$ as $N\rightarrow\infty$.
+
+\begin{lemma}\label{lem:Zconunif}
+Let $G_L(\lambda) = \re(Z(\theta)$.  Then
+\[
+\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} \rightarrow 0
+\]
+as $L\rightarrow 0$.
+\end{lemma}
 

### 2012-05-09 paper.tex
Added to paper

@@
 %opening
-\title{Carrier phase and amplitude estimation using pilots and data}
-\author{Robby McKilliam}
+\title{Carrier phase and amplitude estimation of $M$-ary phase shift keying using pilots and data}
+\author{Robby McKilliam, Andre Pollok, Bill Cowley}
 
@@
 \begin{abstract}
-BLERG
+We describe least squares estimators of carrier phase and signal amplitude from a noisy transmitted communications signal that contains both pilots, known to the receiver, and data, unknown to the receiver.
+We focus on signaling constellations that have symbols lying on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and M-PSK.
+We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of the carrier phase is strongly consistent and asymptotically normaly distributed.
+However, the amplitude estimator is not consistent, and we show that it converges to a positive real number that is a function of true carrier amplitude, the noise distribution and the size of the transmission constellation.
+Our theoretical results can also be applied to the case where no pilot symbols exist (i.e.
+noncoherent detection).
+The results of Monte Carlo simulations are provided and these corroborate the theoretical results.
 \end{abstract}
@@
 \end{IEEEkeywords}
-
-In this document we describe estimators of carrier phase and signal amplitude of a transmitted communications signal that contains both pilots, known to the receiver, and data, unknown to the receiver.
-We focus on signaling constellations that have symbols lying on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and M-PSK.
 
@@
 \end{equation}
-where $w_i$ is noise and $a_0$ is a complex number representing both carrier phase and signal amplitude.
+where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and signal amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
 Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
@@
 Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator and $\abs{x}$ denotes the magnitude of the complex number $x$.
-It will actually be informative to consider the magnitude and complex argument (or phase) of the estimator separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ and $a_0 = \rho_0e^{j\theta_0}$ where both $\rho_0$ and $\hat{\rho}$ are positive real numbers, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
-The function $\fracpart{\cdot}$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\prob\vert\langle\theta_0 - \hat{\theta}\rangle\vert^2$ rather than $\prob\vert\theta_0 - \hat{\theta}\vert^2$ is the appropriate measure of error for the phase parameter.
+It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is positive real numbers, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle_\pi\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
+The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
+It will become apparent why $\prob\vert\langle\theta_0 - \hat{\theta}\rangle_\pi\vert^2$ rather than $\prob\vert\theta_0 - \hat{\theta}\vert^2$ is the appropriate measure of error for the phase parameter.
 
-It is worth commenting on our use of $\prob$ rather than the more usual $\expect$ or $E$ to denote the expected value operator.
+It is worth commenting on our use of $\prob$ rather than the more usual $\expect$ or $E$ for the expected value operator.
 The notation comes from Pollard~\cite[Ch 1]{Pollard_users_guide_prob_2002} and is originally due to de Finetti.
@@
 \begin{align}
-SS(a, \{d_i, i \in D\}) &= \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
+SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
 &= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*. \label{eq:SS}
@@
 To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a positive real.  Now the sum of squares function can be written as
-\begin{align*}
-SS(\rho, \theta, \{d_i, i \in D\}) &= \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \\
-&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2.
-\end{align*}
+\begin{align}
+SS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
+&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:SSallparams}
+\end{align}
 For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
@@
 \[
-\fracpart{x} = x - \frac{2\pi}{M}\operatorname{round}\left(\frac{Mx}{2\pi}\right).
+\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
 \]
+The next two theorems decribe the asymptotic properties of the least squares estimator.
 
-\begin{theorem}\label{thm:consistency}
-Let $\{w_i\}$ be a sequence independent and identically distributed, zero mean, circularly symetric complex random variables.  Let $R_i > 0$ and $\Phi_i \in [-\pi,\pi)$ be random variables satisfying
+\begin{theorem}\label{thm:consistency} (Almost sure convergence)
+Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables and let $\{y_i, i \in P \cup D\}$ be observed noisy symbols given by the signal model~\eqref{eq:sigmod}.
+Let $\hat{a}$ be the least squares estimator of $a_0$ and put $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ and $a_0 = \rho_0 e^{j\theta_0}$ where both $\rho_0$ and $\hat{\rho}$ are positive real numbers.
+Put $L = \abs{P \cup D}$ and define real numbers $p$ and $d$ as the limits
 \[
-R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*.
+\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
 \] 
-and define the functions
+Let $R_i > 0$ and $\Phi_i \in [-\pi,\pi)$ be random variables satisfying
+\[
+R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*,
+\]
+and define the continuous function
+\[
+G(x) = p h_1(x) + d h_2(x)
+\]
+where
 \begin{align*}
@@
 \end{align*}
-Given the observed noisy data symbols $\{y_i, i \in P \cup D\}$ from~\eqref{eq:sigmod} let $\hat{a}$ be the least squares estimator of $a_0$.
-Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}$ and $a_0 = \rho_0 e^{j\theta_0}$.
-Put $L = \abs{P \cup D}$.
-If
-\[
-\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d
-\] 
-as $L \rightarrow \infty$ for positive real numbers $p$ and $d$ and if $h_2(x)$ is uniquely maximised at $x = 0$ over the interval $[\tfrac{-\pi}{M},\tfrac{-\pi}{M})$, then:
+If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
 \begin{enumerate}
 \item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0 \big( p h_1(0) + d h_2(0) \big)$ almost surely as $L \rightarrow \infty$.
+\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
 \end{enumerate}
@@
 
-BLERG discuss assumptions.
+\begin{theorem}\label{thm:normality} (Asymptotic normality)
+Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
+\[
+g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
+\]
+Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0\big(p h_1(0) + d h_2(0) \big)$.
+If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at multiples of $\tfrac{\pi}{M}$, then the the distribution of the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal with zero mean and covariance
+\[
+C = ?
+\]
+as $L \rightarrow \infty$.
+\end{theorem}
+
+%BLERG discuss assumptions.
 We have that $h_1$ is uniquely maximised at $0$ over the interval $[-\pi,\pi)$ since $\{w_i\}$ has zero mean.
@@
 
-\begin{theorem}\label{thm:normality}
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
-\[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
-\]
-Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0\big(p h_1(0) + d h_2(0) \big)$.
-If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at multiples of $\tfrac{\pi}{M}$, then the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal distribution with zero mean and covariance
-\[
-C = ?
-\]
-as $L \rightarrow \infty$.
-\end{theorem}
-
 %BLERG $\hat{a}$ is a biased estimator, particularly when the SNR is low, we show firs that $\hat{\theta}$ is strongly consistent, we are then able to analyse the behaviour of $\hat{a}$.
 
-Again let $a = \rho e^{j\theta}$ and write the sum of squares function  
-\[
-SS(\rho, \theta, \{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{y_i}^2 - \rho y_i e^{-j\theta} s_i^* - \rho y_i e^{-\theta} s_i + \rho^2.
-\]
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives 
-\[
-\rho = \frac{Z + Z^*}{2L} = \frac{1}{L}\Re(Z)
-\]
-where $Z = \sum_{i\in P \cup D}y_i e^{-j\theta} s_i^*$ and $\Re(\cdot)$ denotes the real part.
-Observe that $Z$ is really a function of $\theta$ and $\{d_i, i \in D\}$, but we have supressed the arguments for notational clarity.
-We can now write
-\[
-SS(\rho, \theta, \{d_i, i \in D\}) = A - \rho Z - \rho Z^* + L \rho^2
-\]
-and substituting the above expression for $\rho$ into this gives the sum of squares function conditioned upon minimisation with respect to $\rho$,
-\[
-SS(\theta, \{d_i, i \in D\}) = A - \frac{1}{L}\Re(Z)^2.
-\]
-Substituting~(\ref{eq:hatdfinxtheta}) into this gives the sum of squares function conditioned upon minimisation with respect to $\rho$ and the data sequence $\{d_i, i \in D\}$,
-\[
-SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2
-\]
+\begin{IEEEproof} (Theorem~\ref{thm:consistency})
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function conditioned on minimisation with respect the the data symbols
+ \begin{align*}
+SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
+&= A - \rho Z(\theta) - \rho Z(\theta)^* + L \rho^2,
+\end{align*}
 where
@@
 \]
-We have reused $Z$ here, but this should not cause confusion as $Z(\theta)$ and $Z(\theta, \{d_i, i \in D\})$ are easily told apart from there arguments.
-%From now on, when we write $Z$ without arguments, we will always be refering to $Z(\theta)$.
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives 
+\[
+\hat{\rho} = \frac{Z(\theta) + Z(\theta)^*}{2L} = \frac{1}{L}\Re(Z(\theta))
+\]
+where $\Re(\cdot)$ denote the real part.
+Substituting this expression into $SS(\rho, \theta)$ above gives the sum of squares function conditioned upon minimisation with respect to the amplitude $\rho$ and the data symbols,
+\[
+SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
+\]
 
-We are interested in analysing $\hat{\theta}$, the maximiser of $\abs{Z(\theta)}^2$.
-Because the noise $\{w_i\}$ is circularly symetric then
+We are interested in analysing $\hat{\theta}$, the maximiser of $\Re(Z(\theta))^2$.  Recalling the definition of $R_i$ and $\Phi_i$ from the statement of Theorem~\ref{thm:consistency},
 \begin{align*}
@@
 \[
-Z(\theta) &= \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
+Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
 \]
@@
 
-Put $G_L(\lambda) = \Re(Z(\theta))$ and put
+Put $G_L(\lambda) = \frac{1}{\rho_0}\Re(Z(\theta))$ and observe that
 \[
@@
 \]
-Lemma~\ref{lem:Zconunif} shows that,
+Lemma~\ref{lem:convtoexpGlamL} shows that,
 \[
-\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - \prob G_L(\lambda)} \rightarrow 0
+\prob\{ \abs{G(\lambda_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L})
 \]
-as $N\rightarrow 0$.  As a result of the theorem statement that $G(\lambda)$ is uniquely maximised at $0$ and therefore 
+Since $G(\lambda)$ is continuous and is uniquely maximised at $0$, then for any $\delta > 0$ there is a $\epsilon > 0$ such that 
 \[
-0 \leq G(\lambda_L) - G(0).
+\abs{\lambda_L} > \delta \;\; \Rightarrow \;\; \abs{G(\lambda_L) - G(0)} > \epsilon
 \]
-Since $\lambda_L$ is the maximiser of $G_L$ then
+and therefore for any $\delta > 0$ there is an $\epsilon > 0$ such that
+\[
+\prob\{\abs{\lambda_L} > \delta\} \leq \prob\{ \abs{G(\lambda_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L}).
+\]
+Now the sum 
+\[
+\sum_{L=1}^{\infty}\prob\{ \abs{\lambda_L} > \delta \} = \sum_{L=1}^{\infty}O(e^{-\epsilon^2 L})
+\]
+converges and $\lambda_L$ converges almost surely to zero by the Borel-Cantelli lemma.
+
+BLERG.  Need to show that it's actually sufficient to study the maximiser of $G_L$ rather than $\abs{G_L}$ (atleast asymptotically).
+
+BLERG.  It follows after a bit of discussion and the use of Lemma~\ref{lem:Zconunif} that 
+\[
+\hat{\rho} = \rho_0 G_L(\lambda_L) \rightarrow \rho_0 G(0). 
+\] 
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:convtoexpGlamL}
+\[
+\prob\{ \abs{G(\lambda_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L})
+\]
+\end{lemma}
+\begin{IEEEproof}
+Lemma~\ref{lem:Zconunif} shows BLERG that 
+\[
+0 \leq G(\lambda_L) - G(0),
+\]
+and since $\lambda_L$ is the maximiser of $G_L$ then
 \[ 
@@
 \end{align*}
-and because of~\eqref{eq:supELNLN} the last line converges almost surely to zero as $L \rightarrow \infty$ and therefore $G(\lambda_L) \rightarrow G(0)$ almost surely.
-To complete the proof of strong consistency we use the device of Jenrich~\cite{Jenrich_asymp_nonlin_ulln_1969}.
-We use the
-subscript $N$ to emphasise that $\hat{\lambda}$ depends on $N.$ Suppose that the sequence
-$\left\{ \hat{\lambda}_{N}\right\} $ does not converge almost surely to $0$ as $N\rightarrow\infty,$ and let $A$ be the subset of the sample space on which $\left\{  \hat{\lambda}_{N}\right\}  $ does not converge to 0 as $N\rightarrow\infty$.
-Then $P\left(  A\right)  >0$.
-For any point in $A$, there is an infinite subsequence $\left\{  \hat{\lambda}_{N_{j}}\right\}$ which converges to some $\lambda^{\prime}\neq0$ in the compact interval $[-\nicefrac{1}{2}, \nicefrac{1}{2}]$.
-Thus $Q\left(\hat{\lambda}_{N_{j}}\right)$ converges to $Q\left(\lambda^{\prime}\right)$, since $Q$ is continuous in $\lambda$.
-However, $Q\left(\lambda^{\prime}\right) > Q\left(  0\right)$, contradicting the fact that $Q\left( \hat{\lambda}_{N_{j}}\right)  $ converges to $Q(0)$.
-It follows that $\left\{\hat{\lambda}_{N}\right\}$ converges almost
-surely to $0$ as $N\rightarrow\infty$.
+BLERG
+\end{IEEEproof}
 
@@
 \[
-\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} \rightarrow 0
+\prob\{\abs{G_L(\lambda) - G(\lambda)} > \epsilon \} = O(e^{-\epsilon^2 L}).
 \]
-as $L\rightarrow 0$.
+\end{lemma}
+\begin{IEEEproof}
+\end{IEEEproof}
+
+\begin{IEEEproof} (Theorem~\ref{thm:normality})
+The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.
+This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the minimiser $\lambda}_L$.
+So
+\[
+G_L^\prime(\lambda_L) = 0 = \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+\]
+Write
+\[
+0 = G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) + \prob G_L^\prime(\lambda_L)
+\]
+where
+\[
+\prob G^\prime_L = \sum_{i \in P}
+\]
+\end{IEEEproof}
+
+\begin{lemma}~\label{lem:diffatlambdaL}
+The derivative of $G_L$ existsm, and is equal to zero, at $\lambda_L$.  That is,
+\[
+G_L^\prime(\lambda_L) = \frac{d G_L}{d \lambda}(\lambda_L) = 0.
+\]
 \end{lemma}
@@
 
-\subsection{The coherent setting}
+%\subsection{The coherent setting}
 
@@
 
-\subsection{The noncoherent setting}
+%\subsection{The noncoherent setting}
 

### 2012-05-10 paper.tex
Fixed silly error in the plots and resized them so that they fit 4 to a page.  More edits to proofs.

@@
 $\phi = \angle{y_i}$ \;
-$u = \frac{2\pi}{M}\round{\frac{M}{2\pi}\phi} $ \;
+$u = \round{\phi} $ \;
 $z_i = \phi -  u $ \;
@@
 $\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
+$\eta = e^{j2\pi/M} - 1$ \;
 \For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
-$Y = Y + (e^{j2\pi/M} - 1) g_{\sigma(k)}$ \;
-$g_{\sigma(k)} = g_{\sigma(k)} e^{j2\pi/M} $\;
+$Y = Y + \eta g_{\sigma(k)}$ \;
+$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
 $Q = \frac{1}{L}\abs{Y}^2$\;
@@
 \begin{lemma}\label{lem:Zconunif}
-Let $G_L(\lambda) = \re(Z(\theta)$.  Then
+Let $G_L(\lambda) = \re(Z(\theta))$.  Then
 \[
@@
 This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the minimiser $\lambda}_L$.
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the minimiser $\lambda_L$.
 So
 \[
-G_L^\prime(\lambda_L) = 0 = \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+G_L^\prime(\lambda_L) = 0 = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
 \]
@@
 \[
-\prob G^\prime_L = \sum_{i \in P}
+\prob G^\prime_L(\lambda) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda),
+\]
+and where,
+\begin{align*}
+k_1(x) &= \prob R_1 \sin(x + \Phi_1) \;\;\; \text{and}, \\
+k_2(x) &=  \prob R_1 \sin\sfracpart{x + \Phi_1}.
+\end{align*}
+Lemma~\ref{lem:GprimeLlambdaLconv} shows that
+\[
+\sqrt{L} \prob G^\prime_L(\lambda_L) = \sqrt{L} \lambda_L( G(0)  + o_P(1) )
+\]
+and Lemma~\ref{lem:empiricprocc} shows that
+\[ 
+\sqrt{L} \left( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \right) = \sqrt{L} G_L'(0) + \sqrt{L}\lambda_L( ? +  o_P(1) )
+\]
+It follows that
+\[
+\sqrt{\lambda_L} = o_P(1) + \frac{\sqrt{L}G_L'(0)}{G(0)  + o_P(1)}.
+\]
+Observe that 
+\[
+\sqrt{L} G_L'(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+\]
+It follows from the standard central limit that the distribution of
+\[
+\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{P}}{L}} \frac{1}{\sqrt{\abs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
+\]
+converges to the normal with mean $k_1(0) = 0$ and variance 
+\[
+p \prob R_1^2 \sin^2(\Phi_1).
+\]
+Similarly, the distribution of 
+\[
+\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{D}}{L}} \frac{1}{\sqrt{\abs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
+\]
+converges to the normal with mean $k_2(0) = 0$ and variance
+\[
+d \prob R_1^2 \sin^2\fracpart{\Phi_1}.
+\]
+It follows that the distribution of $\sqrt{L}\lambda_L$ converges to the normal with mean zero and variance
+\[
+\frac{p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}}{G(0)^2}.
 \]
@@
 \begin{lemma}~\label{lem:diffatlambdaL}
-The derivative of $G_L$ existsm, and is equal to zero, at $\lambda_L$.  That is,
+The derivative of $G_L$ exists, and is equal to zero, at $\lambda_L$.  That is,
 \[
@@
 \end{lemma}
+
+\begin{lemma}\label{lem:GprimeLlambdaLconv}
+\[
+\prob G^\prime_L(\lambda_L) = \lambda_L( G(0)  + o_P(1) )
+\]
+\end{lemma}
+\begin{IEEEproof}
+We have (BLERG) that $k_1$ and $k_2$ are continuous and differentiable and that $k_1(0) = k_2(0) = 0$, so by a first order Taylor expansion
+\[
+\prob G^\prime_L(\lambda_L) = \prob G^\prime_L(0) + \lambda_L( \prob G^\prime_L(\lambda_L)  + o_P(1) )
+\]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:empiricprocc} BLERG
+\[ 
+\sqrt{L} \left( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \right) = \sqrt{L} G_L'(0) + \sqrt{L}\lambda_L( ? +  o_P(1) )
+\]
+\end{lemma}
+\begin{IEEEproof}
+\end{IEEEproof}
 
@@
 %BPSK plots
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotM2-1.mps}
+		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
 		\caption{Amplitude error for BPSK}
 		\label{fig:plotamp}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotM2-2.mps}
+		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
 		\caption{Phase error for BPSK}
 		\label{fig:plotphase}
-\end{figure*}
+\end{figure}
 
@@
 %QPSK plots
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotM4-1.mps}
+		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
 		\caption{Amplitude error for QPSK}
 		\label{fig:plotamp}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotM4-2.mps}
+		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
 		\caption{Phase error for QPSK}
 		\label{fig:plotphase}
-\end{figure*}
+\end{figure}
 
-% \begin{figure*}[tp]
+% \begin{figure}[tp]
 % 	\centering
@@
 % 		\label{fig:plotcamp}
-% \end{figure*}
+% \end{figure}
 
 %8PSK plots
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotM8-1.mps}
+		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
 		\caption{Amplitude error for 8PSK}
 		\label{fig:plotamp}
-\end{figure*}
+\end{figure}
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotM8-2.mps}
+		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
 		\caption{Phase error for 8PSK}
 		\label{fig:plotphase}
-\end{figure*}
+\end{figure}
 
-% \begin{figure*}[tp]
+% \begin{figure}[tp]
 % 	\centering
@@
 % 		\label{fig:plotcamp}
-% \end{figure*}
+% \end{figure}
 
@@
 %BPSK plots
- \begin{figure*}[tp]
+ \begin{figure}[tp]
  	\centering
- 		\includegraphics{code/data/plotM2-1.mps}
+ 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
  		\caption{Amplitude error for BPSK}
  		\label{fig:plotamp}
- \end{figure*}
+ \end{figure}
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotncM2-2.mps}
+		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
 		\caption{Phase error for BPSK}
 		\label{fig:plotphase}
-\end{figure*}
+\end{figure}
 
-% \begin{figure*}[tp]
+% \begin{figure}[tp]
 % 	\centering
@@
 % 		\label{fig:plotcamp}
-% \end{figure*}
+% \end{figure}
 
 %QPSK plots
- \begin{figure*}[tp]
+ \begin{figure}[tp]
  	\centering
- 		\includegraphics{code/data/plotM4-1.mps}
+ 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
  		\caption{Amplitude error for QPSK}
  		\label{fig:plotamp}
- \end{figure*}
+ \end{figure}
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotncM4-2.mps}
+		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
 		\caption{Phase error for QPSK}
 		\label{fig:plotphase}
-\end{figure*}
+\end{figure}
 
- % \begin{figure*}[tp]
+ % \begin{figure}[tp]
  % 	\centering
@@
  % 		\label{fig:plotcamp}
- % \end{figure*}
+ % \end{figure}
 
 %8PSK plots
- \begin{figure*}[tp]
+ \begin{figure}[tp]
  	\centering
- 		\includegraphics{code/data/plotM8-1.mps}
+ 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
  		\caption{Amplitude error for 8PSK}
  		\label{fig:plotamp}
- \end{figure*}
+ \end{figure}
 
-\begin{figure*}[tp]
+\begin{figure}[tp]
 	\centering
-		\includegraphics{code/data/plotncM8-2.mps}
+		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
 		\caption{Phase error for 8PSK}
 		\label{fig:plotphase}
-\end{figure*}
+\end{figure}
 
-% \begin{figure*}[tp]
+% \begin{figure}[tp]
 % 	\centering

### 2012-05-10 paper.tex
Added to paper.  Still more to do, but it's moving in the right direction

@@
 We describe least squares estimators of carrier phase and signal amplitude from a noisy transmitted communications signal that contains both pilots, known to the receiver, and data, unknown to the receiver.
-We focus on signaling constellations that have symbols lying on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and M-PSK.
+We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-PSK.
 We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of the carrier phase is strongly consistent and asymptotically normaly distributed.
@@
 
-In this setting, the transmitted symbols take the form,
+\section{Introduction}
+
+With $M$-ary phase shift keying ($M$-PSK), the transmitted symbols take the form,
 \[
@@
 \]
-where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation, i.e.
-we have assumed an $M$-PSK constellation.
+where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation.
 We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.
@@
 Complicating matters is that the data symbols $d_i$ are not known to the reciever.
+Estimation problems of this type have undergone extensive prior study~\cite{Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
+A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
+\begin{align*}
+SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \\
+&= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in P \cup D} \abs{ y_i - a d_i }^2,
+\end{align*}
+where $\abs{x}$ denotes the magnitude of the complex number $x$.
+This is the estimator we study in this paper.
+The least squares estimator is also the maximum likelihood under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
 
-Estimation problems of this type have undergone extensive prior study~\cite{Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
 The existing literature mostly considers what is called \emph{noncoherent detection} where no pilots symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
 In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
-Arguably, the literature culminates in the paper of Mackenthun~\cite{Mackenthun1994} who described a least squares estimator (maximum likelihood under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian) requiring only $O(L \log L)$ arithmetic operations, where $L = \abs{P \cup D}$ is the total number of symbols transmitted.
+Arguably, the literature culminates in the paper of Mackenthun~\cite{Mackenthun1994} who described an algorithm to compute the least squares estimator requiring only $O(L \log L)$ arithmetic operations, where $L = \abs{P \cup D}$ is the total number of symbols transmitted.
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
@@
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-We show in Section~\ref{sec:least-squar-estim}~ that by making a minor modification to Mackenthun's estimator, it is possible to meanifully include pilot symbols.
+We show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
 
@@
 Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
-This is motivated by the fact that in our intended application (narrowband satellite communications) and in most modern communication systems the data symbols are \emph{coded}.
+This is motivated by the fact that in our intended application (narrowband satellite communications) and in many modern communication systems the data symbols are \emph{coded}.
 For this reason raw symbol error rate is not of specific interest at this stage.
@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator and $\abs{x}$ denotes the magnitude of the complex number $x$.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is positive real numbers, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle_\pi\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator.
+It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle_\pi\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
@@
 
+%\subsection{Notation}
 It is worth commenting on our use of $\prob$ rather than the more usual $\expect$ or $E$ for the expected value operator.
-The notation comes from Pollard~\cite[Ch 1]{Pollard_users_guide_prob_2002} and is originally due to de Finetti.
+The notation comes from Pollard~\cite[Ch 1]{Pollard_users_guide_prob_2002}.
 The notation is good because it removes unecessary distinction between `probability' and expectation.
@@
 
-\section{Mackentun's least squares estimator}\label{sec:least-squar-estim}
+\section{Mackentun's algorithm}\label{sec:least-squar-estim}
 
-In this section we derive Mackentun's least squares estimator~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to meaninfully include the pilot symbols, so we also consider the coherent setting here.
+In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
+Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to include the pilot symbols.
 For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 
-To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a positive real.  Now the sum of squares function can be written as
+To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a nonnegative real.  Now the sum of squares function can be written as
 \begin{align}
@@
 and where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
-A word of caution here, it is often the case that $\round{\cdot}$ denotes rounding to the nearest \emph{integer}.
+A word of caution, it is often the case that $\round{\cdot}$ denotes rounding to the nearest \emph{integer}.
 This is not the case here.
@@
 Note that $\hat{d}_i(\theta)$ does not depend on the amplitude $\rho$.
-As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered modulo $2\pi$.
+As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
 With this in mind,
@@
 \]
-It is clear from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i, i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., the minimiser of~\eqref{eq:SSdatasymbols}.
+If $\hat{\theta}$ is the least squares estimator the phase phase, then it is clear from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
 Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
@@
 This requires $O(L \log L)$ operations.
-The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm.
+The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.
 The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
@@
 \SetAlTitleFnt{}
-\caption{Mackenthun's least squares estimator}
+\caption{Mackenthun's algorithm with pilot symbols}
 \DontPrintSemicolon
@@
 Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables and let $\{y_i, i \in P \cup D\}$ be observed noisy symbols given by the signal model~\eqref{eq:sigmod}.
-Let $\hat{a}$ be the least squares estimator of $a_0$ and put $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ and $a_0 = \rho_0 e^{j\theta_0}$ where both $\rho_0$ and $\hat{\rho}$ are positive real numbers.
+Let $\hat{a}$ be the least squares estimator of $a_0$ and put $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number.
 Put $L = \abs{P \cup D}$ and define real numbers $p$ and $d$ as the limits
@@
 \] 
-Let $R_i > 0$ and $\Phi_i \in [-\pi,\pi)$ be random variables satisfying
-\[
+Let $R_i > 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
+\begin{equation}\label{eq:RiandPhii}
 R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*,
-\]
+\end{equation}
 and define the continuous function
@@
 \end{align*}
-If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
+If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
 \begin{enumerate}
@@
 Differentiating with respect to $\rho$ and setting the resulting expression to zero gives 
-\[
+\begin{equation}\label{eq:hatrhoZ}
 \hat{\rho} = \frac{Z(\theta) + Z(\theta)^*}{2L} = \frac{1}{L}\Re(Z(\theta))
-\]
+\end{equation}
 where $\Re(\cdot)$ denote the real part.
@@
 
-We are interested in analysing $\hat{\theta}$, the maximiser of $\Re(Z(\theta))^2$.  Recalling the definition of $R_i$ and $\Phi_i$ from the statement of Theorem~\ref{thm:consistency},
+We are interested in analysing $\hat{\theta}$, the maximiser of $\Re(Z(\theta))^2$ or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
+Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
 \begin{align*}
@@
 \] 
-Let $S_L(\lambda) = L^{-2}\rho_0^{-2}\Re(Z(\theta))^2$ and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
-Then $\lambda_L$ is the maximiser of $S_L(\lambda)$.
+Let 
+\[
+G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
+\]
+and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
+Then $\lambda_L$ is the maximiser of $G_L(\lambda)^2$ and equivalently the maximiser of $\abs{G_L(\lambda)}$.
 We will show that $\lambda_L$ converges almost surely to zero as $L \rightarrow \infty$.
@@
 
-Put $G_L(\lambda) = \frac{1}{\rho_0}\Re(Z(\theta))$ and observe that
+Observe that
+\[
+\prob G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
+\]
+and since $\abs{P}}{L} \rightarrow p$ and $\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
 \[
@@
 
-BLERG.  Need to show that it's actually sufficient to study the maximiser of $G_L$ rather than $\abs{G_L}$ (atleast asymptotically).
-
-BLERG.  It follows after a bit of discussion and the use of Lemma~\ref{lem:Zconunif} that 
-\[
-\hat{\rho} = \rho_0 G_L(\lambda_L) \rightarrow \rho_0 G(0). 
-\] 
+It remains to prove part two of this theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.
+From~\eqref{eq:hatrhoZ} the amplitude estimator satisfies $\hat{\rho} = \rho_0 G_L(\lambda_L)$.
+Lemma~\ref{lem:GLtoG0} shows that $G_L(\lambda_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and consequently $\hat{\rho}$ converges almost surely to $\rho_0 G(0)$ as required.
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:Zconunif}
-Let $G_L(\lambda) = \re(Z(\theta))$.  Then
 \[
-\prob\{\abs{G_L(\lambda) - G(\lambda)} > \epsilon \} = O(e^{-\epsilon^2 L}).
+\prob\left\{\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} > \epsilon \right\} = O(e^{-\epsilon^2 L}).
 \]
@@
 
+\begin{lemma}\label{lem:GLtoG0}
+$G_L(\lambda_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+\end{IEEEproof}
+
 \begin{IEEEproof} (Theorem~\ref{thm:normality})
+BLERG.  Need to show that it's actually sufficient to study the maximiser of $G_L$ rather than $\abs{G_L}$ (atleast asymptotically).
+
 The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.
 This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the minimiser $\lambda_L$.
-So
-\[
-G_L^\prime(\lambda_L) = 0 = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\]
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\lambda_L$, so
+\begin{align*}
+G_L^\prime(\lambda_L) &= 0 \\
+&= \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+\end{align*}
 Write
@@
 \]
-where
-\[
-\prob G^\prime_L(\lambda) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda),
-\]
-and where,
-\begin{align*}
-k_1(x) &= \prob R_1 \sin(x + \Phi_1) \;\;\; \text{and}, \\
-k_2(x) &=  \prob R_1 \sin\sfracpart{x + \Phi_1}.
-\end{align*}
 Lemma~\ref{lem:GprimeLlambdaLconv} shows that
 \[
-\sqrt{L} \prob G^\prime_L(\lambda_L) = \sqrt{L} \lambda_L( G(0)  + o_P(1) )
+\sqrt{L} \prob G^\prime_L(\lambda_L) = \prob G_L^\prime(0) + \sqrt{L} \lambda_L\big( G(0)  + o_P(1) \big)
 \]
-and Lemma~\ref{lem:empiricprocc} shows that
-\[ 
-\sqrt{L} \left( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \right) = \sqrt{L} G_L'(0) + \sqrt{L}\lambda_L( ? +  o_P(1) )
-\]
+where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.  Lemma~\ref{lem:empiricprocc} shows that
+\begin{align*}
+\sqrt{L}&\big( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \big) \\
+&= -\prob G_L^\prime(0) + \sqrt{L} G_L^\prime(0) + \sqrt{L}\lambda_L\big( ? +  o_P(1) \big).
+\end{align*}
 It follows that
 \[
-\sqrt{\lambda_L} = o_P(1) + \frac{\sqrt{L}G_L'(0)}{G(0)  + o_P(1)}.
+\sqrt{L} \lambda_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{G(0)  + o_P(1)}.
 \]
@@
 \[
-\sqrt{L} G_L'(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+\sqrt{L} G_L^\prime((0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
 \]
@@
 \begin{IEEEproof}
+Observe that
+\[
+\prob G^\prime_L(\lambda) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda),
+\]
+where,
+\begin{align*}
+k_1(x) &= \prob R_1 \sin(x + \Phi_1) \;\;\; \text{and}, \\
+k_2(x) &=  \prob R_1 \sin\sfracpart{x + \Phi_1}.
+\end{align*}
 We have (BLERG) that $k_1$ and $k_2$ are continuous and differentiable and that $k_1(0) = k_2(0) = 0$, so by a first order Taylor expansion

### 2012-05-11 paper.tex
minor edits
Added to paper.

@@
 \begin{abstract}
-We describe least squares estimators of carrier phase and signal amplitude from a noisy transmitted communications signal that contains both pilots, known to the receiver, and data, unknown to the receiver.
+We describe least squares estimators of carrier phase and amplitude from a noisy communications signal that contains both pilots, known to the receiver, and data, unknown to the receiver.
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-PSK.
 We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of the carrier phase is strongly consistent and asymptotically normaly distributed.
-However, the amplitude estimator is not consistent, and we show that it converges to a positive real number that is a function of true carrier amplitude, the noise distribution and the size of the transmission constellation.
+However, the amplitude estimator is not consistent, and we show that it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
 Our theoretical results can also be applied to the case where no pilot symbols exist (i.e.
@@
 The sets $P$ and $D$ are disjoint, i.e.
-$P \cap D = \emptyset$, and the union $P \cup D$ contains all those indices where the transmitter is active.
-Outside $P \cup D$ the transmitter is silent (not active).
+$P \cap D = \emptyset$.
 
@@
 \end{equation}
-where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and signal amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
+where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
 Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
-Complicating matters is that the data symbols $d_i$ are not known to the reciever.
+Complicating matters is that the data symbols $d_i$ are not known to the reciever and must also be estimated.
 Estimation problems of this type have undergone extensive prior study~\cite{Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
@@
 where $\abs{x}$ denotes the magnitude of the complex number $x$.
-This is the estimator we study in this paper.
-The least squares estimator is also the maximum likelihood under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
+The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
+However, as we will show, the estimator works well under less stringent assumptions.
+%It is the least squares estimator that we primarily study in this paper.
 
@@
 In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
-Arguably, the literature culminates in the paper of Mackenthun~\cite{Mackenthun1994} who described an algorithm to compute the least squares estimator requiring only $O(L \log L)$ arithmetic operations, where $L = \abs{P \cup D}$ is the total number of symbols transmitted.
+An important paper is by Mackenthun~\cite{Mackenthun1994} who described an algorithm to compute the least squares estimator requiring only $O(L \log L)$ arithmetic operations, where $L = \abs{P \cup D}$ is the total number of symbols transmitted.
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
@@
 
-In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and the complex gain $a_0$ is a nuisance parameter.
+In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
 The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex gain $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator.
 It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle_\pi\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
@@
 where $\indicator \{X \in S\}$ is the indicator function of the set $S$, i.e $\indicator \{X \in S\}(x) = 1$ when the argument $x \in S$ and zero otherwise.
-We will usually drop the $\onebf$ and simply write $\prob \{ X > \delta \}$ to mean $\prob \onebf\{ X > \delta \}$.
-As an example of the utility of this notation, Markov's inequality is plainly
+We will usually drop the $\onebf$ and simply write $\prob \{ X \in S \}$ to mean $\prob \onebf\{ X \in S \}$.
+To illustrate the utility of this notation, Markov's inequality becomes
 \[
-\prob \{ \abs{X} > \delta \}  \leq \prob \frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \} \leq \frac{1}{\delta}\prob\abs{X}.
+\prob \{ \abs{X} > \delta \}  \leq \prob \frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \} \leq \frac{1}{\delta}\prob\abs{X},
 \]
@@
 \]
-If $\hat{\theta}$ is the least squares estimator the phase phase, then it is clear from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
+If $\hat{\theta}$ is the least squares estimator the phase, then it is clear from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
 Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
@@
 \[
-f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}
+f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
 \]  
@@
 We want to find the sequence $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A naive approach would be to compute $SS(f_k)$ for each sequence in $S$ and return the minimiser.
+A na\"{\i}ve approach would be to compute $SS(f_k)$ for each sequence in $S$ and return the minimiser.
 Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
-So, this naive approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
+So, this na\"{\i}ve approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
 Mackenthun~\cite{Mackenthun1994} showed how $SS(f_k)$ can be computed recursively.
@@
 where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in the sequence $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
-Letting $g_{k} =\{ y_i f_{ki}^*, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
+Letting $g_{k}$ be the sequence $\{ y_i f_{ki}^*, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
 \[
@@
 \[
-\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
+\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right),
 \]
@@
 \[
-\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
+\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right) = x - \round{x}.
 \]
@@
 \] 
-Let $R_i > 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
+Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
 \begin{equation}\label{eq:RiandPhii}
@@
 \]
-Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0\big(p h_1(0) + d h_2(0) \big)$.
-If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at multiples of $\tfrac{\pi}{M}$, then the the distribution of the vector $[\sqrt{L}\lambda_L, \sqrt{L}m_L]$ converges to the bivariate normal with zero mean and covariance
+Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0 G(0)$.
+If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at multiples of $\tfrac{\pi}{M}$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
 \[
@@
 \begin{IEEEproof} (Theorem~\ref{thm:consistency})
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function conditioned on minimisation with respect the the data symbols
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function conditioned on minimisation with respect the the data symbols,
  \begin{align*}
 SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
-&= A - \rho Z(\theta) - \rho Z(\theta)^* + L \rho^2,
+&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
 \end{align*}
@@
 \[
-Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta).
+Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
 \]
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives 
+and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.  Differentiating with respect to $\rho$ and setting the resulting expression to zero gives 
 \begin{equation}\label{eq:hatrhoZ}
@@
 \end{equation}
-where $\Re(\cdot)$ denote the real part.
+where $\Re(\cdot)$ denotes the real part.
 Substituting this expression into $SS(\rho, \theta)$ above gives the sum of squares function conditioned upon minimisation with respect to the amplitude $\rho$ and the data symbols,
@@
 \end{align*}
-where $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  When $i \in D$,
+where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, then, when $i \in D$,
 \begin{align*}
- y_i e^{-j\theta} \hat{d}_i(\theta)^* &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
+ y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
 &= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
@@
 \]
-and put $\lambda_L = \langle\theta_0 - \hat{\theta}\rangle$.
-Then $\lambda_L$ is the maximiser of $G_L(\lambda)^2$ and equivalently the maximiser of $\abs{G_L(\lambda)}$.
-We will show that $\lambda_L$ converges almost surely to zero as $L \rightarrow \infty$.
+and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle$.
+Then $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)^2$ and equivalently the maximiser of $\abs{G_L(\lambda)}$.
+We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
 From this, the proof of strong consistency follows.
@@
 \]
-and since $\abs{P}}{L} \rightarrow p$ and $\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
+and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
 \[
@@
 \[
-\prob\{ \abs{G(\lambda_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L})
+\prob\{ \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L})
 \]
@@
 \[
-\abs{\lambda_L} > \delta \;\; \Rightarrow \;\; \abs{G(\lambda_L) - G(0)} > \epsilon
+\sabs{\hat{\lambda}_L} > \delta \;\; \Rightarrow \;\; \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon
 \]
@@
 \[
-\prob\{\abs{\lambda_L} > \delta\} \leq \prob\{ \abs{G(\lambda_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L}).
+\prob\{\sabs{\hat{\lambda}_L} > \delta\} \leq \prob\{ \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L}).
 \]
@@
 \[
-\sum_{L=1}^{\infty}\prob\{ \abs{\lambda_L} > \delta \} = \sum_{L=1}^{\infty}O(e^{-\epsilon^2 L})
+\sum_{L=1}^{\infty}\prob\{ \sabs{\hat{\lambda}_L} > \delta \} = \sum_{L=1}^{\infty}O(e^{-\epsilon^2 L})
 \]
-converges and $\lambda_L$ converges almost surely to zero by the Borel-Cantelli lemma.
+converges and consequently $\hat{\lambda}_L$ converges almost surely to zero by the Borel-Cantelli lemma.
 
-It remains to prove part two of this theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.
-From~\eqref{eq:hatrhoZ} the amplitude estimator satisfies $\hat{\rho} = \rho_0 G_L(\lambda_L)$.
-Lemma~\ref{lem:GLtoG0} shows that $G_L(\lambda_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and consequently $\hat{\rho}$ converges almost surely to $\rho_0 G(0)$ as required.
+It remains to prove part two of this theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ} the amplitude estimator satisfies 
+\begin{equation}\label{eq:rhoGLZ}
+\hat{\rho} = \frac{1}{L}\Re(Z(\theta)) = \rho_0 G_L(\hat{\lambda}_L).
+\end{equation}  
+Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
 \end{IEEEproof}
@@
 \[
-0 \leq G(\lambda_L) - G(0),
+0 \leq G(\hat{\lambda}_L) - G(0),
 \]
-and since $\lambda_L$ is the maximiser of $G_L$ then
+and since $\hat{\lambda}_L$ is the maximiser of $G_L$ then
 \[ 
-0 \leq G_L(0) - G_L(\lambda_L).
+0 \leq G_L(0) - G_L(\hat{\lambda}_L).
 \]
@@
 \begin{align*}
-0 &\leq G(\lambda_L) - G(0) \\ 
-& \leq G(\lambda_L) - G(0) + G_L(0) - G_L(\lambda_L) \\
-& \leq |G(\lambda_L) - G_L(\lambda_L)| + |G_L(0) - G(0)|,
+0 &\leq G(\hat{\lambda}_L) - G(0) \\ 
+& \leq G(\hat{\lambda}_L) - G(0) + G_L(0) - G_L(\hat{\lambda}_L) \\
+& \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)|,
 \end{align*}
@@
 \begin{lemma}\label{lem:GLtoG0}
-$G_L(\lambda_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
 \end{lemma}
@@
 \begin{IEEEproof} (Theorem~\ref{thm:normality})
-BLERG.  Need to show that it's actually sufficient to study the maximiser of $G_L$ rather than $\abs{G_L}$ (atleast asymptotically).
-
 The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.
 This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\lambda_L$, so
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
 \begin{align*}
-G_L^\prime(\lambda_L) &= 0 \\
+G_L^\prime(\hat{\lambda}_L) &= 0 \\
 &= \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
 \end{align*}
-Write
+Define the function $Q(\lambda) = \prob G_L^\prime(\lambda)$ and write
+\begin{align*}
+0 &= G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) + Q(\hat{\lambda}_L) \\
+&= \sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) + \sqrt{L}Q(\hat{\lambda}_L).
+\end{align*}
+Lemma~\ref{lem:Qconv} shows that
 \[
-0 = G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) + \prob G_L^\prime(\lambda_L)
-\]
-Lemma~\ref{lem:GprimeLlambdaLconv} shows that
-\[
-\sqrt{L} \prob G^\prime_L(\lambda_L) = \prob G_L^\prime(0) + \sqrt{L} \lambda_L\big( G(0)  + o_P(1) \big)
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} Q(0) + \sqrt{L} \hat{\lambda}_L\big( ?  + o_P(1) \big)
 \]
 where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.  Lemma~\ref{lem:empiricprocc} shows that
-\begin{align*}
-\sqrt{L}&\big( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \big) \\
-&= -\prob G_L^\prime(0) + \sqrt{L} G_L^\prime(0) + \sqrt{L}\lambda_L\big( ? +  o_P(1) \big).
-\end{align*}
+\[
+\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} \big( G_L^\prime(0) - Q(0) \big).
+\]
 It follows that
 \[
-\sqrt{L} \lambda_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{G(0)  + o_P(1)}.
+0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( ?  + o_P(1) \big)
 \]
+and rearranging gives,
+\[
+\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{?  + o_P(1)}.
+\]
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance
+\[
+p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}
+\]
+and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
+\[
+\frac{p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}}{?}
+\]
+as required.
+
+Define the function $H_L(\lambda) = \prob G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
+\begin{align*}
+\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
+&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) + H_L(\hat{\lambda}_L) - G(0) \big)
+\end{align*}
+Lemma~\ref{lem:empiricprocforrho} shows that 
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
+\]
+where $X_L = \sqrt{L}\big( G_L(0) - H_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
+\[
+\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \hat{\lambda}_L \big( ? + o_P(1) \big).
+\]
+It follows that
+\[
+\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
+\]
+Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and covariance
+\[
+?
+\]
+and Lemma~\ref{lem:XLcovlambdaL} shows that the covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ goes to zero as $L\rightarrow\infty$.
+So, the distribution of $\sqrt{L}m_L$ converges to the normal with zero mean and variance
+\[
+\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = BLERG.
+\]
+The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondinly
+\[
+?
+\]
+\end{IEEEproof}
+
+\begin{lemma}~\label{lem:diffatlambdaL}
+The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
+\[
+G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
+\]
+\end{lemma}
+
+\begin{lemma}\label{lem:Qconv}
+\[
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} Q(0) + \sqrt{L} \hat{\lambda}_L\big( ?  + o_P(1) \big)
+\]
+\end{lemma}
+\begin{IEEEproof}
+
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:empiricprocc} BLERG
+\[
+\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} \big( G_L^\prime(0) - Q(0) \big).
+\]
+\end{lemma}
+\begin{IEEEproof}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:convdistGLdash}
+The distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance
+\[
+p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}.
+\]
+\end{lemma}
+\begin{proof}
 Observe that 
 \[
-\sqrt{L} G_L^\prime((0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
+\sqrt{L} G_L^\prime(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
 \]
@@
 \]
-It follows that the distribution of $\sqrt{L}\lambda_L$ converges to the normal with mean zero and variance
+\end{proof}
+
+\begin{lemma}\label{lem:empiricprocforrho} BLERG
 \[
-\frac{p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}}{G(0)^2}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}~\label{lem:diffatlambdaL}
-The derivative of $G_L$ exists, and is equal to zero, at $\lambda_L$.  That is,
-\[
-G_L^\prime(\lambda_L) = \frac{d G_L}{d \lambda}(\lambda_L) = 0.
-\]
-\end{lemma}
-
-\begin{lemma}\label{lem:GprimeLlambdaLconv}
-\[
-\prob G^\prime_L(\lambda_L) = \lambda_L( G(0)  + o_P(1) )
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} \big( G_L^\prime(0) - H_L(0) \big).
 \]
@@
 \begin{IEEEproof}
-Observe that
-\[
-\prob G^\prime_L(\lambda) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda),
-\]
-where,
-\begin{align*}
-k_1(x) &= \prob R_1 \sin(x + \Phi_1) \;\;\; \text{and}, \\
-k_2(x) &=  \prob R_1 \sin\sfracpart{x + \Phi_1}.
-\end{align*}
-We have (BLERG) that $k_1$ and $k_2$ are continuous and differentiable and that $k_1(0) = k_2(0) = 0$, so by a first order Taylor expansion
-\[
-\prob G^\prime_L(\lambda_L) = \prob G^\prime_L(0) + \lambda_L( \prob G^\prime_L(\lambda_L)  + o_P(1) )
-\]
 \end{IEEEproof}
 
-\begin{lemma}\label{lem:empiricprocc} BLERG
+\begin{lemma}\label{lem:HLtoG} BLERG
 \[
-\sqrt{L} \left( G_L^\prime(\lambda_L) - \prob G_L^\prime(\lambda_L) \right) = \sqrt{L} G_L'(0) + \sqrt{L}\lambda_L( ? +  o_P(1) )
+\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \hat{\lambda}_L \big( ? + o_P(1) \big).
 \]
+\end{lemma}
+\begin{IEEEproof}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:XL} 
+The distribution of $X_L$ converges to the normal with zero mean and covariance
+\[
+?
+\]
+as $L \rightarrow\infty$.
+\end{lemma}
+\begin{IEEEproof}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:XLcovlambdaL} 
+The covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ converges to zero as $L\rightarrow\infty$.
 \end{lemma}
@@
 %         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{Complex gain error for BPSK}
+% 		\caption{complex amplitude error for BPSK}
 % 		\label{fig:plotcamp}
@@
 %         \includegraphics{code/data/plotM4-3.mps}
-% 		\caption{Complex gain error for QPSK}
+% 		\caption{complex amplitude error for QPSK}
 % 		\label{fig:plotcamp}
@@
 %         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{Complex gain error for 8PSK}
+% 		\caption{complex amplitude error for 8PSK}
 % 		\label{fig:plotcamp}
@@
  %         \includegraphics{code/data/plotM4-3.mps}
- % 		\caption{Complex gain error for QPSK}
+ % 		\caption{complex amplitude error for QPSK}
  % 		\label{fig:plotcamp}

### 2012-05-12 paper.tex
further edits to paper.  Just realised some simplifications too!

@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables and let $\{y_i, i \in P \cup D\}$ be observed noisy symbols given by the signal model~\eqref{eq:sigmod}.
-Let $\hat{a}$ be the least squares estimator of $a_0$ and put $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables with $\prob \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be observed noisy symbols given by the signal model~\eqref{eq:sigmod}.
+Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
 Put $L = \abs{P \cup D}$ and define real numbers $p$ and $d$ as the limits
@@
 where
-\begin{align*}
-h_1(x) &= \prob R_1 \cos(x + \Phi_1) \;\;\; \text{and}, \\
-h_2(x) &=  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
-\end{align*}
+\[
+h_1(x) = \prob R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
+\]
 If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
@@
 \end{align*}
-Define the function $Q(\lambda) = \prob G_L^\prime(\lambda)$ and write
+Define the function $Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0)$ and write
 \begin{align*}
@@
 \[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} Q(0) + \sqrt{L} \hat{\lambda}_L\big( ?  + o_P(1) \big)
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( ?  + o_P(1) \big)
 \]
@@
 \[
-\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} \big( G_L^\prime(0) - Q(0) \big).
+\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
 \]
@@
 
+\begin{lemma}
+\[
+\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{1}{2M} - \frac{1}{2LM ?}
+\]
+for all $i \in D$.
+\end{lemma}
+\begin{proof}
+Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
+The proof now proceeds by contradiction.
+Assume that $\sfracpart{\hat{\lambda}_L + \Phi_i} \leq \frac{1}{2M} - \frac{1}{2LM ?}$ for some $k$.
+Put $r_i = \hat{d}_i e_k$, then
+\begin{align*}
+Y(\{ r_i, i \in D\}) = \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) = \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
+\end{align*}
+where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
+\begin{align*}
+SS&(\{r_i, i \in D\}) \\
+&= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
+&= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
+&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
+&= SS(\{\hat{d}_i, i \in D\}) - R
+\end{align*}
+where 
+\[
+R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
+\]
+We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+
+To see that $R$ is positive write
+ \begin{align*}
+R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
+ &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
+ \end{align*}
+ so it is sufficient to show that
+ \[
+ \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
+ \]
+ Unfinished, but this proof will work
+\end{proof}
+
 \begin{lemma}\label{lem:Qconv}
 \[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} Q(0) + \sqrt{L} \hat{\lambda}_L\big( ?  + o_P(1) \big)
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( ph_1(0) + ?  + o_P(1) \big)
 \]
@@
 \begin{IEEEproof}
-
+Write
+\[
+Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda)
+\]
+where
+\begin{align*}
+k_1(\lambda) &= \prob R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \\
+k_2(\lambda) &= \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big).
+\end{align*}
+Since
+\[
+\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big)
+\]
+where $\zeta(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, then
+\begin{align*}
+k_1(\lambda) &= \prob R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
+&= \lambda \big( \prob R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \prob R_1 \big) \\
+&= \lambda \big( h_1(0) \; + \; \zeta(\lambda) K \big),
+\end{align*}
+where we put $K = \prob R_1$, a finite constant.
+Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
+$\zeta(\hat{\lambda}_L) = o_P(1)$.
+So,
+\begin{align*}
+k_1(\hat{\lambda}_L) &= \hat{\lambda}_L \big( h_1(0) \; + \; \zeta(\hat{\lambda}_L) K \big) \\
+&= \hat{\lambda}_L \big( h_1(0)  + o_P(1)\big).
+\end{align*}
 \end{IEEEproof}
@@
 \end{lemma}
-\begin{proof}
+\begin{IEEEproof}
 Observe that 
@@
 \]
-\end{proof}
+\end{IEEEproof}
 
@@
 \end{IEEEproof}
-
-% \section{A linear-time algorithm}\label{sec:line-time-algor}
-
-% \begin{lemma}
-% Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ so that $\hat{\theta}$ is the least squares estimator of the carrier phase and $\hat{\rho}$ is the least squares estimator of the amplitude.  Let 
-% \[
-% \lambda_i = \phi_i - \hat{\theta} - \frac{2\pi}{M}\round{\frac{M}{2\pi}(\phi_i - \hat{\theta})}.
-% \]
-% Then 
-% \[
-% \abs{\lambda_i} \leq \frac{1}{2M} - \frac{1}{2LM BLERG}
-% \]
-% for all $i \in P \cup D$.
-% \end{lemma}
-% \begin{proof}
-% Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
-The proof now proceeds by contradiction.
-Assume that $\lambda_i > \frac{1}{2M} - \frac{1}{2LM BLERG}$ for some $k$.
-Put $r_i = \hat{d}_i e_k$, then
-% \begin{align*}
-% Y(\{ r_i, i \in D\}) &= \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) \\
-% &= \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
-% \end{align*}
-% where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
-% \begin{align*}
-% SS(\{r_i, i \in D\}) &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
-% &= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
-% &= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
-% &= SS(\{\hat{d}_i, i \in D\}) - R
-% \end{align*}
-% where 
-% \[
-% R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
-% \]
-% We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-
-% To see that $R$ is positive write
-% \begin{align*}
-% R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
-% &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
-% \end{align*}
-% so it is sufficient to show that
-% \[
-% \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
-% \]
-% Unfinished.
-This proof will work, but the problem is that $\abs{y_k}$ can be really small.
-The would appear to neccesitate a very large number of buckets.
-In some sense, this is no problem, but you have to be able to allocate memory for all of these?
-Atleast you need a way of ordering them, up to a certain resolution, without sorting.
-I'm not convinced this will ever be possible.
-% \end{proof}
 

### 2012-05-12 paper.tex
Added to paper.

@@
 \[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( ph_1(0) + ?  + o_P(1) \big)
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + ?  + o_P(1) \big)
 \]
@@
 \end{align*}
-Since
+Since,
 \[
-\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big)
+\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big),
 \]
@@
 &= \lambda \big( \prob R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \prob R_1 \big) \\
-&= \lambda \big( h_1(0) \; + \; \zeta(\lambda) K \big),
+&= \lambda \big( 1 \; + \; \zeta(\lambda) K \big),
 \end{align*}
-where we put $K = \prob R_1$, a finite constant.
+since $h_1(0) = \prob R_1\cos(\Phi_1) = 1$ and where we put $K = \prob R_1$, a finite constant.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 So,
+\[
+k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) K \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big).
+\]
+Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
 \begin{align*}
-k_1(\hat{\lambda}_L) &= \hat{\lambda}_L \big( h_1(0) \; + \; \zeta(\hat{\lambda}_L) K \big) \\
-&= \hat{\lambda}_L \big( h_1(0)  + o_P(1)\big).
+k_2(\lambda) &= \int_0^{2\pi}\int_{0}^\infty r \big( \sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} \big) f(r,\phi) dr d\phi \\
+&= \int_0^{2\pi} \big( \sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} \big) g(\phi) d\phi.
 \end{align*}
+Now, if $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$,
+\begin{align*}
+\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} &= \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi k}{M}) \\
+&= \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big)
+\end{align*}
+when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{2k\pi}{M} - \lambda )$ and 
+\begin{align*}
+\sin&\fracpart{\lambda + \phi} - \sin\fracpart{\phi} \\
+&= \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi (k-1)}{M}) \\
+&=  v_k(\phi) + \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) 
+\end{align*}
+when $\phi \in [ \tfrac{2k\pi}{M} - \lambda,\tfrac{2k\pi}{M} )$, and where
+\[
+v_k(\phi) = \sin(\phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi (k-1)}{M}).
+\]
 \end{IEEEproof}
@@
 \begin{IEEEproof}
+
 \end{IEEEproof}

### 2012-05-13 paper.tex
Added to paper. Also added mathemtica file for computing some specific formula when the noise is Gaussian.

--- 2012-05-12 paper.tex
@@
 Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0 G(0)$.
-If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at multiples of $\tfrac{\pi}{M}$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for every integer $k$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
 \[
-C = ?
+C = \frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
+A_1 + A_2 & ? \\
+? & ?
+\end{array} \right)
 \]
-as $L \rightarrow \infty$.
+as $L \rightarrow \infty$, where
+\[
+H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
+\]
+and
+\[
+A_1 = \prob R_1^2\sin^2(\Phi_1), \qquad A_2 = \prob R_1^2\sin^2\fracpart{\Phi_1}.
+\]
 \end{theorem}
@@
 \[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( ?  + o_P(1) \big)
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
 \]
@@
 \[
-0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( ?  + o_P(1) \big)
+0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
 \]
@@
 \[
-\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{?  + o_P(1)}.
+\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
 \]
@@
 \end{lemma}
-\begin{proof}
+\begin{IEEEproof}
 Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
@@
  Unfinished, but this proof will work
-\end{proof}
+\end{IEEEproof}
 
@@
 \[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( ph_1(0) + ?  + o_P(1) \big)
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\]
+where 
+\[
+H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
 \]
@@
 where
-\begin{align*}
-k_1(\lambda) &= \prob R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \\
-k_2(\lambda) &= \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big).
-\end{align*}
-Since
+\begin{align}
+k_1(\lambda) &= \prob R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
+k_2(\lambda) &= \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
+\end{align}
+Since,
 \[
-\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big)
+\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big),
 \]
@@
 So,
+\[
+k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) K \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big).
+\]
+Lemma~\ref{lem:k2} shows that
+\[
+k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( ? \; + \; o_P(1) \big).
+\]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:k2}
+With the function $k_2(\lambda)$ defined as in~\eqref{eq:k2def}, then
+\[
+k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H \; + \; o_P(1) \big).
+\]
+where
+\[
+H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
+\]
+\end{lemma}
+\begin{IEEEproof}
+Now, if $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$,
 \begin{align*}
-k_1(\hat{\lambda}_L) &= \hat{\lambda}_L \big( h_1(0) \; + \; \zeta(\hat{\lambda}_L) K \big) \\
-&= \hat{\lambda}_L \big( h_1(0)  + o_P(1)\big).
+\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} &= \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi k}{M}) \\
+&= \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
+&= \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big)
 \end{align*}
+when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$ and 
+\begin{align*}
+\sin&\fracpart{\lambda + \phi} - \sin\fracpart{\phi} \\
+&= \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi (k-1)}{M}) \\
+&=  v_k(\phi) + \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
+&= v_k(\phi) + \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big)
+\end{align*}
+when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} )$, and where
+\[
+v_k(\phi) = \sin(\phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi (k-1)}{M}).
+\]
+Observe that $v_k(\tfrac{(2k+1)\pi}{M}) = 2\sin(\tfrac{\pi}{M})$.  We can now write
+\[
+\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} = v(\lambda, \phi) + \lambda\cos\fracpart{\phi}
+\]
+where $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$ and where
+\[
+v_k(\lambda,\phi) = \begin{cases}
+v_k(\phi) + \lambda\zeta_k(\lambda) &  \phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} ) \\
+\lambda\zeta_k(\lambda) & \phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda ) \\
+0 & \text{otherwise}.
+\end{cases}
+\]
+Now
+\begin{align*}
+k_2(\lambda) &=  \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
+&= \prob R_1\big( v(\lambda, \Phi_1) + \lambda\cos\fracpart{\Phi_1} \big) \\
+&= \prob R_1 v(\lambda, \Phi_1) + \lambda h_2(0).
+\end{align*}
+It remains to evaluate $\prob R_1 v(\lambda, \Phi_1)$.  Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
+\begin{align*}
+k_2(\lambda) &= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \\
+&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi.
+\end{align*}
+Since $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$,
+\[
+k_2(\lambda) = \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi
+\]
+and since $v_k(\lambda,\phi) = 0$ for $\phi \in [0,2\pi)$ when $k \notin \{0,1,\dots,M\}$ the infinite sum above can be truncated so that
+\[
+k_2(\lambda) = \sum_{k = 0}^M \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi
+\]
+Now
+\begin{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = &\int_{\tfrac{(2k+1)\pi}{M} - \lambda}^{\tfrac{(2k+1)\pi}{M}}v_k(\phi)g(\phi)d\phi \\
+&\hspace{0.5cm}+ \lambda\zeta_k(\lambda) \int_{\tfrac{(2k-1)\pi}{M}}^{\tfrac{(2k+1)\pi}{M}}g(\phi)d\phi.
+\end{align*}
+Now, since $g\fracpart{\phi}_\pi$ is continuous at $(2k+1)\tfrac{\pi}{M}$ for each integer $k$, and $v_k(\phi)$ is continuous and $v_k(\tfrac{(2k+1)\pi}{M}) =  2\sin(\tfrac{\pi}{M})$,
+\begin{align*}
+\int_{\tfrac{(2k+1)\pi}{M} - \lambda}^{\tfrac{(2k+1)\pi}{M}}v_k(\phi)g(\phi)d\phi &= \lambda( v_k(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) ) \\
+&= \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) )
+\end{align*}
+where $\nu_k(\lambda)$ is function that goes to zero as $\lambda$ goes to zero.
+To keep the notation clean put $A_k = \int_{\tfrac{(2k-1)\pi}{M}}^{\tfrac{(2k+1)\pi}{M}}g(\phi)d\phi$.
+For each $k\in\ints$, $A_k$ is a nonnegative finite constant less than one.
+Now
+\begin{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi \\
+&= \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k )
+\end{align*}
+and
+\begin{align*}
+k_2(\lambda) &= \sum_{k = 0}^M \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k ) \\
+ &= \epsilon(\lambda) + 2 \lambda \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
+\end{align*}
+where $\epsilon(\lambda) = \sum_{k=0}^M\big( \nu_k(\lambda) + \zeta_k(\lambda) A_k \big)$ is a function that goes to zero as $\lambda$ goes to zero.
+Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\epsilon(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
+$\epsilon(\hat{\lambda}_L) = o_P(1)$.
+So
+\[
+k_2(\hat{\lambda}_L) = o_P(1) + 2 \hat{\lambda}_L \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
+\]
+as required.
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:convdistGLdash}
-The distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance
-\[
-p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}.
-\]
+The distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $A_1 + A_2$.
 \end{lemma}
@@
 \begin{IEEEproof}
+
 \end{IEEEproof}
+
+\section{The Gaussian noise case}
+
+Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.
+The joint density function of the real and imaginary parts is then
+\[
+p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
+\]
+Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+%and the joint density function of the real and imaginary parts of $Z$ is
+%\[
+%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
+%\] 
+%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
+It can be shown that
+\[
+g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+\]
+where
+\[
+\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
+\]
+The value of $A_1$ and $A_2$ can be computed by numerical integration.
 

### 2012-05-13 paper.tex
Added to paper.

@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables with $\prob \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be observed noisy symbols given by the signal model~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables with $\prob \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be noisy symbols given by the signal model~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
@@
 Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0 G(0)$.
-If the periodic function $g\fracpart{\phi}_{\pi}$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for every integer $k$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for every integer $k$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
 \[
@@
 \[
-H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
+H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M})
 \]
@@
 
-We are interested in analysing $\hat{\theta}$, the maximiser of $\Re(Z(\theta))^2$ or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
+We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$ or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
 Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
@@
 and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle$.
-Then $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)^2$ and equivalently the maximiser of $\abs{G_L(\lambda)}$.
+Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
 We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-From this, the proof of strong consistency follows.
+The proof of part 1 of the theorem follows from this.
 
@@
 \]
-and since $v_k(\lambda,\phi) = 0$ for $\phi \in [0,2\pi)$ when $k \notin \{0,1,\dots,M\}$ the infinite sum above can be truncated so that
+and since $v_k(\lambda,\phi) = 0$ for $\phi \in [0,2\pi)$ when $k \notin \{0,1,\dots,M-1\}$ the infinite sum above can be truncated so that
 \[
-k_2(\lambda) = \sum_{k = 0}^M \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi
+k_2(\lambda) = \sum_{k = 0}^{M-1} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi
 \]
+(BLERG this needs modification, more care)
 Now
@@
 \end{align*}
-Now, since $g\fracpart{\phi}_\pi$ is continuous at $(2k+1)\tfrac{\pi}{M}$ for each integer $k$, and $v_k(\phi)$ is continuous and $v_k(\tfrac{(2k+1)\pi}{M}) =  2\sin(\tfrac{\pi}{M})$,
+Now, since $g(\phi)$ is continuous at $(2k+1)\tfrac{\pi}{M}$ for each integer $k$, and $v_k(\phi)$ is continuous and $v_k(\tfrac{(2k+1)\pi}{M}) =  2\sin(\tfrac{\pi}{M})$,
 \begin{align*}
@@
 \begin{align*}
-k_2(\lambda) &= \sum_{k = 0}^M \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k ) \\
+k_2(\lambda) &= \sum_{k = 0}^{M-1} \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k ) \\
  &= \epsilon(\lambda) + 2 \lambda \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
@@
 \[
-k_2(\hat{\lambda}_L) = o_P(1) + 2 \hat{\lambda}_L \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
+k_2(\hat{\lambda}_L) = o_P(1) + 2 \hat{\lambda}_L \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M})
 \]

### 2012-05-13 paper.tex
Added to paper.

@@
 \]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $A_1 + A_2$ and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
 \[
-p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}
+\frac{A_1 + A_2}{(p + Hd)^2}.
 \]
-and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
-\[
-\frac{p \prob R_1^2 \sin^2(\Phi_1) + d \prob R_1^2 \sin^2\fracpart{\Phi_1}}{?}
-\]
-as required.
 
-Define the function $H_L(\lambda) = \prob G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
+We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Define the function $H_L(\lambda) = \prob G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
 \begin{align*}
@@
 \]
-Lemma~\ref{lem:k2} shows that
-\[
-k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( ? \; + \; o_P(1) \big).
-\]
-\end{IEEEproof}
 
-\begin{lemma}\label{lem:k2}
-With the function $k_2(\lambda)$ defined as in~\eqref{eq:k2def}, then
-\[
-k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H \; + \; o_P(1) \big).
-\]
-where
-\[
-H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
-\]
-\end{lemma}
-\begin{IEEEproof}
-Now, if $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$,
+We now analyse the behaviour of $k_2(\lambda)$.
+Firsty consider when, $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$ and when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$,
 \begin{align*}
@@
 &= \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
-&= \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big)
+&= \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big).
 \end{align*}
-when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$ and 
+where $\zeta_l(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} )$,
 \begin{align*}
@@
 \end{align*}
-when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} )$, and where
+where
 \[
@@
 \]
-Observe that $v_k(\tfrac{(2k+1)\pi}{M}) = 2\sin(\tfrac{\pi}{M})$.  We can now write
+Observe that $v_k(\tfrac{(2k+1)\pi}{M}) = 2\sin(\tfrac{\pi}{M})$.  Now, for all $\phi \in \reals$ we have
 \[
@@
 \]
-and since $v_k(\lambda,\phi) = 0$ for $\phi \in [0,2\pi)$ when $k \notin \{0,1,\dots,M-1\}$ the infinite sum above can be truncated so that
+and since $v_k(\lambda,\phi) = 0$ for $\phi$ in the interval $[0,2\pi)$ when $k \notin \{0,1,\dots,M-1\}$ the infinite sum above can be truncated so that
 \[

### 2012-05-14 paper.tex
Added to paper a little

@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\prob\abs{a_0 - \hat{a}}^2$ where $\prob$ is the expected value operator.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\prob\vert\langle\theta_0 - \hat{\theta}\rangle_\pi\vert^2$ and $\prob\abs{\rho_0 - \hat{\rho}}^2$.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ is the expected value operator.
+It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\prob\vert\langle\theta_0 - \hat{\theta}\rangle_\pi\vert^2$ rather than $\prob\vert\theta_0 - \hat{\theta}\vert^2$ is the appropriate measure of error for the phase parameter.
+It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
 
-%\subsection{Notation}
-It is worth commenting on our use of $\prob$ rather than the more usual $\expect$ or $E$ for the expected value operator.
+%It is worth commenting on our use of $\prob$ rather than the more usual $\expect$ or $E$ for the expected value operator.
 The notation comes from Pollard~\cite[Ch 1]{Pollard_users_guide_prob_2002}.
@@
 Given a random variable $X$ with cumulative density function $F$, the probability of an event, say $X \in S$, where $S$ is some subset of the range of $X$, is
-\[
-\prob \indicator \{X \in S\} = \int \indicator \{X \in S\}(x) dF(x) = \int_{S} dF(x)
-\]
-where $\indicator \{X \in S\}$ is the indicator function of the set $S$, i.e $\indicator \{X \in S\}(x) = 1$ when the argument $x \in S$ and zero otherwise.
+%\[
+%\prob \indicator \{X \in S\} = \int \indicator \{X \in S\}(x) dF(x) = \int_{S} dF(x)
+%\]
+%where $\indicator \{X \in S\}$ is the indicator function of the set $S$, i.e $\indicator \{X \in S\}(x) = 1$ when the argument $x \in S$ and zero otherwise.
 We will usually drop the $\onebf$ and simply write $\prob \{ X \in S \}$ to mean $\prob \onebf\{ X \in S \}$.
 To illustrate the utility of this notation, Markov's inequality becomes
-\[
-\prob \{ \abs{X} > \delta \}  \leq \prob \frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \} \leq \frac{1}{\delta}\prob\abs{X},
-\]
-where $\frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \}(x)$ is the function equal to $\abs{x}/\delta$ when the argument $x > \delta$ and zero otherwise.
+%\[
+%\prob \{ \abs{X} > \delta \}  \leq \prob \frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \} \leq \frac{1}{\delta}\prob\abs{X},
+%\]
+%where $\frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \}(x)$ is the function equal to $\abs{x}/\delta$ when the argument $x > \delta$ and zero otherwise.
 
@@
 
-\begin{algorithm} \label{alg:loglinear}
+\begin{algorithm}[t] \label{alg:loglinear}
 \SetAlCapFnt{\small}
@@
 We now analyse the behaviour of $k_2(\lambda)$.
-Firsty consider when, $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$ and when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$,
+Firstly consider when, $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$ and when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$,
 \begin{align*}
@@
 \begin{align*}
-\int_{\tfrac{(2k+1)\pi}{M} - \lambda}^{\tfrac{(2k+1)\pi}{M}}v_k(\phi)g(\phi)d\phi &= \lambda( v_k(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) ) \\
+\int_{\tfrac{(2k+1)\pi}{M} - \lambda}^{\tfrac{(2k+1)\pi}{M}}&v_k(\phi)g(\phi)d\phi \\
+&= \lambda( v_k(\tfrac{(2k+1)\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) ) \\
 &= \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) )

### 2012-05-14 paper.tex
Added to paper.

@@
 
-\section{Statistical properties}
+\section{Circularly symmetric complex random variables}
+
+Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
+A complex random variable $X$ is called \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
+That is, if we define real random variables $Z > 0$ and $\Theta \in [0,2\pi)$ such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
+If the probability density function of $Z$ is $f_Z(z)$, then the joint density function of $\Theta$ and $Z$ is
+\[
+f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
+\]  
+Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariate to phase rotation.
+If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean since
+\begin{align*}
+\expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
+&= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
+&= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
+\end{align*}
+
+We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
+\[
+R e^{j\Phi} = 1 + X.
+\]
+The joint distribution of $R$ and $\Phi$ can be show to be
+\[
+f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}
+\]
+The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.
+
+\section{Statistical properties of the least square estimator}
 

### 2012-05-15 paper.tex
Added a bit more to paper.

@@
 
-We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
+We will have particular use of complex random variables of the form $1 + X$.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
 \[
@@
 \[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}
-\]
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.
-
-\section{Statistical properties of the least square estimator}
+f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}.
+\]
+The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
+\[
+\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
+\]
+where $\Re(\cdot)$ denotes the real part, and
+\[
+\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
+\]
+where $\Im(\cdot)$ denotes the imaginary part.
+
+\section{Statistical properties of the least squares estimator}
 
@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symetric complex random variables with $\prob \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be noisy symbols given by the signal model~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be noisy symbols given by the signal model~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
@@
 \[
-G(x) = p h_1(x) + d h_2(x)
-\]
-where
-\[
-h_1(x) = \prob R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \prob R_1 \cos\sfracpart{ x + \Phi_1}.
+G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
+\]
+\[
+h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
 \]
@@
 Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0 G(0)$.
-If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for every integer $k$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
 \[
@@
 \[
-H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M})
-\]
-and
-\[
-A_1 = \prob R_1^2\sin^2(\Phi_1), \qquad A_2 = \prob R_1^2\sin^2\fracpart{\Phi_1}.
+H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M}),
+\]
+\[
+A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1}.
 \]
@@
 
-\begin{IEEEproof} (Theorem~\ref{thm:consistency})
+Proofs of these theorems are given in the follows subsections~\ref{sec:proof-almost-sure}~and~\ref{sec:proof-asympt-norm}.  We first discussion the assumptions made by these theorems.
+
+\begin{lemma}
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+Let $h_2$ and $k_2$ be functions
+\[
+h_2(x) = \expect R \cos\sfracpart{x + \Phi}, \;\;\; k_2(x) =  \expect R \sin\sfracpart{ x + \Phi}.
+\]
+Then $h_2(x)$ is uniquely maximised at $x=0$ other the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$ and $k_2(0) = 0$.
+\end{lemma}
+
+\subsection{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
+
 Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function conditioned on minimisation with respect the the data symbols,
@@
 \]
-and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.  Differentiating with respect to $\rho$ and setting the resulting expression to zero gives 
+and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of the amplitude as a function of $\theta$,
 \begin{equation}\label{eq:hatrhoZ}
@@
 &= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
-&\equiv \round{ \fracpart{\theta_0 - \theta} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
+&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
 &= \round{ \lambda + \Phi_i + \angle{s_i} },
@@
 \[
-\prob G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
+\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
 \]
@@
 \[
-G(\theta) = \lim_{L \rightarrow\infty} \prob G_L(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
+G(\lambda) = \lim_{L \rightarrow\infty} \expect G_L(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
 \]
@@
 \]
-Since $G(\lambda)$ is continuous and is uniquely maximised at $0$, then for any $\delta > 0$ there is a $\epsilon > 0$ such that 
+where $\prob$ denotes probability.  Since $G(\lambda)$ is continuous and is uniquely maximised at $0$, then for any $\delta > 0$ there is a $\epsilon > 0$ such that 
 \[
@@
 
-It remains to prove part two of this theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ} the amplitude estimator satisfies 
+It remains to prove part two of this theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
 \begin{equation}\label{eq:rhoGLZ}
-\hat{\rho} = \frac{1}{L}\Re(Z(\theta)) = \rho_0 G_L(\hat{\lambda}_L).
+\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
 \end{equation}  
 Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
-\end{IEEEproof}
 
@@
 
-\begin{IEEEproof} (Theorem~\ref{thm:normality})
+\subsection{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
+
 The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.
@@
 G_L^\prime(\hat{\lambda}_L) &= 0 \\
-&= \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\end{align*}
-Define the function $Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0)$ and write
+&= \frac{1}{L} \sum_{i \in P} R_i \sin(\hat{\lambda}_L + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\hat{\lambda}_L + \Phi_i}.
+\end{align*}
+Define the function $Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0)$ and write
 \begin{align*}
@@
 
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Define the function $H_L(\lambda) = \prob G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
+We now analyse the asymptotic distribution of $m_L = \hat{\rho} - \rho_0 G(0)$.  Define the function $H_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
 \begin{align*}
@@
 \[
-\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
 \]
@@
 \]
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and covariance
+Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance
 \[
@@
 \]
-\end{IEEEproof}
 
@@
 \[
-Q(\lambda) = \prob G_L^\prime(\lambda) - \prob G_L^\prime(0) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda)
+Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda)
 \]
@@
 \begin{align}
-k_1(\lambda) &= \prob R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
-k_2(\lambda) &= \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
+k_1(\lambda) &= \expect R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
+k_2(\lambda) &= \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
 \end{align}
@@
 \begin{align*}
-k_1(\lambda) &= \prob R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
-&= \lambda \big( \prob R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \prob R_1 \big) \\
+k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
+&= \lambda \big( \expect R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \expect R_1 \big) \\
 &= \lambda \big( 1 \; + \; \zeta(\lambda) K \big),
 \end{align*}
-since $h_1(0) = \prob R_1\cos(\Phi_1) = 1$ and where we put $K = \prob R_1$, a finite constant.
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$, and where we put $K = \expect R_1$ is a finite constant.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 \begin{align*}
-k_2(\lambda) &=  \prob R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
-&= \prob R_1\big( v(\lambda, \Phi_1) + \lambda\cos\fracpart{\Phi_1} \big) \\
-&= \prob R_1 v(\lambda, \Phi_1) + \lambda h_2(0).
-\end{align*}
-It remains to evaluate $\prob R_1 v(\lambda, \Phi_1)$.  Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
+k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
+&= \expect R_1\big( v(\lambda, \Phi_1) + \lambda\cos\fracpart{\Phi_1} \big) \\
+&= \expect R_1 v(\lambda, \Phi_1) + \lambda h_2(0).
+\end{align*}
+It remains to evaluate $\expect R_1 v(\lambda, \Phi_1)$.  Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
 \begin{align*}
@@
 \[
-p \prob R_1^2 \sin^2(\Phi_1).
+p \expect R_1^2 \sin^2(\Phi_1).
 \]
@@
 \[
-d \prob R_1^2 \sin^2\fracpart{\Phi_1}.
+d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
 \]

### 2012-05-16 code/scala/test/cam/psk/CLTTest.scala
added clt computing code.  A bit broken at the moment!

@@ -1,53 +1,0 @@
-/*
- * Test for the CLT computers
- */
-
-package cam.psk
-
-import org.junit._
-import Assert._
-import pubsim.distributions.complex.ComplexRandomVariable
-import pubsim.distributions.complex.SymmetricComplexNormal
-
-
-class CLTTest {
-
-  @Before
-  def setUp: Unit = {
-  }
-
-  @After
-  def tearDown: Unit = {
-  }
-
-  @Test
-  def example = {
-  }
-  
-  @Test
-  def testCoherentCLT = {
-    val M = 4
-    val p = 0.1
-    val rho0 = 1.0
-    val sigma = 1.0
-    val snr = rho0*rho0/2/sigma/sigma
-    val kappa = rho0/sigma
-    val X = new SymmetricComplexNormal(sigma*sigma)
-    val clt = new CoherentCLT(M,X,p)
-    println("H " + clt.H)
-    println("A1 " + clt.A1)
-    println("A2 " + clt.A2)
-    println("var " + clt.variance)
-    println("Z var " + clt.Z.getVariance)
-    println("g")
-    for( x <- 0.0 to 2.0*scala.math.Pi by 0.2 ) println(clt.g(x))
-    println()
-    println("fZ")
-    for( z <- 0.0 to 1000*clt.Z.getVariance by 10*clt.Z.getVariance ) println(clt.fZ(z))
-    //println()
-    //println("f")
-    //for( z <- 0.0 to 10.0 by 0.2 ) println(clt.f(z,0.0))
-    
-  }
-
-}

### 2012-05-17 paper.tex
added clt computing code.  A bit broken at the moment!
Added a stack of new simulations.  Fixed bug in clt code.  This now appears to work, but I would like to look at it a little closer.
Added to paper

@@
 s_i = \begin{cases}
-p_i & i \in P, \\
-d_i & i \in D, \\
-0 & otherwise,
+p_i & i \in P \\
+d_i & i \in D,
 \end{cases}
@@
 The sets $P$ and $D$ are disjoint, i.e.
-$P \cap D = \emptyset$.
+$P \cap D = \emptyset$, and $L = \abs{P \cup D}$ is the total number of symbols transmitted.
 
@@
 Complicating matters is that the data symbols $d_i$ are not known to the reciever and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
+Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
 A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
@@
 In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
-An important paper is by Mackenthun~\cite{Mackenthun1994} who described an algorithm to compute the least squares estimator requiring only $O(L \log L)$ arithmetic operations, where $L = \abs{P \cup D}$ is the total number of symbols transmitted.
+A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+The idea is the `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
+A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and $A$ is the average
+\[
+A = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
+\]
+Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
+A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included, so as to made the estimator coherent.
+Techniques are available but they are somewhat adhoc~\cite{Cowley_ref_sym_carr_1998}.
+This problem does not occur with the least square estimator.
+
+An important paper is by Mackenthun~\cite{Mackenthun1994} who described an algorithm to compute the least squares estimator requiring only $O(L \log L)$ arithmetic operations.
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Here, we consider a slightly more general model than is usual, one that allows the utilisation of pilot symbols, i.e., \emph{coherent detection}.
+Both Mackenthun and Swelden's considered the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-We show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
 
@@
 SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
-&= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*. \label{eq:SS}
+&= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*, \label{eq:SS}
 \end{align}
+where $*$ denote the complex congugate.
 Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find that the least squares estimator of $a_0$, satisfies
@@
 and where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
-A word of caution, it is often the case that $\round{\cdot}$ denotes rounding to the nearest \emph{integer}.
+A word of caution, the notation $\round{\cdot}$ is often used to denote rounding to the nearest \emph{integer}.
 This is not the case here.
@@
 \] 
-Note that $\hat{d}_i(\theta)$ does not depend on the amplitude $\rho$.
+Note that $\hat{d}_i(\theta)$ does not depend on $\rho$.
 As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
@@
 
-Let $S$ be the set of all $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$, that is,
+Let $S$ be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$, that is,
 \[
@@
 \]
-If $\hat{\theta}$ is the least squares estimator the phase, then it is clear from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
+If $\hat{\theta}$ is the least squares estimator the phase, then it follows from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
 Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
@@
 where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in the sequence $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
-Letting $g_{k}$ be the sequence $\{ y_i f_{ki}^*, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
+Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
 \[
@@
 Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given the sequence $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ of the indices in $D$ as described in~\eqref{eq:sigmasortind}.
-Computing the $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
+The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
 This requires $O(L \log L)$ operations.
@@
 
-\section{Circularly symmetric complex random variables}
+\section{Circularly symmetric complex random variables}\label{sec:circ-symm-compl}
 
 Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
-A complex random variable $X$ is called \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
+A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
 That is, if we define real random variables $Z > 0$ and $\Theta \in [0,2\pi)$ such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
@@
 \]  
-Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariate to phase rotation.
+Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariant to phase rotation.
 If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean since
@@
 \]
-The joint distribution of $R$ and $\Phi$ can be show to be
+The joint distribution of $R$ and $\Phi$ can be shown to be
 \[
@@
 \]
-where $\Im(\cdot)$ denotes the imaginary part.
+where $\Im(\cdot)$ denotes the imaginary part.  The next lemma will be useful for the analysis of the least squares estimator.
+
+\begin{lemma}\label{lem:h1minedcircsym}
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+Let $h_1(x)$ be the function
+\[
+h_1(x) = \expect R \cos(x + \Phi).
+\]
+Then $h_1(x)$ is uniquley maximised as $0$ over the interval $[-\pi,\pi)$.
+\end{lemma}
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
 
@@
 \]
-Put $\lambda_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $m_L = \hat{\rho} - \rho_0 G(0)$.
+Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
 If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
@@
 C = \frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
-A_1 + A_2 & ? \\
+pA_1 + dA_2 & ? \\
 ? & ?
@@
 \[
-H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M}),
+H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M}),
 \]
@@
 
-Proofs of these theorems are given in the follows subsections~\ref{sec:proof-almost-sure}~and~\ref{sec:proof-asympt-norm}.  We first discussion the assumptions made by these theorems.
+Proofs of these theorems are given in subsections~\ref{sec:proof-almost-sure}~and~\ref{sec:proof-asympt-norm}.  We first discuss the assumptions made by these theorems.
 
 \begin{lemma}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
-Let $h_2$ and $k_2$ be functions
+Under the same conditions as Lemma~\ref{lem:h1minedcircsym}, define the functions $h_2$ and $k_2$, 
 \[
@@
 We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-The proof of part 1 of the theorem follows from this.
+The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.
 
@@
 \[
-G(\lambda) = \lim_{L \rightarrow\infty} \expect G_L(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
-\]
-Lemma~\ref{lem:convtoexpGlamL} shows that,
+\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
+\]
+where the function $G$, $h_1$ and $h_2$ are defined in the statement of Theorem~\ref{thm:consistency}.  Lemma~\ref{lem:convtoexpGlamL} shows that,
 \[
@@
 
-\begin{lemma}\label{lem:convtoexpGlamL}
+\begin{lemma}\label{lem:convtoexpGlamL} For any $\epsilon > 0$,
 \[
@@
 \begin{IEEEproof}
-Lemma~\ref{lem:Zconunif} shows BLERG that 
-\[
-0 \leq G(\hat{\lambda}_L) - G(0),
-\]
-and since $\hat{\lambda}_L$ is the maximiser of $G_L$ then
-\[ 
-0 \leq G_L(0) - G_L(\hat{\lambda}_L).
-\]
-Adding both inequalities and taking appropriate absolute values gives
-\begin{align*}
-0 &\leq G(\hat{\lambda}_L) - G(0) \\ 
-& \leq G(\hat{\lambda}_L) - G(0) + G_L(0) - G_L(\hat{\lambda}_L) \\
-& \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)|,
-\end{align*}
+BLERG
+% Lemma~\ref{lem:Zconunif} shows BLERG that 
+% \[
+% 0 \leq G(\hat{\lambda}_L) - G(0),
+% \]
+% and since $\hat{\lambda}_L$ is the maximiser of $G_L$ then
+% \[ 
+% 0 \leq G_L(0) - G_L(\hat{\lambda}_L).
+% \]
+% Adding both inequalities and taking appropriate absolute values gives
+% \begin{align*}
+% 0 &\leq G(\hat{\lambda}_L) - G(0) \\ 
+% & \leq G(\hat{\lambda}_L) - G(0) + G_L(0) - G_L(\hat{\lambda}_L) \\
+% & \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)|,
+% \end{align*}
+\end{IEEEproof}
+
+% \begin{lemma}\label{lem:Zconunif}
+% \[
+% \prob\left\{\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} > \epsilon \right\} = O(e^{-\epsilon^2 L}).
+% \]
+% \end{lemma}
+% \begin{IEEEproof}
+% \end{IEEEproof}
+
+\begin{lemma}\label{lem:GLtoG0}
+$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
 BLERG
@@
 
-\begin{lemma}\label{lem:Zconunif}
-\[
-\prob\left\{\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} > \epsilon \right\} = O(e^{-\epsilon^2 L}).
+\subsection{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
+
+The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
+This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
+\begin{align*}
+G_L^\prime(\hat{\lambda}_L) &= 0 \\
+&= \frac{1}{L} \sum_{i \in P} R_i \sin(\hat{\lambda}_L + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\hat{\lambda}_L + \Phi_i}.
+\end{align*}
+Define the function $Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0)$ and write
+\begin{align*}
+0 &= G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) + Q(\hat{\lambda}_L) \\
+&= \sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) + \sqrt{L}Q(\hat{\lambda}_L).
+\end{align*}
+Lemma~\ref{lem:Qconv} shows that
+\[
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\]
+where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.  Lemma~\ref{lem:empiricprocc} shows that
+\[
+\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
+\]
+It follows that
+\[
+0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
+\]
+and rearranging gives,
+\[
+\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
+\]
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
+\[
+\frac{pA_1 + dA_2}{(p + Hd)^2}.
+\]
+
+We now analyse the asymptotic distribution of $m_L = \hat{\rho} - \rho_0 G(0)$.  Define the function $H_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
+\begin{align*}
+\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
+&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) + H_L(\hat{\lambda}_L) - G(0) \big)
+\end{align*}
+Lemma~\ref{lem:empiricprocforrho} shows that 
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
+\]
+where $X_L = \sqrt{L}\big( G_L(0) - H_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
+\[
+\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
+\]
+It follows that
+\[
+\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
+\]
+Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance
+\[
+?
+\]
+and Lemma~\ref{lem:XLcovlambdaL} shows that the covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ goes to zero as $L\rightarrow\infty$.
+So, the distribution of $\sqrt{L}m_L$ converges to the normal with zero mean and variance
+\[
+\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = BLERG.
+\]
+The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondinly
+\[
+?
+\]
+
+\begin{lemma}~\label{lem:diffatlambdaL}
+The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
+\[
+G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
 \]
@@
 
-\begin{lemma}\label{lem:GLtoG0}
-$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-\end{IEEEproof}
-
-\subsection{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
-
-The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at half integers.
-This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
-\begin{align*}
-G_L^\prime(\hat{\lambda}_L) &= 0 \\
-&= \frac{1}{L} \sum_{i \in P} R_i \sin(\hat{\lambda}_L + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\hat{\lambda}_L + \Phi_i}.
-\end{align*}
-Define the function $Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0)$ and write
-\begin{align*}
-0 &= G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) + Q(\hat{\lambda}_L) \\
-&= \sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) + \sqrt{L}Q(\hat{\lambda}_L).
-\end{align*}
-Lemma~\ref{lem:Qconv} shows that
-\[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.  Lemma~\ref{lem:empiricprocc} shows that
-\[
-\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
-\]
-It follows that
-\[
-0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
-\]
-and rearranging gives,
-\[
-\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
-\]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $A_1 + A_2$ and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
-\[
-\frac{A_1 + A_2}{(p + Hd)^2}.
-\]
-
-We now analyse the asymptotic distribution of $m_L = \hat{\rho} - \rho_0 G(0)$.  Define the function $H_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
-\begin{align*}
-\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) + H_L(\hat{\lambda}_L) - G(0) \big)
-\end{align*}
-Lemma~\ref{lem:empiricprocforrho} shows that 
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - H_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
-\]
-It follows that
-\[
-\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
-\]
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance
-\[
-?
-\]
-and Lemma~\ref{lem:XLcovlambdaL} shows that the covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ goes to zero as $L\rightarrow\infty$.
-So, the distribution of $\sqrt{L}m_L$ converges to the normal with zero mean and variance
-\[
-\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = BLERG.
-\]
-The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondinly
-\[
-?
-\]
-
-\begin{lemma}~\label{lem:diffatlambdaL}
-The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
-\[
-G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
-\]
-\end{lemma}
-
-\begin{lemma}
-\[
-\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{1}{2M} - \frac{1}{2LM ?}
-\]
-for all $i \in D$.
-\end{lemma}
-\begin{IEEEproof}
-Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
+% \begin{lemma}
+% \[
+% \sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{1}{2M} - \frac{1}{2LM ?}
+% \]
+% for all $i \in D$.
+% \end{lemma}
+% \begin{IEEEproof}
+% Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
 The proof now proceeds by contradiction.
@@
 Put $r_i = \hat{d}_i e_k$, then
-\begin{align*}
-Y(\{ r_i, i \in D\}) = \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) = \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
-\end{align*}
-where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
-\begin{align*}
-SS&(\{r_i, i \in D\}) \\
-&= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
-&= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
-&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
-&= SS(\{\hat{d}_i, i \in D\}) - R
-\end{align*}
+% \begin{align*}
+% Y(\{ r_i, i \in D\}) = \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) = \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
+% \end{align*}
+% where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
+% \begin{align*}
+% SS&(\{r_i, i \in D\}) \\
+% &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
+% &= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
+% &= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
+% &= SS(\{\hat{d}_i, i \in D\}) - R
+% \end{align*}
+% where 
+% \[
+% R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
+% \]
+% We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+
+% To see that $R$ is positive write
+%  \begin{align*}
+% R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
+%  &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
+%  \end{align*}
+%  so it is sufficient to show that
+%  \[
+%  \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
+%  \]
+%  Unfinished, but this proof will work
+% \end{IEEEproof}
+
+\begin{lemma}\label{lem:Qconv}
+\[
+\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\]
 where 
 \[
-R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
-\]
-We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-
-To see that $R$ is positive write
- \begin{align*}
-R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
- &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
- \end{align*}
- so it is sufficient to show that
- \[
- \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
- \]
- Unfinished, but this proof will work
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:Qconv}
-\[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where 
-\[
-H = 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
+H = h_2(0) - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
 \]
@@
 \end{align}
-Since,
+By a Taylor expansion of $\sin(\lambda + \phi)$ about $\phi$ we obtain
 \[
@@
 \]
-where $\zeta(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, then
+where $\zeta(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Then,
 \begin{align*}
@@
 \end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$, and where we put $K = \expect R_1$ is a finite constant.
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$, and $K = \expect R_1$ is a finite constant.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 
-We now analyse the behaviour of $k_2(\lambda)$.
-Firstly consider when, $0 \leq \lambda < \tfrac{\pi}{M}$, then for each integer $k$ and when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$,
-\begin{align*}
-\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} &= \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi k}{M}) \\
-&= \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
+We now analyse the behaviour of $k_2(\lambda)$.  Firstly consider when, $0 \leq \lambda < \tfrac{\pi}{M}$.  For each integer $k$, if $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M})$, then
+\[
+\fracpart{\phi} = \phi - \tfrac{2\pi k}{M}.
+\]
+Similarly if $\phi \in  [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda)$ then
+\[
+\fracpart{\phi + \lambda} = \phi + \lambda - \tfrac{2\pi k}{M},
+\]
+whilst if $\phi \in [\tfrac{ (2k+1)\pi}{M} - \lambda, \tfrac{(2k+1)\pi}{M})$ then
+\[
+\fracpart{\phi + \lambda} = \phi + \lambda - \tfrac{2\pi (k+1)}{M}.
+\]
+So, when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$,
+\[
+\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} = \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi k}{M}),
+\]
+and by a Taylor expansion of $\sin(\lambda + \phi - \tfrac{2\pi k}{M})$ about $\phi - \tfrac{2\pi k}{M}$ we obtain 
+\begin{align*}
+\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} &= \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
 &= \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big).
 \end{align*}
-where $\zeta_l(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} )$,
+where $\zeta_k(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively, when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M})$,
+\[
+\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} = \sin(\lambda + \phi - \tfrac{2\pi(k+1)}{M}) - \sin(\phi - \tfrac{2\pi k}{M})
+\]
+and by a Taylor expansion of $\sin(\lambda + \phi - \tfrac{2\pi(k+1)}{M})$ about $\phi - \tfrac{2\pi(k+1)}{M}$,
+\[
+\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} =  s_k(\phi) + \lambda\big(\cos(\phi - \tfrac{2\pi(k+1)}{M}) + \zeta_{k+1}(\lambda)\big) 
+\] 
+where
+\[
+s_k(\phi) = \sin(\phi - \tfrac{2\pi(k+1)}{M}) - \sin(\phi - \tfrac{2\pi k}{M}).
+\]
+Putting
+\[
+c_k(\phi) = \cos(\phi - \tfrac{2\pi(k+1)}{M}) - \cos(\phi - \tfrac{2\pi k}{M})
+\]
+we obtain
 \begin{align*}
 \sin&\fracpart{\lambda + \phi} - \sin\fracpart{\phi} \\
-&= \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi (k-1)}{M}) \\
-&=  v_k(\phi) + \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
-&= v_k(\phi) + \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big)
-\end{align*}
-where
-\[
-v_k(\phi) = \sin(\phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi (k-1)}{M}).
-\]
-Observe that $v_k(\tfrac{(2k+1)\pi}{M}) = 2\sin(\tfrac{\pi}{M})$.  Now, for all $\phi \in \reals$ we have
+&= v_k(\phi) + \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + c_k(\phi) + \zeta_{k+1}(\lambda)\big) \\
+&= s_k(\phi) + \lambda\big(\cos\fracpart{\phi} + c_k(\phi) + \zeta_{k+1}(\lambda)\big).
+\end{align*}
+
+Observe that both $s_k$ and $c_k$ are continuous functions and that $s_k(\tfrac{(2k+1)\pi}{M}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\tfrac{(2k+1)\pi}{M}) = 0$.  Now, for all $\phi \in \reals$ we have
 \[
@@
 %BPSK plots
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-		\caption{Amplitude error for BPSK}
-		\label{fig:plotamp}
-\end{figure}
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
+%		\caption{Amplitude error for BPSK}
+%		\label{fig:plotamp}
+%\end{figure}
 
@@
 %QPSK plots
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-		\caption{Amplitude error for QPSK}
-		\label{fig:plotamp}
-\end{figure}
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+%		\caption{Amplitude error for QPSK}
+%		\label{fig:plotamp}
+%\end{figure}
 
@@
 %8PSK plots
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-		\caption{Amplitude error for 8PSK}
-		\label{fig:plotamp}
-\end{figure}
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+%		\caption{Amplitude error for 8PSK}
+%		\label{fig:plotamp}
+%\end{figure}
 
@@
 %BPSK plots
- \begin{figure}[tp]
- 	\centering
- 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
- 		\caption{Amplitude error for BPSK}
- 		\label{fig:plotamp}
- \end{figure}
+% \begin{figure}[tp]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
+% 		\caption{Amplitude error for BPSK}
+% 		\label{fig:plotamp}
+ %\end{figure}
 
@@
 %QPSK plots
- \begin{figure}[tp]
- 	\centering
- 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
- 		\caption{Amplitude error for QPSK}
- 		\label{fig:plotamp}
- \end{figure}
+ %\begin{figure}[tp]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+% 		\caption{Amplitude error for QPSK}
+% 		\label{fig:plotamp}
+% \end{figure}
 
@@
 %8PSK plots
- \begin{figure}[tp]
- 	\centering
- 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
- 		\caption{Amplitude error for 8PSK}
- 		\label{fig:plotamp}
- \end{figure}
+% \begin{figure}[tp]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+% 		\caption{Amplitude error for 8PSK}
+% 		\label{fig:plotamp}
+% \end{figure}
 

### 2012-05-17 code/montecarlo.scala

@@ -24,7 +24,7 @@ val starttime = (new java.util.Date).getTime
 for( L <- Ls; M <- Ms ) {
 
   //for a range of different numbers of pilots
-  for( numpilots <- List( L/16, L/4, L ).removeDuplicates ) {
+  for( numpilots <- List( L/8 ).removeDuplicates ) {
 
     val P = 0 until numpilots //pilots at the front
     val D = numpilots until L //data at the back

### 2012-05-18 paper.tex
Added to paper, and edited some of the plots.  Looking good, but there is still a fair bit to do!

@@
 A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
-That is, if we define real random variables $Z > 0$ and $\Theta \in [0,2\pi)$ such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
+That is, if we define real random variables $Z \geq 0$ and $\Theta \in [0,2\pi)$ such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
 If the probability density function of $Z$ is $f_Z(z)$, then the joint density function of $\Theta$ and $Z$ is
@@
 
-We will have particular use of complex random variables of the form $1 + X$.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
+We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
 \[
@@
 \begin{lemma}\label{lem:h1minedcircsym}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
@@
 \]
-Then $h_1(x)$ is uniquley maximised as $0$ over the interval $[-\pi,\pi)$.
-\end{lemma}
+Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
+\end{lemma}
+Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
+That is, the pdf of $Z e^{j \Theta}$ in rectangular coordinates is given by $z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$ and $x$ and $y$ are the real and imaginary parts of $Z e^{j \Theta}$, and this pdf is non increasing with $z$.
+For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
+
+By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if 
+\[
+\abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
+\]
+We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
+
 \begin{IEEEproof}
@@
 
-\section{Statistical properties of the least squares estimator}
-
-We use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
+\section{Statistical properties of the least squares estimator}\label{sec:stat-prop-least}
+
+In this section we decribe the asymptotic properties of the least squares estimator.
+In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
 \[
@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, zero mean, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be noisy symbols given by the signal model~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be noisy symbols given by the signal model~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
@@
 \] 
-Let $R_i \geq 0$ and $\Phi_i \in [-\pi,\pi)$ be real random variables satisfying
+Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
 \begin{equation}\label{eq:RiandPhii}
@@
 \begin{lemma}
-Under the same conditions as Lemma~\ref{lem:h1minedcircsym}, define the functions $h_2$ and $k_2$, 
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+Let $h_2$ and $k_2$ be the functions as,
 \[
@@
 \]
-Then $h_2(x)$ is uniquely maximised at $x=0$ other the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$ and $k_2(0) = 0$.
-\end{lemma}
+Then $h_2(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$ and $k_2(0) = 0$.
+\end{lemma}
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
 
@@
 \begin{equation}\label{eq:hatrhoZ}
-\hat{\rho} = \frac{Z(\theta) + Z(\theta)^*}{2L} = \frac{1}{L}\Re(Z(\theta))
+\hat{\rho} = \frac{Z(\theta) + Z(\theta)^*}{2L} = \frac{1}{L}\Re(Z(\theta)),
 \end{equation}
@@
 \end{align*}
-since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  When $i \in P$,  
+since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
 \[
@@
 Let 
-\[
+\begin{equation}\label{eq:GLdefn}
 G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
-\]
+\end{equation}
 and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle$.
@@
 
-Observe that
+Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
 \[
@@
 \]
-where the function $G$, $h_1$ and $h_2$ are defined in the statement of Theorem~\ref{thm:consistency}.  Lemma~\ref{lem:convtoexpGlamL} shows that,
+Lemma~\ref{lem:convtoexpGlamL} shows that,
 \[
@@
 \]
-where $\prob$ denotes probability.  Since $G(\lambda)$ is continuous and is uniquely maximised at $0$, then for any $\delta > 0$ there is a $\epsilon > 0$ such that 
-\[
-\sabs{\hat{\lambda}_L} > \delta \;\; \Rightarrow \;\; \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon
+where $\prob$ denotes probability.  Since $G(\lambda)$ is continuous and is uniquely maximised at $0$, then for any $\delta > 0$ there is a $\epsilon > 0$ such that if 
+\[
+\sabs{\hat{\lambda}_L} > \delta \qquad \text{then} \qquad \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon,
 \]
@@
 
-It remains to prove part two of this theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
+It remains to prove part two of this theorem regarding the convergence of the least squares amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
 \begin{equation}\label{eq:rhoGLZ}
@@
 
+We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
+Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
 The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
@@
 \]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.  Lemma~\ref{lem:empiricprocc} shows that
+where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$ and where the constant real number $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+Lemma~\ref{lem:empiricprocc} shows that
 \[
@@
 \]
-It follows that
+It follows from the three equations above that,
 \[
@@
 \]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ and it follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constant defined in the statement of Thereom~\ref{thm:normality}.
+It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
 \[
@@
 
-We now analyse the asymptotic distribution of $m_L = \hat{\rho} - \rho_0 G(0)$.  Define the function $H_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
+We now analyse the asymptotic distribution of $m_L = \hat{\rho} - \rho_0 G(0)$.  Define the function $T(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
 \begin{align*}
 \sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) + H_L(\hat{\lambda}_L) - G(0) \big)
+&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L) + T(\hat{\lambda}_L) - G(0) \big)
 \end{align*}
@@
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - H_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
+\]
+where $X_L = \sqrt{L}\big( G_L(0) - T(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
+\[
+\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
 \]
@@
 \]
-The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondinly
+The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondingly
 \[
@@
 \begin{IEEEproof}
+BLERG
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:Qconv}
-\[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where 
-\[
-H = h_2(0) - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
-\]
-\end{lemma}
-\begin{IEEEproof}
-Write
-\[
-Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0) = \frac{\abs{P}}{L} k_1(\lambda) + \frac{\abs{D}}{L} k_2(\lambda)
+$\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)$
+\end{lemma}
+\begin{IEEEproof}
+We have
+\[
+Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0) = \tfrac{\abs{P}}{L} k_1(\lambda) + \tfrac{\abs{D}}{L} k_2(\lambda)
 \]
@@
 \end{align}
-By a Taylor expansion of $\sin(\lambda + \phi)$ about $\phi$ we obtain
-\[
-\sin(\lambda + \phi) - \sin(\phi) = \lambda \big( \cos(\phi) + \zeta(\lambda) \big),
+We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\Phi_1$ we obtain
+\[
+\sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
 \]
@@
 \end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$, and $K = \expect R_1$ is a finite constant.
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ is circularly symmetric, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite (by assumption in Theorem~\ref{thm:consistency}).
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 \]
-
-We now analyse the behaviour of $k_2(\lambda)$.  Firstly consider when, $0 \leq \lambda < \tfrac{\pi}{M}$.  For each integer $k$, if $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M})$, then
-\[
-\fracpart{\phi} = \phi - \tfrac{2\pi k}{M}.
-\]
-Similarly if $\phi \in  [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda)$ then
-\[
-\fracpart{\phi + \lambda} = \phi + \lambda - \tfrac{2\pi k}{M},
-\]
-whilst if $\phi \in [\tfrac{ (2k+1)\pi}{M} - \lambda, \tfrac{(2k+1)\pi}{M})$ then
-\[
-\fracpart{\phi + \lambda} = \phi + \lambda - \tfrac{2\pi (k+1)}{M}.
-\]
-So, when $\phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda )$,
-\[
-\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} = \sin(\lambda + \phi - \tfrac{2\pi k}{M}) - \sin(\phi - \tfrac{2\pi k}{M}),
-\]
-and by a Taylor expansion of $\sin(\lambda + \phi - \tfrac{2\pi k}{M})$ about $\phi - \tfrac{2\pi k}{M}$ we obtain 
-\begin{align*}
-\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} &= \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
-&= \lambda\big(\cos\fracpart{\phi} + \zeta_k(\lambda)\big).
-\end{align*}
-where $\zeta_k(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively, when $\phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M})$,
-\[
-\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} = \sin(\lambda + \phi - \tfrac{2\pi(k+1)}{M}) - \sin(\phi - \tfrac{2\pi k}{M})
-\]
-and by a Taylor expansion of $\sin(\lambda + \phi - \tfrac{2\pi(k+1)}{M})$ about $\phi - \tfrac{2\pi(k+1)}{M}$,
-\[
-\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} =  s_k(\phi) + \lambda\big(\cos(\phi - \tfrac{2\pi(k+1)}{M}) + \zeta_{k+1}(\lambda)\big) 
-\] 
+Lemma~\ref{lem:k2conv} shows that
+\[
+k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).
+\]
+So,
+\begin{align*}
+Q(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( h_2(0) + H + o_P(1) ) \\
+&= \hat{\lambda}_L \left( \tfrac{\abs{P}}{L}  + \tfrac{\abs{D}}{L} H + o_P(1) \right).
+\end{align*}
+The proof follows since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:k2conv}
+$k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
+\end{lemma}
+\begin{IEEEproof}
+Since $\hat{\lambda}_L \in [-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ it is only the behaviour of the function $k_2(\lambda)$ for $\lambda\in [-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ that is relevant.
+We will analyse $k_2(\lambda)$ for non negative $\lambda$, i.e., when $0 \leq \lambda < \tfrac{\pi}{M}$.
+An analagous argument follows when $\lambda$ is negative, i.e., $-\tfrac{\pi}{M} \leq \lambda < 0$.
+
+To keep our notation clean put
+\[
+\psi_k = (2k+1)\tfrac{\pi}{M}.
+\]
+For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
+\[
+\fracpart{\Phi_1} = \Phi_1 - \tfrac{2\pi k}{M}.
+\]
+Similarly if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$ then
+\[
+\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi k}{M},
+\]
+whilst if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$ then
+\[
+\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi (k+1)}{M}.
+\]
+So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
+&= \sin(\lambda + \Phi_1 - \tfrac{2\pi k}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}),
+\end{align*}
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi k}{M})$ about $\Phi_1 - \tfrac{2\pi k}{M}$ we obtain 
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
+&= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
+\end{align*}
+where $\zeta_k(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
+&= \sin(\lambda + \Phi_1 - \tfrac{2\pi(k+1)}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M})
+\end{align*}
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi(k+1)}{M})$ about $\Phi_1 - \tfrac{2\pi(k+1)}{M}$,
+\begin{align*}
+\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
+&=  s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi(k+1)}{M}) + \zeta_{k+1}(\lambda)\big) 
+\end{align*}
 where
 \[
-s_k(\phi) = \sin(\phi - \tfrac{2\pi(k+1)}{M}) - \sin(\phi - \tfrac{2\pi k}{M}).
+s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi(k+1)}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}).
 \]
@@
 \[
-c_k(\phi) = \cos(\phi - \tfrac{2\pi(k+1)}{M}) - \cos(\phi - \tfrac{2\pi k}{M})
-\]
-we obtain
-\begin{align*}
-\sin&\fracpart{\lambda + \phi} - \sin\fracpart{\phi} \\
-&= v_k(\phi) + \lambda\big(\cos(\phi - \tfrac{2\pi k}{M}) + c_k(\phi) + \zeta_{k+1}(\lambda)\big) \\
-&= s_k(\phi) + \lambda\big(\cos\fracpart{\phi} + c_k(\phi) + \zeta_{k+1}(\lambda)\big).
-\end{align*}
-
-Observe that both $s_k$ and $c_k$ are continuous functions and that $s_k(\tfrac{(2k+1)\pi}{M}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\tfrac{(2k+1)\pi}{M}) = 0$.  Now, for all $\phi \in \reals$ we have
-\[
-\sin\fracpart{\lambda + \phi} - \sin\fracpart{\phi} = v(\lambda, \phi) + \lambda\cos\fracpart{\phi}
-\]
-where $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$ and where
-\[
-v_k(\lambda,\phi) = \begin{cases}
-v_k(\phi) + \lambda\zeta_k(\lambda) &  \phi \in [ \tfrac{(2k+1)\pi}{M} - \lambda,\tfrac{(2k+1)\pi}{M} ) \\
-\lambda\zeta_k(\lambda) & \phi \in [\tfrac{ (2k-1)\pi}{M}, \tfrac{(2k+1)\pi}{M} - \lambda ) \\
+c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi(k+1)}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M})
+\]
+we can write
+\begin{align*}
+\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
+&= s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi k}{M}) + c_k(\Phi_1) + \zeta_{k+1}(\lambda)\big) \\
+&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1) \big).
+\end{align*}
+where, for convenience we put $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.
+Observe that both $s_k$ and $c_k$ are continuous functions and that $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\psi_{k}) = 0$.
+
+Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
+\]
+whilst if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
+\]
+Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$ then, for $\Phi_1 \in \reals$ we can write
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = v(\lambda, \Phi_1) + \lambda\cos\fracpart{\Phi_1}
+\]
+where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and where
+\[
+v_k(\lambda,\Phi_1) = \begin{cases}
+v_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
 0 & \text{otherwise}.
@@
 \end{align*}
-It remains to evaluate $\expect R_1 v(\lambda, \Phi_1)$.  Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
+Define the function $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where
+\[
+H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g\big((2k+1)\tfrac{\pi}{M}\big),
+\]
+and where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.
+Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\gamma(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
+$\gamma(\hat{\lambda}_L) = o_P(1)$ and therefore
+\[
+k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( h_2(0) + H_1 + o_P(1) ) = \hat{\lambda}_L ( H + o_P(1) )
+\]
+as required
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:expRvlamphi}
+$W(\hat{\lambda}_L) =  \hat{\lambda}_L(H + o_P(1))$
+\end{lemma}
+\begin{IEEEproof}
+Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
 \begin{align*}

### 2012-05-18 code/montecarlo.scala

@@ -24,7 +24,7 @@ val starttime = (new java.util.Date).getTime
 for( L <- Ls; M <- Ms ) {
 
   //for a range of different numbers of pilots
-  for( numpilots <- List( L/8 ).removeDuplicates ) {
+  for( numpilots <- List( L/8, L/4, L ).removeDuplicates ) {
 
     val P = 0 until numpilots //pilots at the front
     val D = numpilots until L //data at the back

### 2012-05-19 paper.tex
Added to paper.  Continued cleaning.  Keeps getting better, but there's alot of work to do here!

@@
 We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
+Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
 Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
 The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This precludes the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
+This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
 \begin{align*}
@@
 \]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$ and where the constant real number $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.
+The constant real numbers $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
 Lemma~\ref{lem:empiricprocc} shows that
@@
 \]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constant defined in the statement of Thereom~\ref{thm:normality}.
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
 It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
@@
 
-We now analyse the asymptotic distribution of $m_L = \hat{\rho} - \rho_0 G(0)$.  Define the function $T(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ} we have
+We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Define the function $T(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
 \begin{align*}
 \sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L) + T(\hat{\lambda}_L) - G(0) \big)
+&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L) + T(\hat{\lambda}_L) - G(0) \big).
 \end{align*}
@@
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L)  \big) = o_P(1) + X_L 
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
 \]
@@
 \]
+It remains to prove the lemmas that we have used.
 
@@
 \]
-Lemma~\ref{lem:k2conv} shows that
-\[
-k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).
-\]
-So,
-\begin{align*}
-Q(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( h_2(0) + H + o_P(1) ) \\
+Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
+\begin{align*}
+Q(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
 &= \hat{\lambda}_L \left( \tfrac{\abs{P}}{L}  + \tfrac{\abs{D}}{L} H + o_P(1) \right).
@@
 We will analyse $k_2(\lambda)$ for non negative $\lambda$, i.e., when $0 \leq \lambda < \tfrac{\pi}{M}$.
-An analagous argument follows when $\lambda$ is negative, i.e., $-\tfrac{\pi}{M} \leq \lambda < 0$.
-
+An analagous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
 To keep our notation clean put
 \[
-\psi_k = (2k+1)\tfrac{\pi}{M}.
+\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
 \]
@@
 \[
-\fracpart{\Phi_1} = \Phi_1 - \tfrac{2\pi k}{M}.
-\]
-Similarly if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$ then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi k}{M},
-\]
-whilst if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$ then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi (k+1)}{M}.
+\fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
+\]
+Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
+\[
+\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
+\]
+whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
+\[
+\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
 \]
@@
 \sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi k}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}),
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi k}{M})$ about $\Phi_1 - \tfrac{2\pi k}{M}$ we obtain 
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_k(\lambda)\big) \\
+&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
+\end{align*}
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\Phi_1 - \tfrac{2\pi}{M}k$ we obtain 
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
 &= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
@@
 \begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi(k+1)}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M})
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi(k+1)}{M})$ about $\Phi_1 - \tfrac{2\pi(k+1)}{M}$,
+\sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
+&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
+\end{align*}
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}$,
 \begin{align*}
 \sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&=  s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi(k+1)}{M}) + \zeta_{k+1}(\lambda)\big) 
+&=  s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) + \zeta_{k+1}(\lambda)\big) 
 \end{align*}
@@
 \[
-s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi(k+1)}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}).
-\]
-Putting
-\[
-c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi(k+1)}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M})
+s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k).
+\]
+Putting,
+\[
+c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M})
 \]
@@
 \end{align*}
-where, for convenience we put $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.
+where, for convenience, we put $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.
 Observe that both $s_k$ and $c_k$ are continuous functions and that $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\psi_{k}) = 0$.
@@
 \]
-whilst if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
+whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
 \[
@@
 \]
-Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$ then, for $\Phi_1 \in \reals$ we can write
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = v(\lambda, \Phi_1) + \lambda\cos\fracpart{\Phi_1}
-\]
-where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and where
-\[
+Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, then, for all $\Phi_1 \in \reals$,
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1) 
+\]
+where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
+\begin{equation}\label{eq:vk}
 v_k(\lambda,\Phi_1) = \begin{cases}
-v_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
 \lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
@@
 \end{cases}
-\]
+\end{equation}
 Now
@@
 k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
-&= \expect R_1\big( v(\lambda, \Phi_1) + \lambda\cos\fracpart{\Phi_1} \big) \\
-&= \expect R_1 v(\lambda, \Phi_1) + \lambda h_2(0).
-\end{align*}
-Define the function $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where
-\[
-H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g\big((2k+1)\tfrac{\pi}{M}\big),
-\]
-and where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.
+&= \expect R_1\big( \lambda\cos\fracpart{\Phi_1} + v(\lambda, \Phi_1)  \big) \\
+&= \lambda h_2(0) + \expect R_1 v(\lambda, \Phi_1).
+\end{align*}
+Define the function $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, and
+\[
+H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0).
+\]
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\gamma(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
-$\gamma(\hat{\lambda}_L) = o_P(1)$ and therefore
+$\gamma(\hat{\lambda}_L) = o_P(1)$, and therefore
 \[
@@
 \]
-as required
+as required.  It remains to prove Lemma~\ref{lem:expRvlamphi}.
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:expRvlamphi}
-$W(\hat{\lambda}_L) =  \hat{\lambda}_L(H + o_P(1))$
-\end{lemma}
-\begin{IEEEproof}
-Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
-\begin{align*}
-k_2(\lambda) &= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \\
-&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi.
-\end{align*}
-Since $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$,
-\[
-k_2(\lambda) = \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi
-\]
-and since $v_k(\lambda,\phi) = 0$ for $\phi$ in the interval $[0,2\pi)$ when $k \notin \{0,1,\dots,M-1\}$ the infinite sum above can be truncated so that
-\[
-k_2(\lambda) = \sum_{k = 0}^{M-1} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi
-\]
-(BLERG this needs modification, more care)
-Now
+$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.
+\end{lemma}
+\begin{IEEEproof}
+In this lemma, as Lemma~\ref{lem:k2conv}, we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
+An analgous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
+Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$, then,
+\begin{align*}
+W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) = \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \\
+&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi,
+\end{align*}
+and, since $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$,
+\begin{equation}\label{eq:Winfsum}
+W(\lambda) = \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi.
+\end{equation}
+Now, when $0 < k < M$, we have $0 < \psi_{k-1} < \psi_k < 2\pi$, so, using~\eqref{eq:vk},
+\begin{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
+&\hspace{1cm} +  \int_{\psi_{k-1}}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi.
+\end{align*}
+When $k = 0$, we have $\psi_{k-1} < 0 < \psi_k - \lambda < \psi_k < 2\pi$, so that
+\begin{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{0}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
+&\hspace{1cm} +  \int_{0}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi ,
+\end{align*}
+and when $k = M-1$, we have $0 < \psi_{k-1} < 2\pi < \psi_k - \lambda$, so that
+\begin{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \int_{\psi_{k-1}}^{2\pi}\lambda\zeta_k(\lambda)g(\phi) d\phi.
+\end{align*}
+Otherwise, when $k < 0$ or $k > M$, 
+\[
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0
+\]
+So the infinite sum in~\eqref{eq:Winfsum} is actually a finite sum,
+\begin{align*}
+W(\lambda) &= \sum_{k=0}^{M} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \sum_{k=0}^{M-1}\int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
+&\hspace{0.5cm} + \sum_{k=0}^{M-1} \int_{\psi_{k-1}}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi \\
+&\hspace{0.5cm} +  \int_{0}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi + \int_{\psi_{k-1}}^{2\pi}\lambda\zeta_k(\lambda)g(\phi) d\phi + 
+\end{align*}
 \begin{align*}

### 2012-05-20 paper.tex
Added to paper.

@@
 \begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
-&\hspace{1cm} +  \int_{\psi_{k-1}}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi.
-\end{align*}
-When $k = 0$, we have $\psi_{k-1} < 0 < \psi_k - \lambda < \psi_k < 2\pi$, so that
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{0}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
-&\hspace{1cm} +  \int_{0}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi ,
-\end{align*}
-and when $k = M-1$, we have $0 < \psi_{k-1} < 2\pi < \psi_k - \lambda$, so that
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \int_{\psi_{k-1}}^{2\pi}\lambda\zeta_k(\lambda)g(\phi) d\phi.
-\end{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi,
+\end{align*}
+where, for each integer $k = 1, \dots, M-1$, the functions $a_k$ and $b_k$ are
+\begin{equation}\label{eq:ak}
+a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
+\end{equation}
+and 
+\[
+b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi.
+\]
+Observe that $a_k(\lambda)$ and $b_k(\lambda)$ are functions that go to zero as $\lambda$ goes to zero.
+Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
+\begin{align*}
+\int_0^{2\pi} v_0(\lambda, \phi) &g(\phi) d\phi = \int_{0}^{\psi_{k}} v_0(\lambda, \phi) g(\phi) d\phi \\
+&= \lambda a_0(\lambda) + \lambda b_0(\lambda) + \int_{\psi_{0} - \lambda}^{\psi_{0}} s_0(\phi) g(\phi) d\phi,
+\end{align*}
+where $a_0(\lambda)$ is defined as in~\eqref{eq:ak} and 
+\[
+b_0(\lambda) = \zeta_0(\lambda) \int_{0}^{\psi_k-\lambda} g(\phi) d\phi.
+\]
+When $k = M$, we have $0 < \psi_{M-1} < 2\pi < \psi_{M} - \lambda$, so that
+\begin{align*}
+\int_0^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{M-1}}^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi \\
+&= \lambda b_{M}(\lambda)
+\end{align*}
+where 
+\[
+b_{M}(\lambda) = \zeta_{M}(\lambda) \int_{\psi_{M-1}}^{2\pi}g(\phi) d\phi.
+\]
 Otherwise, when $k < 0$ or $k > M$, 
 \[
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0
-\]
-So the infinite sum in~\eqref{eq:Winfsum} is actually a finite sum,
-\begin{align*}
-W(\lambda) &= \sum_{k=0}^{M} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \sum_{k=0}^{M-1}\int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0.
+\]
+So the infinite sum in~\eqref{eq:Winfsum} can be written as the finite sum,
+\begin{align*}
+W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \sum_{k=0}^{M-1}\int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi)g(\phi)d\phi + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
 &\hspace{0.5cm} + \sum_{k=0}^{M-1} \int_{\psi_{k-1}}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi \\

### 2012-05-20 paper.tex
further small edits

--- 2012-05-19 paper.tex
@@
 \[
-H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M}),
+H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
 \]
@@
 \begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
-&\hspace{1cm} +  \int_{\psi_{k-1}}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi.
-\end{align*}
-When $k = 0$, we have $\psi_{k-1} < 0 < \psi_k - \lambda < \psi_k < 2\pi$, so that
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{0}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
-&\hspace{1cm} +  \int_{0}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi ,
-\end{align*}
-and when $k = M-1$, we have $0 < \psi_{k-1} < 2\pi < \psi_k - \lambda$, so that
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \int_{\psi_{k-1}}^{2\pi}\lambda\zeta_k(\lambda)g(\phi) d\phi.
-\end{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
+\end{align*}
+where, for each integer $k = 1, \dots, M-1$, the functions $a_k$ and $b_k$ are
+\begin{equation}\label{eq:ak}
+a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
+\end{equation}
+\[
+b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
+\]
+and
+\begin{equation}\label{eq:ck}
+c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
+\end{equation}
+Observe that $a_k(\lambda)$ and $b_k(\lambda)$ are functions that converge to zero as $\lambda$ goes to zero, whilst $c_k(\lambda)$ converges to 
+\[
+s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
+\]
+Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
+\[
+\int_0^{2\pi} v_0(\lambda, \phi) g(\phi) d\phi = \lambda a_0(\lambda) + \lambda b_0(\lambda) +  \lambda c_0(\lambda),
+\]
+where $a_0(\lambda)$ and $c_0(\lambda)$ are defined as in~\eqref{eq:ak}~and~\eqref{eq:ck}, and
+\[
+b_0(\lambda) = \zeta_0(\lambda) \int_{0}^{\psi_k-\lambda} g(\phi) d\phi.
+\]
+When $k = M$, we have $0 < \psi_{M-1} < 2\pi < \psi_{M} - \lambda$, so that
+\[
+\int_0^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi = \lambda b_{M}(\lambda)
+\]
+where 
+\[
+b_{M}(\lambda) = \zeta_{M}(\lambda) \int_{\psi_{M-1}}^{2\pi}g(\phi) d\phi.
+\]
 Otherwise, when $k < 0$ or $k > M$, 
 \[
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0
-\]
-So the infinite sum in~\eqref{eq:Winfsum} is actually a finite sum,
-\begin{align*}
-W(\lambda) &= \sum_{k=0}^{M} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \sum_{k=0}^{M-1}\int_{\psi_{k} - \lambda}^{\psi_{k}} (s_k(\phi) + \lambda\eta_k(\lambda, \phi)) g(\phi) d\phi \\
-&\hspace{0.5cm} + \sum_{k=0}^{M-1} \int_{\psi_{k-1}}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi \\
-&\hspace{0.5cm} +  \int_{0}^{\psi_{k} - \lambda}\lambda\zeta_k(\lambda)g(\phi) d\phi + \int_{\psi_{k-1}}^{2\pi}\lambda\zeta_k(\lambda)g(\phi) d\phi + 
-\end{align*}
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = &\int_{\tfrac{(2k+1)\pi}{M} - \lambda}^{\tfrac{(2k+1)\pi}{M}}v_k(\phi)g(\phi)d\phi \\
-&\hspace{0.5cm}+ \lambda\zeta_k(\lambda) \int_{\tfrac{(2k-1)\pi}{M}}^{\tfrac{(2k+1)\pi}{M}}g(\phi)d\phi.
-\end{align*}
-Now, since $g(\phi)$ is continuous at $(2k+1)\tfrac{\pi}{M}$ for each integer $k$, and $v_k(\phi)$ is continuous and $v_k(\tfrac{(2k+1)\pi}{M}) =  2\sin(\tfrac{\pi}{M})$,
-\begin{align*}
-\int_{\tfrac{(2k+1)\pi}{M} - \lambda}^{\tfrac{(2k+1)\pi}{M}}&v_k(\phi)g(\phi)d\phi \\
-&= \lambda( v_k(\tfrac{(2k+1)\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) ) \\
-&= \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) )
-\end{align*}
-where $\nu_k(\lambda)$ is function that goes to zero as $\lambda$ goes to zero.
-To keep the notation clean put $A_k = \int_{\tfrac{(2k-1)\pi}{M}}^{\tfrac{(2k+1)\pi}{M}}g(\phi)d\phi$.
-For each $k\in\ints$, $A_k$ is a nonnegative finite constant less than one.
-Now
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi \\
-&= \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k )
-\end{align*}
-and
-\begin{align*}
-k_2(\lambda) &= \sum_{k = 0}^{M-1} \lambda( 2\sin(\tfrac{\pi}{M})g(\tfrac{(2k+1)\pi}{M}) + \nu_k(\lambda) + \zeta_k(\lambda) A_k ) \\
- &= \epsilon(\lambda) + 2 \lambda \sin(\tfrac{\pi}{M}) \sum_{k = 0}^M g(\tfrac{(2k+1)\pi}{M})
-\end{align*}
-where $\epsilon(\lambda) = \sum_{k=0}^M\big( \nu_k(\lambda) + \zeta_k(\lambda) A_k \big)$ is a function that goes to zero as $\lambda$ goes to zero.
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\epsilon(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
-$\epsilon(\hat{\lambda}_L) = o_P(1)$.
-So
-\[
-k_2(\hat{\lambda}_L) = o_P(1) + 2 \hat{\lambda}_L \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{(2k+1)\pi}{M})
-\]
-as required.
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0.
+\]
+So the infinite sum in~\eqref{eq:Winfsum} can be written as the finite sum,
+\begin{align*}
+W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
+&= \lambda \left( C_k(\lambda)  + A_k(\lambda) + B_k(\lambda) \right).
+\end{align*}
+where 
+\[
+A_k(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B_k(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
+\]
+\[
+C_k(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
+\]
+Observe both $A_k(\lambda)$ and $B_k(\lambda)$ converges to zero as as $\lambda$ goes to zero, and $C_k(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
+\[
+\gamma(\lambda) = A_k(\lambda) + B_k(\lambda) + C_k(\lambda) - H_1.
+\]
+Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$.  The proof follows.
 \end{IEEEproof}

### 2012-05-22 paper.tex
Added a bunch of new data files.  Added a bit more to the paper.

@@
 Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+%If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, Then
+Then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
 \[
@@
 \[
-\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
+\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
 \]
@@
 \[
-\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = BLERG.
+\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = ?.
 \]
@@
 \]
-where $\zeta(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Then,
+where $\zeta$ is a continuous function with $\zeta(0) = 0$.  So,
 \begin{align*}
@@
 since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ is circularly symmetric, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite (by assumption in Theorem~\ref{thm:consistency}).
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
+Because $\zeta$ with $\zeta(0) = 0$ and because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
 $\zeta(\hat{\lambda}_L) = o_P(1)$.
@@
 \end{align*}
-where $\zeta_k(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+where, for each integer $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
 \begin{align*}
@@
 \end{align*}
-where, for convenience, we put $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.
+where, for convenience, we have defined the function, 
+\[
+\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda).
+\]  
 Observe that both $s_k$ and $c_k$ are continuous functions and that $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\psi_{k}) = 0$.
+%It follows, since $\zeta_{k+1}$ is also continuous, that the function $\eta_k(\lambda, \Phi_1)$ is continuous in both $\lambda$ and $\Phi_1$, and $\eta_k(0, \psi_k) = 0$.
 
@@
 \end{equation}
-Now
+Now we can write,
 \begin{align*}
@@
 \end{align*}
-Define the function $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, and
+Define the function $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$, and
 \[
@@
 \begin{lemma}\label{lem:expRvlamphi}
-$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.
-\end{lemma}
-\begin{IEEEproof}
-In this lemma, as Lemma~\ref{lem:k2conv}, we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
+$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
+\end{lemma}
+\begin{IEEEproof}
+In this lemma, as in Lemma~\ref{lem:k2conv}, we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
 An analgous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
@@
 \end{align*}
-and, since $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$,
+and, since $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$, then,
 \begin{equation}\label{eq:Winfsum}
@@
 \begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
 &= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
 \end{align*}
-where, for each integer $k = 1, \dots, M-1$, the functions $a_k$ and $b_k$ are
+where, for each integer $k = 1, \dots, M-1$, the functions $a_k$, $b_k$, and $c_k$ are
 \begin{equation}\label{eq:ak}
@@
 \]
-and
 \begin{equation}\label{eq:ck}
@@
 \end{equation}
-Observe that $a_k(\lambda)$ and $b_k(\lambda)$ are functions that converge to zero as $\lambda$ goes to zero, whilst $c_k(\lambda)$ converges to 
-\[
+Observe that $a_k(\lambda)$ and $b_k(\lambda)$ are functions that converge to zero as $\lambda$ goes to zero from above.  Since $g$ is continuous at $\psi_k$ then $c_k(\lambda)$ converges to 
+\begin{equation}\label{eq:skconverge}
 s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
-\]
+\end{equation}
+as $\lambda$ goes to zero from above~\footnote{We actually only require $g$ to be continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$, and the requirements of Theorem~\ref{thm:normality} could be weakened in this sense.
+Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.}.
+We are only interested in the limits from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
+The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
+
 Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
@@
 \]
-Otherwise, when $k < 0$ or $k > M$, 
+Both $b_0(\lambda)$ and $b_M(\lambda)$ are functions that converge to zero as $\lambda$ approaches zero from above.
+
+Finally, when $k < 0$ or $k > M$, 
 \[
@@
 &= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \left( C_k(\lambda)  + A_k(\lambda) + B_k(\lambda) \right).
+&= \lambda \left( C(\lambda)  + A(\lambda) + B(\lambda) \right).
 \end{align*}
@@
 \[
-A_k(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B_k(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
-\]
-\[
-C_k(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
-\]
-Observe both $A_k(\lambda)$ and $B_k(\lambda)$ converges to zero as as $\lambda$ goes to zero, and $C_k(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
-\[
-\gamma(\lambda) = A_k(\lambda) + B_k(\lambda) + C_k(\lambda) - H_1.
-\]
-Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$.  The proof follows.
+A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
+\]
+\[
+C(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
+\]
+Observe, that the function $A(\lambda)$ and $B(\lambda)$ converge to to zero as as $\lambda$ approaches zero from above, and, using~\eqref{eq:skconverge}, the function $C(\lambda)$ converges to 
+\[
+H_1 =  - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
+\]
+as $\lambda$ converges to zero from above.  Put
+\[
+\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1.
+\]
+Then $\gamma(\lambda)$ converges to zero as $\lambda$ approaches to zero from above and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$.  The proof follows.
 \end{IEEEproof}

### 2012-07-30 paper.tex
Added a bit more data, minor paper edits.
Edits.

--- 2012-05-20 paper.tex
@@
 \begin{abstract}
-We describe least squares estimators of carrier phase and amplitude from a noisy communications signal that contains both pilots, known to the receiver, and data, unknown to the receiver.
+We describe least squares estimators of carrier phase and amplitude from a noisy communications signal that contains both pilot signals, known to the receiver, and data signals, unknown to the receiver.
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-PSK.
@@
 Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included, so as to made the estimator coherent.
+A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included, so as to make the estimator coherent.
 Techniques are available but they are somewhat adhoc~\cite{Cowley_ref_sym_carr_1998}.
@@
 Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
-This is motivated by the fact that in our intended application (narrowband satellite communications) and in many modern communication systems the data symbols are \emph{coded}.
+This is motivated by the fact that in our motivating application (narrowband satellite communications) and in many modern communication systems the data symbols are \emph{coded}.
 For this reason raw symbol error rate is not of specific interest at this stage.
@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ is the expected value operator.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
 It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
@@
 \end{align}
-where $*$ denote the complex congugate.
+where $*$ denotes the complex congugate.
 Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find that the least squares estimator of $a_0$, satisfies
@@
 Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to surpress the argument $(\{ d_i, i \in D\})$ for notational clarity.
-Substituting $\hat{a} = \frac{1}{L}Y$ into~\eqref{eq:SS} we obtain the sum of squares function, conditioned on minimisation with respect to $a$,
+Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain the $SS$ conditioned on minimisation with respect to $a$,
 \begin{equation}\label{eq:SSdatasymbols}
@@
 where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.
-Observe that given a candidate value for the data symbols, we can compute the corresponding sum of squares function $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
+Observe that given a candidate value for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
 It turns out that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
@@
 
-To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a nonnegative real.  Now the sum of squares function can be written as
+To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a nonnegative real.  Now,
 \begin{align}
@@
 \end{equation}
-and where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
+where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
 A word of caution, the notation $\round{\cdot}$ is often used to denote rounding to the nearest \emph{integer}.
@@
 
-We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Define $b_i = \hat{d}_i(0)$ and the real number 
-\[
-z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}},
-\]
-then,
+We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Let $b_i = \hat{d}_i(0)$ and let 
+\[
+z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}.
+\]
+Then,
 \begin{equation}\label{eq:uicombos}
@@
 
-Let $f(\theta)$ be the sequence $\{ \hat{d}_i(\theta), i \in D \}$.
-So, $f(\theta)$ is a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
+Let 
+\[
+f(\theta) = \{ \hat{d}_i(\theta), i \in D \}
+\]
+be a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
 Observe that $f(\theta)$ is piecewise continuous.
 The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains contant are determined by the values of $\{z_i, i \in D\}$.
-
-Let $S$ be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$, that is,
-\[
-S = \{ f(\theta) \mid \theta \in [0, 2 \pi) \}.
-\]
-If $\hat{\theta}$ is the least squares estimator the phase, then it follows from~\eqref{eq:hatdfinxtheta}, that $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
+Let
+\[
+S = \{ f(\theta) \mid \theta \in [0, 2 \pi) \}
+\]
+be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$.
+If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
 Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
 
-The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indicies in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
+The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indices in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
 \begin{equation}\label{eq:sigmasortind}
@@
 
-We want to find the sequence $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A na\"{\i}ve approach would be to compute $SS(f_k)$ for each sequence in $S$ and return the minimiser.
+We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
+A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,(M+1)\abs{D}-1\}$.
 Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
@@
 \end{align*}
-where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in the sequence $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
+where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
 Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
@@
 \end{align*}
-where $\eta = (e^{j2\pi/M} - 1)$.  In general we have,
+where $\eta = (e^{j2\pi/M} - 1)$.  In general,
 \[
@@
 Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.
-Let $\hat{k}$ be the integer that minimises $SS(f_k)$, i.e.
-$\hat{k} = \arg\min SS(f_k)$.
-The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata}, as
+Let $\hat{k} = \arg\min SS(f_k)$.
+The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata},
 \begin{equation}\label{eq:ahatYhat}
@@
 Psuedocode is given in Algorithm~\ref{alg:loglinear}.
-Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given the sequence $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ of the indices in $D$ as described in~\eqref{eq:sigmasortind}.
+Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
 The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
@@
 A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
-That is, if we define real random variables $Z \geq 0$ and $\Theta \in [0,2\pi)$ such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
-If the probability density function of $Z$ is $f_Z(z)$, then the joint density function of $\Theta$ and $Z$ is
+That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
+If the probability density function of $Z$ is $f_Z(z)$, then the joint probability density function (pdf) of $\Theta$ and $Z$ is
 \[
@@
 Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariant to phase rotation.
-If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean since
+If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
 \begin{align*}
@@
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
-Let $h_1(x)$ be the function
+Let
 \[
@@
 \end{lemma}
-Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
+% Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
 That is, the pdf of $Z e^{j \Theta}$ in rectangular coordinates is given by $z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$ and $x$ and $y$ are the real and imaginary parts of $Z e^{j \Theta}$, and this pdf is non increasing with $z$.
@@
 
-By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if 
-\[
-\abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
-\]
-We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
-
+% By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if
+% \[
+% \abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
+% \]
+% We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
 \begin{IEEEproof}
@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be noisy symbols given by the signal model~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be given by the signal model~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
@@
 Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-If $g(\phi)$ is continuous at $\phi = (2k+1)\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
-\[
-C = \frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
+If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+\[
+\frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
 pA_1 + dA_2 & ? \\
@@
 
-Proofs of these theorems are given in subsections~\ref{sec:proof-almost-sure}~and~\ref{sec:proof-asympt-norm}.  We first discuss the assumptions made by these theorems.
+Proofs of these theorems are given in Sections~\ref{sec:proof-almost-sure}~and~\ref{sec:proof-asympt-norm}.  We first discuss the assumptions made by these theorems.
 
@@
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
-Let $h_2$ and $k_2$ be the functions as,
+Let,
 \[
@@
 
-\subsection{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
-
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function conditioned on minimisation with respect the the data symbols,
+\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
+
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function $SS$ conditioned on minimisation with respect the the data symbols,
  \begin{align*}
@@
 and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of the amplitude as a function of $\theta$,
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of the amplitude $\rho_0$ as a function of $\theta$,
 \begin{equation}\label{eq:hatrhoZ}
@@
 \end{equation}
-where $\Re(\cdot)$ denotes the real part.
-Substituting this expression into $SS(\rho, \theta)$ above gives the sum of squares function conditioned upon minimisation with respect to the amplitude $\rho$ and the data symbols,
+where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives the $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
 \[
@@
 %The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
-Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta}, we have
+Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
 \begin{align*}
@@
 \]
-Now we can write
-\[
-Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}
+Now,
+\[
+Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
 \]
@@
 \end{equation}
-and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle$.
+and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
 Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
@@
 \]
-Lemma~\ref{lem:convtoexpGlamL} shows that,
-\[
-\prob\{ \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L})
-\]
-where $\prob$ denotes probability.  Since $G(\lambda)$ is continuous and is uniquely maximised at $0$, then for any $\delta > 0$ there is a $\epsilon > 0$ such that if 
-\[
-\sabs{\hat{\lambda}_L} > \delta \qquad \text{then} \qquad \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon,
-\]
-and therefore for any $\delta > 0$ there is an $\epsilon > 0$ such that
-\[
-\prob\{\sabs{\hat{\lambda}_L} > \delta\} \leq \prob\{ \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L}).
-\]
-Now the sum 
-\[
-\sum_{L=1}^{\infty}\prob\{ \sabs{\hat{\lambda}_L} > \delta \} = \sum_{L=1}^{\infty}O(e^{-\epsilon^2 L})
-\]
-converges and consequently $\hat{\lambda}_L$ converges almost surely to zero by the Borel-Cantelli lemma.
-
-It remains to prove part two of this theorem regarding the convergence of the least squares amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
+As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
+Let $A$ be the subset of the sample space $\Omega$ upon which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
+Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
+Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
+Because $G$ is uniquely maximised at zero, then $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
+Part one of Theorem~\ref{thm:consistency} follows.
+
+It remains to prove part two of the theorem regarding the convergence of the least squares amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
 \begin{equation}\label{eq:rhoGLZ}
@@
 
-\begin{lemma}\label{lem:convtoexpGlamL} For any $\epsilon > 0$,
-\[
-\prob\{ \sabs{G(\hat{\lambda}_L) - G(0)} > \epsilon \} = O(e^{-\epsilon^2 L})
-\]
+\begin{lemma}\label{lem:convtoexpGlamL} 
+$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
 \end{lemma}
@@
 
-\subsection{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
+\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
 
@@
 This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at the maximiser $\hat{\lambda}_L$, so
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$, so
 \begin{align*}
@@
 \]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$.
-The constant real numbers $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
 Lemma~\ref{lem:empiricprocc} shows that
@@
 
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Define the function $T(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
+We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
 \begin{align*}
@@
 \end{align}
-We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\Phi_1$ we obtain
+We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\Phi_1$,
 \[
@@
 \end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ is circularly symmetric, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite (by assumption in Theorem~\ref{thm:consistency}).
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $\expect w_1$, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 Q(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
-&= \hat{\lambda}_L \left( \tfrac{\abs{P}}{L}  + \tfrac{\abs{D}}{L} H + o_P(1) \right).
-\end{align*}
-The proof follows since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.
+&= \hat{\lambda}_L \left( \tfrac{\abs{P}}{L}  + \tfrac{\abs{D}}{L} H + o_P(1) \right) \\
+&= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
+\end{align*}
+since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.
 \end{IEEEproof}
@@
 \begin{IEEEproof}
-Since $\hat{\lambda}_L \in [-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ it is only the behaviour of the function $k_2(\lambda)$ for $\lambda\in [-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ that is relevant.
-We will analyse $k_2(\lambda)$ for non negative $\lambda$, i.e., when $0 \leq \lambda < \tfrac{\pi}{M}$.
+Since $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $k_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
+We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
 An analagous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
@@
 \end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\Phi_1 - \tfrac{2\pi}{M}k$ we obtain 
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\Phi_1 - \tfrac{2\pi}{M}k$, 
 \begin{align*}
@@
 \sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&=  s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) + \zeta_{k+1}(\lambda)\big) 
+&= s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi k}{M}) + c_k(\Phi_1) + \zeta_{k+1}(\lambda)\big)
 \end{align*}
@@
 \[
-s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k).
-\]
-Putting,
-\[
-c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M})
-\]
-we can write
-\begin{align*}
-\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&= s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi k}{M}) + c_k(\Phi_1) + \zeta_{k+1}(\lambda)\big) \\
-&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1) \big).
-\end{align*}
-where, for convenience, we put $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.
-Observe that both $s_k$ and $c_k$ are continuous functions and that $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\psi_{k}) = 0$.
+s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
+\]
+and
+\[
+c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M}).
+\]
+Now,
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1}\\ 
+&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1) \big),
+\end{align*}
+where $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.  Both $s_k$ and $c_k$ are continuous functions and $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\psi_{k}) = 0$.
 
@@
 \end{align*}
-Define the function $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, and
+Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, and
 \[
@@
 \]
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\gamma(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
-$\gamma(\hat{\lambda}_L) = o_P(1)$, and therefore
+Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, it follows that $\gamma(\hat{\lambda}_L) = o_P(1)$, and so
 \[
@@
 \begin{IEEEproof}
-In this lemma, as Lemma~\ref{lem:k2conv}, we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
+In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
 An analgous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
-Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$, then,
-\begin{align*}
-W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) = \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \\
-&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi,
-\end{align*}
-and, since $v(\lambda, \phi) = \sum_{k\in\ints} v_k(\lambda, \phi)$,
-\begin{equation}\label{eq:Winfsum}
-W(\lambda) = \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi.
-\end{equation}
+Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
+\begin{align}
+W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
+&= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
+&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
+&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi. \label{eq:Winfsum}
+\end{align}
 Now, when $0 < k < M$, we have $0 < \psi_{k-1} < \psi_k < 2\pi$, so, using~\eqref{eq:vk},
@@
 \int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
-\end{align*}
-where, for each integer $k = 1, \dots, M-1$, the functions $a_k$ and $b_k$ are
+&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda d_k(\lambda),
+\end{align*}
+where, for $k = 1, \dots, M-1$,
 \begin{equation}\label{eq:ak}
@@
 \begin{equation}\label{eq:ck}
-c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
-\end{equation}
-Observe that $a_k(\lambda)$ and $b_k(\lambda)$ are functions that converge to zero as $\lambda$ goes to zero, whilst $c_k(\lambda)$ converges to 
-\[
-s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
+d_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
+\end{equation}
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $d_k(\lambda)$ converges to 
+\[
+s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}).
 \]
@@
 W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \left( C_k(\lambda)  + A_k(\lambda) + B_k(\lambda) \right).
+&= \sum_{k=0}^{M-1} \lambda d_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
+&= \lambda \left( D_k(\lambda)  + A_k(\lambda) + B_k(\lambda) \right).
 \end{align*}
@@
 \[
-C_k(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
-\]
-Observe both $A_k(\lambda)$ and $B_k(\lambda)$ converges to zero as as $\lambda$ goes to zero, and $C_k(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
-\[
-\gamma(\lambda) = A_k(\lambda) + B_k(\lambda) + C_k(\lambda) - H_1.
-\]
-Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$.  The proof follows.
+D_k(\lambda) = \sum_{k=0}^{M-1} d_k(\lambda).
+\]
+Both $A_k(\lambda)$ and $B_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and $D_k(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
+\[
+\gamma(\lambda) = A_k(\lambda) + B_k(\lambda) + D_k(\lambda) - H_1.
+\]
+Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$ as required.
 \end{IEEEproof}

### 2012-07-30 paper.tex

@@
 Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+%If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
+Then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
 \[
@@
 \begin{lemma}\label{lem:expRvlamphi}
-$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.
+$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
 \end{lemma}
@@
 \end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $d_k(\lambda)$ converges to 
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous, $d_k(\lambda)$ converges to 
 \[
@@
 \]
+as $\lambda$ goes to zero from above\footnote{We actually only require $g$ to be continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$, and the requirements of Theorem~\ref{thm:normality} could be weakened in this sense.
+Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.}.
+We are only interested in the limits from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
+The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
+
 Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
@@
 &= \sum_{k=0}^{M-1} \lambda d_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \left( D_k(\lambda)  + A_k(\lambda) + B_k(\lambda) \right).
+&= \lambda \left( D(\lambda)  + A(\lambda) + B(\lambda) \right),
 \end{align*}
@@
 \[
-A_k(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B_k(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
-\]
-\[
-D_k(\lambda) = \sum_{k=0}^{M-1} d_k(\lambda).
-\]
-Both $A_k(\lambda)$ and $B_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and $D_k(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
-\[
-\gamma(\lambda) = A_k(\lambda) + B_k(\lambda) + D_k(\lambda) - H_1.
+A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
+\]
+\[
+D(\lambda) = \sum_{k=0}^{M-1} d_k(\lambda).
+\]
+Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $D(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
+\[
+\gamma(\lambda) = A(\lambda) + B(\lambda) + D(\lambda) - H_1.
 \]

### 2012-07-31 other/confnoncoherent/paper.tex
Added tex.

--- 2012-07-30 paper.tex
@@
 %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal]{IEEEtran}
+\documentclass[conference]{IEEEtran}
 
@@
 %opening
-\title{Carrier phase and amplitude estimation of $M$-ary phase shift keying using pilots and data}
+\title{Noncoherent least squares estimators of carrier phase and amplitude}
 \author{Robby McKilliam, Andre Pollok, Bill Cowley}
@@
 \begin{abstract}
-We describe least squares estimators of carrier phase and amplitude from a noisy communications signal that contains both pilot signals, known to the receiver, and data signals, unknown to the receiver.
+We consider least squares estimators of carrier phase and amplitude from a noisy communications signal.
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-PSK.
-We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of the carrier phase is strongly consistent and asymptotically normaly distributed.
+Under reasonably mild conditions on the distribution of the noise, the least squares estimator of carrier phase is strongly consistent and asymptotically normaly distributed.
 However, the amplitude estimator is not consistent, and we show that it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
-Our theoretical results can also be applied to the case where no pilot symbols exist (i.e.
-noncoherent detection).
 The results of Monte Carlo simulations are provided and these corroborate the theoretical results.
@@
 where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation.
-We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.
-So,
-\[
-s_i = \begin{cases}
-p_i & i \in P \\
-d_i & i \in D,
-\end{cases}
-\]
-where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.
-The sets $P$ and $D$ are disjoint, i.e.
-$P \cap D = \emptyset$, and $L = \abs{P \cup D}$ is the total number of symbols transmitted.
 
-We assume that time offset estimation (symbol timing) and matched filtering have already been performed.  The recieved symbols are then given by,
+We assume that time offset estimation (symbol timing) and matched filtering have already been performed at the reciever.  The recieved symbols take the form,
 \begin{equation}\label{eq:sigmod}
-y_i = a_0 s_i + w_i, \qquad i \in P \cup D,
+y_i = a_0 s_i + w_i, \qquad i = 1, \dots, L,
 \end{equation}
 where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
-Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
-Complicating matters is that the data symbols $d_i$ are not known to the reciever and must also be estimated.
+Our aim is to estimate $a_0$ from the noisy symbols $y_1, \dots, y_L$.
+Complicating matters is that the transmitted symbols $s_i$ are not known to the reciever and must also be estimated.
 Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
 A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
-\begin{align*}
-SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \\
-&= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in P \cup D} \abs{ y_i - a d_i }^2,
-\end{align*}
+\[
+SS(a, \{s_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 
+\]
 where $\abs{x}$ denotes the magnitude of the complex number $x$.
@@
 However, as we will show, the estimator works well under less stringent assumptions.
-%It is the least squares estimator that we primarily study in this paper.
+Mackenthun~\cite{Mackenthun1994} described an algorithm to compute the least squares estimator that requires only $O(L \log L)$ arithmetic operations.
+Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
 
-The existing literature mostly considers what is called \emph{noncoherent detection} where no pilots symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
-In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
-A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+% An alternative approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
 The idea is the `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
 A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and $A$ is the average
-\[
-A = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
-\]
-Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included, so as to make the estimator coherent.
-Techniques are available but they are somewhat adhoc~\cite{Cowley_ref_sym_carr_1998}.
-This problem does not occur with the least square estimator.
+% \[
+% A = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
+% \]
+% Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
 
-An important paper is by Mackenthun~\cite{Mackenthun1994} who described an algorithm to compute the least squares estimator requiring only $O(L \log L)$ arithmetic operations.
-Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
-Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-
-In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
+In the literature it has been common to assume that the symbols $\{s_i, i \in D\}$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
 The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
 Whilst estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite approach here.
-Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
+Our aim is to estimate $a_0$, and we treat the unknown symbols as nuisance parameters.
 This is motivated by the fact that in our motivating application (narrowband satellite communications) and in many modern communication systems the data symbols are \emph{coded}.
@@
 Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
-The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
+The function $\fracpart{\cdot}$ denotes its argument taken `modulo $\tfrac{2\pi}{M}$' into the interval $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$.
+It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
 
@@
 
-\section{Mackentun's algorithm}\label{sec:least-squar-estim}
+% \section{The least squares estimator}\label{sec:least-squar-estim}
 
-In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to include the pilot symbols.
-For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
-In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.
-This complexity arises from the need to sort a list of $\abs{D}$ elements.
-%In Section~\ref{sec:line-time-algor} we will show that a full sort is not neccesary, and that the least squares estimator can be implemented in $O(L)$ operations.
-
-Define the sum of squares function
-\begin{align}
-SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
-&= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*, \label{eq:SS}
-\end{align}
-where $*$ denotes the complex congugate.
-Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find that the least squares estimator of $a_0$, satisfies
-\begin{equation}\label{eq:hata}
-\hat{a} = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y
-\end{equation}
-where $L = \abs{P \cup D}$ is the total number of symbols transmitted, and to simplify our notation we have put 
-\[
-Y = \sum_{i \in P \cup D} y_i s_i^* = \sum_{i \in P } y_i p_i^* + \sum_{i \in D } y_i d_i^*.
-\]  
-Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to surpress the argument $(\{ d_i, i \in D\})$ for notational clarity.
-Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain the $SS$ conditioned on minimisation with respect to $a$,
-\begin{equation}\label{eq:SSdatasymbols}
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y}^2,
-\end{equation}
-where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.
-Observe that given a candidate value for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-It turns out that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
-%When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
-
-To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a nonnegative real.  Now,
-\begin{align}
-SS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
-&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:SSallparams}
-\end{align}
-For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
-\begin{equation}\label{eq:hatdfinxtheta}
-\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
-\end{equation}
-where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
-A word of caution, the notation $\round{\cdot}$ is often used to denote rounding to the nearest \emph{integer}.
-This is not the case here.
-If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer then,
-\[
-\round{x} = \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
-\] 
-Note that $\hat{d}_i(\theta)$ does not depend on $\rho$.
-As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
-With this in mind,
-\[
-\hat{u}_i(\theta) = \round{\angle{y_i} - \theta }
-\]
-which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $2\pi$.
-
-We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Let $b_i = \hat{d}_i(0)$ and let 
-\[
-z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}.
-\]
-Then,
-\begin{equation}\label{eq:uicombos}
-\hat{d}_i(\theta) = 
-\begin{cases}
-b_i, &  0 \leq \theta < z_i + \frac{\pi}{M} \\
-b_i e^{-j2\pi/M}, & z_i + \frac{\pi}{M} \leq \theta < z_i + \frac{3\pi}{M} \\ 
-\vdots & \\
-b_i e^{-j2\pi k /M}, & z_i + \frac{\pi(2k - 1)}{M} \leq \theta < z_i + \frac{\pi(2k + 1)}{M}  \\ 
-\vdots & \\
-b_i e^{-j2\pi}, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
-\end{cases}
-\end{equation}
-
-Let 
-\[
-f(\theta) = \{ \hat{d}_i(\theta), i \in D \}
-\]
-be a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
-Observe that $f(\theta)$ is piecewise continuous.
-The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains contant are determined by the values of $\{z_i, i \in D\}$.
-Let
-\[
-S = \{ f(\theta) \mid \theta \in [0, 2 \pi) \}
-\]
-be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$.
-If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
-
-The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indices in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
-\begin{equation}\label{eq:sigmasortind}
-z_{\sigma(i)} \leq z_{\sigma(k)}
-\end{equation}
-whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}-1\}$.
-It is convenient to define the indices into $\sigma$ to be taken modulo $\abs{D}$, that is, if $m$ is an integer not from $\{0, 1, \dots, \abs{D}-1\}$ then we define $\sigma(m) = \sigma(k)$ where $k \equiv m \mod \abs{D}$ and $k \in  \{0, 1, \dots, \abs{D}-1\}$.
-The first sequence in $S$ is
-\[
-f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
-\]  
-The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/M}$.
-Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.
-Using this notation,
-\[
-f_1 = f_0 e_{\sigma(0)}.
-\] 
-The next sequence in $S$ is correspondingly 
-\[
-f_2 = f_0 e_{\sigma(0)} e_{\sigma(1)} = f_1 e_{\sigma(1)},
-\]
-and the $k$th sequence is
-\begin{equation}\label{eq:fkrec}
-f_{k+1} = f_{k} e_{\sigma(k)}.
-\end{equation}
-In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively enumerated.
-
-We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,(M+1)\abs{D}-1\}$.
-Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
-So, this na\"{\i}ve approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
-Mackenthun~\cite{Mackenthun1994} showed how $SS(f_k)$ can be computed recursively.
-
-Let,
-\begin{equation}\label{eq:SSfk}
-SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
-\end{equation}
-where, 
-\begin{align*}
-Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
-&= B + \sum_{i \in D}g_{ki},
-\end{align*}
-where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
-Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
-\[
-g_{k+1} = g_{k} e_{\sigma(k)}^*,
-\]
-where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/M}$.  Now,
-\[
-Y_0 = B + \sum_{i \in D} g_{0i}
-\] 
-can be computed in $O(L)$ operations, and
-\begin{align*}
-Y_1 &= B + \sum_{i \in D} g_{1i} \\
-&= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
-&= Y_0 + \eta g_{0\sigma(0)},
-\end{align*}
-where $\eta = (e^{j2\pi/M} - 1)$.  In general,
-\[
-Y_{k+1} = Y_k + \eta g_{k\sigma(k)}.
-\]
-So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a constant number of operations.
-Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.
-Let $\hat{k} = \arg\min SS(f_k)$.
-The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata},
-\begin{equation}\label{eq:ahatYhat}
-\hat{a} = \frac{1}{L} Y_{\hat{k}}.
-\end{equation}
-Psuedocode is given in Algorithm~\ref{alg:loglinear}.
-Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
-The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
-This requires $O(L \log L)$ operations.
-The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.
-The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
-%In the next sections we will show how to the sortinces function can be avoided.
-This leads to an algorithm that requires only $O(L)$ operations.
-
-\begin{algorithm}[t] \label{alg:loglinear}
-\SetAlCapFnt{\small}
-\SetAlTitleFnt{}
-\caption{Mackenthun's algorithm with pilot symbols}
-\DontPrintSemicolon
-\KwIn{$\{y_i, i \in P \cup D \}$}
-\For{$i \in D$ \nllabel{alg_loop_setup}}{
-$\phi = \angle{y_i}$ \;
-$u = \round{\phi} $ \;
-$z_i = \phi -  u $ \;
-$g_i = y_i e^{-j u}$ \;
-}
-$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
-$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
-$\hat{a} = \frac{1}{L} Y$ \;
-$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
-$\eta = e^{j2\pi/M} - 1$ \;
-\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
-$Y = Y + \eta g_{\sigma(k)}$ \;
-$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
-$Q = \frac{1}{L}\abs{Y}^2$\;
-\If{$Q > \hat{Q}$}{
- 	$\hat{Q} = Q$ \;
- 	$\hat{a} =  \frac{1}{L} Y$ \;
- }
-}
-\Return{$\hat{a}$ \nllabel{alg_return}}
-\end{algorithm}
+% Define the sum of squares function
+% \begin{align}
+% SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
+% &= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*, \label{eq:SS}
+% \end{align}
+% where $*$ denotes the complex congugate.
 
@@
 That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
-If the probability density function of $Z$ is $f_Z(z)$, then the joint probability density function (pdf) of $\Theta$ and $Z$ is
+If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf (pdf) of $\Theta$ and $Z$ is
 \[
@@
 \]
-where $\Im(\cdot)$ denotes the imaginary part.  The next lemma will be useful for the analysis of the least squares estimator.
+where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
 
-\begin{lemma}\label{lem:h1minedcircsym}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
+% \begin{lemma}\label{lem:h1minedcircsym}
+% Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
 Let
-\[
-h_1(x) = \expect R \cos(x + \Phi).
-\]
-Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
-\end{lemma}
+% \[
+% h_1(x) = \expect R \cos(x + \Phi).
+% \]
+% Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
+% \end{lemma}
+
 % Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
@@
 % We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
+
+% \begin{IEEEproof}
+% BLERG
+% \end{IEEEproof}
 
@@
 
-In this section we decribe the asymptotic properties of the least squares estimator.
-In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
-\[
-\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right),
-\]
-where $\operatorname{round}(\cdot)$ takes its argument to the nearest integer.
-The direction of rounding for half-integers is not important so long as it is consistent.
-We have chosen to round up half-integers here.
-Similarly we use $\sfracpart{x}$ to denote $x$ taken `modulo $\tfrac{2\pi}{M}$' into the interval $\left[-\tfrac{\pi}{M}, \tfrac{\pi}{M}\right)$, that is
-\[
-\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right) = x - \round{x}.
-\]
-The next two theorems decribe the asymptotic properties of the least squares estimator.
+The next two theorems decribe the asymptotic properties of the least squares estimator.  We ommit the proofs due to space constraints.  Full proofs will be provided in an upcomming paper.
 
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be given by the signal model~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
-Put $L = \abs{P \cup D}$ and define real numbers $p$ and $d$ as the limits
-\[
-\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
-\] 
 Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
@@
 \[
-G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
+G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
 \]
-\[
-h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
-\]
-If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
+If $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
 \begin{enumerate}
@@
 Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+%If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
+Then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
 \[
-\frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
-pA_1 + dA_2 & ? \\
+\frac{1}{H^2} \left( \begin{array}{cc} 
+A & ? \\
 ? & ?
@@
 \[
-H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
+H = G(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
 \]
+and
 \[
-A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1}.
+A = \expect R_1^2\sin^2\fracpart{\Phi_1}.
 \]
@@
 
-Proofs of these theorems are given in Sections~\ref{sec:proof-almost-sure}~and~\ref{sec:proof-asympt-norm}.  We first discuss the assumptions made by these theorems.
+We now make some discussion regarding the assumptions made by these theorems.
 
@@
 \[
-h_2(x) = \expect R \cos\sfracpart{x + \Phi}, \;\;\; k_2(x) =  \expect R \sin\sfracpart{ x + \Phi}.
+G(x) = \expect R \cos\sfracpart{x + \Phi}.
 \]
-Then $h_2(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$ and $k_2(0) = 0$.
+Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
 \end{lemma}
@@
 
-\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
-
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function $SS$ conditioned on minimisation with respect the the data symbols,
- \begin{align*}
-SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
-&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
-\end{align*}
-where
-\[
-Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
-\]
-and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of the amplitude $\rho_0$ as a function of $\theta$,
-\begin{equation}\label{eq:hatrhoZ}
-\hat{\rho} = \frac{Z(\theta) + Z(\theta)^*}{2L} = \frac{1}{L}\Re(Z(\theta)),
-\end{equation}
-where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives the $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
-\[
-SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
-\]
-
-We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$ or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
-Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
-\begin{align*}
-y_i &= a_0 s_i + w_i \\
-&= \rho_0 s_i e^{j \theta_0}\left( 1 + \frac{1}{\rho_0} w_i s_i^* e^{-j \theta_0} \right) \\
-&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
-\end{align*}
-%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
-Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
-\begin{align*}
-\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
-&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
-&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
-&= \round{ \lambda + \Phi_i + \angle{s_i} },
-\end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, then, when $i \in D$,
-\begin{align*}
- y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
-&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
-&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}}
-\end{align*}
-since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
-\[
-y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
-\]
-Now,
-\[
-Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
-\]
-and
-\[
-\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
-\] 
-Let 
-\begin{equation}\label{eq:GLdefn}
-G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
-\end{equation}
-and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
-Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
-We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.
-
-Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
-\[
-\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
-\]
-and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
-\[
-\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
-\]
-As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ upon which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
-Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
-Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $G$ is uniquely maximised at zero, then $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
-Part one of Theorem~\ref{thm:consistency} follows.
-
-It remains to prove part two of the theorem regarding the convergence of the least squares amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
-\begin{equation}\label{eq:rhoGLZ}
-\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
-\end{equation}  
-Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
-
-\begin{lemma}\label{lem:convtoexpGlamL} 
-$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-BLERG
-% Lemma~\ref{lem:Zconunif} shows BLERG that 
-% \[
-% 0 \leq G(\hat{\lambda}_L) - G(0),
-% \]
-% and since $\hat{\lambda}_L$ is the maximiser of $G_L$ then
-% \[ 
-% 0 \leq G_L(0) - G_L(\hat{\lambda}_L).
-% \]
-% Adding both inequalities and taking appropriate absolute values gives
-% \begin{align*}
-% 0 &\leq G(\hat{\lambda}_L) - G(0) \\ 
-% & \leq G(\hat{\lambda}_L) - G(0) + G_L(0) - G_L(\hat{\lambda}_L) \\
-% & \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)|,
-% \end{align*}
-\end{IEEEproof}
-
-% \begin{lemma}\label{lem:Zconunif}
-% \[
-% \prob\left\{\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} > \epsilon \right\} = O(e^{-\epsilon^2 L}).
-% \]
-% \end{lemma}
-% \begin{IEEEproof}
-% \end{IEEEproof}
-
-\begin{lemma}\label{lem:GLtoG0}
-$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
-\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
-
-We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
-Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
-Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
-The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$, so
-\begin{align*}
-G_L^\prime(\hat{\lambda}_L) &= 0 \\
-&= \frac{1}{L} \sum_{i \in P} R_i \sin(\hat{\lambda}_L + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\hat{\lambda}_L + \Phi_i}.
-\end{align*}
-Define the function $Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0)$ and write
-\begin{align*}
-0 &= G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) + Q(\hat{\lambda}_L) \\
-&= \sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) + \sqrt{L}Q(\hat{\lambda}_L).
-\end{align*}
-Lemma~\ref{lem:Qconv} shows that
-\[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
-Lemma~\ref{lem:empiricprocc} shows that
-\[
-\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
-\]
-It follows from the three equations above that,
-\[
-0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
-\]
-and rearranging gives,
-\[
-\sqrt{L} \hat{\lambda}_L = o_P(1) + \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
-\]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
-It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
-\[
-\frac{pA_1 + dA_2}{(p + Hd)^2}.
-\]
-
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
-\begin{align*}
-\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L) + T(\hat{\lambda}_L) - G(0) \big).
-\end{align*}
-Lemma~\ref{lem:empiricprocforrho} shows that 
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - T(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
-\]
-It follows that
-\[
-\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
-\]
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance
-\[
-?
-\]
-and Lemma~\ref{lem:XLcovlambdaL} shows that the covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ goes to zero as $L\rightarrow\infty$.
-So, the distribution of $\sqrt{L}m_L$ converges to the normal with zero mean and variance
-\[
-\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = BLERG.
-\]
-The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondingly
-\[
-?
-\]
-It remains to prove the lemmas that we have used.
-
-\begin{lemma}~\label{lem:diffatlambdaL}
-The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
-\[
-G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
-\]
-\end{lemma}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
-% \begin{lemma}
-% \[
-% \sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{1}{2M} - \frac{1}{2LM ?}
-% \]
-% for all $i \in D$.
-% \end{lemma}
-% \begin{IEEEproof}
-% Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
-The proof now proceeds by contradiction.
-Assume that $\sfracpart{\hat{\lambda}_L + \Phi_i} \leq \frac{1}{2M} - \frac{1}{2LM ?}$ for some $k$.
-Put $r_i = \hat{d}_i e_k$, then
-% \begin{align*}
-% Y(\{ r_i, i \in D\}) = \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) = \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
-% \end{align*}
-% where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
-% \begin{align*}
-% SS&(\{r_i, i \in D\}) \\
-% &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
-% &= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
-% &= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
-% &= SS(\{\hat{d}_i, i \in D\}) - R
-% \end{align*}
-% where 
-% \[
-% R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
-% \]
-% We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-
-% To see that $R$ is positive write
-%  \begin{align*}
-% R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
-%  &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
-%  \end{align*}
-%  so it is sufficient to show that
-%  \[
-%  \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
-%  \]
-%  Unfinished, but this proof will work
-% \end{IEEEproof}
-
-\begin{lemma}\label{lem:Qconv}
-$\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)$
-\end{lemma}
-\begin{IEEEproof}
-We have
-\[
-Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0) = \tfrac{\abs{P}}{L} k_1(\lambda) + \tfrac{\abs{D}}{L} k_2(\lambda)
-\]
-where
-\begin{align}
-k_1(\lambda) &= \expect R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
-k_2(\lambda) &= \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
-\end{align}
-We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\Phi_1$,
-\[
-\sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
-\]
-where $\zeta(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Then,
-\begin{align*}
-k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
-&= \lambda \big( \expect R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \expect R_1 \big) \\
-&= \lambda \big( 1 \; + \; \zeta(\lambda) K \big),
-\end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $\expect w_1$, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
-$\zeta(\hat{\lambda}_L) = o_P(1)$.
-So,
-\[
-k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) K \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big).
-\]
-Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
-\begin{align*}
-Q(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
-&= \hat{\lambda}_L \left( \tfrac{\abs{P}}{L}  + \tfrac{\abs{D}}{L} H + o_P(1) \right) \\
-&= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
-\end{align*}
-since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:k2conv}
-$k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
-\end{lemma}
-\begin{IEEEproof}
-Since $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $k_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
-We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
-An analagous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
-To keep our notation clean put
-\[
-\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
-\]
-For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
-\[
-\fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
-\]
-Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
-\]
-whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
-\]
-So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\Phi_1 - \tfrac{2\pi}{M}k$, 
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
-&= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
-\end{align*}
-where $\zeta_k(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}$,
-\begin{align*}
-\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&= s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi k}{M}) + c_k(\Phi_1) + \zeta_{k+1}(\lambda)\big)
-\end{align*}
-where
-\[
-s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
-\]
-and
-\[
-c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M}).
-\]
-Now,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1}\\ 
-&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1) \big),
-\end{align*}
-where $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.  Both $s_k$ and $c_k$ are continuous functions and $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\psi_{k}) = 0$.
-
-Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
-\]
-whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
-\]
-Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, then, for all $\Phi_1 \in \reals$,
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1) 
-\]
-where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
-\begin{equation}\label{eq:vk}
-v_k(\lambda,\Phi_1) = \begin{cases}
-s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
-\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
-0 & \text{otherwise}.
-\end{cases}
-\end{equation}
-Now
-\begin{align*}
-k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
-&= \expect R_1\big( \lambda\cos\fracpart{\Phi_1} + v(\lambda, \Phi_1)  \big) \\
-&= \lambda h_2(0) + \expect R_1 v(\lambda, \Phi_1).
-\end{align*}
-Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, and
-\[
-H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0).
-\]
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, it follows that $\gamma(\hat{\lambda}_L) = o_P(1)$, and so
-\[
-k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( h_2(0) + H_1 + o_P(1) ) = \hat{\lambda}_L ( H + o_P(1) )
-\]
-as required.  It remains to prove Lemma~\ref{lem:expRvlamphi}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:expRvlamphi}
-$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero.
-\end{lemma}
-\begin{IEEEproof}
-In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
-An analgous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
-Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
-\begin{align}
-W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
-&= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
-&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
-&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi. \label{eq:Winfsum}
-\end{align}
-Now, when $0 < k < M$, we have $0 < \psi_{k-1} < \psi_k < 2\pi$, so, using~\eqref{eq:vk},
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda d_k(\lambda),
-\end{align*}
-where, for $k = 1, \dots, M-1$,
-\begin{equation}\label{eq:ak}
-a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
-\end{equation}
-\[
-b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
-\]
-and
-\begin{equation}\label{eq:ck}
-d_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
-\end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $d_k(\lambda)$ converges to 
-\[
-s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}).
-\]
-Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
-\[
-\int_0^{2\pi} v_0(\lambda, \phi) g(\phi) d\phi = \lambda a_0(\lambda) + \lambda b_0(\lambda) +  \lambda c_0(\lambda),
-\]
-where $a_0(\lambda)$ and $c_0(\lambda)$ are defined as in~\eqref{eq:ak}~and~\eqref{eq:ck}, and
-\[
-b_0(\lambda) = \zeta_0(\lambda) \int_{0}^{\psi_k-\lambda} g(\phi) d\phi.
-\]
-When $k = M$, we have $0 < \psi_{M-1} < 2\pi < \psi_{M} - \lambda$, so that
-\[
-\int_0^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi = \lambda b_{M}(\lambda)
-\]
-where 
-\[
-b_{M}(\lambda) = \zeta_{M}(\lambda) \int_{\psi_{M-1}}^{2\pi}g(\phi) d\phi.
-\]
-Otherwise, when $k < 0$ or $k > M$, 
-\[
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0.
-\]
-So the infinite sum in~\eqref{eq:Winfsum} can be written as the finite sum,
-\begin{align*}
-W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \sum_{k=0}^{M-1} \lambda d_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \left( D_k(\lambda)  + A_k(\lambda) + B_k(\lambda) \right).
-\end{align*}
-where 
-\[
-A_k(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B_k(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
-\]
-\[
-D_k(\lambda) = \sum_{k=0}^{M-1} d_k(\lambda).
-\]
-Both $A_k(\lambda)$ and $B_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and $D_k(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
-\[
-\gamma(\lambda) = A_k(\lambda) + B_k(\lambda) + D_k(\lambda) - H_1.
-\]
-Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$ as required.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:empiricprocc} BLERG
-\[
-\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
-\]
-\end{lemma}
-\begin{IEEEproof}
-
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:convdistGLdash}
-The distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $A_1 + A_2$.
-\end{lemma}
-\begin{IEEEproof}
-Observe that 
-\[
-\sqrt{L} G_L^\prime(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
-\]
-It follows from the standard central limit that the distribution of
-\[
-\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{P}}{L}} \frac{1}{\sqrt{\abs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
-\]
-converges to the normal with mean $k_1(0) = 0$ and variance 
-\[
-p \expect R_1^2 \sin^2(\Phi_1).
-\]
-Similarly, the distribution of 
-\[
-\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{D}}{L}} \frac{1}{\sqrt{\abs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
-\]
-converges to the normal with mean $k_2(0) = 0$ and variance
-\[
-d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:empiricprocforrho} BLERG
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} \big( G_L^\prime(0) - H_L(0) \big).
-\]
-\end{lemma}
-\begin{IEEEproof}
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:HLtoG} BLERG
-\[
-\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \hat{\lambda}_L \big( ? + o_P(1) \big).
-\]
-\end{lemma}
-\begin{IEEEproof}
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:XL} 
-The distribution of $X_L$ converges to the normal with zero mean and covariance
-\[
-?
-\]
-as $L \rightarrow\infty$.
-\end{lemma}
-\begin{IEEEproof}
-
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:XLcovlambdaL} 
-The covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ converges to zero as $L\rightarrow\infty$.
-\end{lemma}
-\begin{IEEEproof}
-
-\end{IEEEproof}
-
 \section{The Gaussian noise case}
 
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.
-The joint density function of the real and imaginary parts is then
+Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint pdf of the real and imaginary parts is then
 \[
@@
 \]
-The value of $A_1$ and $A_2$ can be computed by numerical integration.
+The value of $A$ can be computed by numerical integration.
 
 \section{Simulations}\label{sec:simulations}
-
-%\subsection{The coherent setting}
-
-Our first set of simulations considers the case where there are a number of pilot symbols.
-
-%BPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-%		\caption{Amplitude error for BPSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error for BPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure*}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure*}
-
-%QPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-%		\caption{Amplitude error for QPSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error for QPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM4-3.mps}
-% 		\caption{complex amplitude error for QPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%8PSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-%		\caption{Amplitude error for 8PSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error for 8PSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%\subsection{The noncoherent setting}
-
-%BPSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-% 		\caption{Amplitude error for BPSK}
-% 		\label{fig:plotamp}
- %\end{figure}
 
@@
 
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%QPSK plots
- %\begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-% 		\caption{Amplitude error for QPSK}
-% 		\label{fig:plotamp}
-% \end{figure}
-
 \begin{figure}[tp]
@@
 \end{figure}
-
- % \begin{figure}[tp]
- % 	\centering
- %         \includegraphics{code/data/plotM4-3.mps}
- % 		\caption{complex amplitude error for QPSK}
- % 		\label{fig:plotcamp}
- % \end{figure}
-
-%8PSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-% 		\caption{Amplitude error for 8PSK}
-% 		\label{fig:plotamp}
-% \end{figure}
 
@@
 
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
+\small
 \bibliography{bib}

### 2012-07-31 paper.tex
minor edits.
Added to paper.  Putting in the amplitude extimator CLT.  I think I can see the result.  Might be time to simulate.
Added to simulation script so that unbiased clt for the amplitude can be tested.

@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be given by the signal model~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
-Put $L = \abs{P \cup D}$ and define real numbers $p$ and $d$ as the limits
+Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
 \[
@@
 \[
-\frac{1}{(p + H d)^2} \left( \begin{array}{cc} 
-pA_1 + dA_2 & ? \\
-? & ?
+\left( \begin{array}{cc} 
+\frac{pA_1 + dA_2}{(p + H d)^2} & 0 \\
+0 & pB_1 + dB_2
 \end{array} \right)
@@
 \[
-A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1}.
+A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
+\]
+\[
+B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
 \]
@@
 
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain the sum of squares function $SS$ conditioned on minimisation with respect the the data symbols,
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect the the data symbols,
  \begin{align*}
@@
 \end{equation}
-where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives the $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
+where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
 \[
@@
 
-We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$ or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
+We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
 Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
@@
 \end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, then, when $i \in D$,
+where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that when $i \in D$,
 \begin{align*}
@@
 Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $G$ is uniquely maximised at zero, then $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
+Because $G$ is uniquely maximised at zero, it follows that $G(\widehat{\lambda}_L) \rightarrow G(0)$ only if $\widehat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
+So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
 Part one of Theorem~\ref{thm:consistency} follows.
 
-It remains to prove part two of the theorem regarding the convergence of the least squares amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
+It remains to prove part two of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
 \begin{equation}\label{eq:rhoGLZ}
@@
 \[
-?
+ p B_1 + d B_2 
+\]
+where
+\[
+B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0)
 \]
@@
 \]
-It follows from the standard central limit that the distribution of
-\[
-\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{P}}{L}} \frac{1}{\sqrt{\abs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
+From the standard central limit the distribution of
+\[
+\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) = \sqrt{\frac{\abs{P}}{L}} \frac{1}{\sqrt{\abs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
 \]
@@
 \[
-p \expect R_1^2 \sin^2(\Phi_1).
+A_1 = p \expect R_1^2 \sin^2(\Phi_1).
 \]
@@
 \[
-\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin(\lambda + \Phi_i) = \sqrt{\frac{\abs{D}}{L}} \frac{1}{\sqrt{\abs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
+\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} = \sqrt{\frac{\abs{D}}{L}} \frac{1}{\sqrt{\abs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
 \]
@@
 \[
-d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
+A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
 \]

### 2012-07-31 code/montecarlo.scala

@@ -45,9 +45,11 @@ for( L <- Ls; M <- Ms ) {
 	//generate some PSK symbols 
 	val rand = new scala.util.Random
 	val s = (1 to L).map(m => new PolarComplex(1, 2*scala.math.Pi*rand.nextInt(M)/M)) 
-	val est = estf(s) //construct an estimator	  
+	val est = estf(s) //construct an estimator (the estimator gets all the symbols, but only makes use of the pilots!)	  
+        val p = P.length.toDouble/L
+        val G0 = noise.clt(M,p).G(0) //value of G0 for unbiasing the amplitude estimator
 
-	var msec = 0.0; var msea = 0.0; var msep = 0.0;
+	var msec = 0.0; var msea = 0.0; var msep = 0.0; var msepunb = 0.0;
 	for( itr <- 1 to iters ) {
 	  //generate a recieved signal
 	  val y = s.map(si => a0*si + noise.noise )
@@ -55,10 +57,11 @@ for( L <- Ls; M <- Ms ) {
 	  val (ae, pe) = est.error(ahat, a0) //compute the error
 	  msep += pe
 	  msea += ae
-	  msec += (ahat - a0).mag2 //compute the mean square error 
+	  msec += (ahat - a0).mag2 //compute the mean square error
+          msepunb += (ahat.magnitude - G0)*(ahat.magnitude - G0) //rho0 is assumed equal to 1 
 	}
 	print(".")		      
-	(msea/iters, msep/iters, msec/iters) //last thing is what gets returned
+	(msea/iters, msep/iters, msec/iters, msepunb/iters) //last thing is what gets returned
       }.toList
 
       val estruntime = (new java.util.Date).getTime - eststarttime
@@ -67,13 +70,15 @@ for( L <- Ls; M <- Ms ) {
       val filea = new java.io.FileWriter("data/" + estname + "a")
       val filep = new java.io.FileWriter("data/" + estname + "p")
       val filec = new java.io.FileWriter("data/" + estname + "c")
+      val filepunb = new java.io.FileWriter("data/" + estname + "punb")
       (mselist, SNRdBs).zipped.foreach{ (mse, snr) =>
-	val (ma,mp,mc) = mse
+	val (ma,mp,mc,mu) = mse
 				       filea.write(snr.toString.replace('E', 'e') + "\t" + ma.toString.replace('E', 'e')  + "\n") 
 				       filep.write(snr.toString.replace('E', 'e') + "\t" + mp.toString.replace('E', 'e')  + "\n") 
[…]

### 2012-08-01 paper.tex
Added a bit to paper.

@@
 Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $G$ is uniquely maximised at zero, it follows that $G(\widehat{\lambda}_L) \rightarrow G(0)$ only if $\widehat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
+Because $G$ is uniquely maximised at zero, it follows that $G(\hat{\lambda}_L) \rightarrow G(0)$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
 So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
@@
 \end{align*}
-Define the function $Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0)$ and write
+Let $Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0)$ and write
 \begin{align*}
@@
 \[
-\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
+%\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = o_P(1).
 \]
@@
 \[
-\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
+%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L} m_L =  X_L + o_P(1). 
 \]
@@
 \]
-where
-\[
-B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0)
-\]
-and Lemma~\ref{lem:XLcovlambdaL} shows that the covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ goes to zero as $L\rightarrow\infty$.
-So, the distribution of $\sqrt{L}m_L$ converges to the normal with zero mean and variance
-\[
-\var X_L + \var \sqrt{L} \hat{\lambda}_L ?^2 = ?.
-\]
-The covariance between $\sqrt{L}m_L$ and $\sqrt{L}\hat{\lambda}_L$ is correspondingly
-\[
-?
-\]
-It remains to prove the lemmas that we have used.
+where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  As $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$.  Thus,
+\[
+\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
+\]
+as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.
 
@@
 \end{align}
-We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\Phi_1$,
+We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\lambda = 0$,
 \[
@@
 \end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $\expect w_1$, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ has zero mean, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 \begin{IEEEproof}
-Since $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $k_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
+As $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $k_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
 We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.

### 2012-08-01 paper.tex
Added new data.  Looks like I have all the results now.  Just need to start filling in all the Lemmas!

@@
 \end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\Phi_1 - \tfrac{2\pi}{M}k$, 
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
 \begin{align*}
@@
 \end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}$,
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
 \begin{align*}
 \sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&= s_k(\Phi_1) + \lambda\big(\cos(\Phi_1 - \tfrac{2\pi k}{M}) + c_k(\Phi_1) + \zeta_{k+1}(\lambda)\big)
+&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
 \end{align*}
@@
 \[
-c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M}).
-\]
-Now,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1}\\ 
-&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1) \big),
-\end{align*}
-where $\eta_k(\lambda, \Phi_1) = c_k(\Phi_1) + \zeta_{k+1}(\lambda)$.  Both $s_k$ and $c_k$ are continuous functions and $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $c_k(\psi_{k}) = 0$.
+\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_{k+1}(\lambda).
+\]
+Observe that $s_k$ is continuous and $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$.
 
@@
 \]
-Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, then, for all $\Phi_1 \in \reals$,
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1) 
+Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
 \]
@@
 \end{equation}
-Now we can write,
+Now,
 \begin{align*}
@@
 \end{align*}
-Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a function that goes to zero as $\lambda$ goes to zero, and
-\[
-H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0).
-\]
+Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a continuous function with $\gamma(0) = 0$, and
+\[
+H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0),
+\]
+where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, it follows that $\gamma(\hat{\lambda}_L) = o_P(1)$, and so
@@
 An analgous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
-Recalling that $f(r,\phi)$ is the joint probability density function of $R_1$ and $\Phi_1$,
+Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$,
 \begin{align}
@@
 \int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda d_k(\lambda),
+&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
 \end{align*}
@@
 \begin{equation}\label{eq:ck}
-d_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
-\end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous, $d_k(\lambda)$ converges to 
-\[
-s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}).
+c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
+\end{equation}
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous, $c_k(\lambda)$ converges to 
+\[
+s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
 \]
@@
 Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.}.
-We are only interested in the limits from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
+We are only interested in the limit from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
 The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
@@
 \]
-Both $b_0(\lambda)$ and $b_M(\lambda)$ are functions that converge to zero as $\lambda$ approaches zero from above.
+Both $b_0(\lambda)$ and $b_M(\lambda)$ are functions that converge to zero as $\lambda$ converges to zero.
 
@@
 W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \sum_{k=0}^{M-1} \lambda d_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \left( D(\lambda)  + A(\lambda) + B(\lambda) \right),
+&= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
+&= \lambda \left( C(\lambda)  + A(\lambda) + B(\lambda) \right),
 \end{align*}
@@
 \[
-D(\lambda) = \sum_{k=0}^{M-1} d_k(\lambda).
-\]
-Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $D(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
+C(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
+\]
+Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
 \[
@@
 \begin{lemma}\label{lem:convdistGLdash}
-The distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $A_1 + A_2$.
+The distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$.
 \end{lemma}
@@
 \]
-converges to the normal with mean $k_1(0) = 0$ and variance 
-\[
-A_1 = p \expect R_1^2 \sin^2(\Phi_1).
+converges to the normal with mean $\sqrt{p} k_1(0) = 0$ and variance
+\[
+p A_1 = p \expect R_1^2 \sin^2(\Phi_1).
 \]
@@
 \]
-converges to the normal with mean $k_2(0) = 0$ and variance
-\[
-A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
+converges to the normal with mean $\sqrt{d} k_2(0) = 0$ and variance
+\[
+d A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
 \]
@@
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - H_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} \big( G_L^\prime(0) - H_L(0) \big).
-\]
-\end{lemma}
-\begin{IEEEproof}
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L) \big) = o_P(1) + X_L,
+\]
+where $X_L = \sqrt{L} \big( G_L(0) - T(0) \big)$.
+\end{lemma}
+\begin{IEEEproof}
+The is the empirical process proof!
 \end{IEEEproof}
@@
 \[
-\sqrt{L}\big( H_L(\hat{\lambda}_L) - G(0) \big) = \hat{\lambda}_L \big( ? + o_P(1) \big).
-\]
-\end{lemma}
-\begin{IEEEproof}
+\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = o_P(1).
+\]
+\end{lemma}
+\begin{IEEEproof}
+The argument is similar to that used in Lemma~\ref{lem:Qconv}.  First observe that
+\begin{align*}
+T(\lambda) &= \frac{\abs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\abs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
+%&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
+%& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
+&=  p \expect R_1\cos(\lambda + \Phi_1) + d\expect R_1\cos\fracpart{\lambda + \Phi_1} + o(1),
+\end{align*}
+because $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L\rightarrow\infty$. Put
+\begin{align*}
+L(\lambda) &= \sqrt{L}\big( T(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
+\end{align*}
+where
+\[
+q_1(\lambda) = \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big),
+\]
+\begin{equation}
+q_2(\lambda) = \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
+\end{equation}
+We analyse $q_1(\lambda)$ first.  By a Taylor expansion of $\cos(\lambda + \Phi_1)$ about $\lambda = 0$,
+\[
+\cos(\lambda + \Phi_1) - \cos(\Phi_1) = -\lambda \big( \sin(\Phi_1) + \zeta(\lambda) \big),
+\]
+where $\zeta$ is a continuous function with $\zeta(0) = 0$. So,
+\begin{align*}
+q_1(\lambda) &= -\expect R_1\lambda(\sin(\Phi_1) +  \zeta(\lambda)) \\
+&= -\lambda \big( \expect R_1\sin(\Phi_1) +  \zeta(\lambda) \expect R_1 \big) \\
+&= \lambda \zeta(\lambda) K,
+\end{align*}
+since $\expect R_1\sin(\Phi_1) = 0$ because $w_1$ has zero mean,  and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.  So, 
+\begin{align*}
+\sqrt{L} q_1(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L \zeta(\hat{\lambda}_L) K = o_P(1)
+\end{align*}
+because $\sqrt{L} \hat{\lambda}_L$ converges in distribution and $\zeta(\hat{\lambda}_L) = o_P(1)$.  Lemma~\ref{} shows that $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$ also.  Thus 
+\[
+L(\lambda) = p q_1(\lambda) + d q_2(\lambda) + o(1) = o_P(1)
+\]
+as reuqired.
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:XL} 
-The distribution of $X_L$ converges to the normal with zero mean and covariance
-\[
-?
+$\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
+\end{lemma}
+\begin{IEEEproof}
+The argument is similar to that used in Lemma~\ref{lem:k2conv}.
+As $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $q_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
+We will analyse $q_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
+An analagous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
+Recall that
+\[
+\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
+\]
+when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
+\begin{align*}
+\cos\fracpart{\lambda + \Phi_1} - &\cos\fracpart{\Phi_1} \\
+&= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
+\end{align*}
+and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
+\begin{align*}
+\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} &= -\lambda\big(\sin(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
+&= -\lambda\big(\sin\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
+\end{align*}
+where, for each integer $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.   Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+\begin{align*}
+\cos\fracpart{\lambda + \Phi_1} &- \cos\fracpart{\Phi_1} \\
+&= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k)
+\end{align*}
+and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
+\begin{align*}
+\cos&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
+&= c_k(\Phi_1) - \lambda\big(\sin\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
+\end{align*}
+where
+\[
+c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
+\]
+and
+\[
+\eta_k(\lambda, \Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_{k+1}(\lambda).
+\]
+Observe that $c_k$ is continuous and $c_k(\psi_{k}) = 0$.
+
+Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
+\[
+\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
+\]
+whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
+\[
+\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda \sin\fracpart{\Phi_1} + c_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
+\]
+Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
+\[
+\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
+\]
+where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
+\begin{equation}\label{eq:vk}
+v_k(\lambda,\Phi_1) = \begin{cases}
+c_k(\Phi_1) - \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
+0 & \text{otherwise}.
+\end{cases}
+\end{equation}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:XL} 
+The distribution of 
+\[
+X_L = \sqrt{L} \big( G_L(0) - T(0) \big) = \sqrt{L} \big( G_L(0) - \expect G_L(0) \big)
+\] 
+converges to the normal with zero mean and covariance 
+\[
+pB_1 + d B_2
 \] 
@@
 \begin{IEEEproof}
-
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:XLcovlambdaL} 
-The covariance between $X_L$ and $\sqrt{L}\hat{\lambda}_L$ converges to zero as $L\rightarrow\infty$.
-\end{lemma}
-\begin{IEEEproof}
-
+Observe that $X_L = C_L + D_L$ where
+\[
+C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} \big( R_i \cos(\Phi_i) - h_1(0) \big),
+\]
+\[
+D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ).
+\]
+From the standard central limit the distribution of $C_L$ converges to the normal with zero mean and variance
+\[
+p B_1 = p \expect R_1^2 \cos^2(\Phi_1) - p h_1^2(0) =  p \expect R_1^2 \cos^2(\Phi_1) - p 
+\]
+since $h_1(0) = 1$.  The distribution of $D_L$ converges to the normal with zero mean and variance
+\[
+d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0)
+\]
+The proof follows since $C_L$ and $D_L$ are indpendent.
 \end{IEEEproof}

### 2012-08-01 code/montecarlo.scala

@@ -11,7 +11,8 @@ import pubsim.distributions.complex.ComplexRandomVariable
 import pubsim.Util
 
 val Ms = List(2,4,8) //BPSK, QPSK, 8-PSK
-val Ls = List(2,4,8,16,32,64,128,256,512,1024,2048)
+//val Ls = List(2,4,8,16,32,64,128,256,512,1024,2048)
+val Ls = List(32,256,1024)
 val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
 val iters = 5000
 
@@ -49,7 +50,7 @@ for( L <- Ls; M <- Ms ) {
         val p = P.length.toDouble/L
         val G0 = noise.clt(M,p).G(0) //value of G0 for unbiasing the amplitude estimator
 
-	var msec = 0.0; var msea = 0.0; var msep = 0.0; var msepunb = 0.0;
+	var msec = 0.0; var msea = 0.0; var msep = 0.0; var mseaunb = 0.0;
 	for( itr <- 1 to iters ) {
 	  //generate a recieved signal
 	  val y = s.map(si => a0*si + noise.noise )
@@ -58,10 +59,10 @@ for( L <- Ls; M <- Ms ) {
 	  msep += pe
 	  msea += ae
 	  msec += (ahat - a0).mag2 //compute the mean square error
-          msepunb += (ahat.magnitude - G0)*(ahat.magnitude - G0) //rho0 is assumed equal to 1 
+          mseaunb += (ahat.magnitude - G0)*(ahat.magnitude - G0) //rho0 is assumed equal to 1
 	}
 	print(".")		      
-	(msea/iters, msep/iters, msec/iters, msepunb/iters) //last thing is what gets returned
+	(msea/iters, msep/iters, msec/iters, mseaunb/iters) //last thing is what gets returned
       }.toList
 
       val estruntime = (new java.util.Date).getTime - eststarttime
@@ -70,15 +71,15 @@ for( L <- Ls; M <- Ms ) {
       val filea = new java.io.FileWriter("data/" + estname + "a")
       val filep = new java.io.FileWriter("data/" + estname + "p")
       val filec = new java.io.FileWriter("data/" + estname + "c")
-      val filepunb = new java.io.FileWriter("data/" + estname + "punb")
+      val fileaunb = new java.io.FileWriter("data/" + estname + "aunb")
       (mselist, SNRdBs).zipped.foreach{ (mse, snr) =>
 	val (ma,mp,mc,mu) = mse
[…]

### 2012-08-01 code/scala/src/cam/noise/ComplexNoise.scala

@@ -1,4 +1,4 @@
-/*
+  /*
  * To change this template, choose Tools | Templates
  * and open the template in the editor.
  */

### 2012-08-01 other/confnoncoherent/code/montecarlo.scala

@@ -1,90 +1,0 @@
-/**
-* Run montecarlo.
-*/
-import cam.psk.MackenthunCoherent
-import cam.psk.ComplexAmplitudeEstimator
-import cam.noise.ComplexGaussian
-import numbers.finite.RectComplex
-import numbers.finite.Complex
-import numbers.finite.PolarComplex
-import pubsim.distributions.complex.ComplexRandomVariable
-import pubsim.Util
-
-val Ms = List(2,4,8) //BPSK, QPSK, 8-PSK
-val Ls = List(2,4,8,16,32,64,128,256,512,1024,2048)
-val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
-val iters = 5000
-
-//construct an array of noise distributions with a logarithmic scale
-val SNRdBs = -20 to 20 by 1
-val SNRs = SNRdBs.map(db => scala.math.pow(10.0, db/10.0))
-val noises = SNRs.map( snr => new ComplexGaussian(a0.mag2/snr/2) ) //variance for real and imaginay parts (divide by 2)
-
-val starttime = (new java.util.Date).getTime
-for( L <- Ls; M <- Ms ) {
-
-  //for a range of different numbers of pilots
-  for( numpilots <- List( L/64, L/32, L/16, L/8, L/4, L/2, L ).distinct.filterNot(b => b == 0) ) {
-
-    val P = 0 until numpilots //pilots at the front
-    val D = numpilots until L //data at the back
-
-    //factory method to enable parallelism
-    def estfactory = List(
-      (p : Seq[Complex]) => new MackenthunCoherent(M,P,D,p)
-    )
-
-    for( estf <- estfactory ) {
-      
-      val estname =  estf(null).getClass.getSimpleName + "M" + M + "L" + L.toString + "absP" + P.length
-      print("Running " + estname)
-      val eststarttime = (new java.util.Date).getTime
-
-      //for all the noise distributions (in parallel threads)
-      val mselist = noises.par.map { noise =>	
-	//generate some PSK symbols 
-	val rand = new scala.util.Random
-	val s = (1 to L).map(m => new PolarComplex(1, 2*scala.math.Pi*rand.nextInt(M)/M)) 
-	val est = estf(s) //construct an estimator	  
-
-	var msec = 0.0; var msea = 0.0; var msep = 0.0;
-	for( itr <- 1 to iters ) {
-	  //generate a recieved signal
[…]

### 2012-08-02 other/confnoncoherent/code/runsim.sh
Added more data and other stuff for conference paper.

@@ -1,2 +1,2 @@
 #export JAVA_OPTS="-Xprof"
-scala -nocompdaemon -cp ../../lib/ComplexAmplitudePSK.jar:../../lib/ScalaNumbers.jar:../../lib/PubSim.jar:../../lib/Jama-1.0.2.jar:../../lib/flanagan.jar:../../lib/colt.jar:../../lib/RngPack.jar montecarlo.scala
+scala -nocompdaemon -cp ../../../code/lib/ComplexAmplitudePSK.jar:../../../code/lib/ScalaNumbers.jar:../../../code/lib/PubSim.jar:../../../code/lib/Jama-1.0.2.jar:../../../code/lib/flanagan.jar:../../../code/lib/colt.jar:../../../code/lib/RngPack.jar montecarlonc.scala

### 2012-08-02 other/confnoncoherent/paper.tex
Added new data.  Looks like I have all the results now.  Just need to start filling in all the Lemmas!
Added more data and other stuff for conference paper.
Cleaned up data for conf

@@
 \title{Noncoherent least squares estimators of carrier phase and amplitude}
-\author{Robby McKilliam, Andre Pollok, Bill Cowley}
+\author{Robby McKilliam$^{1}$, Andre Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \vspace{0.2cm} \\
+\small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.1cm} \\
+\small $^{2}$School of Information Technology \& Electrical Engineering, The University of Queensland, QLD 4067, \textsc{Australia} \vspace{0.1cm} \\
+\small $^{3}$Department of Statistics, Macquarie University, Sydney, \textsc{Australia} \\}
 
@@
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-PSK.
-Under reasonably mild conditions on the distribution of the noise, the least squares estimator of carrier phase is strongly consistent and asymptotically normaly distributed.
+We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normaly distributed.
 However, the amplitude estimator is not consistent, and we show that it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
@@
 
-With $M$-ary phase shift keying ($M$-PSK), the transmitted symbols take the form,
+In passband communications systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation.
+These effects must be compensated for at the reciever.
+In this paper we assume that the time offset has been previously handled, and we focus on estimating the phase shift and attenuation.
+We consider signalling constellations that have symbols evenly distributed on the complex unit circle, commonly refered to as $M$-ary phase shift keying ($M$-PSK).
+In this case, the transmitted symbols take the form,
 \[
@@
 
-We assume that time offset estimation (symbol timing) and matched filtering have already been performed at the reciever.  The recieved symbols take the form,
+We assume that time offset estimation and matched filtering have already been performed at the reciever.  The recieved symbols then take the form,
 \begin{equation}\label{eq:sigmod}
@@
 where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
-Our aim is to estimate $a_0$ from the noisy symbols $y_1, \dots, y_L$.
+Our aim is to estimate $a_0$ from $y_1, \dots, y_L$.
 Complicating matters is that the transmitted symbols $s_i$ are not known to the reciever and must also be estimated.
@@
 \[
-SS(a, \{s_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 
+SS(a, \{s_i, i \in D\}) = \sum_{i = 1}^L \abs{ y_i - a s_i }^2 
 \]
@@
 The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
-However, as we will show, the estimator works well under less stringent assumptions.
+As we will show, the estimator works well under less stringent assumptions.
 Mackenthun~\cite{Mackenthun1994} described an algorithm to compute the least squares estimator that requires only $O(L \log L)$ arithmetic operations.
@@
 Our aim is to estimate $a_0$, and we treat the unknown symbols as nuisance parameters.
-This is motivated by the fact that in our motivating application (narrowband satellite communications) and in many modern communication systems the data symbols are \emph{coded}.
+This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.
 For this reason raw symbol error rate is not of specific interest at this stage.
@@
 
-% \section{The least squares estimator}\label{sec:least-squar-estim}
+\section{The least squares estimator}\label{sec:least-squar-estim}
 
-% Define the sum of squares function
-% \begin{align}
-% SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
-% &= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*, \label{eq:SS}
-% \end{align}
-% where $*$ denotes the complex congugate.
+As discussed, the least squares estimator is given by the minimiser of $SS(a, \{s_i, i \in D\})$ over all $a \in \complex$ and $s_1, \dots, s_L$ in the $M$-PSK constellation.
+For fixed $a$ the least squares estimators of $s_1, \dots, s_L$, as functions $a = \rho e^{j \theta}$, are
+\[
+\hat{s}_i(a) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
+\]
+where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
+Substituting $\hat{s}_i(a)$ into $SS(a, \{s_i, i \in D\})$, give the sum of squares, conditioned on minimisation with respect to the transmitted symbols,
+\[
+SS(a) = \sum_{i = 1}^L \abs{ y_i - a \hat{s}_i(a) }^2.
+\]
+The least squares estimator of the carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$ then satisfy $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{a} = \arg\min_{a \in \complex} SS(a)$.
+Mackenthun~\cite{Mackenthun1994} described an algorithm that computes $\hat{a}$ using $O(L\log L)$ operations.
 
@@
 \]  
-Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariant to phase rotation.
-If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
-\begin{align*}
-\expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
-&= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
-&= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
-\end{align*}
+Observe that for any real number $\phi$, the distribution of $X$ is that same as that of $e^{j\phi}X$.  %If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
+% \begin{align*}
+% \expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
+% &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
+% &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
+% \end{align*}
 
@@
 \]
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
-\[
-\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
-\]
-where $\Re(\cdot)$ denotes the real part, and
-\[
-\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
-\]
-where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
+% The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
+% \[
+% \expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
+% \]
+% where $\Re(\cdot)$ denotes the real part, and
+% \[
+% \expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
+% \]
+% where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
 
@@
 
-The next two theorems decribe the asymptotic properties of the least squares estimator.  We ommit the proofs due to space constraints.  Full proofs will be provided in an upcomming paper.
+The next two theorems decribe the asymptotic properties of the least squares estimator.  We ommit the proofs due to space constraints.  Proofs will be provided in an upcomming paper.
 
@@
 Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod}.
-Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
+Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
 Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
@@
 \begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
 \item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
@@
 \]
-Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
+Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
@@
 \[
-\frac{1}{H^2} \left( \begin{array}{cc} 
-A & ? \\
-? & ?
+\left( \begin{array}{cc} 
+A H^{-2} & 0 \\
+0 & B
 \end{array} \right)
@@
 \]
-and
 \[
-A = \expect R_1^2\sin^2\fracpart{\Phi_1}.
+A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1}. 
 \]
 \end{theorem}
+
+\begin{corollary}
+The mean square error of the amplitude estimator $\hat{\rho}$ is $\expect (\rho_0 - \hat{\rho})^2 = \frac{B}{L} + \rho_0^2(1 - G(0))^2$.
+\end{corollary}
 
@@
 
-\begin{figure}[tp]
+\begin{figure}[p]
 	\centering
@@
 
-\begin{figure}[tp]
+\begin{figure}[p]
 	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
-		\caption{Phase error for 8PSK}
+		\includegraphics[width=\linewidth]{code/data/plotncM2-1.mps}
+		\caption{Amplitude error for BPSK}
 		\label{fig:plotphase}
 \end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotncM4-1.mps}
+		\caption{Amplitude error for QPSK}
+		\label{fig:plotphase}
+\end{figure}
+
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
+%		\caption{Phase error for 8PSK}
+%		\label{fig:plotphase}
+%\end{figure}
 

### 2012-08-02 other/confnoncoherent/build.sh

@@ -1,7 +1,6 @@
 cd code/data/
 mpost plotncM2.mp
 mpost plotncM4.mp
-mpost plotncM8.mp
 cd ../..
 
 pdflatex paper.tex

### 2012-08-02 other/confnoncoherent/code/clt.scala

@@ -10,12 +10,12 @@ import numbers.finite.PolarComplex
 import pubsim.distributions.complex.ComplexRandomVariable
 import pubsim.Util
 
-val Ms = List(2,4,8) //BPSK, QPSK, 8-PSK
+val Ms = List(2,4) //BPSK, QPSK, 8-PSK
 val Ls = List(16,256,4096)
 val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
 
 //construct an array of noise distributions with a logarithmic scale
-val SNRdBs = -15.0 to 20.0 by 0.5
+val SNRdBs = -15.0 to 20.0 by 0.2
 val SNRs = SNRdBs.map(db => scala.math.pow(10.0, db/10.0))
 val noises = SNRs.map( snr => new ComplexGaussian(a0.mag2/snr/2) ) //variance for real and imaginay parts (divide by 2)
 

### 2012-08-02 other/confnoncoherent/paper.tex
Added to conf

@@
 \begin{IEEEkeywords}
-BLERG
+Noncoherent detection, differential phase shift keying, asymptotic statistics
 \end{IEEEkeywords}
@@
 \section{Introduction}
+
+\newcommand{\calC}{\mathcal C}
 
@@
 where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation.
+%We denote the set of symbols contained in the $M$-PSK constellation by $\calC$.
 
@@
 Our aim is to estimate $a_0$ from $y_1, \dots, y_L$.
-Complicating matters is that the transmitted symbols $s_i$ are not known to the reciever and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
-A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
+If the transmitted symbols $s_1, \dots, s_L$ are known apriory at the reciever then the least squares estimator is typically used,
 \[
-SS(a, \{s_i, i \in D\}) = \sum_{i = 1}^L \abs{ y_i - a s_i }^2 
+\hat{a}_{\text{uc}} = \arg \min_{a \in \complex} \sum_{i = 1}^L \abs{ y_i - a s_i }^2  = \frac{1}{L} \sum_{i = 1}^L y_i s_i^*,
 \]
-where $\abs{x}$ denotes the magnitude of the complex number $x$.
+where $\complex$ is the set of complex numbers and $x^*$ and $\abs{x}$ denotes the complex conjugate and complex magnitude of the complex number $x$.
+Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} call this the \emph{unmodulated carrier} estimator.
+This estimator can be used if the tranmitter includes \emph{pilot} symbols, known to the reciever, i.e.
+with~\emph{coherent detection}.
+
+Here we are interested in \emph{noncoherent detection}, when $s_1, \dots, s_L$ are not known at the reciever, and must also be estimated.
+Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
+A practical approach is the least squares estimator,
+\[
+\hat{a} = \arg\min_{a \in \complex} \min_{s_1, \dots, s_L \in \calC} \sum_{i = 1}^L \abs{ y_i - a s_i }^2,
+\]
+where $\calC$ is the set of symbols from the $M$-PSK constellation.
 The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
 As we will show, the estimator works well under less stringent assumptions.
-Mackenthun~\cite{Mackenthun1994} described an algorithm to compute the least squares estimator that requires only $O(L \log L)$ arithmetic operations.
+Mackenthun~\cite{Mackenthun1994} described an algorithm to compute the least squares estimator $\hat{a}$ that requires only $O(L \log L)$ arithmetic operations.
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
@@
 
-\section{The least squares estimator}\label{sec:least-squar-estim}
+% \section{The least squares estimator}\label{sec:least-squar-estim}
 
-As discussed, the least squares estimator is given by the minimiser of $SS(a, \{s_i, i \in D\})$ over all $a \in \complex$ and $s_1, \dots, s_L$ in the $M$-PSK constellation.
+% As discussed, the least squares estimator is given by the minimiser of $SS(a, \{s_i, i \in D\})$ over all $a \in \complex$ and $s_1, \dots, s_L$ in the $M$-PSK constellation.
 For fixed $a$ the least squares estimators of $s_1, \dots, s_L$, as functions $a = \rho e^{j \theta}$, are
-\[
-\hat{s}_i(a) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
-\]
-where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
+% \[
+% \hat{s}_i(a) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
+% \]
+% where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
 Substituting $\hat{s}_i(a)$ into $SS(a, \{s_i, i \in D\})$, give the sum of squares, conditioned on minimisation with respect to the transmitted symbols,
-\[
-SS(a) = \sum_{i = 1}^L \abs{ y_i - a \hat{s}_i(a) }^2.
-\]
-The least squares estimator of the carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$ then satisfy $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{a} = \arg\min_{a \in \complex} SS(a)$.
+% \[
+% SS(a) = \sum_{i = 1}^L \abs{ y_i - a \hat{s}_i(a) }^2.
+% \]
+% The least squares estimator of the carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$ then satisfy $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{a} = \arg\min_{a \in \complex} SS(a)$.
 Mackenthun~\cite{Mackenthun1994} described an algorithm that computes $\hat{a}$ using $O(L\log L)$ operations.
@@
 \begin{theorem}\label{thm:normality} (Asymptotic normality)
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
+Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint pdf of $R_1$ and $\Phi_1$, and let
 \[
@@
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-Then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
 \[
 \left( \begin{array}{cc} 
-A H^{-2} & 0 \\
+H^{-2} A & 0 \\
 0 & B
@@
 
-\begin{corollary}
-The mean square error of the amplitude estimator $\hat{\rho}$ is $\expect (\rho_0 - \hat{\rho})^2 = \frac{B}{L} + \rho_0^2(1 - G(0))^2$.
+\begin{corollary}\label{cor:ampmse}
+The mean square error of the amplitude estimator $\hat{\rho}$ is 
+\[
+\expect (\rho_0 - \hat{\rho})^2 = \frac{B}{L} + \rho_0^2(1 - G(0))^2 + o(1)
+\] 
+where $o(1)$ goes to zero as $L \rightarrow \infty$.
 \end{corollary}
+\begin{IEEEproof} Write
+\begin{align*}
+\expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
+&= \expect \big( \rho_0(1 - G(0)) + \hat{m}_L \big)^2 \\
+&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 + 2\rho_0(1 - G(0))\expect\hat{m}_L,
+\end{align*}
+and the corollary holds since $\expect \hat{m}_L^2 = B/L + o(1)$ and $\expect\hat{m}_L = o(1)$ from Theorem~\ref{thm:normality}.
+\end{IEEEproof}
 
@@
 
-We now make some discussion regarding the assumptions made by these theorems.
+We now discuss the assumptions made by these theorems.
+The assumption that the $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statement more complicated.
+Primarily this is because if $w_i$ are not circularly symmetric then the distribution of the random variables $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $s_1, \dots, s_L$.
+The result of this is that the asymptotic variance described in Theorem~\ref{thm:normality} is more complicated, and depends on $a_0$ and $s_1, \dots, s_L$.
+The circularly symmetric assumption may not always hold in practice, but it does lead to simple, and useful expressions for the asymptotic variance.
+
+A key assumption in Theorem~\ref{thm:consistency} is that the function $\abs{G(x)}$ is uniquely maximised at $x = 0$.
 
@@
 \]
-The value of $A$ can be computed by numerical integration.
+The value of $A$, $B$ and $H$ can be efficiently computed by numerical integration using the above formula for $g(\phi)$.
 
 \section{Simulations}\label{sec:simulations}
+
+We present the results of Monte-Carlo simulations that evaluate the performance of the least squares estimator.
+In all simulations the noise $\{w_i\}$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+Simulations are run with $M=2,4$ (BPSK and QPSK) and $L=16,256,4096$ with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
+Five thousand replications are used per experiment.
+
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the mean square error (MSE) of the phase estimator $\hat{\theta}$.
+The dots are the Monte-Carlo simulations of the least square estiamtor.
+The dashed line is the simulated performance of least squares estimator when all the data symbols $s_1, \dots, s_L$ are known apriory at the reciever.
+Viterbi calls this the \emph{unmodulated carrier} and we use this name also.
+The dashed line acts as a lower bound on performance.
+The solid line is the MSE predicted by Theorem~\ref{thm:normality}.
+The Theorem accurately predicts the behaviour of the phase estimator when $L$ is large and when the SNR is not too small.
+As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi/M, \pi/M)$, and Theorem~\ref{thm:normality} does not model this behaviour.
+
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the mean square error (MSE) of the amplitude estimator $\hat{\rho}$.
+The dots are again the Monte-Carlo simulations, and the dahsed line is the unmodulated carrier.
+The solid line is the asymptotic MSE predicted by Corollary~\ref{cor:ampmse}.
+The corollary accurately models the behaviour of the amplitude estimator when $L$ is sufficiently large.
+The estimator is biased, and as $L\rightarrow\infty$, the MSE converges to $\rho_0^2(1 - G(0))^2$.
+The effect is clearly seen in the Figures when the SNR is small.
+
+%\section{Discussion}
+
+\section{Conclusion}
 
@@
 		\caption{Phase error for BPSK}
-		\label{fig:plotphase}
+		\label{fig:plotphaseM2}
 \end{figure}
@@
 		\caption{Phase error for QPSK}
-		\label{fig:plotphase}
+		\label{fig:plotphaseM4}
 \end{figure}
@@
 		\caption{Amplitude error for BPSK}
-		\label{fig:plotphase}
+		\label{fig:plotampM2}
 \end{figure}
@@
 		\caption{Amplitude error for QPSK}
-		\label{fig:plotphase}
+		\label{fig:plotampM}
 \end{figure}

### 2012-08-02 other/confnoncoherent/code/montecarlonc.scala

@@ -17,7 +17,7 @@ val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
 val iters = 5000
 
 //construct an array of noise distributions with a logarithmic scale
-val SNRdBs = -15 to 20 by 1
+val SNRdBs = -20 to 20 by 1
 val SNRs = SNRdBs.map(db => scala.math.pow(10.0, db/10.0))
 val noises = SNRs.map( snr => new ComplexGaussian(a0.mag2/snr/2) ) //variance for real and imaginay parts (divide by 2)
 

### 2012-08-02 other/confnoncoherent/paper.tex
more conf edits.
Added to paper

@@
 We consider least squares estimators of carrier phase and amplitude from a noisy communications signal.
-We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-PSK.
+We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-ary phase shift keying.
 We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normaly distributed.
-However, the amplitude estimator is not consistent, and we show that it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
+However, the amplitude estimator is not consistent, it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
 The results of Monte Carlo simulations are provided and these corroborate the theoretical results.
@@
 In this paper we assume that the time offset has been previously handled, and we focus on estimating the phase shift and attenuation.
-We consider signalling constellations that have symbols evenly distributed on the complex unit circle, commonly refered to as $M$-ary phase shift keying ($M$-PSK).
+We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK) and $M$-ary phase shift keying ($M$-PSK).
 In this case, the transmitted symbols take the form,
@@
 \]
-where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation.
+where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$.
+%and $M$ is the size of the constellation.
 %We denote the set of symbols contained in the $M$-PSK constellation by $\calC$.
 
-We assume that time offset estimation and matched filtering have already been performed at the reciever.  The recieved symbols then take the form,
+We assume that time offset estimation and matched filtering have already been performed at the reciever.  The recieved symbols are then,
 \begin{equation}\label{eq:sigmod}
@@
 \]
-where $\complex$ is the set of complex numbers and $x^*$ and $\abs{x}$ denotes the complex conjugate and complex magnitude of the complex number $x$.
+where $\complex$ is the set of complex numbers and $x^*$ and $\abs{x}$ denotes the conjugate and magnitude of the complex number $x$.
 Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} call this the \emph{unmodulated carrier} estimator.
-This estimator can be used if the tranmitter includes \emph{pilot} symbols, known to the reciever, i.e.
-with~\emph{coherent detection}.
+This estimator can be used if the tranmitter includes \emph{pilot} symbols, known to the reciever, i.e.~\emph{coherent detection}.
 
 Here we are interested in \emph{noncoherent detection}, when $s_1, \dots, s_L$ are not known at the reciever, and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
+This estimation problem has undergone extensive prior study and is often called \emph{multiple symbol differential detection}~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
 A practical approach is the least squares estimator,
@@
 where $\calC$ is the set of symbols from the $M$-PSK constellation.
-The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
-As we will show, the estimator works well under less stringent assumptions.
+The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is white and Gaussian.
+As we will show, the estimator can work well even when the noise is not Gaussian.
 Mackenthun~\cite{Mackenthun1994} described an algorithm to compute the least squares estimator $\hat{a}$ that requires only $O(L \log L)$ arithmetic operations.
@@
 It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+
+The paper is organised in the following way.
+Section~\ref{sec:circ-symm-compl} describes some properties of complex random variables that we need.
+Section~\ref{sec:stat-prop-least} describes the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
+We show, under some assumptions about the distribution of the noise sequence $\{w_i\}$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed.
+However, the amplitude estimator $\hat{\rho_0}$ is not a consistent estimator of the true amplitude $\rho_0$.
+The bias is small if the signal to noise ratio (SNR) is large, but is significant when the SNR is small.
+In Section~\ref{sec:simulations} we present the results of simulations.
+These simulations agree with the derived asymptotic properties.
 
@@
 \expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
-&= \expect \big( \rho_0(1 - G(0)) + \hat{m}_L \big)^2 \\
-&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 + 2\rho_0(1 - G(0))\expect\hat{m}_L,
+&= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
+&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L,
 \end{align*}
@@
 A key assumption in Theorem~\ref{thm:consistency} is that the function $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+The next Lemma describes a class of circularly symmetric distributions that satisfy this property.
 
@@
 \end{IEEEproof}
+
+The theorems place conditions on the $\sfracpart{\hat{\theta} - \theta_0}$ rather than directly on $\hat{\theta} - \theta_0$.  This makes practical sense and BLERG.
 
@@
 
+\small
+\bibliography{bib}
+
 \begin{figure}[p]
@@
 
-\small
-\bibliography{bib}
-
 \end{document}

### 2012-08-03 other/confnoncoherent/paper.tex
Added to icassp version.  First draft has been given to Andre and Bill.

@@
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-ary phase shift keying.
-We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normaly distributed.
+We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
 However, the amplitude estimator is not consistent, it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
@@
 \begin{IEEEkeywords}
-Noncoherent detection, differential phase shift keying, asymptotic statistics
+Noncoherent detection, phase shift keying, asymptotic statistics
 \end{IEEEkeywords}
@@
 
-In passband communications systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation.
-These effects must be compensated for at the reciever.
+In passband communication systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation.
+These effects must be compensated for at the receiver.
 In this paper we assume that the time offset has been previously handled, and we focus on estimating the phase shift and attenuation.
@@
 
-We assume that time offset estimation and matched filtering have already been performed at the reciever.  The recieved symbols are then,
+We assume that time offset estimation and matched filtering have been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
 \begin{equation}\label{eq:sigmod}
@@
 Our aim is to estimate $a_0$ from $y_1, \dots, y_L$.
-If the transmitted symbols $s_1, \dots, s_L$ are known apriory at the reciever then the least squares estimator is typically used,
-\[
+If the transmitted symbols $s_1, \dots, s_L$ are known a priori at the receiver then the least squares estimator is typically used,
+\begin{equation}\label{eq:hataumod}
 \hat{a}_{\text{uc}} = \arg \min_{a \in \complex} \sum_{i = 1}^L \abs{ y_i - a s_i }^2  = \frac{1}{L} \sum_{i = 1}^L y_i s_i^*,
-\]
-where $\complex$ is the set of complex numbers and $x^*$ and $\abs{x}$ denotes the conjugate and magnitude of the complex number $x$.
+\end{equation}
+where $\complex$ is the set of complex numbers and $x^*$ and $\abs{x}$ denote the conjugate and magnitude of the complex number $x$.
 Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} call this the \emph{unmodulated carrier} estimator.
-This estimator can be used if the tranmitter includes \emph{pilot} symbols, known to the reciever, i.e.~\emph{coherent detection}.
+This estimator can be used if the transmitter includes \emph{pilot} symbols, known to the receiver, i.e.~\emph{coherent detection}.
 
-Here we are interested in \emph{noncoherent detection}, when $s_1, \dots, s_L$ are not known at the reciever, and must also be estimated.
+Here we are interested in \emph{noncoherent detection}, where $s_1, \dots, s_L$ are not known at the receiver, and must also be estimated.
 This estimation problem has undergone extensive prior study and is often called \emph{multiple symbol differential detection}~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
 A practical approach is the least squares estimator,
-\[
+\begin{equation}\label{eq:hatadef}
 \hat{a} = \arg\min_{a \in \complex} \min_{s_1, \dots, s_L \in \calC} \sum_{i = 1}^L \abs{ y_i - a s_i }^2,
-\]
+\end{equation}
 where $\calC$ is the set of symbols from the $M$-PSK constellation.
@@
 
-In the literature it has been common to assume that the symbols $\{s_i, i \in D\}$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
+In the literature it has been common to assume that the symbols $s_1, \dots, s_L$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
 The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
@@
 This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.
-For this reason raw symbol error rate is not of specific interest at this stage.
-Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
-The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
+For this reason raw symbol error rate is not of interest at this stage.
+Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated received symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
+The additive noise channel is a common assumption for subsequent receiver operations, such as decoding.
+Consequently, our metric of performance is mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
+It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
 The function $\fracpart{\cdot}$ denotes its argument taken `modulo $\tfrac{2\pi}{M}$' into the interval $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$.
-It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate error for the phase parameter.
 
 The paper is organised in the following way.
-Section~\ref{sec:circ-symm-compl} describes some properties of complex random variables that we need.
+Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
 Section~\ref{sec:stat-prop-least} describes the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some assumptions about the distribution of the noise sequence $\{w_i\}$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed.
-However, the amplitude estimator $\hat{\rho_0}$ is not a consistent estimator of the true amplitude $\rho_0$.
-The bias is small if the signal to noise ratio (SNR) is large, but is significant when the SNR is small.
-In Section~\ref{sec:simulations} we present the results of simulations.
-These simulations agree with the derived asymptotic properties.
+We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
+However, $\hat{\rho_0}$ is not a consistent estimator of the amplitude $\rho_0$.
+The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
+%In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
+In this case, our formula for the asymptotic variance of can be simplified.
+Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.  These simulations agree with the derived asymptotic properties.
 
@@
 
-Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
+%Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.  
 A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
-That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
-If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf (pdf) of $\Theta$ and $Z$ is
+That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
+If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
 \[
@@
 \]  
-Observe that for any real number $\phi$, the distribution of $X$ is that same as that of $e^{j\phi}X$.  %If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
+If $X$ is circularly symmetric, then for any real number $\phi$, the distribution of $X$ is the same as that of $e^{j\phi}X$.
+%If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
 % \begin{align*}
@@
 
-The next two theorems decribe the asymptotic properties of the least squares estimator.  We ommit the proofs due to space constraints.  Proofs will be provided in an upcomming paper.
+The next two theorems describe the asymptotic properties of the least squares estimator.  We omit the proofs due to space constraints.  Proofs will be provided in an upcoming paper.
 
@@
 \] 
-If $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
+If $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
 \begin{enumerate}
@@
 We now discuss the assumptions made by these theorems.
-The assumption that the $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statement more complicated.
-Primarily this is because if $w_i$ are not circularly symmetric then the distribution of the random variables $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $s_1, \dots, s_L$.
-The result of this is that the asymptotic variance described in Theorem~\ref{thm:normality} is more complicated, and depends on $a_0$ and $s_1, \dots, s_L$.
-The circularly symmetric assumption may not always hold in practice, but it does lead to simple, and useful expressions for the asymptotic variance.
+The assumption that $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
+If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $s_1, \dots, s_L$.
+In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $s_1, \dots, s_L$, rather than just $\rho_0$.
+The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
 
@@
 The next Lemma describes a class of circularly symmetric distributions that satisfy this property.
+We omit the proof due to space constraints.
 
@@
 \]
-Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
+Then $\abs{G(x)}$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
 \end{lemma}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
 
-The theorems place conditions on the $\sfracpart{\hat{\theta} - \theta_0}$ rather than directly on $\hat{\theta} - \theta_0$.  This makes practical sense and BLERG.
+The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the probability density function of $X$ decreases as we move away from the origin.
+That is, the pdf of $X$ in rectangular coordinates is $p(x,y) = z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$, and this is required to be non increasing with $z$.
+For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
 
-\section{The Gaussian noise case}
+The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}$ rather than directly on $\hat{\theta} - \theta_0$.
+This makes practical sense, and to see why let $s_i' = e^{j2\pi k/M}s_i$ be $M$-PSK symbols obtained by rotating $s_1, \dots, s_L$ by $e^{j2\pi k/M}$ for some integer $k$.
+Then
+\[
+a s_i =  \rho e^{j\theta}s_i = \rho e^{j(\theta - 2\pi k/M)} s_i',
+\]
+and so, if $a = \hat{\rho}e^{j\hat{\theta}}$ minimises~\eqref{eq:hatadef}, then so does $\hat{\rho} e^{j(\hat{\theta} - 2\pi k/M)}$.
+Thus, $\theta_0 - \hat{\theta}$ should be attributed the same error as $\theta_0 - \hat{\theta} - \tfrac{2\pi}{M}k$, that is, the error should be computed `modulo $\tfrac{2\pi}{M}$'.
+It is for this reason that \emph{differential encoding} is often used with noncoherent $M$-PSK  detectors.
 
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint pdf of the real and imaginary parts is then
+\section{The Gaussian noise case}\label{sec:gaussian-noise-case}
+
+Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint pdf of the real and imaginary parts is
 \[
-p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
+p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}.
 \]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
 %and the joint density function of the real and imaginary parts of $Z$ is
@@
 
-We present the results of Monte-Carlo simulations that evaluate the performance of the least squares estimator.
-In all simulations the noise $\{w_i\}$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+We present the results of Monte-Carlo simulations with the least squares estimator.
+In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
 Simulations are run with $M=2,4$ (BPSK and QPSK) and $L=16,256,4096$ with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
-Five thousand replications are used per experiment.
+The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
+For each value of SNR $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
 
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the mean square error (MSE) of the phase estimator $\hat{\theta}$.
-The dots are the Monte-Carlo simulations of the least square estiamtor.
-The dashed line is the simulated performance of least squares estimator when all the data symbols $s_1, \dots, s_L$ are known apriory at the reciever.
-Viterbi calls this the \emph{unmodulated carrier} and we use this name also.
-The dashed line acts as a lower bound on performance.
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the mean square error (MSE) of the phase estimator $\hat{\theta}$ computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$.
+The dots are the Monte-Carlo simulations of the least square estimator.
+The dashed line the performance of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
+%The dashed line acts as a lower bound on performance.
 The solid line is the MSE predicted by Theorem~\ref{thm:normality}.
-The Theorem accurately predicts the behaviour of the phase estimator when $L$ is large and when the SNR is not too small.
-As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi/M, \pi/M)$, and Theorem~\ref{thm:normality} does not model this behaviour.
+The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
+As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$, and Theorem~\ref{thm:normality} does not model this behaviour.
+%An interesting feature is that the phase estimator is not accurate at low SNR unless $L$ is very large.
 
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the mean square error (MSE) of the amplitude estimator $\hat{\rho}$.
-The dots are again the Monte-Carlo simulations, and the dahsed line is the unmodulated carrier.
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the MSE of the amplitude estimator $\hat{\rho}$.
+The dots are again the Monte-Carlo simulations, and the dashed line is the unmodulated carrier.
 The solid line is the asymptotic MSE predicted by Corollary~\ref{cor:ampmse}.
-The corollary accurately models the behaviour of the amplitude estimator when $L$ is sufficiently large.
+The corollary accurately models the behaviour of $\hat{\rho}$ when $L$ is sufficiently large.
 The estimator is biased, and as $L\rightarrow\infty$, the MSE converges to $\rho_0^2(1 - G(0))^2$.
-The effect is clearly seen in the Figures when the SNR is small.
+The bias is clearly seen in the figures when the SNR is small.
 
@@
 \section{Conclusion}
+
+We have studied the least squares estimator of carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols.
+The estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
+We shows that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
+This bias is large when the signal to noise ratio is small.
+It would be interesting to investigate methods for correcting this bias.
 

### 2012-08-04 paper.tex
Added to conf
added a bit to main paper.

@@
 %opening
-\title{Carrier phase and amplitude estimation of $M$-ary phase shift keying using pilots and data}
+\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
 \author{Robby McKilliam, Andre Pollok, Bill Cowley}
@@
 \begin{abstract}
-We describe least squares estimators of carrier phase and amplitude from a noisy communications signal that contains both pilot signals, known to the receiver, and data signals, unknown to the receiver.
-We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-PSK.
-We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of the carrier phase is strongly consistent and asymptotically normaly distributed.
-However, the amplitude estimator is not consistent, and we show that it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
-Our theoretical results can also be applied to the case where no pilot symbols exist (i.e.
-noncoherent detection).
-The results of Monte Carlo simulations are provided and these corroborate the theoretical results.
+We consider least squares estimators of carrier phase and amplitude from a noisy communications signal that contains both pilot signals, known to the receiver, and data signals, unknown to the receiver.
+We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-ary phase shift keying.
+We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
+However, the amplitude estimator is not consistent, it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
+Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.
+The results of Monte Carlo simulations are provided and these agree with the theoretical results.
 \end{abstract}
 \begin{IEEEkeywords}
-BLERG
+Coherent detection, noncoherent detection, phase shift keying, asymptotic statistics
 \end{IEEEkeywords}
@@
 
-With $M$-ary phase shift keying ($M$-PSK), the transmitted symbols take the form,
+In passband communication systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation.
+These effects must be compensated for at the receiver.
+In this paper we assume that the time offset has been previously handled, and we focus on estimating the phase shift and attenuation.
+We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK) and $M$-ary phase shift keying ($M$-PSK).
+In this case, the transmitted symbols take the form,
 \[
@@
 
-We assume that time offset estimation (symbol timing) and matched filtering have already been performed.  The recieved symbols are then given by,
+We assume that time offset estimation and matched filtering have been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
 \begin{equation}\label{eq:sigmod}
@@
 Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
-Complicating matters is that the data symbols $d_i$ are not known to the reciever and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001}.
+Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the reciever and must also be estimated.
+Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
 A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
@@
 Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
-This is motivated by the fact that in our motivating application (narrowband satellite communications) and in many modern communication systems the data symbols are \emph{coded}.
+This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.
 For this reason raw symbol error rate is not of specific interest at this stage.
@@
 Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.
-Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$ where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
+Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
+%where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
 Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
@@
 \begin{equation}\label{eq:RiandPhii}
-R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*,
+R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
 \end{equation}
@@
 \]
-If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,-\pi)$, then:
+If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
 \begin{enumerate}
@@
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-Then the distribution of $(\sqrt{L}\lambda_L, \sqrt{L}m_L)$ converges to the bivariate normal with zero mean and covariance
+Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
 \[
@@
 
-%BLERG discuss assumptions.
-We have that $h_1$ is uniquely maximised at $0$ over the interval $[-\pi,\pi)$ since $\{w_i\}$ has zero mean.
-The requirement that $h_2$ is uniquely maximised at 0 is stronger that we need.
-We actually only require that the sum $p h_1(x) + dh_2(x)$ is maximised uniquely at 0 over the interval $[-\pi, \pi)$.
-However, we have stated the theorem in this stronger form as it makes the proof in the noncoherent case (i.e.
-when there are no pilot symbols) follow more transparently.
-
-%BLERG $\hat{a}$ is a biased estimator, particularly when the SNR is low, we show firs that $\hat{\theta}$ is strongly consistent, we are then able to analyse the behaviour of $\hat{a}$.
-
-Proofs of these theorems are given in Sections~\ref{sec:proof-almost-sure}~and~\ref{sec:proof-asympt-norm}.  We first discuss the assumptions made by these theorems.
+The proof of Theorem~\ref{thm:consistency} is in Section~\ref{sec:proof-almost-sure} and the proof of Theorem~\ref{thm:normality} is in Section~\ref{sec:proof-asympt-norm}.
+Before giving the proofs we discuss the assumptions made by these theorems.
+The assumption that $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
+If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $s_1, \dots, s_L$.
+In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $s_1, \dots, s_L$, rather than just $\rho_0$.
+The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
+
+A key assumption in Theorem~\ref{thm:consistency} is that the function $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+The next Lemma describes a class of circularly symmetric distributions that satisfy this property.
 
@@
 \end{IEEEproof}
+
+The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the probability density function of $X$ decreases as we move away from the origin.
+That is, the pdf of $X$ in rectangular coordinates is $p(x,y) = z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$, and this is required to be non increasing with $z$.
+For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
+
+The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta} - \theta_0$.
+This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
+So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be \emph{close} together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and \emph{not} $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
 
@@
 
+\section{Conclusion}
+
 \bibliography{bib}

### 2012-08-06 other/confnoncoherent/paper.tex
minor
minor additions.

@@
 \begin{equation}\label{eq:RiandPhii}
-R_ie^{j\Phi_i} = 1 + \frac{1}{\rho_0} w_i e^{-j\theta_0} s_i^*,
+R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
 \end{equation}
@@
 
-A key assumption in Theorem~\ref{thm:consistency} is that the function $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$.
 The next Lemma describes a class of circularly symmetric distributions that satisfy this property.
@@
 \]
-where
+where $\kappa = \rho_0/\sigma$ and
+\vspace{-0.1cm}
 \[
-\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
+\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt.
 \]
-The value of $A$, $B$ and $H$ can be efficiently computed by numerical integration using the above formula for $g(\phi)$.
+Each of $A$, $B$ and $H$ can be computed efficiently by numerical integration using the above formula for $g(\phi)$.
 

### 2012-08-06 paper.tex
minor
minor additions.

@@
 
-In passband communication systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation.
+In passband communication systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation (amplitude change).
 These effects must be compensated for at the receiver.
@@
 The idea is the `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
-A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and $A$ is the average
+A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
 \[
@@
 It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+
+The paper is organised as follows BLERG.
 
@@
 That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
-If the probability density function of $Z$ is $f_Z(z)$, then the joint probability density function (pdf) of $\Theta$ and $Z$ is
+If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
 \[
@@
 \]
-where $\Im(\cdot)$ denotes the imaginary part.  The next lemma will be useful for the analysis of the least squares estimator.
-
-\begin{lemma}\label{lem:h1minedcircsym}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
+where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
+
+% \begin{lemma}\label{lem:h1minedcircsym}
+% Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
 Let
-\[
-h_1(x) = \expect R \cos(x + \Phi).
-\]
-Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
-\end{lemma}
-% Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
+% \[
+% h_1(x) = \expect R \cos(x + \Phi).
+% \]
+% Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
+% \end{lemma}
+% % Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
 That is, the pdf of $Z e^{j \Theta}$ in rectangular coordinates is given by $z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$ and $x$ and $y$ are the real and imaginary parts of $Z e^{j \Theta}$, and this pdf is non increasing with $z$.
@@
 
-% By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if
-% \[
-% \abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
-% \]
-% We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
+% % By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if
+% % \[
+% % \abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
+% % \]
+% % We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
+% \begin{IEEEproof}
+% BLERG
+% \end{IEEEproof}
 
@@
 
+\begin{corollary}\label{cor:ampmse}
+The mean square error of the amplitude estimator $\hat{\rho}$ is 
+\[
+\expect (\rho_0 - \hat{\rho})^2 = \frac{B}{L} + \rho_0^2(1 - G(0))^2 + o(1)
+\] 
+where $o(1)$ goes to zero as $L \rightarrow \infty$.
+\end{corollary}
+\begin{IEEEproof} Write
+\begin{align*}
+\expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
+&= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
+&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L,
+\end{align*} 
+and the corollary holds since $\expect \hat{m}_L^2 = B/L + o(1)$ and $\expect\hat{m}_L = o(1)$ from Theorem~\ref{thm:normality}.
+\end{IEEEproof}
+
 The proof of Theorem~\ref{thm:consistency} is in Section~\ref{sec:proof-almost-sure} and the proof of Theorem~\ref{thm:normality} is in Section~\ref{sec:proof-asympt-norm}.
@@
 
-A key assumption in Theorem~\ref{thm:consistency} is that the function $\abs{G(x)}$ is uniquely maximised at $x = 0$.
-The next Lemma describes a class of circularly symmetric distributions that satisfy this property.
+The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta} - \theta_0$.
+This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
+So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be \emph{close} together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and \emph{not} $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
+
+A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
+This assumption requires that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence such that $\abs{G(x_i)} \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+Although we will not prove it here, this assumption is not only sufficient, but nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
 
@@
 \[
-h_2(x) = \expect R \cos\sfracpart{x + \Phi}, \;\;\; k_2(x) =  \expect R \sin\sfracpart{ x + \Phi}.
-\]
-Then $h_2(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$ and $k_2(0) = 0$.
+h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{ x + \Phi}.
+\]
+Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
 \end{lemma}
@@
 
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta} - \theta_0$.
-This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
-So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be \emph{close} together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and \emph{not} $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
-
 \section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
 
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect the the data symbols,
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
  \begin{align*}
@@
 and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of the amplitude $\rho_0$ as a function of $\theta$,
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
 \begin{equation}\label{eq:hatrhoZ}
@@
 \]
-where
+where $\kappa = \rho_0/\sigma$ and
 \[

### 2012-08-06 paper.tex
Added to paper.  Looking good.

@@
 As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ upon which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
+Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.
 Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
 Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $G$ is uniquely maximised at zero, it follows that $G(\hat{\lambda}_L) \rightarrow G(0)$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
+Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
 So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
@@
 \begin{lemma}\label{lem:convtoexpGlamL} 
-$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-BLERG
-% Lemma~\ref{lem:Zconunif} shows BLERG that 
-% \[
-% 0 \leq G(\hat{\lambda}_L) - G(0),
-% \]
-% and since $\hat{\lambda}_L$ is the maximiser of $G_L$ then
-% \[ 
-% 0 \leq G_L(0) - G_L(\hat{\lambda}_L).
-% \]
-% Adding both inequalities and taking appropriate absolute values gives
-% \begin{align*}
-% 0 &\leq G(\hat{\lambda}_L) - G(0) \\ 
-% & \leq G(\hat{\lambda}_L) - G(0) + G_L(0) - G_L(\hat{\lambda}_L) \\
-% & \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)|,
-% \end{align*}
+$\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Since $\sabs{G(x)}$ is uniquely maximised at $x=0$,
+\[
+0 \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)},
+\]
+and since $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(x)}$,
+\[ 
+0 \leq \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)}.
+\]
+Thus,
+\begin{align*}
+0 &\leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} \\ 
+& \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} + \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)} \\
+& \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)| \\
+&\leq 2\sup_{x \in [-\pi, \pi)} \sabs{G(x) - G_L(x)},
+\end{align*}
+and the last line converges amost surely to zero by Lemma~\ref{lem:uniflawGGL}.
 \end{IEEEproof}
@@
 $G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+By the triangle inequality,
+\[
+\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
+\]
+Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:uniflawGGL}
+The function $G_L$ converges almost surely and uniformly on $[-\pi, \pi)$ to the function $G$, that is,
+\[
+\sup_{x \in [-\pi, \pi)}\sabs{G_L(x) - G(x)} \rightarrow 0
+\] 
+almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Put $T(x) = \expect G_L(x)$ and observe that
+\begin{align*}
+T(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
+&= o(1) h_1(\lambda) + o(1) h_2(\lambda)
+\end{align*}
+where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.
+Since $\sabs{h_1(\lambda)} \leq 1$ and $\sabs{h_2(\lambda)} \leq 1$ for all $\lambda \in [-\pi, \pi)$, then $\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \rightarrow 0$ as $L\rightarrow \infty$.
+Now
+\begin{align*}
+\sup_{\lambda \in [-\pi, \pi)}&\sabs{G_L(\lambda) - G(\lambda)} \\
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda) + T(\lambda) - G(\lambda)} \\
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)}
+\end{align*}
+and Lemma~\ref{lem:uniflawTGL} shows that $\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0$ as $L\rightarrow \infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:uniflawTGL} 
+Put $T(x) = \expect G_L(x)$, then
+\[
+\sup_{x \in [-\pi, \pi)}\sabs{G_L(x) - T(x)} \rightarrow 0
+\] 
+almost surely as $L \rightarrow \infty$.
 \end{lemma}

### 2012-08-07 other/confnoncoherent/paper.tex
added to convergence proof. Getting there, but needs some cleanup.

@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.

### 2012-08-07 paper.tex
added to convergence proof. Getting there, but needs some cleanup.

@@
 
-\section{Mackentun's algorithm}\label{sec:least-squar-estim}
+\section{Mackentun's algorithm with pilots}\label{sec:least-squar-estim}
 
@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.
 Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
@@
 The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
+
+The assumption that $\expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.
+This is required in Theorem~\ref{thm:normality} so that the constants $A_1$, $A_2$, $B_1$ and $B_2$ exist.
+We will also use that $R_i$ has finite variance to simplify the proof of Theorem~\ref{thm:consistency} by use of Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}
 
@@
 & \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)| \\
-&\leq 2\sup_{x \in [-\pi, \pi)} \sabs{G(x) - G_L(x)},
+&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G(\lambda) - G_L(\lambda)},
 \end{align*}
@@
 \[
-\sup_{x \in [-\pi, \pi)}\sabs{G_L(x) - G(x)} \rightarrow 0
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0
 \] 
@@
 \begin{IEEEproof}
-Put $T(x) = \expect G_L(x)$ and observe that
+Put $T(\lambda) = \expect G_L(\lambda)$ and write
+\begin{align*}
+&\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda) + T(\lambda) - G(\lambda)} \\
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)}.
+\end{align*}
+Now,
 \begin{align*}
@@
 \end{align*}
-where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.
-Since $\sabs{h_1(\lambda)} \leq 1$ and $\sabs{h_2(\lambda)} \leq 1$ for all $\lambda \in [-\pi, \pi)$, then $\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \rightarrow 0$ as $L\rightarrow \infty$.
-Now
-\begin{align*}
-\sup_{\lambda \in [-\pi, \pi)}&\sabs{G_L(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda) + T(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)}
-\end{align*}
-and Lemma~\ref{lem:uniflawTGL} shows that $\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0$ as $L\rightarrow \infty$.
+where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since $\sabs{h_1(\lambda)} \leq 1$ and $\sabs{h_2(\lambda)} \leq 1$ for all $\lambda \in [-\pi, \pi)$, it follows that 
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \rightarrow 0
+\]  
+almost surely as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0
+\]
+almost surely as $L\rightarrow \infty$.
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:uniflawTGL} 
-Put $T(x) = \expect G_L(x)$, then
-\[
-\sup_{x \in [-\pi, \pi)}\sabs{G_L(x) - T(x)} \rightarrow 0
+Put $T(\lambda) = \expect G_L(\lambda)$, then
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0
 \] 
@@
 \begin{IEEEproof}
-BLERG
-\end{IEEEproof}
+Put $D_L(\lambda) = G_L(\lambda) - T(\lambda)$ and let
+\[
+\lambda_n = \tfrac{2\pi}{N}(n-1) - \pi, \qquad n = 1, \dots, N
+\]
+be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n$ be the interval $[\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
+\begin{align*}
+\sup_{\lambda \in [-\pi, \pi)}&\sabs{D_L(\lambda)} \\
+&= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n) + D_L(\lambda_n)} \\
+&\leq A_L + B_L,
+\end{align*}
+where 
+\[
+A_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)},
+\]
+\[
+B_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
+\]
+Lemma~\ref{lem:supDLlambdan} shows that for any finite $N$, and any $\epsilon > 0$, 
+\[
+\prob\left\{ \lim_{L\rightarrow\infty} A_L > \epsilon \right\} = 0,
+\]
+i.e., $A_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
+\] 
+Choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ and
+\begin{align*}
+\prob& \left\{ \lim_{L\rightarrow\infty} \sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} > 3\epsilon \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty}(A_L + B_L) > 3\epsilon \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty} (A_L + B_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty} A_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} B_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+&= 0.
+\end{align*}
+Thus, $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:supDLlambdan}
+If $N$ is finite then $A_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Put
+\[
+Z_i(\lambda) = \begin{cases}
+R_i\cos(\lambda + \Phi_i) , & i \in P \\
+R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
+\end{cases}
+\]
+so that
+\[
+D_L(\lambda) = G_L(\lambda) - T(\lambda) = \frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
+\]
+Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
+\[
+\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
+\]
+almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Now, since $N$ is finite,
+\[
+A_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
+\]
+almost surely to zero as $L \rightarrow \infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+\]
+\end{lemma}
+\begin{IEEEproof}
+Observe that
+\begin{align*}
+&\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
+&\leq  \abs{G_L(\lambda) - \expect G_L(\lambda) - G_L(\lambda_n) + \expect G_L(\lambda_n)} \\
+&\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
+&\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
+\end{align*}
+the last line following from Jensens inequality.  Put
+\[
+C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) }
+\]
+then
+\[
+B_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \leq C_L + \expect C_L.
+\]
+because 
+\[
+ \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect \abs{G_L(\lambda) - G_L(\lambda_n)} \leq \expect C_L
+\]
+follows by moving expectation outside the supremum.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_i$ and also that
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+\]
+So,
+\begin{align*}
+\prob&\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+\end{align*}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:CL}
+$\expect C_L \leq \tfrac{2\pi}{N}\expect R_i$ and for any $\epsilon > 0$,
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+\]
+\end{lemma}
+\begin{IEEEproof}
+If $\lambda \in L_n$, then $\lambda = \lambda_n + \delta$ with $\delta < \tfrac{2\pi}{N}$, and from Lemma~\ref{lem:coslipshitz},
+\[
+\abs{\cos(\lambda) - \cos(\lambda_n)} \leq \tfrac{2\pi}{N} \;\;\; \text{and} \;\;\; \abs{\cos\sfracpart{\lambda} - \cos\sfracpart{\lambda_n}} \leq \tfrac{2\pi}{N}.
+\]
+Because this results does not depend on $n$,
+\[
+\sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda)} \leq R_i \frac{2\pi}{N}
+\]
+for all $i = 1, \dots, L$.  Now,
+\begin{align*}
+C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i=1}^L Z_i(\lambda) - Z_i(\lambda_n)} \\
+&\leq \frac{1}{L}\sum_{i=1}^L \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
+&\leq \frac{2\pi}{N L}\sum_{i=1}^L R_i \rightarrow \frac{2\pi}{N}\expect R_i
+\end{align*}
+almost surely to $\frac{2\pi}{N}\expect R_i$ as $L \rightarrow\infty$ by the strong law of large numbers, and so
+\begin{align*}
+\prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i=1}^L R_i > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+\end{align*}
+Also $\expect C_L \leq \frac{2\pi}{N}\expect R_i$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:coslipshitz}
+Let $x$ and $\delta$ be real numbers.  Then
+\[
+\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta},
+\]
+and 
+\[
+\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
+\]
+\end{lemma}
 

### 2012-08-08 paper.tex
more edits to the convergence section.  Starting edits to normality now.  Will need to, at some point, work out what the simulations section should say.

@@
 This complexity arises from the need to sort a list of $\abs{D}$ elements.
+%We use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
+
 %In Section~\ref{sec:line-time-algor} we will show that a full sort is not neccesary, and that the least squares estimator can be implemented in $O(L)$ operations.
@@
 \[
-\expect (\rho_0 - \hat{\rho})^2 = \frac{B}{L} + \rho_0^2(1 - G(0))^2 + o(1)
-\] 
-where $o(1)$ goes to zero as $L \rightarrow \infty$.
+\expect (\rho_0 - \hat{\rho})^2 = \frac{p B_1 + d B_2}{L} + \rho_0^2(1 - G(0))^2 + o(L^{-1/2}).
+\] 
 \end{corollary}
@@
 &= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
-&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L,
-\end{align*} 
-and the corollary holds since $\expect \hat{m}_L^2 = B/L + o(1)$ and $\expect\hat{m}_L = o(1)$ from Theorem~\ref{thm:normality}.
+&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L.
+\end{align*} 
+From Theorem~\ref{thm:normality} it follows that $\expect \hat{m}_L^2 = \tfrac{B}{L} + o(L^{-1})$ and that $\expect \hat{m}_L = o(L^{-1/2})$. 
 \end{IEEEproof}
@@
 
-The assumption that $\expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.
+The assumption that $\expect \sabs{w_1}^2 = \expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.
 This is required in Theorem~\ref{thm:normality} so that the constants $A_1$, $A_2$, $B_1$ and $B_2$ exist.
@@
 This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
-So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be \emph{close} together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and \emph{not} $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
+So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
 
 A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption requires that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence such that $\abs{G(x_i)} \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+Although we will not prove it here, this assumption is not only sufficient, but also nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
 The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
@@
 \end{lemma}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
-The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the probability density function of $X$ decreases as we move away from the origin.
+Before giving the proof we discuss the requirements of this lemma.
+The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the pdf of $X$ decreases as we move away from the origin.
 That is, the pdf of $X$ in rectangular coordinates is $p(x,y) = z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$, and this is required to be non increasing with $z$.
 For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
 
@@
 \begin{equation}\label{eq:hatrhoZ}
-\hat{\rho} = \frac{Z(\theta) + Z(\theta)^*}{2L} = \frac{1}{L}\Re(Z(\theta)),
+\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
 \end{equation}
@@
 y_i &= a_0 s_i + w_i \\
-&= \rho_0 s_i e^{j \theta_0}\left( 1 + \frac{1}{\rho_0} w_i s_i^* e^{-j \theta_0} \right) \\
+&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
+&= a_0 s_i R_i e^{j \Phi_i} \\
 &= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
@@
 \end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that when $i \in D$,
+where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
 \begin{align*}
@@
 So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
-Part one of Theorem~\ref{thm:consistency} follows.
-
-It remains to prove part two of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
+Part 1 of Theorem~\ref{thm:consistency} follows.
+
+It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
 \begin{equation}\label{eq:rhoGLZ}
@@
 \[
-0 \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)},
+0 \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)},
 \]
@@
 \[ 
-0 \leq \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)}.
+0 \leq \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}.
 \]
@@
 \begin{align*}
-0 &\leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} \\ 
-& \leq \sabs{G(\hat{\lambda}_L)} - \sabs{G(0)} + \sabs{G_L(0)} - \sabs{G_L(\hat{\lambda}_L)} \\
-& \leq |G(\hat{\lambda}_L) - G_L(\hat{\lambda}_L)| + |G_L(0) - G(0)| \\
-&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G(\lambda) - G_L(\lambda)},
+0 &\leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} \\ 
+& \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} + \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}\\
+& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
+&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
 \end{align*}
@@
 \end{align*}
-where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since $\sabs{h_1(\lambda)} \leq 1$ and $\sabs{h_2(\lambda)} \leq 1$ for all $\lambda \in [-\pi, \pi)$, it follows that 
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \rightarrow 0
-\]  
-almost surely as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
+where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since 
+\[
+h_1(\lambda) = \expect R_1\cos(\lambda + \Phi_1) \leq \expect R_1,
+\]
+and 
+\[
+h_2(\lambda) = \expect R_1\cos\fracpart{\lambda + \Phi_1} \leq \expect R_1
+\] 
+for all $\lambda \in [-\pi, \pi)$, it follows that 
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \leq o(1) \expect R_1 \rightarrow 0
+\]  
+as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
 \[
@@
 &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n) + D_L(\lambda_n)} \\
-&\leq A_L + B_L,
+&\leq U_L + V_L,
 \end{align*}
@@
 \[
-A_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)},
-\]
-\[
-B_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
+U_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)},
+\]
+\[
+V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
 \]
@@
 \[
-\prob\left\{ \lim_{L\rightarrow\infty} A_L > \epsilon \right\} = 0,
-\]
-i.e., $A_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
+\prob\left\{ \lim_{L\rightarrow\infty} U_L > \epsilon \right\} = 0,
+\]
+i.e., $U_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
 \] 
@@
 \prob& \left\{ \lim_{L\rightarrow\infty} \sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} > 3\epsilon \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty}(A_L + B_L) > 3\epsilon \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty} (A_L + B_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty} A_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} B_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty}(U_L + V_L) > 3\epsilon \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty} (U_L + V_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty} U_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
 &= 0.
@@
 \begin{lemma}\label{lem:supDLlambdan}
-If $N$ is finite then $A_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+If $N$ is finite then $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
 \end{lemma}
@@
 \[
-A_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
+U_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
 \]
@@
 \[
-\prob\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
 \]
@@
 &\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
-&\leq  \abs{G_L(\lambda) - \expect G_L(\lambda) - G_L(\lambda_n) + \expect G_L(\lambda_n)} \\
+&=  \abs{G_L(\lambda) - T(\lambda) - G_L(\lambda_n) + T(\lambda_n)} \\
 &\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
@@
 \[
-C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) }
-\]
-then
-\[
-B_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \leq C_L + \expect C_L.
+C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) },
+\]
+so that
+\[
+V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \leq C_L + \expect C_L,
 \]
@@
 \]
-follows by moving expectation outside the supremum.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_i$ and also that
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+follows by moving expectation outside the supremum.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
 \]
@@
 \begin{align*}
-\prob&\left\{ \lim_{L \rightarrow \infty} B_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
-\end{align*}
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:CL}
-$\expect C_L \leq \tfrac{2\pi}{N}\expect R_i$ and for any $\epsilon > 0$,
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
-\]
+\prob&\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+\end{align*}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:CL} The following statements hold:
+\begin{enumerate}
+\item $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$,
+\item for any $\epsilon > 0$, $\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0$.
+\end{enumerate}
 \end{lemma}
@@
 \[
-\abs{\cos(\lambda) - \cos(\lambda_n)} \leq \tfrac{2\pi}{N} \;\;\; \text{and} \;\;\; \abs{\cos\sfracpart{\lambda} - \cos\sfracpart{\lambda_n}} \leq \tfrac{2\pi}{N}.
+\abs{\cos(\lambda + \Phi_i) - \cos(\lambda_n + \Phi_i)} \leq \tfrac{2\pi}{N},
+\]
+and
+\[
+\abs{\cos\sfracpart{\lambda + \Phi_i} - \cos\sfracpart{\lambda_n + \Phi_i}} \leq \tfrac{2\pi}{N}.
 \]
@@
 \[
-\sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda)} \leq R_i \frac{2\pi}{N}
+\sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \leq R_i \frac{2\pi}{N}
 \]
@@
 &\leq \frac{1}{L}\sum_{i=1}^L \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
-&\leq \frac{2\pi}{N L}\sum_{i=1}^L R_i \rightarrow \frac{2\pi}{N}\expect R_i
-\end{align*}
-almost surely to $\frac{2\pi}{N}\expect R_i$ as $L \rightarrow\infty$ by the strong law of large numbers, and so
-\begin{align*}
-\prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i=1}^L R_i > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
-\end{align*}
-Also $\expect C_L \leq \frac{2\pi}{N}\expect R_i$.
+&\leq \frac{2\pi}{N L}\sum_{i=1}^L R_i .
+\end{align*}
+Thus, 
+\[
+\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i=1}^L R_i = \tfrac{2\pi}{N}\expect R_1
+\]
+and the first statement holds.  Now, 
+\[
+\frac{2\pi}{N L}\sum_{i=1}^L R_i \rightarrow \frac{2\pi}{N}\expect R_1
+\] 
+almost surely to $\frac{2\pi}{N}\expect R_1$ as $L \rightarrow\infty$ by the strong law of large numbers, and so, for any $\epsilon > 0$,
+\begin{align*}
+\prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i=1}^L R_i > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+\end{align*}
 \end{IEEEproof}
@@
 \end{lemma}
+\begin{IEEEproof}
+Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipshitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
+\[
+\abs{\cos(y) - \cos(x)} \leq K\abs{x-y} = \abs{x-y},
+\]
+and
+\[
+\abs{\cos\fracpart{y} - \cos\fracpart{x}} \leq K\abs{x-y} = \abs{x-y}.
+\]
+ The lemma follows by putting $y = x + \delta$.
+\end{IEEEproof}
 

### 2012-08-09 paper.tex
added to journal version.  Filling in normality lemmas.

@@
 If $\lambda \in L_n$, then $\lambda = \lambda_n + \delta$ with $\delta < \tfrac{2\pi}{N}$, and from Lemma~\ref{lem:coslipshitz},
-\[
-\abs{\cos(\lambda + \Phi_i) - \cos(\lambda_n + \Phi_i)} \leq \tfrac{2\pi}{N},
-\]
-and
-\[
-\abs{\cos\sfracpart{\lambda + \Phi_i} - \cos\sfracpart{\lambda_n + \Phi_i}} \leq \tfrac{2\pi}{N}.
-\]
+\begin{align*}
+&\abs{\cos(\lambda + \Phi_i) - \cos(\lambda_n + \Phi_i)} \leq \tfrac{2\pi}{N}, \;\;\; \text{and} \\
+&\abs{\cos\sfracpart{\lambda + \Phi_i} - \cos\sfracpart{\lambda_n + \Phi_i}} \leq \tfrac{2\pi}{N}.
+\end{align*}
 Because this results does not depend on $n$,
@@
 \]
-for all $i = 1, \dots, L$.  Now,
+for all $i = P \cup D$.  Now,
 \begin{align*}
@@
 Let $x$ and $\delta$ be real numbers.  Then
-\[
-\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta},
-\]
-and 
-\[
-\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
-\]
+\begin{align*}
+&\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta}, \;\;\; \text{and} \\
+&\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
+\end{align*}
 \end{lemma}
@@
 Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipshitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
-\[
-\abs{\cos(y) - \cos(x)} \leq K\abs{x-y} = \abs{x-y},
-\]
-and
-\[
-\abs{\cos\fracpart{y} - \cos\fracpart{x}} \leq K\abs{x-y} = \abs{x-y}.
-\]
+\begin{align*}
+&\abs{\cos(y) - \cos(x)} \leq K\abs{x-y} = \abs{x-y}, \;\;\; \text{and} \\
+&\abs{\cos\fracpart{y} - \cos\fracpart{x}} \leq K\abs{x-y} = \abs{x-y}.
+\end{align*}
  The lemma follows by putting $y = x + \delta$.
@@
 \begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
-% \begin{lemma}
-% \[
-% \sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{1}{2M} - \frac{1}{2LM ?}
-% \]
-% for all $i \in D$.
-% \end{lemma}
-% \begin{IEEEproof}
-% Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
+Observe that $G_L$ is differentiable at all points $\lambda \in [-\pi , \pi)$ those which staisfy $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i = P \cup D$.
+Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
+\[
+\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{2M} - \frac{\pi}{2LM ?}
+\]
+for all $i \in P \cup D$.  The proof follows since $y_i = 0$ with probability zero.  
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:fracpartlambdahatnotpi}
+\[
+\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{2M} - \frac{\pi}{2LM ?}
+\]
+for all $i \in D$.
+\end{lemma}
+\begin{IEEEproof}
+Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
 The proof now proceeds by contradiction.
-Assume that $\sfracpart{\hat{\lambda}_L + \Phi_i} \leq \frac{1}{2M} - \frac{1}{2LM ?}$ for some $k$.
+Assume that $\sfracpart{\hat{\lambda}_L + \Phi_i} > \frac{1}{2M} - \frac{1}{2LM ?}$ for some $k$.
 Put $r_i = \hat{d}_i e_k$, then
-% \begin{align*}
-% Y(\{ r_i, i \in D\}) = \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) = \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
-% \end{align*}
-% where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
-% \begin{align*}
-% SS&(\{r_i, i \in D\}) \\
-% &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
-% &= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
-% &= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
-% &= SS(\{\hat{d}_i, i \in D\}) - R
-% \end{align*}
-% where 
-% \[
-% R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
-% \]
-% We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-
-% To see that $R$ is positive write
-%  \begin{align*}
-% R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
-%  &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
-%  \end{align*}
-%  so it is sufficient to show that
-%  \[
-%  \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
-%  \]
-%  Unfinished, but this proof will work
-% \end{IEEEproof}
+\begin{align*}
+Y(\{ r_i, i \in D\}) = \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) = \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
+\end{align*}
+where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
+\begin{align*}
+SS&(\{r_i, i \in D\}) \\
+&= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
+&= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
+&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
+&= SS(\{\hat{d}_i, i \in D\}) - R
+\end{align*}
+where 
+\[
+R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
+\]
+We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+
+To see that $R$ is positive write
+ \begin{align*}
+R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
+ &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
+ \end{align*}
+ so it is sufficient to show that
+ \[
+ \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
+ \]
+ Unfinished, but this proof will work.  BLERG.
+\end{IEEEproof}
 
@@
 k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
-&= \lambda \big( \expect R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \expect R_1 \big) \\
-&= \lambda \big( 1 \; + \; \zeta(\lambda) K \big),
-\end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ has zero mean, and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.
+&= \lambda \big( \expect R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \expect R_1 \big) 
+\end{align*}
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ has zero mean.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 \begin{IEEEproof}
-As $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $k_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
+As $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
 We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
-An analagous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
+An analagous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
 To keep our notation clean put
@@
 \end{align*}
-where, for each integer $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+where, for each $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
 \begin{align*}
@@
 \[
-\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_{k+1}(\lambda).
-\]
-Observe that $s_k$ is continuous and $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$.
+\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos\fracpart{\Phi_1} + \zeta_{k+1}(\lambda).
+\]
+Observe that both $s_k$ and $\eta_k$ are continuous and that $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $\eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda)$.
 
@@
 \end{align}
-Now, when $0 < k < M$, we have $0 < \psi_{k-1} < \psi_k < 2\pi$, so, using~\eqref{eq:vk},
+Now, when $0 < k < M$, we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
 \begin{align*}
@@
 \[
-\gamma(\lambda) = A(\lambda) + B(\lambda) + D(\lambda) - H_1.
+\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1.
 \]
@@
 where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
-\begin{equation}\label{eq:vk}
+\begin{equation}\label{eq:cvk}
 v_k(\lambda,\Phi_1) = \begin{cases}

### 2012-08-10 paper.tex
modified some notation.  It's important that I put the variances etc in the right place.
Added a bit more the compamp.  The differentiability proof is a bit fiddly!

@@
 where $*$ denotes the complex congugate.
-Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find that the least squares estimator of $a_0$, satisfies
+Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find the least squares estimator of $a_0$ as a function of $\{d_i, i \in D\}$,
 \begin{equation}\label{eq:hata}
@@
 \end{align*}
-where $\eta = (e^{j2\pi/M} - 1)$.  In general,
+where $\eta = e^{j2\pi/M} - 1$.  In general,
 \[
@@
 \[
-A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
+A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \var\big( R_1^2\sin^2\fracpart{\Phi_1} \big),
 \]
@@
 where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
-\begin{align*}
- y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \\
-&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \\
-&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}}
-\end{align*}
+\begin{align}
+ y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
+&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
+&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
+\end{align}
 since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
@@
 \[
-G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
-\]
-\end{lemma}
-\begin{IEEEproof}
-Observe that $G_L$ is differentiable at all points $\lambda \in [-\pi , \pi)$ those which staisfy $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i = P \cup D$.
+G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0,
+\]
+\end{lemma}
+\begin{IEEEproof}
+Observe that 
+\[
+G_L(\lambda) = \frac{1}{L}\sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}
+\] 
+is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ and $R_i > 0$ for any $i \in D$.  Let $q$ be the largest negative real number such that
+\[
+\sin(q) > -\frac{\rho_o \sabs{\eta}^2 R_{\text{min}}}{4 L \hat{\rho} \sin(\pi/M)}
+\]
+where $\eta = e^{-j2\pi/M} - 1$ and $R_{\text{min}}$ is the smallest strictly postive value of $R_i,$ for $i \in D$, that is
+\[
+R_{\text{min}} = \min_{i \in D, R_i \neq 0} R_i.
+\]  
 Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
 \[
-\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{2M} - \frac{\pi}{2LM ?}
-\]
-for all $i \in P \cup D$.  The proof follows since $y_i = 0$ with probability zero.  
+\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q
+\]
+for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq \tfrac{\pi}{M}$ if $R_i > 0$.
 \end{IEEEproof}
@@
 \[
-\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{2M} - \frac{\pi}{2LM ?}
+\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} - \frac{\pi}{2LM ?}
 \]
@@
 \begin{IEEEproof}
-Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta})$ are the minimsers of $SS$.
-The proof now proceeds by contradiction.
-Assume that $\sfracpart{\hat{\lambda}_L + \Phi_i} > \frac{1}{2M} - \frac{1}{2LM ?}$ for some $k$.
-Put $r_i = \hat{d}_i e_k$, then
-\begin{align*}
-Y(\{ r_i, i \in D\}) = \frac{1}{L}\left( B + \sum_{i \in D} y_i\hat{r}_i^* \right) = \hat{Y} + \frac{1}{L}\eta y_k\hat{d}_k^*.
-\end{align*}
-where we put $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
+Recall that 
+\[
+\hat{d}_i = \hat{d}_i(\hat{\theta}) = e^{j\round{\angle y_i - \hat{\theta}}}, \;\;\; i \in D
+\] 
+defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
+\[
+SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y(\{d_i, i \in D\})}^2,
+\]
+defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
+\begin{equation}\label{eq:fraclambassumption}
+\sfracpart{\hat{\lambda}_L + \Phi_i} > \frac{\pi}{M} - ?
+\end{equation}
+for some $k$ for which $\abs{R_k} > 0$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$, then
+\begin{align*}
+Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*.
+\end{align*}
+where $\eta = e^{-j2\pi/M} - 1$ and $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
 \begin{align*}
@@
 &= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
-&= A - \frac{1}{L} \abs{ \hat{Y} + \frac{1}{L} \eta y_k\hat{d}_k^* }^2 \\
-&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2\\
-&= SS(\{\hat{d}_i, i \in D\}) - R
+&= A - \frac{1}{L} \abs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
+&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\abs{ \eta y_k}^2\\
+&= SS(\{\hat{d}_i, i \in D\}) - D
 \end{align*}
@@
 \[
-R = \frac{2}{L^2}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta y_k \hat{d}_k^* }^2.
-\]
-We will show that $R$ is positive, from which is will follow that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-
-To see that $R$ is positive write
- \begin{align*}
-R &\geq  \frac{2}{L^2}\abs{\hat{Y} y_k}\Re\left(\eta e^{j(\phi_i - \hat{\theta})}\hat{d}_k^* \right) +  \frac{1}{L^3}\abs{ \eta}^2 \abs{ y_k }^2 \\
- &= \frac{1}{L^2}\abs{y_k} \left( 2\abs{\hat{Y}}\Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) + \frac{1}{L}\abs{ \eta}^2 \abs{ y_k }  \right) 
- \end{align*}
- so it is sufficient to show that
- \[
- \Re\left(\eta e^{j(\phi_i - \hat{\theta})} \hat{d}_k^* \right) > - \frac{\abs{ \eta}^2 \abs{ y_k }}{2 L \abs{\hat{Y}}}.
- \]
- Unfinished, but this proof will work.  BLERG.
+D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\abs{ \eta y_k }^2.
+\]
+Now $\frac{1}{L}\hat{Y} = \hat{a} = \hat{\rho} e^{j\hat{\theta}}$ from~\eqref{eq:hata} and using~\eqref{eq:yethetadhat},
+\[
+\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}
+\]
+so that
+\[
+D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\abs{\eta}^2\rho_0^2 R_k^2.
+\]
+Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
+\begin{align*}
+\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
+&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
+&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) }
+\end{align*}
+and
+\begin{align*}
+\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
+ &= 2 \sin(\tfrac{\pi}{M}) \sin(v)
+\end{align*}
+Because we assumed~\eqref{eq:fraclambassumption},
+\[
+0 > v > -?
+\]
+and therefore
+\begin{align*}
+0 &> 2 \sin(\tfrac{\pi}{M}) \sin(v) \\
+&> -\frac{\rho_0 R_k \abs{\eta}^2}{2L\hat{\rho}}
+\end{align*}
+
+So $D$ is positive.  But then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+
 \end{IEEEproof}

### 2012-08-10 paper.tex
further minor edits.

@@
 %\documentclass[a4paper,10pt]{article}
-%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[journal]{IEEEtran}
+\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+%\documentclass[journal]{IEEEtran}
 
@@
 \] 
-is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ and $R_i > 0$ for any $i \in D$.  Let $q$ be the largest negative real number such that
-\[
-\sin(q) > -\frac{\rho_o \sabs{\eta}^2 R_{\text{min}}}{4 L \hat{\rho} \sin(\pi/M)}
-\]
-where $\eta = e^{-j2\pi/M} - 1$ and $R_{\text{min}}$ is the smallest strictly postive value of $R_i,$ for $i \in D$, that is
-\[
-R_{\text{min}} = \min_{i \in D, R_i \neq 0} R_i.
-\]  
-Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
-\[
-\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q
-\]
-for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq \tfrac{\pi}{M}$ if $R_i > 0$.
+is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ where $R_i > 0$.  Let $q_i$ be the smallest number from $[-\tfrac{\pi}{M}, 0]$ such that
+\[
+ \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_k}{4 L \hat{\rho} \sin(\pi/M)}
+\]
+where $\eta = e^{-j2\pi/M} - 1$.  Observe that if $R_i > 0$ then $q_i < 0$.  Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
+\[
+\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i
+\]
+for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ if $R_i > 0$.  The lemma follows.
 \end{IEEEproof}

### 2012-08-10 paper.tex
Added a bit more to diff proof.  Looking good.

@@
 %\documentclass[a4paper,10pt]{article}
-\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[journal]{IEEEtran}
+%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
+\documentclass[journal]{IEEEtran}
 
@@
 The assumption that $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
-If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $s_1, \dots, s_L$.
-In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $s_1, \dots, s_L$, rather than just $\rho_0$.
+If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $\{s_i,i \in P \cup D\}$.
+In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $\{s_i,i \in P \cup D\}$, rather than just $\rho_0$.
 The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
@@
 \] 
-is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ where $R_i > 0$.  Let $q_i$ be the smallest number from $[-\tfrac{\pi}{M}, 0]$ such that
-\[
- \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_k}{4 L \hat{\rho} \sin(\pi/M)}
+is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ where $R_i > 0$.
+Let $q_i$ be the smallest number from the interval $[-\tfrac{\pi}{M}, 0]$ such that
+\[
+ \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
 \]
@@
 
-\begin{lemma}\label{lem:fracpartlambdahatnotpi}
-\[
-\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} - \frac{\pi}{2LM ?}
-\]
-for all $i \in D$.
-\end{lemma}
-\begin{IEEEproof}
-Recall that 
-\[
-\hat{d}_i = \hat{d}_i(\hat{\theta}) = e^{j\round{\angle y_i - \hat{\theta}}}, \;\;\; i \in D
-\] 
-defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
+\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i, i \in D$ be defined as in Lemma~\ref{lem:diffatlambdaL}.
+Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.
+\end{lemma}
+\begin{IEEEproof}
+Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D$ defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
 \[
@@
 \begin{equation}\label{eq:fraclambassumption}
-\sfracpart{\hat{\lambda}_L + \Phi_i} > \frac{\pi}{M} - ?
-\end{equation}
-for some $k$ for which $\abs{R_k} > 0$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$, then
+\sfracpart{\hat{\lambda}_L + \Phi_k} > \frac{\pi}{M} + q_k
+\end{equation}
+for some $k \in D$.
+Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.
+We will show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+First observe that,
 \begin{align*}
@@
 \[
-\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}
+\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}},
 \]
 so that
-\[
+\begin{equation}\label{eq:Ddefn}
 D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\abs{\eta}^2\rho_0^2 R_k^2.
-\]
+\end{equation}
 Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
@@
 \[
-0 > v > -?
-\]
-and therefore
-\begin{align*}
-0 &> 2 \sin(\tfrac{\pi}{M}) \sin(v) \\
-&> -\frac{\rho_0 R_k \abs{\eta}^2}{2L\hat{\rho}}
-\end{align*}
-
-So $D$ is positive.  But then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+0 > v > q_k
+\]
+and, from the definition of $q_k$,
+\[
+0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
+\]
+Subtituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+Thus,~\eqref{eq:fraclambassumption} is false.
+
+To show that $\sfracpart{\hat{\lambda}_L + \Phi_k} \geq -\frac{\pi}{M} - q_k$ we use contradiction again.
+Assume that $\sfracpart{\hat{\lambda}_L + \Phi_k} < -\frac{\pi}{M} - q_k$.
+Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.
+Now an argument analagous to that used previously shows that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
 

### 2012-08-12 paper.tex
minor edits
Added to paper.  Minor code edits.

@@
 \]
-where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M$ is the size of the constellation.
+where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M \geq 2$ is the size of the constellation.
 We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.
@@
 \frac{pA_1 + dA_2}{(p + H d)^2} & 0 \\
-0 & pB_1 + dB_2
+0 & \rho_0^2(pB_1 + dB_2)
 \end{array} \right)
@@
 \[
-A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \var\big( R_1^2\sin^2\fracpart{\Phi_1} \big),
-\]
-\[
-B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
+A_1 = \var\big( R_1^2\sin^2(\Phi_1) \big), \qquad A_2 = \var\big( R_1^2\sin^2\fracpart{\Phi_1} \big),
+\]
+\[
+B_1 = \var\big( R_1^2 \cos^2(\Phi_1) \big), \;\;\; B_2 = \var\big( R_1^2 \cos^2\fracpart{\Phi_1} \big).
 \]
@@
 \] 
-is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ where $R_i > 0$.
+is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ with $R_i > 0$.
 Let $q_i$ be the smallest number from the interval $[-\tfrac{\pi}{M}, 0]$ such that
@@
 \begin{IEEEproof}
-Recall that $\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D$ defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
-\[
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y(\{d_i, i \in D\})}^2,
+Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
+\[
+SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
 \]
@@
 SS&(\{r_i, i \in D\}) \\
-&= A - \frac{1}{L} \abs{ Y(\{ r_i, i \in D\}) }^2 \\
-&= A - \frac{1}{L} \abs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
-&= A - \frac{1}{L}\abs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\abs{ \eta y_k}^2\\
+&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
+&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
+&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
 &= SS(\{\hat{d}_i, i \in D\}) - D
@@
 \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
- &= 2 \sin(\tfrac{\pi}{M}) \sin(v)
-\end{align*}
-Because we assumed~\eqref{eq:fraclambassumption},
-\[
-0 > v > q_k
-\]
-and, from the definition of $q_k$,
+ &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
+\end{align*}
+Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
 \[
@@
 Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.
-Now an argument analagous to that used previously shows that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
+Now an analagous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
 
@@
 
-\begin{lemma}\label{lem:empiricprocc} BLERG
+\begin{lemma}\label{lem:empiricprocc}
 \[
@@
 \begin{IEEEproof}
-
+Put 
+\[
+Z_L(\lambda) = \sqrt{L}\big(G_L^\prime(\lambda) - \expect G_L^\prime(\lambda) - G_L^\prime(0) + \expect G_L^\prime(0)\big)
+\] 
+so that
+\[
+\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = Z_L(\hat{\lambda}_L) + \sqrt{L} G_L^\prime(0)
+\]
+Lemma~? shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+\[
+\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Z_L(\lambda)} > \delta  \right\} < \nu
+\]
+for all positive integers $L$.  Since $\hat{\lambda}_L$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
+\[
+\lim_{L\rightarrow\infty}\prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} = 0
+\] 
+and therefore $\prob\{ \sabs{\hat{\lambda}_L} \geq \epsilon \} < \nu$ for all sufficiently large $L$.  Now
+\begin{align*}
+  \prob&\left\{\sabs{ Z_L(\hat{\lambda}_L) } > \delta \right\} \\
+&= \prob\left\{ \sabs{Z_L(\hat{\lambda}_L)} > \delta \;\text{and} \; \sabs{\hat{\lambda}_L} < \epsilon \right\} \\
+&\hspace{0.7cm} + \prob\left\{ \sabs{Z_L(\lambda_L)} > \delta  \;\text{and} \; \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\lambda < \epsilon} \sabs{ Z_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+&\leq 2\nu.
+\end{align*}
+for all sufficiently large $L$.  Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $Z_L(\hat{\lambda}_L)$ converges in probability to zero as $N\rightarrow\infty$.
 \end{IEEEproof}

### 2012-08-12 code/scala/test/cam/noise/CLTComputerTest.scala

@@ -20,10 +20,6 @@ class CLTComputerTest {
   }
 
   @Test
-  def example = {
-  }
-  
-  @Test
   def testGaussianCLT = {
     val M = 4
     val p = 0.1

### 2012-08-12 paper.tex
Added paper.

@@
 This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$, so
-\begin{align*}
-G_L^\prime(\hat{\lambda}_L) &= 0 \\
-&= \frac{1}{L} \sum_{i \in P} R_i \sin(\hat{\lambda}_L + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\hat{\lambda}_L + \Phi_i}.
-\end{align*}
-Let $Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0)$ and write
-\begin{align*}
-0 &= G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) + Q(\hat{\lambda}_L) \\
-&= \sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) + \sqrt{L}Q(\hat{\lambda}_L).
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
+Define the function
+\[
+R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+\]
+Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
+Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
+\begin{align*}
+0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
+&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
 \end{align*}
@@
 \[
-\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
 \]
@@
 \[
-\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} G_L^\prime(0).
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
 \]
@@
 \[
-0 = o_P(1) + \sqrt{L}G_L^\prime(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
+0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
 \]
@@
 \[
-\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}G_L'(0)}{p + Hd  + o_P(1)}.
-\]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
+\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
+\]
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
 It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
@@
 \begin{lemma}\label{lem:Qconv}
-$\sqrt{L} Q(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)$
+$\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)$
 \end{lemma}
@@
 \[
-Q(\lambda) = \expect G_L^\prime(\lambda) - \expect G_L^\prime(0) = \tfrac{\abs{P}}{L} k_1(\lambda) + \tfrac{\abs{D}}{L} k_2(\lambda)
+Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0) = \tfrac{\abs{P}}{L} k_1(\lambda) + \tfrac{\abs{D}}{L} k_2(\lambda)
 \]
@@
 \begin{align*}
-Q(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
+Q_L(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
 &= \hat{\lambda}_L \left( \tfrac{\abs{P}}{L}  + \tfrac{\abs{D}}{L} H + o_P(1) \right) \\
@@
 \[
-Z_L(\lambda) = \sqrt{L}\big(G_L^\prime(\lambda) - \expect G_L^\prime(\lambda) - G_L^\prime(0) + \expect G_L^\prime(0)\big)
+Z_L(\lambda) = \sqrt{L}\big( R_L(\lambda) - \expect R_L(\lambda) - R_L(0) + \expect R_L(0)\big)
 \] 
@@
 \[
-\sqrt{L}\big( G_L^\prime(\hat{\lambda}_L) - Q(\hat{\lambda}_L) \big) = Z_L(\hat{\lambda}_L) + \sqrt{L} G_L^\prime(0)
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = Z_L(\hat{\lambda}_L) + \sqrt{L} R_L(0)
 \]

### 2012-08-13 paper.tex
Added to paper.  Starting the empirical process proofs.  It's been moved to the appendix.

@@
 \title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
-\author{Robby McKilliam, Andre Pollok, Bill Cowley}
+\author{Robby McKilliam, Andre Pollok, Bill Cowley 
+\thanks{
+Supported under the Australian Government’s Australian Space Research Program.
+Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+}}
 
@@
 The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
-\[
+\begin{equation}\label{eq:expectRepartRphi}
 \expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
-\]
+\end{equation}
 where $\Re(\cdot)$ denotes the real part, and
-\[
+\begin{equation}\label{eq:expectImpartRphi}
 \expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
-\]
+\end{equation}
 where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
@@
 \[
-\expect (\rho_0 - \hat{\rho})^2 = \frac{p B_1 + d B_2}{L} + \rho_0^2(1 - G(0))^2 + o(L^{-1/2}).
+\expect (\rho_0 - \hat{\rho})^2 = \rho_0^2\frac{p B_1 + d B_2}{L} + \rho_0^2(1 - G(0))^2 + o(L^{-1/2}).
 \] 
@@
 \begin{IEEEproof}
-Put $T(\lambda) = \expect G_L(\lambda)$ and write
+Put $T_L(\lambda) = \expect G_L(\lambda)$ and write
 \begin{align*}
 &\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda) + T(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)}.
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda) + T_L(\lambda) - G(\lambda)} \\
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)}.
 \end{align*}
@@
 \begin{align*}
-T(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
+T_L(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
 &= o(1) h_1(\lambda) + o(1) h_2(\lambda)
@@
 \[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{T(\lambda) - G(\lambda)} \leq o(1) \expect R_1 \rightarrow 0
+\sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)} \leq o(1) \expect R_1 \rightarrow 0
 \]  
@@
 \[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
 \]
@@
 \begin{lemma}\label{lem:uniflawTGL} 
-Put $T(\lambda) = \expect G_L(\lambda)$, then
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T(\lambda)} \rightarrow 0
+Put $T_L(\lambda) = \expect G_L(\lambda)$, then
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
 \] 
@@
 \begin{IEEEproof}
-Put $D_L(\lambda) = G_L(\lambda) - T(\lambda)$ and let
+Put $D_L(\lambda) = G_L(\lambda) - T_L(\lambda)$ and let
 \[
@@
 \[
-D_L(\lambda) = G_L(\lambda) - T(\lambda) = \frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
+D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
 \]
@@
 &\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
-&=  \abs{G_L(\lambda) - T(\lambda) - G_L(\lambda_n) + T(\lambda_n)} \\
+&=  \abs{G_L(\lambda) - T_L(\lambda) - G_L(\lambda_n) + T_L(\lambda_n)} \\
 &\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
@@
 Define the function
-\[
+\begin{equation}\label{eq:RLdef}
 R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\]
+\end{equation}
 Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
@@
 
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
+We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
 \begin{align*}
 \sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L) + T(\hat{\lambda}_L) - G(0) \big).
+&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
 \end{align*}
@@
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - T(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-%\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = o_P(1).
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
+\]
+where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
+\[
+%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
 \]
@@
 %\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L} m_L =  X_L + o_P(1). 
-\]
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance
-\[
- p B_1 + d B_2 
-\]
-where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.  As $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$.  Thus,
-\[
-\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
+\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
+\]
+Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
+Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
+Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
+\[
+\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
 \]
@@
 \]
-for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ if $R_i > 0$.  The lemma follows.
+for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.
+
+It remains to show that $G_L^\prime(\hat{\lambda}_L) = 0$.
+If $\sabs{G_L(\hat{\lambda}_L)} > 0$ then $\sabs{G_L(\lambda)}$ is differentiable at $\hat{\lambda}_L$ and, since $\hat{\lambda}_L$ is a maximiser of $\sabs{G_L(\lambda)}$,
+\[
+0 = \frac{d \sabs{G_L(\lambda)}}{d \lambda}(\hat{\lambda}_L) = \begin{cases} 
+G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) > 0\\
+-G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) < 0.
+\end{cases}
+\]
+Thus, $G_L^\prime(\hat{\lambda}_L) = 0$ if $\sabs{G_L(\hat{\lambda}_L)} > 0$.
+On the other hand $\sabs{G_L(\hat{\lambda}_L)} = 0$ only if $G_L(\lambda) = 0$ for all $\lambda \in \reals$, inwhich case $G_L^\prime(\hat{\lambda}_L) = 0$.
 \end{IEEEproof}
@@
 \[
-D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\abs{ \eta y_k }^2.
+D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
 \]
@@
 \begin{equation}\label{eq:Ddefn}
-D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\abs{\eta}^2\rho_0^2 R_k^2.
+D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
 \end{equation}
@@
 \begin{lemma}\label{lem:Qconv}
-$\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)$
+Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
+\[ 
+\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\]
+where $p, d$ and $H$ are real numbers defined in the statements of Thereoms~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
 \end{lemma}
@@
 \[
-Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0) = \tfrac{\abs{P}}{L} k_1(\lambda) + \tfrac{\abs{D}}{L} k_2(\lambda)
+Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0) = \tfrac{\sabs{P}}{L} k_1(\lambda) + \tfrac{\sabs{D}}{L} k_2(\lambda)
 \]
@@
 k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
-&= \lambda \big( \expect R_1\cos(\Phi_1) \; + \; \zeta(\lambda) \expect R_1 \big) 
+&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big) 
 \end{align*}
@@
 \[
-k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) K \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big).
-\]
-Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
-\begin{align*}
-Q_L(\hat{\lambda}_L) &=  \tfrac{\abs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\abs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
-&= \hat{\lambda}_L \left( \tfrac{\abs{P}}{L}  + \tfrac{\abs{D}}{L} H + o_P(1) \right) \\
+k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) \expect R_1 \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big)
+\]
+because $\expect R_1$ is finite.  Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
+\begin{align*}
+Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
+&= \hat{\lambda}_L \left( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \right) \\
 &= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
 \end{align*}
-since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.
+since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the above equation by $\sqrt{L}$.
 \end{IEEEproof}
@@
 \begin{IEEEproof}
-As $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
+Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
 We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
@@
 \]
-Observe that both $s_k$ and $\eta_k$ are continuous and that $s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M})$ and $\eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda)$.
+Observe that both $s_k$ and $\eta_k$ are continuous and that 
+\[
+s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
+\]
 
@@
 \end{align}
-Now, when $0 < k < M$, we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
+Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
 \begin{align*}
@@
 \]
-as $\lambda$ goes to zero from above\footnote{We actually only require $g$ to be continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$, and the requirements of Theorem~\ref{thm:normality} could be weakened in this sense.
-Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.}.
+as $\lambda$ goes to zero from above.
 We are only interested in the limit from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
 The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
+We actually only require $g$ to be continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$, and the requirements of Theorem~\ref{thm:normality} could be weakened in this sense.
+Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.
 
@@
 
-\begin{lemma}\label{lem:empiricprocc}
+\begin{lemma}\label{lem:empiricprocc} Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
 \[
@@
 Put 
-\[
-Z_L(\lambda) = \sqrt{L}\big( R_L(\lambda) - \expect R_L(\lambda) - R_L(0) + \expect R_L(0)\big)
-\] 
+\begin{equation}\label{eq:WLdef}
+W_L(\lambda) = \sqrt{L}\big( R_L(\lambda) - Q_L(\lambda) - R_L(0) \big)
+\end{equation}
 so that
 \[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = Z_L(\hat{\lambda}_L) + \sqrt{L} R_L(0)
-\]
-Lemma~? shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
-\[
-\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Z_L(\lambda)} > \delta  \right\} < \nu
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = W_L(\hat{\lambda}_L) + \sqrt{L} R_L(0)
+\]
+Lemma~\ref{lem:ZYLempiricproc} in the Appendix shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+\[
+\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu
 \]
@@
 \begin{align*}
-  \prob&\left\{\sabs{ Z_L(\hat{\lambda}_L) } > \delta \right\} \\
-&= \prob\left\{ \sabs{Z_L(\hat{\lambda}_L)} > \delta \;\text{and} \; \sabs{\hat{\lambda}_L} < \epsilon \right\} \\
-&\hspace{0.7cm} + \prob\left\{ \sabs{Z_L(\lambda_L)} > \delta  \;\text{and} \; \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
-&\leq \prob\left\{  \sup_{\lambda < \epsilon} \sabs{ Z_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+  \prob&\left\{\sabs{ W_L(\hat{\lambda}_L) } > \delta \right\} \\
+&= \prob\left\{ \sabs{W_L(\hat{\lambda}_L)} > \delta \;\text{and} \; \sabs{\hat{\lambda}_L} < \epsilon \right\} \\
+&\hspace{0.7cm} + \prob\left\{ \sabs{W_L(\lambda_L)} > \delta  \;\text{and} \; \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\lambda < \epsilon} \sabs{ W_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
 &\leq 2\nu.
 \end{align*}
-for all sufficiently large $L$.  Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $Z_L(\hat{\lambda}_L)$ converges in probability to zero as $N\rightarrow\infty$.
+for all sufficiently large $L$.  Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $W_L(\hat{\lambda}_L)$ converges in probability to zero as $N\rightarrow\infty$.
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:convdistGLdash}
-The distribution of $\sqrt{L}G_L'(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$.
+The distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$.
 \end{lemma}
@@
 \[
-\sqrt{L} G_L^\prime(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
+\sqrt{L} R_L(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
 \]
@@
 \[
-\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) = \sqrt{\frac{\abs{P}}{L}} \frac{1}{\sqrt{\abs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
-\]
-converges to the normal with mean $\sqrt{p} k_1(0) = 0$ and variance
+\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) = \sqrt{\frac{\sabs{P}}{L}} \frac{1}{\sqrt{\sabs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
+\]
+converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
 \[
@@
 \[ 
-\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} = \sqrt{\frac{\abs{D}}{L}} \frac{1}{\sqrt{\abs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
-\]
-converges to the normal with mean $\sqrt{d} k_2(0) = 0$ and variance
+\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} = \sqrt{\frac{\sabs{D}}{L}} \frac{1}{\sqrt{\sabs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
+\]
+converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~? BLERG, and variance
 \[
@@
 
-\begin{lemma}\label{lem:empiricprocforrho} BLERG
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T(\hat{\lambda}_L) \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L} \big( G_L(0) - T(0) \big)$.
-\end{lemma}
-\begin{IEEEproof}
-The is the empirical process proof!
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:HLtoG} BLERG
-\[
-\sqrt{L}\big( T(\hat{\lambda}_L) - G(0) \big) = o_P(1).
-\]
+\begin{lemma}\label{lem:empiricprocforrho} Let $T_L(\lambda) = \expect G_L(\lambda)$.  Then
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = o_P(1) + X_L,
+\]
+where $X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big)$.
+\end{lemma}
+\begin{IEEEproof}
+Let
+\begin{equation}\label{eq:YLdef}
+Y_L(\lambda) = \sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) - X_L
+\end{equation}
+so that
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\lambda) + X_L
+\]
+Lemma~\ref{lem:ZYLempiricproc} in the appendix 
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:HLtoG} $\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1)$.
 \end{lemma}
@@
 \begin{align*}
-T(\lambda) &= \frac{\abs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\abs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
+T_L(\lambda) &= \frac{\sabs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\sabs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
 %&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
@@
 \end{align*}
-because $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L\rightarrow\infty$. Put
-\begin{align*}
-L(\lambda) &= \sqrt{L}\big( T(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
+because $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L\rightarrow\infty$. Put
+\begin{align*}
+L(\lambda) &= \sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
 \end{align*}
@@
 \end{align*}
-since $\expect R_1\sin(\Phi_1) = 0$ because $w_1$ has zero mean,  and $K = \expect R_1$ is a finite constant because $\expect\abs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.  So, 
+since $\expect R_1\sin(\Phi_1) = 0$ because $w_1$ has zero mean,  and $K = \expect R_1$ is a finite constant because $\expect\sabs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.  So, 
 \begin{align*}
@@
 \end{align*}
-because $\sqrt{L} \hat{\lambda}_L$ converges in distribution and $\zeta(\hat{\lambda}_L) = o_P(1)$.  Lemma~\ref{} shows that $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$ also.  Thus 
+because $\sqrt{L} \hat{\lambda}_L$ converges in distribution and $\zeta(\hat{\lambda}_L) = o_P(1)$.
+An argument analogous to that used in Lemma~\ref{lem:k2conv} shows that $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
+Thus,
 \[
@@
 
-\begin{lemma}\label{lem:XL} 
-$\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
-\end{lemma}
-\begin{IEEEproof}
-The argument is similar to that used in Lemma~\ref{lem:k2conv}.
+% \begin{lemma}\label{lem:XL} 
+% $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
+% \end{lemma}
+% \begin{IEEEproof}
+% The argument is similar to that used in Lemma~\ref{lem:k2conv}.
 As $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $q_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
@@
 Recall that
-\[
-\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
-\]
-when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
-\begin{align*}
-\cos\fracpart{\lambda + \Phi_1} - &\cos\fracpart{\Phi_1} \\
-&= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
-\end{align*}
-and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
-\begin{align*}
-\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} &= -\lambda\big(\sin(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
-&= -\lambda\big(\sin\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
-\end{align*}
-where, for each integer $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.   Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
-\begin{align*}
-\cos\fracpart{\lambda + \Phi_1} &- \cos\fracpart{\Phi_1} \\
-&= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k)
-\end{align*}
-and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
-\begin{align*}
-\cos&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&= c_k(\Phi_1) - \lambda\big(\sin\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
-\end{align*}
-where
-\[
-c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
-\]
-and
-\[
-\eta_k(\lambda, \Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_{k+1}(\lambda).
-\]
-Observe that $c_k$ is continuous and $c_k(\psi_{k}) = 0$.
-
-Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
-\[
-\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
-\]
-whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
-\[
-\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda \sin\fracpart{\Phi_1} + c_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
-\]
-Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
-\[
-\cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
-\]
-where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
-\begin{equation}\label{eq:cvk}
-v_k(\lambda,\Phi_1) = \begin{cases}
-c_k(\Phi_1) - \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
-\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
-0 & \text{otherwise}.
-\end{cases}
-\end{equation}
-\end{IEEEproof}
+% \[
+% \psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
+% \]
+% when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
+% \begin{align*}
+% \cos\fracpart{\lambda + \Phi_1} - &\cos\fracpart{\Phi_1} \\
+% &= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
+% \end{align*}
+% and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
+% \begin{align*}
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} &= -\lambda\big(\sin(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
+% &= -\lambda\big(\sin\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
+% \end{align*}
+% where, for each integer $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.   Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+% \begin{align*}
+% \cos\fracpart{\lambda + \Phi_1} &- \cos\fracpart{\Phi_1} \\
+% &= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k)
+% \end{align*}
+% and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
+% \begin{align*}
+% \cos&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
+% &= c_k(\Phi_1) - \lambda\big(\sin\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
+% \end{align*}
+% where
+% \[
+% c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
+% \]
+% and
+% \[
+% \eta_k(\lambda, \Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_{k+1}(\lambda).
+% \]
+% Observe that $c_k$ is continuous and $c_k(\psi_{k}) = 0$.
+
+% Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
+% \[
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
+% \]
+% whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
+% \[
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda \sin\fracpart{\Phi_1} + c_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
+% \]
+% Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
+% \[
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
+% \]
+% where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
+% \begin{equation}\label{eq:cvk}
+% v_k(\lambda,\Phi_1) = \begin{cases}
+% c_k(\Phi_1) - \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+% \lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
+% 0 & \text{otherwise}.
+% \end{cases}
+% \end{equation}
+% \end{IEEEproof}
 
@@
 \[
-X_L = \sqrt{L} \big( G_L(0) - T(0) \big) = \sqrt{L} \big( G_L(0) - \expect G_L(0) \big)
+X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big) = \sqrt{L} \big( G_L(0) - \expect G_L(0) \big)
 \] 
@@
 
+\small
 \bibliography{bib}
 
+\normalsize
+\appendix
+
+%\section{Tricks with fractional parts}
+%Throughout the paper, and particularly in  have made use of a number of results involving the fractional part function
+
+\subsection{Empirical process results}
+
+During the proof of asymptotic normality in Section~\ref{sec:proof-asympt-norm} we made use of the following result regarding the functions $W_L(\lambda)$ and $Y_L(\lambda)$ defined in~\eqref{eq:WLdef}~and~\eqref{eq:YLdef}.
+
+\begin{lemma}\label{lem:ZYLempiricproc}
+For any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that:
+\begin{enumerate}
+\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu$, and,
+\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu$,
+\end{enumerate}
+for all positive integers $L$
+\end{lemma}
+\begin{IEEEproof}
+These results are related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
+The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
+We will only give a proof of the first statement regarding the function $W_L$.
+The proof for the second statement regarding $Y_L$ is similar.
+
+Put 
+\[
+f_i(\lambda, R_i, \Phi_i) = \begin{cases}
+R_i( \sin(\lambda + \Phi_i) - \sin(\Phi_i)), & i \in P \\
+R_i( \sin\sfracpart{\lambda + \Phi_i} - \sin\sfracpart{\Phi_i}), & i \in D \\
+\end{cases}
+\]
+so that
+\[
+W_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i = 1}^{L} \big( f_i(\lambda, R_i, \Phi_i) - \expect f_i(\lambda, R_i, \Phi_i) \big).
+\]
+Let $\{g_i\}$ be a sequence of independent standard normal random variables, independent of $\{R_i\}$ and $\{\Phi_i\}$.  Lemma~\ref{lem:symmetrisation} shows that
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) },
+\]
+where 
+\begin{equation}\label{eq:ZpsiCondGaussProc}
+X_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{n=1}^{L} g_i f_i(\lambda, R_i, \Phi_i),
+\end{equation}
+and where $\expect$ runs over all of $\{g_i\}$, $\{R_i\}$ and $\{\Phi_i\}$.
+Conditionally on $\{\Phi_i\}$ and $\{R_i\}$, $\{X_L(\lambda), \lambda \in \reals \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
+Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists $\epsilon > 0$ such that
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } < \kappa.
+\]
+It follows that,
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)}  <  \sqrt{2\pi} \; \kappa,
+\]
+and by Markov's inequality,
+\[
+\prob \cubr{  \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
+\]
+for any $\delta > 0$.  The proof follows by choosing $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:symmetrisation}
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) }
+\]
+\end{lemma}
+\begin{IEEEproof}
+Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$, both $\{R_i'\}$ and $\{\Phi_i'\}$ chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
+Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
+Then
+\begin{align*}
+W_L(\lambda) &= \frac{1}{\sqrt{L}} \sum_{i = 1}^{L} \big( f_i(\lambda, R_i, \Phi_i) - \expect_{R,\Phi} f_i(\lambda, R_i', \Phi_i') \big) \\
+&= \expect_{R,\Phi} \frac{1}{\sqrt{L}} \sum_{i=1}^{L} ( f_{i} - f_{i}^\prime ),
+\end{align*}
+where, for notational convenience, we put 
+\[
+f_{i} = f_i(\lambda, R_i, \Phi_i) \qquad \text{and} \qquad  f_{i}^\prime = f_i(\lambda, R_i', \Phi_i').
+\] 
+Taking absolute values followed by supremums,
+\begin{align*}
+ \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &= \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \expect_{R,\Phi} \sum_{i=1}^{L} (f_{i} - f_{i}^\prime )} \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi}  \abs{ \frac{1}{\sqrt{L}} \sum_{i=1}^{L}  (f_{i} - f_{i}^\prime) },
+\end{align*}
+the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, the bound can be increased by moving $\expect_{R,\Phi}$ outside the supremum,
+\begin{equation}\label{eq:Eoutsidesup}
+ \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} ( f_{i} - f_{i}^\prime ) }.
+\end{equation}
+Applying $\expect$ to both sides gives the inequality
+\begin{equation}\label{eq:Gninequfin}
+ \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L}  (f_{i} - f_{i}^\prime) },
+\end{equation}
+since $\expect \expect_{R,\Phi}$ is equivalent to $\expect$.   Let
+\[
+\sigma_i = \frac{g_i}{\abs{g_i}},
+\]
+and put $\sigma_i = 1$ if $g_i = 0$.  The $\sigma_i$ are thus independent with 
+\[
+\prob\{ \sigma_i = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_i = 1 \} = \tfrac{1}{2}.
+\]  
+As $R_i$ and $\Phi_i$ are independent and identically distributed to $R_i^\prime$ and $\Phi_i^\prime$, the random variable $f_{i} - f_{i}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_i( f_{i} - f_{i}^\prime)$.
+Thus,
+%\[
+%\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
+%\]
+%are identically distributed, and therefore
+\begin{align*}
+ \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{  \frac{1}{\sqrt{L}}\sum_{i=1}^{L}  (f_{i} - f_{i}^\prime) } \\
+&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{  \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i ( f_{i} - f_{i}^\prime) }.
+\end{align*}
+By the triangle inequality
+\[
+\abs{ \sum_{i=1}^{L} \sigma_i (f_{i} - f_{i}^\prime)} \leq \abs{ \sum_{i=1}^{L} \sigma_i f_{i} } + \abs{ \sum_{i=1}^{L}  \sigma_i f_{i}^\prime },
+\]
+and it follows from~\eqref{eq:Gninequfin}, that
+\begin{align*}
+ \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &\leq \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} } \\
+&\hspace{1cm} + \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \sum_{i=1}^{L}  \sigma_i  f_{i}^\prime } \\
+&= 2 \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} }.
+\end{align*}
+Since $g_i$ is a standard normal random variable, $\expect \abs{g_i} = \expect_{R,\Phi} \abs{g_i} = \sqrt{2/\pi}$, and
+\begin{align*}
+\expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} } \\
+&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} \sqrt{\frac{\pi}{2}}\expect_{R,\Phi} \abs{g_i}} \\
+&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{ \sqrt{L}}\sum_{i=1}^{L} \sigma_i \abs{g_i}  f_{i}},
+\end{align*}
+the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$, it follows that
+\begin{align*}
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} g_i  f_{i}} \\
+&= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) } 
+\end{align*}
+as required.
+\end{IEEEproof}
+
+\begin{lemma} \label{lem:chaining}
+For any $\kappa > 0$ there exists $\epsilon > 0$ such that
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } < \kappa,
+\]
+where $X_L(\lambda)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
+\end{lemma}
+\begin{IEEEproof}
+
+\end{IEEEproof}
+
 \end{document}

### 2012-08-13 paper.tex
Added to paper.  Started added the empirical proc stuff to the appendix.  Needs more work.  I'm not sure I've got the notation right.

@@
 \[
-D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
+D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
 \]
@@
 \[
-\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i=1}^{L} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
+\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
 \]
@@
 \begin{align*}
-C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i=1}^L Z_i(\lambda) - Z_i(\lambda_n)} \\
-&\leq \frac{1}{L}\sum_{i=1}^L \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
-&\leq \frac{2\pi}{N L}\sum_{i=1}^L R_i .
+C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i \in P \cup D} Z_i(\lambda) - Z_i(\lambda_n)} \\
+&\leq \frac{1}{L}\sum_{i \in P \cup D} \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
+&\leq \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i .
 \end{align*}
@@
 \[
-\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i=1}^L R_i = \tfrac{2\pi}{N}\expect R_1
+\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i = \tfrac{2\pi}{N}\expect R_1
 \]
@@
 \[
-\frac{2\pi}{N L}\sum_{i=1}^L R_i \rightarrow \frac{2\pi}{N}\expect R_1
+\frac{2\pi}{N L}\sum_{i \in P \cup D} R_i \rightarrow \frac{2\pi}{N}\expect R_1
 \] 
@@
 \prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i=1}^L R_i > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
 \end{align*}
@@
 \end{enumerate}
-for all positive integers $L$
+for all positive integers $L$.
 \end{lemma}
@@
 \[
-W_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i = 1}^{L} \big( f_i(\lambda, R_i, \Phi_i) - \expect f_i(\lambda, R_i, \Phi_i) \big).
+W_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect f_i(\lambda, R_i, \Phi_i) \big).
 \]
@@
 \begin{equation}\label{eq:ZpsiCondGaussProc}
-X_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{n=1}^{L} g_i f_i(\lambda, R_i, \Phi_i),
+X_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} g_i f_i(\lambda, R_i, \Phi_i),
 \end{equation}
@@
 \begin{align*}
-W_L(\lambda) &= \frac{1}{\sqrt{L}} \sum_{i = 1}^{L} \big( f_i(\lambda, R_i, \Phi_i) - \expect_{R,\Phi} f_i(\lambda, R_i', \Phi_i') \big) \\
-&= \expect_{R,\Phi} \frac{1}{\sqrt{L}} \sum_{i=1}^{L} ( f_{i} - f_{i}^\prime ),
+W_L(\lambda) &= \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect_{R,\Phi} f_i(\lambda, R_i', \Phi_i') \big) \\
+&= \expect_{R,\Phi} \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ),
 \end{align*}
@@
 \begin{align*}
- \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &= \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \expect_{R,\Phi} \sum_{i=1}^{L} (f_{i} - f_{i}^\prime )} \\
-&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi}  \abs{ \frac{1}{\sqrt{L}} \sum_{i=1}^{L}  (f_{i} - f_{i}^\prime) },
+ \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &= \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \expect_{R,\Phi} \sum_{i \in P \cup D} (f_{i} - f_{i}^\prime )} \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi}  \abs{ \frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) },
 \end{align*}
@@
 \begin{equation}\label{eq:Eoutsidesup}
- \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} ( f_{i} - f_{i}^\prime ) }.
+ \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ) }.
 \end{equation}
@@
 \begin{equation}\label{eq:Gninequfin}
- \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L}  (f_{i} - f_{i}^\prime) },
+ \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} (f_{i} - f_{i}^\prime) },
 \end{equation}
@@
 \begin{align*}
- \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{  \frac{1}{\sqrt{L}}\sum_{i=1}^{L}  (f_{i} - f_{i}^\prime) } \\
-&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{  \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i ( f_{i} - f_{i}^\prime) }.
+ \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) } \\
+&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i ( f_{i} - f_{i}^\prime) }.
 \end{align*}
@@
 \[
-\abs{ \sum_{i=1}^{L} \sigma_i (f_{i} - f_{i}^\prime)} \leq \abs{ \sum_{i=1}^{L} \sigma_i f_{i} } + \abs{ \sum_{i=1}^{L}  \sigma_i f_{i}^\prime },
+\abs{ \sum_{i\in P \cup D} \sigma_i (f_{i} - f_{i}^\prime)} \leq \abs{ \sum_{i \in P \cup D} \sigma_i f_{i} } + \abs{ \sum_{i\in P \cup D}  \sigma_i f_{i}^\prime },
 \]
@@
 \begin{align*}
- \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &\leq \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} } \\
-&\hspace{1cm} + \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \sum_{i=1}^{L}  \sigma_i  f_{i}^\prime } \\
-&= 2 \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} }.
+ \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &\leq \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} \sigma_i f_{i} } \\
+&\hspace{1cm} + \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  \sigma_i  f_{i}^\prime } \\
+&= 2 \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} }.
 \end{align*}
@@
 \begin{align*}
-\expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} } \\
-&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} \sigma_i f_{i} \sqrt{\frac{\pi}{2}}\expect_{R,\Phi} \abs{g_i}} \\
-&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{ \sqrt{L}}\sum_{i=1}^{L} \sigma_i \abs{g_i}  f_{i}},
+\expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} } \\
+&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} \sqrt{\frac{\pi}{2}}\expect_{R,\Phi} \abs{g_i}} \\
+&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{ \sqrt{L}}\sum_{i \in P \cup D} \sigma_i \abs{g_i}  f_{i}},
 \end{align*}
@@
 \begin{align*}
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i=1}^{L} g_i  f_{i}} \\
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} g_i  f_{i}} \\
 &= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) } 
@@
 \begin{IEEEproof}
+Without loss of generality, assume that $\epsilon < \tfrac{\pi}{M}$.  Lemma~\ref{lem:chaining2} shows that
+\begin{equation}\label{eq:supZCK1} 
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi\})}
+\end{equation}
+where $K_1$ is a finite, positive constant, and $C_\epsilon(\{\Phi\})$ is average number of times $\Phi_i, i \in D$ are inside the set
+\[
+S_\epsilon = \bigcup_{k \in \ints} [\psi_k - \epsilon, \psi_k + \epsilon],
+\]
+where $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  That is,
+\[
+C_\epsilon(\{\Phi\}) = \frac{1}{L} \sum_{i \in P \cup D} I_\epsilon(\Phi_i)
+\]
+where $I_\epsilon(\Phi_i) = 1$ if $\Phi_i \in S_\epsilon$ and zero otherwise.
+\end{IEEEproof}
+
+The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
+The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
+%Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.
+For notational convenience we use the abbreviations 
+\[
+C_\epsilon = C_\epsilon(\{\Phi\}) \qquad \text{and} \qquad f_{i}(\lambda) = f_{i}(\lambda, R_i, \Phi_i)
+\]
+in what follows.
+As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
+%This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.
+
+\begin{lemma}\label{lem:chaining2}
+There exists a positive constant $K_1$ such that
+\[
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } \leq K_1 \sqrt{C_\epsilon}.
+\]
+\end{lemma}
+\begin{IEEEproof}
+Let $B_\epsilon = \{x \in \reals \mid \abs{x} < \epsilon\}$ be the set of real numbers with magnitude less than $\epsilon$.
+For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $\reals$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda^* \in T_\epsilon(k)$ such that the pseudometric
+\[
+d(\lambda, \lambda^*) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \leq 2^{-k} C_\epsilon N.
+\]
+We specifically define $T_\epsilon(0)$ to contain only zero.  Defined this way 
+\end{IEEEproof}
+
+\begin{lemma}
+Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
+\begin{align*}
+d(\lambda, \lambda^*) &= \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \\
+&\leq R_i^2 L \left( (\lambda - \lambda^*)^2 + 4 C_\epsilon \right).
+\end{align*}
+\end{lemma}
+
+\begin{lemma}
+If $i \in P$, then $\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2.$
+\end{lemma}
+\begin{IEEEproof} 
+We have
+\[
+\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 = R_i^2\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)}^2
+\]
+and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+\[
+\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)} \leq K\sabs{\lambda - \lambda^*} = \sabs{\lambda - \lambda^*}.
+\]
+\end{IEEEproof}
+
+\begin{lemma}
+Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$.  If $i \in D$, then
+\begin{equation}\label{eq:fismall}
+\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2
+\end{equation}
+when $\Phi_i \notin S_\epsilon$, otherwise,
+\begin{equation}\label{eq:fibig}
+\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq 4 R_i^2.
+\end{equation}
+\end{lemma}
+\begin{IEEEproof}
+Because $\sin(\cdot) \leq 1$ it follows that
+\begin{align*}
+\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 &= R_i^2\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}}^2 \\
+&\leq 4 R_i^2
+\end{align*}
+for any $\Phi_i \in \reals$, and so~\eqref{eq:fibig} holds.  If $\Phi_i \notin S_\epsilon$ then $\Phi_i \in (\psi_{k-1} + \epsilon, \psi_{k} - \epsilon)$ for some $k\in\ints$ and 
+\[
+\fracpart{\Phi_i + \lambda} = \Phi_i + \lambda - \tfrac{2\pi}{M}k
+\] 
+and
+\[
+\fracpart{\Phi_i + \lambda^*} = \Phi_i + \lambda^* - \tfrac{2\pi}{M}k 
+\]
+and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+\begin{align*}
+\vert \sin&\fracpart{\Phi_i + \lambda} - \sin\fracpart{\Phi_i + \lambda^*}\vert \\
+&=  \sabs{\sin(\Phi_i + \lambda - \tfrac{2\pi}{M}k) - \sin(\Phi_i + \lambda^* - \tfrac{2\pi}{M}k)} \\
+&\leq K \sabs{\lambda - \lambda^*} \\
+&= \sabs{\lambda - \lambda^*}.
+\end{align*}
+Thus~\eqref{eq:fismall} holds. 
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:chaining3}(Chaining)
+For all positive integers $k$,
+\[
+\expect_{R,\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ X_L(\lambda) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{?}{2^{i/2}}.
+\]
+\end{lemma}
+\begin{IEEEproof}
 

### 2012-08-14 paper.tex
added to paper.

@@
 SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \\
-&= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in P \cup D} \abs{ y_i - a d_i }^2,
+&= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2,
 \end{align*}
@@
 where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
+
+\begin{lemma}\label{lem:gg2cont}
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with pdf $f(r,\phi)$.
+Let
+\[
+g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
+\]
+and
+\[
+g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr.
+\]
+Then $g$ and $g_2$ are continuous functions from $\reals$ to $\reals$.
+\end{lemma}
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
 
@@
 
-\subsection{Empirical process results}
+\subsection{An empirical process result}
 
@@
 \begin{IEEEproof}
-Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$, both $\{R_i'\}$ and $\{\Phi_i'\}$ chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
+Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$.
+Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
 Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
@@
 \end{align*}
-Since $g_i$ is a standard normal random variable, $\expect \abs{g_i} = \expect_{R,\Phi} \abs{g_i} = \sqrt{2/\pi}$, and
+Because $g_i$ is a standard normal random variable it follows that $\expect \abs{g_i} = \expect_{R,\Phi} \abs{g_i} = \sqrt{2/\pi}$, and
 \begin{align*}
@@
 \]
-where $X_L(\lambda)$ is defined by~\eqref{eq:ZpsiCondGaussProc}.
+where $X_L(\lambda)$ is defined in~\eqref{eq:ZpsiCondGaussProc}.
 \end{lemma}
@@
 \begin{equation}\label{eq:supZCK1} 
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{C_\epsilon(\{\Phi\})}
-\end{equation}
-where $K_1$ is a finite, positive constant, and $C_\epsilon(\{\Phi\})$ is average number of times $\Phi_i, i \in D$ are inside the set
-\[
-S_\epsilon = \bigcup_{k \in \ints} [\psi_k - \epsilon, \psi_k + \epsilon],
-\]
-where $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  That is,
-\[
-C_\epsilon(\{\Phi\}) = \frac{1}{L} \sum_{i \in P \cup D} I_\epsilon(\Phi_i)
-\]
-where $I_\epsilon(\Phi_i) = 1$ if $\Phi_i \in S_\epsilon$ and zero otherwise.
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{\frac{C_\epsilon}{L}}
+\end{equation}
+where $K_1$ is a finite, positive constant, and
+\[
+C_\epsilon = 2\epsilon\sum_{i\in P \cup D}R_i^2 + 4\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
+\]
+where $I_{\epsilon}(\Phi_i)$ returns one when $\Phi_i \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon]$ and zero otherwise.  Now,
+\[
+\tfrac{1}{L}\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
+\]
+Recalling that $f(r,\phi)$ is the joint probability density function of $R_i$ and $\Phi_i$,
+\begin{align*}
+\expect R_i^2 I_\epsilon(\Phi_i) &= \int_0^{2\pi} \int_{0}^\infty  r^2 f(r,\phi) I_\epsilon(\phi) dr d\phi \\ 
+&= \int_0^{2\pi}g_2(\phi) I_\epsilon(\phi) d\phi 
+\end{align*}
+where the function 
+\[
+g_2(\phi) = \int_{0}^\infty  r^2 f(r,\phi) dr d\phi 
+\]
+is continuous due to Lemma~\ref{lem:gg2cont}.  From the definition of $I_\epsilon$,
+\begin{align*}
+\expect R_i^2 I_\epsilon(\Phi_i) &= \sum_{k=0}^{M-1}\int_{\psi_k-\epsilon}^{\psi_k-\epsilon}g_2(\phi) d\phi \\
+&= 2\epsilon \sum_{k=0}^{M-1} (g_2(\psi_k) + o(1))
+\end{align*}
+where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.
+Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$.
+Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive, then
+\[
+\expect \sqrt{\frac{C_\epsilon}{L}} \leq  \sqrt{\tfrac{1}{L}\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
+\]
+by Jensen's inequality.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
+\]
+Choosing 
+\[
+\epsilon = \frac{\kappa^2}{ K_1^2 ( K_2 + 2\expect R_1^2)}
+\] 
+completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
 \end{IEEEproof}
@@
 %Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.
-For notational convenience we use the abbreviations 
-\[
-C_\epsilon = C_\epsilon(\{\Phi\}) \qquad \text{and} \qquad f_{i}(\lambda) = f_{i}(\lambda, R_i, \Phi_i)
+For notational convenience we use the abbreviation 
+\[
+f_{i}(\lambda) = f_{i}(\lambda, R_i, \Phi_i)
 \]
@@
 \[
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } \leq K_1 \sqrt{C_\epsilon}.
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } \leq K_1 \sqrt{\frac{C_\epsilon}{L}}.
 \]
@@
 Let $B_\epsilon = \{x \in \reals \mid \abs{x} < \epsilon\}$ be the set of real numbers with magnitude less than $\epsilon$.
-For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $\reals$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda^* \in T_\epsilon(k)$ such that the pseudometric
-\[
-d(\lambda, \lambda^*) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \leq 2^{-k} C_\epsilon N.
-\]
-We specifically define $T_\epsilon(0)$ to contain only zero.  Defined this way 
-\end{IEEEproof}
-
-\begin{lemma}
+For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $B_\epsilon$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the psuedometric
+\[
+d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon,
+\]
+and the psuedometric
+\[
+d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon
+\]
+where 
+\[
+D_\epsilon = \epsilon\sum_{i\in P \cup D}R_i + \sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
+\]
+We specifically define $T_\epsilon(0)$ to contain only zero.
+Defined this way, $T_\epsilon(0)$ satisfies the inequality above because both $d_1(\lambda, 0)$ and $d_2(\lambda, 0)$ are less than $C_\epsilon$ for all $\lambda \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
+The existance of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{}.
+
+Lemma~\ref{lem:limexpTk} shows that
+\[
+\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}.
+\]
+Lemma~\ref{lem:chaining3} shows that
+\[
+\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} \leq \sqrt{\frac{C_\epsilon}{L}} \sum_{i=1}^{k} \frac{?}{2^{i/2}}
+\]
+The sum above converges as $k\rightarrow\infty$ so lemma holds with
+\[
+K_1 = \sum_{i=1}^{\infty} \frac{?}{2^{i/2}}.
+\]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:limexpTk}
+\[
+\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}
+\]
+\end{lemma}
+\begin{IEEEproof}
+Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
+\[
+\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}.
+\]
+Let $\lambda^* \in B_\epsilon$ be a maximiser of $\sabs{X_L(\lambda)}$, that is
+\[
+\sabs{X_L(\lambda^*)} = \sup_{\sabs{\lambda} < \epsilon}\sabs{X_L(\lambda)}.
+\]
+By definition $T_\epsilon(k)$ contains a point $\lambda_2$ such that $d_2(\lambda^*,\lambda_2) \leq 2^{-k}D_\epsilon$.  Now
+\begin{align*}
+\sabs{X_L(\lambda^*) - X_L(\lambda_2)} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda^*) - f_i(\lambda_2))} \\
+&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} \sum_{i\in P \cup D}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} d(\lambda^*,\lambda_2) \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon.
+\end{align*}
+where $g_\text{max} = \max_{i\in P\cup D} \abs{g_i}$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:epslmlemma}
 Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
@@
 \begin{IEEEproof}
-
+Let $b_k(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that 
+\[
+d(\lambda, b_k(\lambda)) \leq 2^{1-k}C_\epsilon.
+\]
+The existance of $b_k$ is guaranteed by the definition of $T_\epsilon(k-1)$. 
 \end{IEEEproof}

### 2012-08-15 paper.tex
Added to paper.  Empirical process proof is comming along.

@@
 \begin{IEEEproof}
+Let $b_k(\lambda)$ be a function that maps $\lambda \in B_\epsilon$ to a point in $T_\epsilon(k)$ such that  $d_2(\lambda,b_k(\lambda)) \leq 2^{-k}D_\epsilon$.
+The existance of $b_k$ is guaranteed by the definition of $T_\epsilon(k-1)$.
+So, for all $\lambda \in B_\epsilon$,
+\begin{align*}
+\sabs{X_L(\lambda) - X_L(b_k(\lambda))} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda) - f_i(b_k(\lambda)))} \\
+&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda) - f_i(b_k(\lambda))} \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} \sum_{i\in P \cup D}\sabs{f_i(\lambda) - f_i(b_k(\lambda))} \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} d_2(\lambda,b_k(\lambda)) \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon.
+\end{align*}
+where $g_\text{max} = \max_{i\in P\cup D} \abs{g_i}$. The bound applies uniformly over $\lambda \in B_\epsilon$, so
+\[
+\sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda) - X_L(b_k(\lambda))} \leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon.
+\]
+Applying $\expect_{R,\Phi}$ to both sides of this inequality,
+\[
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda) - X_L(b_k(\lambda))} \leq K_4 2^{-k} D_\epsilon
+\]
+where $K_4 = \tfrac{1}{\sqrt{L}}\expect_{R,\Phi} g_\text{max}$ is a positive constant as a result of Lemma~\ref{} (BLERG, a maximal inequality lemma is needed).
 Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
 \[
-\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}.
-\]
-Let $\lambda^* \in B_\epsilon$ be a maximiser of $\sabs{X_L(\lambda)}$, that is
-\[
-\sabs{X_L(\lambda^*)} = \sup_{\sabs{\lambda} < \epsilon}\sabs{X_L(\lambda)}.
-\]
-By definition $T_\epsilon(k)$ contains a point $\lambda_2$ such that $d_2(\lambda^*,\lambda_2) \leq 2^{-k}D_\epsilon$.  Now
-\begin{align*}
-\sabs{X_L(\lambda^*) - X_L(\lambda_2)} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda^*) - f_i(\lambda_2))} \\
-&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} \sum_{i\in P \cup D}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} d(\lambda^*,\lambda_2) \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon.
-\end{align*}
-where $g_\text{max} = \max_{i\in P\cup D} \abs{g_i}$.
+\sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X_L(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X_L(\lambda)},
+\]
+and so,
+\begin{align*}
+0 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X_L(\lambda)} - \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X_L(\lambda)} \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \big( \sabs{X_L(\lambda)} - \sabs{X_L(b_k(\lambda))} \big) \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X_L(\lambda) - X_L(b_k(\lambda))} \\
+&\leq K_4 2^{-k} D_\epsilon.
+\end{align*}
+The lemma follows since this converges to zero as $k\rightarrow\infty$.
 \end{IEEEproof}
@@
 \begin{IEEEproof}
-Let $b_k(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that 
-\[
-d(\lambda, b_k(\lambda)) \leq 2^{1-k}C_\epsilon.
-\]
-The existance of $b_k$ is guaranteed by the definition of $T_\epsilon(k-1)$. 
+Let $b_k(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_k(\lambda)) \leq 2^{1-k}C_\epsilon$. 
 \end{IEEEproof}

### 2012-08-15 paper.tex
Added to paper.  Need the covering lemma now.

--- 2012-08-14 paper.tex
@@
 \[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) },
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) },
 \]
@@
 \begin{equation}\label{eq:ZpsiCondGaussProc}
-X_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} g_i f_i(\lambda, R_i, \Phi_i),
+X(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} g_i f_i(\lambda, R_i, \Phi_i),
 \end{equation}
 and where $\expect$ runs over all of $\{g_i\}$, $\{R_i\}$ and $\{\Phi_i\}$.
-Conditionally on $\{\Phi_i\}$ and $\{R_i\}$, $\{X_L(\lambda), \lambda \in \reals \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
+Conditionally on $\{\Phi_i\}$ and $\{R_i\}$, $\{X(\lambda), \lambda \in \reals \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
 Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists $\epsilon > 0$ such that
 \[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } < \kappa.
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa.
 \]
@@
 \[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) }
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) }
 \]
@@
 \end{align*}
-By the triangle inequality
+By the triangle inequality,
 \[
@@
 \expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} g_i  f_{i}} \\
-&= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) } 
+&= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) } 
 \end{align*}
@@
 \[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } < \kappa,
-\]
-where $X_L(\lambda)$ is defined in~\eqref{eq:ZpsiCondGaussProc}.
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa,
+\]
+where $X(\lambda)$ is the Gaussian process defined in~\eqref{eq:ZpsiCondGaussProc}.
 \end{lemma}
@@
 \begin{equation}\label{eq:supZCK1} 
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{\frac{C_\epsilon}{L}}
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}
 \end{equation}
 where $K_1$ is a finite, positive constant, and
-\[
-C_\epsilon = 2\epsilon\sum_{i\in P \cup D}R_i^2 + 4\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
-\]
-where $I_{\epsilon}(\Phi_i)$ returns one when $\Phi_i \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon]$ and zero otherwise.  Now,
-\[
-\tfrac{1}{L}\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
+\begin{equation}\label{eq:Cepsdefn}
+C_\epsilon = \frac{2}{L}\epsilon\sum_{i\in P \cup D}R_i^2 + \frac{4}{L}\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
+\end{equation}
+where $I_\epsilon$ is the indicator function
+\[
+I_{\epsilon}(x) = \begin{cases}
+1, & x \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon] \\
+0, & \text{otherwise},
+\end{cases}
+\] 
+and $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  Now,
+\[
+\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
 \]
@@
 Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive, then
-\[
-\expect \sqrt{\frac{C_\epsilon}{L}} \leq  \sqrt{\tfrac{1}{L}\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
+Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive, it follows that
+\[
+\expect \sqrt{C_\epsilon} \leq  \sqrt{\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
 \]
@@
 \[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
 \]
@@
 in what follows.
-As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
+As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{\pi}{M}$.
 %This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.
@@
 \[
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } \leq K_1 \sqrt{\frac{C_\epsilon}{L}}.
-\]
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}.
+\]
+where $C_\epsilon$ is defined in~\eqref{eq:Cepsdefn}
 \end{lemma}
@@
 \[
-d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon,
+d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
 \]
@@
 \[
-d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon
+d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L
 \]
 where 
-\[
-D_\epsilon = \epsilon\sum_{i\in P \cup D}R_i + \sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
-\]
+\begin{equation}\label{eq:Depsdefn}
+D_\epsilon = \frac{1}{L}\epsilon\sum_{i\in P \cup D}R_i + \frac{1}{L}\sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
+\end{equation}
 We specifically define $T_\epsilon(0)$ to contain only zero.
-Defined this way, $T_\epsilon(0)$ satisfies the inequality above because both $d_1(\lambda, 0)$ and $d_2(\lambda, 0)$ are less than $C_\epsilon$ for all $\lambda \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
-The existance of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{}.
+Defined this way, $T_\epsilon(0)$ satisfies the inequality above because $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
+The existance of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{lem:metricentropy}.
+Lemma~\ref{lem:metricentropy} also shows that the number of elements in $T_\epsilon(k)$ is no more that $K_3 2^k$ where $K_3$ is a constant, independent of $L$, $\epsilon$ and $k$.
 
@@
 \[
-\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}.
+\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}.
 \]
@@
 \[
-\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} \leq \sqrt{\frac{C_\epsilon}{L}} \sum_{i=1}^{k} \frac{?}{2^{i/2}}
+\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}
 \]
@@
 \[
-K_1 = \sum_{i=1}^{\infty} \frac{?}{2^{i/2}}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:limexpTk}
-\[
-\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}
-\]
-\end{lemma}
-\begin{IEEEproof}
-Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
-\[
-\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}.
-\]
-Let $\lambda^* \in B_\epsilon$ be a maximiser of $\sabs{X_L(\lambda)}$, that is
-\[
-\sabs{X_L(\lambda^*)} = \sup_{\sabs{\lambda} < \epsilon}\sabs{X_L(\lambda)}.
-\]
-By definition $T_\epsilon(k)$ contains a point $\lambda_2$ such that $d_2(\lambda^*,\lambda_2) \leq 2^{-k}D_\epsilon$.  Now
-\begin{align*}
-\sabs{X_L(\lambda^*) - X_L(\lambda_2)} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda^*) - f_i(\lambda_2))} \\
-&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} \sum_{i\in P \cup D}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} d(\lambda^*,\lambda_2) \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon.
-\end{align*}
-where $g_\text{max} = \max_{i\in P\cup D} \abs{g_i}$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:epslmlemma}
-Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
-\begin{align*}
-d(\lambda, \lambda^*) &= \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \\
-&\leq R_i^2 L \left( (\lambda - \lambda^*)^2 + 4 C_\epsilon \right).
-\end{align*}
-\end{lemma}
+K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}
+\]
+\end{IEEEproof}
 
 \begin{lemma}
-If $i \in P$, then $\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2.$
-\end{lemma}
-\begin{IEEEproof} 
-We have
-\[
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 = R_i^2\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)}^2
-\]
-and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
-\[
-\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)} \leq K\sabs{\lambda - \lambda^*} = \sabs{\lambda - \lambda^*}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}
-Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$.  If $i \in D$, then
-\begin{equation}\label{eq:fismall}
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2
-\end{equation}
-when $\Phi_i \notin S_\epsilon$, otherwise,
-\begin{equation}\label{eq:fibig}
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq 4 R_i^2.
-\end{equation}
-\end{lemma}
-\begin{IEEEproof}
-Because $\sin(\cdot) \leq 1$ it follows that
-\begin{align*}
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 &= R_i^2\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}}^2 \\
-&\leq 4 R_i^2
-\end{align*}
-for any $\Phi_i \in \reals$, and so~\eqref{eq:fibig} holds.  If $\Phi_i \notin S_\epsilon$ then $\Phi_i \in (\psi_{k-1} + \epsilon, \psi_{k} - \epsilon)$ for some $k\in\ints$ and 
-\[
-\fracpart{\Phi_i + \lambda} = \Phi_i + \lambda - \tfrac{2\pi}{M}k
-\] 
-and
-\[
-\fracpart{\Phi_i + \lambda^*} = \Phi_i + \lambda^* - \tfrac{2\pi}{M}k 
-\]
-and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
-\begin{align*}
-\vert \sin&\fracpart{\Phi_i + \lambda} - \sin\fracpart{\Phi_i + \lambda^*}\vert \\
-&=  \sabs{\sin(\Phi_i + \lambda - \tfrac{2\pi}{M}k) - \sin(\Phi_i + \lambda^* - \tfrac{2\pi}{M}k)} \\
-&\leq K \sabs{\lambda - \lambda^*} \\
-&= \sabs{\lambda - \lambda^*}.
-\end{align*}
-Thus~\eqref{eq:fismall} holds. 
-\end{IEEEproof}
+Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.
+\end{lemma}
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:limexpTk} With $T_\epsilon(k)$ defined as in Lemma~\ref{lem:chaining2},
+\[
+\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}
+\]
+\end{lemma}
+\begin{IEEEproof}
+Let $b_2(\lambda)$ be a function that maps $\lambda \in B_\epsilon$ to a point in $T_\epsilon(k)$ such that  $d_2(\lambda,b_2(\lambda)) \leq 2^{-k}D_\epsilon L$.
+The existance of $b_2$ is guaranteed by the definition of $T_\epsilon(k-1)$.
+So, for all $\lambda \in B_\epsilon$,
+\begin{align*}
+\sabs{X(\lambda) - X(b_2(\lambda))} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda) - f_i(b_2(\lambda)))} \\
+&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} \sum_{i\in P \cup D}\sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} d_2(\lambda,b_2(\lambda)) \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L,
+\end{align*}
+where $g_\text{max} = \max_{i\in P\cup D} \abs{g_i}$. The bound applies uniformly over $\lambda \in B_\epsilon$, so
+\[
+\sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L.
+\]
+Applying $\expect_{R,\Phi}$ to both sides of this inequality,
+\[
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq K_4 2^{-k} D_\epsilon L,
+\]
+where, as a result of Lemma~\ref{lem:maxineq},
+\[
+K_4 = \tfrac{1}{\sqrt{L}}\expect_{R,\Phi} g_\text{max} \leq 3 \sqrt{\frac{\log L}{L}}
+\]
+is a constant, independent of $\epsilon$ and $k$.  Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
+\[
+\sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)},
+\]
+and so,
+\begin{align}
+0 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \nonumber \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \nonumber \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \big( \sabs{X(\lambda)} - \sabs{X(b_2(\lambda))} \big) \nonumber \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda) - X(b_2(\lambda))} \nonumber \\
+&\leq K_4 2^{-k} D_\epsilon L, \label{eq:Deps2nktozero}
+\end{align}
+the second line because,
+\[
+\sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)},
+\]
+since $b_2(\lambda) \in T_\epsilon(k)$.  The lemma follows from~\eqref{eq:Deps2nktozero} because $K_4 2^{-k} D_\epsilon L$ converges to zero as $k\rightarrow\infty$.
+\end{IEEEproof}
+
+% \begin{lemma}\label{lem:epslmlemma}
+% Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
+% \begin{align*}
+% d(\lambda, \lambda^*) &= \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \\
+% &\leq R_i^2 L \left( (\lambda - \lambda^*)^2 + 4 C_\epsilon \right).
+% \end{align*}
+% \end{lemma}
+
+% \begin{lemma}
+% If $i \in P$, then $\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2.$
+% \end{lemma}
+% \begin{IEEEproof} 
+% We have
+% \[
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 = R_i^2\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)}^2
+% \]
+% and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+% \[
+% \sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)} \leq K\sabs{\lambda - \lambda^*} = \sabs{\lambda - \lambda^*}.
+% \]
+% \end{IEEEproof}
+
+% \begin{lemma}
+% Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$.  If $i \in D$, then
+% \begin{equation}\label{eq:fismall}
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2
+% \end{equation}
+% when $\Phi_i \notin S_\epsilon$, otherwise,
+% \begin{equation}\label{eq:fibig}
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq 4 R_i^2.
+% \end{equation}
+% \end{lemma}
+% \begin{IEEEproof}
+% Because $\sin(\cdot) \leq 1$ it follows that
+% \begin{align*}
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 &= R_i^2\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}}^2 \\
+% &\leq 4 R_i^2
+% \end{align*}
+% for any $\Phi_i \in \reals$, and so~\eqref{eq:fibig} holds.  If $\Phi_i \notin S_\epsilon$ then $\Phi_i \in (\psi_{k-1} + \epsilon, \psi_{k} - \epsilon)$ for some $k\in\ints$ and 
+% \[
+% \fracpart{\Phi_i + \lambda} = \Phi_i + \lambda - \tfrac{2\pi}{M}k
+% \] 
+% and
+% \[
+% \fracpart{\Phi_i + \lambda^*} = \Phi_i + \lambda^* - \tfrac{2\pi}{M}k 
+% \]
+% and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+% \begin{align*}
+% \vert \sin&\fracpart{\Phi_i + \lambda} - \sin\fracpart{\Phi_i + \lambda^*}\vert \\
+% &=  \sabs{\sin(\Phi_i + \lambda - \tfrac{2\pi}{M}k) - \sin(\Phi_i + \lambda^* - \tfrac{2\pi}{M}k)} \\
+% &\leq K \sabs{\lambda - \lambda^*} \\
+% &= \sabs{\lambda - \lambda^*}.
+% \end{align*}
+% Thus~\eqref{eq:fismall} holds. 
+% \end{IEEEproof}
 
@@
 \[
-\expect_{R,\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ X_L(\lambda) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{?}{2^{i/2}}.
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $b_k(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that 
-\[
-d(\lambda, b_k(\lambda)) \leq 2^{1-k}C_\epsilon.
-\]
-The existance of $b_k$ is guaranteed by the definition of $T_\epsilon(k-1)$. 
+\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k) } \abs{ X(\lambda) } \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
+\]
+\end{lemma}
+\begin{IEEEproof}
+Let $b_1(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
+The existance of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
+By the triangle inequality
+\[
+\sabs{X(\lambda)} \leq \sabs{X(b_2(\lambda))} + \sabs{X(\lambda) - X(b_2(\lambda))}  
+\]
+and taking supremums on both sides,
+\begin{align}
+&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \nonumber \\
+&\leq \sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}   \nonumber \\
+&\leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}, \label{eq:Xb2recpreexp}
+\end{align}
+the last line because $b_2(\lambda) \in T_\epsilon(k-1)$ and so
+\[
+\sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)}.
+\]
+Conditional on $\{R_i\}$ and $\{\Phi_i\}$ the random variable
+\[
+U(\lambda) = X(\lambda) - X(b_2(\lambda))
+\]
+has zero mean, and it normally distributed with variance
+\begin{align*}
+\sigma_U^2 &= \expect_{R,\Phi} \frac{1}{L} \sum_{i \in P \cup D} g_i^2 \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
+&=  \frac{1}{L} \sum_{i \in P \cup D} \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
+&= d(\lambda, b_2(\lambda) \leq 2^{1-k} C_\epsilon,
+\end{align*}
+since $\expect_{R,\Phi}g_i^2 = 1$.  So, from Lemma~\ref{lem:maxineq},
+\begin{align*}
+\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{U(\lambda)} &\leq 3 \sqrt{ 2^{1-k} C_\epsilon \log \abs{T_\epsilon(k)} } \\
+&\leq \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}  
+\end{align*}
+Taking expectations on both sides of~\eqref{eq:Xb2recpreexp},
+\begin{align*}
+\expect_{R,\Phi}&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \\
+&\leq \expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}  
+\end{align*}
+which involves a recursion in $k$.  By unravelling the recursion and using that $T_\epsilon(k)$ contains only zero, and therefore
+\[
+\expect_{R,\Phi} \sup_{\lambda \in T_k(0)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sabs{X(0)} = 0
+\]
+we obtain,
+\[
+\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \leq  \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}  
+\]
+as required.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:maxineq}(Maximal inequalities)
+Suppose $X_1, \dots, X_N$ are $N$ zero mean Gaussian random variables each with variance less than some positive constant $K$.  Then
+\[
+\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}
+\]
+where $\log N$ is the natural logarithm of $N$.
+\end{lemma}
+\begin{IEEEproof}
+This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:metricentropy}(Covering numbers)
+For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset B_\epsilon$ with the property that, for every $\lambda \in B_\epsilon$, there is a $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the psuedometric
+\[
+d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
+\]
+and the psuedometric
+\[
+d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L,
+\]
+where $C_\epsilon$ and $D_\epsilon$ are defined in~\eqref{eq:Cepsdefn} and~\eqref{eq:Depsdefn}.
+The number of elements in $T_\epsilon(k)$ is no more than $K_3 2^{k}$ where $K_3$ is a positive constant, independent of $L$, $\epsilon$ and $k$.
+\end{lemma}
+\begin{IEEEproof}
+
 \end{IEEEproof}

### 2012-08-16 paper1.tex
Added paper1 and and paper2 just to see how this would look split into too.  Should be fine.

--- 2012-08-14 paper.tex
@@
 
+\usepackage{xr}
+\externaldocument{paper2}
+
 \usepackage{amsmath,amsfonts,amssymb, amsthm, bm}
@@
 %opening
-\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
+\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data 1: \\ Estimators and simulations}
 \author{Robby McKilliam, Andre Pollok, Bill Cowley  
@@
 
-The existing literature mostly considers what is called \emph{noncoherent detection} where no pilots symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
+The existing literature mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
 In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
+Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
@@
 This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.
-For this reason raw symbol error rate is not of specific interest at this stage.
+For this reason raw symbol error rate is not of interest at this stage.
 Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
@@
 \begin{equation}\label{eq:hata}
-\hat{a} = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y
+\hat{a}(\{d_i, i \in D\}) = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y
 \end{equation}
@@
 where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.
-Observe that given a candidate value for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
+Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
 It turns out that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
@@
 Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with pdf $f(r,\phi)$.
-Let
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with joint pdf $f(r,\phi)$.
+Then
 \[
@@
 \]
-Then $g$ and $g_2$ are continuous functions from $\reals$ to $\reals$.
+are continuous functions from $\reals$ to $\reals$.
 \end{lemma}
@@
 
-\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
-
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
- \begin{align*}
-SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
-&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
-\end{align*}
-where
-\[
-Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
-\]
-and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
-\begin{equation}\label{eq:hatrhoZ}
-\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
-\end{equation}
-where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
-\[
-SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
-\]
-
-We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
-Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
-\begin{align*}
-y_i &= a_0 s_i + w_i \\
-&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
-&= a_0 s_i R_i e^{j \Phi_i} \\
-&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
-\end{align*}
-%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
-Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
-\begin{align*}
-\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
-&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
-&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
-&= \round{ \lambda + \Phi_i + \angle{s_i} },
-\end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
-\begin{align}
- y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
-&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
-&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
-\end{align}
-since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
-\[
-y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
-\]
-Now,
-\[
-Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
-\]
-and
-\[
-\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
-\] 
-Let 
-\begin{equation}\label{eq:GLdefn}
-G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
-\end{equation}
-and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
-Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
-We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.
-
-Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
-\[
-\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
-\]
-and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
-\[
-\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
-\]
-As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.
-Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
-Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
-So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
-Part 1 of Theorem~\ref{thm:consistency} follows.
-
-It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
-\begin{equation}\label{eq:rhoGLZ}
-\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
-\end{equation}  
-Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
-
-\begin{lemma}\label{lem:convtoexpGlamL} 
-$\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Since $\sabs{G(x)}$ is uniquely maximised at $x=0$,
-\[
-0 \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)},
-\]
-and since $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(x)}$,
-\[ 
-0 \leq \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}.
-\]
-Thus,
-\begin{align*}
-0 &\leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} \\ 
-& \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} + \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}\\
-& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
-&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
-\end{align*}
-and the last line converges amost surely to zero by Lemma~\ref{lem:uniflawGGL}.
-\end{IEEEproof}
-
-% \begin{lemma}\label{lem:Zconunif}
-% \[
-% \prob\left\{\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} > \epsilon \right\} = O(e^{-\epsilon^2 L}).
-% \]
-% \end{lemma}
-% \begin{IEEEproof}
-% \end{IEEEproof}
-
-\begin{lemma}\label{lem:GLtoG0}
-$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-By the triangle inequality,
-\[
-\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
-\]
-Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:uniflawGGL}
-The function $G_L$ converges almost surely and uniformly on $[-\pi, \pi)$ to the function $G$, that is,
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0
-\] 
-almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Put $T_L(\lambda) = \expect G_L(\lambda)$ and write
-\begin{align*}
-&\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda) + T_L(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)}.
-\end{align*}
-Now,
-\begin{align*}
-T_L(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
-&= o(1) h_1(\lambda) + o(1) h_2(\lambda)
-\end{align*}
-where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since 
-\[
-h_1(\lambda) = \expect R_1\cos(\lambda + \Phi_1) \leq \expect R_1,
-\]
-and 
-\[
-h_2(\lambda) = \expect R_1\cos\fracpart{\lambda + \Phi_1} \leq \expect R_1
-\] 
-for all $\lambda \in [-\pi, \pi)$, it follows that 
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)} \leq o(1) \expect R_1 \rightarrow 0
-\]  
-as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
-\]
-almost surely as $L\rightarrow \infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:uniflawTGL} 
-Put $T_L(\lambda) = \expect G_L(\lambda)$, then
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
-\] 
-almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Put $D_L(\lambda) = G_L(\lambda) - T_L(\lambda)$ and let
-\[
-\lambda_n = \tfrac{2\pi}{N}(n-1) - \pi, \qquad n = 1, \dots, N
-\]
-be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n$ be the interval $[\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
-\begin{align*}
-\sup_{\lambda \in [-\pi, \pi)}&\sabs{D_L(\lambda)} \\
-&= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n) + D_L(\lambda_n)} \\
-&\leq U_L + V_L,
-\end{align*}
-where 
-\[
-U_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)},
-\]
-\[
-V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
-\]
-Lemma~\ref{lem:supDLlambdan} shows that for any finite $N$, and any $\epsilon > 0$, 
-\[
-\prob\left\{ \lim_{L\rightarrow\infty} U_L > \epsilon \right\} = 0,
-\]
-i.e., $U_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
-\] 
-Choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ and
-\begin{align*}
-\prob& \left\{ \lim_{L\rightarrow\infty} \sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} > 3\epsilon \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty}(U_L + V_L) > 3\epsilon \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty} (U_L + V_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty} U_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
-&= 0.
-\end{align*}
-Thus, $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:supDLlambdan}
-If $N$ is finite then $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Put
-\[
-Z_i(\lambda) = \begin{cases}
-R_i\cos(\lambda + \Phi_i) , & i \in P \\
-R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
-\end{cases}
-\]
-so that
-\[
-D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
-\]
-Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
-\[
-\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
-\]
-almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Now, since $N$ is finite,
-\[
-U_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
-\]
-almost surely to zero as $L \rightarrow \infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
-\]
-\end{lemma}
-\begin{IEEEproof}
-Observe that
-\begin{align*}
-&\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
-&=  \abs{G_L(\lambda) - T_L(\lambda) - G_L(\lambda_n) + T_L(\lambda_n)} \\
-&\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
-&\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
-\end{align*}
-the last line following from Jensens inequality.  Put
-\[
-C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) },
-\]
-so that
-\[
-V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \leq C_L + \expect C_L,
-\]
-because 
-\[
- \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect \abs{G_L(\lambda) - G_L(\lambda_n)} \leq \expect C_L
-\]
-follows by moving expectation outside the supremum.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
-\]
-So,
-\begin{align*}
-\prob&\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
-\end{align*}
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:CL} The following statements hold:
-\begin{enumerate}
-\item $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$,
-\item for any $\epsilon > 0$, $\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0$.
-\end{enumerate}
-\end{lemma}
-\begin{IEEEproof}
-If $\lambda \in L_n$, then $\lambda = \lambda_n + \delta$ with $\delta < \tfrac{2\pi}{N}$, and from Lemma~\ref{lem:coslipshitz},
-\begin{align*}
-&\abs{\cos(\lambda + \Phi_i) - \cos(\lambda_n + \Phi_i)} \leq \tfrac{2\pi}{N}, \;\;\; \text{and} \\
-&\abs{\cos\sfracpart{\lambda + \Phi_i} - \cos\sfracpart{\lambda_n + \Phi_i}} \leq \tfrac{2\pi}{N}.
-\end{align*}
-Because this results does not depend on $n$,
-\[
-\sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \leq R_i \frac{2\pi}{N}
-\]
-for all $i = P \cup D$.  Now,
-\begin{align*}
-C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i \in P \cup D} Z_i(\lambda) - Z_i(\lambda_n)} \\
-&\leq \frac{1}{L}\sum_{i \in P \cup D} \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
-&\leq \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i .
-\end{align*}
-Thus, 
-\[
-\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i = \tfrac{2\pi}{N}\expect R_1
-\]
-and the first statement holds.  Now, 
-\[
-\frac{2\pi}{N L}\sum_{i \in P \cup D} R_i \rightarrow \frac{2\pi}{N}\expect R_1
-\] 
-almost surely to $\frac{2\pi}{N}\expect R_1$ as $L \rightarrow\infty$ by the strong law of large numbers, and so, for any $\epsilon > 0$,
-\begin{align*}
-\prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
-\end{align*}
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:coslipshitz}
-Let $x$ and $\delta$ be real numbers.  Then
-\begin{align*}
-&\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta}, \;\;\; \text{and} \\
-&\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
-\end{align*}
-\end{lemma}
-\begin{IEEEproof}
-Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipshitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
-\begin{align*}
-&\abs{\cos(y) - \cos(x)} \leq K\abs{x-y} = \abs{x-y}, \;\;\; \text{and} \\
-&\abs{\cos\fracpart{y} - \cos\fracpart{x}} \leq K\abs{x-y} = \abs{x-y}.
-\end{align*}
- The lemma follows by putting $y = x + \delta$.
-\end{IEEEproof}
-
-\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
-
-We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
-Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
-Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
-The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
-Define the function
-\begin{equation}\label{eq:RLdef}
-R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\end{equation}
-Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
-Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
-\begin{align*}
-0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
-&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
-\end{align*}
-Lemma~\ref{lem:Qconv} shows that
-\[
-\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
-Lemma~\ref{lem:empiricprocc} shows that
-\[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
-\]
-It follows from the three equations above that,
-\[
-0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
-\]
-and rearranging gives,
-\[
-\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
-\]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
-It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
-\[
-\frac{pA_1 + dA_2}{(p + Hd)^2}.
-\]
-
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
-\begin{align*}
-\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
-\end{align*}
-Lemma~\ref{lem:empiricprocforrho} shows that 
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
-\]
-It follows that
-\[
-%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
-\]
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
-Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
-Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
-\[
-\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
-\]
-as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.
-
-\begin{lemma}~\label{lem:diffatlambdaL}
-The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
-\[
-G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0,
-\]
-\end{lemma}
-\begin{IEEEproof}
-Observe that 
-\[
-G_L(\lambda) = \frac{1}{L}\sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}
-\] 
-is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ with $R_i > 0$.
-Let $q_i$ be the smallest number from the interval $[-\tfrac{\pi}{M}, 0]$ such that
-\[
- \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
-\]
-where $\eta = e^{-j2\pi/M} - 1$.  Observe that if $R_i > 0$ then $q_i < 0$.  Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
-\[
-\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i
-\]
-for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.
-
-It remains to show that $G_L^\prime(\hat{\lambda}_L) = 0$.
-If $\sabs{G_L(\hat{\lambda}_L)} > 0$ then $\sabs{G_L(\lambda)}$ is differentiable at $\hat{\lambda}_L$ and, since $\hat{\lambda}_L$ is a maximiser of $\sabs{G_L(\lambda)}$,
-\[
-0 = \frac{d \sabs{G_L(\lambda)}}{d \lambda}(\hat{\lambda}_L) = \begin{cases} 
-G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) > 0\\
--G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) < 0.
-\end{cases}
-\]
-Thus, $G_L^\prime(\hat{\lambda}_L) = 0$ if $\sabs{G_L(\hat{\lambda}_L)} > 0$.
-On the other hand $\sabs{G_L(\hat{\lambda}_L)} = 0$ only if $G_L(\lambda) = 0$ for all $\lambda \in \reals$, inwhich case $G_L^\prime(\hat{\lambda}_L) = 0$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i, i \in D$ be defined as in Lemma~\ref{lem:diffatlambdaL}.
-Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.
-\end{lemma}
-\begin{IEEEproof}
-Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
-\[
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
-\]
-defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
-\begin{equation}\label{eq:fraclambassumption}
-\sfracpart{\hat{\lambda}_L + \Phi_k} > \frac{\pi}{M} + q_k
-\end{equation}
-for some $k \in D$.
-Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.
-We will show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-First observe that,
-\begin{align*}
-Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*.
-\end{align*}
-where $\eta = e^{-j2\pi/M} - 1$ and $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
-\begin{align*}
-SS&(\{r_i, i \in D\}) \\
-&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
-&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
-&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
-&= SS(\{\hat{d}_i, i \in D\}) - D
-\end{align*}
-where 
-\[
-D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
-\]
-Now $\frac{1}{L}\hat{Y} = \hat{a} = \hat{\rho} e^{j\hat{\theta}}$ from~\eqref{eq:hata} and using~\eqref{eq:yethetadhat},
-\[
-\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}},
-\]
-so that
-\begin{equation}\label{eq:Ddefn}
-D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
-\end{equation}
-Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
-\begin{align*}
-\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
-&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
-&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) }
-\end{align*}
-and
-\begin{align*}
-\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
- &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
-\end{align*}
-Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
-\[
-0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
-\]
-Subtituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-Thus,~\eqref{eq:fraclambassumption} is false.
-
-To show that $\sfracpart{\hat{\lambda}_L + \Phi_k} \geq -\frac{\pi}{M} - q_k$ we use contradiction again.
-Assume that $\sfracpart{\hat{\lambda}_L + \Phi_k} < -\frac{\pi}{M} - q_k$.
-Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.
-Now an analagous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
-
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:Qconv}
-Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
-\[ 
-\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where $p, d$ and $H$ are real numbers defined in the statements of Thereoms~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
-\end{lemma}
-\begin{IEEEproof}
-We have
-\[
-Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0) = \tfrac{\sabs{P}}{L} k_1(\lambda) + \tfrac{\sabs{D}}{L} k_2(\lambda)
-\]
-where
-\begin{align}
-k_1(\lambda) &= \expect R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
-k_2(\lambda) &= \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
-\end{align}
-We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\lambda = 0$,
-\[
-\sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
-\]
-where $\zeta$ is a continuous function with $\zeta(0) = 0$.  So,
-\begin{align*}
-k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
-&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big) 
-\end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ has zero mean.
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
-$\zeta(\hat{\lambda}_L) = o_P(1)$.
-So,
-\[
-k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) \expect R_1 \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big)
-\]
-because $\expect R_1$ is finite.  Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
-\begin{align*}
-Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
-&= \hat{\lambda}_L \left( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \right) \\
-&= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
-\end{align*}
-since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the above equation by $\sqrt{L}$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:k2conv}
-$k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
-\end{lemma}
-\begin{IEEEproof}
-Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
-We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
-An analagous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
-To keep our notation clean put
-\[
-\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
-\]
-For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
-\[
-\fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
-\]
-Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
-\]
-whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
-\]
-So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
-&= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
-\end{align*}
-where, for each $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
-\begin{align*}
-\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
-\end{align*}
-where
-\[
-s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
-\]
-and
-\[
-\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos\fracpart{\Phi_1} + \zeta_{k+1}(\lambda).
-\]
-Observe that both $s_k$ and $\eta_k$ are continuous and that 
-\[
-s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
-\]
-
-Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
-\]
-whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
-\]
-Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
-\]
-where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
-\begin{equation}\label{eq:vk}
-v_k(\lambda,\Phi_1) = \begin{cases}
-s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
-\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
-0 & \text{otherwise}.
-\end{cases} 
-\end{equation}
-Now,
-\begin{align*}
-k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
-&= \expect R_1\big( \lambda\cos\fracpart{\Phi_1} + v(\lambda, \Phi_1)  \big) \\
-&= \lambda h_2(0) + \expect R_1 v(\lambda, \Phi_1).
-\end{align*}
-Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a continuous function with $\gamma(0) = 0$, and
-\[
-H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0),
-\]
-where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, it follows that $\gamma(\hat{\lambda}_L) = o_P(1)$, and so
-\[
-k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( h_2(0) + H_1 + o_P(1) ) = \hat{\lambda}_L ( H + o_P(1) )
-\]
-as required.  It remains to prove Lemma~\ref{lem:expRvlamphi}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:expRvlamphi}
-$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
-\end{lemma}
-\begin{IEEEproof}
-In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
-An analgous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
-Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$,
-\begin{align}
-W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
-&= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
-&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
-&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi. \label{eq:Winfsum}
-\end{align}
-Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
-\end{align*}
-where, for $k = 1, \dots, M-1$,
-\begin{equation}\label{eq:ak}
-a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
-\end{equation}
-\[
-b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
-\]
-\begin{equation}\label{eq:ck}
-c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
-\end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous, $c_k(\lambda)$ converges to 
-\[
-s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
-\]
-as $\lambda$ goes to zero from above.
-We are only interested in the limit from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
-The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
-We actually only require $g$ to be continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$, and the requirements of Theorem~\ref{thm:normality} could be weakened in this sense.
-Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.
-
-Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
-\[
-\int_0^{2\pi} v_0(\lambda, \phi) g(\phi) d\phi = \lambda a_0(\lambda) + \lambda b_0(\lambda) +  \lambda c_0(\lambda),
-\]
-where $a_0(\lambda)$ and $c_0(\lambda)$ are defined as in~\eqref{eq:ak}~and~\eqref{eq:ck}, and
-\[
-b_0(\lambda) = \zeta_0(\lambda) \int_{0}^{\psi_k-\lambda} g(\phi) d\phi.
-\]
-When $k = M$, we have $0 < \psi_{M-1} < 2\pi < \psi_{M} - \lambda$, so that
-\[
-\int_0^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi = \lambda b_{M}(\lambda)
-\]
-where 
-\[
-b_{M}(\lambda) = \zeta_{M}(\lambda) \int_{\psi_{M-1}}^{2\pi}g(\phi) d\phi.
-\]
-Both $b_0(\lambda)$ and $b_M(\lambda)$ are functions that converge to zero as $\lambda$ converges to zero.
-
-Finally, when $k < 0$ or $k > M$, 
-\[
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0.
-\]
-So the infinite sum in~\eqref{eq:Winfsum} can be written as the finite sum,
-\begin{align*}
-W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \left( C(\lambda)  + A(\lambda) + B(\lambda) \right),
-\end{align*}
-where 
-\[
-A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
-\]
-\[
-C(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
-\]
-Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
-\[
-\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1.
-\]
-Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$ as required.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:empiricprocc} Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
-\[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
-\]
-\end{lemma}
-\begin{IEEEproof}
-Put 
-\begin{equation}\label{eq:WLdef}
-W_L(\lambda) = \sqrt{L}\big( R_L(\lambda) - Q_L(\lambda) - R_L(0) \big)
-\end{equation}
-so that
-\[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = W_L(\hat{\lambda}_L) + \sqrt{L} R_L(0)
-\]
-Lemma~\ref{lem:ZYLempiricproc} in the Appendix shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
-\[
-\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu
-\]
-for all positive integers $L$.  Since $\hat{\lambda}_L$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
-\[
-\lim_{L\rightarrow\infty}\prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} = 0
-\] 
-and therefore $\prob\{ \sabs{\hat{\lambda}_L} \geq \epsilon \} < \nu$ for all sufficiently large $L$.  Now
-\begin{align*}
-  \prob&\left\{\sabs{ W_L(\hat{\lambda}_L) } > \delta \right\} \\
-&= \prob\left\{ \sabs{W_L(\hat{\lambda}_L)} > \delta \;\text{and} \; \sabs{\hat{\lambda}_L} < \epsilon \right\} \\
-&\hspace{0.7cm} + \prob\left\{ \sabs{W_L(\lambda_L)} > \delta  \;\text{and} \; \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
-&\leq \prob\left\{  \sup_{\lambda < \epsilon} \sabs{ W_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
-&\leq 2\nu.
-\end{align*}
-for all sufficiently large $L$.  Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $W_L(\hat{\lambda}_L)$ converges in probability to zero as $N\rightarrow\infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:convdistGLdash}
-The distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$.
-\end{lemma}
-\begin{IEEEproof}
-Observe that 
-\[
-\sqrt{L} R_L(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
-\]
-From the standard central limit the distribution of
-\[
-\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) = \sqrt{\frac{\sabs{P}}{L}} \frac{1}{\sqrt{\sabs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
-\]
-converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
-\[
-p A_1 = p \expect R_1^2 \sin^2(\Phi_1).
-\]
-Similarly, the distribution of 
-\[ 
-\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} = \sqrt{\frac{\sabs{D}}{L}} \frac{1}{\sqrt{\sabs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
-\]
-converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~? BLERG, and variance
-\[
-d A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:empiricprocforrho} Let $T_L(\lambda) = \expect G_L(\lambda)$.  Then
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big)$.
-\end{lemma}
-\begin{IEEEproof}
-Let
-\begin{equation}\label{eq:YLdef}
-Y_L(\lambda) = \sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) - X_L
-\end{equation}
-so that
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\lambda) + X_L
-\]
-Lemma~\ref{lem:ZYLempiricproc} in the appendix 
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:HLtoG} $\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1)$.
-\end{lemma}
-\begin{IEEEproof}
-The argument is similar to that used in Lemma~\ref{lem:Qconv}.  First observe that
-\begin{align*}
-T_L(\lambda) &= \frac{\sabs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\sabs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
-%&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
-%& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
-&=  p \expect R_1\cos(\lambda + \Phi_1) + d\expect R_1\cos\fracpart{\lambda + \Phi_1} + o(1),
-\end{align*}
-because $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L\rightarrow\infty$. Put
-\begin{align*}
-L(\lambda) &= \sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
-\end{align*}
-where
-\[
-q_1(\lambda) = \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big),
-\]
-\begin{equation}
-q_2(\lambda) = \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
-\end{equation}
-We analyse $q_1(\lambda)$ first.  By a Taylor expansion of $\cos(\lambda + \Phi_1)$ about $\lambda = 0$,
-\[
-\cos(\lambda + \Phi_1) - \cos(\Phi_1) = -\lambda \big( \sin(\Phi_1) + \zeta(\lambda) \big),
-\]
-where $\zeta$ is a continuous function with $\zeta(0) = 0$. So,
-\begin{align*}
-q_1(\lambda) &= -\expect R_1\lambda(\sin(\Phi_1) +  \zeta(\lambda)) \\
-&= -\lambda \big( \expect R_1\sin(\Phi_1) +  \zeta(\lambda) \expect R_1 \big) \\
-&= \lambda \zeta(\lambda) K,
-\end{align*}
-since $\expect R_1\sin(\Phi_1) = 0$ because $w_1$ has zero mean,  and $K = \expect R_1$ is a finite constant because $\expect\sabs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.  So, 
-\begin{align*}
-\sqrt{L} q_1(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L \zeta(\hat{\lambda}_L) K = o_P(1)
-\end{align*}
-because $\sqrt{L} \hat{\lambda}_L$ converges in distribution and $\zeta(\hat{\lambda}_L) = o_P(1)$.
-An argument analogous to that used in Lemma~\ref{lem:k2conv} shows that $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
-Thus,
-\[
-L(\lambda) = p q_1(\lambda) + d q_2(\lambda) + o(1) = o_P(1)
-\]
-as reuqired.
-\end{IEEEproof}
-
-% \begin{lemma}\label{lem:XL} 
-% $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
-% \end{lemma}
-% \begin{IEEEproof}
-% The argument is similar to that used in Lemma~\ref{lem:k2conv}.
-As $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $q_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
-We will analyse $q_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
-An analagous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
-Recall that
-% \[
-% \psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
-% \]
-% when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
-% \begin{align*}
-% \cos\fracpart{\lambda + \Phi_1} - &\cos\fracpart{\Phi_1} \\
-% &= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
-% \end{align*}
-% and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
-% \begin{align*}
-% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} &= -\lambda\big(\sin(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
-% &= -\lambda\big(\sin\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
-% \end{align*}
-% where, for each integer $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.   Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
-% \begin{align*}
-% \cos\fracpart{\lambda + \Phi_1} &- \cos\fracpart{\Phi_1} \\
-% &= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k)
-% \end{align*}
-% and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
-% \begin{align*}
-% \cos&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-% &= c_k(\Phi_1) - \lambda\big(\sin\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
-% \end{align*}
-% where
-% \[
-% c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
-% \]
-% and
-% \[
-% \eta_k(\lambda, \Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_{k+1}(\lambda).
-% \]
-% Observe that $c_k$ is continuous and $c_k(\psi_{k}) = 0$.
-
-% Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
-% \[
-% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
-% \]
-% whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
-% \[
-% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda \sin\fracpart{\Phi_1} + c_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
-% \]
-% Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
-% \[
-% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
-% \]
-% where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
-% \begin{equation}\label{eq:cvk}
-% v_k(\lambda,\Phi_1) = \begin{cases}
-% c_k(\Phi_1) - \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
-% \lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
-% 0 & \text{otherwise}.
-% \end{cases}
-% \end{equation}
-% \end{IEEEproof}
-
-\begin{lemma}\label{lem:XL} 
-The distribution of 
-\[
-X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big) = \sqrt{L} \big( G_L(0) - \expect G_L(0) \big)
-\] 
-converges to the normal with zero mean and covariance 
-\[
-pB_1 + d B_2
-\] 
-as $L \rightarrow\infty$.
-\end{lemma}
-\begin{IEEEproof}
-Observe that $X_L = C_L + D_L$ where
-\[
-C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} \big( R_i \cos(\Phi_i) - h_1(0) \big),
-\]
-\[
-D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ).
-\]
-From the standard central limit the distribution of $C_L$ converges to the normal with zero mean and variance
-\[
-p B_1 = p \expect R_1^2 \cos^2(\Phi_1) - p h_1^2(0) =  p \expect R_1^2 \cos^2(\Phi_1) - p 
-\]
-since $h_1(0) = 1$.  The distribution of $D_L$ converges to the normal with zero mean and variance
-\[
-d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0)
-\]
-The proof follows since $C_L$ and $D_L$ are indpendent.
-\end{IEEEproof}
-
 \section{The Gaussian noise case}
@@
 
-\subsection{An empirical process result}
-
-During the proof of asymptotic normality in Section~\ref{sec:proof-asympt-norm} we made use of the following result regarding the functions $W_L(\lambda)$ and $Y_L(\lambda)$ defined in~\eqref{eq:WLdef}~and~\eqref{eq:YLdef}.
-
-\begin{lemma}\label{lem:ZYLempiricproc}
-For any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that:
-\begin{enumerate}
-\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu$, and,
-\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu$,
-\end{enumerate}
-for all positive integers $L$.
-\end{lemma}
-\begin{IEEEproof}
-These results are related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
-The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
-We will only give a proof of the first statement regarding the function $W_L$.
-The proof for the second statement regarding $Y_L$ is similar.
-
-Put 
-\[
-f_i(\lambda, R_i, \Phi_i) = \begin{cases}
-R_i( \sin(\lambda + \Phi_i) - \sin(\Phi_i)), & i \in P \\
-R_i( \sin\sfracpart{\lambda + \Phi_i} - \sin\sfracpart{\Phi_i}), & i \in D \\
-\end{cases}
-\]
-so that
-\[
-W_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect f_i(\lambda, R_i, \Phi_i) \big).
-\]
-Let $\{g_i\}$ be a sequence of independent standard normal random variables, independent of $\{R_i\}$ and $\{\Phi_i\}$.  Lemma~\ref{lem:symmetrisation} shows that
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) },
-\]
-where 
-\begin{equation}\label{eq:ZpsiCondGaussProc}
-X_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} g_i f_i(\lambda, R_i, \Phi_i),
-\end{equation}
-and where $\expect$ runs over all of $\{g_i\}$, $\{R_i\}$ and $\{\Phi_i\}$.
-Conditionally on $\{\Phi_i\}$ and $\{R_i\}$, $\{X_L(\lambda), \lambda \in \reals \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
-Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists $\epsilon > 0$ such that
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } < \kappa.
-\]
-It follows that,
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)}  <  \sqrt{2\pi} \; \kappa,
-\]
-and by Markov's inequality,
-\[
-\prob \cubr{  \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
-\]
-for any $\delta > 0$.  The proof follows by choosing $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:symmetrisation}
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) }
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$.
-Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
-Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
-Then
-\begin{align*}
-W_L(\lambda) &= \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect_{R,\Phi} f_i(\lambda, R_i', \Phi_i') \big) \\
-&= \expect_{R,\Phi} \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ),
-\end{align*}
-where, for notational convenience, we put 
-\[
-f_{i} = f_i(\lambda, R_i, \Phi_i) \qquad \text{and} \qquad  f_{i}^\prime = f_i(\lambda, R_i', \Phi_i').
-\] 
-Taking absolute values followed by supremums,
-\begin{align*}
- \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &= \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \expect_{R,\Phi} \sum_{i \in P \cup D} (f_{i} - f_{i}^\prime )} \\
-&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi}  \abs{ \frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) },
-\end{align*}
-the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, the bound can be increased by moving $\expect_{R,\Phi}$ outside the supremum,
-\begin{equation}\label{eq:Eoutsidesup}
- \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ) }.
-\end{equation}
-Applying $\expect$ to both sides gives the inequality
-\begin{equation}\label{eq:Gninequfin}
- \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} (f_{i} - f_{i}^\prime) },
-\end{equation}
-since $\expect \expect_{R,\Phi}$ is equivalent to $\expect$.   Let
-\[
-\sigma_i = \frac{g_i}{\abs{g_i}},
-\]
-and put $\sigma_i = 1$ if $g_i = 0$.  The $\sigma_i$ are thus independent with 
-\[
-\prob\{ \sigma_i = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_i = 1 \} = \tfrac{1}{2}.
-\]  
-As $R_i$ and $\Phi_i$ are independent and identically distributed to $R_i^\prime$ and $\Phi_i^\prime$, the random variable $f_{i} - f_{i}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_i( f_{i} - f_{i}^\prime)$.
-Thus,
-%\[
-%\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
-%\]
-%are identically distributed, and therefore
-\begin{align*}
- \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) } \\
-&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i ( f_{i} - f_{i}^\prime) }.
-\end{align*}
-By the triangle inequality
-\[
-\abs{ \sum_{i\in P \cup D} \sigma_i (f_{i} - f_{i}^\prime)} \leq \abs{ \sum_{i \in P \cup D} \sigma_i f_{i} } + \abs{ \sum_{i\in P \cup D}  \sigma_i f_{i}^\prime },
-\]
-and it follows from~\eqref{eq:Gninequfin}, that
-\begin{align*}
- \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &\leq \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} \sigma_i f_{i} } \\
-&\hspace{1cm} + \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  \sigma_i  f_{i}^\prime } \\
-&= 2 \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} }.
-\end{align*}
-Because $g_i$ is a standard normal random variable it follows that $\expect \abs{g_i} = \expect_{R,\Phi} \abs{g_i} = \sqrt{2/\pi}$, and
-\begin{align*}
-\expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} } \\
-&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} \sqrt{\frac{\pi}{2}}\expect_{R,\Phi} \abs{g_i}} \\
-&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{ \sqrt{L}}\sum_{i \in P \cup D} \sigma_i \abs{g_i}  f_{i}},
-\end{align*}
-the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$, it follows that
-\begin{align*}
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} g_i  f_{i}} \\
-&= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X_L(\lambda) } 
-\end{align*}
-as required.
-\end{IEEEproof}
-
-\begin{lemma} \label{lem:chaining}
-For any $\kappa > 0$ there exists $\epsilon > 0$ such that
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } < \kappa,
-\]
-where $X_L(\lambda)$ is defined in~\eqref{eq:ZpsiCondGaussProc}.
-\end{lemma}
-\begin{IEEEproof}
-Without loss of generality, assume that $\epsilon < \tfrac{\pi}{M}$.  Lemma~\ref{lem:chaining2} shows that
-\begin{equation}\label{eq:supZCK1} 
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{\frac{C_\epsilon}{L}}
-\end{equation}
-where $K_1$ is a finite, positive constant, and
-\[
-C_\epsilon = 2\epsilon\sum_{i\in P \cup D}R_i^2 + 4\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
-\]
-where $I_{\epsilon}(\Phi_i)$ returns one when $\Phi_i \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon]$ and zero otherwise.  Now,
-\[
-\tfrac{1}{L}\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
-\]
-Recalling that $f(r,\phi)$ is the joint probability density function of $R_i$ and $\Phi_i$,
-\begin{align*}
-\expect R_i^2 I_\epsilon(\Phi_i) &= \int_0^{2\pi} \int_{0}^\infty  r^2 f(r,\phi) I_\epsilon(\phi) dr d\phi \\ 
-&= \int_0^{2\pi}g_2(\phi) I_\epsilon(\phi) d\phi 
-\end{align*}
-where the function 
-\[
-g_2(\phi) = \int_{0}^\infty  r^2 f(r,\phi) dr d\phi 
-\]
-is continuous due to Lemma~\ref{lem:gg2cont}.  From the definition of $I_\epsilon$,
-\begin{align*}
-\expect R_i^2 I_\epsilon(\Phi_i) &= \sum_{k=0}^{M-1}\int_{\psi_k-\epsilon}^{\psi_k-\epsilon}g_2(\phi) d\phi \\
-&= 2\epsilon \sum_{k=0}^{M-1} (g_2(\psi_k) + o(1))
-\end{align*}
-where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.
-Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive, then
-\[
-\expect \sqrt{\frac{C_\epsilon}{L}} \leq  \sqrt{\tfrac{1}{L}\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
-\]
-by Jensen's inequality.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\psibf) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
-\]
-Choosing 
-\[
-\epsilon = \frac{\kappa^2}{ K_1^2 ( K_2 + 2\expect R_1^2)}
-\] 
-completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
-\end{IEEEproof}
-
-The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
-The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
-%Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.
-For notational convenience we use the abbreviation 
-\[
-f_{i}(\lambda) = f_{i}(\lambda, R_i, \Phi_i)
-\]
-in what follows.
-As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{1}{m+1}$.
-%This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.
-
-\begin{lemma}\label{lem:chaining2}
-There exists a positive constant $K_1$ such that
-\[
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X_L(\lambda) } \leq K_1 \sqrt{\frac{C_\epsilon}{L}}.
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $B_\epsilon = \{x \in \reals \mid \abs{x} < \epsilon\}$ be the set of real numbers with magnitude less than $\epsilon$.
-For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $B_\epsilon$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the psuedometric
-\[
-d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon,
-\]
-and the psuedometric
-\[
-d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon
-\]
-where 
-\[
-D_\epsilon = \epsilon\sum_{i\in P \cup D}R_i + \sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
-\]
-We specifically define $T_\epsilon(0)$ to contain only zero.
-Defined this way, $T_\epsilon(0)$ satisfies the inequality above because both $d_1(\lambda, 0)$ and $d_2(\lambda, 0)$ are less than $C_\epsilon$ for all $\lambda \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
-The existance of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{}.
-
-Lemma~\ref{lem:limexpTk} shows that
-\[
-\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}.
-\]
-Lemma~\ref{lem:chaining3} shows that
-\[
-\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} \leq \sqrt{\frac{C_\epsilon}{L}} \sum_{i=1}^{k} \frac{?}{2^{i/2}}
-\]
-The sum above converges as $k\rightarrow\infty$ so lemma holds with
-\[
-K_1 = \sum_{i=1}^{\infty} \frac{?}{2^{i/2}}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:limexpTk}
-\[
-\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}
-\]
-\end{lemma}
-\begin{IEEEproof}
-Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
-\[
-\sup_{\lambda \in T_\epsilon(k)} \sabs{X_L(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \sabs{X_L(\lambda)}.
-\]
-Let $\lambda^* \in B_\epsilon$ be a maximiser of $\sabs{X_L(\lambda)}$, that is
-\[
-\sabs{X_L(\lambda^*)} = \sup_{\sabs{\lambda} < \epsilon}\sabs{X_L(\lambda)}.
-\]
-By definition $T_\epsilon(k)$ contains a point $\lambda_2$ such that $d_2(\lambda^*,\lambda_2) \leq 2^{-k}D_\epsilon$.  Now
-\begin{align*}
-\sabs{X_L(\lambda^*) - X_L(\lambda_2)} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda^*) - f_i(\lambda_2))} \\
-&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} \sum_{i\in P \cup D}\sabs{f_i(\lambda^*) - f_i(\lambda_2)} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} d(\lambda^*,\lambda_2) \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon.
-\end{align*}
-where $g_\text{max} = \max_{i\in P\cup D} \abs{g_i}$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:epslmlemma}
-Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
-\begin{align*}
-d(\lambda, \lambda^*) &= \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \\
-&\leq R_i^2 L \left( (\lambda - \lambda^*)^2 + 4 C_\epsilon \right).
-\end{align*}
-\end{lemma}
-
-\begin{lemma}
-If $i \in P$, then $\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2.$
-\end{lemma}
-\begin{IEEEproof} 
-We have
-\[
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 = R_i^2\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)}^2
-\]
-and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
-\[
-\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)} \leq K\sabs{\lambda - \lambda^*} = \sabs{\lambda - \lambda^*}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}
-Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$.  If $i \in D$, then
-\begin{equation}\label{eq:fismall}
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2
-\end{equation}
-when $\Phi_i \notin S_\epsilon$, otherwise,
-\begin{equation}\label{eq:fibig}
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq 4 R_i^2.
-\end{equation}
-\end{lemma}
-\begin{IEEEproof}
-Because $\sin(\cdot) \leq 1$ it follows that
-\begin{align*}
-\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 &= R_i^2\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}}^2 \\
-&\leq 4 R_i^2
-\end{align*}
-for any $\Phi_i \in \reals$, and so~\eqref{eq:fibig} holds.  If $\Phi_i \notin S_\epsilon$ then $\Phi_i \in (\psi_{k-1} + \epsilon, \psi_{k} - \epsilon)$ for some $k\in\ints$ and 
-\[
-\fracpart{\Phi_i + \lambda} = \Phi_i + \lambda - \tfrac{2\pi}{M}k
-\] 
-and
-\[
-\fracpart{\Phi_i + \lambda^*} = \Phi_i + \lambda^* - \tfrac{2\pi}{M}k 
-\]
-and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
-\begin{align*}
-\vert \sin&\fracpart{\Phi_i + \lambda} - \sin\fracpart{\Phi_i + \lambda^*}\vert \\
-&=  \sabs{\sin(\Phi_i + \lambda - \tfrac{2\pi}{M}k) - \sin(\Phi_i + \lambda^* - \tfrac{2\pi}{M}k)} \\
-&\leq K \sabs{\lambda - \lambda^*} \\
-&= \sabs{\lambda - \lambda^*}.
-\end{align*}
-Thus~\eqref{eq:fismall} holds. 
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:chaining3}(Chaining)
-For all positive integers $k$,
-\[
-\expect_{R,\Phi} \sup_{\psibf \in T_\epsilon(k) } \abs{ X_L(\lambda) } \leq \sqrt{C_\epsilon} \sum_{i=1}^{k}\frac{?}{2^{i/2}}.
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $b_k(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that 
-\[
-d(\lambda, b_k(\lambda)) \leq 2^{1-k}C_\epsilon.
-\]
-The existance of $b_k$ is guaranteed by the definition of $T_\epsilon(k-1)$. 
-\end{IEEEproof}
-
 \end{document}

### 2012-08-16 paper2.tex
Added paper1 and and paper2 just to see how this would look split into too.  Should be fine.

--- 2012-08-15 paper.tex
@@
 
+\usepackage{xr}
+\externaldocument{paper1}
+
 \usepackage{amsmath,amsfonts,amssymb, amsthm, bm}
@@
 %opening
-\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
+\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data 2: Proofs}
 \author{Robby McKilliam, Andre Pollok, Bill Cowley 
@@
 \section{Introduction}
-
-In passband communication systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation (amplitude change).
-These effects must be compensated for at the receiver.
-In this paper we assume that the time offset has been previously handled, and we focus on estimating the phase shift and attenuation.
-We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK) and $M$-ary phase shift keying ($M$-PSK).
-In this case, the transmitted symbols take the form,
-\[
-s_i = e^{j u_i},
-\]
-where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M \geq 2$ is the size of the constellation.
-We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.
-So,
-\[
-s_i = \begin{cases}
-p_i & i \in P \\
-d_i & i \in D,
-\end{cases}
-\]
-where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.
-The sets $P$ and $D$ are disjoint, i.e.
-$P \cap D = \emptyset$, and $L = \abs{P \cup D}$ is the total number of symbols transmitted.
-
-We assume that time offset estimation and matched filtering have been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
-\begin{equation}\label{eq:sigmod}
-y_i = a_0 s_i + w_i, \qquad i \in P \cup D,
-\end{equation}
-where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
-Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
-Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the reciever and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
-A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
-\begin{align*}
-SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \\
-&= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2,
-\end{align*}
-where $\abs{x}$ denotes the magnitude of the complex number $x$.
-The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
-However, as we will show, the estimator works well under less stringent assumptions.
-%It is the least squares estimator that we primarily study in this paper.
-
-The existing literature mostly considers what is called \emph{noncoherent detection} where no pilots symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
-In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
-A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-The idea is the `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
-A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
-\[
-A = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
-\]
-Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included, so as to make the estimator coherent.
-Techniques are available but they are somewhat adhoc~\cite{Cowley_ref_sym_carr_1998}.
-This problem does not occur with the least square estimator.
-
-An important paper is by Mackenthun~\cite{Mackenthun1994} who described an algorithm to compute the least squares estimator requiring only $O(L \log L)$ arithmetic operations.
-Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
-Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-
-In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
-The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
-Whilst estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite approach here.
-Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
-This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.
-For this reason raw symbol error rate is not of specific interest at this stage.
-Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
-The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
-The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
-
-The paper is organised as follows BLERG.
-
-%It is worth commenting on our use of $\prob$ rather than the more usual $\expect$ or $E$ for the expected value operator.
-The notation comes from Pollard~\cite[Ch 1]{Pollard_users_guide_prob_2002}.
-The notation is good because it removes unecessary distinction between `probability' and expectation.
-Given a random variable $X$ with cumulative density function $F$, the probability of an event, say $X \in S$, where $S$ is some subset of the range of $X$, is
-%\[
-%\prob \indicator \{X \in S\} = \int \indicator \{X \in S\}(x) dF(x) = \int_{S} dF(x)
-%\]
-%where $\indicator \{X \in S\}$ is the indicator function of the set $S$, i.e $\indicator \{X \in S\}(x) = 1$ when the argument $x \in S$ and zero otherwise.
-We will usually drop the $\onebf$ and simply write $\prob \{ X \in S \}$ to mean $\prob \onebf\{ X \in S \}$.
-To illustrate the utility of this notation, Markov's inequality becomes
-%\[
-%\prob \{ \abs{X} > \delta \}  \leq \prob \frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \} \leq \frac{1}{\delta}\prob\abs{X},
-%\]
-%where $\frac{\abs{X}}{\delta}\onebf\{ \abs{X} > \delta \}(x)$ is the function equal to $\abs{x}/\delta$ when the argument $x > \delta$ and zero otherwise.
-
-\section{Mackentun's algorithm with pilots}\label{sec:least-squar-estim}
-
-In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to include the pilot symbols.
-For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
-In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.
-This complexity arises from the need to sort a list of $\abs{D}$ elements.
-%We use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
-
-%In Section~\ref{sec:line-time-algor} we will show that a full sort is not neccesary, and that the least squares estimator can be implemented in $O(L)$ operations.
-
-Define the sum of squares function
-\begin{align}
-SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
-&= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*, \label{eq:SS}
-\end{align}
-where $*$ denotes the complex congugate.
-Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find the least squares estimator of $a_0$ as a function of $\{d_i, i \in D\}$,
-\begin{equation}\label{eq:hata}
-\hat{a} = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y
-\end{equation}
-where $L = \abs{P \cup D}$ is the total number of symbols transmitted, and to simplify our notation we have put 
-\[
-Y = \sum_{i \in P \cup D} y_i s_i^* = \sum_{i \in P } y_i p_i^* + \sum_{i \in D } y_i d_i^*.
-\]  
-Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to surpress the argument $(\{ d_i, i \in D\})$ for notational clarity.
-Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain the $SS$ conditioned on minimisation with respect to $a$,
-\begin{equation}\label{eq:SSdatasymbols}
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y}^2,
-\end{equation}
-where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.
-Observe that given a candidate value for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-It turns out that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
-%When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
-
-To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a nonnegative real.  Now,
-\begin{align}
-SS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
-&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:SSallparams}
-\end{align}
-For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
-\begin{equation}\label{eq:hatdfinxtheta}
-\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
-\end{equation}
-where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
-A word of caution, the notation $\round{\cdot}$ is often used to denote rounding to the nearest \emph{integer}.
-This is not the case here.
-If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer then,
-\[
-\round{x} = \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
-\] 
-Note that $\hat{d}_i(\theta)$ does not depend on $\rho$.
-As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
-With this in mind,
-\[
-\hat{u}_i(\theta) = \round{\angle{y_i} - \theta }
-\]
-which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $2\pi$.
-
-We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Let $b_i = \hat{d}_i(0)$ and let 
-\[
-z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}.
-\]
-Then,
-\begin{equation}\label{eq:uicombos}
-\hat{d}_i(\theta) = 
-\begin{cases}
-b_i, &  0 \leq \theta < z_i + \frac{\pi}{M} \\
-b_i e^{-j2\pi/M}, & z_i + \frac{\pi}{M} \leq \theta < z_i + \frac{3\pi}{M} \\ 
-\vdots & \\
-b_i e^{-j2\pi k /M}, & z_i + \frac{\pi(2k - 1)}{M} \leq \theta < z_i + \frac{\pi(2k + 1)}{M}  \\ 
-\vdots & \\
-b_i e^{-j2\pi}, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
-\end{cases}
-\end{equation}
-
-Let 
-\[
-f(\theta) = \{ \hat{d}_i(\theta), i \in D \}
-\]
-be a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
-Observe that $f(\theta)$ is piecewise continuous.
-The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains contant are determined by the values of $\{z_i, i \in D\}$.
-Let
-\[
-S = \{ f(\theta) \mid \theta \in [0, 2 \pi) \}
-\]
-be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$.
-If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
-
-The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indices in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
-\begin{equation}\label{eq:sigmasortind}
-z_{\sigma(i)} \leq z_{\sigma(k)}
-\end{equation}
-whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}-1\}$.
-It is convenient to define the indices into $\sigma$ to be taken modulo $\abs{D}$, that is, if $m$ is an integer not from $\{0, 1, \dots, \abs{D}-1\}$ then we define $\sigma(m) = \sigma(k)$ where $k \equiv m \mod \abs{D}$ and $k \in  \{0, 1, \dots, \abs{D}-1\}$.
-The first sequence in $S$ is
-\[
-f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
-\]  
-The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/M}$.
-Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.
-Using this notation,
-\[
-f_1 = f_0 e_{\sigma(0)}.
-\] 
-The next sequence in $S$ is correspondingly 
-\[
-f_2 = f_0 e_{\sigma(0)} e_{\sigma(1)} = f_1 e_{\sigma(1)},
-\]
-and the $k$th sequence is
-\begin{equation}\label{eq:fkrec}
-f_{k+1} = f_{k} e_{\sigma(k)}.
-\end{equation}
-In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively enumerated.
-
-We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,(M+1)\abs{D}-1\}$.
-Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
-So, this na\"{\i}ve approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
-Mackenthun~\cite{Mackenthun1994} showed how $SS(f_k)$ can be computed recursively.
-
-Let,
-\begin{equation}\label{eq:SSfk}
-SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
-\end{equation}
-where, 
-\begin{align*}
-Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
-&= B + \sum_{i \in D}g_{ki},
-\end{align*}
-where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
-Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
-\[
-g_{k+1} = g_{k} e_{\sigma(k)}^*,
-\]
-where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/M}$.  Now,
-\[
-Y_0 = B + \sum_{i \in D} g_{0i}
-\] 
-can be computed in $O(L)$ operations, and
-\begin{align*}
-Y_1 &= B + \sum_{i \in D} g_{1i} \\
-&= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
-&= Y_0 + \eta g_{0\sigma(0)},
-\end{align*}
-where $\eta = e^{j2\pi/M} - 1$.  In general,
-\[
-Y_{k+1} = Y_k + \eta g_{k\sigma(k)}.
-\]
-So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a constant number of operations.
-Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.
-Let $\hat{k} = \arg\min SS(f_k)$.
-The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata},
-\begin{equation}\label{eq:ahatYhat}
-\hat{a} = \frac{1}{L} Y_{\hat{k}}.
-\end{equation}
-Psuedocode is given in Algorithm~\ref{alg:loglinear}.
-Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
-The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
-This requires $O(L \log L)$ operations.
-The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.
-The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
-%In the next sections we will show how to the sortinces function can be avoided.
-This leads to an algorithm that requires only $O(L)$ operations.
-
-\begin{algorithm}[t] \label{alg:loglinear}
-\SetAlCapFnt{\small}
-\SetAlTitleFnt{}
-\caption{Mackenthun's algorithm with pilot symbols}
-\DontPrintSemicolon
-\KwIn{$\{y_i, i \in P \cup D \}$}
-\For{$i \in D$ \nllabel{alg_loop_setup}}{
-$\phi = \angle{y_i}$ \;
-$u = \round{\phi} $ \;
-$z_i = \phi -  u $ \;
-$g_i = y_i e^{-j u}$ \;
-}
-$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
-$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
-$\hat{a} = \frac{1}{L} Y$ \;
-$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
-$\eta = e^{j2\pi/M} - 1$ \;
-\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
-$Y = Y + \eta g_{\sigma(k)}$ \;
-$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
-$Q = \frac{1}{L}\abs{Y}^2$\;
-\If{$Q > \hat{Q}$}{
- 	$\hat{Q} = Q$ \;
- 	$\hat{a} =  \frac{1}{L} Y$ \;
- }
-}
-\Return{$\hat{a}$ \nllabel{alg_return}}
-\end{algorithm}
-
-\section{Circularly symmetric complex random variables}\label{sec:circ-symm-compl}
-
-Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
-A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
-That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
-If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
-\[
-f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
-\]  
-Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariant to phase rotation.
-If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
-\begin{align*}
-\expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
-&= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
-&= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
-\end{align*}
-
-We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
-\[
-R e^{j\Phi} = 1 + X.
-\]
-The joint distribution of $R$ and $\Phi$ can be shown to be
-\[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}.
-\]
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
-\begin{equation}\label{eq:expectRepartRphi}
-\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
-\end{equation}
-where $\Re(\cdot)$ denotes the real part, and
-\begin{equation}\label{eq:expectImpartRphi}
-\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
-\end{equation}
-where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
-
-\begin{lemma}\label{lem:gg2cont}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with pdf $f(r,\phi)$.
-Let
-\[
-g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
-\]
-and
-\[
-g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr.
-\]
-Then $g$ and $g_2$ are continuous functions from $\reals$ to $\reals$.
-\end{lemma}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
-% \begin{lemma}\label{lem:h1minedcircsym}
-% Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
-Let
-% \[
-% h_1(x) = \expect R \cos(x + \Phi).
-% \]
-% Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
-% \end{lemma}
-% % Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
-That is, the pdf of $Z e^{j \Theta}$ in rectangular coordinates is given by $z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$ and $x$ and $y$ are the real and imaginary parts of $Z e^{j \Theta}$, and this pdf is non increasing with $z$.
-For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
-
-% % By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if
-% % \[
-% % \abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
-% % \]
-% % We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
-% \begin{IEEEproof}
-% BLERG
-% \end{IEEEproof}
-
-\section{Statistical properties of the least squares estimator}\label{sec:stat-prop-least}
-
-In this section we decribe the asymptotic properties of the least squares estimator.
-In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
-\[
-\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right),
-\]
-where $\operatorname{round}(\cdot)$ takes its argument to the nearest integer.
-The direction of rounding for half-integers is not important so long as it is consistent.
-We have chosen to round up half-integers here.
-Similarly we use $\sfracpart{x}$ to denote $x$ taken `modulo $\tfrac{2\pi}{M}$' into the interval $\left[-\tfrac{\pi}{M}, \tfrac{\pi}{M}\right)$, that is
-\[
-\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right) = x - \round{x}.
-\]
-The next two theorems decribe the asymptotic properties of the least squares estimator.
-
-\begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.
-Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
-%where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
-Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
-\[
-\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
-\] 
-Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
-\begin{equation}\label{eq:RiandPhii}
-R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
-\end{equation}
-and define the continuous function
-\[
-G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
-\]
-\[
-h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
-\]
-If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
-\begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
-\end{enumerate}
-\end{theorem}
-
-\begin{theorem}\label{thm:normality} (Asymptotic normality)
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
-\[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
-\]
-Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-%If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
-\[
-\left( \begin{array}{cc} 
-\frac{pA_1 + dA_2}{(p + H d)^2} & 0 \\
-0 & \rho_0^2(pB_1 + dB_2)
-\end{array} \right)
-\]
-as $L \rightarrow \infty$, where
-\[
-H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
-\]
-\[
-A_1 = \var\big( R_1^2\sin^2(\Phi_1) \big), \qquad A_2 = \var\big( R_1^2\sin^2\fracpart{\Phi_1} \big),
-\]
-\[
-B_1 = \var\big( R_1^2 \cos^2(\Phi_1) \big), \;\;\; B_2 = \var\big( R_1^2 \cos^2\fracpart{\Phi_1} \big).
-\]
-\end{theorem}
-
-\begin{corollary}\label{cor:ampmse}
-The mean square error of the amplitude estimator $\hat{\rho}$ is 
-\[
-\expect (\rho_0 - \hat{\rho})^2 = \rho_0^2\frac{p B_1 + d B_2}{L} + \rho_0^2(1 - G(0))^2 + o(L^{-1/2}).
-\] 
-\end{corollary}
-\begin{IEEEproof} Write
-\begin{align*}
-\expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
-&= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
-&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L.
-\end{align*} 
-From Theorem~\ref{thm:normality} it follows that $\expect \hat{m}_L^2 = \tfrac{B}{L} + o(L^{-1})$ and that $\expect \hat{m}_L = o(L^{-1/2})$. 
-\end{IEEEproof}
-
-The proof of Theorem~\ref{thm:consistency} is in Section~\ref{sec:proof-almost-sure} and the proof of Theorem~\ref{thm:normality} is in Section~\ref{sec:proof-asympt-norm}.
-Before giving the proofs we discuss the assumptions made by these theorems.
-The assumption that $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
-If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $\{s_i,i \in P \cup D\}$.
-In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $\{s_i,i \in P \cup D\}$, rather than just $\rho_0$.
-The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
-
-The assumption that $\expect \sabs{w_1}^2 = \expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.
-This is required in Theorem~\ref{thm:normality} so that the constants $A_1$, $A_2$, $B_1$ and $B_2$ exist.
-We will also use that $R_i$ has finite variance to simplify the proof of Theorem~\ref{thm:consistency} by use of Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}
-
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta} - \theta_0$.
-This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
-So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
-
-A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
-
-\begin{lemma}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
-Let,
-\[
-h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{ x + \Phi}.
-\]
-Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-\end{lemma}
-Before giving the proof we discuss the requirements of this lemma.
-The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the pdf of $X$ decreases as we move away from the origin.
-That is, the pdf of $X$ in rectangular coordinates is $p(x,y) = z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$, and this is required to be non increasing with $z$.
-For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
 
@@
 
-\section{The Gaussian noise case}
-
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.
-The joint density function of the real and imaginary parts is then
-\[
-p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
-\]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
-%and the joint density function of the real and imaginary parts of $Z$ is
-%\[
-%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
-%\] 
-%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
-It can be shown that
-\[
-g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-\]
-where $\kappa = \rho_0/\sigma$ and
-\[
-\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
-\]
-The value of $A_1$ and $A_2$ can be computed by numerical integration.
-
-\section{Simulations}\label{sec:simulations}
-
-%\subsection{The coherent setting}
-
-Our first set of simulations considers the case where there are a number of pilot symbols.
-
-%BPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-%		\caption{Amplitude error for BPSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error for BPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure*}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure*}
-
-%QPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-%		\caption{Amplitude error for QPSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error for QPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM4-3.mps}
-% 		\caption{complex amplitude error for QPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%8PSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-%		\caption{Amplitude error for 8PSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error for 8PSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%\subsection{The noncoherent setting}
-
-%BPSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-% 		\caption{Amplitude error for BPSK}
-% 		\label{fig:plotamp}
- %\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
-		\caption{Phase error for BPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%QPSK plots
- %\begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-% 		\caption{Amplitude error for QPSK}
-% 		\label{fig:plotamp}
-% \end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
-		\caption{Phase error for QPSK}
-		\label{fig:plotphase}
-\end{figure}
-
- % \begin{figure}[tp]
- % 	\centering
- %         \includegraphics{code/data/plotM4-3.mps}
- % 		\caption{complex amplitude error for QPSK}
- % 		\label{fig:plotcamp}
- % \end{figure}
-
-%8PSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-% 		\caption{Amplitude error for 8PSK}
-% 		\label{fig:plotamp}
-% \end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
-		\caption{Phase error for 8PSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
 \section{Conclusion}

### 2012-08-16 paper1.tex
modified the 2 parters.  The second paper is not just appendix.  I think this makes alot of sense.

@@
 %opening
-\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data 1: \\ Estimators and simulations}
+\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
 \author{Robby McKilliam, Andre Pollok, Bill Cowley  
@@
 
-The paper is organised as follows BLERG.
+The paper is organised in the following way.
+Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when pilot symbols and data symbols exist.
+Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
+Section~\ref{sec:stat-prop-least} states two theorms that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
+We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
+However, $\hat{\rho_0}$ is not a consistent estimator of the amplitude $\rho_0$.
+The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
+Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.
+Due to space constraints, the lemmas needed by these proofs have been placed in the supplementary paper~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
+In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
+In this case, our expressions for the asymptotic variance can be simplified.
+Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.
+These simulations agree with the derived asymptotic properties.
 
@@
 
-\section{The Gaussian noise case}
+\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
+
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
+ \begin{align*}
+SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
+&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
+\end{align*}
+where
+\[
+Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
+\]
+and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
+\begin{equation}\label{eq:hatrhoZ}
+\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
+\end{equation}
+where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
+\[
+SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
+\]
+
+We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
+Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
+\begin{align*}
+y_i &= a_0 s_i + w_i \\
+&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
+&= a_0 s_i R_i e^{j \Phi_i} \\
+&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
+\end{align*}
+%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
+Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
+\begin{align*}
+\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
+&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
+&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
+&= \round{ \lambda + \Phi_i + \angle{s_i} },
+\end{align*}
+where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
+\begin{align}
+ y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
+&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
+&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
+\end{align}
+since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
+\[
+y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
+\]
+Now,
+\[
+Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
+\]
+and
+\[
+\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
+\] 
+Let 
+\begin{equation}\label{eq:GLdefn}
+G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
+\end{equation}
+and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
+Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
+We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
+The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.
+
+Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
+\[
+\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
+\]
+and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
+\[
+\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
+\]
+As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
+Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.
+Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
+Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
+Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
+So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
+Part 1 of Theorem~\ref{thm:consistency} follows.
+
+It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
+\begin{equation}\label{eq:rhoGLZ}
+\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
+\end{equation}  
+Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+
+\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
+
+We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
+Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
+Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
+The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
+This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
+Define the function
+\begin{equation}\label{eq:RLdef}
+R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+\end{equation}
+Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
+Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
+\begin{align*}
+0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
+&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
+\end{align*}
+Lemma~\ref{lem:Qconv} shows that
+\[
+\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\]
+where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+Lemma~\ref{lem:empiricprocc} shows that
+\[
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
+\]
+It follows from the three equations above that,
+\[
+0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
+\]
+and rearranging gives,
+\[
+\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
+\]
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
+It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
+\[
+\frac{pA_1 + dA_2}{(p + Hd)^2}.
+\]
+
+We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
+\begin{align*}
+\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
+&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
+\end{align*}
+Lemma~\ref{lem:empiricprocforrho} shows that 
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
+\]
+where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
+\[
+%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
+\]
+It follows that
+\[
+%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
+\]
+Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
+Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
+Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
+\[
+\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
+\]
+as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.
+
+\section{The Gaussian noise case}\label{sec:gaussian-noise-case}
 
@@
 
-\normalsize
-\appendix
-
 %\section{Tricks with fractional parts}

### 2012-08-16 paper2.tex
modified the 2 parters.  The second paper is not just appendix.  I think this makes alot of sense.

@@
 %opening
-\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data 2: Proofs}
+\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data (Appendix)}
 \author{Robby McKilliam, Andre Pollok, Bill Cowley 
@@
 
-\section{Introduction}
-
-\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
-
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
- \begin{align*}
-SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
-&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
-\end{align*}
-where
-\[
-Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
-\]
-and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
-\begin{equation}\label{eq:hatrhoZ}
-\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
-\end{equation}
-where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
-\[
-SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
-\]
-
-We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
-Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
-\begin{align*}
-y_i &= a_0 s_i + w_i \\
-&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
-&= a_0 s_i R_i e^{j \Phi_i} \\
-&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
-\end{align*}
-%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
-Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
-\begin{align*}
-\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
-&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
-&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
-&= \round{ \lambda + \Phi_i + \angle{s_i} },
-\end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
-\begin{align}
- y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
-&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
-&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
-\end{align}
-since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
-\[
-y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
-\]
-Now,
-\[
-Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
-\]
-and
-\[
-\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
-\] 
-Let 
-\begin{equation}\label{eq:GLdefn}
-G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
-\end{equation}
-and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
-Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
-We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.
-
-Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
-\[
-\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
-\]
-and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
-\[
-\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
-\]
-As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.
-Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
-Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
-So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
-Part 1 of Theorem~\ref{thm:consistency} follows.
-
-It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
-\begin{equation}\label{eq:rhoGLZ}
-\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
-\end{equation}  
-Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+\section{Introduction to this appendix}
+
+Here we compile lemmas required for the proof almost sure convergence (Theorem~\ref{thm:consistency}) and asymptotic normality (Theorem~\ref{thm:normality}) from the paper~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+This document is not self contained and it is intended that the reader start at~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+The equation reference numbers in this document start at (?).
+All reference numbers less than (?) refer back to~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+This document is separated into 3 sections.
+Section~\ref{sec:proof-almost-sureappendix} contains those lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}).
+Section~\ref{sec:proof-asympt-normappendix} contains those lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}).
+Section~\ref{sec:an-empirical-process} contains proofs for lemmas that arise during Section~\ref{sec:proof-asympt-normappendix}.
+These involve the proof of~\emph{tightness} for certain empirical processes~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
+
+\section{Lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sureappendix}
 
@@
 
-\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
-
-We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
-Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
-Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
-The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
-Define the function
-\begin{equation}\label{eq:RLdef}
-R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\end{equation}
-Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
-Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
-\begin{align*}
-0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
-&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
-\end{align*}
-Lemma~\ref{lem:Qconv} shows that
-\[
-\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
-Lemma~\ref{lem:empiricprocc} shows that
-\[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
-\]
-It follows from the three equations above that,
-\[
-0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
-\]
-and rearranging gives,
-\[
-\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
-\]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
-It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
-\[
-\frac{pA_1 + dA_2}{(p + Hd)^2}.
-\]
-
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
-\begin{align*}
-\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
-\end{align*}
-Lemma~\ref{lem:empiricprocforrho} shows that 
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
-\]
-It follows that
-\[
-%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
-\]
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
-Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
-Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
-\[
-\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
-\]
-as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.
+\section{Lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-normappendix}
 
@@
 
-\section{Conclusion}
-
-\small
-\bibliography{bib}
-
-\normalsize
-\appendix
-
 %\section{Tricks with fractional parts}
@@
 
-\subsection{An empirical process result}
+\section{An empirical process result}\label{sec:an-empirical-process}
 
@@
 
+\small
+\bibliography{bib}
+
 \end{document}

### 2012-08-16 paper1.tex
added to papers.  There's some fiddly stuff to do.  Should probably think about running the simulations if I'm tied of proving stuff.

@@
 \]  
-Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariant to phase rotation.
+Observe that for any real number $\phi$, the pdf of $X$ and $e^{j\phi}X$ are the same, that is, the pdf is invariant to phase rotation.
 If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
@@
 \]
-The joint distribution of $R$ and $\Phi$ can be shown to be
-\[
+The joint pdf of $R$ and $\Phi$ can be shown to be
+\begin{equation}\label{eq:pdfRPhi}
 f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}.
-\]
+\end{equation}
 The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
@@
 \end{equation}
-where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
+where $\Im(\cdot)$ denotes the imaginary part.  The next result will be useful for the analysis of the least squares estimator.  %Me ommit the proof, which is straightforward, but tedious.
 
-\begin{lemma}\label{lem:gg2cont}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
+\begin{proposition}\label{lem:gg2cont}
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ that is continuous on the positive reals and with $\expect \abs{X}^2 = \expect Z^2$ finite.
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with joint pdf $f(r,\phi)$.
@@
 \[
-g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr.
+g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
 \]
 are continuous functions from $\reals$ to $\reals$.
-\end{lemma}
+\end{proposition}
 \begin{IEEEproof}
-BLERG
+From~\eqref{eq:pdfRPhi},
+\[
+g(\phi) = \int_{0}^\infty \frac{r^2 f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}} dr,
+\]
+and
+\[
+g_2(\phi) = \int_{0}^\infty \frac{r^3 f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}} dr.
+\]
+The continuity of these functions follows immediately from the continuity of $f_Z$.
 \end{IEEEproof}
@@
 
-\begin{lemma}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
+\begin{proposition}
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ that is continuous and differentiable on the positive reals.
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
@@
 \[
-h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{ x + \Phi}.
+h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
 \]
-Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-\end{lemma}
+If $z^{-1} f_Z(z)$ is nonincreasing with $z$, then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
+\end{proposition}
 Before giving the proof we discuss the requirements of this lemma.
@@
 \]
-as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.
+as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
 

### 2012-08-16 paper2.tex
added to papers.  There's some fiddly stuff to do.  Should probably think about running the simulations if I'm tied of proving stuff.

@@
 
+%set this to the number of the last equation in the previous document.
+\setcounter{equation}{20}
+\newcounter{lasteqplusone}
+\setcounter{lasteqplusone}{\value{equation}}
+\addtocounter{lasteqplusone}{1}
+
 \section{Introduction to this appendix}
@@
 This document is not self contained and it is intended that the reader start at~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
-The equation reference numbers in this document start at (?).
-All reference numbers less than (?) refer back to~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+The equation reference numbers in this document start at (\arabic{lasteqplusone}).
+All equation numbers less than (\arabic{lasteqplusone}) refer back to~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
 This document is separated into 3 sections.
@@
 \begin{IEEEproof}
-BLERG
+We prove the lemma for $d_2$.  The proof for $d_1$ is similar.  Observe that $d_2(\lambda, 0) = c_1 + c_2$ where 
+ \[
+ c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) }
+ \]
+ and 
+ \[
+ c_2 = \sum_{i \in D} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+ \]
+ As $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+ \[
+ \sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i)} \leq K\sabs{\lambda} = \sabs{\lambda} < \epsilon
+ \]
+ for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let $S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}$. Now $c_2 = c_3 + c_4$ where 
+ \[
+ c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}
+ \]
+ and
+ \[
+ c_4 =  \sum_{i \in D, i \notin S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+ \]
+ If $i \notin S$ then $\round{\lambda + \Phi_i} = \round{\Phi_i}$ for all $\sabs{\lambda} < \epsilon$ and therefore
+ \begin{align*}
+ \vert\sin&\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}\vert \\
+ &= \sabs{\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\Phi_i - \round{\Phi_i})} \\
+ &\leq \sabs{\lambda} < \epsilon
+ \end{align*}
+and so, 
+\[
+c_4 \leq \epsilon \sum_{i \in D, i \notin S} R_i \leq \epsilon \sum_{i \in D} R_i.
+\]
+Otherwise, if $i \in S$, then $\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}} \leq 2$ because $\sin(\cdot) \leq 1$.  Thus 
+\[
+c_3 \leq 2 \sum_{i \in D, i \in C} R_i \leq 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i).
+\]
+Now
+\begin{align*}
+d_2(\lambda, 0) &= c_1 + c_2 \\
+&\leq c_1 + c_3 + c_4 \\
+&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i)
+\end{align*}
+as required
 \end{IEEEproof}

### 2012-08-17 paper.tex
Added to paper parts.

@@
 \]
-The joint distribution of $R$ and $\Phi$ can be shown to be
-\[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}.
+The joint pdf of $R$ and $\Phi$ can be shown to be
+\[
+f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
 \]
@@
 \begin{lemma}\label{lem:gg2cont}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with pdf $f(r,\phi)$.
-Let
+Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with continuous pdf $\tfrac{1}{2\pi}f_Z(z)$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with joint pdf $f(r,\phi)$.
+Then
 \[
@@
 \]
-Then $g$ and $g_2$ are continuous functions from $\reals$ to $\reals$.
-\end{lemma}
-\begin{IEEEproof}
-BLERG
+are continuous functions from $\reals$ to $\reals$.
+\end{lemma}
+\begin{IEEEproof}
+We only give a proof for $g_2$, the proof for $g$ being similar.  From 
 \end{IEEEproof}

### 2012-08-17 paper1.tex
Added to paper parts.

--- 2012-08-16 paper1.tex
@@
 Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
-A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
-That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
+A complex random variable $W$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{W}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{W}$.
+That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
 If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
@@
 \]
-Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariant to phase rotation.
-If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
+If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
+\[
+f_{X,Y}(x,y) = \frac{f_Z(z)}{2\pi z} \qquad \text{where $z = \sqrt{x^2 + y^2}$.}
+\]
+Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
+If $\expect\abs{W} = \expect Z$ is finite, then $W$ has zero mean because
 \begin{align*}
-\expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
+ \expect W &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
  &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
@@
 
-We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
+We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
 \[
-R e^{j\Phi} = 1 + X.
+R e^{j\Phi} = 1 + W.
 \]
-The joint distribution of $R$ and $\Phi$ can be shown to be
-\[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}.
-\]
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
+The joint pdf of $R$ and $\Phi$ can be shown to be
+\begin{equation}\label{eq:pdfRPhi}
+f(r,\phi) = \frac{r f_Z(\sqrt{\abs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\abs{r^2 - 2r\cos\phi + 1}}}.
+\end{equation}
+The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
 \begin{equation}\label{eq:expectRepartRphi}
@@
 \end{equation}
-where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
-
-\begin{lemma}\label{lem:gg2cont}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with joint pdf $f(r,\phi)$.
-Then
-\[
-g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
-\]
-and
-\[
-g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr.
-\]
-are continuous functions from $\reals$ to $\reals$.
-\end{lemma}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
+where $\Im(\cdot)$ denotes the imaginary part.
 
@@
 \[
-A_1 = \var\big( R_1^2\sin^2(\Phi_1) \big), \qquad A_2 = \var\big( R_1^2\sin^2\fracpart{\Phi_1} \big),
+A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
 \]
 \[
-B_1 = \var\big( R_1^2 \cos^2(\Phi_1) \big), \;\;\; B_2 = \var\big( R_1^2 \cos^2\fracpart{\Phi_1} \big).
+B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
 \]
@@
 
-\begin{corollary}\label{cor:ampmse}
-The mean square error of the amplitude estimator $\hat{\rho}$ is 
-\[
-\expect (\rho_0 - \hat{\rho})^2 = \rho_0^2\frac{p B_1 + d B_2}{L} + \rho_0^2(1 - G(0))^2 + o(L^{-1/2}).
-\] 
-\end{corollary}
-\begin{IEEEproof} Write
-\begin{align*}
-\expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
-&= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
-&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L.
-\end{align*} 
-From Theorem~\ref{thm:normality} it follows that $\expect \hat{m}_L^2 = \tfrac{B}{L} + o(L^{-1})$ and that $\expect \hat{m}_L = o(L^{-1/2})$. 
-\end{IEEEproof}
+% \begin{corollary}\label{cor:ampmse}
+% The mean square error of the amplitude estimator $\hat{\rho}$ is 
+% \[
+% \expect (\rho_0 - \hat{\rho})^2 = \rho_0^2\frac{p B_1 + d B_2}{L} + \rho_0^2(1 - G(0))^2 + o(L^{-1/2}).
+% \] 
+% \end{corollary}
+% \begin{IEEEproof} Write
+% \begin{align*}
+% \expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
+% &= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
+% &= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L.
+% \end{align*} 
+% From Theorem~\ref{thm:normality} it follows that $\expect \hat{m}_L^2 = \tfrac{B}{L} + o(L^{-1})$ and that $\expect \hat{m}_L = o(L^{-1/2})$. 
+% \end{IEEEproof}
 
@@
 
+For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ in order the constant $H$ to exist.
+The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.
+At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.
+
+\begin{proposition}\label{prop:gg2cont}
+Let $W$ be a circularly symmetric complex random variable with $\expect\abs{W}^2$ finite.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
+Let $f(r,\phi)$ be the joint pdf of $R$ and $\Phi$.
+Then
+\[
+g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
+\]
+and
+\[
+g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
+\]
+are even functions on $(-\pi,\pi)$ and are continuous at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$.
+\end{proposition}
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
+
 A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
@@
 The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+The zero mean complex Gaussian distribution with independent real and imaginary parts is apart of this class.
 
-\begin{lemma}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+\begin{proposition}\label{prop:contgg2}
+Let $W = X + jY$ be a circularly symmetric complex random variable with pdf $f_{X,Y}(x,y)$ of the real and imaginary parts nonincreasing with $x^2 + y^2$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
 Let,
 \[
-h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{ x + \Phi}.
+h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
 \]
 Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-\end{lemma}
-Before giving the proof we discuss the requirements of this lemma.
-The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the pdf of $X$ decreases as we move away from the origin.
-That is, the pdf of $X$ in rectangular coordinates is $p(x,y) = z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$, and this is required to be non increasing with $z$.
-For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
+\end{proposition}
 \begin{IEEEproof}
@@
 \end{IEEEproof}
+
+% \begin{IEEEproof}
+% We only give a proof for $g_2$, the proof for $g$ begin similar.  From~\eqref{eq:pdfRPhi},
+% \[
+% g_2(\phi) = \int_{0}^\infty \frac{r^3 f_Z(\sqrt{\sabs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\sabs{r^2 - 2r\cos\phi + 1}}} dr.
+% \]
+% That $g_2$ is even on $[-\pi,\pi]$ is now obvious because $\cos\phi$ is even on $[-\pi,\pi]$.  Put $a = \cos\phi$ and $z^2 = r^2 - 2ra + 1$ and by a change of variable from $r$ to $z$,
+% \[
+% g_2(\phi) = \int_{1}^{\sqrt{1-a^2}} \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz + \int_{\sqrt{1-a^2}}^\infty \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz.
+% \]
+% when $0 < a \leq 1$ and
+% \[
+% g_2(\phi) = \int_{1}^\infty \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz.
+% \]
+% when $-1 \leq a \leq 0$.  Because
+% \[
+% \frac{r^3}{(r - a)} \leq a z^2 +  b
+% \]
+% for all $\phi \in \reals$, it follows that
+% \[
+% g_2(\phi) \leq \int_{0}^\infty \frac{(a z^2 + d) f_Z(z)}{2\pi} dz = a \expect Z^2 + b
+% \]
+% which is finite.  BLERG work out a and b.
+% \end{IEEEproof}
 
@@
 \]
-The value of $A_1$ and $A_2$ can be computed by numerical integration.
+The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
 

### 2012-08-17 paper2.tex
Added to paper parts.

@@
 \title{Carrier phase and amplitude estimation for phase shift keying using pilots and data (Appendix)}
-\author{Robby McKilliam, Andre Pollok, Bill Cowley 
+\author{Robby McKilliam, Andre Pollok, Bill Cowley, I. Vaughan L. Clarkson and Barry Quinn
 \thanks{
@@
 
-%set this to the number of the last equation in the previous document.
+%set sections and equation counters based on the previous document.
 \setcounter{equation}{20}
@@
 \addtocounter{lasteqplusone}{1}
+\setcounter{section}{9}
+\newcounter{lastsecplusone}
+\setcounter{lastsecplusone}{\value{section}}
+\addtocounter{lastsecplusone}{1}
 
@@
 Here we compile lemmas required for the proof almost sure convergence (Theorem~\ref{thm:consistency}) and asymptotic normality (Theorem~\ref{thm:normality}) from the paper~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
-This document is not self contained and it is intended that the reader start at~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+This document is not self contained and it is intended that the reader start at~\cite{McKilliam_leastsqPSKpilotsdata_2012} and refer to this document when prompted in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
 The equation reference numbers in this document start at (\arabic{lasteqplusone}).
 All equation numbers less than (\arabic{lasteqplusone}) refer back to~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+The section numbers in this document start at~\arabic{lastsecplusone} and all sections before~\arabic{lastsecplusone} are contain in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
 This document is separated into 3 sections.
@@
 Section~\ref{sec:proof-asympt-normappendix} contains those lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}).
-Section~\ref{sec:an-empirical-process} contains proofs for lemmas that arise during Section~\ref{sec:proof-asympt-normappendix}.
+Section~\ref{sec:an-empirical-process} contains proofs for lemmas that arise in Section~\ref{sec:proof-asympt-normappendix}.
 These involve the proof of~\emph{tightness} for certain empirical processes~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
@@
 Thus, $G_L^\prime(\hat{\lambda}_L) = 0$ if $\sabs{G_L(\hat{\lambda}_L)} > 0$.
-On the other hand $\sabs{G_L(\hat{\lambda}_L)} = 0$ only if $G_L(\lambda) = 0$ for all $\lambda \in \reals$, inwhich case $G_L^\prime(\hat{\lambda}_L) = 0$.
+On the other hand $\sabs{G_L(\hat{\lambda}_L)} = 0$ only if $G_L(\lambda) = 0$ for all $\lambda \in \reals$, and in this case $G_L^\prime(\hat{\lambda}_L) = 0$ also.
 \end{IEEEproof}
@@
 \end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous, $c_k(\lambda)$ converges to 
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont} in~\cite{McKilliam_leastsqPSKpilotsdata_2012}, it follows $c_k(\lambda)$ converges to
 \[
@@
 as $\lambda$ goes to zero from above.
-We are only interested in the limit from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
+We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
 The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
-We actually only require $g$ to be continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$, and the requirements of Theorem~\ref{thm:normality} could be weakened in this sense.
-Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.
 
@@
 \]
-Lemma~\ref{lem:ZYLempiricproc} in the Appendix shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+Lemma~\ref{lem:ZYLempiricproc} in Section~\ref{sec:an-empirical-process} shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
 \[
@@
 \]
-converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~? BLERG, and variance
+converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~\ref{lem:expectImRfracpart}, and variance
 \[
 d A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
+\]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:expectImRfracpart}
+$\expect R_1 \sin\fracpart{\Phi_1} = 0$.
+\end{lemma}
+\begin{IEEEproof}
+Recalling the $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
+\begin{align*}
+\expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
+&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi,
+\end{align*}
+where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ are periodic with period $2\pi$.
+Propostion~\ref{prop:gg2cont} in \cite{McKilliam_leastsqPSKpilotsdata_2012} shows that $g$ is even on $(-\pi,\pi)$.
+Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.
+So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
+\[
+g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
+\]
+and
+\[
+\sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
+\]
+Now,
+\begin{align*}
+I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
+&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
+&\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+&= 0
+\end{align*}
+for all integers $k$ because
+\begin{align*}
+\int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
+&= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
+\end{align*}
+Observe that the intervals
+\[
+ [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
+\]
+partition $[-\pi, \pi]$, and so
+\[
+\expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
 \]
@@
 \]
-Lemma~\ref{lem:ZYLempiricproc} in the appendix 
+Lemma~\ref{lem:ZYLempiricproc} in Section~\ref{sec:an-empirical-process} shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+\[
+\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu.
+\]
+The proof now follows by an argument analogous to that in Lemma~\ref{lem:empiricprocc}.
 \end{IEEEproof}
@@
 \]
-is continuous due to Lemma~\ref{lem:gg2cont}.  From the definition of $I_\epsilon$,
+is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont} in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+From the definition of $I_\epsilon$,
 \begin{align*}

### 2012-08-17 paper2.tex
minor edits

@@
 
-\begin{lemma}\label{lem:symmetrisation}
+\begin{lemma}\label{lem:symmetrisation} (Symmetrisation)
 \[
@@
 Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
-Then
+Now,
 \begin{align*}
@@
 \end{align*}
-the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$, it follows that
+the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$,
 \begin{align*}
@@
 Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive, it follows that
+Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive,
 \[
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem:epslmlemma}
 Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.

### 2012-08-18 paper2.tex
added to papers

@@
 
-Here we compile lemmas required for the proof almost sure convergence (Theorem~\ref{thm:consistency}) and asymptotic normality (Theorem~\ref{thm:normality}) from the paper~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
-This document is not self contained and it is intended that the reader start at~\cite{McKilliam_leastsqPSKpilotsdata_2012} and refer to this document when prompted in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+Here we compile lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) and asymptotic normality (Theorem~\ref{thm:normality}) from the paper~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+This document is not self contained.
+It is intended that the reader start at~\cite{McKilliam_leastsqPSKpilotsdata_2012} and refer to this document when prompted in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
 The equation reference numbers in this document start at (\arabic{lasteqplusone}).
 All equation numbers less than (\arabic{lasteqplusone}) refer back to~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
-The section numbers in this document start at~\arabic{lastsecplusone} and all sections before~\arabic{lastsecplusone} are contain in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
+The section numbers in this document start at~\arabic{lastsecplusone} and all sections before~\arabic{lastsecplusone} are contained in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
 This document is separated into 3 sections.
@@
 \begin{equation}\label{eq:Depsdefn}
-D_\epsilon = \frac{1}{L}\epsilon\sum_{i\in P \cup D}R_i + \frac{1}{L}\sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
+D_\epsilon = \frac{1}{L}\epsilon\sum_{i\in P \cup D}R_i + \frac{2}{L}\sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
 \end{equation}
@@
 \[
-c_3 \leq 2 \sum_{i \in D, i \in C} R_i \leq 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i).
-\]
-Now
+c_3 \leq 2 \sum_{i \in D, i \in S} R_i \leq 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i).
+\]
+Now,
 \begin{align*}
@@
 &\leq c_1 + c_3 + c_4 \\
-&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i)
+&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i) \\
+&= D_\epsilon L
 \end{align*}
@@
 &\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} \sum_{i\in P \cup D}\sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
+&\leq \frac{1}{\sqrt{L}} \sum_{i\in P \cup D} g_\text{max} \sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
 &\leq \frac{g_\text{max}}{\sqrt{L}} d_2(\lambda,b_2(\lambda)) \\
@@
 \end{align*}
-where $g_\text{max} = \max_{i\in P\cup D} \abs{g_i}$. The bound applies uniformly over $\lambda \in B_\epsilon$, so
+where $g_\text{max} = \sup_{i\in P\cup D} \abs{g_i}$. The bound applies uniformly over $\lambda \in B_\epsilon$, so
 \[
@@
 \]
-Applying $\expect_{R,\Phi}$ to both sides of this inequality,
+Applying $\expect_{R,\Phi}$ to both sides,
 \[
@@
 \]
-is a constant, independent of $\epsilon$ and $k$.  Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
+is independent of $\epsilon$ and $k$.  Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
 \[
@@
 and so,
-\begin{align}
+\begin{align*}
 0 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \nonumber \\
@@
 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda) - X(b_2(\lambda))} \nonumber \\
-&\leq K_4 2^{-k} D_\epsilon L, \label{eq:Deps2nktozero}
-\end{align}
-the second line because,
+&\leq K_4 2^{-k} D_\epsilon L. \label{eq:Deps2nktozero}
+\end{align*}
+The second line in the inequaliy above follows because,
 \[
@@
 \]
-since $b_2(\lambda) \in T_\epsilon(k)$.  The lemma follows from~\eqref{eq:Deps2nktozero} because $K_4 2^{-k} D_\epsilon L$ converges to zero as $k\rightarrow\infty$.
+since $b_2(\lambda) \in T_\epsilon(k)$.  The lemma follows because $K_4 2^{-k} D_\epsilon L$ goes to zero as $k\rightarrow\infty$.
 \end{IEEEproof}
@@
 \begin{IEEEproof}
-Let $b_1(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
+Let $b_1(\lambda)$ map each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
 The existance of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
@@
 \]
-has zero mean, and it normally distributed with variance
+has zero mean and is normally distributed with variance
 \begin{align*}
@@
 &=  \frac{1}{L} \sum_{i \in P \cup D} \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
-&= d(\lambda, b_2(\lambda) \leq 2^{1-k} C_\epsilon,
+&= d(\lambda, b_2(\lambda)) \leq 2^{1-k} C_\epsilon,
 \end{align*}
@@
 \expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{U(\lambda)} &\leq 3 \sqrt{ 2^{1-k} C_\epsilon \log \abs{T_\epsilon(k)} } \\
-&\leq \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}  
+&\leq \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}.
 \end{align*}
@@
 \expect_{R,\Phi}&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \\
-&\leq \expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}  
+&\leq \expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}} ,
 \end{align*}
@@
 \begin{lemma}\label{lem:maxineq}(Maximal inequalities)
-Suppose $X_1, \dots, X_N$ are $N$ zero mean Gaussian random variables each with variance less than some positive constant $K$.  Then
+Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$.  Then
 \[

### 2012-08-18 paper1.tex
added to papers
started filling in the simulation section.

@@
 Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
 }}
@@
 
-For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ in order the constant $H$ to exist.
+For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.
 The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.
@@
 Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
+These are proved in Section~\ref{sec:proof-almost-sureappendix} of \cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
 
@@
 \]
-as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
+as $L \rightarrow \infty$.
+It remains to prove the lemmas that we have used.
+These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
 
@@
 
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.
-The joint density function of the real and imaginary parts is then
+Let the noise sequence $\{w_i\}$ is complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
 \[
-p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
+\frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
 \]
@@
 
-%\subsection{The coherent setting}
-
-Our first set of simulations considers the case where there are a number of pilot symbols.
+\begin{itemize}
+\item Run simulations looking at what happens as the number of pilots changes.
+\item Simulations showing what happens as $L$ changes for fixed number of pilots.
+\item Show that sometimes just using pilots is better!
+\end{itemize}
 
@@
 
-%\subsection{The noncoherent setting}
-
-%BPSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-% 		\caption{Amplitude error for BPSK}
-% 		\label{fig:plotamp}
- %\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
-		\caption{Phase error for BPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%QPSK plots
- %\begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-% 		\caption{Amplitude error for QPSK}
-% 		\label{fig:plotamp}
-% \end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
-		\caption{Phase error for QPSK}
-		\label{fig:plotphase}
-\end{figure}
-
- % \begin{figure}[tp]
- % 	\centering
- %         \includegraphics{code/data/plotM4-3.mps}
- % 		\caption{complex amplitude error for QPSK}
- % 		\label{fig:plotcamp}
- % \end{figure}
-
-%8PSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-% 		\caption{Amplitude error for 8PSK}
-% 		\label{fig:plotamp}
-% \end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
-		\caption{Phase error for 8PSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
 \section{Conclusion}

### 2012-08-18 code/clt.scala

@@ -11,7 +11,8 @@ import pubsim.distributions.complex.ComplexRandomVariable
 import pubsim.Util
 
 val Ms = List(2,4,8) //BPSK, QPSK, 8-PSK
-val Ls = List(2,4,8,16,32,64,128,256,512,1024,2048)
+val Ls = List(32,128,256,1024,4096)
+//val Ls = List(2,4,8,16,32,64,128,256,512,1024,2048)
 val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
 
 //construct an array of noise distributions with a logarithmic scale
@@ -23,7 +24,7 @@ val starttime = (new java.util.Date).getTime
 for( L <- Ls; M <- Ms ) {
 
   //for a range of different numbers of pilots
-  for( numpilots <- List( L/64, L/32, L/16, L/8, L/4, L/2, L ).distinct.filterNot(b => b == 0) ) {
+  for( numpilots <- List( 0, L/64, L/32, L/16, L/8, L/4, L/2, L ).distinct ) {
 
     val P = 0 until numpilots //pilots at the front
     val D = numpilots until L //data at the back

### 2012-08-18 code/montecarlonc.scala

@@ -11,7 +11,7 @@ import cam.noise.ComplexGaussian
 import pubsim.Util
 
 val Ms = List(2,4,8) //BPSK, QPSK, 8-PSK
-val Ls = List(32, 256, 2048)
+val Ls = List(32,256,1024,4096)
 val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
 val iters = 5000
 

### 2012-08-19 code/clt.scala
Added new L=128 data.  Modified figures in the paper.  Still needs a bit more

@@ -11,8 +11,9 @@ import pubsim.distributions.complex.ComplexRandomVariable
 import pubsim.Util
 
 val Ms = List(2,4,8) //BPSK, QPSK, 8-PSK
-val Ls = List(32,128,256,1024,4096)
 //val Ls = List(2,4,8,16,32,64,128,256,512,1024,2048)
+//val Ls = List(32,128,256,1024,4096)
+val Ls = List(128)
 val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
 
 //construct an array of noise distributions with a logarithmic scale

### 2012-08-19 code/montecarlonc.scala

@@ -11,7 +11,8 @@ import cam.noise.ComplexGaussian
 import pubsim.Util
 
 val Ms = List(2,4,8) //BPSK, QPSK, 8-PSK
-val Ls = List(32,256,1024,4096)
+//val Ls = List(32,128,256,1024,4096)
+val Ls = List(128)
 val a0 = new PolarComplex(1,2*scala.math.Pi*(new scala.util.Random).nextDouble)
 val iters = 5000
 

### 2012-08-20 paper1.tex
Added new L=128 data.  Modified figures in the paper.  Still needs a bit more
minor scala edits.

@@
 
-%BPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-%		\caption{Amplitude error for BPSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
 \begin{figure}[tp]
@@
 		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error for BPSK}
+		\caption{Phase error versus SNR for BPSK with $L=4096$.}
 		\label{fig:plotphase}
 \end{figure}
-
-% \begin{figure*}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure*}
-
-%QPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-%		\caption{Amplitude error for QPSK}
-%		\label{fig:plotamp}
-%\end{figure}
 
@@
 		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error for QPSK}
+		\caption{Phase error versus SNR for QPSK with $L=4096$.}
 		\label{fig:plotphase}
 \end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM4-3.mps}
-% 		\caption{complex amplitude error for QPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%8PSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-%		\caption{Amplitude error for 8PSK}
-%		\label{fig:plotamp}
-%\end{figure}
 
@@
 		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error for 8PSK}
+		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
 		\label{fig:plotphase}
@@
 
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
+\begin{figure}[tp]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
+		\caption{Unbiased amplitude error versus SNR for BPSK with $L=4096$.}
+		\label{fig:plotphase}
+\end{figure}
+
+\begin{figure}[tp]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+		\caption{Unbiased amplitude error versus SNR for QPSK with $L=4096$.}
+		\label{fig:plotphase}
+\end{figure}
+
+\begin{figure}[tp]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+		\caption{Unbiased amplitude error versus SNR for $8$-PSK with $L=4096$.}
+		\label{fig:plotphase}
+\end{figure}
 

### 2012-08-20 other/confnoncoherent/paper.tex
modified some notation.  It's important that I put the variances etc in the right place.
Added to paper.  Minor code edits.
Added to paper.  Starting the empirical process proofs.  It's been moved to the appendix.
Added more simulations and conclusion.  I think I just need to fill the last Lemma and the two Propositions now.

@@
 \small $^{2}$School of Information Technology \& Electrical Engineering, The University of Queensland, QLD 4067, \textsc{Australia} \vspace{0.1cm} \\
-\small $^{3}$Department of Statistics, Macquarie University, Sydney, \textsc{Australia} \\}
+\small $^{3}$Department of Statistics, Macquarie University, Sydney, \textsc{Australia} 
+\thanks{Supported under the Australian Government’s Australian Space Research Program.}
+}
 
+\IEEEoverridecommandlockouts
 \begin{document}
@@
 Here we are interested in \emph{noncoherent detection}, where $s_1, \dots, s_L$ are not known at the receiver, and must also be estimated.
-This estimation problem has undergone extensive prior study and is often called \emph{multiple symbol differential detection}~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
+This estimation problem has undergone extensive prior study and is often called \emph{multiple symbol differential detection}~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
 A practical approach is the least squares estimator,
@@
 H^{-2} A & 0 \\
-0 & B
+0 & \rho_0^2 B
 \end{array} \right)
@@
 \[
-A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1}. 
+A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1} - G^2(0). 
 \]
@@
 The estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
-We shows that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+We show that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
 However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.

### 2012-08-20 paper1.tex
Added more simulations and conclusion.  I think I just need to fill the last Lemma and the two Propositions now.

@@
 \title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
-\author{Robby McKilliam, Andre Pollok, Bill Cowley  
+\author{Robby McKilliam, Andre Pollok, Bill Cowley, I. Vaughan L. Clarkson and Barry Quinn  
 \thanks{
@@
 Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the reciever and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
+Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
 A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
-\begin{align*}
+\begin{equation}\label{eq:SSdefn}
+\begin{split}
 SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2  \\
 &= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2,
-\end{align*}
+\end{split}
+\end{equation}
 where $\abs{x}$ denotes the magnitude of the complex number $x$.
@@
 
-Let the noise sequence $\{w_i\}$ is complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
+Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
 \[
@@
 
-\begin{itemize}
-\item Run simulations looking at what happens as the number of pilots changes.
-\item Simulations showing what happens as $L$ changes for fixed number of pilots.
-\item Show that sometimes just using pilots is better!
-\end{itemize}
+We present the results of Monte-Carlo simulations with the least squares estimator.
+In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+Under these conditions the least squares estimator is also the maximum likelihood estimator.
+Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
+The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
+For each value of SNR, $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
+
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.
+When $\abs{P} \neq 0$ (i.e.
+coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
+Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
+The solid lines are the MSEs predicted by Theorem~\ref{thm:normality}.
+The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
+As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+Theorem~\ref{thm:normality} does not model this behaviour.
+
+Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
+The dots and crosses show the results of Monte Carlo simulations.
+Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+This requires $G(0)$ to be known.
+In practice $G(0)$ not be known at the reciever, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the estimator.
+When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
+However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+
+Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
+The figure depicts an interesting phenomenon.
+When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.
+When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator combining the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
+A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
+%When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
+This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
+For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
+\[
+SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
+\]
+where the weight $\beta$ would be small when SNR is small and $\beta$ would be near $1$ when SNR is large.
+For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
+One caveat is that the receiver would require knowledge about the noise distribution in order to correctly choose $\beta$.
+We do not invesitigate this further here.
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
+		\caption{Phase error versus SNR for BPSK with $L=4096$.}
+		\label{fig:plotphaseBPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
+		\caption{Phase error versus SNR for QPSK with $L=4096$.}
+		\label{fig:plotphaseQPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
+		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
+		\label{fig:plotphase8PSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
+		\caption{Unbiased amplitude error versus SNR for BPSK.}
+		\label{fig:plotampBPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+		\caption{Unbiased amplitude error versus SNR for QPSK.}
+		\label{fig:plotampQPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+		\caption{Unbiased amplitude error versus SNR for $8$-PSK.}
+		\label{fig:plotamp8PSK}
+\end{figure}
+
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{code/data/plotM2-3.mps}
+%		\caption{Phase error versus SNR for BPSK.}
+%		\label{fig:plotphase}
+%\end{figure}
 
@@
 	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error versus SNR for BPSK with $L=4096$.}
-		\label{fig:plotphase}
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error versus SNR for QPSK with $L=4096$.}
-		\label{fig:plotphase}
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
-		\label{fig:plotphase}
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-		\caption{Unbiased amplitude error versus SNR for BPSK with $L=4096$.}
-		\label{fig:plotphase}
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-		\caption{Unbiased amplitude error versus SNR for QPSK with $L=4096$.}
-		\label{fig:plotphase}
-\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-		\caption{Unbiased amplitude error versus SNR for $8$-PSK with $L=4096$.}
-		\label{fig:plotphase}
+		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
+		\caption{Phase error versus SNR for QPSK.}
+		\label{fig:plotphaseQPSKmultL}
 \end{figure}
@@
 \section{Conclusion}
+
+We considered least squares estimators of carrier phase and amplitude from noisy communications signals that contain both pilot signals, known to the receiver, and data signals, unknown to the receiver.
+We focussed on $M$-ary phase shift keying constellations.
+The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
+
+We showed, under some conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
+This bias is large when the signal to noise ratio is small.
+It would be interesting to investigate methods for correcting this bias.
+A method for estimating $G(0)$ at the reciever appears to be required.
+
+Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
+Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
+This suggests the use of a weigthed objective function.
+In order to select the weight it appears that the reciever would need some knowledge about the noise distribution.
+It would be interesting to investigate methods for selecting, or estimating, the weight at the reciever.
 

### 2012-08-20 paper1.tex
Added some edits

@@
 
-In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
+In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.
 The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be square error between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\abs{a_0 - \hat{a}}^2$.
+It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
 
 The paper is organised in the following way.
-Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when pilot symbols and data symbols exist.
+Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when both pilot symbols and data symbols exist.
 Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
@@
 We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
-However, $\hat{\rho_0}$ is not a consistent estimator of the amplitude $\rho_0$.
+However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
 The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
@@
 In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
-In this case, our expressions for the asymptotic variance can be simplified.
+In this case, our expressions for the asymptotic distribution can be simplified.
 Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.
@@
 \]
-If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
-\[
-f_{X,Y}(x,y) = \frac{f_Z(z)}{2\pi z} \qquad \text{where $z = \sqrt{x^2 + y^2}$.}
-\]
 Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
@@
  \end{align*}
+If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
+\[
+f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
+\]
 
@@
 A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
 Although we will not prove it here, this assumption is not only sufficient, but also nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
@@
 \begin{proposition}\label{prop:contgg2}
-Let $W = X + jY$ be a circularly symmetric complex random variable with pdf $f_{X,Y}(x,y)$ of the real and imaginary parts nonincreasing with $x^2 + y^2$.
+Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.

### 2012-08-20 paper2.tex
started filling in the simulation section.
Added some edits

@@
 \end{align*}
-and the last line converges amost surely to zero by Lemma~\ref{lem:uniflawGGL}.
+and the last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}.
 \end{IEEEproof}
@@
 \[
-\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{2\pi}{N}\expect R_i \right\} = 0.
+\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} = 0.
 \]
@@
 so that
-\[
-V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \leq C_L + \expect C_L,
-\]
-because 
-\[
- \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect \abs{G_L(\lambda) - G_L(\lambda_n)} \leq \expect C_L
-\]
-follows by moving expectation outside the supremum.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
+\begin{align*}
+V_L &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
+&\leq  C_L +  \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect\abs{G_L(\lambda) - G_L(\lambda_n) } \\
+&\leq C_L + \expect C_L,
+\end{align*}
+where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
 \[
@@
 \]
-So,
+Thus,
 \begin{align*}
@@
 \[
-\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i = \tfrac{2\pi}{N}\expect R_1
+\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i = \frac{2\pi}{N}\expect R_1
 \]

### 2012-08-20 paper1.tex
minor edits

@@
 \]
-It follows that
-\[
-%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
-\]
+It follows that $\sqrt{L} m_L =  \rho_0 X_L + o_P(1)$.
 Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
@@
 The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
-The dots and crosses show the results of Monte Carlo simulations.
+The dots and crosses show the results of Monte-Carlo simulations.
 Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
 This requires $G(0)$ to be known.
-In practice $G(0)$ not be known at the reciever, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the estimator.
+In practice $G(0)$ may not be known at the reciever, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
 When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
@@
 When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.
-When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator combining the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
+When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
 A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
@@
 \]
-where the weight $\beta$ would be small when SNR is small and $\beta$ would be near $1$ when SNR is large.
+where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
 For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
-One caveat is that the receiver would require knowledge about the noise distribution in order to correctly choose $\beta$.
+This would enable a rigorous theory for selection of $\beta$ at the reciever.
+One caveat is that the receiver would require knowledge about the noise distribution in order to advantageously choose $\beta$.
 We do not invesitigate this further here.
@@
 Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
-This suggests the use of a weigthed objective function.
-In order to select the weight it appears that the reciever would need some knowledge about the noise distribution.
-It would be interesting to investigate methods for selecting, or estimating, the weight at the reciever.
+This suggests the use of a weigthed objective function, which would be an interesting topic for future research.
 

### 2012-08-21 paper1.tex
more edits

@@
 \begin{theorem}\label{thm:normality} (Asymptotic normality)
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
+Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, let
 \[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
+g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr,
 \]
+and assume that $\frac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\frac{\abs{D}}{L} = d + o(L^{-1/2})$.
 Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
@@
 \end{IEEEproof}
+
+Theorem~\ref{thm:consistency} defines real numbers $p$ and $d$ to represent the proportion of pilot symbols and data symbols in the limit as $L$ goes to infinity.
+For Theorem~\ref{thm:normality} we need the slightly stronger condition that
+\[
+\frac{\abs{P}}{L} = p + o(L^{-1/2}) \qquad \text{and} \qquad \frac{\abs{D}}{L} = d + o(L^{-1/2})
+\] 
+where $o(L^{-1/2})$ denotes a number such that $\sqrt{L} o(L^{-1/2})$ goes to zero as $L\rightarrow \infty$.
+This stronger condition is required to prove the asymptotic normality of $\sqrt{L}\hat{m}_L$.
 

### 2012-08-21 paper2.tex
minor edits
more edits

@@
 %set sections and equation counters based on the previous document.
-\setcounter{equation}{20}
+\setcounter{equation}{21}
 \newcounter{lasteqplusone}
@@
 \[
-G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0,
+G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
 \]
@@
 \end{equation}
-for some $k \in D$.
-Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.
-We will show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-First observe that,
-\begin{align*}
-Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*.
+for some $k \in D$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.  We will show that 
+\[
+SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
+\]
+violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  First observe that,
+\begin{align*}
+Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*,
 \end{align*}
@@
 &= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
-&= SS(\{\hat{d}_i, i \in D\}) - D
+&= SS(\{\hat{d}_i, i \in D\}) - D,
 \end{align*}
@@
 &= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
-&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) }
+&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) },
 \end{align*}
@@
 \end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ because $w_1$ has zero mean.
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ as a result of~\eqref{eq:expectRepartRphi}.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
@@
 &= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
-&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi. \label{eq:Winfsum}
+&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi, \label{eq:Winfsum}
 \end{align}
+where $g$ is defined in the statement of Theorem~\ref{thm:normality}
 Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
@@
 \end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont} in~\cite{McKilliam_leastsqPSKpilotsdata_2012}, it follows $c_k(\lambda)$ converges to
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont} in~\cite{McKilliam_leastsqPSKpilotsdata_2012}, it follows that $c_k(\lambda)$ converges to
 \[
@@
 \begin{IEEEproof}
-Observe that 
-\[
-\sqrt{L} R_L(0) = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) + \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
-\]
-From the standard central limit the distribution of
-\[
-\frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i) = \sqrt{\frac{\sabs{P}}{L}} \frac{1}{\sqrt{\sabs{P}}} \sum_{i \in P} R_i \sin(\Phi_i) 
-\]
-converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
+Observe that $\sqrt{L} R_L(0) = C_L + D_L$ where
+\[
+C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i), \qquad D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
+\]
+From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
 \[
@@
 \]
-Similarly, the distribution of 
-\[ 
-\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} = \sqrt{\frac{\sabs{D}}{L}} \frac{1}{\sqrt{\sabs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
-\]
-converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~\ref{lem:expectImRfracpart}, and variance
+Similarly, the distribution of $D_L$ converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~\ref{lem:expectImRfracpart}, and variance
 \[
@@
 \]
+The lemma holds because $C_L$ and $D_L$ are independent. 
 \end{IEEEproof}
@@
 \begin{IEEEproof}
-Recalling the $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
+Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
 \begin{align*}
@@
 \end{align*}
-where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ are periodic with period $2\pi$.
+where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
 Propostion~\ref{prop:gg2cont} in \cite{McKilliam_leastsqPSKpilotsdata_2012} shows that $g$ is even on $(-\pi,\pi)$.
@@
 \end{align*}
-Observe that the intervals
+Observe that the sets
 \[
@@
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\lambda) + X_L
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\hat{\lambda}_L) + X_L
 \]
@@
 %& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
-&=  p \expect R_1\cos(\lambda + \Phi_1) + d\expect R_1\cos\fracpart{\lambda + \Phi_1} + o(1),
-\end{align*}
-because $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L\rightarrow\infty$. Put
-\begin{align*}
-L(\lambda) &= \sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
+&=  \big( p + o(L^{-1/2}) \big) \expect R_1\cos(\lambda + \Phi_1) \\
+&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1},
+\end{align*}
+because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ is an assumption in the statement of Theorem~\ref{thm:normality}.  Now,
+\begin{align*}
+\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
 \end{align*}
@@
 &= -\lambda \big( \expect R_1\sin(\Phi_1) +  \zeta(\lambda) \expect R_1 \big) \\
-&= \lambda \zeta(\lambda) K,
-\end{align*}
-since $\expect R_1\sin(\Phi_1) = 0$ because $w_1$ has zero mean,  and $K = \expect R_1$ is a finite constant because $\expect\sabs{w_1}$ is finite by assumption in Theorem~\ref{thm:consistency}.  So, 
-\begin{align*}
-\sqrt{L} q_1(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L \zeta(\hat{\lambda}_L) K = o_P(1)
+&= -\lambda \zeta(\lambda) \expect R_1,
+\end{align*}
+since $\expect R_1\sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}.  Now, 
+\begin{align*}
+\sqrt{L} q_1(\hat{\lambda}_L) = -\sqrt{L} \hat{\lambda}_L \zeta(\hat{\lambda}_L) \expect R_1 = o_P(1)
 \end{align*}
@@
 \[
-L(\lambda) = p q_1(\lambda) + d q_2(\lambda) + o(1) = o_P(1)
-\]
-as reuqired.
+\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1) = o_P(1)
+\]
+as required.
 \end{IEEEproof}
@@
 \[
-D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ).
-\]
-From the standard central limit the distribution of $C_L$ converges to the normal with zero mean and variance
+D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ),
+\]
+where $h_1$ and $h_2$ are defined in the statement of Theorem~\ref{thm:consistency}.
+From the standard central limit theorem the distribution of $C_L$ converges to the normal with zero mean and variance
 \[
@@
 \]
-since $h_1(0) = 1$.  The distribution of $D_L$ converges to the normal with zero mean and variance
-\[
-d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0)
-\]
-The proof follows since $C_L$ and $D_L$ are indpendent.
+since $h_1(0) = 1$.  Similarly the distribution of $D_L$ converges to the normal with zero mean and variance
+\[
+d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0).
+\]
+The lemma follows since $C_L$ and $D_L$ are indpendent.
 \end{IEEEproof}
@@
 These results are related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
-The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989}.
+The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989,Gine_Zinn_symmetrisation_1984,Ossiander_clt_bracketing_1984}.
 We will only give a proof of the first statement regarding the function $W_L$.
@@
 \begin{IEEEproof}
-Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$.
+Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be a sequence identically distributed to $\{\Phi_i\}$.
 Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
@@
 \]
-is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont} in~\cite{McKilliam_leastsqPSKpilotsdata_2012}.
-From the definition of $I_\epsilon$,
+is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}.  From the definition of $I_\epsilon$,
 \begin{align*}
@@
 \[
-K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}
+K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
 \]
@@
 \[
-c_3 \leq 2 \sum_{i \in D, i \in S} R_i \leq 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i).
+c_3 \leq 2 \sum_{i \in D, i \in S} R_i \leq 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i).
 \]
@@
 &\leq c_1 + c_3 + c_4 \\
-&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in D} R_i I_\epsilon(\Phi_i) \\
+&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i) \\
 &= D_\epsilon L

### 2012-08-21 paper2.tex
more edits.  Filling in the covering Lemma now.

@@
  \[
- c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) }
- \]
- and 
+ c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) },
+ \] 
  \[
@@
  \]
- for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let $S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}$. Now $c_2 = c_3 + c_4$ where 
- \[
- c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}
- \]
- and
+ for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let 
+\[
+S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}.
+\] 
+Now $c_2 = c_3 + c_4$ where 
+ \[
+ c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}},
+ \]
  \[
@@
  \]
- If $i \notin S$ then $\round{\lambda + \Phi_i} = \round{\Phi_i}$ for all $\sabs{\lambda} < \epsilon$ and therefore
+Observe that $I_\epsilon(\Phi_i) = 1$ if and only if $\abs{\fracpart{\Phi_i}} \geq \tfrac{\pi}{M} - \epsilon$.  So, if $i \notin S$ and $\sabs{\lambda} < \epsilon$, then
+\begin{align*}
+\round{\lambda + \Phi_i} &= \round{\lambda + \fracpart{\Phi_i} + \round{\Phi_i}} \\ 
+&= \round{\lambda + \fracpart{\Phi_i}} + \round{\Phi_i} = \round{\Phi_i}
+\end{align*}
+because
+\[
+\sabs{\lambda + \fracpart{\Phi_i}} < \epsilon + \tfrac{\pi}{M} - \epsilon = \tfrac{\pi}{M}
+\] 
+and therefore $\round{\lambda + \fracpart{\Phi_i}} = 0$.  Now,
  \begin{align*}
@@
 \begin{align*}
-d_2(\lambda, 0) &= c_1 + c_2 \\
-&\leq c_1 + c_3 + c_4 \\
-&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i) \\
-&= D_\epsilon L
+d_2(\lambda, 0) &= c_1 + c_2 = c_1 + c_3 + c_4 \\
+&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i) = D_\epsilon L
 \end{align*}
@@
 \begin{lemma}\label{lem:maxineq}(Maximal inequalities)
-Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$.  Then
+Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
 \[
@@
 \begin{IEEEproof}
+We will first construct a set $T^2_\epsilon(k)$ containing no more than $K_{23}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_2 \in T^2_\epsilon(k)$ such that $d_2(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{23}$ a constant, independent of $L$, $\epsilon$ and $k$.
+Using a similar construction we find a set $T^1_\epsilon(k)$ containing no more than $K_{13}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_1 \in T^1_\epsilon(k)$ such that $d_1(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{13}$ a constant, independent of $L$, $\epsilon$ and $k$.
+Puting $T_\epsilon(k) = T^1_\epsilon(k) \cup T^2_\epsilon(k)$ will then complete the proof with $K_3 = K_{13} + K_{23}$.
+
+We will only give a construction for $T^2_\epsilon(k)$, the construction for $T^1_\epsilon(k)$ being similar.  First observe that
+\[
+f_{i}(\lambda) - f_{i}(\lambda_2) = \begin{cases}
+R_i\big( \sin(\lambda + \Phi_i) - \sin(\lambda_2 + \Phi_i) \big) & i \in P \\
+R_i\big(\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda_2 + \Phi_i}\big) & i \in D
+\end{cases}
+\]
+Let 
+\[
+U = \{\epsilon(\ell 2^{-k} - 1) \mid \ell = 0, 1, \dots, 2^{k+1} \} 
+\]
+be a set containing $2^{k+1} + 1$ elements uniformly spaced over $B_\epsilon$.
+For every $\lambda \in B_\epsilon$ there exists a $\lambda^* \in U$ such that $\abs{\lambda - \lambda^*} < \epsilon 2^{-k}$.
+Since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$, it follows that if $i \in P$ then,
+Let 
+\begin{align*}
+\abs{f_{i}(\lambda) - f_{i}(\lambda_2)} &= R_i\abs{ \sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i) } \\
+&\leq R_i\sabs{\lambda - \lambda^*} \\
+&< \epsilon 2^{-k} R_i.
+\end{align*}
+Thus
+\[
+\sum_{i \in P} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i\in P} R_i.
+\]
+As in Lemma~\ref{lem:epslmlemma} let $S = \{ i \in D \mid I_\epsilon(\Phi_i) \}$. Now, if $i \in D$ but $i \notin S$ then 
+\[
+\round{\lambda + \Phi_i} = \round{\lambda^* + \Phi_i} = \round{\Phi_i},
+\] 
+so
+ \begin{align*}
+ \vert &f_{i}(\lambda) - f_{i}(\lambda_2)\vert \\
+&= R_i\vert\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}\vert \\
+ &= R_i\vert\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\lambda^* + \Phi_i - \round{\Phi_i})\vert  \\
+ &\leq R_i \sabs{\lambda - \lambda^*} \\
+&< \epsilon 2^{-k} R_i.
+ \end{align*}
+Thus,
+\[
+\sum_{i \in D, i \notin S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i \in D, i \notin S} R_i \leq \epsilon 2^{-k} \sum_{i \in D} R_i
+\]
+
+To handle the case when $i \in S$ we need to add some extra elements to the set $U$.  Let $\sigma$ be a permutation of the indices in $S$ such that $\fracpart{\Phi_i}$ are in ascending order.
 

### 2012-08-22 paper2.tex
minor edits to second paper.

--- 2012-08-21 paper2.tex
@@
 \begin{IEEEproof}
+% We will first construct a set $T^2_\epsilon(k)$ containing no more than $K_{23}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_2 \in T^2_\epsilon(k)$ such that $d_2(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{23}$ a constant, independent of $L$, $\epsilon$ and $k$.
+Using a similar construction we find a set $T^1_\epsilon(k)$ containing no more than $K_{13}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_1 \in T^1_\epsilon(k)$ such that $d_1(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{13}$ a constant, independent of $L$, $\epsilon$ and $k$.
+Puting $T_\epsilon(k) = T^1_\epsilon(k) \cup T^2_\epsilon(k)$ will then complete the proof with $K_3 = K_{13} + K_{23}$.
+
+% We will only give a construction for $T^2_\epsilon(k)$, the construction for $T^1_\epsilon(k)$ being similar.  First observe that
+% \[
+% f_{i}(\lambda) - f_{i}(\lambda_2) = \begin{cases}
+% R_i\big( \sin(\lambda + \Phi_i) - \sin(\lambda_2 + \Phi_i) \big) & i \in P \\
+% R_i\big(\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda_2 + \Phi_i}\big) & i \in D
+% \end{cases}
+% \]
+% Let 
+% \[
+% U = \{\epsilon(\ell 2^{-k} - 1) \mid \ell = 0, 1, \dots, 2^{k+1} \} 
+% \]
+% be a set containing $2^{k+1} + 1$ elements uniformly spaced over $B_\epsilon$.
+For every $\lambda \in B_\epsilon$ there exists a $\lambda^* \in U$ such that $\abs{\lambda - \lambda^*} < \epsilon 2^{-k}$.
+Since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$, it follows that if $i \in P$ then,
+% Let 
+% \begin{align*}
+% \abs{f_{i}(\lambda) - f_{i}(\lambda_2)} &= R_i\abs{ \sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i) } \\
+% &\leq R_i\sabs{\lambda - \lambda^*} \\
+% &< \epsilon 2^{-k} R_i.
+% \end{align*}
+% Thus
+% \[
+% \sum_{i \in P} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i\in P} R_i.
+% \]
+% As in Lemma~\ref{lem:epslmlemma} let $S = \{ i \in D \mid I_\epsilon(\Phi_i) \}$. Now, if $i \in D$ but $i \notin S$ then 
+% \[
+% \round{\lambda + \Phi_i} = \round{\lambda^* + \Phi_i} = \round{\Phi_i},
+% \] 
+% so
+%  \begin{align*}
+%  \vert f_{i}(\lambda) - &f_{i}(\lambda^*)\vert \\
+% &= R_i\vert\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}\vert \\
+%  &= R_i\vert\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\lambda^* + \Phi_i - \round{\Phi_i})\vert  \\
+%  &\leq R_i \sabs{\lambda - \lambda^*} \\
+% &< \epsilon 2^{-k} R_i.
+%  \end{align*}
+% Thus,
+% \[
+% \sum_{i \in D, i \notin S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i \in D, i \notin S} R_i 
+% \]
+
+% To handle the case when $i \in S$ we need to add some extra elements to the set $U$.  Let 
+% \[
+% r_i = \begin{cases}
+% \tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} > 0 \\
+% -\tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} \leq 0
+% \end{cases}
+% \]
+% and let $\sigma$ be a permutation of the indices in $S$ such that $r_i$ are in ascending order, that is,
+% \begin{equation}\label{eq:sigmasortind}
+% r_{\sigma(i)} \leq r_{\sigma(k)}
+% \end{equation}
+% whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{S}-1\}$.  Now,
+% \begin{align*}
+% \sum_{i\in S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} = \sum_{i = 0}^{\abs{S}-1} \sabs{ f_{\sigma(i)}(\lambda) - f_{\sigma(i)}(\lambda^*)}.
+% \end{align*}
+% If $\epsilon > \lambda \geq r_{\sigma(k)} > 0$ for some $k$, then
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = \tfrac{\pi}{M}
+% \]
+% for all $i \leq k$, whilst
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
+% \]
+% for all $i > k$.  Otherwise if $-\epsilon < \lambda < r_{\sigma(k)} \leq 0$ for some $k$, then
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + -\tfrac{\pi}{M} - r_{\sigma(i)}} = -\tfrac{\pi}{M}
+% \]
+% for all $i \leq k$, whilst
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
+% \]
+% for all $i > k$.
+
+% Let $R = \sum_{i \in S} R_i$ and let $d(\ell)$ be the largest integer such that
+% \[
+% \sum_{i=0}^{d(\ell)} R_{\sigma(i)} \leq \frac{\ell}{2^k} R.
+% \]
+% for integers $\ell = 1, \dots, 2^k$.  Define the intervals
+% \[
+% L_\ell = \in [r_{d(\ell)}, r_{d(\ell+1)})
+% \]
+% where $r_{0} = -\epsilon$ and $r_{2^k+1} = \epsilon$.
+Then $L_0, L_1, \dots, L_{2^k}$ partition the closure of $B_\epsilon$.
+If $\lambda$ and $\lambda^*$ are from interval $L_\ell$, then for all $i \in \{0, 1, \dots, \abs{S}-1\}$ less than $d(\ell)$,
+% \begin{align*}
+% \fracpart{\lambda + \Phi_{\sigma(i)}} &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}} \\
+% &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}}
+% \end{align*}
+% \begin{align*}
+% \vert f_{\sigma(i)}(\lambda) - &f_{\sigma(i)}(\lambda^*) \vert \\
+% &= R_{\sigma(i)}\vert\sin\fracpart{\lambda + \Phi_{\sigma(i)}} - \sin\fracpart{\lambda^* + \Phi_{\sigma(i)}}\vert 
+% \end{align*}
+BLERG
 

### 2012-08-28 algonly.tex
minor additions

--- 2012-08-20 paper1.tex
@@
 \title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
-\author{Robby McKilliam, Andre Pollok, Bill Cowley, I. Vaughan L. Clarkson and Barry Quinn  
+\author{Robby McKilliam, Andre Pollok, Bill Cowley
 \thanks{
@@
 
-We assume that time offset estimation and matched filtering have been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
+We assume that time offset estimation has been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
 \begin{equation}\label{eq:sigmod}
@@
 
-The existing literature mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
+The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
 In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
 A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-The idea is the `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
-A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
+The idea is to `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
+A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
 \[
@@
 Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included, so as to make the estimator coherent.
+A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included.
 Techniques are available but they are somewhat adhoc~\cite{Cowley_ref_sym_carr_1998}.
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
+Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-
-In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.
-The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
-Whilst estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite approach here.
-Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
-This is motivated by the fact that in many modern communication systems the data symbols are \emph{coded}.
-For this reason raw symbol error rate is not of interest at this stage.
-Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
-The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be square error between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\abs{a_0 - \hat{a}}^2$.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.
-The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
-
-The paper is organised in the following way.
-Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when both pilot symbols and data symbols exist.
-Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
-Section~\ref{sec:stat-prop-least} states two theorms that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
-However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
-The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
-Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.
-Due to space constraints, the lemmas needed by these proofs have been placed in the supplementary paper~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
-In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
-In this case, our expressions for the asymptotic distribution can be simplified.
-Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.
-These simulations agree with the derived asymptotic properties.
 
@@
 In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to include the pilot symbols.
+Mackenthun specifically considered the noncoherent setting, so we modify the algorithm to include the pilot symbols.
 For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 So, this na\"{\i}ve approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
-Mackenthun~\cite{Mackenthun1994} showed how $SS(f_k)$ can be computed recursively.
+Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.
 
@@
 
-\section{Circularly symmetric complex random variables}\label{sec:circ-symm-compl}
-
-Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
-A complex random variable $W$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{W}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{W}$.
-That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
-If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
-\[
-f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
-\]
-Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
-If $\expect\abs{W} = \expect Z$ is finite, then $W$ has zero mean because
-\begin{align*}
- \expect W &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
- &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
- &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
- \end{align*}
-If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
-\[
-f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
-\]
-
-We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
-\[
-R e^{j\Phi} = 1 + W.
-\]
-The joint pdf of $R$ and $\Phi$ can be shown to be
-\begin{equation}\label{eq:pdfRPhi}
-f(r,\phi) = \frac{r f_Z(\sqrt{\abs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\abs{r^2 - 2r\cos\phi + 1}}}.
-\end{equation}
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
-\begin{equation}\label{eq:expectRepartRphi}
-\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
-\end{equation}
-where $\Re(\cdot)$ denotes the real part, and
-\begin{equation}\label{eq:expectImpartRphi}
-\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
-\end{equation}
-where $\Im(\cdot)$ denotes the imaginary part.
-
-% \begin{lemma}\label{lem:h1minedcircsym}
-% Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
-Let
-% \[
-% h_1(x) = \expect R \cos(x + \Phi).
-% \]
-% Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
-% \end{lemma}
-% % Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
-That is, the pdf of $Z e^{j \Theta}$ in rectangular coordinates is given by $z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$ and $x$ and $y$ are the real and imaginary parts of $Z e^{j \Theta}$, and this pdf is non increasing with $z$.
-For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
-
-% % By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if
-% % \[
-% % \abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
-% % \]
-% % We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
-% \begin{IEEEproof}
-% BLERG
-% \end{IEEEproof}
-
-\section{Statistical properties of the least squares estimator}\label{sec:stat-prop-least}
-
-In this section we decribe the asymptotic properties of the least squares estimator.
-In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
-\[
-\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right),
-\]
-where $\operatorname{round}(\cdot)$ takes its argument to the nearest integer.
-The direction of rounding for half-integers is not important so long as it is consistent.
-We have chosen to round up half-integers here.
-Similarly we use $\sfracpart{x}$ to denote $x$ taken `modulo $\tfrac{2\pi}{M}$' into the interval $\left[-\tfrac{\pi}{M}, \tfrac{\pi}{M}\right)$, that is
-\[
-\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right) = x - \round{x}.
-\]
-The next two theorems decribe the asymptotic properties of the least squares estimator.
-
-\begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $\{y_i, i \in P \cup D\}$ be given by~\eqref{eq:sigmod}.
-Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
-%where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
-Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
-\[
-\frac{\abs{P}}{L} \rightarrow p \qquad \text{and} \qquad \frac{\abs{D}}{L} \rightarrow d \qquad \text{as $L \rightarrow \infty$.}
-\] 
-Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
-\begin{equation}\label{eq:RiandPhii}
-R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
-\end{equation}
-and define the continuous function
-\[
-G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
-\]
-\[
-h_1(x) = \expect R_1 \cos(x + \Phi_1), \;\;\; h_2(x) =  \expect R_1 \cos\sfracpart{ x + \Phi_1}.
-\]
-If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
-\begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
-\end{enumerate}
-\end{theorem}
-
-\begin{theorem}\label{thm:normality} (Asymptotic normality)
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
-\[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
-\]
-Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-%If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
-\[
-\left( \begin{array}{cc} 
-\frac{pA_1 + dA_2}{(p + H d)^2} & 0 \\
-0 & \rho_0^2(pB_1 + dB_2)
-\end{array} \right)
-\]
-as $L \rightarrow \infty$, where
-\[
-H = h_2(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
-\]
-\[
-A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
-\]
-\[
-B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
-\]
-\end{theorem}
-
-% \begin{corollary}\label{cor:ampmse}
-% The mean square error of the amplitude estimator $\hat{\rho}$ is 
-% \[
-% \expect (\rho_0 - \hat{\rho})^2 = \rho_0^2\frac{p B_1 + d B_2}{L} + \rho_0^2(1 - G(0))^2 + o(L^{-1/2}).
-% \] 
-% \end{corollary}
-% \begin{IEEEproof} Write
-% \begin{align*}
-% \expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
-% &= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
-% &= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L.
-% \end{align*} 
-% From Theorem~\ref{thm:normality} it follows that $\expect \hat{m}_L^2 = \tfrac{B}{L} + o(L^{-1})$ and that $\expect \hat{m}_L = o(L^{-1/2})$. 
-% \end{IEEEproof}
-
-The proof of Theorem~\ref{thm:consistency} is in Section~\ref{sec:proof-almost-sure} and the proof of Theorem~\ref{thm:normality} is in Section~\ref{sec:proof-asympt-norm}.
-Before giving the proofs we discuss the assumptions made by these theorems.
-The assumption that $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
-If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $\{s_i,i \in P \cup D\}$.
-In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $\{s_i,i \in P \cup D\}$, rather than just $\rho_0$.
-The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
-
-The assumption that $\expect \sabs{w_1}^2 = \expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.
-This is required in Theorem~\ref{thm:normality} so that the constants $A_1$, $A_2$, $B_1$ and $B_2$ exist.
-We will also use that $R_i$ has finite variance to simplify the proof of Theorem~\ref{thm:consistency} by use of Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}
-
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta} - \theta_0$.
-This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
-So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
-
-For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.
-The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.
-At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.
-
-\begin{proposition}\label{prop:gg2cont}
-Let $W$ be a circularly symmetric complex random variable with $\expect\abs{W}^2$ finite.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
-Let $f(r,\phi)$ be the joint pdf of $R$ and $\Phi$.
-Then
-\[
-g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
-\]
-and
-\[
-g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
-\]
-are even functions on $(-\pi,\pi)$ and are continuous at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$.
-\end{proposition}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
-A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
-The zero mean complex Gaussian distribution with independent real and imaginary parts is apart of this class.
-
-\begin{proposition}\label{prop:contgg2}
-Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
-Let,
-\[
-h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
-\]
-Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-\end{proposition}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
-% \begin{IEEEproof}
-% We only give a proof for $g_2$, the proof for $g$ begin similar.  From~\eqref{eq:pdfRPhi},
-% \[
-% g_2(\phi) = \int_{0}^\infty \frac{r^3 f_Z(\sqrt{\sabs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\sabs{r^2 - 2r\cos\phi + 1}}} dr.
-% \]
-% That $g_2$ is even on $[-\pi,\pi]$ is now obvious because $\cos\phi$ is even on $[-\pi,\pi]$.  Put $a = \cos\phi$ and $z^2 = r^2 - 2ra + 1$ and by a change of variable from $r$ to $z$,
-% \[
-% g_2(\phi) = \int_{1}^{\sqrt{1-a^2}} \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz + \int_{\sqrt{1-a^2}}^\infty \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz.
-% \]
-% when $0 < a \leq 1$ and
-% \[
-% g_2(\phi) = \int_{1}^\infty \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz.
-% \]
-% when $-1 \leq a \leq 0$.  Because
-% \[
-% \frac{r^3}{(r - a)} \leq a z^2 +  b
-% \]
-% for all $\phi \in \reals$, it follows that
-% \[
-% g_2(\phi) \leq \int_{0}^\infty \frac{(a z^2 + d) f_Z(z)}{2\pi} dz = a \expect Z^2 + b
-% \]
-% which is finite.  BLERG work out a and b.
-% \end{IEEEproof}
-
-\section{Proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sure}
-
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
- \begin{align*}
-SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
-&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
-\end{align*}
-where
-\[
-Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
-\]
-and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
-\begin{equation}\label{eq:hatrhoZ}
-\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
-\end{equation}
-where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
-\[
-SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
-\]
-
-We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
-Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
-\begin{align*}
-y_i &= a_0 s_i + w_i \\
-&= a_0 s_i \left( 1 + \frac{w_i}{a_0 s_i} \right) \\
-&= a_0 s_i R_i e^{j \Phi_i} \\
-&= \rho_0 R_i e^{j ( \Phi_i + \theta_0 + \angle{s_i}) }.
-\end{align*}
-%The $\Phi_i$ can be identified as \emph{circular~random~variables}~\cite{Mardia_directional_statistics,Fisher1993,McKilliam_mean_dir_est_sq_arc_length2010}.  
-Recalling the definition of $\hat{d}_i(\theta)$ and $\hat{u}_i(\theta)$ from~\eqref{eq:hatdfinxtheta},
-\begin{align*}
-\hat{u}_i(\theta) &= \round{\angle{y_i} - \theta} \\
-&= \round{\theta_0 + \Phi_i + \angle{s_i} - \theta} \\
-&\equiv \round{ \fracpart{\theta_0 - \theta}_{\pi} + \Phi_i + \angle{s_i}} \pmod{2\pi} \\
-&= \round{ \lambda + \Phi_i + \angle{s_i} },
-\end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
-\begin{align}
- y_i e^{-j\theta} \hat{d}_i^*(\theta) &= \rho_0 R_i e^{j(\lambda + \Phi_i + \angle{s_i} - \round{\lambda + \Phi_i + \angle{s_i}})} \nonumber \\
-&= \rho_0 R_i e^{j(\lambda + \Phi_i - \round{\lambda + \Phi_i})} \nonumber  \\
-&= \rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}} \label{eq:yethetadhat}
-\end{align}
-since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
-\[
-y_i e^{-j\theta} p_i^* = \rho_0 R_i e^{j(\lambda + \Phi_i)}.
-\]
-Now,
-\[
-Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
-\]
-and
-\[
-\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
-\] 
-Let 
-\begin{equation}\label{eq:GLdefn}
-G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
-\end{equation}
-and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
-Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
-We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.
-
-Recall the functions $G$, $h_1$ and $h_2$ defined in the statement of Theorem~\ref{thm:consistency}.  Observe that
-\[
-\expect G_L(\lambda) = \frac{\abs{P}}{L} h_1(\lambda) + \frac{\abs{D}}{L} h_2(\lambda)
-\]
-and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
-\[
-\lim_{L \rightarrow\infty} \expect G_L(\lambda) = G(\lambda) = p h_1(\lambda)   +  d h_2(\lambda).
-\]
-As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.
-Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
-Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
-So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
-Part 1 of Theorem~\ref{thm:consistency} follows.
-
-It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
-\begin{equation}\label{eq:rhoGLZ}
-\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
-\end{equation}  
-Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
-It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
-These are proved in Section~\ref{sec:proof-almost-sureappendix} of \cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
-
-\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
-
-We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
-Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
-Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
-The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
-Define the function
-\begin{equation}\label{eq:RLdef}
-R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\end{equation}
-Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
-Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
-\begin{align*}
-0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
-&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
-\end{align*}
-Lemma~\ref{lem:Qconv} shows that
-\[
-\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
-Lemma~\ref{lem:empiricprocc} shows that
-\[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
-\]
-It follows from the three equations above that,
-\[
-0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
-\]
-and rearranging gives,
-\[
-\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
-\]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
-It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
-\[
-\frac{pA_1 + dA_2}{(p + Hd)^2}.
-\]
-
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
-\begin{align*}
-\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
-\end{align*}
-Lemma~\ref{lem:empiricprocforrho} shows that 
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
-\]
-It follows that $\sqrt{L} m_L =  \rho_0 X_L + o_P(1)$.
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
-Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
-Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
-\[
-\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
-\]
-as $L \rightarrow \infty$.
-It remains to prove the lemmas that we have used.
-These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
-
-\section{The Gaussian noise case}\label{sec:gaussian-noise-case}
-
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
-\[
-\frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
-\]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
-%and the joint density function of the real and imaginary parts of $Z$ is
-%\[
-%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
-%\] 
-%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
-It can be shown that
-\[
-g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-\]
-where $\kappa = \rho_0/\sigma$ and
-\[
-\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
-\]
-The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
-
-\section{Simulations}\label{sec:simulations}
-
-We present the results of Monte-Carlo simulations with the least squares estimator.
-In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
-Under these conditions the least squares estimator is also the maximum likelihood estimator.
-Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
-The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
-For each value of SNR, $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
-
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.
-When $\abs{P} \neq 0$ (i.e.
-coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
-Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
-The solid lines are the MSEs predicted by Theorem~\ref{thm:normality}.
-The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
-As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-Theorem~\ref{thm:normality} does not model this behaviour.
-
-Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
-The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
-The dots and crosses show the results of Monte-Carlo simulations.
-Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
-This requires $G(0)$ to be known.
-In practice $G(0)$ may not be known at the reciever, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
-When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
-However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
-
-Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
-The figure depicts an interesting phenomenon.
-When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.
-When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
-A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
-%When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
-This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
-For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
-\[
-SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
-\]
-where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
-For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
-This would enable a rigorous theory for selection of $\beta$ at the reciever.
-One caveat is that the receiver would require knowledge about the noise distribution in order to advantageously choose $\beta$.
-We do not invesitigate this further here.
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error versus SNR for BPSK with $L=4096$.}
-		\label{fig:plotphaseBPSK}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error versus SNR for QPSK with $L=4096$.}
-		\label{fig:plotphaseQPSK}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
-		\label{fig:plotphase8PSK}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-		\caption{Unbiased amplitude error versus SNR for BPSK.}
-		\label{fig:plotampBPSK}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-		\caption{Unbiased amplitude error versus SNR for QPSK.}
-		\label{fig:plotampQPSK}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-		\caption{Unbiased amplitude error versus SNR for $8$-PSK.}
-		\label{fig:plotamp8PSK}
-\end{figure}
-
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM2-3.mps}
-%		\caption{Phase error versus SNR for BPSK.}
-%		\label{fig:plotphase}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
-		\caption{Phase error versus SNR for QPSK.}
-		\label{fig:plotphaseQPSKmultL}
-\end{figure}
-
-\section{Conclusion}
-
-We considered least squares estimators of carrier phase and amplitude from noisy communications signals that contain both pilot signals, known to the receiver, and data signals, unknown to the receiver.
-We focussed on $M$-ary phase shift keying constellations.
-The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
-
-We showed, under some conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
-This bias is large when the signal to noise ratio is small.
-It would be interesting to investigate methods for correcting this bias.
-A method for estimating $G(0)$ at the reciever appears to be required.
-
-Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
-Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
-This suggests the use of a weigthed objective function, which would be an interesting topic for future research.
-
 \small

### 2012-08-28 paper1.tex
minor additions

@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
+Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be square error between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\abs{a_0 - \hat{a}}^2$.
+The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be $\abs{a_0 - \hat{a}}^2$.
 It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.

### 2012-08-28 paper1.tex
fixed Andre's name.

@@
 \title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
-\author{Robby McKilliam, Andre Pollok, Bill Cowley, I. Vaughan L. Clarkson and Barry Quinn  
+\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, I. Vaughan L. Clarkson and Barry Quinn  
 \thanks{
 Supported under the Australian Government’s Australian Space Research Program.
-Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Robby McKilliam, Andr\'{e} Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
 Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.

### 2012-09-01 paper1.tex
minor edits to java options, should make the simulation go a little faster.

--- 2012-08-21 paper1.tex
@@
 Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
-Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the reciever and must also be estimated.
+Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the receiver and must also be estimated.
 Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
@@
 The existing literature mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
-In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
+In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
 A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-The idea is the `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
+The idea is the `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.
 A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
@@
 For this reason raw symbol error rate is not of interest at this stage.
-Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
-The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
+Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated received symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
+The additive noise channel is a common assumption for subsequent receiver operations, such as decoding.
 Consequently, our metric of performance will not be symbol or bit error rate, it will be square error between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\abs{a_0 - \hat{a}}^2$.
@@
 Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
-Section~\ref{sec:stat-prop-least} states two theorms that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
+Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
 We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
@@
 
-\section{Mackentun's algorithm with pilots}\label{sec:least-squar-estim}
+\section{Mackenthun's algorithm with pilots}\label{sec:least-squar-estim}
 
@@
 Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to include the pilot symbols.
-For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
+For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
 In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.
@@
 \end{align}
-where $*$ denotes the complex congugate.
+where $*$ denotes the complex conjugate.
 Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find the least squares estimator of $a_0$ as a function of $\{d_i, i \in D\}$,
@@
 \]  
-Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to surpress the argument $(\{ d_i, i \in D\})$ for notational clarity.
+Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to suppress the argument $(\{ d_i, i \in D\})$ for notational clarity.
 Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain the $SS$ conditioned on minimisation with respect to $a$,
@@
 Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-It turns out that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
+It turns out that there are at most $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
 %When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
@@
 Observe that $f(\theta)$ is piecewise continuous.
-The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains contant are determined by the values of $\{z_i, i \in D\}$.
+The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains constant are determined by the values of $\{z_i, i \in D\}$.
 Let
@@
 If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
+Observe from~\eqref{eq:uicombos} that there are at most $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
 
@@
 \end{equation}
-Psuedocode is given in Algorithm~\ref{alg:loglinear}.
+Pseudocode is given in Algorithm~\ref{alg:loglinear}.
 Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
@@
 A complex random variable $W$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{W}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{W}$.
-That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
+That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
 If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
@@
 
-In this section we decribe the asymptotic properties of the least squares estimator.
+In this section we describe the asymptotic properties of the least squares estimator.
 In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
@@
 \]
-The next two theorems decribe the asymptotic properties of the least squares estimator.
+The next two theorems describe the asymptotic properties of the least squares estimator.
 
@@
 This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
 The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
@@
 \]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+where $o_P(1)$ denotes a number converging in probability to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
 Lemma~\ref{lem:empiricprocc} shows that
@@
 \]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Theorem~\ref{thm:normality}.
 It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
@@
 It follows that $\sqrt{L} m_L =  \rho_0 X_L + o_P(1)$.
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
+Lemma~\ref{lem:XL} shows that the distribution of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
 Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
@@
 \]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
 %and the joint density function of the real and imaginary parts of $Z$ is
@@
 This requires $G(0)$ to be known.
-In practice $G(0)$ may not be known at the reciever, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
+In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
 When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
@@
 For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
-This would enable a rigorous theory for selection of $\beta$ at the reciever.
+This would enable a rigorous theory for selection of $\beta$ at the receiver.
 One caveat is that the receiver would require knowledge about the noise distribution in order to advantageously choose $\beta$.
-We do not invesitigate this further here.
+We do not investigate this further here.
 
@@
 We considered least squares estimators of carrier phase and amplitude from noisy communications signals that contain both pilot signals, known to the receiver, and data signals, unknown to the receiver.
-We focussed on $M$-ary phase shift keying constellations.
+We focused on $M$-ary phase shift keying constellations.
 The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
@@
 It would be interesting to investigate methods for correcting this bias.
-A method for estimating $G(0)$ at the reciever appears to be required.
+A method for estimating $G(0)$ at the receiver appears to be required.
 
@@
 Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
-This suggests the use of a weigthed objective function, which would be an interesting topic for future research.
+This suggests the use of a weighed objective function, which would be an interesting topic for future research.
 

### 2012-09-01 paper2.tex
minor edits to java options, should make the simulation go a little faster.

@@
 \end{align*}
-the last line following from Jensens inequality.  Put
+the last line following from Jensen's inequality.  Put
 \[
@@
 \begin{IEEEproof}
-Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimsers of the function 
+Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimisers of the function 
 \[
@@
 \]
-Subtituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
 Thus,~\eqref{eq:fraclambassumption} is false.
@@
 Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.
-Now an analagous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
+Now an analogous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
 
@@
 \]
-where $p, d$ and $H$ are real numbers defined in the statements of Thereoms~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
+where $p, d$ and $H$ are real numbers defined in the statements of Theorems~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
 \end{lemma}
@@
 We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
-An analagous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
+An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
 To keep our notation clean put
@@
 \]
-Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
+Since the intervals $[\psi_{k-1}, \psi_{k})$ partition the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
 \[
@@
 In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
-An analgous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
+An analogous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
 Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$,
@@
 We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
-The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
+The analogous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
 
@@
 where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
-Propostion~\ref{prop:gg2cont} in \cite{McKilliam_leastsqPSKpilotsdata_2012} shows that $g$ is even on $(-\pi,\pi)$.
+Proposition~\ref{prop:gg2cont} in \cite{McKilliam_leastsqPSKpilotsdata_2012} shows that $g$ is even on $(-\pi,\pi)$.
 Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.
@@
 \]
-The lemma follows since $C_L$ and $D_L$ are indpendent.
+The lemma follows since $C_L$ and $D_L$ are independent.
 \end{IEEEproof}
@@
 Let $B_\epsilon = \{x \in \reals \mid \abs{x} < \epsilon\}$ be the set of real numbers with magnitude less than $\epsilon$.
-For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $B_\epsilon$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the psuedometric
+For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $B_\epsilon$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the pseudometric
 \[
@@
 \]
-and the psuedometric
+and the pseudometric
 \[
@@
 Defined this way, $T_\epsilon(0)$ satisfies the inequality above because $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
-The existance of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{lem:metricentropy}.
+The existence of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{lem:metricentropy}.
 Lemma~\ref{lem:metricentropy} also shows that the number of elements in $T_\epsilon(k)$ is no more that $K_3 2^k$ where $K_3$ is a constant, independent of $L$, $\epsilon$ and $k$.
@@
 Let $b_2(\lambda)$ be a function that maps $\lambda \in B_\epsilon$ to a point in $T_\epsilon(k)$ such that  $d_2(\lambda,b_2(\lambda)) \leq 2^{-k}D_\epsilon L$.
-The existance of $b_2$ is guaranteed by the definition of $T_\epsilon(k-1)$.
+The existence of $b_2$ is guaranteed by the definition of $T_\epsilon(k-1)$.
 So, for all $\lambda \in B_\epsilon$,
@@
 \end{align*}
-The second line in the inequaliy above follows because,
+The second line in the inequality above follows because,
 \[
@@
 Let $b_1(\lambda)$ map each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
-The existance of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
+The existence of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
 By the triangle inequality
@@
 \begin{lemma}\label{lem:metricentropy}(Covering numbers)
-For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset B_\epsilon$ with the property that, for every $\lambda \in B_\epsilon$, there is a $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the psuedometric
+For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset B_\epsilon$ with the property that, for every $\lambda \in B_\epsilon$, there is a $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the pseudometric
 \[

### 2012-09-01 code/runsim.sh

@@ -1,4 +1,5 @@
-#export JAVA_OPTS="-Xprof"
+#export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
+export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
 scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar clt.scala
 scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar montecarlo.scala
 scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar montecarlonc.scala

### 2012-09-04 paper1.tex

@@
 
-The existing literature mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
-In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
+The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
+In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
 A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-The idea is the `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.
-A function $F$, mapping $\reals$ to $\reals$, is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
+The idea is to `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
+A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
 \[
@@
 For this reason raw symbol error rate is not of interest at this stage.
-Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated received symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
-The additive noise channel is a common assumption for subsequent receiver operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be square error between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\abs{a_0 - \hat{a}}^2$.
+Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
+The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
+The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be $\abs{a_0 - \hat{a}}^2$.
 It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.
@@
 In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to include the pilot symbols.
-For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
+Mackenthun specifically considered the noncoherent setting, so we modify the algorithm to include the pilot symbols.
+For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
 In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.

### 2012-09-12 algonly.tex
updated algonly

@@
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-ary phase shift keying.
-We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator is not consistent, it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
-Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.
-The results of Monte Carlo simulations are provided and these agree with the theoretical results.
+We describe an algorithm to compute the least suqares estimator that requires only $O(L \log L)$ arithmetic operations, where $L$ is the number of recieved symbols.
 \end{abstract}
@@
 
+\section{Simulations}\label{sec:simulations}
+
+We present the results of Monte-Carlo simulations with the least squares estimator.
+In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+Under these conditions the least squares estimator is also the maximum likelihood estimator.
+Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
+The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
+For each value of SNR, $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
+
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.
+When $\abs{P} \neq 0$ (i.e.
+coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
+Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
+The solid lines are the MSEs predicted a theoretical result not described here.
+%The theory accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
+As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+
+% Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
+The dots and crosses show the results of Monte-Carlo simulations.
+Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+This requires $G(0)$ to be known.
+In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
+When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
+However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+
+Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
+The figure depicts an interesting phenomenon.
+When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.
+When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
+A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
+%When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
+This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
+For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
+\[
+SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
+\]
+where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
+%For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
+This would enable a rigorous theory for selection of $\beta$ at the receiver.
+One caveat is that the receiver would require knowledge about the noise distribution in order to advantageously choose $\beta$.
+We do not investigate this further here.
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
+		\caption{Phase error versus SNR for BPSK with $L=4096$.}
+		\label{fig:plotphaseBPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
+		\caption{Phase error versus SNR for QPSK with $L=4096$.}
+		\label{fig:plotphaseQPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
+		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
+		\label{fig:plotphase8PSK}
+\end{figure}
+
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
+% 		\caption{Unbiased amplitude error versus SNR for BPSK.}
+% 		\label{fig:plotampBPSK}
+% \end{figure}
+
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+% 		\caption{Unbiased amplitude error versus SNR for QPSK.}
+% 		\label{fig:plotampQPSK}
+% \end{figure}
+
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+% 		\caption{Unbiased amplitude error versus SNR for $8$-PSK.}
+% 		\label{fig:plotamp8PSK}
+% \end{figure}
+
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{code/data/plotM2-3.mps}
+%		\caption{Phase error versus SNR for BPSK.}
+%		\label{fig:plotphase}
+%\end{figure}
+
+\begin{figure}[tp]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
+		\caption{Phase error versus SNR for QPSK.}
+		\label{fig:plotphaseQPSKmultL}
+\end{figure}
+
+\section{Conclusion}
+
+We considered least squares estimators of carrier phase and amplitude from noisy communications signals that contain both pilot signals, known to the receiver, and data signals, unknown to the receiver.
+We focused on $M$-ary phase shift keying constellations.
+The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
+
+%We showed, under some conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
+This bias is large when the signal to noise ratio is small.
+It would be interesting to investigate methods for correcting this bias.
+A method for estimating $G(0)$ at the receiver appears to be required.
+
+Monte Carlo simulations were used to assess the performance of the least squares estimator.
+Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
+This suggests the use of a weighed objective function, which would be an interesting topic for future research.
+
 \small
@@
 
-%\section{Tricks with fractional parts}
-%Throughout the paper, and particularly in  have made use of a number of results involving the fractional part function
-
 \end{document}

### 2012-09-12 algonly.tex
minor

--- 2012-08-28 algonly.tex
@@
 Supported under the Australian Government’s Australian Space Research Program.
-Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
-Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
-Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
-}}
+Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.}}
 

### 2012-10-09 paper1.tex
merged to one paper.

@@
 where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.
-The sets $P$ and $D$ are disjoint, i.e.
-$P \cap D = \emptyset$, and $L = \abs{P \cup D}$ is the total number of symbols transmitted.
-
-We assume that time offset estimation has been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
+The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$, and we let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
+
+We assume that time offset estimation has been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is,
 \begin{equation}\label{eq:sigmod}
@@
 Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.
-Due to space constraints, the lemmas needed by these proofs have been placed in the supplementary paper~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
 In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
@@
 Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
-The zero mean complex Gaussian distribution with independent real and imaginary parts is apart of this class.
+The next proposition describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
 
@@
 This stronger condition is required to prove the asymptotic normality of $\sqrt{L}\hat{m}_L$.
+The next two sections give proofs for Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+The proofs make use of various lemmas.
+The lemmas are proved in the appendix.
 
@@
 \end{equation}  
-Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+Lemma~\ref{lem:GLtoG0} in the appendix shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
 It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
-These are proved in Section~\ref{sec:proof-almost-sureappendix} of \cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
+These are proved in Section~\ref{sec:proof-almost-sureappendix} of the appendix.
 
@@
 This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
+However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
 Define the function
@@
 \]
-as $L \rightarrow \infty$.
-It remains to prove the lemmas that we have used.
-These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
+as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of the appendix.
 
@@
 where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
-For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
+For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm}.
 This would enable a rigorous theory for selection of $\beta$ at the receiver.
@@
 
+\appendix
+
+\subsection{Lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sureappendix}
+
+\begin{lemma}\label{lem:convtoexpGlamL} 
+$\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Since $\sabs{G(x)}$ is uniquely maximised at $x=0$,
+\[
+0 \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)},
+\]
+and since $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(x)}$,
+\[ 
+0 \leq \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}.
+\]
+Thus,
+\begin{align*}
+0 &\leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} \\ 
+& \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} + \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}\\
+& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
+&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
+\end{align*}
+and the last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}.
+\end{IEEEproof}
+
+% \begin{lemma}\label{lem:Zconunif}
+% \[
+% \prob\left\{\sup_{\lambda \in [-\pi, \pi)}\abs{G_L(\lambda) - G(\lambda)} > \epsilon \right\} = O(e^{-\epsilon^2 L}).
+% \]
+% \end{lemma}
+% \begin{IEEEproof}
+% \end{IEEEproof}
+
+\begin{lemma}\label{lem:GLtoG0}
+$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+By the triangle inequality,
+\[
+\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
+\]
+Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:uniflawGGL}
+The function $G_L$ converges almost surely and uniformly on $[-\pi, \pi)$ to the function $G$, that is,
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0
+\] 
+almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Put $T_L(\lambda) = \expect G_L(\lambda)$ and write
+\begin{align*}
+&\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda) + T_L(\lambda) - G(\lambda)} \\
+&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)}.
+\end{align*}
+Now,
+\begin{align*}
+T_L(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
+&= o(1) h_1(\lambda) + o(1) h_2(\lambda)
+\end{align*}
+where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since 
+\[
+h_1(\lambda) = \expect R_1\cos(\lambda + \Phi_1) \leq \expect R_1,
+\]
+and 
+\[
+h_2(\lambda) = \expect R_1\cos\fracpart{\lambda + \Phi_1} \leq \expect R_1
+\] 
+for all $\lambda \in [-\pi, \pi)$, it follows that 
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)} \leq o(1) \expect R_1 \rightarrow 0
+\]  
+as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
+\]
+almost surely as $L\rightarrow \infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:uniflawTGL} 
+Put $T_L(\lambda) = \expect G_L(\lambda)$, then
+\[
+\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
+\] 
+almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Put $D_L(\lambda) = G_L(\lambda) - T_L(\lambda)$ and let
+\[
+\lambda_n = \tfrac{2\pi}{N}(n-1) - \pi, \qquad n = 1, \dots, N
+\]
+be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n$ be the interval $[\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
+\begin{align*}
+\sup_{\lambda \in [-\pi, \pi)}&\sabs{D_L(\lambda)} \\
+&= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n) + D_L(\lambda_n)} \\
+&\leq U_L + V_L,
+\end{align*}
+where 
+\[
+U_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)},
+\]
+\[
+V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
+\]
+Lemma~\ref{lem:supDLlambdan} shows that for any finite $N$, and any $\epsilon > 0$, 
+\[
+\prob\left\{ \lim_{L\rightarrow\infty} U_L > \epsilon \right\} = 0,
+\]
+i.e., $U_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
+\] 
+Choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ and
+\begin{align*}
+\prob& \left\{ \lim_{L\rightarrow\infty} \sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} > 3\epsilon \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty}(U_L + V_L) > 3\epsilon \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty} (U_L + V_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+&\leq \prob\left\{\lim_{L\rightarrow\infty} U_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+&= 0.
+\end{align*}
+Thus, $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:supDLlambdan}
+If $N$ is finite then $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Put
+\[
+Z_i(\lambda) = \begin{cases}
+R_i\cos(\lambda + \Phi_i) , & i \in P \\
+R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
+\end{cases}
+\]
+so that
+\[
+D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
+\]
+Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
+\[
+\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
+\]
+almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Now, since $N$ is finite,
+\[
+U_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
+\]
+almost surely to zero as $L \rightarrow \infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} = 0.
+\]
+\end{lemma}
+\begin{IEEEproof}
+Observe that
+\begin{align*}
+&\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
+&=  \abs{G_L(\lambda) - T_L(\lambda) - G_L(\lambda_n) + T_L(\lambda_n)} \\
+&\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
+&\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
+\end{align*}
+the last line following from Jensen's inequality.  Put
+\[
+C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) },
+\]
+so that
+\begin{align*}
+V_L &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
+&\leq  C_L +  \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect\abs{G_L(\lambda) - G_L(\lambda_n) } \\
+&\leq C_L + \expect C_L,
+\end{align*}
+where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
+\[
+\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+\]
+Thus,
+\begin{align*}
+\prob&\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+\end{align*}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:CL} The following statements hold:
+\begin{enumerate}
+\item $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$,
+\item for any $\epsilon > 0$, $\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0$.
+\end{enumerate}
+\end{lemma}
+\begin{IEEEproof}
+If $\lambda \in L_n$, then $\lambda = \lambda_n + \delta$ with $\delta < \tfrac{2\pi}{N}$, and from Lemma~\ref{lem:coslipshitz},
+\begin{align*}
+&\abs{\cos(\lambda + \Phi_i) - \cos(\lambda_n + \Phi_i)} \leq \tfrac{2\pi}{N}, \;\;\; \text{and} \\
+&\abs{\cos\sfracpart{\lambda + \Phi_i} - \cos\sfracpart{\lambda_n + \Phi_i}} \leq \tfrac{2\pi}{N}.
+\end{align*}
+Because this results does not depend on $n$,
+\[
+\sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \leq R_i \frac{2\pi}{N}
+\]
+for all $i = P \cup D$.  Now,
+\begin{align*}
+C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i \in P \cup D} Z_i(\lambda) - Z_i(\lambda_n)} \\
+&\leq \frac{1}{L}\sum_{i \in P \cup D} \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
+&\leq \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i .
+\end{align*}
+Thus, 
+\[
+\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i = \frac{2\pi}{N}\expect R_1
+\]
+and the first statement holds.  Now, 
+\[
+\frac{2\pi}{N L}\sum_{i \in P \cup D} R_i \rightarrow \frac{2\pi}{N}\expect R_1
+\] 
+almost surely to $\frac{2\pi}{N}\expect R_1$ as $L \rightarrow\infty$ by the strong law of large numbers, and so, for any $\epsilon > 0$,
+\begin{align*}
+\prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} \\
+&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+\end{align*}
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:coslipshitz}
+Let $x$ and $\delta$ be real numbers.  Then
+\begin{align*}
+&\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta}, \;\;\; \text{and} \\
+&\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
+\end{align*}
+\end{lemma}
+\begin{IEEEproof}
+Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipshitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
+\begin{align*}
+&\abs{\cos(y) - \cos(x)} \leq K\abs{x-y} = \abs{x-y}, \;\;\; \text{and} \\
+&\abs{\cos\fracpart{y} - \cos\fracpart{x}} \leq K\abs{x-y} = \abs{x-y}.
+\end{align*}
+ The lemma follows by putting $y = x + \delta$.
+\end{IEEEproof}
+
+\subsection{Lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-normappendix}
+
+\begin{lemma}~\label{lem:diffatlambdaL}
+The derivative of $G_L$ exists, and is equal to zero, at $\hat{\lambda}_L$.  That is,
+\[
+G_L^\prime(\hat{\lambda}_L) = \frac{d G_L}{d \lambda}(\hat{\lambda}_L) = 0.
+\]
+\end{lemma}
+\begin{IEEEproof}
+Observe that 
+\[
+G_L(\lambda) = \frac{1}{L}\sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}
+\] 
+is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ with $R_i > 0$.
+Let $q_i$ be the smallest number from the interval $[-\tfrac{\pi}{M}, 0]$ such that
+\[
+ \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
+\]
+where $\eta = e^{-j2\pi/M} - 1$.  Observe that if $R_i > 0$ then $q_i < 0$.  Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
+\[
+\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i
+\]
+for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.
+
+It remains to show that $G_L^\prime(\hat{\lambda}_L) = 0$.
+If $\sabs{G_L(\hat{\lambda}_L)} > 0$ then $\sabs{G_L(\lambda)}$ is differentiable at $\hat{\lambda}_L$ and, since $\hat{\lambda}_L$ is a maximiser of $\sabs{G_L(\lambda)}$,
+\[
+0 = \frac{d \sabs{G_L(\lambda)}}{d \lambda}(\hat{\lambda}_L) = \begin{cases} 
+G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) > 0\\
+-G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) < 0.
+\end{cases}
+\]
+Thus, $G_L^\prime(\hat{\lambda}_L) = 0$ if $\sabs{G_L(\hat{\lambda}_L)} > 0$.
+On the other hand $\sabs{G_L(\hat{\lambda}_L)} = 0$ only if $G_L(\lambda) = 0$ for all $\lambda \in \reals$, and in this case $G_L^\prime(\hat{\lambda}_L) = 0$ also.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i, i \in D$ be defined as in Lemma~\ref{lem:diffatlambdaL}.
+Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.
+\end{lemma}
+\begin{IEEEproof}
+Recall that $\{\hat{d}_i = \hat{d}_i(\hat{\theta}), i \in D\}$ defined in \eqref{eq:hatdfinxtheta} are the minimisers of the function 
+\[
+SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
+\]
+defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
+\begin{equation}\label{eq:fraclambassumption}
+\sfracpart{\hat{\lambda}_L + \Phi_k} > \frac{\pi}{M} + q_k
+\end{equation}
+for some $k \in D$.  Recalling the notation $e_k$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k$.  We will show that 
+\[
+SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
+\]
+violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  First observe that,
+\begin{align*}
+Y(\{ r_i, i \in D\}) = \sum_{i \in P} y_ip_i^* + \sum_{i \in D} y_i\hat{r}_i^* = \hat{Y} + \eta y_k\hat{d}_k^*,
+\end{align*}
+where $\eta = e^{-j2\pi/M} - 1$ and $\hat{Y} = Y(\{ \hat{d}_i, i \in D\})$.  Now,
+\begin{align*}
+SS&(\{r_i, i \in D\}) \\
+&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
+&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
+&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
+&= SS(\{\hat{d}_i, i \in D\}) - D,
+\end{align*}
+where 
+\[
+D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
+\]
+Now $\frac{1}{L}\hat{Y} = \hat{a} = \hat{\rho} e^{j\hat{\theta}}$ from~\eqref{eq:hata} and using~\eqref{eq:yethetadhat},
+\[
+\frac{1}{L} \hat{Y}^* y_k \hat{d}_k^* = \hat{\rho} \rho_0 R_k e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}},
+\]
+so that
+\begin{equation}\label{eq:Ddefn}
+D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
+\end{equation}
+Let $v = \sfracpart{\hat{\lambda}_L + \Phi_k} - \tfrac{\pi}{M}$ so that
+\begin{align*}
+\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
+&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
+&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) },
+\end{align*}
+and
+\begin{align*}
+\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
+ &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
+\end{align*}
+Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
+\[
+0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
+\]
+Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
+Thus,~\eqref{eq:fraclambassumption} is false.
+
+To show that $\sfracpart{\hat{\lambda}_L + \Phi_k} \geq -\frac{\pi}{M} - q_k$ we use contradiction again.
+Assume that $\sfracpart{\hat{\lambda}_L + \Phi_k} < -\frac{\pi}{M} - q_k$.
+Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.
+Now an analogous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
+
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:Qconv}
+Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
+\[ 
+\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\]
+where $p, d$ and $H$ are real numbers defined in the statements of Theorems~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
+\end{lemma}
+\begin{IEEEproof}
+We have
+\[
+Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0) = \tfrac{\sabs{P}}{L} k_1(\lambda) + \tfrac{\sabs{D}}{L} k_2(\lambda)
+\]
+where
+\begin{align}
+k_1(\lambda) &= \expect R_1\big( \sin(\lambda + \Phi_1) - \sin(\Phi_1) \big), \;\;\; \text{and} \nonumber\\
+k_2(\lambda) &= \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big). \label{eq:k2def}
+\end{align}
+We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\lambda = 0$,
+\[
+\sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
+\]
+where $\zeta$ is a continuous function with $\zeta(0) = 0$.  So,
+\begin{align*}
+k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
+&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big) 
+\end{align*}
+since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ as a result of~\eqref{eq:expectRepartRphi}.
+Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
+$\zeta(\hat{\lambda}_L) = o_P(1)$.
+So,
+\[
+k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) \expect R_1 \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big)
+\]
+because $\expect R_1$ is finite.  Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
+\begin{align*}
+Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
+&= \hat{\lambda}_L \left( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \right) \\
+&= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
+\end{align*}
+since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the above equation by $\sqrt{L}$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:k2conv}
+$k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
+\end{lemma}
+\begin{IEEEproof}
+Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
+We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
+An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
+To keep our notation clean put
+\[
+\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
+\]
+For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
+\[
+\fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
+\]
+Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
+\[
+\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
+\]
+whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
+\[
+\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
+\]
+So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
+&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
+\end{align*}
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
+&= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
+\end{align*}
+where, for each $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+\begin{align*}
+\sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
+&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
+\end{align*}
+and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
+\begin{align*}
+\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
+&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
+\end{align*}
+where
+\[
+s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
+\]
+and
+\[
+\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos\fracpart{\Phi_1} + \zeta_{k+1}(\lambda).
+\]
+Observe that both $s_k$ and $\eta_k$ are continuous and that 
+\[
+s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
+\]
+
+Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
+\]
+whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
+\]
+Since the intervals $[\psi_{k-1}, \psi_{k})$ partition the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
+\[
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
+\]
+where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
+\begin{equation}\label{eq:vk}
+v_k(\lambda,\Phi_1) = \begin{cases}
+s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
+0 & \text{otherwise}.
+\end{cases} 
+\end{equation}
+Now,
+\begin{align*}
+k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
+&= \expect R_1\big( \lambda\cos\fracpart{\Phi_1} + v(\lambda, \Phi_1)  \big) \\
+&= \lambda h_2(0) + \expect R_1 v(\lambda, \Phi_1).
+\end{align*}
+Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a continuous function with $\gamma(0) = 0$, and
+\[
+H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0),
+\]
+where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
+Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, it follows that $\gamma(\hat{\lambda}_L) = o_P(1)$, and so
+\[
+k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( h_2(0) + H_1 + o_P(1) ) = \hat{\lambda}_L ( H + o_P(1) )
+\]
+as required.  It remains to prove Lemma~\ref{lem:expRvlamphi}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:expRvlamphi}
+$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
+\end{lemma}
+\begin{IEEEproof}
+In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
+An analogous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
+Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$,
+\begin{align}
+W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
+&= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
+&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
+&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi, \label{eq:Winfsum}
+\end{align}
+where $g$ is defined in the statement of Theorem~\ref{thm:normality}
+Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
+\begin{align*}
+\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
+\end{align*}
+where, for $k = 1, \dots, M-1$,
+\begin{equation}\label{eq:ak}
+a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
+\end{equation}
+\[
+b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
+\]
+\begin{equation}\label{eq:ck}
+c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
+\end{equation}
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}, it follows that $c_k(\lambda)$ converges to
+\[
+s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
+\]
+as $\lambda$ goes to zero from above.
+We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
+The analogous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
+
+Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
+\[
+\int_0^{2\pi} v_0(\lambda, \phi) g(\phi) d\phi = \lambda a_0(\lambda) + \lambda b_0(\lambda) +  \lambda c_0(\lambda),
+\]
+where $a_0(\lambda)$ and $c_0(\lambda)$ are defined as in~\eqref{eq:ak}~and~\eqref{eq:ck}, and
+\[
+b_0(\lambda) = \zeta_0(\lambda) \int_{0}^{\psi_k-\lambda} g(\phi) d\phi.
+\]
+When $k = M$, we have $0 < \psi_{M-1} < 2\pi < \psi_{M} - \lambda$, so that
+\[
+\int_0^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi = \lambda b_{M}(\lambda)
+\]
+where 
+\[
+b_{M}(\lambda) = \zeta_{M}(\lambda) \int_{\psi_{M-1}}^{2\pi}g(\phi) d\phi.
+\]
+Both $b_0(\lambda)$ and $b_M(\lambda)$ are functions that converge to zero as $\lambda$ converges to zero.
+
+Finally, when $k < 0$ or $k > M$, 
+\[
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0.
+\]
+So the infinite sum in~\eqref{eq:Winfsum} can be written as the finite sum,
+\begin{align*}
+W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
+&= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
+&= \lambda \left( C(\lambda)  + A(\lambda) + B(\lambda) \right),
+\end{align*}
+where 
+\[
+A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
+\]
+\[
+C(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
+\]
+Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
+\[
+\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1.
+\]
+Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$ as required.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:empiricprocc} Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
+\[
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
+\]
+\end{lemma}
+\begin{IEEEproof}
+Put 
+\begin{equation}\label{eq:WLdef}
+W_L(\lambda) = \sqrt{L}\big( R_L(\lambda) - Q_L(\lambda) - R_L(0) \big)
+\end{equation}
+so that
+\[
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = W_L(\hat{\lambda}_L) + \sqrt{L} R_L(0)
+\]
+Lemma~\ref{lem:ZYLempiricproc} in Section~\ref{sec:an-empirical-process} shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+\[
+\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu
+\]
+for all positive integers $L$.  Since $\hat{\lambda}_L$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
+\[
+\lim_{L\rightarrow\infty}\prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} = 0
+\] 
+and therefore $\prob\{ \sabs{\hat{\lambda}_L} \geq \epsilon \} < \nu$ for all sufficiently large $L$.  Now
+\begin{align*}
+  \prob&\left\{\sabs{ W_L(\hat{\lambda}_L) } > \delta \right\} \\
+&= \prob\left\{ \sabs{W_L(\hat{\lambda}_L)} > \delta \;\text{and} \; \sabs{\hat{\lambda}_L} < \epsilon \right\} \\
+&\hspace{0.7cm} + \prob\left\{ \sabs{W_L(\lambda_L)} > \delta  \;\text{and} \; \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\lambda < \epsilon} \sabs{ W_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+&\leq 2\nu.
+\end{align*}
+for all sufficiently large $L$.  Since $\nu$ and $\delta$ can be chosen arbitrarily small, it follows that $W_L(\hat{\lambda}_L)$ converges in probability to zero as $N\rightarrow\infty$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:convdistGLdash}
+The distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$.
+\end{lemma}
+\begin{IEEEproof}
+Observe that $\sqrt{L} R_L(0) = C_L + D_L$ where
+\[
+C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} R_i \sin(\Phi_i), \qquad D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\sfracpart{\Phi_i}.
+\]
+From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
+\[
+p A_1 = p \expect R_1^2 \sin^2(\Phi_1).
+\]
+Similarly, the distribution of $D_L$ converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~\ref{lem:expectImRfracpart}, and variance
+\[
+d A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
+\]
+The lemma holds because $C_L$ and $D_L$ are independent. 
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:expectImRfracpart}
+$\expect R_1 \sin\fracpart{\Phi_1} = 0$.
+\end{lemma}
+\begin{IEEEproof}
+Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
+\begin{align*}
+\expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
+&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi,
+\end{align*}
+where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
+Proposition~\ref{prop:gg2cont} shows that $g$ is even on $(-\pi,\pi)$.
+Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.
+So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
+\[
+g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
+\]
+and
+\[
+\sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
+\]
+Now,
+\begin{align*}
+I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
+&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
+&\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+&= 0
+\end{align*}
+for all integers $k$ because
+\begin{align*}
+\int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
+&= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
+\end{align*}
+Observe that the sets
+\[
+ [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
+\]
+partition $[-\pi, \pi]$, and so
+\[
+\expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
+\]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:empiricprocforrho} Let $T_L(\lambda) = \expect G_L(\lambda)$.  Then
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = o_P(1) + X_L,
+\]
+where $X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big)$.
+\end{lemma}
+\begin{IEEEproof}
+Let
+\begin{equation}\label{eq:YLdef}
+Y_L(\lambda) = \sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) - X_L
+\end{equation}
+so that
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\hat{\lambda}_L) + X_L
+\]
+Lemma~\ref{lem:ZYLempiricproc} in Section~\ref{sec:an-empirical-process} shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+\[
+\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu.
+\]
+The proof now follows by an argument analogous to that in Lemma~\ref{lem:empiricprocc}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:HLtoG} $\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1)$.
+\end{lemma}
+\begin{IEEEproof}
+The argument is similar to that used in Lemma~\ref{lem:Qconv}.  First observe that
+\begin{align*}
+T_L(\lambda) &= \frac{\sabs{P}}{L}\expect R_1\cos(\lambda + \Phi_1) + \frac{\sabs{D}}{L} \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
+%&= \big( p + o(1) \big) \expect R_1\cos(\lambda + \Phi_1) \\
+%& \hspace{2cm} + \big( d + o(1) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1} \\
+&=  \big( p + o(L^{-1/2}) \big) \expect R_1\cos(\lambda + \Phi_1) \\
+&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1},
+\end{align*}
+because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ is an assumption in the statement of Theorem~\ref{thm:normality}.  Now,
+\begin{align*}
+\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
+\end{align*}
+where
+\[
+q_1(\lambda) = \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big),
+\]
+\begin{equation}
+q_2(\lambda) = \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
+\end{equation}
+We analyse $q_1(\lambda)$ first.  By a Taylor expansion of $\cos(\lambda + \Phi_1)$ about $\lambda = 0$,
+\[
+\cos(\lambda + \Phi_1) - \cos(\Phi_1) = -\lambda \big( \sin(\Phi_1) + \zeta(\lambda) \big),
+\]
+where $\zeta$ is a continuous function with $\zeta(0) = 0$. So,
+\begin{align*}
+q_1(\lambda) &= -\expect R_1\lambda(\sin(\Phi_1) +  \zeta(\lambda)) \\
+&= -\lambda \big( \expect R_1\sin(\Phi_1) +  \zeta(\lambda) \expect R_1 \big) \\
+&= -\lambda \zeta(\lambda) \expect R_1,
+\end{align*}
+since $\expect R_1\sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}.  Now, 
+\begin{align*}
+\sqrt{L} q_1(\hat{\lambda}_L) = -\sqrt{L} \hat{\lambda}_L \zeta(\hat{\lambda}_L) \expect R_1 = o_P(1)
+\end{align*}
+because $\sqrt{L} \hat{\lambda}_L$ converges in distribution and $\zeta(\hat{\lambda}_L) = o_P(1)$.
+An argument analogous to that used in Lemma~\ref{lem:k2conv} shows that $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
+Thus,
+\[
+\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1) = o_P(1)
+\]
+as required.
+\end{IEEEproof}
+
+% \begin{lemma}\label{lem:XL} 
+% $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
+% \end{lemma}
+% \begin{IEEEproof}
+% The argument is similar to that used in Lemma~\ref{lem:k2conv}.
+As $\hat{\lambda}_L \in [-\pi, \pi)$ it is only the behaviour of the function $q_2(\lambda)$ for $\lambda\in [-\pi, \pi)$ that is relevant.
+We will analyse $q_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
+An analagous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
+Recall that
+% \[
+% \psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
+% \]
+% when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
+% \begin{align*}
+% \cos\fracpart{\lambda + \Phi_1} - &\cos\fracpart{\Phi_1} \\
+% &= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
+% \end{align*}
+% and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
+% \begin{align*}
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} &= -\lambda\big(\sin(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
+% &= -\lambda\big(\sin\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
+% \end{align*}
+% where, for each integer $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.   Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+% \begin{align*}
+% \cos\fracpart{\lambda + \Phi_1} &- \cos\fracpart{\Phi_1} \\
+% &= \cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k)
+% \end{align*}
+% and by a Taylor expansion of $\cos(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
+% \begin{align*}
+% \cos&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
+% &= c_k(\Phi_1) - \lambda\big(\sin\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
+% \end{align*}
+% where
+% \[
+% c_k(\Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos(\Phi_1 - \tfrac{2\pi}{M}k),
+% \]
+% and
+% \[
+% \eta_k(\lambda, \Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi k}{M}) + \zeta_{k+1}(\lambda).
+% \]
+% Observe that $c_k$ is continuous and $c_k(\psi_{k}) = 0$.
+
+% Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
+% \[
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
+% \]
+% whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
+% \[
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda \sin\fracpart{\Phi_1} + c_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
+% \]
+% Since the intervals $[\psi_{k-1}, \psi_{k})$ partion the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
+% \[
+% \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} = -\lambda\sin\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
+% \]
+% where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
+% \begin{equation}\label{eq:cvk}
+% v_k(\lambda,\Phi_1) = \begin{cases}
+% c_k(\Phi_1) - \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+% \lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
+% 0 & \text{otherwise}.
+% \end{cases}
+% \end{equation}
+% \end{IEEEproof}
+
+\begin{lemma}\label{lem:XL} 
+The distribution of 
+\[
+X_L = \sqrt{L} \big( G_L(0) - T_L(0) \big) = \sqrt{L} \big( G_L(0) - \expect G_L(0) \big)
+\] 
+converges to the normal with zero mean and covariance 
+\[
+pB_1 + d B_2
+\] 
+as $L \rightarrow\infty$.
+\end{lemma}
+\begin{IEEEproof}
+Observe that $X_L = C_L + D_L$ where
+\[
+C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} \big( R_i \cos(\Phi_i) - h_1(0) \big),
+\]
+\[
+D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ),
+\]
+where $h_1$ and $h_2$ are defined in the statement of Theorem~\ref{thm:consistency}.
+From the standard central limit theorem the distribution of $C_L$ converges to the normal with zero mean and variance
+\[
+p B_1 = p \expect R_1^2 \cos^2(\Phi_1) - p h_1^2(0) =  p \expect R_1^2 \cos^2(\Phi_1) - p 
+\]
+since $h_1(0) = 1$.  Similarly the distribution of $D_L$ converges to the normal with zero mean and variance
+\[
+d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0).
+\]
+The lemma follows since $C_L$ and $D_L$ are independent.
+\end{IEEEproof}
+
 %\section{Tricks with fractional parts}
@@
 
+\subsection{An empirical process result}\label{sec:an-empirical-process}
+
+During the proof of asymptotic normality in Section~\ref{sec:proof-asympt-norm} we made use of the following result regarding the functions $W_L(\lambda)$ and $Y_L(\lambda)$ defined in~\eqref{eq:WLdef}~and~\eqref{eq:YLdef}.
+
+\begin{lemma}\label{lem:ZYLempiricproc}
+For any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that:
+\begin{enumerate}
+\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu$, and,
+\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu$,
+\end{enumerate}
+for all positive integers $L$.
+\end{lemma}
+\begin{IEEEproof}
+These results are related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
+The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989,Gine_Zinn_symmetrisation_1984,Ossiander_clt_bracketing_1984}.
+We will only give a proof of the first statement regarding the function $W_L$.
+The proof for the second statement regarding $Y_L$ is similar.
+
+Put 
+\[
+f_i(\lambda, R_i, \Phi_i) = \begin{cases}
+R_i( \sin(\lambda + \Phi_i) - \sin(\Phi_i)), & i \in P \\
+R_i( \sin\sfracpart{\lambda + \Phi_i} - \sin\sfracpart{\Phi_i}), & i \in D \\
+\end{cases}
+\]
+so that
+\[
+W_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect f_i(\lambda, R_i, \Phi_i) \big).
+\]
+Let $\{g_i\}$ be a sequence of independent standard normal random variables, independent of $\{R_i\}$ and $\{\Phi_i\}$.  Lemma~\ref{lem:symmetrisation} shows that
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) },
+\]
+where 
+\begin{equation}\label{eq:ZpsiCondGaussProc}
+X(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} g_i f_i(\lambda, R_i, \Phi_i),
+\end{equation}
+and where $\expect$ runs over all of $\{g_i\}$, $\{R_i\}$ and $\{\Phi_i\}$.
+Conditionally on $\{\Phi_i\}$ and $\{R_i\}$, $\{X(\lambda), \lambda \in \reals \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
+Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists $\epsilon > 0$ such that
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa.
+\]
+It follows that,
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)}  <  \sqrt{2\pi} \; \kappa,
+\]
+and by Markov's inequality,
+\[
+\prob \cubr{  \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
+\]
+for any $\delta > 0$.  The proof follows by choosing $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:symmetrisation} (Symmetrisation)
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) }
+\]
+\end{lemma}
+\begin{IEEEproof}
+Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be a sequence identically distributed to $\{\Phi_i\}$.
+Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
+Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
+Now,
+\begin{align*}
+W_L(\lambda) &= \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect_{R,\Phi} f_i(\lambda, R_i', \Phi_i') \big) \\
+&= \expect_{R,\Phi} \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ),
+\end{align*}
+where, for notational convenience, we put 
+\[
+f_{i} = f_i(\lambda, R_i, \Phi_i) \qquad \text{and} \qquad  f_{i}^\prime = f_i(\lambda, R_i', \Phi_i').
+\] 
+Taking absolute values followed by supremums,
+\begin{align*}
+ \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &= \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \expect_{R,\Phi} \sum_{i \in P \cup D} (f_{i} - f_{i}^\prime )} \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi}  \abs{ \frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) },
+\end{align*}
+the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, the bound can be increased by moving $\expect_{R,\Phi}$ outside the supremum,
+\begin{equation}\label{eq:Eoutsidesup}
+ \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ) }.
+\end{equation}
+Applying $\expect$ to both sides gives the inequality
+\begin{equation}\label{eq:Gninequfin}
+ \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} (f_{i} - f_{i}^\prime) },
+\end{equation}
+since $\expect \expect_{R,\Phi}$ is equivalent to $\expect$.   Let
+\[
+\sigma_i = \frac{g_i}{\abs{g_i}},
+\]
+and put $\sigma_i = 1$ if $g_i = 0$.  The $\sigma_i$ are thus independent with 
+\[
+\prob\{ \sigma_i = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_i = 1 \} = \tfrac{1}{2}.
+\]  
+As $R_i$ and $\Phi_i$ are independent and identically distributed to $R_i^\prime$ and $\Phi_i^\prime$, the random variable $f_{i} - f_{i}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_i( f_{i} - f_{i}^\prime)$.
+Thus,
+%\[
+%\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
+%\]
+%are identically distributed, and therefore
+\begin{align*}
+ \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) } \\
+&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i ( f_{i} - f_{i}^\prime) }.
+\end{align*}
+By the triangle inequality,
+\[
+\abs{ \sum_{i\in P \cup D} \sigma_i (f_{i} - f_{i}^\prime)} \leq \abs{ \sum_{i \in P \cup D} \sigma_i f_{i} } + \abs{ \sum_{i\in P \cup D}  \sigma_i f_{i}^\prime },
+\]
+and it follows from~\eqref{eq:Gninequfin}, that
+\begin{align*}
+ \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &\leq \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} \sigma_i f_{i} } \\
+&\hspace{1cm} + \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  \sigma_i  f_{i}^\prime } \\
+&= 2 \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} }.
+\end{align*}
+Because $g_i$ is a standard normal random variable it follows that $\expect \abs{g_i} = \expect_{R,\Phi} \abs{g_i} = \sqrt{2/\pi}$, and
+\begin{align*}
+\expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} } \\
+&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} \sqrt{\frac{\pi}{2}}\expect_{R,\Phi} \abs{g_i}} \\
+&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{ \sqrt{L}}\sum_{i \in P \cup D} \sigma_i \abs{g_i}  f_{i}},
+\end{align*}
+the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$,
+\begin{align*}
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} g_i  f_{i}} \\
+&= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) } 
+\end{align*}
+as required.
+\end{IEEEproof}
+
+\begin{lemma} \label{lem:chaining}
+For any $\kappa > 0$ there exists $\epsilon > 0$ such that
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa,
+\]
+where $X(\lambda)$ is the Gaussian process defined in~\eqref{eq:ZpsiCondGaussProc}.
+\end{lemma}
+\begin{IEEEproof}
+Without loss of generality, assume that $\epsilon < \tfrac{\pi}{M}$.  Lemma~\ref{lem:chaining2} shows that
+\begin{equation}\label{eq:supZCK1} 
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}
+\end{equation}
+where $K_1$ is a finite, positive constant, and
+\begin{equation}\label{eq:Cepsdefn}
+C_\epsilon = \frac{2}{L}\epsilon\sum_{i\in P \cup D}R_i^2 + \frac{4}{L}\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
+\end{equation}
+where $I_\epsilon$ is the indicator function
+\[
+I_{\epsilon}(x) = \begin{cases}
+1, & x \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon] \\
+0, & \text{otherwise},
+\end{cases}
+\] 
+and $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  Now,
+\[
+\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
+\]
+Recalling that $f(r,\phi)$ is the joint probability density function of $R_i$ and $\Phi_i$,
+\begin{align*}
+\expect R_i^2 I_\epsilon(\Phi_i) &= \int_0^{2\pi} \int_{0}^\infty  r^2 f(r,\phi) I_\epsilon(\phi) dr d\phi \\ 
+&= \int_0^{2\pi}g_2(\phi) I_\epsilon(\phi) d\phi 
+\end{align*}
+where the function
+\[
+g_2(\phi) = \int_{0}^\infty  r^2 f(r,\phi) dr d\phi 
+\]
+is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}.  From the definition of $I_\epsilon$,
+\begin{align*}
+\expect R_i^2 I_\epsilon(\Phi_i) &= \sum_{k=0}^{M-1}\int_{\psi_k-\epsilon}^{\psi_k-\epsilon}g_2(\phi) d\phi \\
+&= 2\epsilon \sum_{k=0}^{M-1} (g_2(\psi_k) + o(1))
+\end{align*}
+where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.
+Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$.
+Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive,
+\[
+\expect \sqrt{C_\epsilon} \leq  \sqrt{\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
+\]
+by Jensen's inequality.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
+\[
+\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
+\]
+Choosing 
+\[
+\epsilon = \frac{\kappa^2}{ K_1^2 ( K_2 + 2\expect R_1^2)}
+\] 
+completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
+\end{IEEEproof}
+
+The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
+The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
+%Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.
+For notational convenience we use the abbreviation 
+\[
+f_{i}(\lambda) = f_{i}(\lambda, R_i, \Phi_i)
+\]
+in what follows.
+As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{\pi}{M}$.
+%This comes at no loss since Lemma~\ref{lem:chaining} only asks for the existence of an $\epsilon > 0$.
+
+\begin{lemma}\label{lem:chaining2}
+There exists a positive constant $K_1$ such that
+\[
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}.
+\]
+where $C_\epsilon$ is defined in~\eqref{eq:Cepsdefn}
+\end{lemma}
+\begin{IEEEproof}
+Let $B_\epsilon = \{x \in \reals \mid \abs{x} < \epsilon\}$ be the set of real numbers with magnitude less than $\epsilon$.
+For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $B_\epsilon$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the pseudometric
+\[
+d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
+\]
+and the pseudometric
+\[
+d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L
+\]
+where 
+\begin{equation}\label{eq:Depsdefn}
+D_\epsilon = \frac{1}{L}\epsilon\sum_{i\in P \cup D}R_i + \frac{2}{L}\sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
+\end{equation}
+We specifically define $T_\epsilon(0)$ to contain only zero.
+Defined this way, $T_\epsilon(0)$ satisfies the inequality above because $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
+The existence of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{lem:metricentropy}.
+Lemma~\ref{lem:metricentropy} also shows that the number of elements in $T_\epsilon(k)$ is no more that $K_3 2^k$ where $K_3$ is a constant, independent of $L$, $\epsilon$ and $k$.
+
+Lemma~\ref{lem:limexpTk} shows that
+\[
+\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}.
+\]
+Lemma~\ref{lem:chaining3} shows that
+\[
+\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}
+\]
+The sum above converges as $k\rightarrow\infty$ so lemma holds with
+\[
+K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
+\]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:epslmlemma}
+Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.
+\end{lemma}
+\begin{IEEEproof}
+We prove the lemma for $d_2$.  The proof for $d_1$ is similar.  Observe that $d_2(\lambda, 0) = c_1 + c_2$ where 
+ \[
+ c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) },
+ \] 
+ \[
+ c_2 = \sum_{i \in D} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+ \]
+ As $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+ \[
+ \sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i)} \leq K\sabs{\lambda} = \sabs{\lambda} < \epsilon
+ \]
+ for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let 
+\[
+S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}.
+\] 
+Now $c_2 = c_3 + c_4$ where 
+ \[
+ c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}},
+ \]
+ \[
+ c_4 =  \sum_{i \in D, i \notin S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+ \]
+Observe that $I_\epsilon(\Phi_i) = 1$ if and only if $\abs{\fracpart{\Phi_i}} \geq \tfrac{\pi}{M} - \epsilon$.  So, if $i \notin S$ and $\sabs{\lambda} < \epsilon$, then
+\begin{align*}
+\round{\lambda + \Phi_i} &= \round{\lambda + \fracpart{\Phi_i} + \round{\Phi_i}} \\ 
+&= \round{\lambda + \fracpart{\Phi_i}} + \round{\Phi_i} = \round{\Phi_i}
+\end{align*}
+because
+\[
+\sabs{\lambda + \fracpart{\Phi_i}} < \epsilon + \tfrac{\pi}{M} - \epsilon = \tfrac{\pi}{M}
+\] 
+and therefore $\round{\lambda + \fracpart{\Phi_i}} = 0$.  Now,
+ \begin{align*}
+ \vert\sin&\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}\vert \\
+ &= \sabs{\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\Phi_i - \round{\Phi_i})} \\
+ &\leq \sabs{\lambda} < \epsilon
+ \end{align*}
+and so, 
+\[
+c_4 \leq \epsilon \sum_{i \in D, i \notin S} R_i \leq \epsilon \sum_{i \in D} R_i.
+\]
+Otherwise, if $i \in S$, then $\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}} \leq 2$ because $\sin(\cdot) \leq 1$.  Thus 
+\[
+c_3 \leq 2 \sum_{i \in D, i \in S} R_i \leq 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i).
+\]
+Now,
+\begin{align*}
+d_2(\lambda, 0) &= c_1 + c_2 = c_1 + c_3 + c_4 \\
+&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i) = D_\epsilon L
+\end{align*}
+as required
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:limexpTk} With $T_\epsilon(k)$ defined as in Lemma~\ref{lem:chaining2},
+\[
+\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}
+\]
+\end{lemma}
+\begin{IEEEproof}
+Let $b_2(\lambda)$ be a function that maps $\lambda \in B_\epsilon$ to a point in $T_\epsilon(k)$ such that  $d_2(\lambda,b_2(\lambda)) \leq 2^{-k}D_\epsilon L$.
+The existence of $b_2$ is guaranteed by the definition of $T_\epsilon(k-1)$.
+So, for all $\lambda \in B_\epsilon$,
+\begin{align*}
+\sabs{X(\lambda) - X(b_2(\lambda))} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda) - f_i(b_2(\lambda)))} \\
+&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
+&\leq \frac{1}{\sqrt{L}} \sum_{i\in P \cup D} g_\text{max} \sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} d_2(\lambda,b_2(\lambda)) \\
+&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L,
+\end{align*}
+where $g_\text{max} = \sup_{i\in P\cup D} \abs{g_i}$. The bound applies uniformly over $\lambda \in B_\epsilon$, so
+\[
+\sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L.
+\]
+Applying $\expect_{R,\Phi}$ to both sides,
+\[
+\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq K_4 2^{-k} D_\epsilon L,
+\]
+where, as a result of Lemma~\ref{lem:maxineq},
+\[
+K_4 = \tfrac{1}{\sqrt{L}}\expect_{R,\Phi} g_\text{max} \leq 3 \sqrt{\frac{\log L}{L}}
+\]
+is independent of $\epsilon$ and $k$.  Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
+\[
+\sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)},
+\]
+and so,
+\begin{align*}
+0 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \nonumber \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \nonumber \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \big( \sabs{X(\lambda)} - \sabs{X(b_2(\lambda))} \big) \nonumber \\
+&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda) - X(b_2(\lambda))} \nonumber \\
+&\leq K_4 2^{-k} D_\epsilon L. \label{eq:Deps2nktozero}
+\end{align*}
+The second line in the inequality above follows because,
+\[
+\sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)},
+\]
+since $b_2(\lambda) \in T_\epsilon(k)$.  The lemma follows because $K_4 2^{-k} D_\epsilon L$ goes to zero as $k\rightarrow\infty$.
+\end{IEEEproof}
+
+% \begin{lemma}\label{lem:epslmlemma}
+% Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
+% \begin{align*}
+% d(\lambda, \lambda^*) &= \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \\
+% &\leq R_i^2 L \left( (\lambda - \lambda^*)^2 + 4 C_\epsilon \right).
+% \end{align*}
+% \end{lemma}
+
+% \begin{lemma}
+% If $i \in P$, then $\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2.$
+% \end{lemma}
+% \begin{IEEEproof} 
+% We have
+% \[
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 = R_i^2\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)}^2
+% \]
+% and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+% \[
+% \sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)} \leq K\sabs{\lambda - \lambda^*} = \sabs{\lambda - \lambda^*}.
+% \]
+% \end{IEEEproof}
+
+% \begin{lemma}
+% Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$.  If $i \in D$, then
+% \begin{equation}\label{eq:fismall}
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2
+% \end{equation}
+% when $\Phi_i \notin S_\epsilon$, otherwise,
+% \begin{equation}\label{eq:fibig}
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq 4 R_i^2.
+% \end{equation}
+% \end{lemma}
+% \begin{IEEEproof}
+% Because $\sin(\cdot) \leq 1$ it follows that
+% \begin{align*}
+% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 &= R_i^2\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}}^2 \\
+% &\leq 4 R_i^2
+% \end{align*}
+% for any $\Phi_i \in \reals$, and so~\eqref{eq:fibig} holds.  If $\Phi_i \notin S_\epsilon$ then $\Phi_i \in (\psi_{k-1} + \epsilon, \psi_{k} - \epsilon)$ for some $k\in\ints$ and 
+% \[
+% \fracpart{\Phi_i + \lambda} = \Phi_i + \lambda - \tfrac{2\pi}{M}k
+% \] 
+% and
+% \[
+% \fracpart{\Phi_i + \lambda^*} = \Phi_i + \lambda^* - \tfrac{2\pi}{M}k 
+% \]
+% and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+% \begin{align*}
+% \vert \sin&\fracpart{\Phi_i + \lambda} - \sin\fracpart{\Phi_i + \lambda^*}\vert \\
+% &=  \sabs{\sin(\Phi_i + \lambda - \tfrac{2\pi}{M}k) - \sin(\Phi_i + \lambda^* - \tfrac{2\pi}{M}k)} \\
+% &\leq K \sabs{\lambda - \lambda^*} \\
+% &= \sabs{\lambda - \lambda^*}.
+% \end{align*}
+% Thus~\eqref{eq:fismall} holds. 
+% \end{IEEEproof}
+
+\begin{lemma}\label{lem:chaining3}(Chaining)
+For all positive integers $k$,
+\[
+\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k) } \abs{ X(\lambda) } \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
+\]
+\end{lemma}
+\begin{IEEEproof}
+Let $b_1(\lambda)$ map each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
+The existence of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
+By the triangle inequality
+\[
+\sabs{X(\lambda)} \leq \sabs{X(b_2(\lambda))} + \sabs{X(\lambda) - X(b_2(\lambda))}  
+\]
+and taking supremums on both sides,
+\begin{align}
+&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \nonumber \\
+&\leq \sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}   \nonumber \\
+&\leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}, \label{eq:Xb2recpreexp}
+\end{align}
+the last line because $b_2(\lambda) \in T_\epsilon(k-1)$ and so
+\[
+\sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)}.
+\]
+Conditional on $\{R_i\}$ and $\{\Phi_i\}$ the random variable
+\[
+U(\lambda) = X(\lambda) - X(b_2(\lambda))
+\]
+has zero mean and is normally distributed with variance
+\begin{align*}
+\sigma_U^2 &= \expect_{R,\Phi} \frac{1}{L} \sum_{i \in P \cup D} g_i^2 \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
+&=  \frac{1}{L} \sum_{i \in P \cup D} \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
+&= d(\lambda, b_2(\lambda)) \leq 2^{1-k} C_\epsilon,
+\end{align*}
+since $\expect_{R,\Phi}g_i^2 = 1$.  So, from Lemma~\ref{lem:maxineq},
+\begin{align*}
+\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{U(\lambda)} &\leq 3 \sqrt{ 2^{1-k} C_\epsilon \log \abs{T_\epsilon(k)} } \\
+&\leq \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}.
+\end{align*}
+Taking expectations on both sides of~\eqref{eq:Xb2recpreexp},
+\begin{align*}
+\expect_{R,\Phi}&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \\
+&\leq \expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}} ,
+\end{align*}
+which involves a recursion in $k$.  By unravelling the recursion and using that $T_\epsilon(k)$ contains only zero, and therefore
+\[
+\expect_{R,\Phi} \sup_{\lambda \in T_k(0)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sabs{X(0)} = 0
+\]
+we obtain,
+\[
+\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \leq  \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}  
+\]
+as required.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:maxineq}(Maximal inequalities)
+Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
+\[
+\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}
+\]
+where $\log N$ is the natural logarithm of $N$.
+\end{lemma}
+\begin{IEEEproof}
+This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:metricentropy}(Covering numbers)
+For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset B_\epsilon$ with the property that, for every $\lambda \in B_\epsilon$, there is a $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the pseudometric
+\[
+d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
+\]
+and the pseudometric
+\[
+d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L,
+\]
+where $C_\epsilon$ and $D_\epsilon$ are defined in~\eqref{eq:Cepsdefn} and~\eqref{eq:Depsdefn}.
+The number of elements in $T_\epsilon(k)$ is no more than $K_3 2^{k}$ where $K_3$ is a positive constant, independent of $L$, $\epsilon$ and $k$.
+\end{lemma}
+\begin{IEEEproof}
+% We will first construct a set $T^2_\epsilon(k)$ containing no more than $K_{23}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_2 \in T^2_\epsilon(k)$ such that $d_2(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{23}$ a constant, independent of $L$, $\epsilon$ and $k$.
+Using a similar construction we find a set $T^1_\epsilon(k)$ containing no more than $K_{13}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_1 \in T^1_\epsilon(k)$ such that $d_1(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{13}$ a constant, independent of $L$, $\epsilon$ and $k$.
+Puting $T_\epsilon(k) = T^1_\epsilon(k) \cup T^2_\epsilon(k)$ will then complete the proof with $K_3 = K_{13} + K_{23}$.
+
+% We will only give a construction for $T^2_\epsilon(k)$, the construction for $T^1_\epsilon(k)$ being similar.  First observe that
+% \[
+% f_{i}(\lambda) - f_{i}(\lambda_2) = \begin{cases}
+% R_i\big( \sin(\lambda + \Phi_i) - \sin(\lambda_2 + \Phi_i) \big) & i \in P \\
+% R_i\big(\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda_2 + \Phi_i}\big) & i \in D
+% \end{cases}
+% \]
+% Let 
+% \[
+% U = \{\epsilon(\ell 2^{-k} - 1) \mid \ell = 0, 1, \dots, 2^{k+1} \} 
+% \]
+% be a set containing $2^{k+1} + 1$ elements uniformly spaced over $B_\epsilon$.
+For every $\lambda \in B_\epsilon$ there exists a $\lambda^* \in U$ such that $\abs{\lambda - \lambda^*} < \epsilon 2^{-k}$.
+Since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$, it follows that if $i \in P$ then,
+% Let 
+% \begin{align*}
+% \abs{f_{i}(\lambda) - f_{i}(\lambda_2)} &= R_i\abs{ \sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i) } \\
+% &\leq R_i\sabs{\lambda - \lambda^*} \\
+% &< \epsilon 2^{-k} R_i.
+% \end{align*}
+% Thus
+% \[
+% \sum_{i \in P} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i\in P} R_i.
+% \]
+% As in Lemma~\ref{lem:epslmlemma} let $S = \{ i \in D \mid I_\epsilon(\Phi_i) \}$. Now, if $i \in D$ but $i \notin S$ then 
+% \[
+% \round{\lambda + \Phi_i} = \round{\lambda^* + \Phi_i} = \round{\Phi_i},
+% \] 
+% so
+%  \begin{align*}
+%  \vert f_{i}(\lambda) - &f_{i}(\lambda^*)\vert \\
+% &= R_i\vert\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}\vert \\
+%  &= R_i\vert\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\lambda^* + \Phi_i - \round{\Phi_i})\vert  \\
+%  &\leq R_i \sabs{\lambda - \lambda^*} \\
+% &< \epsilon 2^{-k} R_i.
+%  \end{align*}
+% Thus,
+% \[
+% \sum_{i \in D, i \notin S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i \in D, i \notin S} R_i 
+% \]
+
+% To handle the case when $i \in S$ we need to add some extra elements to the set $U$.  Let 
+% \[
+% r_i = \begin{cases}
+% \tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} > 0 \\
+% -\tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} \leq 0
+% \end{cases}
+% \]
+% and let $\sigma$ be a permutation of the indices in $S$ such that $r_i$ are in ascending order, that is,
+% \begin{equation}\label{eq:sigmasortind}
+% r_{\sigma(i)} \leq r_{\sigma(k)}
+% \end{equation}
+% whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{S}-1\}$.  Now,
+% \begin{align*}
+% \sum_{i\in S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} = \sum_{i = 0}^{\abs{S}-1} \sabs{ f_{\sigma(i)}(\lambda) - f_{\sigma(i)}(\lambda^*)}.
+% \end{align*}
+% If $\epsilon > \lambda \geq r_{\sigma(k)} > 0$ for some $k$, then
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = \tfrac{\pi}{M}
+% \]
+% for all $i \leq k$, whilst
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
+% \]
+% for all $i > k$.  Otherwise if $-\epsilon < \lambda < r_{\sigma(k)} \leq 0$ for some $k$, then
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + -\tfrac{\pi}{M} - r_{\sigma(i)}} = -\tfrac{\pi}{M}
+% \]
+% for all $i \leq k$, whilst
+% \[
+% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
+% \]
+% for all $i > k$.
+
+% Let $R = \sum_{i \in S} R_i$ and let $d(\ell)$ be the largest integer such that
+% \[
+% \sum_{i=0}^{d(\ell)} R_{\sigma(i)} \leq \frac{\ell}{2^k} R.
+% \]
+% for integers $\ell = 1, \dots, 2^k$.  Define the intervals
+% \[
+% L_\ell = \in [r_{d(\ell)}, r_{d(\ell+1)})
+% \]
+% where $r_{0} = -\epsilon$ and $r_{2^k+1} = \epsilon$.
+Then $L_0, L_1, \dots, L_{2^k}$ partition the closure of $B_\epsilon$.
+If $\lambda$ and $\lambda^*$ are from interval $L_\ell$, then for all $i \in \{0, 1, \dots, \abs{S}-1\}$ less than $d(\ell)$,
+% \begin{align*}
+% \fracpart{\lambda + \Phi_{\sigma(i)}} &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}} \\
+% &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}}
+% \end{align*}
+% \begin{align*}
+% \vert f_{\sigma(i)}(\lambda) - &f_{\sigma(i)}(\lambda^*) \vert \\
+% &= R_{\sigma(i)}\vert\sin\fracpart{\lambda + \Phi_{\sigma(i)}} - \sin\fracpart{\lambda^* + \Phi_{\sigma(i)}}\vert 
+% \end{align*}
+BLERG
+
+\end{IEEEproof}
+
 \end{document}

### 2012-10-10 other/confnoncoherent/paper.tex
Added to conf

@@
 \title{Noncoherent least squares estimators of carrier phase and amplitude}
-\author{Robby McKilliam$^{1}$, Andre Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \vspace{0.2cm} \\
+\author{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \vspace{0.2cm} \\
 \small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.1cm} \\
@@
 For this reason raw symbol error rate is not of interest at this stage.
-Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated received symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
-The additive noise channel is a common assumption for subsequent receiver operations, such as decoding.
-Consequently, our metric of performance is mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
-It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
-The function $\fracpart{\cdot}$ denotes its argument taken `modulo $\tfrac{2\pi}{M}$' into the interval $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$.
-It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate error for the phase parameter.
+Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
+The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
+The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be $\abs{a_0 - \hat{a}}^2$.
+It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle^2$ and $(\rho_0 - \hat{\rho})^2$.
+The function $\fracpart{\cdot}$ denotes its argument taken `modulo $\tfrac{2\pi}{M}$' into the interval $[-\pi/M, \pi/M)$.
+It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
 
@@
 Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
-Section~\ref{sec:stat-prop-least} describes the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
+Section~\ref{sec:stat-prop-least} describes the statistical properties of the least squares estimators of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
 We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
-However, $\hat{\rho_0}$ is not a consistent estimator of the amplitude $\rho_0$.
+However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
 The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
@@
 
-\begin{corollary}\label{cor:ampmse}
-The mean square error of the amplitude estimator $\hat{\rho}$ is 
-\[
-\expect (\rho_0 - \hat{\rho})^2 = \frac{B}{L} + \rho_0^2(1 - G(0))^2 + o(1)
-\] 
-where $o(1)$ goes to zero as $L \rightarrow \infty$.
-\end{corollary}
-\begin{IEEEproof} Write
-\begin{align*}
-\expect (\rho_0 - \hat{\rho})^2 &= \expect \big( \rho_0(1 - G(0)) + G(0)\rho_0 - \hat{\rho} \big)^2 \\
-&= \expect \big( \rho_0(1 - G(0)) - \hat{m}_L \big)^2 \\
-&= \rho_0^2(1 - G(0))^2 + \expect \hat{m}_L^2 - 2\rho_0(1 - G(0))\expect\hat{m}_L,
-\end{align*} 
-and the corollary holds since $\expect \hat{m}_L^2 = B/L + o(1)$ and $\expect\hat{m}_L = o(1)$ from Theorem~\ref{thm:normality}.
-\end{IEEEproof}
-
-%BLERG discuss assumptions.
-We have that $h_1$ is uniquely maximised at $0$ over the interval $[-\pi,\pi)$ since $\{w_i\}$ has zero mean.
-The requirement that $h_2$ is uniquely maximised at 0 is stronger that we need.
-We actually only require that the sum $p h_1(x) + dh_2(x)$ is maximised uniquely at 0 over the interval $[-\pi, \pi)$.
-However, we have stated the theorem in this stronger form as it makes the proof in the noncoherent case (i.e.
-when there are no pilot symbols) follow more transparently.
-
-%BLERG $\hat{a}$ is a biased estimator, particularly when the SNR is low, we show firs that $\hat{\theta}$ is strongly consistent, we are then able to analyse the behaviour of $\hat{a}$.
-
 We now discuss the assumptions made by these theorems.
@@
 A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$.
-The next Lemma describes a class of circularly symmetric distributions that satisfy this property.
+The next proposition describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
 We omit the proof due to space constraints.
 
-\begin{lemma}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+\begin{proposition}\label{prop:contgg2}
+Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
 Let,
@@
 \]
-Then $\abs{G(x)}$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
-\end{lemma}
-
-The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the probability density function of $X$ decreases as we move away from the origin.
-That is, the pdf of $X$ in rectangular coordinates is $p(x,y) = z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$, and this is required to be non increasing with $z$.
-For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
+Then $\abs{G(x)}$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
+\end{proposition}
 
@@
 \]
-Each of $A$, $B$ and $H$ can be computed efficiently by numerical integration using the above formula for $g(\phi)$.
+Each of $A$, $B$ and $H$ can be computed by numerical integration using the above formula for $g(\phi)$.
 
@@
 In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
-Simulations are run with $M=2,4$ (BPSK and QPSK) and $L=16,256,4096$ with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
+Simulations are run with $M=2,4$ (BPSK and QPSK) and $L=16,256,4096$ with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.
 The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
-For each value of SNR $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
+For each value of SNR $T = 10000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
 
@@
 The dots are the Monte-Carlo simulations of the least square estimator.
-The dashed line the performance of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
+The dashed line the MSE of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
 %The dashed line acts as a lower bound on performance.
@@
 
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the MSE of the amplitude estimator $\hat{\rho}$.
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+This estimator requires the selection of a function $F : \reals \mapsto \reals$.
+We have chosen $F(x) = 1$.
+The MSE of the least squares estimator and the Viterbi and Viterbi estimator is similar.
+The least squares estimator appears slightly more accurate at low SNR.
+
+Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4} show the variance of the amplitude estimator $\hat{\rho}$.
+The solid line is the asymptotic variance predicted by Theorem~\ref{thm:normality}.
 The dots are again the Monte-Carlo simulations, and the dashed line is the unmodulated carrier.
-The solid line is the asymptotic MSE predicted by Corollary~\ref{cor:ampmse}.
-The corollary accurately models the behaviour of $\hat{\rho}$ when $L$ is sufficiently large.
-The estimator is biased, and as $L\rightarrow\infty$, the MSE converges to $\rho_0^2(1 - G(0))^2$.
-The bias is clearly seen in the figures when the SNR is small.
+Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+This requires $G(0)$ to be known.
+In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
+When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
+However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
 
@@
 		\caption{Amplitude error for QPSK}
-		\label{fig:plotampM}
+		\label{fig:plotampM4}
 \end{figure}

### 2012-10-10 other/confnoncoherent/paper.tex
Added to paper, added new data

@@
 
-We assume that time offset estimation and matched filtering have been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
+We assume that time offset estimation and matched filtering have been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is,
 \begin{equation}\label{eq:sigmod}
@@
 For this reason raw symbol error rate is not of interest at this stage.
-Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated recieved symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
-The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
+Instead, we desire an accurate estimator $\hat{a}$ of $a_0$, so that the compensated received symbols $\hat{a}^{-1}y_i$ can be accurately modelled using an additive noise channel.
+The additive noise channel is a common assumption for subsequent receiver operations, such as decoding.
 The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
@@
 \[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{\sqrt{r^2 - 2r\cos\phi + 1}}.
+f(r,\phi) = \frac{r f_Z(\sqrt{\sabs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\sabs{r^2 - 2r\cos\phi + 1}}}.
 \]
@@
 Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
+Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
 Let,
@@
 The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
-For each value of SNR $T = 10000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
+For each value of signal to noise ratio $T = 10000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
 
 Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the mean square error (MSE) of the phase estimator $\hat{\theta}$ computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$.
-The dots are the Monte-Carlo simulations of the least square estimator.
-The dashed line the MSE of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
+The dots show the simulated MSE of the least square estimator.
+The dashed line is the simulated MSE of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
 %The dashed line acts as a lower bound on performance.
@@
 
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the simulated MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
 This estimator requires the selection of a function $F : \reals \mapsto \reals$.
@@
 Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4} show the variance of the amplitude estimator $\hat{\rho}$.
-The solid line is the asymptotic variance predicted by Theorem~\ref{thm:normality}.
-The dots are again the Monte-Carlo simulations, and the dashed line is the unmodulated carrier.
+The solid line is the asymptotic variance predicted by Theorem~\ref{thm:normality}, the dots are the Monte-Carlo simulations with the least squares estimator, and the dashed line the simulations with the unmodulated carrier estimator.
 Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
@@
 We show that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
+However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.
 This bias is large when the signal to noise ratio is small.

### 2012-10-10 paper.tex
Added to paper, added new data

@@
 \title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
-\author{Robby McKilliam, Andre Pollok, Bill Cowley 
+\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, I. Vaughan L. Clarkson and Barry Quinn  
 \thanks{
 Supported under the Australian Government’s Australian Space Research Program.
-Robby McKilliam, Andre Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Robby McKilliam, Andr\'{e} Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
+Vaughan~Clarkson is with the School of Information Technology \& Electrical Engineering, The University of Queensland, QLD., 4072, Australia.
+Barry~Quinn is with the Department of Statistics, Macquarie University, Sydney, NSW, 2109, Australia.
 }}
@@
 where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.
-The sets $P$ and $D$ are disjoint, i.e.
-$P \cap D = \emptyset$, and $L = \abs{P \cup D}$ is the total number of symbols transmitted.
-
-We assume that time offset estimation and matched filtering have been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is then,
+The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$, and we let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
+
+We assume that time offset estimation has been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is,
 \begin{equation}\label{eq:sigmod}
@@
 Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
-Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the reciever and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009}.
+Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the receiver and must also be estimated.
+Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
 A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
-\begin{align*}
+\begin{equation}\label{eq:SSdefn}
+\begin{split}
 SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2  \\
 &= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2,
-\end{align*}
+\end{split}
+\end{equation}
 where $\abs{x}$ denotes the magnitude of the complex number $x$.
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that only a minor modification of Mackenthun's algorithm is required to include pilot symbols.
+Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
 
-In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
+In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.
 The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
@@
 The additive noise channel is a common assumption for subsequent reciever operations, such as decoding.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be mean square error (MSE) between complex amplitude $a_0$ and its estimator $\hat{a}$, that is, our metric is $\expect\abs{a_0 - \hat{a}}^2$ where $\expect$ denotes the expected value.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $\expect(\rho_0 - \hat{\rho})^2$.
+The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
+Consequently, our metric of performance will not be symbol or bit error rate, it will be $\abs{a_0 - \hat{a}}^2$.
+It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\expect\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $\expect(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
-
-The paper is organised as follows BLERG.
+It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+
+The paper is organised in the following way.
+Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when both pilot symbols and data symbols exist.
+Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
+Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
+We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
+However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
+The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
+Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.
+In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
+In this case, our expressions for the asymptotic distribution can be simplified.
+Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.
+These simulations agree with the derived asymptotic properties.
 
@@
 
-\section{Mackentun's algorithm with pilots}\label{sec:least-squar-estim}
+\section{Mackenthun's algorithm with pilots}\label{sec:least-squar-estim}
 
 In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, but only minor modifications are required to include the pilot symbols.
+Mackenthun specifically considered the noncoherent setting, so we modify the algorithm to include the pilot symbols.
 For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
-A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
-That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is indepedent of $Z$.
+A complex random variable $W$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{W}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{W}$.
+That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
 If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
@@
 \]
-Observe that for any real number $\phi$, the distribution of $X$ and $e^{j\phi}X$ are the same, that is, the distribution is invariant to phase rotation.
-If $\expect\abs{X} = \expect Z$ is finite, then $X$ has zero mean because
-\begin{align*}
-\expect X &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
+Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
+If $\expect\abs{W} = \expect Z$ is finite, then $W$ has zero mean because
+\begin{align*}
+ \expect W &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
  &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
@@
  \end{align*}
-
-We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
-\[
-R e^{j\Phi} = 1 + X.
+If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
+\[
+f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
+\]
+
+We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
+\[
+R e^{j\Phi} = 1 + W.
 \]
 The joint pdf of $R$ and $\Phi$ can be shown to be
-\[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
-\]
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
+\begin{equation}\label{eq:pdfRPhi}
+f(r,\phi) = \frac{r f_Z(\sqrt{\abs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\abs{r^2 - 2r\cos\phi + 1}}}.
+\end{equation}
+The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
 \begin{equation}\label{eq:expectRepartRphi}
@@
 \end{equation}
-where $\Im(\cdot)$ denotes the imaginary part.  %The next lemma will be useful for the analysis of the least squares estimator.
-
-\begin{lemma}\label{lem:gg2cont}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with continuous pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$ and with joint pdf $f(r,\phi)$.
-Then
-\[
-g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
-\]
-and
-\[
-g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr.
-\]
-are continuous functions from $\reals$ to $\reals$.
-\end{lemma}
-\begin{IEEEproof}
-We only give a proof for $g_2$, the proof for $g$ being similar.  From 
-\end{IEEEproof}
+where $\Im(\cdot)$ denotes the imaginary part.
 
@@
 \begin{theorem}\label{thm:normality} (Asymptotic normality)
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, and let
-\[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
-\]
+Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, let
+\[
+g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr,
+\]
+and assume that $\frac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\frac{\abs{D}}{L} = d + o(L^{-1/2})$.
 Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
@@
 \[
-A_1 = \var\big( R_1^2\sin^2(\Phi_1) \big), \qquad A_2 = \var\big( R_1^2\sin^2\fracpart{\Phi_1} \big),
-\]
-\[
-B_1 = \var\big( R_1^2 \cos^2(\Phi_1) \big), \;\;\; B_2 = \var\big( R_1^2 \cos^2\fracpart{\Phi_1} \big).
+A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
+\]
+\[
+B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
 \]
@@
 
+For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.
+The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.
+At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.
+
+\begin{proposition}\label{prop:gg2cont}
+Let $W$ be a circularly symmetric complex random variable with $\expect\abs{W}^2$ finite.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
+Let $f(r,\phi)$ be the joint pdf of $R$ and $\Phi$.
+Then
+\[
+g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
+\]
+and
+\[
+g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
+\]
+are even functions on $(-\pi,\pi)$ and are continuous at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$.
+\end{proposition}
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
+
 A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also nescessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-The next Lemma describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
-
-\begin{lemma}
-Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$ such that $z^{-1} f_Z(z)$ is nonincreasing with $z$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+The next proposition describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
+
+\begin{proposition}\label{prop:contgg2}
+Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
+Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
 Let,
 \[
-h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{ x + \Phi}.
+h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
 \]
 Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-\end{lemma}
-Before giving the proof we discuss the requirements of this lemma.
-The requirement that $z^{-1}f_{Z}(z)$ be non increasing implies that the pdf of $X$ decreases as we move away from the origin.
-That is, the pdf of $X$ in rectangular coordinates is $p(x,y) = z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$, and this is required to be non increasing with $z$.
-For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
+\end{proposition}
 \begin{IEEEproof}
@@
 \end{IEEEproof}
+
+Theorem~\ref{thm:consistency} defines real numbers $p$ and $d$ to represent the proportion of pilot symbols and data symbols in the limit as $L$ goes to infinity.
+For Theorem~\ref{thm:normality} we need the slightly stronger condition that
+\[
+\frac{\abs{P}}{L} = p + o(L^{-1/2}) \qquad \text{and} \qquad \frac{\abs{D}}{L} = d + o(L^{-1/2})
+\] 
+where $o(L^{-1/2})$ denotes a number such that $\sqrt{L} o(L^{-1/2})$ goes to zero as $L\rightarrow \infty$.
+This stronger condition is required to prove the asymptotic normality of $\sqrt{L}\hat{m}_L$.
+The next two sections give proofs for Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+The proofs make use of various lemmas.
+The lemmas are proved in the appendix.
+
+% \begin{IEEEproof}
+% We only give a proof for $g_2$, the proof for $g$ begin similar.  From~\eqref{eq:pdfRPhi},
+% \[
+% g_2(\phi) = \int_{0}^\infty \frac{r^3 f_Z(\sqrt{\sabs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\sabs{r^2 - 2r\cos\phi + 1}}} dr.
+% \]
+% That $g_2$ is even on $[-\pi,\pi]$ is now obvious because $\cos\phi$ is even on $[-\pi,\pi]$.  Put $a = \cos\phi$ and $z^2 = r^2 - 2ra + 1$ and by a change of variable from $r$ to $z$,
+% \[
+% g_2(\phi) = \int_{1}^{\sqrt{1-a^2}} \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz + \int_{\sqrt{1-a^2}}^\infty \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz.
+% \]
+% when $0 < a \leq 1$ and
+% \[
+% g_2(\phi) = \int_{1}^\infty \frac{r^3 f_Z(z)}{2\pi ( r - a)} dz.
+% \]
+% when $-1 \leq a \leq 0$.  Because
+% \[
+% \frac{r^3}{(r - a)} \leq a z^2 +  b
+% \]
+% for all $\phi \in \reals$, it follows that
+% \[
+% g_2(\phi) \leq \int_{0}^\infty \frac{(a z^2 + d) f_Z(z)}{2\pi} dz = a \expect Z^2 + b
+% \]
+% which is finite.  BLERG work out a and b.
+% \end{IEEEproof}
 
@@
 \end{equation}  
-Lemma~\ref{lem:GLtoG0} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+Lemma~\ref{lem:GLtoG0} in the appendix shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
+These are proved in Section~\ref{sec:proof-almost-sureappendix} of the appendix.
+
+\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
+
+We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
+Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
+Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
+The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
+This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
+Define the function
+\begin{equation}\label{eq:RLdef}
+R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
+\end{equation}
+Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
+Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
+\begin{align*}
+0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
+&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
+\end{align*}
+Lemma~\ref{lem:Qconv} shows that
+\[
+\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
+\]
+where $o_P(1)$ denotes a number converging in probability to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+Lemma~\ref{lem:empiricprocc} shows that
+\[
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
+\]
+It follows from the three equations above that,
+\[
+0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
+\]
+and rearranging gives,
+\[
+\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
+\]
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Theorem~\ref{thm:normality}.
+It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
+\[
+\frac{pA_1 + dA_2}{(p + Hd)^2}.
+\]
+
+We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
+\begin{align*}
+\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
+&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
+\end{align*}
+Lemma~\ref{lem:empiricprocforrho} shows that 
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
+\]
+where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
+\[
+%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
+\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
+\]
+It follows that $\sqrt{L} m_L =  \rho_0 X_L + o_P(1)$.
+Lemma~\ref{lem:XL} shows that the distribution of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
+Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
+Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
+\[
+\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
+\]
+as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of the appendix.
+
+\section{The Gaussian noise case}\label{sec:gaussian-noise-case}
+
+Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
+\[
+\frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
+\]
+Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+%and the joint density function of the real and imaginary parts of $Z$ is
+%\[
+%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
+%\] 
+%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
+It can be shown that
+\[
+g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+\]
+where $\kappa = \rho_0/\sigma$ and
+\[
+\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
+\]
+The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
+
+\section{Simulations}\label{sec:simulations}
+
+We present the results of Monte-Carlo simulations with the least squares estimator.
+In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+Under these conditions the least squares estimator is also the maximum likelihood estimator.
+Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
+The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
+For each value of SNR, $T = 5000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
+
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.
+When $\abs{P} \neq 0$ (i.e.
+coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
+Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
+The solid lines are the MSEs predicted by Theorem~\ref{thm:normality}.
+The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
+As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+Theorem~\ref{thm:normality} does not model this behaviour.
+
+Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
+The dots and crosses show the results of Monte-Carlo simulations.
+Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+This requires $G(0)$ to be known.
+In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
+When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
+However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+
+Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
+The figure depicts an interesting phenomenon.
+When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.
+When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
+A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
+%When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
+This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
+For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
+\[
+SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
+\]
+where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
+For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm}.
+This would enable a rigorous theory for selection of $\beta$ at the receiver.
+One caveat is that the receiver would require knowledge about the noise distribution in order to advantageously choose $\beta$.
+We do not investigate this further here.
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
+		\caption{Phase error versus SNR for BPSK with $L=4096$.}
+		\label{fig:plotphaseBPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
+		\caption{Phase error versus SNR for QPSK with $L=4096$.}
+		\label{fig:plotphaseQPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
+		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
+		\label{fig:plotphase8PSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
+		\caption{Unbiased amplitude error versus SNR for BPSK.}
+		\label{fig:plotampBPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+		\caption{Unbiased amplitude error versus SNR for QPSK.}
+		\label{fig:plotampQPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+		\caption{Unbiased amplitude error versus SNR for $8$-PSK.}
+		\label{fig:plotamp8PSK}
+\end{figure}
+
+%\begin{figure}[tp]
+%	\centering
+%		\includegraphics[width=\linewidth]{code/data/plotM2-3.mps}
+%		\caption{Phase error versus SNR for BPSK.}
+%		\label{fig:plotphase}
+%\end{figure}
+
+\begin{figure}[tp]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
+		\caption{Phase error versus SNR for QPSK.}
+		\label{fig:plotphaseQPSKmultL}
+\end{figure}
+
+\section{Conclusion}
+
+We considered least squares estimators of carrier phase and amplitude from noisy communications signals that contain both pilot signals, known to the receiver, and data signals, unknown to the receiver.
+We focused on $M$-ary phase shift keying constellations.
+The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
+
+We showed, under some conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
+This bias is large when the signal to noise ratio is small.
+It would be interesting to investigate methods for correcting this bias.
+A method for estimating $G(0)$ at the receiver appears to be required.
+
+Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
+Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
+This suggests the use of a weighed objective function, which would be an interesting topic for future research.
+
+\small
+\bibliography{bib}
+
+\appendix
+
+\subsection{Lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sureappendix}
 
@@
 
-\section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
-
-We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
-Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
-Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
-The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
-However, we show in Lemma~\ref{lem:diffatlambdaL} that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
-Define the function
-\begin{equation}\label{eq:RLdef}
-R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\end{equation}
-Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
-Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
-\begin{align*}
-0 &= R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) + Q_L(\hat{\lambda}_L) \\
-&= \sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) + \sqrt{L}Q_L(\hat{\lambda}_L).
-\end{align*}
-Lemma~\ref{lem:Qconv} shows that
-\[
-\sqrt{L} Q_L(\hat{\lambda}_L) = \sqrt{L} \hat{\lambda}_L\big( p + Hd  + o_P(1) \big)
-\]
-where $o_P(1)$ denotes a number converging in probablity to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
-Lemma~\ref{lem:empiricprocc} shows that
-\[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
-\]
-It follows from the three equations above that,
-\[
-0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
-\]
-and rearranging gives,
-\[
-\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
-\]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Thereom~\ref{thm:normality}.
-It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
-\[
-\frac{pA_1 + dA_2}{(p + Hd)^2}.
-\]
-
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
-\begin{align*}
-\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
-&= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
-\end{align*}
-Lemma~\ref{lem:empiricprocforrho} shows that 
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
-\]
-where $X_L = \sqrt{L}\big( G_L(0) - T_L(0)  \big)$.  Lemma~\ref{lem:HLtoG} shows that
-\[
-%\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = \sqrt{L} \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L}\big( T_L(\hat{\lambda}_L) - G(0) \big) = o_P(1).
-\]
-It follows that
-\[
-%\sqrt{L} m_L = o_P(1) + X_L  + \hat{\lambda}_L \big( ? + o_P(1) \big).
-\sqrt{L} m_L =  \rho_0 X_L + o_P(1). 
-\]
-Lemma~\ref{lem:XL} shows that the distrubtion of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
-Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
-Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
-\[
-\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
-\]
-as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.
+\subsection{Lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-normappendix}
 
@@
 \end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous, $c_k(\lambda)$ converges to 
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}, it follows that $c_k(\lambda)$ converges to
 \[
@@
 as $\lambda$ goes to zero from above.
-We are only interested in the limit from above because we are working under the assumption that that $0 \leq \lambda < \frac{\pi}{M}$.
-The analgous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
-We actually only require $g$ to be continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$, and the requirements of Theorem~\ref{thm:normality} could be weakened in this sense.
-Here, the continuity of $g$ is garaunteed by the assumption that the noise $w_1$ is circularly symetric.
+We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
+The analogous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
 
@@
 \]
-Similarly, the distribution of 
-\[ 
-\frac{1}{\sqrt{L}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} = \sqrt{\frac{\sabs{D}}{L}} \frac{1}{\sqrt{\sabs{D}}} \sum_{i \in D} R_i \sin\fracpart{\Phi_i} 
-\]
-converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~? BLERG, and variance
+Similarly, the distribution of $D_L$ converges to the normal with mean $\sqrt{d}\expect R_1 \sin\fracpart{\Phi_1} = 0$ as a result of Lemma~\ref{lem:expectImRfracpart}, and variance
 \[
 d A_2 = d \expect R_1^2 \sin^2\fracpart{\Phi_1}.
+\]
+The lemma holds because $C_L$ and $D_L$ are independent. 
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:expectImRfracpart}
+$\expect R_1 \sin\fracpart{\Phi_1} = 0$.
+\end{lemma}
+\begin{IEEEproof}
+Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
+\begin{align*}
+\expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
+&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi,
+\end{align*}
+where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
+Proposition~\ref{prop:gg2cont} shows that $g$ is even on $(-\pi,\pi)$.
+Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.
+So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
+\[
+g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
+\]
+and
+\[
+\sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
+\]
+Now,
+\begin{align*}
+I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
+&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
+&\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+&= 0
+\end{align*}
+for all integers $k$ because
+\begin{align*}
+\int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
+&= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
+\end{align*}
+Observe that the sets
+\[
+ [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
+\]
+partition $[-\pi, \pi]$, and so
+\[
+\expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
 \]
@@
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\lambda) + X_L
-\]
-Lemma~\ref{lem:ZYLempiricproc} in the appendix 
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\hat{\lambda}_L) + X_L
+\]
+Lemma~\ref{lem:ZYLempiricproc} in Section~\ref{sec:an-empirical-process} shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+\[
+\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu.
+\]
+The proof now follows by an argument analogous to that in Lemma~\ref{lem:empiricprocc}.
 \end{IEEEproof}
@@
 \]
-since $h_1(0) = 1$.  The distribution of $D_L$ converges to the normal with zero mean and variance
-\[
-d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0)
-\]
-The proof follows since $C_L$ and $D_L$ are indpendent.
-\end{IEEEproof}
-
-\section{The Gaussian noise case}
-
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.
-The joint density function of the real and imaginary parts is then
-\[
-p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
-\]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
-%and the joint density function of the real and imaginary parts of $Z$ is
-%\[
-%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
-%\] 
-%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
-It can be shown that
-\[
-g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-\]
-where $\kappa = \rho_0/\sigma$ and
-\[
-\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
-\]
-The value of $A_1$ and $A_2$ can be computed by numerical integration.
-
-\section{Simulations}\label{sec:simulations}
-
-%\subsection{The coherent setting}
-
-Our first set of simulations considers the case where there are a number of pilot symbols.
-
-%BPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-%		\caption{Amplitude error for BPSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error for BPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure*}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure*}
-
-%QPSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-%		\caption{Amplitude error for QPSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error for QPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM4-3.mps}
-% 		\caption{complex amplitude error for QPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%8PSK plots
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-%		\caption{Amplitude error for 8PSK}
-%		\label{fig:plotamp}
-%\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error for 8PSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%\subsection{The noncoherent setting}
-
-%BPSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-% 		\caption{Amplitude error for BPSK}
-% 		\label{fig:plotamp}
- %\end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
-		\caption{Phase error for BPSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM2-3.mps}
-% 		\caption{complex amplitude error for BPSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-%QPSK plots
- %\begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-% 		\caption{Amplitude error for QPSK}
-% 		\label{fig:plotamp}
-% \end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
-		\caption{Phase error for QPSK}
-		\label{fig:plotphase}
-\end{figure}
-
- % \begin{figure}[tp]
- % 	\centering
- %         \includegraphics{code/data/plotM4-3.mps}
- % 		\caption{complex amplitude error for QPSK}
- % 		\label{fig:plotcamp}
- % \end{figure}
-
-%8PSK plots
-% \begin{figure}[tp]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-% 		\caption{Amplitude error for 8PSK}
-% 		\label{fig:plotamp}
-% \end{figure}
-
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM8-2.mps}
-		\caption{Phase error for 8PSK}
-		\label{fig:plotphase}
-\end{figure}
-
-% \begin{figure}[tp]
-% 	\centering
-%         \includegraphics{code/data/plotM8-3.mps}
-% 		\caption{complex amplitude error for 8PSK}
-% 		\label{fig:plotcamp}
-% \end{figure}
-
-\section{Conclusion}
-
-\small
-\bibliography{bib}
-
-\normalsize
-\appendix
+since $h_1(0) = 1$.  Similarly the distribution of $D_L$ converges to the normal with zero mean and variance
+\[
+d B_2 = d \expect R_1^2 \cos^2\fracpart{\Phi_1} - d h_2^2(0).
+\]
+The lemma follows since $C_L$ and $D_L$ are independent.
+\end{IEEEproof}
 
@@
 
-\subsection{An empirical process result}
+\subsection{An empirical process result}\label{sec:an-empirical-process}
 
@@
 \begin{IEEEproof}
-Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be sequence identically distributed to $\{\Phi_i\}$.
+Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be a sequence identically distributed to $\{\Phi_i\}$.
 Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
 Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
-Then
+Now,
 \begin{align*}
@@
 \]
-is continuous due to Lemma~\ref{lem:gg2cont}.  From the definition of $I_\epsilon$,
+is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}.  From the definition of $I_\epsilon$,
 \begin{align*}
@@
 \[
-K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}
-\]
-\end{IEEEproof}
-
-\begin{lemma}
+K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
+\]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:epslmlemma}
 Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.
@@
 \begin{IEEEproof}
-BLERG
+We prove the lemma for $d_2$.  The proof for $d_1$ is similar.  Observe that $d_2(\lambda, 0) = c_1 + c_2$ where 
+ \[
+ c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) },
+ \] 
+ \[
+ c_2 = \sum_{i \in D} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+ \]
+ As $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+ \[
+ \sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i)} \leq K\sabs{\lambda} = \sabs{\lambda} < \epsilon
+ \]
+ for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let 
+\[
+S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}.
+\] 
+Now $c_2 = c_3 + c_4$ where 
+ \[
+ c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}},
+ \]
+ \[
+ c_4 =  \sum_{i \in D, i \notin S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+ \]
+Observe that $I_\epsilon(\Phi_i) = 1$ if and only if $\abs{\fracpart{\Phi_i}} \geq \tfrac{\pi}{M} - \epsilon$.  So, if $i \notin S$ and $\sabs{\lambda} < \epsilon$, then
+\begin{align*}
+\round{\lambda + \Phi_i} &= \round{\lambda + \fracpart{\Phi_i} + \round{\Phi_i}} \\ 
+&= \round{\lambda + \fracpart{\Phi_i}} + \round{\Phi_i} = \round{\Phi_i}
+\end{align*}
+because
+\[
+\sabs{\lambda + \fracpart{\Phi_i}} < \epsilon + \tfrac{\pi}{M} - \epsilon = \tfrac{\pi}{M}
+\] 
+and therefore $\round{\lambda + \fracpart{\Phi_i}} = 0$.  Now,
+ \begin{align*}
+ \vert\sin&\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}\vert \\
+ &= \sabs{\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\Phi_i - \round{\Phi_i})} \\
+ &\leq \sabs{\lambda} < \epsilon
+ \end{align*}
+and so, 
+\[
+c_4 \leq \epsilon \sum_{i \in D, i \notin S} R_i \leq \epsilon \sum_{i \in D} R_i.
+\]
+Otherwise, if $i \in S$, then $\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}} \leq 2$ because $\sin(\cdot) \leq 1$.  Thus 
+\[
+c_3 \leq 2 \sum_{i \in D, i \in S} R_i \leq 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i).
+\]
+Now,
+\begin{align*}
+d_2(\lambda, 0) &= c_1 + c_2 = c_1 + c_3 + c_4 \\
+&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i) = D_\epsilon L
+\end{align*}
+as required
 \end{IEEEproof}
@@
 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda) - X(b_2(\lambda))} \nonumber \\
-&\leq K_4 2^{-k} D_\epsilon L, \label{eq:Deps2nktozero}
-\end{align}
-the second line because,
+&\leq K_4 2^{-k} D_\epsilon L. \label{eq:Deps2nktozero}
+\end{align*}
+The second line in the inequality above follows because,
 \[
@@
 \begin{IEEEproof}
-Let $b_1(\lambda)$ maps each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
-The existance of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
+Let $b_1(\lambda)$ map each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
+The existence of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
 By the triangle inequality
@@
 \begin{lemma}\label{lem:maxineq}(Maximal inequalities)
-Suppose $X_1, \dots, X_N$ are $N$ zero mean Gaussian random variables each with variance less than some positive constant $K$.  Then
+Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
 \[

### 2012-10-10 paper.tex
Added new data

@@
 At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.
+The proof is given in Appendix~\ref{appendix:prop1proof}.
 
@@
 \end{proposition}
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
 
@@
 The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
+The proof is given in Appendix~\ref{appendix:prop2proof}.
 
@@
 The solid lines are the MSEs predicted by Theorem~\ref{thm:normality}.
-The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
+The theorem accurately predicts the behaviour of the phase estimator when the SNR is sufficiently large.
 As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
@@
 \begin{IEEEproof}
-This result is well known, see for example~\cite[Section 3]{Pollard_asymp_empi_proc_1989}  
+This result is well known, see for example~\cite[Section~3]{Pollard_asymp_empi_proc_1989}  
 \end{IEEEproof}
@@
 
+\subsection{Proof of Proposition~\ref{prop:gg2cont}} \label{appendix:prop1proof}
+
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
+
+\subsection{Proof of Proposition~\ref{prop:contgg2}} \label{appendix:prop2proof}
+
+\begin{IEEEproof}
+BLERG
+\end{IEEEproof}
+
 \end{document}

### 2012-10-10 paper.tex
fixed appendix text size
fixed appendix size
fixed build script
added a bit to appendix

@@
 
+\normalsize
 \appendix
@@
 
-\begin{IEEEproof}
-BLERG
+\begin{lemma}
+Let $W$ be a circularly symmetric random variable with constant magnitude $Z = \abs{W} = c > 0$ and phase uniformly distributed on $[-\pi,pi)$.
+Let $R \geq 0$ and $\Phi \in $ be real random variables such that $Re^{-j2\pi \Phi} = 1 + W$.
+Then the pdf of $\Phi$ is
+\begin{equation} \label{eq:fphicleq1}
+f_{\Phi}(\phi) = \begin{cases} 
+\frac{\cos(\Phi)}{\pi c} &  [-\tfrac{\pi}{2}, \tfrac{\pi}{2}] \\
+0 & \text{otherwise}
+\end{cases}
+\end{equation}
+when $0 < c \leq 1$ and
+\begin{equation} \label{eq:fphicg1}
+f_{\Phi}(\phi) = 
+\end{equation}
+when $c > 1$.
+\end{lemma}
+\begin{IEEEproof}
+We make use of delta `function' notation and write the `pdf' of $Z$ as $\delta(z - c)$.  The pdf of $\Phi$ is then
+\[
+f_{\Phi}(\phi) = \int_{0}^{\infty} r \frac{\delta(z - c)}{ 2\pi z} dr = \sum_{r} \frac{r}{2\pi c}
+\]
+where $z^2 = r^2 - 2r\cos(\phi) + 1$ and the sum is over the positive real solutions of $c^2 = r^2 - 2r\cos(\phi) + 1$.
+If $0 < c \leq 1$ and $phi \in [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are two positive real solutions, these being
+\[
+r = \cos(\phi) \pm \sqrt{c^2 - 1 + \cos^2(\phi)},
+\]
+whilst if $\phi \notin [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are no real solutions.  Thus~\eqref{eq:fphicleq1} holds.
 \end{IEEEproof}
@@
 
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
-
 \end{document}

### 2012-10-11 paper.tex
minor mods.  This needs work, I need to read Barry's unimodal proof.

--- 2012-10-10 paper.tex
@@
 \end{equation}
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.  So,
+Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi]$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$.
+The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.
+So,
 \begin{equation}\label{eq:expectRepartRphi}
@@
 \[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr,
+g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr, \qquad g_2(\phi) = \int_{0}^{\infty} r^2 f(r,\phi) dr,
 \]
@@
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
+If the functions $g$ and $g_2$ are continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$, then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
 \[
@@
 
-For Theorem~\ref{thm:normality} the integral defining $g(\phi)$ must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.
+Theorem~\ref{thm:normality} requires functions $g$ and $g_2$ to be continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$.
+This places mild restrictions on the distribution of the noise $w_i$.
+For example, the requirements are satisfied if the joint pdf of the real and imaginary parts of $w_i$ is continuous, since in this case $f(r,\phi)$ is continuous.
+Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$ it follows that both $g$ and $g_2$ have period $2\pi$ and are even on $[-\pi, \pi]$.
+
+%must exist at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$ so that the constant $H$ exists.
 The next proposition shows that the assumptions made in Theorem~\ref{thm:normality} are strong enough to ensure that the integral does exist at these points.
 At the same time we derive some properties that will be useful during the proof of Theorem~\ref{thm:normality}.
-The proof is given in Appendix~\ref{appendix:prop1proof}.
-
-\begin{proposition}\label{prop:gg2cont}
-Let $W$ be a circularly symmetric complex random variable with $\expect\abs{W}^2$ finite.
+%The proof is given in Appendix~\ref{appendix:prop1proof}.
+
+% \begin{proposition}\label{prop:gg2cont}
+% Let $W$ be a circularly symmetric complex random variable with $\expect\abs{W}^2$ finite.
 Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
@@
 Then
-\[
-g(\phi) = \int_{0}^\infty r f(r, \phi) dr,
-\]
-and
-\[
-g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
-\]
-are even functions on $(-\pi,\pi)$ and are continuous at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$.
-\end{proposition}
+% \[
+% g_1(\phi) = \int_{0}^\infty r f(r, \phi) dr,
+% \]
+% and
+% \[
+% g_2(\phi) = \int_{0}^\infty r^2 f(r, \phi) dr
+% \]
+% are even functions on $(-\pi,\pi)$ and are continuous at $\phi = \tfrac{2\pi}{M}k + \frac{\pi}{M}$ for $k = 0, \dots, M-1$.
+% \end{proposition}
+% \begin{IEEEproof}
+% Let $f_Z(z)$ be the cumulative distribution function of $Z = \abs{W}$.  Now $g_1$ and $g_2$ can be written as
+% \[
+% g_k(\phi) = \int_{0}^{\infty} \frac{\beta(\phi))^{k+1}}{z} f_Z(z) dz
+% \]
+% where $k = 1, 2$ and $\beta(\phi) = \cos(\phi) + \sqrt{z^2 - 1 + \cos^2(\phi)}$.  Since $\expect \abs{Z}^2$ is finite and non-zero $\int_0^\infty z^c dF(z)$ is finite for all $c = -1,0,1,2$.
+% \end{IEEEproof}
 
@@
 \begin{proposition}\label{prop:contgg2}
-Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
+Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable joint pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
 Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
@@
 
-\subsection{Proof of Proposition~\ref{prop:gg2cont}} \label{appendix:prop1proof}
-
-\begin{IEEEproof}
-BLERG
-\end{IEEEproof}
+% \subsection{Proof of Proposition~\ref{prop:gg2cont}} \label{appendix:prop1proof}
+
+% \begin{lemma}
+% Let $W$ be a circularly symmetric random variable with constant magnitude $Z = \abs{W} = c > 0$ and phase uniformly distributed on $[-\pi,\pi)$.
+Let $R \geq 0$ and $\Phi \in [-\pi, \pi)$ be real random variables such that $Re^{j\Phi} = 1 + W$.
+Then the pdf of $\Phi$ is
+% \begin{equation} \label{eq:fphicleq1}
+% f_{\Phi}(\phi) = \begin{cases} 
+% \frac{\cos(\Phi)}{\pi c} &  [-\tfrac{\pi}{2}, \tfrac{\pi}{2}] \\
+% 0 & \text{otherwise}
+% \end{cases}
+% \end{equation}
+% when $0 < c \leq 1$, and
+% \begin{equation} \label{eq:fphicg1}
+% f_{\Phi}(\phi) = \frac{\cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}}{2\pi c}
+% \end{equation}
+% when $c > 1$.
+% \end{lemma}
+% \begin{IEEEproof}
+% We make use of delta `function' notation and write the `pdf' of $Z$ as $\delta(z - c)$.  Using~\eqref{eq:pdfRPhi} the pdf of $\Phi$ is
+% \[
+% f_{\Phi}(\phi) = \int_{0}^{\infty} r \frac{\delta(z - c)}{ 2\pi z} dr = \sum_{r} \frac{r}{2\pi c}
+% \]
+% where $z^2 = r^2 - 2r\cos(\phi) + 1$ and the sum is over the positive real solutions of $c^2 = r^2 - 2r\cos(\phi) + 1$.
+If $0 < c \leq 1$ and $\phi \in [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are two positive real solutions, these being
+% \[
+% r = \cos(\phi) \pm \sqrt{c^2 - 1 + \cos^2(\phi)},
+% \]
+% whilst if $\phi \notin [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are no real solutions.  Thus~\eqref{eq:fphicleq1} holds.  If $c > 1$ there is a single real solution
+% \[
+% r = \cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}
+% \]
+% so that~\eqref{eq:fphicg1} holds.
+% \end{IEEEproof}
+
+% Now if $W$ is a circularly symmetric complex random variable with $Z = \abs{W}$ having pdf $f_Z$ and let $R \geq 0$ and $\Phi \in [-\pi, \pi)$ be real random variables such that $Re^{j\Phi} = 1 + W$.
+The pdf of $\Phi$ is
+% \[
+% f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{f_Z(z)}{ 2\pi z} dr.
+% \]
+% Let
+% \[
+% I_{N,c}(z) = \begin{cases}
+% N \frac{f_Z(z)}{ 2\pi z} & z \in [cN, c(N-1)) \\
+% 0 & \text{otherwise}.
+% \end{cases}
+% \]
+% Now
+% \[
+% \frac{f_Z(z)}{ 2\pi z} = \frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z)
+% \]
+% so that
+% \[
+% f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z) dr = \frac{1}{N} \sum_{c = 0}^{\infty} \int_{0}^{\infty} r I_{N,c}(z) dr 
+% \]
 

### 2012-10-16 paper.tex
fixed silly abs error
minor paper edits.

@@
 $z_i = \phi -  u $ \;
-$g_i = y_i e^{-j u}$ \;
+$g_i = y_i e^{-j u}$ \nllabel{alg_gset} \;
 }
@@
 $Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
-$\hat{a} = \frac{1}{L} Y$ \;
+$\hat{a} = \frac{1}{L} Y$ \nllabel{alg_hata1} \;
 $\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
@@
  	$\hat{Q} = Q$ \;
- 	$\hat{a} =  \frac{1}{L} Y$ \;
+ 	$\hat{a} =  \frac{1}{L} Y$ \nllabel{alg_hata2} \;
  }
@@
 \begin{equation}\label{eq:pdfRPhi}
-f(r,\phi) = \frac{r f_Z(\sqrt{\abs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\abs{r^2 - 2r\cos\phi + 1}}}.
+f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
 \end{equation}
@@
 \]
-If $p > 0$ and if $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
+If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
 \begin{enumerate}
@@
 
-A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption asserts that $\abs{G(x)} \leq \abs{G(0)}$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $\abs{G(x_i)} \rightarrow \sabs{G(0)}$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $\abs{G(x)}$ is uniquely maximised at some point $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $\abs{G(x)}$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-The next proposition describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
+This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
 The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
@@
 \]
-Then $\abs{h_1(x)}$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $\abs{h_2(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
+Then $h_1(x)$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $h_2(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
 \end{proposition}
@@
 and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$ conditioned on minimisation with respect to the data symbols,
 \begin{equation}\label{eq:hatrhoZ}
-\hat{\rho} = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
-\end{equation}
-where $\Re(\cdot)$ denotes the real part.  Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
+\hat{\rho}(\theta) = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
+\end{equation}
+where $\Re(\cdot)$ denotes the real part.
+Since $\hat{\rho}(\theta)$ is nonnegative it follows that $\Re(Z(\theta))$ is nonnegative for all $\theta \in [-\pi, \pi)$.
+Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
 \[
@@
 
-We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\abs{\Re(Z(\theta))}$.
+We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\Re(Z(\theta))$.
 Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
@@
 and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
-Since $\hat{\theta}$ is the maximiser of $\abs{\Re(Z(\theta))}$ then  $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(\lambda)}$.
+Since $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ then  $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)$.
 We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
@@
 As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ upon which $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ as $L\rightarrow\infty$.
+Let $A$ be the subset of the sample space $\Omega$ upon which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
 Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
 Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
-Because $\sabs{G(x)}$ is uniquely maximised at $x=0$, it follows that $\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
+Because $G(x)$ is uniquely maximised at $x=0$, it follows that $G(\hat{\lambda}_L) \rightarrow G(0)$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
 So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
@@
 \end{equation}
-Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
+Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$, and so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
 Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
@@
 Under these conditions the least squares estimator is also the maximum likelihood estimator.
-Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
+Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.
 The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
@@
 where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
+Computing the $\hat{a}$ that minimises $SS_\beta$ can be achieved with only a minor modification to algorithm~\ref{alg:loglinear}.
+Line~\ref{alg_gset} is modified to $g_i = \beta y_i e^{-j u}$ and lines~\ref{alg_hata1}~and~\ref{alg_hata2} are modified to $\hat{a} =  \frac{1}{\abs{P} + \beta\abs{D}} Y$.
 For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm}.

### 2012-10-16 paper.tex
mods to paper.

@@
 \begin{lemma}\label{lem:convtoexpGlamL} 
-$\sabs{G(\hat{\lambda}_L)} \rightarrow \sabs{G(0)}$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Since $\sabs{G(x)}$ is uniquely maximised at $x=0$,
-\[
-0 \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)},
-\]
-and since $\hat{\lambda}_L$ is the maximiser of $\abs{G_L(x)}$,
-\[ 
-0 \leq \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}.
+$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Since $G(x)$ is uniquely maximised at $x=0$,
+\[
+0 \leq G(0) - G(\hat{\lambda}_L),
+\]
+and since $\hat{\lambda}_L$ is the maximiser of $G_L(x)$,
+\[ 
+0 \leq G_L(\hat{\lambda}_L) - G_L(0).
 \]
@@
 \begin{align*}
-0 &\leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} \\ 
-& \leq \sabs{G(0)} - \sabs{G(\hat{\lambda}_L)} + \sabs{G_L(\hat{\lambda}_L)} - \sabs{G_L(0)}\\
+0 &\leq G(0) - G(\hat{\lambda}_L) \\ 
+& \leq G(0) - G(\hat{\lambda}_L) + G_L(\hat{\lambda}_L) - G_L(0) \\
 & \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
@@
 \]
-Lemma~\ref{lem:supDLlambdan} shows that for any finite $N$, and any $\epsilon > 0$, 
+Lemma~\ref{lem:supDLlambdan} shows that for any finite constant $N$, and any $\epsilon > 0$, 
 \[
@@
 \]
-i.e., $U_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
+that is, $U_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
 \[
@@
 \begin{lemma}\label{lem:supDLlambdan}
-If $N$ is finite then $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+With $N$ a finite constant $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
 \end{lemma}
@@
 \]
-almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Now, since $N$ is finite,
+almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Since $N$ is a constant,
 \[

### 2012-10-16 paper.tex
minor

@@
 \]
-almost surely as $L\rightarrow\infty$ by the Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Since $N$ is a constant,
+almost surely as $L\rightarrow\infty$ by Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Since $N$ is a constant,
 \[
@@
 \]
-for all $i \in D$.  Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.
-
-It remains to show that $G_L^\prime(\hat{\lambda}_L) = 0$.
-If $\sabs{G_L(\hat{\lambda}_L)} > 0$ then $\sabs{G_L(\lambda)}$ is differentiable at $\hat{\lambda}_L$ and, since $\hat{\lambda}_L$ is a maximiser of $\sabs{G_L(\lambda)}$,
-\[
-0 = \frac{d \sabs{G_L(\lambda)}}{d \lambda}(\hat{\lambda}_L) = \begin{cases} 
-G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) > 0\\
--G_L^\prime(\hat{\lambda}_L), & G_L(\hat{\lambda}_L) < 0.
-\end{cases}
-\]
-Thus, $G_L^\prime(\hat{\lambda}_L) = 0$ if $\sabs{G_L(\hat{\lambda}_L)} > 0$.
-On the other hand $\sabs{G_L(\hat{\lambda}_L)} = 0$ only if $G_L(\lambda) = 0$ for all $\lambda \in \reals$, and in this case $G_L^\prime(\hat{\lambda}_L) = 0$ also.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i, i \in D$ be defined as in Lemma~\ref{lem:diffatlambdaL}.
+for all $i \in D$.
+Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.
+That $G_L^\prime(\hat{\lambda}_L) = 0$ follows since $\hat{\lambda}_L$ is a maximiser of $G_L$.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:fracpartlambdahatnotpi} Let $q_i$ be defined as in Lemma~\ref{lem:diffatlambdaL}.
 Then $\sabs{\sfracpart{\hat{\lambda}_L + \Phi_i}} \leq \frac{\pi}{M} + q_i$ for all $i \in D$.

### 2012-10-16 other/confnoncoherent/paper.tex
fixed silly abs error
fixed abs errors

@@
 \[
-f(r,\phi) = \frac{r f_Z(\sqrt{\sabs{r^2 - 2r\cos\phi + 1}})}{2\pi\sqrt{\sabs{r^2 - 2r\cos\phi + 1}}}.
+f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
 \]
@@
 \begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $\expect \abs{w_1}^2$ finite, and let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod}.
-Let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
+Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $w_1$ having finite variance and continuous pdf.
+Let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod} and let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
 Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
@@
 \] 
-If $\abs{G(x)}$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
+If $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
 \begin{enumerate}
@@
 
-A key assumption in Theorem~\ref{thm:consistency} is that $\abs{G(x)}$ is uniquely maximised at $x = 0$.
-The next proposition describes a class of circularly symmetric distributions for which $\abs{G(x)}$ is uniquely maximised at $x = 0$.
+A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$.
+The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
 The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
@@
 \]
-Then $\abs{G(x)}$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
+Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
 \end{proposition}
@@
 
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the mean square error (MSE) of the phase estimator $\hat{\theta}$ computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$.
-The dots show the simulated MSE of the least square estimator.
-The dashed line is the simulated MSE of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the sample mean square error (MSE) of the phase estimator $\hat{\theta}$ computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$.
+The dots show the sample MSE of the least square estimator.
+The dashed line is the sample MSE of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
 %The dashed line acts as a lower bound on performance.
@@
 
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the simulated MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the sample MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
 This estimator requires the selection of a function $F : \reals \mapsto \reals$.
 We have chosen $F(x) = 1$.
-The MSE of the least squares estimator and the Viterbi and Viterbi estimator is similar.
+The sample MSE of the least squares estimator and the Viterbi and Viterbi estimator is similar.
 The least squares estimator appears slightly more accurate at low SNR.

### 2012-10-23 other/confnoncoherent/paper.tex
modifed conference paper to fit with ICASSP template

@@
 %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-\documentclass[conference]{IEEEtran}
+%\documentclass[conference]{IEEEtran}
+
+\documentclass{article}
+
+\usepackage{spconf}
 
 \usepackage{mathbf-abbrevs}
+
 \input{defs}
@@
 
+%NOTE: If you use the natbib package, you should use natbibspacing.sty instead, and be sure to load it after natbib:
 \usepackage[square,comma,numbers,sort&compress]{natbib}
+\usepackage{natbibspacing}
+\setlength{\bibspacing}{0cm}
 
@@
 
-%opening
+% %opening
+% \title{Noncoherent least squares estimators of carrier phase and amplitude}
+% \author{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \vspace{0.2cm} \\
+% \small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.1cm} \\
+% \small $^{2}$School of Information Technology \& Electrical Engineering, The University of Queensland, QLD 4067, \textsc{Australia} \vspace{0.1cm} \\
+% \small $^{3}$Department of Statistics, Macquarie University, Sydney, \textsc{Australia} 
+% \thanks{Supported under the Australian Government’s Australian Space Research Program.}
+% }
+
  \title{Noncoherent least squares estimators of carrier phase and amplitude}
-\author{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \vspace{0.2cm} \\
-\small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.1cm} \\
-\small $^{2}$School of Information Technology \& Electrical Engineering, The University of Queensland, QLD 4067, \textsc{Australia} \vspace{0.1cm} \\
+\name{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \thanks{Supported under the Australian Government’s Australian Space Research Program.}}
+\address{ \small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.05cm} \\
+\small $^{2}$School of Information Technology \& Electrical Engineering, The University of Queensland, QLD 4067, \textsc{Australia} \vspace{0.05cm} \\
 \small $^{3}$Department of Statistics, Macquarie University, Sydney, \textsc{Australia} 
-\thanks{Supported under the Australian Government’s Australian Space Research Program.}
 }
 
-\IEEEoverridecommandlockouts
+%\IEEEoverridecommandlockouts
 \begin{document}
@@
 \end{abstract}
-\begin{IEEEkeywords}
+\begin{keywords}
 Noncoherent detection, phase shift keying, asymptotic statistics
-\end{IEEEkeywords}
+\end{keywords}
 
@@
 \]
+
 % The mean of $R e^{j\Phi}$ is equal to one because the mean of $X$ is zero.  So,
@@
 
-\section{Conclusion}
+\section{RELATION TO PRIOR WORK}
 
@@
 It would be interesting to investigate methods for correcting this bias.
-
-\small
-\bibliography{bib}
 
@@
 
+\small
+\bibliography{bib}
+
 %\begin{figure}[tp]

### 2012-10-23 other/confnoncoherent/paper.tex
added stuff for silly new icassp fromat

@@
 \usepackage[square,comma,numbers,sort&compress]{natbib}
-\usepackage{natbibspacing}
+\usepackage{bibspacing}
 \setlength{\bibspacing}{0cm}
@@
 
-%Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.  
+Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.  
 A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
@@
 
-We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
-\[
-R e^{j\Phi} = 1 + X.
-\]
-The joint distribution of $R$ and $\Phi$ can be shown to be
-\[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
-\]
+% We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
+% \[
+% R e^{j\Phi} = 1 + X.
+% \]
+% The joint distribution of $R$ and $\Phi$ can be shown to be
+% \[
+% f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
+% \]
 
@@
 
-We have studied the least squares estimator of carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols.
-The estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
-We show that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+The problem of estimating carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols has undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
+The least squares estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and this is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
+The algorithm was later rediscovered by Sweldens~\cite{Sweldens2001}.
+In this paper we described the statistical properties of the least squares estimator.
+We showed that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
 However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.
@@
 
+%\newpage
 \begin{figure}[p]
@@
 
+%\pagebreak
+%\newpage
 \begin{figure}[p]
@@
 
-\small
+%\vfill\pagebreak
+\vfill\pagebreak
+%\newpage
+%\small
 \bibliography{bib}

### 2012-10-23 other/confnoncoherent/paper.tex
minor

@@
 \begin{proposition}\label{prop:contgg2}
-Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
+Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
 Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.

### 2012-10-23 paper.tex
minor

@@
 \end{align}
-where $g$ is defined in the statement of Theorem~\ref{thm:normality}
+where $g$ is defined in the statement of Theorem~\ref{thm:normality}.
 Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
@@
 \begin{lemma}\label{lem:convdistGLdash}
-The distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$.
+The distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ as $L\rightarrow\infty$.
 \end{lemma}
@@
 \] 
-converges to the normal with zero mean and covariance 
-\[
-pB_1 + d B_2
-\] 
-as $L \rightarrow\infty$.
+converges, as $L \rightarrow\infty$,  to the normal with zero mean and covariance $pB_1 + d B_2.$
 \end{lemma}
@@
 
+\begin{IEEEproof}
+We first prove the result for $h_1$.  Write
+\begin{align*}
+h_1(x) &= \int_{-\pi}^{\pi} \int_{0}^\infty r \cos(\phi + x) f(r,\phi) dr d\phi \\
+&= \int_{-\pi}^{\pi}  \cos(\phi + x) g(\phi) d\phi,
+\end{align*}
+where $f(r,\phi)$ is the joint pdf of $R$ and $\Phi$, and 
+\[
+g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
+\]
+Since the pdf of $X$ is circularly symmetric, $f(r,\phi)$ and $g(\phi)$ are even in $\phi$ on $[-\pi,\pi]$.
+
+Write $\cos(\phi + x)$ as a sum of even and odd parts,
+\[
+\cos(\phi + x) = \cos(\phi)\cos(x) - \sin(\phi)\sin(x).
+\]
+Now
+\[
+h_1(x) = \cos(x) \int_{-\pi}^{\pi} \cos(\phi) g(\phi) d\phi
+\]
+since $\int_{-\pi}^{\pi} \sin(\phi) g(\phi) d\phi = 0$ because $\sin(\phi)g(\phi)$ is odd.  Now 
+\[
+K = \int_{-\pi}^{0} \cos(\phi) g(\phi) d\phi > 0
+\] 
+because $g(\phi) > 0$ is increasing on $[-\pi,0]$ and $\cos(\phi)$ is odd and increasing on $[-\pi,0]$.  Thus $h_1(x) = 2K \cos(x)$ is uniquely maximised at zero because $\cos(x)$ is.
+\end{IEEEproof}
+
+\begin{lemma}
+Under the assumptions in Proposition~\ref{prop:contgg2} let $f_Z$ be the pdf of $\abs{W}$.
+Then $z^{-1} f_Z(z)$ is differentiable and nonincreasing in $z > 0$ and the function $g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr$ is symmetric and unimodal with mode at $\phi = 0$.
+\end{lemma}
+We have
+\[
+f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{\sqrt{x^2 + y^2}}.
+\]
+Put $z = \sqrt{x^2 + y^2}$ and, since $X$ is circularly symmetric,
+\[
+f_{X,Y}(x,y) = f_{X,Y}(z,0) = \frac{f_Z(z)}{z}.
+\]
+Since $f_{X,Y}(z,0)$ is differentiable and nonincreasing in $z = \sqrt{x^2 + y^2}$ (by assumption in the proposition statement) we have that $z^{-1} f_Z(z)$ is differentiable and nonincreasing in $z > 0$.
+So the derivative $\frac{d}{dz} \big( z^{-1} f_Z(z) \big) \geq 0$ for all $z > 0$.
+Using~\eqref{eq:pdfRPhi},
+\[
+g(\phi) = \int_{0}^\infty r^2 \frac{f_Z(z)}{2\pi z } dr
+\]
+where $z = \sqrt{r^2 - 2r\cos\phi + 1}$.  Differentiating with respect to $\phi$,
+\[
+g^\prime(\phi) = \sin(\phi) \int_{0}^\infty \frac{r^3}{\pi} \frac{d}{dz}\left( \frac{f_Z(z)}{z} \right) dr
+\]
+and since the term inside the integral is always positive,
+
 \end{document}

### 2012-10-24 other/confnoncoherent/paper.tex
added to paper.  Filled in first part of the proposition.  It appears difficult to do the second part.  Barry might have some good ideas about this.

--- 2012-10-23 other/confnoncoherent/paper.tex
@@
 That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
-If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
-\[
-f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
-\]  
+If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is $f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z)$.
 If $X$ is circularly symmetric, then for any real number $\phi$, the distribution of $X$ is the same as that of $e^{j\phi}X$.
@@
 
-% We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
-% \[
-% R e^{j\Phi} = 1 + X.
-% \]
-% The joint distribution of $R$ and $\Phi$ can be shown to be
-% \[
-% f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
-% \]
+We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying 
+\[
+R e^{j\Phi} = 1 + X.
+\]
+The joint distribution of $R$ and $\Phi$ can be shown to be
+\[
+f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
+\]
 
@@
 Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
-\begin{equation}\label{eq:RiandPhii}
+%\begin{equation}\label{eq:RiandPhii}
+\[
 R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
-\end{equation}
+\]
+%\end{equation}
 and define the continuous function
@@
 
-A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$.
-The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
+A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
+%This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+%The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
 The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
@@
 
-\begin{proposition}\label{prop:contgg2}
-Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
+% \begin{proposition}\label{prop:contgg2}
+% Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
 Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
 Let,
-\[
-G(x) =  \expect R \cos\sfracpart{x + \Phi}.
-\]
-Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-\end{proposition}
+% \[
+% G(x) =  \expect R \cos\sfracpart{x + \Phi}.
+% \]
+% Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
+% \end{proposition}
 
@@
 In this paper we described the statistical properties of the least squares estimator.
-We showed that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+The phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
 However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.

### 2012-10-24 paper.tex
added to paper.  Filled in first part of the proposition.  It appears difficult to do the second part.  Barry might have some good ideas about this.

--- 2012-10-16 paper.tex
@@
 \[
-f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
-\]
+f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}},
+\]
+from which follows
+\begin{equation}\label{eq:fzandfZYwithz}
+f_Z(z) = 2\pi z f_{X,Y}(z,0) = 2\pi z f_{X,Y}(0,z)
+\end{equation}
+for all $z > 0$.
 
@@
 Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
+%The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
 The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
@@
 
-\begin{proposition}\label{prop:contgg2}
-Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable joint pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
+% \begin{proposition}\label{prop:contgg2}
+% Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable joint pdf $f_{X,Y}(x,y)$ that is decreasing with $x^2 + y^2$.
 Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
 Let,
-\[
-h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
-\]
-Then $h_1(x)$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $h_2(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-\end{proposition}
+% \[
+% h_1(x) = \expect R \cos(x + \Phi) \;\;\; \text{and} \;\;\; h_2(x) =  \expect R \cos\sfracpart{x + \Phi}.
+% \]
+% Then $h_1(x)$ is uniquely maximised at $x=0$ over the interval $[-\pi,\pi]$ and $h_2(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
+% \end{proposition}
 
@@
 
-\subsection{Proof of Proposition~\ref{prop:contgg2}} \label{appendix:prop2proof}
+% \subsection{Proof of Proposition~\ref{prop:contgg2}} \label{appendix:prop2proof}
+
+% \begin{IEEEproof}
+% We first prove the result for $h_1$.  Write
+% \begin{align*}
+% h_1(x) &= \int_{-\pi}^{\pi} \int_{0}^\infty r \cos(\phi + x) f(r,\phi) dr d\phi \\
+% &= \int_{-\pi}^{\pi}  \cos(\phi + x) g(\phi) d\phi,
+% \end{align*}
+% where $f(r,\phi)$ is the joint pdf of $R$ and $\Phi$, and 
+% \[
+% g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
+% \]
+% Since the pdf of $X$ is circularly symmetric, $f(r,\phi)$ and $g(\phi)$ are even in $\phi$ on $[-\pi,\pi]$ (see the discussion after~\eqref{eq:pdfRPhi}).
+It is also the case that $g(\phi)$ is unimodal with mode at zero, as a result of Lemma~\ref{lem:gunimodmodzero}.
+Write $\cos(\phi + x)$ as a sum of even and odd parts,
+% \[
+% \cos(\phi + x) = \cos(\phi)\cos(x) - \sin(\phi)\sin(x).
+% \]
+% Now
+% \[
+% h_1(x) = \cos(x) \int_{-\pi}^{\pi} \cos(\phi) g(\phi) d\phi
+% \]
+% since $\int_{-\pi}^{\pi} \sin(\phi) g(\phi) d\phi = 0$ because $\sin(\phi)g(\phi)$ is odd.  Now 
+% \[
+% K = \int_{-\pi}^{0} \cos(\phi) g(\phi) d\phi > 0
+% \] 
+% because $g(\phi) > 0$ is increasing on $(-\pi,0)$ and $\cos(\phi)$ is odd and increasing on $(-\pi,0)$.
+Thus $h_1(x) = 2K \cos(x)$ is uniquely maximised at zero because $\cos(x)$ is uniquely maximised at zero.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:gunimodmodzero}
+% Under the assumptions in Proposition~\ref{prop:contgg2} let $f_Z$ be the pdf of $\abs{W}$.
+Then $z^{-1} f_Z(z)$ is differentiable and decreasing for $z > 0$ and the function $g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr$ is unimodal on $[-\pi,\pi]$ with mode at $\phi = 0$.
+% \end{lemma}
+% \begin{IEEEproof}
+% From~\eqref{eq:fzandfZYwithz} it follows that $2\pi f_{X,Y}(z,0) = z^{-1} f_Z(z)$ for all $z>0$, and since $f_{X,Y}(z,0)$ is differentiable and nonincreasing in $z$ (by assumption in the statement of Proposition~\ref{prop:contgg2}) we have that $z^{-1} f_Z(z)$ is differentiable and decreasing for $z > 0$.
+So the derivative $\frac{d}{dz} \big( z^{-1} f_Z(z) \big) > 0$ for all $z > 0$.
+Using~\eqref{eq:pdfRPhi},
+% \[
+% g(\phi) = \int_{0}^\infty r^2 \frac{f_Z(z)}{2\pi z } dr
+% \]
+% where $z = \sqrt{r^2 - 2r\cos\phi + 1}$.  Differentiating with respect to $\phi$,
+% \[
+% g^\prime(\phi) = \sin(\phi) \int_{0}^\infty \frac{r^3}{\pi} \frac{d}{dz}\left( \frac{f_Z(z)}{z} \right) dr
+% \]
+% and since the term inside the integral is negative, the integral is negative.
+Since $\sin(\phi)$ is negative on $(-\pi,0)$ and positive on $(0,\pi)$ it follows that the derivative $g^\prime(\phi)$ is positive on $(-\pi,0)$ and negative on $(0,\pi)$.
+Thus $g(\phi)$ is unimodal with mode at $\phi = 0$.
+% \end{IEEEproof}
 

### 2012-10-24 other/confunbiased/paper.tex
Added AusCTW version

--- 2012-10-24 other/confnoncoherent/paper.tex
@@
 %\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
-%\documentclass[conference]{IEEEtran}
-
-\documentclass{article}
-
-\usepackage{spconf}
+\documentclass[conference]{IEEEtran}
 
@@
 
-% %opening
-% \title{Noncoherent least squares estimators of carrier phase and amplitude}
-% \author{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \vspace{0.2cm} \\
-% \small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.1cm} \\
-% \small $^{2}$School of Information Technology \& Electrical Engineering, The University of Queensland, QLD 4067, \textsc{Australia} \vspace{0.1cm} \\
-% \small $^{3}$Department of Statistics, Macquarie University, Sydney, \textsc{Australia} 
-% \thanks{Supported under the Australian Government’s Australian Space Research Program.}
-% }
-
- \title{Noncoherent least squares estimators of carrier phase and amplitude}
-\name{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \thanks{Supported under the Australian Government’s Australian Space Research Program.}}
-\address{ \small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.05cm} \\
-\small $^{2}$School of Information Technology \& Electrical Engineering, The University of Queensland, QLD 4067, \textsc{Australia} \vspace{0.05cm} \\
-\small $^{3}$Department of Statistics, Macquarie University, Sydney, \textsc{Australia} 
+%opening
+\title{An unbiased noncoherent estimator amplitude}
+\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley \vspace{0.2cm} \\
+\small Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.1cm} \\
+\thanks{Supported under the Australian Government’s Australian Space Research Program.}
 }
 
-%\IEEEoverridecommandlockouts
+\IEEEoverridecommandlockouts
 \begin{document}
@@
 \begin{abstract}
-We consider least squares estimators of carrier phase and amplitude from a noisy communications signal.
+We consider the problem of estimating the amplitude of communications signal when the observations are pertubed by additive white Gaussian noise.
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-ary phase shift keying.
-We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator is not consistent, it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
-The results of Monte Carlo simulations are provided and these corroborate the theoretical results.
+Recently, the asymptotic properties of the maximum likelihood estimator of the amplitude were derived.
+The estimator is asymptotically biased, and the bias is large when the signal to noise ratio is small.
+In this paper we consider methods for correcting this bias.
+This involves first observing a period of silence within which the statistical properties (i.e.
+the variance) of the noise is estimated.
+This technique strongly hinges on the assumption that the underlying noise distribution is indeed Gaussian.
+We discuss how this method could be extended to handle the situation when the distribution of the noise is not known.
 \end{abstract}
@@
 
-In passband communication systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation.
+In passband communication systems the transmitted signal typically undergoes time offset (delay), phase shift and attenuation (amplitude change).
 These effects must be compensated for at the receiver.
-In this paper we assume that the time offset has been previously handled, and we focus on estimating the phase shift and attenuation.
+In this paper we assume that the time offset has been previously handled, and we focus on estimating ampltude, whilst treating the phase shift as a nuisance parameter.
 We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK) and $M$-ary phase shift keying ($M$-PSK).
@@
 where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
-Our aim is to estimate $a_0$ from $y_1, \dots, y_L$.
+Our aim is to estimate $\rho_0$ from $y_1, \dots, y_L$.
 If the transmitted symbols $s_1, \dots, s_L$ are known a priori at the receiver then the least squares estimator is typically used,
 \begin{equation}\label{eq:hataumod}
-\hat{a}_{\text{uc}} = \arg \min_{a \in \complex} \sum_{i = 1}^L \abs{ y_i - a s_i }^2  = \frac{1}{L} \sum_{i = 1}^L y_i s_i^*,
+\hat{\rho}_{\text{uc}} = \arg \min_{\rho > 0} \min_{\theta \in [-\pi,\pi)} \sum_{i = 1}^L \abs{ y_i -  \rho e^{j\theta} s_i }^2  = \abs{\frac{1}{L} \sum_{i = 1}^L y_i s_i^*},
 \end{equation}
-where $\complex$ is the set of complex numbers and $x^*$ and $\abs{x}$ denote the conjugate and magnitude of the complex number $x$.
+where $x^*$ and $\abs{x}$ denote the conjugate and magnitude of the complex number $x$.
 Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} call this the \emph{unmodulated carrier} estimator.
@@
 
-The next two theorems describe the asymptotic properties of the least squares estimator.  We omit the proofs due to space constraints.  Proofs will be provided in an upcoming paper.
-
-\begin{theorem}\label{thm:consistency} (Almost sure convergence)
-Let $\{w_i\}$ be a sequence of independent and identically distributed, circularly symmetric complex random variables with $w_1$ having finite variance and continuous pdf.
-Let $y_1,\dots, y_L$ be given by~\eqref{eq:sigmod} and let $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ be the least squares estimator of $a_0 = \rho_0e^{j\theta_0}$.
-Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
-%\begin{equation}\label{eq:RiandPhii}
-\[
-R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i},
-\]
-%\end{equation}
-and define the continuous function
-\[
-G(x) = \expect R_1 \cos\sfracpart{ x + \Phi_1}.
-\] 
-If $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M})$, then:
-\begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
-\end{enumerate}
-\end{theorem}
-
-\begin{theorem}\label{thm:normality} (Asymptotic normality)
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint pdf of $R_1$ and $\Phi_1$, and let
-\[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
-\]
-Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
-%If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
-\[
-\left( \begin{array}{cc} 
-H^{-2} A & 0 \\
-0 & \rho_0^2 B
-\end{array} \right)
-\]
-as $L \rightarrow \infty$, where
-\[
-H = G(0) -  2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}),
-\]
-\[
-A = \expect R_1^2\sin^2\fracpart{\Phi_1}, \;\;\; B = \expect R_1^2 \cos^2\fracpart{\Phi_1} - G^2(0). 
-\]
-\end{theorem}
-
-We now discuss the assumptions made by these theorems.
-The assumption that $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
-If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $s_1, \dots, s_L$.
-In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $s_1, \dots, s_L$, rather than just $\rho_0$.
-The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
-
-A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-%This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-%The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
-The zero mean complex Gaussian distribution with independent real and imaginary parts is in this class.
-We omit the proof due to space constraints.
-
-% \begin{proposition}\label{prop:contgg2}
-% Let $W = X + jY$ be a circularly symmetric complex random variable, the real and imaginary parts having continuous and differentiable pdf $f_{X,Y}(x,y)$ that is nonincreasing with $x^2 + y^2$.
-Let $R \geq 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + W$.
-Let,
-% \[
-% G(x) =  \expect R \cos\sfracpart{x + \Phi}.
-% \]
-% Then $G(x)$ is uniquely maximised at $x=0$ over the interval $[-\tfrac{\pi}{M},\tfrac{\pi}{M}]$.
-% \end{proposition}
-
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}$ rather than directly on $\hat{\theta} - \theta_0$.
-This makes practical sense, and to see why let $s_i' = e^{j2\pi k/M}s_i$ be $M$-PSK symbols obtained by rotating $s_1, \dots, s_L$ by $e^{j2\pi k/M}$ for some integer $k$.
-Then
-\[
-a s_i =  \rho e^{j\theta}s_i = \rho e^{j(\theta - 2\pi k/M)} s_i',
-\]
-and so, if $a = \hat{\rho}e^{j\hat{\theta}}$ minimises~\eqref{eq:hatadef}, then so does $\hat{\rho} e^{j(\hat{\theta} - 2\pi k/M)}$.
-Thus, $\theta_0 - \hat{\theta}$ should be attributed the same error as $\theta_0 - \hat{\theta} - \tfrac{2\pi}{M}k$, that is, the error should be computed `modulo $\tfrac{2\pi}{M}$'.
-It is for this reason that \emph{differential encoding} is often used with noncoherent $M$-PSK  detectors.
-
-\section{The Gaussian noise case}\label{sec:gaussian-noise-case}
-
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint pdf of the real and imaginary parts is
-\[
-p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}.
-\]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
-%and the joint density function of the real and imaginary parts of $Z$ is
-%\[
-%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
-%\] 
-%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
-It can be shown that
-\[
-g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-\]
-where $\kappa = \rho_0/\sigma$ and
-\vspace{-0.1cm}
-\[
-\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt.
-\]
-Each of $A$, $B$ and $H$ can be computed by numerical integration using the above formula for $g(\phi)$.
-
 \section{Simulations}\label{sec:simulations}
-
-We present the results of Monte-Carlo simulations with the least squares estimator.
-In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
-Simulations are run with $M=2,4$ (BPSK and QPSK) and $L=16,256,4096$ with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.
-The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
-For each value of signal to noise ratio $T = 10000$ replications are performed to obtain $T$ estimates $\hat{\rho}_1, \dots, \hat{\rho}_T$ and $\hat{\theta}_1, \dots, \hat{\theta}_T$.
-
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} show the sample mean square error (MSE) of the phase estimator $\hat{\theta}$ computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$.
-The dots show the sample MSE of the least square estimator.
-The dashed line is the sample MSE of the unmodulated carrier estimator~\eqref{eq:hataumod} that has a priori knowledge of the transmitted symbols $s_1, \dots, s_L$.
-%The dashed line acts as a lower bound on performance.
-The solid line is the MSE predicted by Theorem~\ref{thm:normality}.
-The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
-As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$, and Theorem~\ref{thm:normality} does not model this behaviour.
-%An interesting feature is that the phase estimator is not accurate at low SNR unless $L$ is very large.
-
-Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the sample MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-This estimator requires the selection of a function $F : \reals \mapsto \reals$.
-We have chosen $F(x) = 1$.
-The sample MSE of the least squares estimator and the Viterbi and Viterbi estimator is similar.
-The least squares estimator appears slightly more accurate at low SNR.
-
-Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4} show the variance of the amplitude estimator $\hat{\rho}$.
-The solid line is the asymptotic variance predicted by Theorem~\ref{thm:normality}, the dots are the Monte-Carlo simulations with the least squares estimator, and the dashed line the simulations with the unmodulated carrier estimator.
-Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
-This requires $G(0)$ to be known.
-In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
-When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
-However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
 
@@
 
-\section{RELATION TO PRIOR WORK}
+\section{Conclusion}
 
-The problem of estimating carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols has undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
-The least squares estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and this is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
-The algorithm was later rediscovered by Sweldens~\cite{Sweldens2001}.
-In this paper we described the statistical properties of the least squares estimator.
-The phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.
-This bias is large when the signal to noise ratio is small.
-It would be interesting to investigate methods for correcting this bias.
+% %\newpage
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
+% 		\caption{Phase error for BPSK}
+% 		\label{fig:plotphaseM2}
+% \end{figure}
 
-%\newpage
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM2-2.mps}
-		\caption{Phase error for BPSK}
-		\label{fig:plotphaseM2}
-\end{figure}
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
+% 		\caption{Phase error for QPSK}
+% 		\label{fig:plotphaseM4}
+% \end{figure}
 
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM4-2.mps}
-		\caption{Phase error for QPSK}
-		\label{fig:plotphaseM4}
-\end{figure}
+% %\pagebreak
+% %\newpage
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotncM2-1.mps}
+% 		\caption{Amplitude error for BPSK}
+% 		\label{fig:plotampM2}
+% \end{figure}
 
-%\pagebreak
-%\newpage
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM2-1.mps}
-		\caption{Amplitude error for BPSK}
-		\label{fig:plotampM2}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotncM4-1.mps}
-		\caption{Amplitude error for QPSK}
-		\label{fig:plotampM4}
-\end{figure}
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotncM4-1.mps}
+% 		\caption{Amplitude error for QPSK}
+% 		\label{fig:plotampM4}
+% \end{figure}
 

### 2012-10-24 other/confunbiased/build.sh

@@ -1,0 +1,9 @@
+cd code/data/
+mpost plotncM2.mp
+mpost plotncM4.mp
+cd ../..
+
+pdflatex paper.tex
+bibtex paper
+pdflatex paper.tex
+pdflatex paper.tex

### 2012-10-24 other/confunbiased/paper.tex
added minor.  This needs a bit of thinking probably.

@@
 \usepackage[square,comma,numbers,sort&compress]{natbib}
-\usepackage{bibspacing}
-\setlength{\bibspacing}{0cm}
+%\usepackage{bibspacing}
+%\setlength{\bibspacing}{0cm}
 
@@
 %opening
-\title{An unbiased noncoherent estimator amplitude}
+\title{Unbiased noncoherent amplitude estimation}
 \author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley \vspace{0.2cm} \\
@@
 The estimator is asymptotically biased, and the bias is large when the signal to noise ratio is small.
-In this paper we consider methods for correcting this bias.
-This involves first observing a period of silence within which the statistical properties (i.e.
-the variance) of the noise is estimated.
-This technique strongly hinges on the assumption that the underlying noise distribution is indeed Gaussian.
-We discuss how this method could be extended to handle the situation when the distribution of the noise is not known.
+We consider a method for correcting this bias.
+This involves the reciever observing a period of `silence' which is used to estimate the statistical properties (i.e.
+the variance) of the noise.
+This approach strongly hinges on the assumption that the underlying noise distribution is indeed Gaussian.
+We discuss how this approach could be extended to handle the situation when the distribution of the noise is not assumed.
 \end{abstract}

### 2012-10-25 other/confnoncoherent/paper.tex
minor

@@
 
-We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying 
-\[
-R e^{j\Phi} = 1 + X.
-\]
+We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.
+Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying $R e^{j\Phi} = 1 + X$.
 The joint distribution of $R$ and $\Phi$ can be shown to be
@@
 Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+One can check that the assumption holds when $w_1$ is circularly symmetric and normally distributed.
+%We will not attempt to further classify those distributions for which the assumption holds here.
 %The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
@@
 
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint pdf of the real and imaginary parts is
+Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
 \[
-p(x,y) = \frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}.
+\frac{1}{2\pi\sigma^2}e^{-\frac{1}{2\sigma^2}(x^2 + y^2)}.
 \]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
 %and the joint density function of the real and imaginary parts of $Z$ is
@@
 \[
-g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+%g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+g(\phi) = \frac{\cos(\phi)}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2(\phi)} \big(1 + \kappa^2\cos^2(\phi) \big)
 \]
 where $\kappa = \rho_0/\sigma$ and
-\vspace{-0.1cm}
 \[
-\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt.
+\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos(\phi)}{\sqrt{2}}\right)
 \]
-Each of $A$, $B$ and $H$ can be computed by numerical integration using the above formula for $g(\phi)$.
+and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
 

### 2012-10-25 paper.tex
minor

@@
 \[
-f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}},
-\]
-from which follows
-\begin{equation}\label{eq:fzandfZYwithz}
-f_Z(z) = 2\pi z f_{X,Y}(z,0) = 2\pi z f_{X,Y}(0,z)
-\end{equation}
-for all $z > 0$.
+f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
+\]
+%from which follows
+%\begin{equation}\label{eq:fzandfZYwithz}
+%f_Z(z) = 2\pi z f_{X,Y}(z,0) = 2\pi z f_{X,Y}(0,z)
+%\end{equation}
+%for all $z > 0$.
 
@@
 Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+One can check that the assumption holds when $w_1$ is circularly symmetric and normally distributed.
+We will not attempt to further classify those distributions for which the assumption holds here.
+
 %The next proposition describes a class of circularly symmetric distributions for which $G(x)$ is uniquely maximised at $x = 0$.
@@
 \end{align}
-since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
+since $\round{x + \angle{s_i}} = \round{x} + \angle{s_i}$ for all $x \in \reals$ as a result of $\angle{s_i}$ being a multiple of $\tfrac{2\pi}{M}$.  Otherwise, when $i \in P$,  
 \[
@@
 We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-The proof of part 1 of the Theorem~\ref{thm:consistency} follows from this.
+The proof of part 1 of Theorem~\ref{thm:consistency} follows from this.
 
@@
 The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989}.
+This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989,van2009empirical}.
 However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
@@
 \[
-\frac{1}{2\pi\sigma^2}e^{-(x^2 + y^2)/\sigma^2}
-\]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold and, since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+\frac{1}{2\pi\sigma^2}e^{-\frac{1}{2\sigma^2}(x^2 + y^2)}.
+\]
+Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
 %and the joint density function of the real and imaginary parts of $Z$ is
@@
 \[
-g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+%g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+g(\phi) = \frac{\cos(\phi)}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2(\phi)} \big(1 + \kappa^2\cos^2(\phi) \big)
 \]
@@
 \[
-\Psi(\phi) = 1 + \erf(\kappa \cos(\phi)) = 1 + \frac{2}{\sqrt{\pi}}\int_0^{\kappa \cos(\phi)} e^{-t^2} dt .
-\]
-The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
+\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos(\phi)}{\sqrt{2}}\right)
+\]
+and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
 
@@
 
-\begin{lemma}\label{lem:GLtoG0}
-$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-By the triangle inequality,
-\[
-\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
-\]
-Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
-\end{IEEEproof}
-
 \begin{lemma}\label{lem:uniflawGGL} 
-The function $G_L$ converges almost surely and uniformly on $[-\pi, \pi)$ to the function $G$, that is,
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0
-\] 
-almost surely as $L \rightarrow \infty$.
+$\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
 \end{lemma}
@@
 \[
-\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda)\big)} \rightarrow 0
+\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda_n)\big)} \rightarrow 0
 \]
@@
 \begin{enumerate}
-\item $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$,
+\item $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ for all positive integers $L$,
 \item for any $\epsilon > 0$, $\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0$.
@@
 \] 
-almost surely to $\frac{2\pi}{N}\expect R_1$ as $L \rightarrow\infty$ by the strong law of large numbers, and so, for any $\epsilon > 0$,
+almost surely as $L \rightarrow\infty$ by the strong law of large numbers, and so, for any $\epsilon > 0$,
 \begin{align*}
@@
 
+\begin{lemma}\label{lem:GLtoG0}
+$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+By the triangle inequality,
+\[
+\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
+\]
+Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{IEEEproof}
+
 \subsection{Lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-normappendix}
@@
 \[
- \sin(q_i) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
+ \sin(q_i) > -\frac{\rho_0 \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
 \]
@@
 for all $i \in D$.
-Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ when $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.
+Thus, $\sfracpart{\hat{\lambda}_L + \Phi_i} \neq -\tfrac{\pi}{M}$ for $i \in D$ such that $R_i > 0$ and therefore $G_L$ is differentiable at $\hat{\lambda}_L$.
 That $G_L^\prime(\hat{\lambda}_L) = 0$ follows since $\hat{\lambda}_L$ is a maximiser of $G_L$.
@@
 \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
- &= 2 \sin(\tfrac{\pi}{M}) \sin(v),
+ &= 2 \sin(\tfrac{\pi}{M}) \sin(v).
 \end{align*}
@@
 \[
-0 > \sin(v) > -\frac{\rho_o \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
-\]
-Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$, violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.
-Thus,~\eqref{eq:fraclambassumption} is false.
+0 > \sin(v) > -\frac{\rho_0 \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
+\]
+Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then 
+\[
+SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
+\] 
+violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  So~\eqref{eq:fraclambassumption} is false.
 
@@
 Now an analogous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
-
 \end{IEEEproof}
@@
 k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
-&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big) 
+&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big),
 \end{align*}
 since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ as a result of~\eqref{eq:expectRepartRphi}.
-Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$, i.e.
-$\zeta(\hat{\lambda}_L) = o_P(1)$.
+Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$.
 So,
@@
 \]
-Observe that both $s_k$ and $\eta_k$ are continuous and that 
+Observe that both $s_k$ is continuous and that 
 \[
@@
 \begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) &g(\phi) d\phi = \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
 &= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
@@
 \end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}, it follows that $c_k(\lambda)$ converges to
+Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ (by assumption in Theorem~\ref{thm:normality}), it follows that $c_k(\lambda)$ converges to
 \[
@@
 &= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \left( C(\lambda)  + A(\lambda) + B(\lambda) \right),
+&= \lambda \big( C(\lambda)  + A(\lambda) + B(\lambda) \big),
 \end{align*}
@@
 \[
-A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \;\;\; B(\lambda) = \sum_{k=0}^{M} b_k(\lambda), \;\;\; \text{and},
+A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \qquad B(\lambda) = \sum_{k=0}^{M} b_k(\lambda),  
 \]
@@
 \]
-Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda$ goes to zero, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.  Put
-\[
-\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1.
-\]
-Then $\gamma(\lambda)$ converges to zero as $\lambda$ goes to zero and $W(\lambda) = \lambda(H_1 + \gamma(\lambda))$ as required.
+Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda \rightarrow 0$, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.
+The lemma follows after puting $\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1$.
 \end{IEEEproof}
@@
 \]
-Lemma~\ref{lem:ZYLempiricproc} in Section~\ref{sec:an-empirical-process} shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+Lemma~\ref{lem:ZYLempiricproc} 
+%in Section~\ref{sec:an-empirical-process} 
+shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
 \[
@@
 &= \prob\left\{ \sabs{W_L(\hat{\lambda}_L)} > \delta \;\text{and} \; \sabs{\hat{\lambda}_L} < \epsilon \right\} \\
-&\hspace{0.7cm} + \prob\left\{ \sabs{W_L(\lambda_L)} > \delta  \;\text{and} \; \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
-&\leq \prob\left\{  \sup_{\lambda < \epsilon} \sabs{ W_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+&\hspace{0.7cm} + \prob\left\{ \sabs{W_L(\hat{\lambda}_L)} > \delta  \;\text{and} \; \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
+&\leq \prob\left\{  \sup_{\sabs{\lambda} < \epsilon} \sabs{ W_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
 &\leq 2\nu.
@@
 \expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
-&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi,
-\end{align*}
-where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
+&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi.
+\end{align*}
+The proof is immediate since $g(\phi)$ is even and $\sin\fracpart{\phi}$ is odd.
+% where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
 Proposition~\ref{prop:gg2cont} shows that $g$ is even on $(-\pi,\pi)$.
@@
 So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
-\[
-g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
-\]
-and
-\[
-\sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
-\]
-Now,
-\begin{align*}
-I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
-&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
-&\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
-&= 0
-\end{align*}
-for all integers $k$ because
-\begin{align*}
-\int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
-&= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
-&= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
-\end{align*}
-Observe that the sets
-\[
- [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
-\]
-partition $[-\pi, \pi]$, and so
-\[
-\expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
-\]
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:empiricprocforrho} Let $T_L(\lambda) = \expect G_L(\lambda)$.  Then
+% \[
+% g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
+% \]
+% and
+% \[
+% \sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
+% \]
+% Now,
+% \begin{align*}
+% I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
+% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
+% &\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+% &= 0
+% \end{align*}
+% for all integers $k$ because
+% \begin{align*}
+% \int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
+% &= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
+% \end{align*}
+% Observe that the sets
+% \[
+%  [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
+% \]
+% partition $[-\pi, \pi]$, and so
+% \[
+% \expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
+% \]
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:empiricprocforrho} Let $T_L(\lambda) = \expect G_L(\lambda)$.  We have
 \[
@@
 \end{align*}
-because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ is an assumption in the statement of Theorem~\ref{thm:normality}.  Now,
+and because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ (by assumption in Theorem~\ref{thm:normality}), we have
 \begin{align*}

### 2012-10-25 code/montecarlo.scala

@@ -20,7 +20,7 @@ val iters = 5000
 //construct an array of noise distributions with a logarithmic scale
 val SNRdBs = -20 to 20 by 1
 val SNRs = SNRdBs.map(db => scala.math.pow(10.0, db/10.0))
-val noises = SNRs.map( snr => new ComplexGaussian(a0.mag2/snr/2) ) //variance for real and imaginay parts (divide by 2)
+val noises = SNRs.map( snr => new ComplexGaussian(a0.mag2/snr/2.0) ) //variance for real and imaginay parts (divide by 2)
 
 val starttime = (new java.util.Date).getTime
 for( L <- Ls; M <- Ms ) {

### 2012-10-26 other/confnoncoherent/paper.tex
added some of barry's edits

@@
 We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator is not consistent, it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
+However, the amplitude estimator is not consistent, but converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
 The results of Monte Carlo simulations are provided and these corroborate the theoretical results.
@@
 
-Here we are interested in \emph{noncoherent detection}, where $s_1, \dots, s_L$ are not known at the receiver, and must also be estimated.
+In the paper we are interested in \emph{noncoherent detection}, where $s_1, \dots, s_L$ are not known at the receiver, and must also be estimated.
 This estimation problem has undergone extensive prior study and is often called \emph{multiple symbol differential detection}~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
@@
 The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is white and Gaussian.
-As we will show, the estimator can work well even when the noise is not Gaussian.
+As we show, the estimator can work well even when the noise is not Gaussian.
 Mackenthun~\cite{Mackenthun1994} described an algorithm to compute the least squares estimator $\hat{a}$ that requires only $O(L \log L)$ arithmetic operations.
@@
 In the literature it has been common to assume that the symbols $s_1, \dots, s_L$ are of primary interest and the complex amplitude $a_0$ is a nuisance parameter.
-The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
-Whilst estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite approach here.
+The metric of performance is correspondingly the \emph{symbol error rate}, or \emph{bit error rate}.
+While estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite point of view here.
 Our aim is to estimate $a_0$, and we treat the unknown symbols as nuisance parameters.
@@
 The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be $\abs{a_0 - \hat{a}}^2$.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle^2$ and $(\rho_0 - \hat{\rho})^2$.
-The function $\fracpart{\cdot}$ denotes its argument taken `modulo $\tfrac{2\pi}{M}$' into the interval $[-\pi/M, \pi/M)$.
+Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{a_0 - \hat{a}}^2$.
+It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle^2$ and $(\rho_0 - \hat{\rho})^2$.
+The function $\fracpart{\cdot}$ denotes its argument taken `modulo $\tfrac{2\pi}{M}$' into the interval $[-\pi/M, \pi/M)$, that is
+\[
+\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right),
+\]  
+where $\operatorname{round}(\cdot)$ takes its argument to the nearest integer.
+The direction of rounding for half-integers is not important so long as it is consistent.
+We have chosen to round up half-integers here.
 It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
@@
 
-The next two theorems describe the asymptotic properties of the least squares estimator.  We omit the proofs due to space constraints.  Proofs will be provided in an upcoming paper.
+The next two theorems describe the asymptotic properties of the least squares estimator.  We omit the proofs due to space constraints.  Proofs will be provided in an forthcomming paper.
 

### 2012-10-26 paper.tex
added some of barry's edits

@@
 We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator is not consistent, it converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation.
+However, the amplitude estimator is not consistent, but converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
 Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.
@@
 The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
-However, as we will show, the estimator works well under less stringent assumptions.
+However, as we show, the estimator works well under less stringent assumptions.
 %It is the least squares estimator that we primarily study in this paper.
@@
 In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.
-The metric of performance is correspondingly \emph{symbol error rate}, or \emph{bit error rate}.
-Whilst estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite approach here.
+The metric of performance is correspondingly the \emph{symbol error rate}, or \emph{bit error rate}.
+While estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite point of view here.
 Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
@@
 The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
-Consequently, our metric of performance will not be symbol or bit error rate, it will be $\abs{a_0 - \hat{a}}^2$.
-It will actually be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.
+Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{a_0 - \hat{a}}^2$.
+It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
@@
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
-The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
+The asymptotic bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the asymptotic bias is significant when the SNR is small.
 Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.
@@
 Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
-A complex random variable $W$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{W}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{W}$.
+A complex random variable $W$ is said to be \emph{circularly symmetric} if its phase $\angle{W}$ is independent of its magnitude $\abs{W}$ and if the distribution of $\angle{W}$ is uniform on $[0,2\pi)$.
 That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
@@
 \[
-g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr, \qquad g_2(\phi) = \int_{0}^{\infty} r^2 f(r,\phi) dr,
+g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr %\qquad g_2(\phi) = \int_{0}^{\infty} r^2 f(r,\phi) dr,
 \]
@@
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-If the functions $g$ and $g_2$ are continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$, then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
+If the function $g$ is continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$, then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
 \[
@@
 
-Theorem~\ref{thm:normality} requires functions $g$ and $g_2$ to be continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$.
+Theorem~\ref{thm:normality} requires the function $g$ to be continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$.
 This places mild restrictions on the distribution of the noise $w_i$.
 For example, the requirements are satisfied if the joint pdf of the real and imaginary parts of $w_i$ is continuous, since in this case $f(r,\phi)$ is continuous.
-Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$ it follows that both $g$ and $g_2$ have period $2\pi$ and are even on $[-\pi, \pi]$.
+Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$ it follows that $g$ has period $2\pi$ and is even on $[-\pi, \pi]$.
 
@@
 \begin{IEEEproof}
-Put 
+Write
+\[
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = W_L(\hat{\lambda}_L) + \sqrt{L} R_L(0)
+\]
+where
 \begin{equation}\label{eq:WLdef}
@@
 \end{equation}
-so that
-\[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = W_L(\hat{\lambda}_L) + \sqrt{L} R_L(0)
-\]
-Lemma~\ref{lem:ZYLempiricproc} 
+is what is called an \emph{empirical process} indexed by $\lambda$~\cite{Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986,van2009empirical,Pollard_conv_stat_proc_1984}.
+Techniques from this literature can be used to show that
+%Lemma~\ref{lem:ZYLempiricproc} 
 %in Section~\ref{sec:an-empirical-process} 
-shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+%shows that 
+for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
 \[
@@
 \]
-for all positive integers $L$.  Since $\hat{\lambda}_L$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
+for all positive integers $L$.
+This type of result is typically called \emph{tightness} or \emph{asymptotic continuity}~\cite{Pollard_asymp_empi_proc_1989,van2009empirical,Billingsley1999_convergence_of_probability_measures}.
+We ommit the proof which follows in a straightforward, but lengthy manner using an argument called \emph{symmetrisation} followed by an argument called \emph{chaining}~\cite{Pollard_asymp_empi_proc_1989,van2009empirical}.
+
+Since $\hat{\lambda}_L$ converges almost surely to zero, it follows that for any $\epsilon > 0$,
 \[
@@
 \begin{IEEEproof}
-Let
+Write
+\[
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\hat{\lambda}_L) + X_L
+\]
+where
 \begin{equation}\label{eq:YLdef}
-Y_L(\lambda) = \sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) - X_L
-\end{equation}
-so that
-\[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = Y_L(\hat{\lambda}_L) + X_L
-\]
-Lemma~\ref{lem:ZYLempiricproc} 
+Y_L(\lambda) = \sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) - X_L.
+\end{equation}
+is an empirical process indexed by $\lambda$, similar to $W_L$ from~\eqref{eq:WLdef}.  As with $W_L$ results from the literature on empirical processes can be used to show that  
+%Lemma~\ref{lem:ZYLempiricproc} 
 %in Section~\ref{sec:an-empirical-process} 
-shows that for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
+%shows that 
+for any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that
 \[
@@
 
-\subsection{An empirical process result}\label{sec:an-empirical-process}
-
-During the proof of asymptotic normality in Section~\ref{sec:proof-asympt-norm} we made use of the following result regarding the functions $W_L(\lambda)$ and $Y_L(\lambda)$ defined in~\eqref{eq:WLdef}~and~\eqref{eq:YLdef}.
-
-\begin{lemma}\label{lem:ZYLempiricproc}
-For any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that:
-\begin{enumerate}
-\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu$, and,
-\item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu$,
-\end{enumerate}
-for all positive integers $L$.
-\end{lemma}
-\begin{IEEEproof}
-These results are related to what is called \emph{tightness} in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
-The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989,Gine_Zinn_symmetrisation_1984,Ossiander_clt_bracketing_1984}.
+% \subsection{An empirical process result}\label{sec:an-empirical-process}
+
+% During the proof of asymptotic normality in Section~\ref{sec:proof-asympt-norm} we made use of the following result regarding the functions $W_L(\lambda)$ and $Y_L(\lambda)$ defined in~\eqref{eq:WLdef}~and~\eqref{eq:YLdef}.
+
+% \begin{lemma}\label{lem:ZYLempiricproc}
+% For any $\delta > 0$ and $\nu > 0$, there exists $\epsilon > 0$ such that:
+% \begin{enumerate}
+% \item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{W_L(\lambda)} > \delta  \right\} < \nu$, and,
+% \item $\Pr\left\{ \sup_{\sabs{\lambda}<\epsilon}\sabs{Y_L(\lambda)} > \delta  \right\} < \nu$,
+% \end{enumerate}
+% for all positive integers $L$.
+% \end{lemma}
+% These results are related to what is called \emph{tightness} (or \emph{asymptotic continuity}) in the literature on empirical processes and weak convergence on metric spaces~\cite{Billingsley1999_convergence_of_probability_measures,Dudley_unif_central_lim_th_1999,Shorak_emp_proc_stat_2009}.
+The proof is based on a technique called \emph{symmetrisation} and another technique called \emph{chaining} (also known as \emph{bracketing})~\cite{Pollard_asymp_empi_proc_1989,Gine_Zinn_symmetrisation_1984,Ossiander_clt_bracketing_1984,van2009empirical}.
+These techniques are well studied in the statistics and probability literature, but do not appear well known to engineers, so we give the arguments in full here.
 We will only give a proof of the first statement regarding the function $W_L$.
@@
 
-Put 
-\[
-f_i(\lambda, R_i, \Phi_i) = \begin{cases}
-R_i( \sin(\lambda + \Phi_i) - \sin(\Phi_i)), & i \in P \\
-R_i( \sin\sfracpart{\lambda + \Phi_i} - \sin\sfracpart{\Phi_i}), & i \in D \\
-\end{cases}
-\]
-so that
-\[
-W_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect f_i(\lambda, R_i, \Phi_i) \big).
-\]
-Let $\{g_i\}$ be a sequence of independent standard normal random variables, independent of $\{R_i\}$ and $\{\Phi_i\}$.  Lemma~\ref{lem:symmetrisation} shows that
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) },
-\]
-where 
-\begin{equation}\label{eq:ZpsiCondGaussProc}
-X(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} g_i f_i(\lambda, R_i, \Phi_i),
-\end{equation}
-and where $\expect$ runs over all of $\{g_i\}$, $\{R_i\}$ and $\{\Phi_i\}$.
+% \begin{IEEEproof}
+% Put 
+% \[
+% f_i(\lambda, R_i, \Phi_i) = \begin{cases}
+% R_i( \sin(\lambda + \Phi_i) - \sin(\Phi_i)), & i \in P \\
+% R_i( \sin\sfracpart{\lambda + \Phi_i} - \sin\sfracpart{\Phi_i}), & i \in D \\
+% \end{cases}
+% \]
+% so that
+% \[
+% W_L(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect f_i(\lambda, R_i, \Phi_i) \big).
+% \]
+% Let $\{g_i\}$ be a sequence of independent standard normal random variables, independent of $\{R_i\}$ and $\{\Phi_i\}$.  Using symmetrisation, Lemma~\ref{lem:symmetrisation} shows that
+% \[
+% \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) },
+% \]
+% where 
+% \begin{equation}\label{eq:ZpsiCondGaussProc}
+% X(\lambda) = \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} g_i f_i(\lambda, R_i, \Phi_i),
+% \end{equation}
+% and where $\expect$ runs over all of $\{g_i\}$, $\{R_i\}$ and $\{\Phi_i\}$.
 Conditionally on $\{\Phi_i\}$ and $\{R_i\}$, $\{X(\lambda), \lambda \in \reals \}$ is a \emph{Gaussian process}, and numerous techniques exist for its analysis.
-Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists $\epsilon > 0$ such that
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa.
-\]
-It follows that,
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)}  <  \sqrt{2\pi} \; \kappa,
-\]
-and by Markov's inequality,
-\[
-\prob \cubr{  \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
-\]
-for any $\delta > 0$.  The proof follows by choosing $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:symmetrisation} (Symmetrisation)
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) }
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be a sequence identically distributed to $\{\Phi_i\}$.
+Using chaining, Lemma~\ref{lem:chaining} shows that for any $\kappa > 0$ there exists $\epsilon > 0$ such that
+% \[
+% \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa.
+% \]
+% It follows that,
+% \[
+% \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)}  <  \sqrt{2\pi} \; \kappa,
+% \]
+% and by Markov's inequality,
+% \[
+% \prob \cubr{  \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} > \delta } \leq  \sqrt{2\pi} \; \frac{\kappa}{\delta},
+% \]
+% for any $\delta > 0$.  The proof follows by choosing $\nu =  \sqrt{2\pi} \kappa/\delta$.  It remains to prove Lemmas~\ref{lem:symmetrisation}~and~\ref{lem:chaining}.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:symmetrisation} (Symmetrisation \cite{Pollard_asymp_empi_proc_1989,van2009empirical,Gine_Zinn_symmetrisation_1984})
+% \[
+% \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) }
+% \]
+% \end{lemma}
+% \begin{IEEEproof}
+% Let $\{R_i'\}$ be a sequence identically distributed to $\{R_i\}$ and let $\{\Phi_i'\}$ be a sequence identically distributed to $\{\Phi_i\}$.
 Both $\{R_i'\}$ and $\{\Phi_i'\}$ are chosen independently of $\{R_i\}$ and $\{\Phi_i\}$.
-Let $\expect_{R,\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
+Let $\expect_{R\Phi}$ denote expectation conditional on $\{R_i\}$ and $\{\Phi_i\}$.
 Now,
-\begin{align*}
-W_L(\lambda) &= \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect_{R,\Phi} f_i(\lambda, R_i', \Phi_i') \big) \\
-&= \expect_{R,\Phi} \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ),
-\end{align*}
-where, for notational convenience, we put 
-\[
-f_{i} = f_i(\lambda, R_i, \Phi_i) \qquad \text{and} \qquad  f_{i}^\prime = f_i(\lambda, R_i', \Phi_i').
-\] 
-Taking absolute values followed by supremums,
-\begin{align*}
- \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &= \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \expect_{R,\Phi} \sum_{i \in P \cup D} (f_{i} - f_{i}^\prime )} \\
-&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi}  \abs{ \frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) },
-\end{align*}
-the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, the bound can be increased by moving $\expect_{R,\Phi}$ outside the supremum,
-\begin{equation}\label{eq:Eoutsidesup}
- \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ) }.
-\end{equation}
-Applying $\expect$ to both sides gives the inequality
-\begin{equation}\label{eq:Gninequfin}
- \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} (f_{i} - f_{i}^\prime) },
-\end{equation}
-since $\expect \expect_{R,\Phi}$ is equivalent to $\expect$.   Let
-\[
-\sigma_i = \frac{g_i}{\abs{g_i}},
-\]
-and put $\sigma_i = 1$ if $g_i = 0$.  The $\sigma_i$ are thus independent with 
-\[
-\prob\{ \sigma_i = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_i = 1 \} = \tfrac{1}{2}.
-\]  
+% \begin{align*}
+% W_L(\lambda) &= \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} \big( f_i(\lambda, R_i, \Phi_i) - \expect_{R\Phi} f_i(\lambda, R_i', \Phi_i') \big) \\
+% &= \expect_{R\Phi} \frac{1}{\sqrt{L}} \sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ),
+% \end{align*}
+% where, for notational convenience, we put 
+% \[
+% f_{i} = f_i(\lambda, R_i, \Phi_i) \qquad \text{and} \qquad  f_{i}^\prime = f_i(\lambda, R_i', \Phi_i').
+% \] 
+% Taking absolute values followed by supremums,
+% \begin{align*}
+%  \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &= \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \expect_{R\Phi} \sum_{i \in P \cup D} (f_{i} - f_{i}^\prime )} \\
+% &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi}  \abs{ \frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) },
+% \end{align*}
+% the upper bound following from Jensen's inequality.  Since $\sup \expect \abs{\dots} \leq \expect \sup \abs{\dots}$, the bound can be increased by moving $\expect_{R,\Phi}$ outside the supremum,
+% \begin{equation}\label{eq:Eoutsidesup}
+%  \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} ( f_{i} - f_{i}^\prime ) }.
+% \end{equation}
+% Applying $\expect$ to both sides gives the inequality
+% \begin{equation}\label{eq:Gninequfin}
+%  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} \leq  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} (f_{i} - f_{i}^\prime) },
+% \end{equation}
+% since $\expect \expect_{R\Phi}$ is equivalent to $\expect$.   Let
+% \[
+% \sigma_i = \frac{g_i}{\abs{g_i}},
+% \]
+% and put $\sigma_i = 1$ if $g_i = 0$.  The $\sigma_i$ are thus independent with 
+% \[
+% \prob\{ \sigma_i = -1 \} = \tfrac{1}{2} \qquad \text{and} \qquad  \prob\{ \sigma_i = 1 \} = \tfrac{1}{2}.
+% \]  
+% The symmetry of the distribution of $g_i$ implies that $g_i$ and $\sigma_i$ are independent.
 As $R_i$ and $\Phi_i$ are independent and identically distributed to $R_i^\prime$ and $\Phi_i^\prime$, the random variable $f_{i} - f_{i}^\prime$ is symmetrically distributed about zero, and is therefore distributed identically to $\sigma_i( f_{i} - f_{i}^\prime)$.
 Thus,
-%\[
-%\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
-%\]
-%are identically distributed, and therefore
-\begin{align*}
- \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) } \\
-&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i ( f_{i} - f_{i}^\prime) }.
-\end{align*}
-By the triangle inequality,
-\[
-\abs{ \sum_{i\in P \cup D} \sigma_i (f_{i} - f_{i}^\prime)} \leq \abs{ \sum_{i \in P \cup D} \sigma_i f_{i} } + \abs{ \sum_{i\in P \cup D}  \sigma_i f_{i}^\prime },
-\]
-and it follows from~\eqref{eq:Gninequfin}, that
-\begin{align*}
- \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &\leq \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} \sigma_i f_{i} } \\
-&\hspace{1cm} + \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  \sigma_i  f_{i}^\prime } \\
-&= 2 \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} }.
-\end{align*}
-Because $g_i$ is a standard normal random variable it follows that $\expect \abs{g_i} = \expect_{R,\Phi} \abs{g_i} = \sqrt{2/\pi}$, and
-\begin{align*}
-\expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} } \\
-&= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} \sqrt{\frac{\pi}{2}}\expect_{R,\Phi} \abs{g_i}} \\
-&\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{ \sqrt{L}}\sum_{i \in P \cup D} \sigma_i \abs{g_i}  f_{i}},
-\end{align*}
-the last line following from Jensen's inequality and by moving $\expect_{R,\Phi}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$,
-\begin{align*}
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} g_i  f_{i}} \\
-&= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) } 
-\end{align*}
-as required.
-\end{IEEEproof}
-
-\begin{lemma} \label{lem:chaining}
-For any $\kappa > 0$ there exists $\epsilon > 0$ such that
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa,
-\]
-where $X(\lambda)$ is the Gaussian process defined in~\eqref{eq:ZpsiCondGaussProc}.
-\end{lemma}
-\begin{IEEEproof}
-Without loss of generality, assume that $\epsilon < \tfrac{\pi}{M}$.  Lemma~\ref{lem:chaining2} shows that
-\begin{equation}\label{eq:supZCK1} 
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}
-\end{equation}
-where $K_1$ is a finite, positive constant, and
-\begin{equation}\label{eq:Cepsdefn}
-C_\epsilon = \frac{2}{L}\epsilon\sum_{i\in P \cup D}R_i^2 + \frac{4}{L}\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
-\end{equation}
-where $I_\epsilon$ is the indicator function
-\[
-I_{\epsilon}(x) = \begin{cases}
-1, & x \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon] \\
-0, & \text{otherwise},
-\end{cases}
-\] 
-and $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  Now,
-\[
-\expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
-\]
-Recalling that $f(r,\phi)$ is the joint probability density function of $R_i$ and $\Phi_i$,
-\begin{align*}
-\expect R_i^2 I_\epsilon(\Phi_i) &= \int_0^{2\pi} \int_{0}^\infty  r^2 f(r,\phi) I_\epsilon(\phi) dr d\phi \\ 
-&= \int_0^{2\pi}g_2(\phi) I_\epsilon(\phi) d\phi 
-\end{align*}
-where the function
-\[
-g_2(\phi) = \int_{0}^\infty  r^2 f(r,\phi) dr d\phi 
-\]
-is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ by Proposition~\ref{prop:gg2cont}.  From the definition of $I_\epsilon$,
-\begin{align*}
-\expect R_i^2 I_\epsilon(\Phi_i) &= \sum_{k=0}^{M-1}\int_{\psi_k-\epsilon}^{\psi_k-\epsilon}g_2(\phi) d\phi \\
-&= 2\epsilon \sum_{k=0}^{M-1} (g_2(\psi_k) + o(1))
-\end{align*}
-where $o(1)$ goes to zero as $\epsilon$ goes to zero, by the mean value theorem.
-Thus $\expect R_i^2 I_\epsilon(\Phi_i) \leq K_2\epsilon$ for some constant $K_2$.
-Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive,
-\[
-\expect \sqrt{C_\epsilon} \leq  \sqrt{\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
-\]
-by Jensen's inequality.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
-\[
-\expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
-\]
-Choosing 
-\[
-\epsilon = \frac{\kappa^2}{ K_1^2 ( K_2 + 2\expect R_1^2)}
-\] 
-completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
-\end{IEEEproof}
-
-The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
+% %\[
+% %\sum_{n=1}^{N}  (f_{nN} - f_{nN}^\prime) \qquad \text{and} \qquad \sum_{n=1}^{N}  \sigma_n(f_{nN} - f_{nN}^\prime)
+% %\]
+% %are identically distributed, and therefore
+% \begin{align*}
+%  \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D}  (f_{i} - f_{i}^\prime) } \\
+% &= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{  \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i ( f_{i} - f_{i}^\prime) }.
+% \end{align*}
+% By the triangle inequality,
+% \[
+% \abs{ \sum_{i\in P \cup D} \sigma_i (f_{i} - f_{i}^\prime)} \leq \abs{ \sum_{i \in P \cup D} \sigma_i f_{i} } + \abs{ \sum_{i\in P \cup D}  \sigma_i f_{i}^\prime },
+% \]
+% and it follows from~\eqref{eq:Gninequfin}, that
+% \begin{align*}
+%  \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ W_L(\lambda)} &\leq \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} \sigma_i f_{i} } \\
+% &\hspace{1cm} + \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{\frac{1}{\sqrt{L}} \sum_{i \in P \cup D}  \sigma_i  f_{i}^\prime } \\
+% &= 2 \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} }.
+% \end{align*}
+% Write $\expect_{R\Phi\sigma}$ to denote expectation conditional on $\{R_i\}$, $\{\Phi_i\}$ and $\{\sigma_i\}$.
+Because $g_i$ is a standard normal random variable it follows that $\expect \abs{g_i} = \expect_{R\Phi\sigma} \abs{g_i} = \sqrt{2/\pi}$, and
+% \begin{align*}
+% \expect \sup_{\sabs{\lambda} < \epsilon}  &\abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} } \\
+% &= \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i \in P \cup D} \sigma_i f_{i} \sqrt{\frac{\pi}{2}}\expect_{R\Phi\sigma} \abs{g_i}} \\
+% &\leq \sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{ \sqrt{L}}\sum_{i \in P \cup D} \sigma_i \abs{g_i}  f_{i}},
+% \end{align*}
+% the last line following from Jensen's inequality and by moving $\expect_{R\Phi\sigma}$ outside the supremum similarly to~\eqref{eq:Eoutsidesup}. As $g_i = \sigma_i \abs{g_i}$,
+% \begin{align*}
+% \expect \sup_{\sabs{\lambda} < \epsilon} \abs{W_L(\lambda)}  &\leq 2\sqrt{\frac{\pi}{2}} \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ \frac{1}{\sqrt{L}}\sum_{i\in P \cup D} g_i  f_{i}} \\
+% &= \sqrt{2\pi} \; \expect \sup_{\sabs{\lambda} < \epsilon}  \abs{ X(\lambda) } 
+% \end{align*}
+% as required.
+% \end{IEEEproof}
+
+% \begin{lemma} \label{lem:chaining}
+% For any $\kappa > 0$ there exists $\epsilon > 0$ such that
+% \[
+% \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } < \kappa,
+% \]
+% where $X(\lambda)$ is the Gaussian process defined in~\eqref{eq:ZpsiCondGaussProc}.
+% \end{lemma}
+% \begin{IEEEproof}
+% Without loss of generality, assume that $\epsilon < \tfrac{\pi}{M}$.  Lemma~\ref{lem:chaining2} shows that
+% \begin{equation}\label{eq:supZCK1} 
+% \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}
+% \end{equation}
+% where $K_1$ is a finite, positive constant, and
+% \begin{equation}\label{eq:Cepsdefn}
+% C_\epsilon = \frac{2}{L}\epsilon\sum_{i\in P \cup D}R_i^2 + \frac{4}{L}\sum_{i\in P \cup D}R_i^2 I_{\epsilon}(\Phi_i)
+% \end{equation}
+% where $I_\epsilon$ is the indicator function
+% \[
+% I_{\epsilon}(x) = \begin{cases}
+% 1, & x \in \cup_{k\in\ints}[\psi_k - \epsilon, \psi_k + \epsilon] \\
+% 0, & \text{otherwise},
+% \end{cases}
+% \] 
+% and $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$.  Now,
+% \[
+% \expect C_\epsilon =  2\epsilon\expect R_1^2 + 4 \expect R_i^2 I_{\epsilon}(\Phi_i).
+% \]
+% Recalling that $f(r,\phi)$ is the joint probability density function of $R_i$ and $\Phi_i$,
+% \begin{align*}
+% \expect R_i^2 I_\epsilon(\Phi_i) &= \int_0^{2\pi} \int_{0}^\infty  r^2 f(r,\phi) I_\epsilon(\phi) dr d\phi \\ 
+% &= \int_0^{2\pi}g_2(\phi) I_\epsilon(\phi) d\phi 
+% \end{align*}
+% where the function
+% \[
+% g_2(\phi) = \int_{0}^\infty  r^2 f(r,\phi) dr d\phi 
+% \]
+% From the definition of $I_\epsilon$, and since $\expect \abs{R_i}^2$ is finite,
+% \begin{align*}
+% \expect R_i^2 I_\epsilon(\Phi_i) = \sum_{k=0}^{M-1}\int_{\psi_k-\epsilon}^{\psi_k-\epsilon}g_2(\phi) d\phi \leq K_2 \epsilon
+% \end{align*}
+% for some constant $K_2$. Since $\sqrt{\cdot}$ is a concave function on the positive real line, and since $C_{\epsilon}$ is always positive,
+% \[
+% \expect \sqrt{C_\epsilon} \leq  \sqrt{\expect  C_\epsilon} < \sqrt{\epsilon( K_2 + 2\expect R_1^2)  }
+% \]
+% by Jensen's inequality.  Applying $\expect$ to both sides of~\eqref{eq:supZCK1},
+% \[
+% \expect \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{\tfrac{1}{L}\expect C_\epsilon} < K_1 \sqrt{\epsilon( K_2 + 2\expect R_1^2)}.
+% \]
+% Choosing 
+% \[
+% \epsilon = \frac{\kappa^2}{ K_1^2 ( K_2 + 2\expect R_1^2)}
+% \] 
+% completes the proof.  It remains to prove Lemma~\ref{lem:chaining2}.
+% \end{IEEEproof}
+
+% The proofs of Lemmas~\ref{lem:chaining2}~and~\ref{lem:chaining3} are based on a technique called \emph{chaining} (or \emph{bracketing})~\cite{Dudley_unif_central_lim_th_1999,Ossiander_clt_bracketing_1984,Pollard_asymp_empi_proc_1989,Pollard_new_ways_clts_1986}.
 The proofs here follow those of Pollard~\cite[Section 3]{Pollard_asymp_empi_proc_1989}.
 %Within the remaining proofs, we are interested in expectations conditional on the phase noise sequence $\{\Phi_n\}$, so for the purpose of these lemmas, $\{\Phi_n\}$ is treated as a fixed realisation.
-For notational convenience we use the abbreviation 
-\[
-f_{i}(\lambda) = f_{i}(\lambda, R_i, \Phi_i)
-\]
-in what follows.
+% For notational convenience we use the abbreviation 
+% \[
+% f_{i}(\lambda) = f_{i}(\lambda, R_i, \Phi_i)
+% \]
+% in what follows.
 As in Lemma~\ref{lem:chaining} we assume, without loss of generality, that $\epsilon < \tfrac{\pi}{M}$.
@@
 
-\begin{lemma}\label{lem:chaining2}
-There exists a positive constant $K_1$ such that
-\[
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}.
-\]
-where $C_\epsilon$ is defined in~\eqref{eq:Cepsdefn}
-\end{lemma}
-\begin{IEEEproof}
-Let $B_\epsilon = \{x \in \reals \mid \abs{x} < \epsilon\}$ be the set of real numbers with magnitude less than $\epsilon$.
+% \begin{lemma}\label{lem:chaining2}
+% There exists a positive constant $K_1$ such that
+% \[
+% \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \abs{ X(\lambda) } \leq K_1 \sqrt{C_\epsilon}.
+% \]
+% where $C_\epsilon$ is defined in~\eqref{eq:Cepsdefn}
+% \end{lemma}
+% \begin{IEEEproof}
+% Let $B_\epsilon = \{x \in \reals \mid \abs{x} < \epsilon\}$ be the set of real numbers with magnitude less than $\epsilon$.
 For each non negative integer $k$, let $T_\epsilon(k)$ be a discrete subset of $B_\epsilon$ with the property that for every $\lambda \in B_\epsilon$ there exists $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the pseudometric
-\[
-d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
-\]
-and the pseudometric
-\[
-d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L
-\]
-where 
-\begin{equation}\label{eq:Depsdefn}
-D_\epsilon = \frac{1}{L}\epsilon\sum_{i\in P \cup D}R_i + \frac{2}{L}\sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
-\end{equation}
-We specifically define $T_\epsilon(0)$ to contain only zero.
+% \[
+% d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
+% \]
+% and the pseudometric
+% \[
+% d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L
+% \]
+% where 
+% \begin{equation}\label{eq:Depsdefn}
+% D_\epsilon = \frac{1}{L}\epsilon\sum_{i\in P \cup D}R_i + \frac{2}{L}\sum_{i\in P \cup D}R_i I_{\epsilon}(\Phi_i).
+% \end{equation}
+% We specifically define $T_\epsilon(0)$ to contain only zero.
 Defined this way, $T_\epsilon(0)$ satisfies the inequality above because $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$, as a result of Lemma~\ref{lem:epslmlemma}.
-The existence of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{lem:metricentropy}.
+% The existence of $T_\epsilon(k)$ for every positive integer $k$ will be proved by construction in Lemma~\ref{lem:metricentropy}.
 Lemma~\ref{lem:metricentropy} also shows that the number of elements in $T_\epsilon(k)$ is no more that $K_3 2^k$ where $K_3$ is a constant, independent of $L$, $\epsilon$ and $k$.
 
-Lemma~\ref{lem:limexpTk} shows that
-\[
-\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}.
-\]
-Lemma~\ref{lem:chaining3} shows that
-\[
-\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}
-\]
-The sum above converges as $k\rightarrow\infty$ so lemma holds with
-\[
-K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
-\]
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:epslmlemma}
-Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.
-\end{lemma}
-\begin{IEEEproof}
-We prove the lemma for $d_2$.  The proof for $d_1$ is similar.  Observe that $d_2(\lambda, 0) = c_1 + c_2$ where 
- \[
- c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) },
- \] 
- \[
- c_2 = \sum_{i \in D} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
- \]
- As $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
- \[
- \sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i)} \leq K\sabs{\lambda} = \sabs{\lambda} < \epsilon
- \]
- for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let 
-\[
-S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}.
-\] 
-Now $c_2 = c_3 + c_4$ where 
- \[
- c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}},
- \]
- \[
- c_4 =  \sum_{i \in D, i \notin S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
- \]
-Observe that $I_\epsilon(\Phi_i) = 1$ if and only if $\abs{\fracpart{\Phi_i}} \geq \tfrac{\pi}{M} - \epsilon$.  So, if $i \notin S$ and $\sabs{\lambda} < \epsilon$, then
-\begin{align*}
-\round{\lambda + \Phi_i} &= \round{\lambda + \fracpart{\Phi_i} + \round{\Phi_i}} \\ 
-&= \round{\lambda + \fracpart{\Phi_i}} + \round{\Phi_i} = \round{\Phi_i}
-\end{align*}
-because
-\[
-\sabs{\lambda + \fracpart{\Phi_i}} < \epsilon + \tfrac{\pi}{M} - \epsilon = \tfrac{\pi}{M}
-\] 
-and therefore $\round{\lambda + \fracpart{\Phi_i}} = 0$.  Now,
- \begin{align*}
- \vert\sin&\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}\vert \\
- &= \sabs{\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\Phi_i - \round{\Phi_i})} \\
- &\leq \sabs{\lambda} < \epsilon
- \end{align*}
-and so, 
-\[
-c_4 \leq \epsilon \sum_{i \in D, i \notin S} R_i \leq \epsilon \sum_{i \in D} R_i.
-\]
-Otherwise, if $i \in S$, then $\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}} \leq 2$ because $\sin(\cdot) \leq 1$.  Thus 
-\[
-c_3 \leq 2 \sum_{i \in D, i \in S} R_i \leq 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i).
-\]
-Now,
-\begin{align*}
-d_2(\lambda, 0) &= c_1 + c_2 = c_1 + c_3 + c_4 \\
-&\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i) = D_\epsilon L
-\end{align*}
-as required
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:limexpTk} With $T_\epsilon(k)$ defined as in Lemma~\ref{lem:chaining2},
-\[
-\lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $b_2(\lambda)$ be a function that maps $\lambda \in B_\epsilon$ to a point in $T_\epsilon(k)$ such that  $d_2(\lambda,b_2(\lambda)) \leq 2^{-k}D_\epsilon L$.
+% Lemma~\ref{lem:limexpTk} shows that
+% \[
+% \lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}.
+% \]
+% Lemma~\ref{lem:chaining3} shows that
+% \[
+% \expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}
+% \]
+% The sum above converges as $k\rightarrow\infty$ so lemma holds with
+% \[
+% K_1 = \sum_{i=1}^\infty \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
+% \]
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:epslmlemma}
+% Both $d_1(\lambda, 0) \leq C_\epsilon L $ and $d_2(\lambda, 0) \leq D_\epsilon L$ for all $\lambda \in B_\epsilon$.
+% \end{lemma}
+% \begin{IEEEproof}
+% We prove the lemma for $d_2$.  The proof for $d_1$ is similar.  Observe that $d_2(\lambda, 0) = c_1 + c_2$ where 
+%  \[
+%  c_1 = \sum_{i \in P} R_i\sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i) },
+%  \] 
+%  \[
+%  c_2 = \sum_{i \in D} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+%  \]
+%  As $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+%  \[
+%  \sabs{\sin(\lambda + \Phi_i) - \sin(\Phi_i)} \leq K\sabs{\lambda} = \sabs{\lambda} < \epsilon
+%  \]
+%  for all $i \in P \cup D$, and therefore $c_1 \leq \epsilon \sum_{i \in P} R_i$.  Let 
+% \[
+% S = \{ i \in D \mid I_\epsilon(\Phi_i) = 1\}.
+% \] 
+% Now $c_2 = c_3 + c_4$ where 
+%  \[
+%  c_3 = \sum_{i \in D, i \in S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}},
+%  \]
+%  \[
+%  c_4 =  \sum_{i \in D, i \notin S} R_i \sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}}.
+%  \]
+% Observe that $I_\epsilon(\Phi_i) = 1$ if and only if $\abs{\fracpart{\Phi_i}} \geq \tfrac{\pi}{M} - \epsilon$.  So, if $i \notin S$ and $\sabs{\lambda} < \epsilon$, then
+% \begin{align*}
+% \round{\lambda + \Phi_i} &= \round{\lambda + \fracpart{\Phi_i} + \round{\Phi_i}} \\ 
+% &= \round{\lambda + \fracpart{\Phi_i}} + \round{\Phi_i} = \round{\Phi_i}
+% \end{align*}
+% because
+% \[
+% \sabs{\lambda + \fracpart{\Phi_i}} < \epsilon + \tfrac{\pi}{M} - \epsilon = \tfrac{\pi}{M}
+% \] 
+% and therefore $\round{\lambda + \fracpart{\Phi_i}} = 0$.  Now,
+%  \begin{align*}
+%  \vert\sin&\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}\vert \\
+%  &= \sabs{\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\Phi_i - \round{\Phi_i})} \\
+%  &\leq \sabs{\lambda} < \epsilon
+%  \end{align*}
+% and so, 
+% \[
+% c_4 \leq \epsilon \sum_{i \in D, i \notin S} R_i \leq \epsilon \sum_{i \in D} R_i.
+% \]
+% Otherwise, if $i \in S$, then $\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\Phi_i}} \leq 2$ because $\sin(\cdot) \leq 1$.  Thus 
+% \[
+% c_3 \leq 2 \sum_{i \in D, i \in S} R_i \leq 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i).
+% \]
+% Now,
+% \begin{align*}
+% d_2(\lambda, 0) &= c_1 + c_2 = c_1 + c_3 + c_4 \\
+% &\leq \epsilon \sum_{i \in P \cup D} R_i + 2 \sum_{i \in P \cup D} R_i I_\epsilon(\Phi_i) = D_\epsilon L
+% \end{align*}
+% as required
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:limexpTk} With $T_\epsilon(k)$ defined as in Lemma~\ref{lem:chaining2},
+% \[
+% \lim_{k\rightarrow\infty}\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda)}
+% \]
+% \end{lemma}
+% \begin{IEEEproof}
+% Let $b_2(\lambda)$ be a function that maps $\lambda \in B_\epsilon$ to a point in $T_\epsilon(k)$ such that  $d_2(\lambda,b_2(\lambda)) \leq 2^{-k}D_\epsilon L$.
 The existence of $b_2$ is guaranteed by the definition of $T_\epsilon(k-1)$.
 So, for all $\lambda \in B_\epsilon$,
-\begin{align*}
-\sabs{X(\lambda) - X(b_2(\lambda))} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda) - f_i(b_2(\lambda)))} \\
-&\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
-&\leq \frac{1}{\sqrt{L}} \sum_{i\in P \cup D} g_\text{max} \sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} d_2(\lambda,b_2(\lambda)) \\
-&\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L,
-\end{align*}
-where $g_\text{max} = \sup_{i\in P\cup D} \abs{g_i}$. The bound applies uniformly over $\lambda \in B_\epsilon$, so
-\[
-\sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L.
-\]
-Applying $\expect_{R,\Phi}$ to both sides,
-\[
-\expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq K_4 2^{-k} D_\epsilon L,
-\]
-where, as a result of Lemma~\ref{lem:maxineq},
-\[
-K_4 = \tfrac{1}{\sqrt{L}}\expect_{R,\Phi} g_\text{max} \leq 3 \sqrt{\frac{\log L}{L}}
-\]
-is independent of $\epsilon$ and $k$.  Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
-\[
-\sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)},
-\]
-and so,
-\begin{align*}
-0 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \nonumber \\
-&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \nonumber \\
-&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \big( \sabs{X(\lambda)} - \sabs{X(b_2(\lambda))} \big) \nonumber \\
-&\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda) - X(b_2(\lambda))} \nonumber \\
-&\leq K_4 2^{-k} D_\epsilon L. \label{eq:Deps2nktozero}
-\end{align*}
-The second line in the inequality above follows because,
-\[
-\sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)},
-\]
-since $b_2(\lambda) \in T_\epsilon(k)$.  The lemma follows because $K_4 2^{-k} D_\epsilon L$ goes to zero as $k\rightarrow\infty$.
-\end{IEEEproof}
-
-% \begin{lemma}\label{lem:epslmlemma}
-% Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
 % \begin{align*}
-% d(\lambda, \lambda^*) &= \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \\
-% &\leq R_i^2 L \left( (\lambda - \lambda^*)^2 + 4 C_\epsilon \right).
+% \sabs{X(\lambda) - X(b_2(\lambda))} &= \abs{\frac{1}{\sqrt{L}}\sum_{i\in P \cup D}g_i(f_i(\lambda) - f_i(b_2(\lambda)))} \\
+% &\leq  \frac{1}{\sqrt{L}}\sum_{i\in P \cup D}\abs{g_i}\sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
+% &\leq \frac{1}{\sqrt{L}} \sum_{i\in P \cup D} g_\text{max} \sabs{f_i(\lambda) - f_i(b_2(\lambda))} \\
+% &\leq \frac{g_\text{max}}{\sqrt{L}} d_2(\lambda,b_2(\lambda)) \\
+% &\leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L,
 % \end{align*}
-% \end{lemma}
-
-% \begin{lemma}
-% If $i \in P$, then $\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2.$
+% where $g_\text{max} = \sup_{i\in P\cup D} \abs{g_i}$. The bound applies uniformly over $\lambda \in B_\epsilon$, so
+% \[
+% \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq \frac{g_\text{max}}{\sqrt{L}} 2^{-k} D_\epsilon L.
+% \]
+% Applying $\expect_{R,\Phi}$ to both sides,
+% \[
+% \expect_{R,\Phi} \sup_{\sabs{\lambda} < \epsilon} \sabs{X(\lambda) - X(b_2(\lambda))} \leq K_4 2^{-k} D_\epsilon L,
+% \]
+% where, as a result of Lemma~\ref{lem:maxineq},
+% \[
+% K_4 = \tfrac{1}{\sqrt{L}}\expect_{R,\Phi} g_\text{max} \leq 3 \sqrt{\frac{\log L}{L}}
+% \]
+% is independent of $\epsilon$ and $k$.  Because $T_\epsilon(k)$ is a subset of $B_\epsilon$,
+% \[
+% \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)},
+% \]
+% and so,
+% \begin{align*}
+% 0 &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)} \nonumber \\
+% &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda)} - \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \nonumber \\
+% &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \big( \sabs{X(\lambda)} - \sabs{X(b_2(\lambda))} \big) \nonumber \\
+% &\leq \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(\lambda) - X(b_2(\lambda))} \nonumber \\
+% &\leq K_4 2^{-k} D_\epsilon L. \label{eq:Deps2nktozero}
+% \end{align*}
+% The second line in the inequality above follows because,
+% \[
+% \sup_{\sabs{\lambda} < \epsilon} \expect_{R,\Phi} \sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k)} \expect_{R,\Phi} \sabs{X(\lambda)},
+% \]
+% since $b_2(\lambda) \in T_\epsilon(k)$.  The lemma follows because $K_4 2^{-k} D_\epsilon L$ goes to zero as $k\rightarrow\infty$.
+% \end{IEEEproof}
+
+% % \begin{lemma}\label{lem:epslmlemma}
+% % Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$. Then
+% % \begin{align*}
+% % d(\lambda, \lambda^*) &= \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda^*) \big)^2 \\
+% % &\leq R_i^2 L \left( (\lambda - \lambda^*)^2 + 4 C_\epsilon \right).
+% % \end{align*}
+% % \end{lemma}
+
+% % \begin{lemma}
+% % If $i \in P$, then $\big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2.$
+% % \end{lemma}
+% % \begin{IEEEproof} 
+% % We have
+% % \[
+% % \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 = R_i^2\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)}^2
+% % \]
+% % and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+% % \[
+% % \sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)} \leq K\sabs{\lambda - \lambda^*} = \sabs{\lambda - \lambda^*}.
+% % \]
+% % \end{IEEEproof}
+
+% % \begin{lemma}
+% % Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$.  If $i \in D$, then
+% % \begin{equation}\label{eq:fismall}
+% % \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2
+% % \end{equation}
+% % when $\Phi_i \notin S_\epsilon$, otherwise,
+% % \begin{equation}\label{eq:fibig}
+% % \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq 4 R_i^2.
+% % \end{equation}
+% % \end{lemma}
+% % \begin{IEEEproof}
+% % Because $\sin(\cdot) \leq 1$ it follows that
+% % \begin{align*}
+% % \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 &= R_i^2\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}}^2 \\
+% % &\leq 4 R_i^2
+% % \end{align*}
+% % for any $\Phi_i \in \reals$, and so~\eqref{eq:fibig} holds.  If $\Phi_i \notin S_\epsilon$ then $\Phi_i \in (\psi_{k-1} + \epsilon, \psi_{k} - \epsilon)$ for some $k\in\ints$ and 
+% % \[
+% % \fracpart{\Phi_i + \lambda} = \Phi_i + \lambda - \tfrac{2\pi}{M}k
+% % \] 
+% % and
+% % \[
+% % \fracpart{\Phi_i + \lambda^*} = \Phi_i + \lambda^* - \tfrac{2\pi}{M}k 
+% % \]
+% % and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
+% % \begin{align*}
+% % \vert \sin&\fracpart{\Phi_i + \lambda} - \sin\fracpart{\Phi_i + \lambda^*}\vert \\
+% % &=  \sabs{\sin(\Phi_i + \lambda - \tfrac{2\pi}{M}k) - \sin(\Phi_i + \lambda^* - \tfrac{2\pi}{M}k)} \\
+% % &\leq K \sabs{\lambda - \lambda^*} \\
+% % &= \sabs{\lambda - \lambda^*}.
+% % \end{align*}
+% % Thus~\eqref{eq:fismall} holds. 
+% % \end{IEEEproof}
+
+% \begin{lemma}\label{lem:chaining3}(Chaining)
+% For all positive integers $k$,
+% \[
+% \expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k) } \abs{ X(\lambda) } \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
+% \]
 % \end{lemma}
 % \begin{IEEEproof}
-% We have
-% \[
-% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 = R_i^2\sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)}^2
-% \]
-% and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
-% \[
-% \sabs{\sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i)} \leq K\sabs{\lambda - \lambda^*} = \sabs{\lambda - \lambda^*}.
-% \]
+% Let $b_1(\lambda)$ map each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
+The existence of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
+By the triangle inequality
+% \[
+% \sabs{X(\lambda)} \leq \sabs{X(b_2(\lambda))} + \sabs{X(\lambda) - X(b_2(\lambda))}  
+% \]
+% and taking supremums on both sides,
+% \begin{align}
+% &\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \nonumber \\
+% &\leq \sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}   \nonumber \\
+% &\leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}, \label{eq:Xb2recpreexp}
+% \end{align}
+% the last line because $b_2(\lambda) \in T_\epsilon(k-1)$ and so
+% \[
+% \sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)}.
+% \]
+% Conditional on $\{R_i\}$ and $\{\Phi_i\}$ the random variable
+% \[
+% U(\lambda) = X(\lambda) - X(b_2(\lambda))
+% \]
+% has zero mean and is normally distributed with variance
+% \begin{align*}
+% \sigma_U^2 &= \expect_{R,\Phi} \frac{1}{L} \sum_{i \in P \cup D} g_i^2 \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
+% &=  \frac{1}{L} \sum_{i \in P \cup D} \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
+% &= d(\lambda, b_2(\lambda)) \leq 2^{1-k} C_\epsilon,
+% \end{align*}
+% since $\expect_{R,\Phi}g_i^2 = 1$.  So, from Lemma~\ref{lem:maxineq},
+% \begin{align*}
+% \expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{U(\lambda)} &\leq 3 \sqrt{ 2^{1-k} C_\epsilon \log \abs{T_\epsilon(k)} } \\
+% &\leq \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}.
+% \end{align*}
+% Taking expectations on both sides of~\eqref{eq:Xb2recpreexp},
+% \begin{align*}
+% \expect_{R,\Phi}&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \\
+% &\leq \expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}} ,
+% \end{align*}
+% which involves a recursion in $k$.  By unravelling the recursion and using that $T_\epsilon(k)$ contains only zero, and therefore
+% \[
+% \expect_{R,\Phi} \sup_{\lambda \in T_k(0)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sabs{X(0)} = 0
+% \]
+% we obtain,
+% \[
+% \expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \leq  \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}  
+% \]
+% as required.
 % \end{IEEEproof}
 
-% \begin{lemma}
-% Suppose $\sabs{\lambda} < \epsilon$ and $\sabs{\lambda^*} < \epsilon$ with $\epsilon < \tfrac{\pi}{M}$.  If $i \in D$, then
-% \begin{equation}\label{eq:fismall}
-% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq R_i^2(\lambda - \lambda^*)^2
-% \end{equation}
-% when $\Phi_i \notin S_\epsilon$, otherwise,
-% \begin{equation}\label{eq:fibig}
-% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 \leq 4 R_i^2.
-% \end{equation}
+% \begin{lemma}\label{lem:maxineq}(Maximal inequalities)
+% Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
+% \[
+% \expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}
+% \]
+% where $\log N$ is the natural logarithm of $N$.
 % \end{lemma}
 % \begin{IEEEproof}
-% Because $\sin(\cdot) \leq 1$ it follows that
-% \begin{align*}
-% \big(f_{i}(\lambda) - f_{i}(\lambda^*)\big)^2 &= R_i^2\sabs{\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}}^2 \\
-% &\leq 4 R_i^2
-% \end{align*}
-% for any $\Phi_i \in \reals$, and so~\eqref{eq:fibig} holds.  If $\Phi_i \notin S_\epsilon$ then $\Phi_i \in (\psi_{k-1} + \epsilon, \psi_{k} - \epsilon)$ for some $k\in\ints$ and 
-% \[
-% \fracpart{\Phi_i + \lambda} = \Phi_i + \lambda - \tfrac{2\pi}{M}k
-% \] 
-% and
-% \[
-% \fracpart{\Phi_i + \lambda^*} = \Phi_i + \lambda^* - \tfrac{2\pi}{M}k 
-% \]
-% and, since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$,
-% \begin{align*}
-% \vert \sin&\fracpart{\Phi_i + \lambda} - \sin\fracpart{\Phi_i + \lambda^*}\vert \\
-% &=  \sabs{\sin(\Phi_i + \lambda - \tfrac{2\pi}{M}k) - \sin(\Phi_i + \lambda^* - \tfrac{2\pi}{M}k)} \\
-% &\leq K \sabs{\lambda - \lambda^*} \\
-% &= \sabs{\lambda - \lambda^*}.
-% \end{align*}
-% Thus~\eqref{eq:fismall} holds. 
+% This result is well known, see for example~\cite[Section~3]{Pollard_asymp_empi_proc_1989}  
 % \end{IEEEproof}
 
-\begin{lemma}\label{lem:chaining3}(Chaining)
-For all positive integers $k$,
-\[
-\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k) } \abs{ X(\lambda) } \leq \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}.
-\]
-\end{lemma}
-\begin{IEEEproof}
-Let $b_1(\lambda)$ map each $\lambda \in T_\epsilon(k)$ to an element in $T_\epsilon(k-1)$ such that $d(\lambda, b_1(\lambda)) \leq 2^{1-k}C_\epsilon L$.
-The existence of $b_1$ is guaranteed by the definition of $T_\epsilon(k)$.
-By the triangle inequality
-\[
-\sabs{X(\lambda)} \leq \sabs{X(b_2(\lambda))} + \sabs{X(\lambda) - X(b_2(\lambda))}  
-\]
-and taking supremums on both sides,
-\begin{align}
-&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \nonumber \\
-&\leq \sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}   \nonumber \\
-&\leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sup_{\lambda \in T_\epsilon(k)} \sabs{X(\lambda) - X(b_2(\lambda))}, \label{eq:Xb2recpreexp}
-\end{align}
-the last line because $b_2(\lambda) \in T_\epsilon(k-1)$ and so
-\[
-\sup_{\lambda \in T_\epsilon(k)}\sabs{X(b_2(\lambda))} \leq \sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)}.
-\]
-Conditional on $\{R_i\}$ and $\{\Phi_i\}$ the random variable
-\[
-U(\lambda) = X(\lambda) - X(b_2(\lambda))
-\]
-has zero mean and is normally distributed with variance
-\begin{align*}
-\sigma_U^2 &= \expect_{R,\Phi} \frac{1}{L} \sum_{i \in P \cup D} g_i^2 \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
-&=  \frac{1}{L} \sum_{i \in P \cup D} \big( f_i(\lambda) - f_i(b_2(\lambda)) \big)^2 \\
-&= d(\lambda, b_2(\lambda)) \leq 2^{1-k} C_\epsilon,
-\end{align*}
-since $\expect_{R,\Phi}g_i^2 = 1$.  So, from Lemma~\ref{lem:maxineq},
-\begin{align*}
-\expect_{R,\Phi} \sup_{\lambda \in T_\epsilon(k)} \sabs{U(\lambda)} &\leq 3 \sqrt{ 2^{1-k} C_\epsilon \log \abs{T_\epsilon(k)} } \\
-&\leq \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}}.
-\end{align*}
-Taking expectations on both sides of~\eqref{eq:Xb2recpreexp},
-\begin{align*}
-\expect_{R,\Phi}&\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \\
-&\leq \expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k-1)}\sabs{X(\lambda)} + \sqrt{C_\epsilon} \frac{5\sqrt{k \log 2 + \log K_3}}{2^{k/2}} ,
-\end{align*}
-which involves a recursion in $k$.  By unravelling the recursion and using that $T_\epsilon(k)$ contains only zero, and therefore
-\[
-\expect_{R,\Phi} \sup_{\lambda \in T_k(0)} \sabs{X(\lambda)} = \expect_{R,\Phi} \sabs{X(0)} = 0
-\]
-we obtain,
-\[
-\expect_{R,\Phi}\sup_{\lambda \in T_\epsilon(k)}\sabs{X(\lambda)} \leq  \sqrt{C_\epsilon} \sum_{i=1}^k \frac{5\sqrt{i \log 2 + \log K_3}}{2^{i/2}}  
-\]
-as required.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:maxineq}(Maximal inequalities)
-Suppose $X_1, \dots, X_N$ are zero mean Gaussian random variables each with variance less than some positive constant $K$, then
-\[
-\expect \sup_{n = 1, \dots, N} \abs{X_n} \leq 3  \sqrt{K \log N}
-\]
-where $\log N$ is the natural logarithm of $N$.
-\end{lemma}
-\begin{IEEEproof}
-This result is well known, see for example~\cite[Section~3]{Pollard_asymp_empi_proc_1989}  
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:metricentropy}(Covering numbers)
-For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset B_\epsilon$ with the property that, for every $\lambda \in B_\epsilon$, there is a $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the pseudometric
-\[
-d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
-\]
-and the pseudometric
-\[
-d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L,
-\]
-where $C_\epsilon$ and $D_\epsilon$ are defined in~\eqref{eq:Cepsdefn} and~\eqref{eq:Depsdefn}.
+% \begin{lemma}\label{lem:metricentropy}(Covering numbers)
+% For $k \in \ints$ there exists a discrete set $T_\epsilon(k) \subset B_\epsilon$ with the property that, for every $\lambda \in B_\epsilon$, there is a $\lambda_1$ and $\lambda_2$ in $T_\epsilon(k)$ such that the pseudometric
+% \[
+% d_1(\lambda, \lambda_1) = \sum_{i \in P \cup D} \big( f_{i}(\lambda) - f_{i}(\lambda_1) \big)^2 \leq 2^{-k} C_\epsilon L,
+% \]
+% and the pseudometric
+% \[
+% d_2(\lambda, \lambda_2) = \sum_{i \in P \cup D} \sabs{f_{i}(\lambda) - f_{i}(\lambda_2)} \leq 2^{-k} D_\epsilon L,
+% \]
+% where $C_\epsilon$ and $D_\epsilon$ are defined in~\eqref{eq:Cepsdefn} and~\eqref{eq:Depsdefn}.
 The number of elements in $T_\epsilon(k)$ is no more than $K_3 2^{k}$ where $K_3$ is a positive constant, independent of $L$, $\epsilon$ and $k$.
-\end{lemma}
-\begin{IEEEproof}
-% We will first construct a set $T^2_\epsilon(k)$ containing no more than $K_{23}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_2 \in T^2_\epsilon(k)$ such that $d_2(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{23}$ a constant, independent of $L$, $\epsilon$ and $k$.
+% \end{lemma}
+% \begin{IEEEproof}
+% % We will first construct a set $T^2_\epsilon(k)$ containing no more than $K_{23}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_2 \in T^2_\epsilon(k)$ such that $d_2(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{23}$ a constant, independent of $L$, $\epsilon$ and $k$.
 Using a similar construction we find a set $T^1_\epsilon(k)$ containing no more than $K_{13}2^{k}$ elements and so that for every $\lambda \in B_\epsilon$ there exists $\lambda_1 \in T^1_\epsilon(k)$ such that $d_1(\lambda, \lambda_2) \leq 2^{-k} D_\epsilon L$ with $K_{13}$ a constant, independent of $L$, $\epsilon$ and $k$.
@@
 
-% We will only give a construction for $T^2_\epsilon(k)$, the construction for $T^1_\epsilon(k)$ being similar.  First observe that
-% \[
-% f_{i}(\lambda) - f_{i}(\lambda_2) = \begin{cases}
-% R_i\big( \sin(\lambda + \Phi_i) - \sin(\lambda_2 + \Phi_i) \big) & i \in P \\
-% R_i\big(\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda_2 + \Phi_i}\big) & i \in D
-% \end{cases}
-% \]
-% Let 
-% \[
-% U = \{\epsilon(\ell 2^{-k} - 1) \mid \ell = 0, 1, \dots, 2^{k+1} \} 
-% \]
-% be a set containing $2^{k+1} + 1$ elements uniformly spaced over $B_\epsilon$.
+% % We will only give a construction for $T^2_\epsilon(k)$, the construction for $T^1_\epsilon(k)$ being similar.  First observe that
+% % \[
+% % f_{i}(\lambda) - f_{i}(\lambda_2) = \begin{cases}
+% % R_i\big( \sin(\lambda + \Phi_i) - \sin(\lambda_2 + \Phi_i) \big) & i \in P \\
+% % R_i\big(\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda_2 + \Phi_i}\big) & i \in D
+% % \end{cases}
+% % \]
+% % Let 
+% % \[
+% % U = \{\epsilon(\ell 2^{-k} - 1) \mid \ell = 0, 1, \dots, 2^{k+1} \} 
+% % \]
+% % be a set containing $2^{k+1} + 1$ elements uniformly spaced over $B_\epsilon$.
 For every $\lambda \in B_\epsilon$ there exists a $\lambda^* \in U$ such that $\abs{\lambda - \lambda^*} < \epsilon 2^{-k}$.
 Since $\sin(\cdot)$ is Lipshitz continuous with constant $K=1$, it follows that if $i \in P$ then,
-% Let 
-% \begin{align*}
-% \abs{f_{i}(\lambda) - f_{i}(\lambda_2)} &= R_i\abs{ \sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i) } \\
-% &\leq R_i\sabs{\lambda - \lambda^*} \\
-% &< \epsilon 2^{-k} R_i.
-% \end{align*}
-% Thus
-% \[
-% \sum_{i \in P} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i\in P} R_i.
-% \]
-% As in Lemma~\ref{lem:epslmlemma} let $S = \{ i \in D \mid I_\epsilon(\Phi_i) \}$. Now, if $i \in D$ but $i \notin S$ then 
-% \[
-% \round{\lambda + \Phi_i} = \round{\lambda^* + \Phi_i} = \round{\Phi_i},
-% \] 
-% so
-%  \begin{align*}
-%  \vert f_{i}(\lambda) - &f_{i}(\lambda^*)\vert \\
-% &= R_i\vert\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}\vert \\
-%  &= R_i\vert\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\lambda^* + \Phi_i - \round{\Phi_i})\vert  \\
-%  &\leq R_i \sabs{\lambda - \lambda^*} \\
-% &< \epsilon 2^{-k} R_i.
-%  \end{align*}
-% Thus,
-% \[
-% \sum_{i \in D, i \notin S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i \in D, i \notin S} R_i 
-% \]
-
-% To handle the case when $i \in S$ we need to add some extra elements to the set $U$.  Let 
-% \[
-% r_i = \begin{cases}
-% \tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} > 0 \\
-% -\tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} \leq 0
-% \end{cases}
-% \]
-% and let $\sigma$ be a permutation of the indices in $S$ such that $r_i$ are in ascending order, that is,
-% \begin{equation}\label{eq:sigmasortind}
-% r_{\sigma(i)} \leq r_{\sigma(k)}
-% \end{equation}
-% whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{S}-1\}$.  Now,
-% \begin{align*}
-% \sum_{i\in S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} = \sum_{i = 0}^{\abs{S}-1} \sabs{ f_{\sigma(i)}(\lambda) - f_{\sigma(i)}(\lambda^*)}.
-% \end{align*}
-% If $\epsilon > \lambda \geq r_{\sigma(k)} > 0$ for some $k$, then
-% \[
-% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = \tfrac{\pi}{M}
-% \]
-% for all $i \leq k$, whilst
-% \[
-% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
-% \]
-% for all $i > k$.  Otherwise if $-\epsilon < \lambda < r_{\sigma(k)} \leq 0$ for some $k$, then
-% \[
-% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + -\tfrac{\pi}{M} - r_{\sigma(i)}} = -\tfrac{\pi}{M}
-% \]
-% for all $i \leq k$, whilst
-% \[
-% \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
-% \]
-% for all $i > k$.
-
-% Let $R = \sum_{i \in S} R_i$ and let $d(\ell)$ be the largest integer such that
-% \[
-% \sum_{i=0}^{d(\ell)} R_{\sigma(i)} \leq \frac{\ell}{2^k} R.
-% \]
-% for integers $\ell = 1, \dots, 2^k$.  Define the intervals
-% \[
-% L_\ell = \in [r_{d(\ell)}, r_{d(\ell+1)})
-% \]
-% where $r_{0} = -\epsilon$ and $r_{2^k+1} = \epsilon$.
+% % Let 
+% % \begin{align*}
+% % \abs{f_{i}(\lambda) - f_{i}(\lambda_2)} &= R_i\abs{ \sin(\lambda + \Phi_i) - \sin(\lambda^* + \Phi_i) } \\
+% % &\leq R_i\sabs{\lambda - \lambda^*} \\
+% % &< \epsilon 2^{-k} R_i.
+% % \end{align*}
+% % Thus
+% % \[
+% % \sum_{i \in P} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i\in P} R_i.
+% % \]
+% % As in Lemma~\ref{lem:epslmlemma} let $S = \{ i \in D \mid I_\epsilon(\Phi_i) \}$. Now, if $i \in D$ but $i \notin S$ then 
+% % \[
+% % \round{\lambda + \Phi_i} = \round{\lambda^* + \Phi_i} = \round{\Phi_i},
+% % \] 
+% % so
+% %  \begin{align*}
+% %  \vert f_{i}(\lambda) - &f_{i}(\lambda^*)\vert \\
+% % &= R_i\vert\sin\fracpart{\lambda + \Phi_i} - \sin\fracpart{\lambda^* + \Phi_i}\vert \\
+% %  &= R_i\vert\sin(\lambda + \Phi_i - \round{\Phi_i}) - \sin(\lambda^* + \Phi_i - \round{\Phi_i})\vert  \\
+% %  &\leq R_i \sabs{\lambda - \lambda^*} \\
+% % &< \epsilon 2^{-k} R_i.
+% %  \end{align*}
+% % Thus,
+% % \[
+% % \sum_{i \in D, i \notin S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} < \epsilon 2^{-k} \sum_{i \in D, i \notin S} R_i 
+% % \]
+
+% % To handle the case when $i \in S$ we need to add some extra elements to the set $U$.  Let 
+% % \[
+% % r_i = \begin{cases}
+% % \tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} > 0 \\
+% % -\tfrac{\pi}{M} - \fracpart{\Phi_i} & \fracpart{\Phi_i} \leq 0
+% % \end{cases}
+% % \]
+% % and let $\sigma$ be a permutation of the indices in $S$ such that $r_i$ are in ascending order, that is,
+% % \begin{equation}\label{eq:sigmasortind}
+% % r_{\sigma(i)} \leq r_{\sigma(k)}
+% % \end{equation}
+% % whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{S}-1\}$.  Now,
+% % \begin{align*}
+% % \sum_{i\in S} \sabs{f_{i}(\lambda) - f_{i}(\lambda^*)} = \sum_{i = 0}^{\abs{S}-1} \sabs{ f_{\sigma(i)}(\lambda) - f_{\sigma(i)}(\lambda^*)}.
+% % \end{align*}
+% % If $\epsilon > \lambda \geq r_{\sigma(k)} > 0$ for some $k$, then
+% % \[
+% % \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = \tfrac{\pi}{M}
+% % \]
+% % for all $i \leq k$, whilst
+% % \[
+% % \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
+% % \]
+% % for all $i > k$.  Otherwise if $-\epsilon < \lambda < r_{\sigma(k)} \leq 0$ for some $k$, then
+% % \[
+% % \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + -\tfrac{\pi}{M} - r_{\sigma(i)}} = -\tfrac{\pi}{M}
+% % \]
+% % for all $i \leq k$, whilst
+% % \[
+% % \round{\lambda + \fracpart{\Phi_{\sigma(i)}}} = \round{\lambda + \tfrac{\pi}{M} - r_{\sigma(i)}} = 0
+% % \]
+% % for all $i > k$.
+
+% % Let $R = \sum_{i \in S} R_i$ and let $d(\ell)$ be the largest integer such that
+% % \[
+% % \sum_{i=0}^{d(\ell)} R_{\sigma(i)} \leq \frac{\ell}{2^k} R.
+% % \]
+% % for integers $\ell = 1, \dots, 2^k$.  Define the intervals
+% % \[
+% % L_\ell = \in [r_{d(\ell)}, r_{d(\ell+1)})
+% % \]
+% % where $r_{0} = -\epsilon$ and $r_{2^k+1} = \epsilon$.
 Then $L_0, L_1, \dots, L_{2^k}$ partition the closure of $B_\epsilon$.
 If $\lambda$ and $\lambda^*$ are from interval $L_\ell$, then for all $i \in \{0, 1, \dots, \abs{S}-1\}$ less than $d(\ell)$,
-% \begin{align*}
-% \fracpart{\lambda + \Phi_{\sigma(i)}} &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}} \\
-% &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}}
-% \end{align*}
-% \begin{align*}
-% \vert f_{\sigma(i)}(\lambda) - &f_{\sigma(i)}(\lambda^*) \vert \\
-% &= R_{\sigma(i)}\vert\sin\fracpart{\lambda + \Phi_{\sigma(i)}} - \sin\fracpart{\lambda^* + \Phi_{\sigma(i)}}\vert 
-% \end{align*}
-BLERG
-
-\end{IEEEproof}
-
-% \subsection{Proof of Proposition~\ref{prop:gg2cont}} \label{appendix:prop1proof}
-
-% \begin{lemma}
-% Let $W$ be a circularly symmetric random variable with constant magnitude $Z = \abs{W} = c > 0$ and phase uniformly distributed on $[-\pi,\pi)$.
+% % \begin{align*}
+% % \fracpart{\lambda + \Phi_{\sigma(i)}} &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}} \\
+% % &= \fracpart{\lambda + \fracpart{\Phi_{\sigma(i)}}}
+% % \end{align*}
+% % \begin{align*}
+% % \vert f_{\sigma(i)}(\lambda) - &f_{\sigma(i)}(\lambda^*) \vert \\
+% % &= R_{\sigma(i)}\vert\sin\fracpart{\lambda + \Phi_{\sigma(i)}} - \sin\fracpart{\lambda^* + \Phi_{\sigma(i)}}\vert 
+% % \end{align*}
+% BLERG
+
+% \end{IEEEproof}
+
+% % \subsection{Proof of Proposition~\ref{prop:gg2cont}} \label{appendix:prop1proof}
+
+% % \begin{lemma}
+% % Let $W$ be a circularly symmetric random variable with constant magnitude $Z = \abs{W} = c > 0$ and phase uniformly distributed on $[-\pi,\pi)$.
 Let $R \geq 0$ and $\Phi \in [-\pi, \pi)$ be real random variables such that $Re^{j\Phi} = 1 + W$.
 Then the pdf of $\Phi$ is
-% \begin{equation} \label{eq:fphicleq1}
-% f_{\Phi}(\phi) = \begin{cases} 
-% \frac{\cos(\Phi)}{\pi c} &  [-\tfrac{\pi}{2}, \tfrac{\pi}{2}] \\
-% 0 & \text{otherwise}
-% \end{cases}
-% \end{equation}
-% when $0 < c \leq 1$, and
-% \begin{equation} \label{eq:fphicg1}
-% f_{\Phi}(\phi) = \frac{\cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}}{2\pi c}
-% \end{equation}
-% when $c > 1$.
-% \end{lemma}
-% \begin{IEEEproof}
-% We make use of delta `function' notation and write the `pdf' of $Z$ as $\delta(z - c)$.  Using~\eqref{eq:pdfRPhi} the pdf of $\Phi$ is
-% \[
-% f_{\Phi}(\phi) = \int_{0}^{\infty} r \frac{\delta(z - c)}{ 2\pi z} dr = \sum_{r} \frac{r}{2\pi c}
-% \]
-% where $z^2 = r^2 - 2r\cos(\phi) + 1$ and the sum is over the positive real solutions of $c^2 = r^2 - 2r\cos(\phi) + 1$.
+% % \begin{equation} \label{eq:fphicleq1}
+% % f_{\Phi}(\phi) = \begin{cases} 
+% % \frac{\cos(\Phi)}{\pi c} &  [-\tfrac{\pi}{2}, \tfrac{\pi}{2}] \\
+% % 0 & \text{otherwise}
+% % \end{cases}
+% % \end{equation}
+% % when $0 < c \leq 1$, and
+% % \begin{equation} \label{eq:fphicg1}
+% % f_{\Phi}(\phi) = \frac{\cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}}{2\pi c}
+% % \end{equation}
+% % when $c > 1$.
+% % \end{lemma}
+% % \begin{IEEEproof}
+% % We make use of delta `function' notation and write the `pdf' of $Z$ as $\delta(z - c)$.  Using~\eqref{eq:pdfRPhi} the pdf of $\Phi$ is
+% % \[
+% % f_{\Phi}(\phi) = \int_{0}^{\infty} r \frac{\delta(z - c)}{ 2\pi z} dr = \sum_{r} \frac{r}{2\pi c}
+% % \]
+% % where $z^2 = r^2 - 2r\cos(\phi) + 1$ and the sum is over the positive real solutions of $c^2 = r^2 - 2r\cos(\phi) + 1$.
 If $0 < c \leq 1$ and $\phi \in [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are two positive real solutions, these being
-% \[
-% r = \cos(\phi) \pm \sqrt{c^2 - 1 + \cos^2(\phi)},
-% \]
-% whilst if $\phi \notin [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are no real solutions.  Thus~\eqref{eq:fphicleq1} holds.  If $c > 1$ there is a single real solution
-% \[
-% r = \cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}
-% \]
-% so that~\eqref{eq:fphicg1} holds.
-% \end{IEEEproof}
-
-% Now if $W$ is a circularly symmetric complex random variable with $Z = \abs{W}$ having pdf $f_Z$ and let $R \geq 0$ and $\Phi \in [-\pi, \pi)$ be real random variables such that $Re^{j\Phi} = 1 + W$.
+% % \[
+% % r = \cos(\phi) \pm \sqrt{c^2 - 1 + \cos^2(\phi)},
+% % \]
+% % whilst if $\phi \notin [-\tfrac{\pi}{2}, \tfrac{\pi}{2}]$ there are no real solutions.  Thus~\eqref{eq:fphicleq1} holds.  If $c > 1$ there is a single real solution
+% % \[
+% % r = \cos(\phi) + \sqrt{c^2 - 1 + \cos^2(\phi)}
+% % \]
+% % so that~\eqref{eq:fphicg1} holds.
+% % \end{IEEEproof}
+
+% % Now if $W$ is a circularly symmetric complex random variable with $Z = \abs{W}$ having pdf $f_Z$ and let $R \geq 0$ and $\Phi \in [-\pi, \pi)$ be real random variables such that $Re^{j\Phi} = 1 + W$.
 The pdf of $\Phi$ is
-% \[
-% f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{f_Z(z)}{ 2\pi z} dr.
-% \]
-% Let
-% \[
-% I_{N,c}(z) = \begin{cases}
-% N \frac{f_Z(z)}{ 2\pi z} & z \in [cN, c(N-1)) \\
-% 0 & \text{otherwise}.
-% \end{cases}
-% \]
-% Now
-% \[
-% \frac{f_Z(z)}{ 2\pi z} = \frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z)
-% \]
-% so that
-% \[
-% f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z) dr = \frac{1}{N} \sum_{c = 0}^{\infty} \int_{0}^{\infty} r I_{N,c}(z) dr 
-% \]
-
-% \subsection{Proof of Proposition~\ref{prop:contgg2}} \label{appendix:prop2proof}
-
-% \begin{IEEEproof}
-% We first prove the result for $h_1$.  Write
-% \begin{align*}
-% h_1(x) &= \int_{-\pi}^{\pi} \int_{0}^\infty r \cos(\phi + x) f(r,\phi) dr d\phi \\
-% &= \int_{-\pi}^{\pi}  \cos(\phi + x) g(\phi) d\phi,
-% \end{align*}
-% where $f(r,\phi)$ is the joint pdf of $R$ and $\Phi$, and 
-% \[
-% g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
-% \]
-% Since the pdf of $X$ is circularly symmetric, $f(r,\phi)$ and $g(\phi)$ are even in $\phi$ on $[-\pi,\pi]$ (see the discussion after~\eqref{eq:pdfRPhi}).
+% % \[
+% % f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{f_Z(z)}{ 2\pi z} dr.
+% % \]
+% % Let
+% % \[
+% % I_{N,c}(z) = \begin{cases}
+% % N \frac{f_Z(z)}{ 2\pi z} & z \in [cN, c(N-1)) \\
+% % 0 & \text{otherwise}.
+% % \end{cases}
+% % \]
+% % Now
+% % \[
+% % \frac{f_Z(z)}{ 2\pi z} = \frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z)
+% % \]
+% % so that
+% % \[
+% % f_{\Phi}(\phi) = \int_{0}^{\infty} r\frac{1}{N} \sum_{c = 0}^{\infty}I_{N,c}(z) dr = \frac{1}{N} \sum_{c = 0}^{\infty} \int_{0}^{\infty} r I_{N,c}(z) dr 
+% % \]
+
+% % \subsection{Proof of Proposition~\ref{prop:contgg2}} \label{appendix:prop2proof}
+
+% % \begin{IEEEproof}
+% % We first prove the result for $h_1$.  Write
+% % \begin{align*}
+% % h_1(x) &= \int_{-\pi}^{\pi} \int_{0}^\infty r \cos(\phi + x) f(r,\phi) dr d\phi \\
+% % &= \int_{-\pi}^{\pi}  \cos(\phi + x) g(\phi) d\phi,
+% % \end{align*}
+% % where $f(r,\phi)$ is the joint pdf of $R$ and $\Phi$, and 
+% % \[
+% % g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr.
+% % \]
+% % Since the pdf of $X$ is circularly symmetric, $f(r,\phi)$ and $g(\phi)$ are even in $\phi$ on $[-\pi,\pi]$ (see the discussion after~\eqref{eq:pdfRPhi}).
 It is also the case that $g(\phi)$ is unimodal with mode at zero, as a result of Lemma~\ref{lem:gunimodmodzero}.
 Write $\cos(\phi + x)$ as a sum of even and odd parts,
-% \[
-% \cos(\phi + x) = \cos(\phi)\cos(x) - \sin(\phi)\sin(x).
-% \]
-% Now
-% \[
-% h_1(x) = \cos(x) \int_{-\pi}^{\pi} \cos(\phi) g(\phi) d\phi
-% \]
-% since $\int_{-\pi}^{\pi} \sin(\phi) g(\phi) d\phi = 0$ because $\sin(\phi)g(\phi)$ is odd.  Now 
-% \[
-% K = \int_{-\pi}^{0} \cos(\phi) g(\phi) d\phi > 0
-% \] 
-% because $g(\phi) > 0$ is increasing on $(-\pi,0)$ and $\cos(\phi)$ is odd and increasing on $(-\pi,0)$.
+% % \[
+% % \cos(\phi + x) = \cos(\phi)\cos(x) - \sin(\phi)\sin(x).
+% % \]
+% % Now
+% % \[
+% % h_1(x) = \cos(x) \int_{-\pi}^{\pi} \cos(\phi) g(\phi) d\phi
+% % \]
+% % since $\int_{-\pi}^{\pi} \sin(\phi) g(\phi) d\phi = 0$ because $\sin(\phi)g(\phi)$ is odd.  Now 
+% % \[
+% % K = \int_{-\pi}^{0} \cos(\phi) g(\phi) d\phi > 0
+% % \] 
+% % because $g(\phi) > 0$ is increasing on $(-\pi,0)$ and $\cos(\phi)$ is odd and increasing on $(-\pi,0)$.
 Thus $h_1(x) = 2K \cos(x)$ is uniquely maximised at zero because $\cos(x)$ is uniquely maximised at zero.
-% \end{IEEEproof}
-
-% \begin{lemma}\label{lem:gunimodmodzero}
-% Under the assumptions in Proposition~\ref{prop:contgg2} let $f_Z$ be the pdf of $\abs{W}$.
+% % \end{IEEEproof}
+
+% % \begin{lemma}\label{lem:gunimodmodzero}
+% % Under the assumptions in Proposition~\ref{prop:contgg2} let $f_Z$ be the pdf of $\abs{W}$.
 Then $z^{-1} f_Z(z)$ is differentiable and decreasing for $z > 0$ and the function $g(\phi) = \int_{0}^{\infty} r f(r,\phi) dr$ is unimodal on $[-\pi,\pi]$ with mode at $\phi = 0$.
-% \end{lemma}
-% \begin{IEEEproof}
-% From~\eqref{eq:fzandfZYwithz} it follows that $2\pi f_{X,Y}(z,0) = z^{-1} f_Z(z)$ for all $z>0$, and since $f_{X,Y}(z,0)$ is differentiable and nonincreasing in $z$ (by assumption in the statement of Proposition~\ref{prop:contgg2}) we have that $z^{-1} f_Z(z)$ is differentiable and decreasing for $z > 0$.
+% % \end{lemma}
+% % \begin{IEEEproof}
+% % From~\eqref{eq:fzandfZYwithz} it follows that $2\pi f_{X,Y}(z,0) = z^{-1} f_Z(z)$ for all $z>0$, and since $f_{X,Y}(z,0)$ is differentiable and nonincreasing in $z$ (by assumption in the statement of Proposition~\ref{prop:contgg2}) we have that $z^{-1} f_Z(z)$ is differentiable and decreasing for $z > 0$.
 So the derivative $\frac{d}{dz} \big( z^{-1} f_Z(z) \big) > 0$ for all $z > 0$.
 Using~\eqref{eq:pdfRPhi},
-% \[
-% g(\phi) = \int_{0}^\infty r^2 \frac{f_Z(z)}{2\pi z } dr
-% \]
-% where $z = \sqrt{r^2 - 2r\cos\phi + 1}$.  Differentiating with respect to $\phi$,
-% \[
-% g^\prime(\phi) = \sin(\phi) \int_{0}^\infty \frac{r^3}{\pi} \frac{d}{dz}\left( \frac{f_Z(z)}{z} \right) dr
-% \]
-% and since the term inside the integral is negative, the integral is negative.
+% % \[
+% % g(\phi) = \int_{0}^\infty r^2 \frac{f_Z(z)}{2\pi z } dr
+% % \]
+% % where $z = \sqrt{r^2 - 2r\cos\phi + 1}$.  Differentiating with respect to $\phi$,
+% % \[
+% % g^\prime(\phi) = \sin(\phi) \int_{0}^\infty \frac{r^3}{\pi} \frac{d}{dz}\left( \frac{f_Z(z)}{z} \right) dr
+% % \]
+% % and since the term inside the integral is negative, the integral is negative.
 Since $\sin(\phi)$ is negative on $(-\pi,0)$ and positive on $(0,\pi)$ it follows that the derivative $g^\prime(\phi)$ is positive on $(-\pi,0)$ and negative on $(0,\pi)$.
 Thus $g(\phi)$ is unimodal with mode at $\phi = 0$.
-% \end{IEEEproof}
+% % \end{IEEEproof}
 

### 2012-10-26 other/confnoncoherent/paper.tex
Added more of Barry's edits.

@@
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
-The bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the bias is significant when the SNR is small.
+The asymptotic bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the asymptotic bias is significant when the SNR is small.
 %In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
@@
 Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.  
-A complex random variable $X$ is said to be \emph{circularly symmetric} if the distribution of its phase $\angle{X}$ is uniform on $[0,2\pi)$ and is independent of the distribution of its magnitude $\abs{X}$.
+A complex random variable $W$ is said to be \emph{circularly symmetric} if its phase $\angle{W}$ is independent of its magnitude $\abs{W}$ and if the distribution of $\angle{W}$ is uniform on $[0,2\pi)$.
 That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = X$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
@@
 
-We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.
+% We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.
 Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying $R e^{j\Phi} = 1 + X$.
 The joint distribution of $R$ and $\Phi$ can be shown to be
-\[
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
-\]
+% \[
+% f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
+% \]
 
@@
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance
+Then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
 \[
@@
 
-A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
+A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\tfrac{\pi}{M},\tfrac{\pi}{M})$.
 %This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
@@
 
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}$ rather than directly on $\hat{\theta} - \theta_0$.
+The theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}$ rather than directly on $\theta_0 - \hat{\theta}$.
 This makes practical sense, and to see why let $s_i' = e^{j2\pi k/M}s_i$ be $M$-PSK symbols obtained by rotating $s_1, \dots, s_L$ by $e^{j2\pi k/M}$ for some integer $k$.
@@
 and so, if $a = \hat{\rho}e^{j\hat{\theta}}$ minimises~\eqref{eq:hatadef}, then so does $\hat{\rho} e^{j(\hat{\theta} - 2\pi k/M)}$.
-Thus, $\theta_0 - \hat{\theta}$ should be attributed the same error as $\theta_0 - \hat{\theta} - \tfrac{2\pi}{M}k$, that is, the error should be computed `modulo $\tfrac{2\pi}{M}$'.
+Thus, $\theta_0 - \hat{\theta}$ should be attributed the same error as $\theta_0 - \hat{\theta} - \tfrac{2\pi}{M}k$.
+That is, the error should be computed `modulo $\tfrac{2\pi}{M}$'.
 It is for this reason that \emph{differential encoding} is often used with noncoherent $M$-PSK  detectors.

### 2012-10-26 paper.tex
Added more of Barry's edits.

@@
 The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
-In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
+In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
 A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-The idea is to `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
+The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.
 A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
@@
 Mackenthun specifically considered the noncoherent setting, so we modify the algorithm to include the pilot symbols.
-For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
+For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
 In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.
@@
 
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta} - \theta_0$.
+The theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ rather than directly on $\theta_0 - \hat{\theta}$.
 This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
@@
 Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda \rightarrow 0$, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.
-The lemma follows after puting $\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1$.
+The lemma follows after putting $\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1$.
 \end{IEEEproof}
@@
 This type of result is typically called \emph{tightness} or \emph{asymptotic continuity}~\cite{Pollard_asymp_empi_proc_1989,van2009empirical,Billingsley1999_convergence_of_probability_measures}.
-We ommit the proof which follows in a straightforward, but lengthy manner using an argument called \emph{symmetrisation} followed by an argument called \emph{chaining}~\cite{Pollard_asymp_empi_proc_1989,van2009empirical}.
+We omit the proof which follows in a straightforward, but lengthy manner using an argument called \emph{symmetrisation} followed by an argument called \emph{chaining}~\cite{Pollard_asymp_empi_proc_1989,van2009empirical}.
 

### 2012-10-30 paper.tex
added paragraph about unequal error protection for Bill

@@
 It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+
+It is straightforward to generalise the results we present here to allow data symbols with varying constellation size, i.e.
+varying $M$.
+For example, one might give more importance to certain data symbols and use BPSK ($M=2$) for these, but QPSK ($M=4$) for other less important symbols.
+This is related to what is called \emph{unequal error protection} in the literature~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.
+To keep our ideas and notation focused we don't consider this further here.
 
@@
 This stronger condition is required to prove the asymptotic normality of $\sqrt{L}\hat{m}_L$.
-The next two sections give proofs for Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
-The proofs make use of various lemmas.
-The lemmas are proved in the appendix.
+
+The next two sections give proofs for Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  The proofs make use of various lemmas.  The lemmas are proved in the appendix.
 

### 2012-11-23 other/confnoncoherent/paper.tex
added some minor changes of Vaughan's to the ICASSP paper.

@@
  \title{Noncoherent least squares estimators of carrier phase and amplitude}
-\name{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, Vaughan Clarkson$^{2}$ and Barry Quinn$^{3}$ \thanks{Supported under the Australian Government’s Australian Space Research Program.}}
+\name{Robby McKilliam$^{1}$, Andr\'{e} Pollok$^{1}$, Bill Cowley$^{1}$, I.
+Vaughan L.
+Clarkson$^{2}$ and Barry G.
+Quinn$^{3}$ \thanks{Supported under the Australian Government’s Australian Space Research Program.}}
 \address{ \small $^{1}$Institute for Telecommunications Research, University of South Australia, SA, \textsc{Australia} \vspace{0.05cm} \\
@@
 The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
-Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{a_0 - \hat{a}}^2$.
-It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle^2$ and $(\rho_0 - \hat{\rho})^2$.
+Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{\hat{a} - a_0}^2$.
+It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\hat{\theta} - \theta_0\rangle^2$ and $(\hat{\rho} - \rho_0)^2$.
 The function $\fracpart{\cdot}$ denotes its argument taken `modulo $\tfrac{2\pi}{M}$' into the interval $[-\pi/M, \pi/M)$, that is
@@
 We have chosen to round up half-integers here.
-It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+It will become apparent why $\langle\hat{\theta} - \theta_0\rangle^2$ rather than $(\hat{\theta} - \theta_0)^2$ is the appropriate measure of error for the phase parameter.
 
@@
 Section~\ref{sec:stat-prop-least} describes the statistical properties of the least squares estimators of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
+We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
 The asymptotic bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the asymptotic bias is significant when the SNR is small.
-%In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
+%Section~\ref{sec:gaussian-noise-case} considers the special case when the noise is Gaussian.
 In this case, our formula for the asymptotic variance of can be simplified.
@@
 
-% We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.
-Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying $R e^{j\Phi} = 1 + X$.
+%We will have particular use of complex random variables of the form $1 + X$ where $X$ is circularly symmetric.
+Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real %random variables satisfying $R e^{j\Phi} = 1 + X$.
 The joint distribution of $R$ and $\Phi$ can be shown to be
-% \[
+%\[
 % f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
@@
 
-The next two theorems describe the asymptotic properties of the least squares estimator.  We omit the proofs due to space constraints.  Proofs will be provided in an forthcomming paper.
+The next two theorems describe the asymptotic properties of the least squares estimator.  We omit the proofs due to space constraints.  Proofs will be provided in a forthcomming paper.
 
@@
 \begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item $\sfracpart{\hat{\theta} - \theta_0} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
 \item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
@@
 \]
-Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
+Put $\hat{\lambda}_L = \sfracpart{\hat{\theta} - \theta_0}$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
@@
 %This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\hat{\theta} - \theta_0}_\pi$ will not converge.
 One can check that the assumption holds when $w_1$ is circularly symmetric and normally distributed.
@@
 
-The theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}$ rather than directly on $\theta_0 - \hat{\theta}$.
+The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}$ rather than directly on $\hat{\theta} - \theta_0$.
 This makes practical sense, and to see why let $s_i' = e^{j2\pi k/M}s_i$ be $M$-PSK symbols obtained by rotating $s_1, \dots, s_L$ by $e^{j2\pi k/M}$ for some integer $k$.
@@
 and so, if $a = \hat{\rho}e^{j\hat{\theta}}$ minimises~\eqref{eq:hatadef}, then so does $\hat{\rho} e^{j(\hat{\theta} - 2\pi k/M)}$.
-Thus, $\theta_0 - \hat{\theta}$ should be attributed the same error as $\theta_0 - \hat{\theta} - \tfrac{2\pi}{M}k$.
+Thus, $\hat{\theta} - \theta_0$ should be attributed the same error as $\hat{\theta} - \theta_0 - \tfrac{2\pi}{M}k$.
 That is, the error should be computed `modulo $\tfrac{2\pi}{M}$'.
@@
 \]
-and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
+and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  The value of $A$ and $B$ can be efficiently computed by numerical integration using these formula.
 
@@
 Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the sample MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-This estimator requires the selection of a function $F : \reals \mapsto \reals$.
+This estimator requires the selection of a function $F : \reals \mapsto \reals$ (see equation (2) in~\cite{ViterbiViterbi_phase_est_1983}).
 We have chosen $F(x) = 1$.
@@
 
-\section{RELATION TO PRIOR WORK}
+% \section{RELATION TO PRIOR WORK}
 
-The problem of estimating carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols has undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
+% The problem of estimating carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols has undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
 The least squares estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and this is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
@@
 The phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.
+This bias is large when the signal to noise ratio is small.
+It would be interesting to investigate methods for correcting this bias.
+
+\section{Conclusion}
+
+We have studied the least squares estimator of carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols.
+The estimator can be computed in $O(L\log L)$ operations using the algorithm of Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
+We showed that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
 However, the amplitude estimator $\hat{\rho}$ is biased, and converges to $G(0)\rho_0$.

### 2012-11-25 other/confnoncoherent/paper.tex
added more of Vaughan's edits

@@
 %\begin{equation}\label{eq:RiandPhii}
+\vspace{-0.15cm}
 \[
@@
 \begin{enumerate}
-\item $\sfracpart{\hat{\theta} - \theta_0} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
+\item \vspace{-0.15cm} $\sfracpart{\hat{\theta} - \theta_0} \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item \vspace{-0.2cm} $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
 \end{enumerate}
@@
 as $L \rightarrow \infty$, where
+\vspace{-0.15cm}
 \[
@@
 \]
+\vspace{-0.15cm}
 \[
@@
 
+\vspace{-0.1cm}
 \section{The Gaussian noise case}\label{sec:gaussian-noise-case}
-
+\vspace{-0.1cm}
 Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
+\vspace{-0.15cm}
 \[
@@
 It can be shown that
+\vspace{-0.15cm}
 \[
@@
 where $\kappa = \rho_0/\sigma$ and
+\vspace{-0.15cm}
 \[
@@
 
+\vspace{-0.1cm}
 \section{Simulations}\label{sec:simulations}
-
+\vspace{-0.1cm}
 We present the results of Monte-Carlo simulations with the least squares estimator.
@@
 Figures~\ref{fig:plotphaseM2}~and~\ref{fig:plotphaseM4} also display the sample MSE of the phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-This estimator requires the selection of a function $F : \reals \mapsto \reals$ (see equation (2) in~\cite{ViterbiViterbi_phase_est_1983}).
-We have chosen $F(x) = 1$.
+This estimator requires the selection of a function $F$ that transforms the amplitude of each sample prior to the final estimation step.
+% (see equation (2) in~\cite{ViterbiViterbi_phase_est_1983}).
+They propose several viable alternatives, from which we have chosen $F(x) = 1$.
 The sample MSE of the least squares estimator and the Viterbi and Viterbi estimator is similar.
@@
 The solid line is the asymptotic variance predicted by Theorem~\ref{thm:normality}, the dots are the Monte-Carlo simulations with the least squares estimator, and the dashed line the simulations with the unmodulated carrier estimator.
-Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+For the least squares estimator each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
 This requires $G(0)$ to be known.
@@
 However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+As indicated in Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.
+However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.
 
@@
 
+\vspace{-0.1cm}
 \section{Conclusion}
-
+\vspace{-0.1cm}
 We have studied the least squares estimator of carrier phase and amplitude from the observation of $L$ noisy $M$-PSK symbols.
@@
 
+\nocite{McKilliamLinearTimeBlockPSK2009,Ryan2007}
+
 %\vfill\pagebreak

### 2012-12-03 other/uep/paper.tex

--- 2012-09-12 algonly.tex
@@
 %opening
-\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
+\title{Carrier phase and amplitude estimation for phase shift keying with unequal error protection}
 \author{Robby McKilliam, Andre Pollok, Bill Cowley
@@
 \begin{abstract}
-We consider least squares estimators of carrier phase and amplitude from a noisy communications signal that contains both pilot signals, known to the receiver, and data signals, unknown to the receiver.
-We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-ary phase shift keying.
-We describe an algorithm to compute the least suqares estimator that requires only $O(L \log L)$ arithmetic operations, where $L$ is the number of recieved symbols.
+We consider least squares estimators of carrier phase and amplitude from a noisy communications signal.
+We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-phase shift keying, and consider specifically the case where a number of different constellations sizes $M$ are used simultaneously.
+We describe an algorithm to compute the least squares estimators of carrier phase and amplitude that requires only $O(L \log L)$ arithmetic operations, where $L$ is the number of recieved symbols.
 \end{abstract}
@@
 \]
-where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M \geq 2$ is the size of the constellation.
+where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M \geq 1$ is the size of the constellation.
 We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.
@@
 $P \cap D = \emptyset$, and $L = \abs{P \cup D}$ is the total number of symbols transmitted.
+The data symbols are further separated into subsets according to the size of the constellation used to modulate each symbol.
+Let $D_2,D_3,\dots$ denote a partition of the indices in $D$ such that $D_2$ is the set of indices corresponding to data symbols modulated with $2$-PSK (i.e BSPK), and $D_3$ is the set of symbols modulated with $3$-PSK, and $D_4$ is the set of symbols modulated with $4$-PSK (i.e.
+QPSK) and so on.
+Let $G$ be the set of integers for which $D_m$ is not empty, i.e.
+$D_m \neq \emptyset$ whenever $m \in G$.
+We have
+\[
+D = \bigcup_{m=2}^{\infty}D_m = \bigcup_{m\in G} D_m.
+\]
 
@@
 &= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2, 
+%\\ &= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{m \in G} \sum_{i \in D_m} \abs{ y_i - a d_i }^2
 \end{split}
@@
 
-The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
-In the nocoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
+The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set) and where only one constellation size is used, i.e., $G = \{M\}$ contains precisely one element.
+In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
 A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
@@
 Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included.
-Techniques are available but they are somewhat adhoc~\cite{Cowley_ref_sym_carr_1998}.
-This problem does not occur with the least square estimator.
+It is not entirely obvious how pilot symbols or data symbols from mulitple constellations should be included in this estimator.
+One approach is to choose,
+\[
+A = \frac{1}{L} \sum_{i \in P} F(\abs{y_i}) \tfrac{y_i}{y_i} p_i^* +  \frac{1}{L} \sum_{m \in G} \sum_{i \in D_m} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^m.
+\]
+The simulations we present in Section~\ref{sec:simulations} suggest that the accuracy of this estimator is poor when compared to the least squares estimator.
+%This problem does not occur with the least square estimator.
 
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
+Both Mackenthun and Swelden's considered only the noncoherent setting and where only one constellation size is used, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols and multiple constellation sizes.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
@@
 
-\section{Mackentun's algorithm with pilots}\label{sec:least-squar-estim}
+\section{Mackenthun's algorithm with pilots}\label{sec:least-squar-estim}
 
-In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, so we modify the algorithm to include the pilot symbols.
+In this section we derive an algorithm to compute the least squares estimator of the carrier phase and amplitude.
+Our algorithm is motivated by the algorithm of Mackenthun~\cite{Mackenthun1994} who considered the noncoherent setting and where only one constellation size is used for the data symbols.
 For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
 In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.
-This complexity arises from the need to sort a list of $\abs{D}$ elements.
+
 %We use order notation in the standard way, that is, for functions $h$ and $g$, we write $h(N) = O(g(N))$ to mean that there exists a constant $K > 0$ and a finite $N_0$ such that $h(N) \leq K g(N)$ for all $N > N_0$.
@@
 \end{align}
-For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
+For given $\theta$, the least squares estimator of the $i$th data symbol $d_i \in D_m$ for some $m \in G$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
 \begin{equation}\label{eq:hatdfinxtheta}
-\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
+\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)}_{m},
 \end{equation}
-where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
-A word of caution, the notation $\round{\cdot}$ is often used to denote rounding to the nearest \emph{integer}.
-This is not the case here.
+where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}_{m}$ rounds its argument to the nearest multiple of $\frac{2\pi}{m}$.
 If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer then,
 \[
-\round{x} = \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
+\round{x}_{m} = \tfrac{2\pi}{m}\operatorname{round}\left(\tfrac{m}{2\pi}x\right).
 \] 
 Note that $\hat{d}_i(\theta)$ does not depend on $\rho$.
-As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
+As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{m}, \dots, \tfrac{2\pi(m-1)}{m}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
 With this in mind,
 \[
-\hat{u}_i(\theta) = \round{\angle{y_i} - \theta }
+\hat{u}_i(\theta) = \round{\angle{y_i} - \theta }_{m}
 \]
@@
 \[
-z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}.
+z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}_{m}.
 \]
@@
 \begin{cases}
-b_i, &  0 \leq \theta < z_i + \frac{\pi}{M} \\
-b_i e^{-j2\pi/M}, & z_i + \frac{\pi}{M} \leq \theta < z_i + \frac{3\pi}{M} \\ 
+b_i, &  0 \leq \theta < z_i + \frac{\pi}{m} \\
+b_i e^{-j2\pi/m}, & z_i + \frac{\pi}{m} \leq \theta < z_i + \frac{3\pi}{m} \\ 
 \vdots & \\
-b_i e^{-j2\pi k /M}, & z_i + \frac{\pi(2k - 1)}{M} \leq \theta < z_i + \frac{\pi(2k + 1)}{M}  \\ 
+b_i e^{-j2\pi k /m}, & z_i + \frac{\pi(2k - 1)}{m} \leq \theta < z_i + \frac{\pi(2k + 1)}{m}  \\ 
 \vdots & \\
-b_i e^{-j2\pi}, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
+b_i e^{-j2\pi}, &  z_i + \frac{\pi(2m - 1)}{m} \leq \theta < 2\pi. \\
 \end{cases}
@@
 \]
-be a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
+be a function mapping the interval $[0, 2\pi)$ to a sequence of phase shift keyed symbols indexed by the elements of $D$.
 Observe that $f(\theta)$ is piecewise continuous.
@@
 If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are atmost $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
+Observe from~\eqref{eq:uicombos} that there are atmost $\sum_{m\in G}(m+1)\abs{D_m} = O(L)$ sequences in $S$, because there are $m+1$ sequences for each $i \in D_m$.
 
-The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indices in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
+The sequences in $S$ can be enumerated as follows.  Let 
+\[
+T = \bigcup_{m\in G}\bigcup_{i \in D_m} \{ (z_i + \frac{\pi(2k + 1)}{m}, i, m), k = 0,\dots,m-1 \}
+\]
+be a set of triples with first element a real number, second element from $D$ and third element from $G$.
+Let $H = \abs{T}$ be the number of elements in $T$ and observe that $H \leq \sum_{m\in G}(m+1)\abs{D_m} = O(L)$.
+Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ and denote the three elements of $t_i$ by $a_{i}$, $b_{i}$ and $c_{i}$ respectively, that is,
+\[
+t_i = ( a_{i}, b_{i}, c_{i} ).
+\]
+So 
+Let $\sigma$ denote the permutation of the integers $\{1,\dots,H\}$ such that $t_{\sigma(i)1}$ are in ascending order, that is,
 \begin{equation}\label{eq:sigmasortind}
-z_{\sigma(i)} \leq z_{\sigma(k)}
+z_{\sigma_m(i)} \leq z_{\sigma_m(k)}
 \end{equation}
@@
 
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error versus SNR for BPSK with $L=4096$.}
-		\label{fig:plotphaseBPSK}
-\end{figure}
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
+% 		\caption{Phase error versus SNR for BPSK with $L=4096$.}
+% 		\label{fig:plotphaseBPSK}
+% \end{figure}
 
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error versus SNR for QPSK with $L=4096$.}
-		\label{fig:plotphaseQPSK}
-\end{figure}
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
+% 		\caption{Phase error versus SNR for QPSK with $L=4096$.}
+% 		\label{fig:plotphaseQPSK}
+% \end{figure}
 
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
-		\label{fig:plotphase8PSK}
-\end{figure}
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
+% 		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
+% 		\label{fig:plotphase8PSK}
+% \end{figure}
 
@@
 
-\begin{figure}[tp]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
-		\caption{Phase error versus SNR for QPSK.}
-		\label{fig:plotphaseQPSKmultL}
-\end{figure}
+% \begin{figure}[tp]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
+% 		\caption{Phase error versus SNR for QPSK.}
+% 		\label{fig:plotphaseQPSKmultL}
+% \end{figure}
 

### 2012-12-03 other/uep/paper.tex
added to uep.  Needs code, should be fine now.

@@
 \[
-T = \bigcup_{m\in G}\bigcup_{i \in D_m} \{ (z_i + \frac{\pi(2k + 1)}{m}, i, m), k = 0,\dots,m-1 \}
+T = \bigcup_{m\in G} \bigcup_{ i \in D_m }\bigcup_{k=1}^{m}\left\{ (z_i + \tfrac{\pi(2k - 1)}{m}, i, m) \right\}
 \]
-be a set of triples with first element a real number, second element from $D$ and third element from $G$.
-Let $H = \abs{T}$ be the number of elements in $T$ and observe that $H \leq \sum_{m\in G}(m+1)\abs{D_m} = O(L)$.
-Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ and denote the three elements of $t_i$ by $a_{i}$, $b_{i}$ and $c_{i}$ respectively, that is,
+be a set of triples with first element a real number, second element from $D$ and third element from $G$.  Let 
 \[
-t_i = ( a_{i}, b_{i}, c_{i} ).
+H = \abs{T} = \sum_{m\in G}(m+1)\abs{D_m} = O(L)
 \]  
-So 
-Let $\sigma$ denote the permutation of the integers $\{1,\dots,H\}$ such that $t_{\sigma(i)1}$ are in ascending order, that is,
-\begin{equation}\label{eq:sigmasortind}
-z_{\sigma_m(i)} \leq z_{\sigma_m(k)}
-\end{equation}
-whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}-1\}$.
-It is convenient to define the indices into $\sigma$ to be taken modulo $\abs{D}$, that is, if $m$ is an integer not from $\{0, 1, \dots, \abs{D}-1\}$ then we define $\sigma(m) = \sigma(k)$ where $k \equiv m \mod \abs{D}$ and $k \in  \{0, 1, \dots, \abs{D}-1\}$.
+denote the number of elements in $T$.
+Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ sorted in accending order of the first element in the triple, that is, if $t_{i} = (t_{i1},t_{i2},t_{i3})$ then
+\[
+t_{i1} \leq t_{ki}
+\]
+whenever $i < k$.
+Put $\sigma(i) = t_{i2}$ and $m(i) = t_{i3}$ so that $\sigma(1), \dots, \sigma(H)$ and $m(1),\dots,m(H)$ correspond respectively with the second and third elements from the triples $t_1,\dots,t_H$
+%So $a_1,\dots,a_H$ is a sequence of real numbers, $b_1,\dots,b_H$ is a sequence of integers from $D$ and $c_1,\dots,c_H$ is a sequence
+
 The first sequence in $S$ is 
 \[
-f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
+f_1 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
 \]  
-The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/M}$.
-Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.
+The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/m(1)}$.
+Given a sequence $x$ we use $x e_{k}$ to denote $x$ with the $\sigma(k)$th element replaced by $x_{\sigma(k)} e^{-j2\pi/m(k)}$.
 Using this notation,
 \[
-f_1 = f_0 e_{\sigma(0)}.
+f_2 = f_1 e_1.
 \] 
@@
 \[
-f_2 = f_0 e_{\sigma(0)} e_{\sigma(1)} = f_1 e_{\sigma(1)},
+f_3 = f_1 e_{1} e_{2} = f_2 e_{2},
 \]
@@
 \begin{equation}\label{eq:fkrec}
-f_{k+1} = f_{k} e_{\sigma(k)}.
+f_{k+1} = f_{k} e_{k}.
 \end{equation}
@@
 We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,(M+1)\abs{D}-1\}$.
+A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{1,\dots,H\}$.
 Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
-So, this na\"{\i}ve approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
+So, this na\"{\i}ve approach would require $O(LH) = O(L^2)$ operations in total.
 Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.
@@
 \[
-g_{k+1} = g_{k} e_{\sigma(k)}^*,
+g_{k+1} = g_{k} e_{k}^*,
 \]
-where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/M}$.  Now,
+where $g_{k} e_{k}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/m(k)}$.  Now,
 \[
-Y_0 = B + \sum_{i \in D} g_{0i}
+Y_1 = B + \sum_{i \in D} g_{1i}
 \] 
@@
 \begin{align*}
-Y_1 &= B + \sum_{i \in D} g_{1i} \\
-&= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
-&= Y_0 + \eta g_{0\sigma(0)},
+Y_2 &= B + \sum_{i \in D} g_{1i} \\
+&= B +  (e^{j2\pi/m(1)} - 1)g_{1\sigma(1)} + \sum_{i \in D} g_{1i} \\
+&= Y_0 + (e^{j2\pi/m(1)} - 1)g_{1\sigma(1)},
 \end{align*}
-where $\eta = e^{j2\pi/M} - 1$.  In general,
+In general,
 \[
-Y_{k+1} = Y_k + \eta g_{k\sigma(k)}.
+Y_{k+1} = Y_k + (e^{j2\pi/m(k)} - 1) g_{k\sigma(k)}.
 \]
@@
 \end{equation}
-Psuedocode is given in Algorithm~\ref{alg:loglinear}.
-Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
+Psuedocode is given in Algorithm~\ref{alg:loglinear}.  BLERG
+
+%Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
 The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
@@
 \KwIn{$\{y_i, i \in P \cup D \}$}
-\For{$i \in D$ \nllabel{alg_loop_setup}}{
+\For{$m \in G$ \nllabel{alg_loop_setup}}{
+\For{$i \in D_m$ }{
 $\phi = \angle{y_i}$ \;
-$u = \round{\phi} $ \;
-$z_i = \phi -  u $ \;
+$u = \round{\phi}_m $ \;
 $g_i = y_i e^{-j u}$ \;
+\For{$k = 1, \dots, m$ }{ $t_{mik} =  (z_i + \tfrac{\pi(2k + 1)}{m}, i, m)$} 
 }
-$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
+}
+$\operatorname{sort}(t_1,\dots,t_H)$ \;
 $Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
@@
 $\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
-$\eta = e^{j2\pi/M} - 1$ \;
-\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
-$Y = Y + \eta g_{\sigma(k)}$ \;
-$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
+\For{$k = 1,\dots,H$ \nllabel{alg_loop_search}}{
+$Y = Y + (e^{j2\pi/m(k)} - 1) g_{\sigma(k)}$ \;
+$g_{\sigma(k)} = e^{j2\pi/m(k)} g_{\sigma(k)} $\;
 $Q = \frac{1}{L}\abs{Y}^2$\;

### 2012-12-04 other/uep/paper.tex
added to uep document.

@@
 We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-phase shift keying, and consider specifically the case where a number of different constellations sizes $M$ are used simultaneously.
-We describe an algorithm to compute the least squares estimators of carrier phase and amplitude that requires only $O(L \log L)$ arithmetic operations, where $L$ is the number of recieved symbols.
+This scenario is common in practical communications systems employing what is called \emph{unequal error proection}.
+We describe an algorithm to compute the least squares estimators of carrier phase and amplitude that requires $O(L \log L)$ arithmetic operations, where $L$ is the number of recieved symbols.
 \end{abstract}
@@
 \]
+In practical settings the number of different constellations used $\abs{G}$ is small.
+A practical example might be that BPSK symbols are used for important symbols and QPSK symbols are used for less important symbols.
+In this example $G = \{2,4\}$ and $\abs{G} = 2$.
 
@@
 Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-It is not entirely obvious how pilot symbols or data symbols from mulitple constellations should be included in this estimator.
-One approach is to choose,
-\[
-A = \frac{1}{L} \sum_{i \in P} F(\abs{y_i}) \tfrac{y_i}{y_i} p_i^* +  \frac{1}{L} \sum_{m \in G} \sum_{i \in D_m} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^m.
-\]
+A caveat is that it is not obvious how pilot symbols or data symbols using mulitple constellations should be included in this estimator.
+% One approach is to compute
+% \[
+% A = \frac{1}{L} \sum_{i \in P} F(\abs{y_i}) \tfrac{y_i}{\abs{y_i}} p_i^* +  \frac{1}{L} \sum_{m \in G} \sum_{i \in D_m} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^m
+% \]
+% and choose the estimate of carrier phase to be $\tfrac{1}{K}\angle{A}$ where
+% \[
+% K = \begin{cases}
+% 1 & \abs{P} > 0 \\
+% \min(G) & \text{otherwise},
+% \end{cases}
+% \]
+% is equal to $1$ when pilot symbols exists, and is otherwise equal to the size of the smallest constellation $\min(G)$.
 The simulations we present in Section~\ref{sec:simulations} suggest that the accuracy of this estimator is poor when compared to the least squares estimator.
 %This problem does not occur with the least square estimator.
+This problem does not occur with the least squares estimator.
 
@@
 Both Mackenthun and Swelden's considered only the noncoherent setting and where only one constellation size is used, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols and multiple constellation sizes.
-Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
+%Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
 
@@
 Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-It turns out that there are atmost $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
+Let
+\[
+H = \sum_{m\in G}m\abs{D_m} = O(L).
+\]
+It turns out that there are atmost $H$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
 %When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
@@
 \vdots & \\
-b_i e^{-j2\pi}, &  z_i + \frac{\pi(2m - 1)}{m} \leq \theta < 2\pi. \\
+b_i e^{-j2\pi} = b_i, &  z_i + \frac{\pi(2m - 1)}{m} \leq \theta < 2\pi. \\
 \end{cases}
@@
 If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are atmost $\sum_{m\in G}(m+1)\abs{D_m} = O(L)$ sequences in $S$, because there are $m+1$ sequences for each $i \in D_m$.
+Observe from~\eqref{eq:uicombos} that there are atmost $H = O(L)$ sequences in $S$, because there are $m$ distinct values of $\hat{d}_i(\theta)$ for each $i \in D_m$.
 
@@
 \]
-be a set of triples with first element a real number, second element from $D$ and third element from $G$.  Let 
+be a set of triples with first element a real number, second element from $D$ and third element from $G$.
+The number of elements in $T$ is $H = \abs{T} = O(L)$.
+Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ sorted in accending order of the first element in the triple, that is, if $t_{k} = (t_{k1},t_{k2},t_{k3})$ then $t_{i1} \leq t_{k1}$ whenever $i < k$.
+Put
 \[
-H = \abs{T} = \sum_{m\in G}(m+1)\abs{D_m} = O(L)
+\sigma(k) = t_{k2} \qquad \text{and} \qquad m(k) = t_{k3}
 \]
-denote the number of elements in $T$.
-Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ sorted in accending order of the first element in the triple, that is, if $t_{i} = (t_{i1},t_{i2},t_{i3})$ then
-\[
-t_{i1} \leq t_{ki}
-\]
-whenever $i < k$.
-Put $\sigma(i) = t_{i2}$ and $m(i) = t_{i3}$ so that $\sigma(1), \dots, \sigma(H)$ and $m(1),\dots,m(H)$ correspond respectively with the second and third elements from the triples $t_1,\dots,t_H$
+so that $\sigma(1), \dots, \sigma(H)$ and $m(1),\dots,m(H)$ correspond respectively with the second and third elements from the triples $t_1,\dots,t_H$.
 %So $a_1,\dots,a_H$ is a sequence of real numbers, $b_1,\dots,b_H$ is a sequence of integers from $D$ and $c_1,\dots,c_H$ is a sequence
@@
 \end{equation}
-In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively enumerated.
+In this way, all sequences in $S$ can be recursively enumerated.
 
@@
 \end{equation}
-Psuedocode is given in Algorithm~\ref{alg:loglinear}.  BLERG
-
-%Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
-The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
+Pseudocode is given in Algorithm~\ref{alg:loglinear}.
+The function $\operatorname{sort}$ on line~\ref{alg_sortindices} return the triple $t_1,\dots,t_H$ so that the first element in the triples is in ascending order.
 This requires $O(L \log L)$ operations.
-The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.
+The $\operatorname{sort}$ function is the primary bottleneck in this algorithm when $L$ is large.
 The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
@@
 \KwIn{$\{y_i, i \in P \cup D \}$}
+$c = 1$ \;
 \For{$m \in G$ \nllabel{alg_loop_setup}}{
@@
 $g_i = y_i e^{-j u}$ \;
-\For{$k = 1, \dots, m$ }{ $t_{mik} =  (z_i + \tfrac{\pi(2k + 1)}{m}, i, m)$} 
+$z  = \phi - u$ \;
+\For{$k = 1, \dots, m$ }{ 
+$t_{c} =  (z + \tfrac{\pi(2k - 1)}{m}, i, m)$ \;
+$c = c + 1$ \;
 } 
 }
-$\operatorname{sort}(t_1,\dots,t_H)$ \;
+}
+$\operatorname{sort}(t_1,\dots,t_H)$ \nllabel{alg_sortindices} \;
 $Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
@@
 
-% \begin{figure}[p]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-% 		\caption{Phase error versus SNR for BPSK with $L=4096$.}
-% 		\label{fig:plotphaseBPSK}
-% \end{figure}
+ \begin{figure}[p]
+ 	\centering
+ 		\includegraphics[width=\linewidth]{code/data/plot-2.mps}
+ 		\caption{Phase error versus SNR}
+ 		\label{fig:plotphaseBPSK}
+ \end{figure}
 

### 2012-12-04 paper.tex
minor edit of Kevin
added to uep document.

@@
 
-\usepackage{xr}
-\externaldocument{paper2}
+%\usepackage{xr}
+%\externaldocument{paper2}
 
@@
 SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2  \\
-&= \sum_{i \in P} \abs{ y_i - a s_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2,
+&= \sum_{i \in P} \abs{ y_i - a p_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2,
 \end{split}
@@
 Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-It turns out that there are at most $(M+1)\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
+It turns out that there are at most $M\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
 %When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
@@
 \vdots & \\
-b_i e^{-j2\pi}, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
+b_i e^{-j2\pi} = b_i, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
 \end{cases}
@@
 If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are at most $(M+1)\abs{D}$ sequences in $S$, because there are $M+1$ sequences for each $i \in D$.
+Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $d_i(\theta)$ for each $i \in D$.
 
@@
 \end{equation}
-In this way, all $(M+1)\abs{D}$ sequences in $S$ can be recursively enumerated.
+In this way, all $M\abs{D}$ sequences in $S$ can be recursively enumerated.
 
 We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,(M+1)\abs{D}-1\}$.
+A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,M\abs{D}-1\}$.
 Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
-So, this na\"{\i}ve approach would require $O(L (M+1) \abs{D}) = O(L^2)$ operations in total.
+So, this na\"{\i}ve approach would require $O(L M \abs{D}) = O(L^2)$ operations in total.
 Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.
@@
 $\eta = e^{j2\pi/M} - 1$ \;
-\For{$k= 0$ \emph{\textbf{to}} $(M+1)\abs{D}-1$ \nllabel{alg_loop_search}}{
+\For{$k= 0$ \emph{\textbf{to}} $M\abs{D}-1$ \nllabel{alg_loop_search}}{
 $Y = Y + \eta g_{\sigma(k)}$ \;

### 2012-12-04 code/runsim.sh

@@ -1,5 +1,5 @@
 #export JAVA_OPTS="-Xprof -server -Xms1g -Xmx1g"
 export JAVA_OPTS="-d64 -server -Xms1g -Xmx1g"
-scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar clt.scala
-scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar montecarlo.scala
-scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar montecarlonc.scala
+scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar:lib/jblas-1.2.0.jar:lib/junit-4.11.jar clt.scala
+scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar:lib/jblas-1.2.0.jar:lib/junit-4.11.jar montecarlo.scala
+scala -nocompdaemon -cp lib/ComplexAmplitudePSK.jar:lib/ScalaNumbers.jar:lib/PubSim.jar:lib/Jama-1.0.2.jar:lib/flanagan.jar:lib/colt.jar:lib/RngPack.jar:lib/jblas-1.2.0.jar:lib/junit-4.11.jar montecarlonc.scala

### 2012-12-06 other/uep/paper.tex

@@
 We consider least squares estimators of carrier phase and amplitude from a noisy communications signal.
-We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-phase shift keying, and consider specifically the case where a number of different constellations sizes $M$ are used simultaneously.
-This scenario is common in practical communications systems employing what is called \emph{unequal error proection}.
-We describe an algorithm to compute the least squares estimators of carrier phase and amplitude that requires $O(L \log L)$ arithmetic operations, where $L$ is the number of recieved symbols.
+We focus on signaling constellations that have symbols evenly distributed on the complex unit circle, i.e., $M$-ary phase shift keying, and consider specifically the case where a number of different constellations sizes $M$ are used simultaneously.
+This scenario is common in practical communications systems employing what is called \emph{unequal error protection}.
+We describe an algorithm to compute the least squares estimators of carrier phase and amplitude that requires $O(L \log L)$ arithmetic operations, where $L$ is the number of received symbols.
 \end{abstract}
 \begin{IEEEkeywords}
-Coherent detection, noncoherent detection, phase shift keying, asymptotic statistics
+Coherent detection, noncoherent detection, phase shift keying, unequal error protection
 \end{IEEEkeywords}
@@
 \]
-where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M \geq 1$ is the size of the constellation.
+where $j = \sqrt{-1}$ and $u_i$ is from the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$ and $M \geq 2$ is the size of the constellation.
 We assume that some of the transmitted symbols are \emph{pilot symbols} known to the receiver and the remainder are information carrying \emph{data symbols} with phase that is unknown to the receiver.
@@
 Let $G$ be the set of integers for which $D_m$ is not empty, i.e.
-$D_m \neq \emptyset$ whenever $m \in G$.
+$\abs{D_m} > 0$ if and only if $m \in G$.
 We have
@@
 \]
-In practical settings the number of different constellations used $\abs{G}$ is small.
-A practical example might be that BPSK symbols are used for important symbols and QPSK symbols are used for less important symbols.
+In practical settings the number of different constellations used $\abs{G}$ is typically small.
+A practical example might be that BPSK is used for important symbols and QPSK is used for less important symbols.
 In this example $G = \{2,4\}$ and $\abs{G} = 2$.
@@
 The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
-However, as we will show, the estimator works well under less stringent assumptions.
+However, the estimator works well under less stringent assumptions.
 %It is the least squares estimator that we primarily study in this paper.
@@
 A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
-The idea is to `strip' the modulation from the recieved signal by taking $y_i / \abs{y_i}$ to the power of $M$.
+The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.
 A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
@@
 Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-A caveat is that it is not obvious how pilot symbols or data symbols using mulitple constellations should be included in this estimator.
+A caveat is that it is not obvious how pilot symbols or data symbols using multiple constellations should be included in this estimator.
 % One approach is to compute
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered only the noncoherent setting and where only one constellation size is used, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols and multiple constellation sizes.
+Both Mackenthun and Swelden's considered only the noncoherent setting and where only one constellation size is used.
+We show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols and multiple constellation sizes.
 %Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
@@
 Our algorithm is motivated by the algorithm of Mackenthun~\cite{Mackenthun1994} who considered the noncoherent setting and where only one constellation size is used for the data symbols.
-For the purpose of anaylsing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
-In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.
+For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
+In this case our algorithm requires $O(L \log L)$ arithmetic operations.
 
@@
 \end{equation}
-where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.
+and where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.
 Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
@@
 \]
-It turns out that there are atmost $H$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
+It turns out that there are at most $H$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
 %When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
@@
 The number of elements in $T$ is $H = \abs{T} = O(L)$.
-Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ sorted in accending order of the first element in the triple, that is, if $t_{k} = (t_{k1},t_{k2},t_{k3})$ then $t_{i1} \leq t_{k1}$ whenever $i < k$.
+Let $t_1,\dots,t_H$ be an enumeration of the triples from $T$ sorted in ascending order of the first element in the triple, that is, if $t_{k} = (t_{k1},t_{k2},t_{k3})$ then $t_{i1} \leq t_{k1}$ whenever $i < k$.
 Put
@@
 Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.
-Let $\hat{k} = \arg\min SS(f_k)$.
+Let $\hat{k} = \arg\min_{k\in\{1,\dots,H\}} SS(f_k)$.
 The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata},
@@
 Pseudocode is given in Algorithm~\ref{alg:loglinear}.
-The function $\operatorname{sort}$ on line~\ref{alg_sortindices} return the triple $t_1,\dots,t_H$ so that the first element in the triples is in ascending order.
-This requires $O(L \log L)$ operations.
+The function $\operatorname{sort}$ on line~\ref{alg_sortindices} return the triples $t_1,\dots,t_H$ so that the first element is in ascending order.
+This requires $O(H \log H) = O(L \log L)$ operations.
 The $\operatorname{sort}$ function is the primary bottleneck in this algorithm when $L$ is large.
@@
 \SetAlTitleFnt{}
-\caption{Mackenthun's algorithm with pilot symbols}
+\caption{Mackenthun's algorithm with pilot symbols and multiple constellations sizes}
 \DontPrintSemicolon
@@
 We present the results of Monte-Carlo simulations with the least squares estimator.
-In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+In the simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
 Under these conditions the least squares estimator is also the maximum likelihood estimator.
-Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
+Simulations are run with $G = \{2,4\}$ (BPSK and QPSK symbols) and pilot symbols and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB}.
 The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.
-When $\abs{P} \neq 0$ (i.e.
-coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
-Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
-The solid lines are the MSEs predicted a theoretical result not described here.
-%The theory accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large and when the SNR is not too small.
-As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+Figure~\ref{fig:plotphaseBPSK} shows the mean square error (MSE) of the phase estimator versus SNR.
+The number of symbols if $L = 30, 300$ and $3000$.
+In each case there are $\tfrac{L}{3}$ pilot symbols, $\tfrac{L}{3}$ BPSK symbols and $\tfrac{L}{3}$ QPSK symbols.
+The dots show the sample MSE of the least squares estimator.
+The dashed line is the sample MSE of the unmodulated carrier estimator, that is, the estimator that results when all symbols are known apriory at the receiver, i.e., all symbols are pilot symbols.
 
-% Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
-The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
-The dots and crosses show the results of Monte-Carlo simulations.
-Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
-This requires $G(0)$ to be known.
-In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
-When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
-However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
-
-Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
-The figure depicts an interesting phenomenon.
-When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.
-When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
-A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
-%When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
-This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
-For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
-\[
-SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
-\]
-where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
-%For fixed $\beta$ the asymptotic properties of this weighted estimator could be derived using the techniques we have developed in Sections~\ref{sec:stat-prop-least},~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} and also in~\cite{McKilliam_leastsqPSKpilotsdata_2012appendix}.
-This would enable a rigorous theory for selection of $\beta$ at the receiver.
-One caveat is that the receiver would require knowledge about the noise distribution in order to advantageously choose $\beta$.
-We do not investigate this further here.
-
- \begin{figure}[p]
+ \begin{figure}[htbp]
  	\centering
@@
 
-\section{Conclusion}
-
-We considered least squares estimators of carrier phase and amplitude from noisy communications signals that contain both pilot signals, known to the receiver, and data signals, unknown to the receiver.
-We focused on $M$-ary phase shift keying constellations.
-The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
-
-%We showed, under some conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
-This bias is large when the signal to noise ratio is small.
-It would be interesting to investigate methods for correcting this bias.
-A method for estimating $G(0)$ at the receiver appears to be required.
-
-Monte Carlo simulations were used to assess the performance of the least squares estimator.
-Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
-This suggests the use of a weighed objective function, which would be an interesting topic for future research.
-
 \small

### 2012-12-06 paper.tex

@@
 If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $d_i(\theta)$ for each $i \in D$.
+Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $d_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
 

### 2012-12-19 paper.tex
minor

@@
 where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.
-The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$, and we let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
-
-We assume that time offset estimation has been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal is,
+The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$  where $\emptyset$ is the empty set, and we let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
+
+We assume that time offset estimation has been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal after matched filtering is,
 \begin{equation}\label{eq:sigmod}
@@
 
-The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$ where $\emptyset$ is the empty set).
+The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$).
 In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden's considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
+Both Mackenthun and Swelden considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
@@
 
-It is straightforward to generalise the results we present here to allow data symbols with varying constellation size, i.e.
+It is possible to generalise the results we present here to allow data symbols with varying constellation size, i.e.
 varying $M$.
@@
 Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle^2$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
+We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
@@
 \]  
-Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to suppress the argument $(\{ d_i, i \in D\})$ for notational clarity.
+Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to suppress the argument $(\{ d_i, i \in D\})$ for notational brevity.
 Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain the $SS$ conditioned on minimisation with respect to $a$,
@@
 \[
-z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}.
-\]
-Then,
+z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}
+\]
+be phase difference between the received symbol $y_i$ and the hard decision resulting when $\theta = 0$, i.e. $\round{\angle{y_i}}$.  Then,
 \begin{equation}\label{eq:uicombos}
@@
 be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$.
-If $\hat{\theta}$ is the least squares estimator the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
+If $\hat{\theta}$ is the least squares estimator of the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
 Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $d_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
@@
 \]  
-The next sequence $f_1$ is given by replacing the element $b_{\sigma(1)}$ in $f_0$ with $b_{\sigma(1)}e^{-j2\pi/M}$.
+The next sequence $f_1$ is given by replacing the element $b_{\sigma(0)}$ in $f_0$ with $b_{\sigma(0)}e^{-j2\pi/M}$.
 Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.
@@
 Pseudocode is given in Algorithm~\ref{alg:loglinear}.
-Line~\ref{alg_sortindices} contains the function $\operatorname{sortindicies}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
+Line~\ref{alg_sortindices} contains the function $\operatorname{sortindices}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
 The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
@@
 }
-$\sigma = \operatorname{sortindicies}(z)$ \nllabel{alg_sortindices} \;
 $Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
@@
 $\eta = e^{j2\pi/M} - 1$ \;
+$\sigma = \operatorname{sortindices}(z)$ \nllabel{alg_sortindices} \;
 \For{$k= 0$ \emph{\textbf{to}} $M\abs{D}-1$ \nllabel{alg_loop_search}}{
@@
 \begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
 \item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
@@
 and assume that $\frac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\frac{\abs{D}}{L} = d + o(L^{-1/2})$.
-Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
+Put $\hat{\lambda}_L = -\sfracpart{\hat{\theta} - \theta_0}_\pi = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
@@
 Before giving the proofs we discuss the assumptions made by these theorems.
-The assumption that $w_1, \dots w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
+The assumption that $w_1, \dots, w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
 If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $\{s_i,i \in P \cup D\}$.
@@
 This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
-So, for example, we expect the phases $0.49\pi$ and $-0.49\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.49\pi - 0.49\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.49\pi - 0.49\pi\vert = 0.98\pi$.
+So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.99\pi - 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.99\pi - 0.99\pi\vert = 1.98\pi$.
 
@@
 \end{equation}
-and put $\hat{\lambda}_L = \langle\theta_0 - \hat{\theta}\rangle_\pi$.
+and put $\hat{\lambda}_L = -\sfracpart{\hat{\theta} - \theta_0}_\pi = \sfracpart{\theta_0 - \hat{\theta}}_\pi$.
 Since $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ then  $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)$.
@@
 We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
-Once this is done we will be able to prove the normality of $\sqrt{L} m_L$.
+Once this is done we will be able to prove the normality of $\sqrt{L} \hat{m}_L$.
 Recall that $\hat{\lambda}_L$ is the maximiser of the function $G_L$ defined in~\eqref{eq:GLdefn}.
@@
 
-We now analyse the asymptotic distribution of $\sqrt{L} m_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
-\begin{align*}
-\sqrt{L} m_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
+We now analyse the asymptotic distribution of $\sqrt{L} \hat{m}_L$.  Let $T_L(\lambda) = \expect G_L(\lambda)$.  Using~\eqref{eq:rhoGLZ},
+\begin{align*}
+\sqrt{L} \hat{m}_L &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - G(0) \big) \\
 &= \sqrt{L} \rho_0 \big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) + T_L(\hat{\lambda}_L) - G(0) \big).
@@
 \]
-It follows that $\sqrt{L} m_L =  \rho_0 X_L + o_P(1)$.
+It follows that $\sqrt{L} \hat{m}_L =  \rho_0 X_L + o_P(1)$.
 Lemma~\ref{lem:XL} shows that the distribution of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
-Thus, the distribution of $\sqrt{L} m_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
+Thus, the distribution of $\sqrt{L} \hat{m}_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
 Because $X_L$ does not depend on $\hat{\lambda}_L$, it follows that $\covar(X_L, \sqrt{L}\hat{\lambda}_L) = 0$, and so,
@@
 \]
-Theorem~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+Theorems~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
 %and the joint density function of the real and imaginary parts of $Z$ is
@@
 The dots and crosses show the results of Monte-Carlo simulations.
-Each point is computed as $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+Each point is computed as the unbiased error $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
 This requires $G(0)$ to be known.
@@
 However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+As indicated in Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.
+However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.
 
@@
 This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
-For example, rather the minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
+For example, rather than minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
 \[
@@
 Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
-Interestingly, when the SNR is small, it is counter productive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
+Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
 This suggests the use of a weighed objective function, which would be an interesting topic for future research.
@@
 &\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda) + T_L(\lambda) - G(\lambda)} \\
+&= \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda) + T_L(\lambda) - G(\lambda)} \\
 &\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)}.
@@
 \begin{IEEEproof}
-Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipshitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
+Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipschitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
 \begin{align*}
@@
 &= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
-&= SS(\{\hat{d}_i, i \in D\}) - D,
+&= SS(\{\hat{d}_i, i \in D\}) - C,
 \end{align*}
@@
 \[
-D = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
+C = \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) +  \frac{1}{L}\sabs{ \eta y_k }^2.
 \]
@@
 \begin{equation}\label{eq:Ddefn}
-D = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
+C = 2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) + \frac{1}{L}\sabs{\eta}^2\rho_0^2 R_k^2.
 \end{equation}
@@
 \[
-0 > \sin(v) > -\frac{\rho_0 \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}.
-\]
-Substituting this into~\eqref{eq:Ddefn} gives $D > 0$, but then 
+0 > \sin(v) > -\frac{\rho_0 \sabs{\eta}^2 R_k}{4 L \hat{\rho} \sin(\pi/M)}.
+\]
+Substituting this into~\eqref{eq:Ddefn} gives $C > 0$, but then 
 \[
@@
 \sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&= s_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
+&= r_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
 \end{align*}
@@
 \[
-s_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
+r_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
 \]
@@
 \]
-Observe that both $s_k$ is continuous and that 
-\[
-s_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
+Observe that $r_k$ is continuous and that 
+\[
+r_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
 \]
@@
 \[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
+\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + r_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
 \]
@@
 v_k(\lambda,\Phi_1) = \begin{cases}
-s_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+r_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
 \lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
@@
 \begin{equation}\label{eq:ck}
-c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} s_k(\phi) g(\phi) d\phi.
+c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} r_k(\phi) g(\phi) d\phi.
 \end{equation}
@@
 \[
-s_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
+r_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
 \]
@@
 \begin{equation}\label{eq:YLdef}
-Y_L(\lambda) = \sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) - X_L.
+Y_L(\lambda) = \sqrt{L}\big( G_L(\lambda) - T_L(\lambda) \big) - X_L.
 \end{equation}

### 2012-12-19 paper.tex
fixed silly errors with figures

@@
 However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
-As indicated in Figures~\ref{fig:plotampM2}~and~\ref{fig:plotampM4}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.
+As indicated in Figures~\ref{fig:plotampBPSK}~and~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.
 However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.

### 2012-12-19 paper.tex
fixed a few estimator - parameter, parameter - estimator things.

@@
 The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
-Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{a_0 - \hat{a}}^2$.
-It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ and $(\rho_0 - \hat{\rho})^2$.
+Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{\hat{a} - a_0}^2$.
+It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ and $(\hat{\rho} - \rho_0)^2$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi^2$ rather than $(\theta_0 - \hat{\theta})^2$ is the appropriate measure of error for the phase parameter.
+It will become apparent why $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ rather than $(\hat{\theta} - \theta_0)^2$ is the appropriate measure of error for the phase parameter.
 
@@
 
-The theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ rather than directly on $\theta_0 - \hat{\theta}$.
+The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\theta_0  - \theta_0$.
 This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
@@
 This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\hat{\theta} - \theta_0}_\pi$ will not converge.
 One can check that the assumption holds when $w_1$ is circularly symmetric and normally distributed.

### 2012-12-20 paper.tex
added a few of Barry's edits

@@
 Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
-A caveat of this estimator is that it is not entirely obvious how pilot symbols should be included.
-Techniques are available but they are somewhat adhoc~\cite{Cowley_ref_sym_carr_1998}.
+A caveat of this estimator is that it is not obvious how pilot symbols should be included.
+%Techniques are available but they are somewhat ad hoc~\cite{Cowley_ref_sym_carr_1998}.
 This problem does not occur with the least square estimator.
@@
 Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi^2$ is asymptotically normally distributed as $L\rightarrow \infty$.
+We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
@@
 SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
-&= \sum_{i \in P \cup D} \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^*, \label{eq:SS}
+&= \sum_{i \in P \cup D} \big( \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^* \big), \label{eq:SS}
 \end{align}
-where $*$ denotes the complex conjugate.
-Fixing the data symbols $\{d_i, i \in D\}$, differentiating with respect to $a^*$, and setting the resulting equation to zero we find the least squares estimator of $a_0$ as a function of $\{d_i, i \in D\}$,
+where $*$ denotes complex conjugate.  The minimiser of $SS$ with respect to $a$ as a function of $\{d_i, i \in D\}$ is
 \begin{equation}\label{eq:hata}
@@
 \end{equation}
-where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ is a constant.
+where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ does not depend on the $d_i$.
 Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
@@
 \]
-be phase difference between the received symbol $y_i$ and the hard decision resulting when $\theta = 0$, i.e. $\round{\angle{y_i}}$.  Then,
+be the phase difference between the received symbol $y_i$ and the hard decision resulting when $\theta = 0$, i.e. $\round{\angle{y_i}}$.  Then,
 \begin{equation}\label{eq:uicombos}
@@
 \end{align*}
-where $B = \sum_{i \in P} y_i p_i^*$ is a constant, independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
+where $B = \sum_{i \in P} y_i p_i^*$ is independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
 Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
@@
 \]
-So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a constant number of operations.
-Given $Y_k$, the value of $SS(f_k)$ can be computed in a constant number of operations using~\eqref{eq:SSfk}.
+So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a $O(1)$ arithmetic operations.
+Given $Y_k$, the value of $SS(f_k)$ can be computed in $O(1)$ operations using~\eqref{eq:SSfk}.
 Let $\hat{k} = \arg\min SS(f_k)$.
-The least squares estimator of the complex amplitude is then computed according to~\eqref{eq:hata},
+The least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata},
 \begin{equation}\label{eq:ahatYhat}
@@
 %where both $\hat{\rho}$ and $\rho_0$ are positive real numbers.
-Put $L = \abs{P \cup D}$ and let $p$ and $d$ be the limits
+Put $L = \abs{P \cup D}$ and let $\abs{P}$ and $\abs{D}$ increase in such a way that
 \[
@@
 \[
-G(x) = p h_1(x) + d h_2(x) \qquad \text{where,}
+G(x) = p h_1(x) + d h_2(x) \qquad \text{where}
 \]
@@
 \]
-If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then:
+If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$,
 \begin{enumerate}
@@
 If $w_i$ is not circularly symmetric then the distribution of $R_i$ and $\Phi_i$ may depend on $a_0$ and also on the transmitted symbols $\{s_i,i \in P \cup D\}$.
-In result, the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $\{s_i,i \in P \cup D\}$, rather than just $\rho_0$.
+As a result the asymptotic variance described in Theorem~\ref{thm:normality} depends on $a_0$ and $\{s_i,i \in P \cup D\}$, rather than just $\rho_0$.
 The circularly symmetric assumption may not always hold in practice, but we feel it provides a sensible trade off between simplicity and generality.
@@
 This is required in Theorem~\ref{thm:normality} so that the constants $A_1$, $A_2$, $B_1$ and $B_2$ exist.
-We will also use that $R_i$ has finite variance to simplify the proof of Theorem~\ref{thm:consistency} by use of Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}
-
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\theta_0  - \theta_0$.
+We will also use the fact that $R_i$ has finite variance to simplify the proof of Theorem~\ref{thm:consistency} by use of Kolmogorov's strong law of large numbers~\cite{Billingsley1979_probability_and_measure}.
+
+The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta}  - \theta_0$.
 This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
@@
 \[
-\frac{\abs{P}}{L} = p + o(L^{-1/2}) \qquad \text{and} \qquad \frac{\abs{D}}{L} = d + o(L^{-1/2})
-\] 
-where $o(L^{-1/2})$ denotes a number such that $\sqrt{L} o(L^{-1/2})$ goes to zero as $L\rightarrow \infty$.
+\frac{\abs{P}}{L} = p + o(L^{-1/2}) \qquad \text{and} \qquad \frac{\abs{D}}{L} = d + o(L^{-1/2}).
+\] 
+%where $o(L^{-1/2})$ denotes a number such that $\sqrt{L} o(L^{-1/2})$ goes to zero as $L\rightarrow \infty$.  
 This stronger condition is required to prove the asymptotic normality of $\sqrt{L}\hat{m}_L$.
 
-The next two sections give proofs for Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  The proofs make use of various lemmas.  The lemmas are proved in the appendix.
+The next two sections give proofs of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  The proofs make use of various lemmas, which are proved in the appendix.
 

### 2012-12-28 edits/barryedits.tex
Added some more of Barry's edits.

\documentclass[a4paper,12pt]{article}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

\usepackage[square,comma,numbers,sort&compress]{natbib}

\begin{document}

\begin{enumerate}
\item Page 1. Column 2. I've left the sentence: ``The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.'' in.  If I remove it, Bill and Andre will want to put it back.  I believe it does give the correct intuition, even if it is imprecise.
%\item Regarding the ``conditioned upon minimisation with respect to'' phrases, I have replaced ``conditioned upon'' with ``concentrated by''.  This is perhaps a bit lazy, but I 
\item Page 6, column 2, you want me to add reference for the polynomial phase estimation papers, where we talk about differentiability of the objective function at the estimator.  I've added a reference to the article on arxiv.  We can change it to the trans. sig. proc. papers later.
\item Page 6, column 2, have left in the definition of $o_P(1)$, since it doesn't hurt.
\item Page 10 column 2, I have left ``be $N$ points uniformly spaced on the interval $[-\pi, \pi)$'' since it doesn't hurt.
\item Page 11, column 1, about the phrase ``where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$''.  I want to leave this.  The phrase is taken from one of David Pollard's papers.  It is informal, but I know that if I were personally reading this for the first time, I would be scratching my head at this point until I realised that expectation has been moved outside the supremum.
\item Page 11, column 2, about the Lipschitz continuous statement.  Obviously the term involving $\cos(\phi)$ needs only the mean value theorem.  However the term involving $\cos\fracpart{\phi}$ isn't always differentiable so the mean value theorem doesn't always apply.  It's obviouly Lipschitz continuous though, just plot it.  We can add a proof if you really think it's necessary.
\item Page 12, column 1, ``for $i \in D$ such that $R_i > 0$.'' I think this is necesary, otherwise the surrounding statement is not always true.
\item Lemma 11, I would also like there to be a nicer proof!  I haven't found one though.
\item 
\end{enumerate}

\end{document}

### 2012-12-28 paper.tex
Added some more of Barry's edits.

@@
 The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
-Consequently, our metric of performance will not be symbol or bit error rate, but $\abs{\hat{a} - a_0}^2$.
-It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ and $(\hat{\rho} - \rho_0)^2$.
+Consequently, our metric of performance will not be symbol or bit error rate, but $\sabs{\hat{a} - a_0}$.
+It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\sabs{\langle\hat{\theta} - \theta_0\rangle_\pi}$ and $\sabs{\hat{\rho} - \rho_0}$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ rather than $(\hat{\theta} - \theta_0)^2$ is the appropriate measure of error for the phase parameter.
+It will become apparent why $\langle\hat{\theta} - \theta_0\rangle_\pi$ rather than $\hat{\theta} - \theta_0$ is the appropriate measure of error for the phase parameter.
 
@@
 Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi^2$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
+We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
@@
 Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to suppress the argument $(\{ d_i, i \in D\})$ for notational brevity.
-Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain the $SS$ conditioned on minimisation with respect to $a$,
+Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain $SS$ minimised with respect to $a$,
 \begin{equation}\label{eq:SSdatasymbols}
@@
 where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ does not depend on the $d_i$.
+The least squares estimators of the data symbols are the minimisers of~\eqref{eq:SSdatasymbols}.
 Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
@@
 \end{align}
+We have slightly abused notation here by reusing $SS$.
+This should not cause confusion as $SS(a, \{d_i, i \in D\})$, $SS(\rho, \theta, \{d_i, i \in D\})$, and $SS(\{d_i, i \in D\})$ are easily told apart by their arguments.
 For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
@@
 \]
-If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$,
+If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$ then
 \begin{enumerate}
@@
 
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ conditioned on minimisation with respect to the data symbols,
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ minimised with respect to the data symbols,
  \begin{align*}
@@
 and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
-Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$ conditioned on minimisation with respect to the data symbols,
+Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
 \begin{equation}\label{eq:hatrhoZ}
@@
 \end{equation}
-where $\Re(\cdot)$ denotes the real part.
-Since $\hat{\rho}(\theta)$ is nonnegative it follows that $\Re(Z(\theta))$ is nonnegative for all $\theta \in [-\pi, \pi)$.
-Substituting this expression into $SS(\rho, \theta)$ gives $SS$ conditioned upon minimisation with respect to $\rho$ and the data symbols,
+where $\Re(\cdot)$ denotes the real part.  %Since $\hat{\rho}(\theta)$ is nonnegative it follows that $\Re(Z(\theta))$ is nonnegative for all $\theta \in [-\pi, \pi)$.  
+Substituting this expression into $SS(\rho, \theta)$ gives $SS$ minimised with respect to $\rho$ and the data symbols,
 \[
@@
 \]
-
-We are interested in analysing $\hat{\theta}$, the minimiser of $SS(\theta)$, or equivalently, the maximiser of $\Re(Z(\theta))$.
-Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
+We again abuse notation by reusing $SS$, but this should not cause confusion as $SS(\rho,\theta)$ and $SS(\theta)$ are easily told apart by their inputs.
+By definition the amplitude $\rho_0$ and its estimator $\hat{\rho}$ are positive.
+However, $\hat{\rho}(\theta) = \Re(Z(\theta))$ may take negative values for some $\theta \in [-\pi,\pi)$.
+%In this case we would replace $\theta$ with $\fracpart{\theta + \phi}$ and $\hat{\rho}(\theta)$ with $-\hat{\rho}(\theta)$ and observe that
+%\[
+%SS(\rho,\theta) = SS(-\rho, \fracpart{\theta + \pi})
+%\]
+The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $SS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.  Equivalently
+\[
+\hat{\theta} = \arg\max_{\theta \in [-\pi,\pi)} \Re(Z(\theta))
+\]
+with no constraints required.
+
+We are thus interested in analysing the behaviour of the maximiser of $\Re(Z(\theta))$.  Recalling the definition of $R_i$ and $\Phi_i$ from~\eqref{eq:RiandPhii},
 \begin{align*}
@@
 and put $\hat{\lambda}_L = -\sfracpart{\hat{\theta} - \theta_0}_\pi = \sfracpart{\theta_0 - \hat{\theta}}_\pi$.
-Since $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ then  $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)$.
+Since $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ it follows that $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)$.
 We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
@@
 \]
-and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$, then
+and since $\frac{\abs{P}}{L} \rightarrow p$ and $\frac{\abs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$,
 \[
@@
 \]
-As is customary, let $(\Omega, \mathcal F, \prob)$ be the probability space over which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ upon which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
+As is customary, let $\Omega$ be the sample space on which the random variables $\{w_i\}$ are defined.
+Let $A$ be the subset of the sample space $\Omega$ on which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
 Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
-Let $A'$ be the subset of the sample space upon which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
+Let $A'$ be the subset of the sample space on which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
 Because $G(x)$ is uniquely maximised at $x=0$, it follows that $G(\hat{\lambda}_L) \rightarrow G(0)$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
@@
 However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
+Similar properties have been used to analyse the behaviour of other estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.%\cite{McKilliam_LSU_polyest_part1_2012,McKilliam_LSU_polyest_part2_2012}.
 Define the function
@@
 \]
-where $o_P(1)$ denotes a number converging in probability to zero as $L \rightarrow \infty$, and the constants $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+where $o_P(1)$ denotes a number converging in probability to zero as $L \rightarrow \infty$, and $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
 Lemma~\ref{lem:empiricprocc} shows that
@@
 \]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are real constants defined in the statement of Theorem~\ref{thm:normality}.
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are defined in the statement of Theorem~\ref{thm:normality}.
 It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
@@
 \[
-\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) = \covar(\rho_0 X_L + o_P(1), \sqrt{L}\hat{\lambda}_L) \rightarrow 0
-\]
-as $L \rightarrow \infty$.  It remains to prove the lemmas that we have used.  These lemmas are proved in Section~\ref{sec:proof-asympt-normappendix} of the appendix.
+\covar(\sqrt{L}\hat{m}_L, \sqrt{L}\hat{\lambda}_L) \rightarrow \covar(\rho_0 X_L, \sqrt{L}\hat{\lambda}_L) = 0
+\]
+as $L \rightarrow \infty$.  The lemmas that we have used are proved in Section~\ref{sec:proof-asympt-normappendix} of the appendix.
 
@@
 %g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-g(\phi) = \frac{\cos(\phi)}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2(\phi)} \big(1 + \kappa^2\cos^2(\phi) \big)
-\]
-where $\kappa = \rho_0/\sigma$ and
-\[
-\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos(\phi)}{\sqrt{2}}\right)
-\]
-and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
+g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
+\]
+where $\kappa = \rho_0/\sigma$ and $\Psi(x) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{x} e^{-x^2/2} dx$
+is the cumulative density function of the standard normal.
+%\[
+%\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos\phi}{\sqrt{2}}\right)
+%\]
+%and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  
+The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
 
@@
 \end{align*}
-where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  Since 
-\[
-h_1(\lambda) = \expect R_1\cos(\lambda + \Phi_1) \leq \expect R_1,
+%where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  
+Since 
+\[
+\sabs{h_1(\lambda)} = \sabs{\expect R_1\cos(\lambda + \Phi_1)} \leq \expect R_1,
 \]
@@
 \[
-h_2(\lambda) = \expect R_1\cos\fracpart{\lambda + \Phi_1} \leq \expect R_1
+\sabs{h_2(\lambda)} = \sabs{\expect R_1\cos\fracpart{\lambda + \Phi_1}} \leq \expect R_1
 \] 
@@
 \begin{lemma}\label{lem:uniflawTGL} 
-Put $T_L(\lambda) = \expect G_L(\lambda)$, then
+Put $T_L(\lambda) = \expect G_L(\lambda)$.  Then
 \[
@@
 \]
-be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n$ be the interval $[\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
+be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n = [\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
 \begin{align*}
@@
 \[
-U_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)},
+U_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)} \;\;\; \text{and}
 \]
@@
 \]
-Lemma~\ref{lem:supDLlambdan} shows that for any finite constant $N$, and any $\epsilon > 0$, 
+Lemma~\ref{lem:supDLlambdan} shows that for any $N$ and $\epsilon > 0$, 
 \[
@@
 \] 
-Choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ and
+If we choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ then
 \begin{align*}
@@
 \end{align*}
-Thus, $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
+Thus $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
 \end{IEEEproof}
@@
 \begin{lemma}\label{lem:supDLlambdan}
-With $N$ a finite constant $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+For any $N > 0$, $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$ where $U_L$ is defined in the proof of Lemma~\ref{lem:uniflawTGL}.
 \end{lemma}
@@
 \]
-almost surely as $L\rightarrow\infty$ by Kolmogorov's strong law of large numbers~\cite[Theorem 2.3.10]{SenSinger_large_sample_stats_1993}.  Since $N$ is a constant,
+almost surely as $L\rightarrow\infty$ by Kolmogorov's strong law of large numbers~\cite{Billingsley1979_probability_and_measure}.  Thus
 \[
@@
 \end{align*}
-Because this results does not depend on $n$,
+Because these results do not depend on $n$,
 \[
@@
 \]
-for all $i = P \cup D$.  Now,
+for all $i = P \cup D$.  Also
 \begin{align*}
@@
 \]
-where $\eta = e^{-j2\pi/M} - 1$.  Observe that if $R_i > 0$ then $q_i < 0$.  Lemma~\ref{lem:fracpartlambdahatnotpi} shows that,
+where $\eta = e^{-j2\pi/M} - 1$.  Observe that $q_i < 0$ when $R_i > 0$.  Lemma~\ref{lem:fracpartlambdahatnotpi} shows that
 \[
@@
 &= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
-&= -2 \sin(\tfrac{\pi}{M}) e^{j ( v + \pi/2) },
+&= -2 j \sin(\tfrac{\pi}{M}) e^{j v },
 \end{align*}
 and
-\begin{align*}
-\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) &= -2 \sin(\tfrac{\pi}{M}) \cos(v + \tfrac{\pi}{2}) \\
- &= 2 \sin(\tfrac{\pi}{M}) \sin(v).
-\end{align*}
+\[
+\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) = 2 \sin(\tfrac{\pi}{M}) \sin(v).
+\]
 Because we assumed~\eqref{eq:fraclambassumption}, it follows that $0 > v > q_k$ and, from the definition of $q_k$,
 \[
-0 > \sin(v) > -\frac{\rho_0 \sabs{\eta}^2 R_k}{4 L \hat{\rho} \sin(\pi/M)}.
+ -\frac{\rho_0 \sabs{\eta}^2 R_k}{4 L \hat{\rho} \sin(\pi/M)} <  \sin(v) < 0.
 \]
@@
 \begin{lemma}\label{lem:Qconv}
-Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  Then,
+Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ where the function $R_L$ is defined in~\eqref{eq:RLdef}.  We have
 \[ 
@@
 \]
-where $p, d$ and $H$ are real numbers defined in the statements of Theorems~\ref{thm:consistency}~and~\ref{thm:normality} and $o_P(1)$ denotes a number converging in probability to zero as $L\rightarrow\infty$.
+where $p, d$ and $H$ are defined in the statements of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
 \end{lemma}
@@
 \end{align}
-We analyse $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\lambda = 0$,
-\[
-\sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
-\]
-where $\zeta$ is a continuous function with $\zeta(0) = 0$.  So,
-\begin{align*}
-k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
-&= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big),
-\end{align*}
-since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ as a result of~\eqref{eq:expectRepartRphi}.
+Lemma~\ref{lem:q1k1parts} shows that $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$ and Lemma~\ref{lem:q2k2parts} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H + o_P(1) \big)$ and so
+\begin{align*}
+Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
+&= \hat{\lambda}_L \big( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \big) \\
+ &= \hat{\lambda}_L \big( p  + d H + o_P(1) \big),
+ \end{align*}
+since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the 
+above equation by $\sqrt{L}$.
+% Consider $k_1(\lambda)$ first.  By a Taylor expansion of $\sin(\lambda + \Phi_1)$ about $\lambda = 0$,
+% \[
+% \sin(\lambda + \Phi_1) - \sin(\Phi_1) = \lambda \big( \cos(\Phi_1) + \zeta(\lambda) \big),
+% \]
+% where $\zeta$ is a continuous function with $\zeta(0) = 0$.  So,
+% \begin{align*}
+% k_1(\lambda) &= \expect R_1\lambda(\cos(\Phi_1) \; + \; \zeta(\lambda)) \\
+% &= \lambda \big( 1 \; + \; \zeta(\lambda) \expect R_1 \big),
+% \end{align*}
+% since $h_1(0) = \expect R_1\cos(\Phi_1) = 1$ as a result of~\eqref{eq:expectRepartRphi}.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, then $\zeta(\hat{\lambda}_L)$ converges almost surely (and therefore also in probability) to zero as $L \rightarrow\infty$.
 So,
-\[
-k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) \expect R_1 \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big)
-\]
-because $\expect R_1$ is finite.  Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
-\begin{align*}
-Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
-&= \hat{\lambda}_L \left( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \right) \\
-&= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
-\end{align*}
-since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the above equation by $\sqrt{L}$.
+% \[
+% k_1(\hat{\lambda}_L) = \hat{\lambda}_L \big( 1 \; + \; \zeta(\hat{\lambda}_L) \expect R_1 \big) = \hat{\lambda}_L \big( 1  + o_P(1)\big)
+% \]
+% because $\expect R_1$ is finite.  Lemma~\ref{lem:k2conv} shows that $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) )$.  So,
+% \begin{align*}
+% Q_L(\hat{\lambda}_L) &=  \tfrac{\sabs{P}}{L} \hat{\lambda}_L \big( 1  + o_P(1)\big)  + \tfrac{\sabs{D}}{L} \hat{\lambda}_L ( H + o_P(1) ) \\
+% &= \hat{\lambda}_L \left( \tfrac{\sabs{P}}{L}  + \tfrac{\sabs{D}}{L} H + o_P(1) \right) \\
+% &= \hat{\lambda}_L \left( p  + d H + o_P(1) \right),
+% \end{align*}
+% since $\frac{\sabs{P}}{L} \rightarrow p$ and $\frac{\sabs{D}}{L} \rightarrow d$ as $L \rightarrow \infty$.  The lemma follows by multiplying both sides of the above equation by $\sqrt{L}$.
 \end{IEEEproof}
 
-\begin{lemma}\label{lem:k2conv}
-$k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
+\begin{lemma}\label{lem:q1k1parts}
+Put
+\[
+q_1(\lambda) + j k_1(\lambda) = \expect\big( R_1 e^{j(\lambda + \Phi_1)} - R_1 e^{j\Phi_1}  \big).
+\]
+It follows that $q_1(\lambda) = o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$
+\end{lemma}
+\begin{proof}
+We have
+\[
+q_1(\lambda) + j k_1(\lambda) = (e^{j\lambda} - 1) \expect R_1e^{j\Phi_1} = e^{j\lambda} - 1
+\]
+since the mean of $R_1e^{j\Phi_1}$ is 1.
+Thus $q_1(\lambda) = \cos(\lambda) - 1$ and $k_1(\lambda) = \sin(\lambda)$.
+Now $q_1(\hat{\lambda}_L) = o_P(1)$ since $\hat{\lambda}_L$ converges almost surely to zero as $L\rightarrow\infty$.
+By a Taylor expansion of $\sin(\lambda)$ about $\lambda = 0$ we obtain $k_1(\lambda) = \lambda \big( 1 + \zeta(\lambda) \big)$  where $\zeta(\lambda)$ is a continuous function with $\zeta(0) = 0$.
+Thus $\zeta(\hat{\lambda}_L) = o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$.
+\end{proof}
+
+\begin{lemma}\label{lem:q2k2parts}
+Put
+\[
+q_2(\lambda) + j k_2(\lambda) = \expect\big( R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big).
+\]
+It follows that $q_2(\hat{\lambda}_L) = o_P(1)$ and $k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H + o_P(1) \big)$ where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
+\end{lemma}
+\begin{proof}
+Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $q_2(\lambda)$ and $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
+We will examine $q_2(\lambda)$ and $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
+An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
+To keep our notation clean put
+\[
+\psi_k = \frac{2\pi}{M}k + \frac{\pi}{M}
+\] 
+with $k \in \ints$.  When $\Phi_1 \in [\psi_{k-1}, \psi_k - \lambda)$,
+\begin{align*}
+e^{j\fracpart{\lambda + \Phi_1}} - e^{j\fracpart{\Phi_1}} &=  e^{j\big(\lambda + \Phi_1 - \tfrac{2\pi}{M}k\big)} - e^{j\big( \Phi_1- \tfrac{2\pi}{M}k\big)} \\
+&= \big(e^{j\lambda} - 1 \big) e^{j\big(\Phi_1- \tfrac{2\pi}{M}k\big)} \\
+&= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}},
+\end{align*}
+and when $\Phi_1 \in [\psi_{k} - \lambda, \psi_k)$,
+\begin{align*}
+e^{j\fracpart{\lambda + \Phi_1}} - &e^{j\fracpart{\Phi_1}} \\
+&=  e^{j\big(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}\big)} - e^{j\big( \Phi_1- \tfrac{2\pi}{M}k\big)} \\
+&= \big(e^{j\big(\lambda - \tfrac{2\pi}{M}\big)} - 1 \big) e^{j\fracpart{\Phi_1}} \\
+&= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}}  + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)e^{j\fracpart{\Phi_1}}.
+\end{align*}
+Thus, when $\Phi_1 \in [\psi_{k-1}, \psi_k)$,
+\begin{align*}
+e^{j\fracpart{\lambda + \Phi_1}} - e^{j\fracpart{\Phi_1}} &= \big(e^{j\lambda} - 1 \big) e^{j\fracpart{\Phi_1}} \\
+&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)e^{j\fracpart{\Phi_1}} \chi_k(\Phi_1,\lambda)
+\end{align*}
+where 
+\[
+\chi_k(\Phi_1,\lambda) = \begin{cases}
+1, & \Phi_1 \in [\psi_{k} - \lambda, \psi_k)  \\
+0, & \text{otherwise}.
+\end{cases}
+\]
+Now
+\begin{align*}
+q_2(\lambda) + j k_2(\lambda) &= \big(e^{j\lambda} - 1 \big) h_2(0) \\
+&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big)
+\end{align*}
+since $\expect R_1 e^{j\fracpart{\Phi_1}} = h_2(0)$ as a result of Lemma~\ref{lem:expectImRfracpart} and where $\chi(\Phi_1, \lambda) = \sum_{k \in \ints}\chi_k(\Phi_1, \lambda)$.
+Now
+\begin{align*}
+ \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big) &= \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) \chi(\phi,\lambda) dr d\phi \\
+&= \sum_{k\in\ints}\int_{0}^{2\pi} g(\phi) \chi_k(\phi,\lambda) d\phi \\
+&= \sum_{k=0}^{M-1}\int_{\psi_k-\lambda}^{\psi_k} g(\phi)  d\phi \\
+&= \lambda \sum_{k=1}^{M-1} g(\phi_k) 
+\end{align*}
+where for each $k$, $\phi_k \in (\psi_k-\lambda, \psi_k)$.  The third line above follows from
+\end{proof}
+
+\begin{lemma}\label{lem:expectImRfracpart}
+$\expect R_1 \sin\fracpart{\Phi_1} = 0$.
 \end{lemma}
 \begin{IEEEproof}
-Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
-We will analyse $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
+Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
+\begin{align*}
+\expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
+&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi.
+\end{align*}
+The proof is immediate since $g(\phi)$ is even and $\sin\fracpart{\phi}$ is odd.
+% where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
+Proposition~\ref{prop:gg2cont} shows that $g$ is even on $(-\pi,\pi)$.
+Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.
+So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
+% \[
+% g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
+% \]
+% and
+% \[
+% \sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
+% \]
+% Now,
+% \begin{align*}
+% I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
+% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
+% &\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+% &= 0
+% \end{align*}
+% for all integers $k$ because
+% \begin{align*}
+% \int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
+% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
+% &= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
+% \end{align*}
+% Observe that the sets
+% \[
+%  [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
+% \]
+% partition $[-\pi, \pi]$, and so
+% \[
+% \expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
+% \]
+\end{IEEEproof}
+
+% \begin{lemma}\label{lem:k2conv}
+% $k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( H + o_P(1) ).$
+% \end{lemma}
+% \begin{IEEEproof}
+% Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
+We will examine $k_2(\lambda)$ for $0 \leq \lambda < \tfrac{\pi}{M}$.
 An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
 To keep our notation clean put
-\[
-\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
-\]
-For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
-\[
-\fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
-\]
-Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
-\]
-whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
-\[
-\fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
-\]
-So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
-&= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
-\end{align*}
-where, for each $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
-\begin{align*}
-\sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
-&= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
-\end{align*}
-and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
-\begin{align*}
-\sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
-&= r_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
-\end{align*}
-where
-\[
-r_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
-\]
-and
-\[
-\eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos\fracpart{\Phi_1} + \zeta_{k+1}(\lambda).
-\]
-Observe that $r_k$ is continuous and that 
-\[
-r_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
-\]
-
-Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
-\]
-whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + r_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
-\]
-Since the intervals $[\psi_{k-1}, \psi_{k})$ partition the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
-\[
-\sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
-\]
-where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
-\begin{equation}\label{eq:vk}
-v_k(\lambda,\Phi_1) = \begin{cases}
-r_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
-\lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
-0 & \text{otherwise}.
-\end{cases} 
-\end{equation}
-Now,
-\begin{align*}
-k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
-&= \expect R_1\big( \lambda\cos\fracpart{\Phi_1} + v(\lambda, \Phi_1)  \big) \\
-&= \lambda h_2(0) + \expect R_1 v(\lambda, \Phi_1).
-\end{align*}
-Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a continuous function with $\gamma(0) = 0$, and
-\[
-H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0),
-\]
-where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
+% \[
+% \psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}.
+% \]
+% For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_k)$, then
+% \[
+% \fracpart{\Phi_1} = \Phi_1 - \round{\Phi_1} = \Phi_1 - \tfrac{2\pi}{M}k.
+% \]
+% Similarly, if $\Phi_1 \in  [\psi_{k-1}, \psi_{k} - \lambda)$, then
+% \[
+% \fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k,
+% \]
+% whilst, if $\Phi_1 \in [\psi_{k} - \lambda, \psi_{k})$, then
+% \[
+% \fracpart{\Phi_1 + \lambda} = \Phi_1 + \lambda - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}.
+% \]
+% So, when $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$,
+% \begin{align*}
+% \sin\fracpart{\lambda + \Phi_1} - &\sin\fracpart{\Phi_1} \\
+% &= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
+% \end{align*}
+% and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k)$ about $\lambda = 0$, 
+% \begin{align*}
+% \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} &= \lambda\big(\cos(\Phi_1 - \tfrac{2\pi}{M}k) + \zeta_k(\lambda)\big) \\
+% &= \lambda\big(\cos\fracpart{\Phi_1} + \zeta_k(\lambda)\big).
+% \end{align*}
+% where, for each $k$, the function $\zeta_k$ is continuous with $\zeta_k(0) = 0$.  Alternatively, when $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$,
+% \begin{align*}
+% \sin\fracpart{\lambda + \Phi_1} &- \sin\fracpart{\Phi_1} \\
+% &= \sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k)
+% \end{align*}
+% and by a Taylor expansion of $\sin(\lambda + \Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M})$ about $\lambda = 0$,
+% \begin{align*}
+% \sin&\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \\
+% &= r_k(\Phi_1) + \lambda\big(\cos\fracpart{\Phi_1} + \eta_k(\lambda, \Phi_1)  )\big)
+% \end{align*}
+% where
+% \[
+% r_k(\Phi_1) = \sin(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \sin(\Phi_1 - \tfrac{2\pi}{M}k),
+% \]
+% and
+% \[
+% \eta_k(\lambda, \Phi_1) = \cos(\Phi_1 - \tfrac{2\pi}{M}k - \tfrac{2\pi}{M}) - \cos\fracpart{\Phi_1} + \zeta_{k+1}(\lambda).
+% \]
+% Observe that $r_k$ is continuous and that 
+% \[
+% r_k(\psi_{k}) = -2\sin(\tfrac{\pi}{M}) \;\;\; \text{and} \;\;\; \eta_k(\lambda,\psi_k) = \zeta_{k+1}(\lambda).
+% \]
+
+% Let us now collate what we have.  For each integer $k$, if $\Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda )$, then
+% \[
+% \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} + \lambda\zeta_k(\lambda),
+% \]
+% whilst, if $\Phi_1 \in [ \psi_{k} - \lambda,\psi_{k})$, then
+% \[
+% \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda \cos\fracpart{\Phi_1} + r_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1).
+% \]
+% Since the intervals $[\psi_{k-1}, \psi_{k})$ partition the real line, that is $\reals = \cup_{k\in\ints}[\psi_{k-1}, \psi_{k})$, it follows that for all $\Phi_1 \in \reals$,
+% \[
+% \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} = \lambda\cos\fracpart{\Phi_1} +  v(\lambda, \Phi_1),
+% \]
+% where $v(\lambda, \Phi_1) = \sum_{k\in\ints} v_k(\lambda, \Phi_1)$ and
+% \begin{equation}\label{eq:vk}
+% v_k(\lambda,\Phi_1) = \begin{cases}
+% r_k(\Phi_1) + \lambda\eta_k(\lambda, \Phi_1) &  \Phi_1 \in [ \psi_{k} - \lambda,\psi_{k} ) \\
+% \lambda\zeta_k(\lambda) & \Phi_1 \in [\psi_{k-1}, \psi_{k} - \lambda ) \\
+% 0 & \text{otherwise}.
+% \end{cases} 
+% \end{equation}
+% Now,
+% \begin{align*}
+% k_2(\lambda) &=  \expect R_1\big( \sin\fracpart{\lambda + \Phi_1} - \sin\fracpart{\Phi_1} \big) \\
+% &= \expect R_1\big( \lambda\cos\fracpart{\Phi_1} + v(\lambda, \Phi_1)  \big) \\
+% &= \lambda h_2(0) + \expect R_1 v(\lambda, \Phi_1).
+% \end{align*}
+% Let $W(\lambda) = \expect R_1 v(\lambda, \Phi_1)$ so that $k_2(\lambda) = \lambda h_2(0) + W(\lambda).$  Lemma~\ref{lem:expRvlamphi} shows that $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma(\lambda)$ is a continuous function with $\gamma(0) = 0$, and
+% \[
+% H_1 = - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\psi_k) = H - h_2(0),
+% \]
+% where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
 Because $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow\infty$, it follows that $\gamma(\hat{\lambda}_L) = o_P(1)$, and so
-\[
-k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( h_2(0) + H_1 + o_P(1) ) = \hat{\lambda}_L ( H + o_P(1) )
-\]
-as required.  It remains to prove Lemma~\ref{lem:expRvlamphi}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:expRvlamphi}
-$W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
-\end{lemma}
-\begin{IEEEproof}
-In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
+% \[
+% k_2(\hat{\lambda}_L) = \hat{\lambda}_L ( h_2(0) + H_1 + o_P(1) ) = \hat{\lambda}_L ( H + o_P(1) )
+% \]
+% as required.  It remains to prove Lemma~\ref{lem:expRvlamphi}.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:expRvlamphi}
+% $W(\lambda) =  \lambda(H_1 + \gamma(\lambda))$ where $\gamma$ is a continuous function with $\gamma(0) = 0$.
+% \end{lemma}
+% \begin{IEEEproof}
+% In this lemma, as in Lemma~\ref{lem:k2conv}, we work under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
 An analogous argument follows when $-\tfrac{\pi}{M} \leq \lambda < 0$.
 Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$,
-\begin{align}
-W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
-&= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
-&= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
-&= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi, \label{eq:Winfsum}
-\end{align}
-where $g$ is defined in the statement of Theorem~\ref{thm:normality}.
-Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
-\begin{align*}
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
-\end{align*}
-where, for $k = 1, \dots, M-1$,
-\begin{equation}\label{eq:ak}
-a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
-\end{equation}
-\[
-b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
-\]
-\begin{equation}\label{eq:ck}
-c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} r_k(\phi) g(\phi) d\phi.
-\end{equation}
-Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ (by assumption in Theorem~\ref{thm:normality}), it follows that $c_k(\lambda)$ converges to
-\[
-r_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
-\]
-as $\lambda$ goes to zero from above.
+% \begin{align}
+% W(\lambda) &= \expect R_1 v(\lambda, \Phi_1) \nonumber  \\
+% &= \int_0^{2\pi}\int_{0}^\infty r v(\lambda, \phi) f(r,\phi) dr d\phi \nonumber  \\ 
+% &= \int_0^{2\pi} v(\lambda, \phi) g(\phi) d\phi \nonumber \\
+% &= \sum_{k \in \ints} \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi, \label{eq:Winfsum}
+% \end{align}
+% where $g$ is defined in the statement of Theorem~\ref{thm:normality}.
+% Now, for $k = 1, \dots, M-1$ we have $0 < \psi_{k-1} < \psi_k < 2\pi$, and so, using~\eqref{eq:vk},
+% \begin{align*}
+% \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi &= \int_{\psi_{k-1}}^{\psi_{k}} v_k(\lambda, \phi) g(\phi) d\phi \\
+% &= \lambda a_k(\lambda) + \lambda b_k(\lambda) + \lambda c_k(\lambda),
+% \end{align*}
+% where, for $k = 1, \dots, M-1$,
+% \begin{equation}\label{eq:ak}
+% a_k(\lambda) = \int_{\psi_k-\lambda}^{\psi_k}\eta_k(\lambda, \phi) g(\phi) d\phi,
+% \end{equation}
+% \[
+% b_k(\lambda) = \zeta_k(\lambda) \int_{\psi_{k-1}}^{\psi_k -\lambda}g(\phi) d\phi,
+% \]
+% \begin{equation}\label{eq:ck}
+% c_k(\lambda) = \frac{1}{\lambda} \int_{\psi_{k} - \lambda}^{\psi_{k}} r_k(\phi) g(\phi) d\phi.
+% \end{equation}
+% Both $a_k(\lambda)$ and $b_k(\lambda)$ converge to zero as $\lambda$ goes to zero, and, since $g$ is continuous at $\psi_k = \tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, 1, \dots, M-1$ (by assumption in Theorem~\ref{thm:normality}), it follows that $c_k(\lambda)$ converges to
+% \[
+% r_k(\psi_k)g(\psi_k) = -2\sin(\tfrac{\pi}{M}) g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M})
+% \]
+% as $\lambda$ goes to zero from above.
 We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$.
@@
 
-Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
-\[
-\int_0^{2\pi} v_0(\lambda, \phi) g(\phi) d\phi = \lambda a_0(\lambda) + \lambda b_0(\lambda) +  \lambda c_0(\lambda),
-\]
-where $a_0(\lambda)$ and $c_0(\lambda)$ are defined as in~\eqref{eq:ak}~and~\eqref{eq:ck}, and
-\[
-b_0(\lambda) = \zeta_0(\lambda) \int_{0}^{\psi_k-\lambda} g(\phi) d\phi.
-\]
-When $k = M$, we have $0 < \psi_{M-1} < 2\pi < \psi_{M} - \lambda$, so that
-\[
-\int_0^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi = \lambda b_{M}(\lambda)
-\]
-where 
-\[
-b_{M}(\lambda) = \zeta_{M}(\lambda) \int_{\psi_{M-1}}^{2\pi}g(\phi) d\phi.
-\]
-Both $b_0(\lambda)$ and $b_M(\lambda)$ are functions that converge to zero as $\lambda$ converges to zero.
-
-Finally, when $k < 0$ or $k > M$, 
-\[
-\int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0.
-\]
-So the infinite sum in~\eqref{eq:Winfsum} can be written as the finite sum,
-\begin{align*}
-W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
-&= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
-&= \lambda \big( C(\lambda)  + A(\lambda) + B(\lambda) \big),
-\end{align*}
-where 
-\[
-A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \qquad B(\lambda) = \sum_{k=0}^{M} b_k(\lambda),  
-\]
-\[
-C(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
-\]
-Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda \rightarrow 0$, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.
+% Now, when $k = 0$, we have $\psi_{-1} < 0 < \psi_0 - \lambda < \psi_0 < 2\pi$, so that
+% \[
+% \int_0^{2\pi} v_0(\lambda, \phi) g(\phi) d\phi = \lambda a_0(\lambda) + \lambda b_0(\lambda) +  \lambda c_0(\lambda),
+% \]
+% where $a_0(\lambda)$ and $c_0(\lambda)$ are defined as in~\eqref{eq:ak}~and~\eqref{eq:ck}, and
+% \[
+% b_0(\lambda) = \zeta_0(\lambda) \int_{0}^{\psi_k-\lambda} g(\phi) d\phi.
+% \]
+% When $k = M$, we have $0 < \psi_{M-1} < 2\pi < \psi_{M} - \lambda$, so that
+% \[
+% \int_0^{2\pi} v_{M}(\lambda, \phi) g(\phi) d\phi = \lambda b_{M}(\lambda)
+% \]
+% where 
+% \[
+% b_{M}(\lambda) = \zeta_{M}(\lambda) \int_{\psi_{M-1}}^{2\pi}g(\phi) d\phi.
+% \]
+% Both $b_0(\lambda)$ and $b_M(\lambda)$ are functions that converge to zero as $\lambda$ converges to zero.
+
+% Finally, when $k < 0$ or $k > M$, 
+% \[
+% \int_0^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi = 0.
+% \]
+% So the infinite sum in~\eqref{eq:Winfsum} can be written as the finite sum,
+% \begin{align*}
+% W(\lambda) &= \sum_{k=0}^{M} \int_{0}^{2\pi} v_k(\lambda, \phi) g(\phi) d\phi \\
+% &= \sum_{k=0}^{M-1} \lambda c_k(\lambda) + \sum_{k=0}^{M-1} \lambda a_k(\lambda) + \sum_{k=0}^{M} \lambda b_k(\lambda) \\
+% &= \lambda \big( C(\lambda)  + A(\lambda) + B(\lambda) \big),
+% \end{align*}
+% where 
+% \[
+% A(\lambda) = \sum_{k=0}^{M-1} a_k(\lambda), \qquad B(\lambda) = \sum_{k=0}^{M} b_k(\lambda),  
+% \]
+% \[
+% C(\lambda) = \sum_{k=0}^{M-1} c_k(\lambda).
+% \]
+% Both $A(\lambda)$ and $B(\lambda)$ converge to zero as $\lambda \rightarrow 0$, whilst $C(\lambda)$ converges to $H_1$ as $\lambda$ converges to zero.
 The lemma follows after putting $\gamma(\lambda) = A(\lambda) + B(\lambda) + C(\lambda) - H_1$.
-\end{IEEEproof}
+% \end{IEEEproof}
 
@@
 \[
-\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = o_P(1) + \sqrt{L} R_L(0).
+\sqrt{L}\big( R_L(\hat{\lambda}_L) - Q_L(\hat{\lambda}_L) \big) = \sqrt{L} R_L(0) + o_P(1).
 \]
@@
 \] 
-and therefore $\prob\{ \sabs{\hat{\lambda}_L} \geq \epsilon \} < \nu$ for all sufficiently large $L$.  Now
+and therefore, for any $\nu > 0$, $\prob\{ \sabs{\hat{\lambda}_L} \geq \epsilon \} < \nu$ for all sufficiently large $L$.  Now
 \begin{align*}
@@
 &\leq \prob\left\{  \sup_{\sabs{\lambda} < \epsilon} \sabs{ W_L(\lambda) } > \delta \right\} + \prob\left\{ \sabs{\hat{\lambda}_L} \geq \epsilon \right\} \\
-&\leq 2\nu.
+&\leq 2\nu,
 \end{align*}
@@
 
-\begin{lemma}\label{lem:expectImRfracpart}
-$\expect R_1 \sin\fracpart{\Phi_1} = 0$.
-\end{lemma}
-\begin{IEEEproof}
-Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
-\begin{align*}
-\expect R_1 \sin\fracpart{\Phi_1} &= \int_{0}^{2\pi} \int_{0}^\infty r \sin\fracpart{\phi} f(r,\phi) dr d\phi \\
-&= \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi.
-\end{align*}
-The proof is immediate since $g(\phi)$ is even and $\sin\fracpart{\phi}$ is odd.
-% where $g$ is defined in the statement of Theorem~\ref{thm:normality} and the change in bounds from $0, 2\pi$ to $-\pi,\pi$ is valid because both $\sin\fracpart{\phi}$ and $g(\phi)$ have period $2\pi$.
-Proposition~\ref{prop:gg2cont} shows that $g$ is even on $(-\pi,\pi)$.
-Also $\sin\fracpart{\cdot}$ has period $\tfrac{2\pi}{M}$ and is odd on $(-\frac{\pi}{M}, \tfrac{\pi}{M})$.
-So, for any $k \in \ints$ and $\phi \in (-\tfrac{\pi}{M}, 0)$,
-% \[
-% g(-\phi + k \tfrac{\pi}{M}) = g(\phi - k \tfrac{\pi}{M})
-% \]
-% and
-% \[
-% \sin\fracpart{-\phi + k \tfrac{\pi}{M}} = -\sin\fracpart{\phi - k \tfrac{\pi}{M}}.
-% \]
-% Now,
-% \begin{align*}
-% I_k &= \int_{-(k+1)\tfrac{\pi}{M}}^{-k\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi + \int_{k\tfrac{\pi}{M}}^{(k+1)\tfrac{\pi}{M}}\sin\fracpart{\phi} g(\phi) d\phi \\
-% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi -  k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi \\
-% &\hspace{1cm}+ \int_{0}^{\tfrac{\pi}{M}}\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
-% &= 0
-% \end{align*}
-% for all integers $k$ because
-% \begin{align*}
-% \int_{0}^{\tfrac{\pi}{M}}&\sin\fracpart{\phi + k\tfrac{\pi}{M}} g(\phi + k\tfrac{\pi}{M}) d\phi \\
-% &= \int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{-\phi + k\tfrac{\pi}{M}} g(-\phi + k\tfrac{\pi}{M}) d\phi \\
-% &= -\int_{-\tfrac{\pi}{M}}^{0}\sin\fracpart{\phi - k\tfrac{\pi}{M}} g(\phi - k\tfrac{\pi}{M}) d\phi.
-% \end{align*}
-% Observe that the sets
-% \[
-%  [-(k+1)\tfrac{\pi}{M},-k\tfrac{\pi}{M}] \cup [k\tfrac{\pi}{M},(k+1)\tfrac{\pi}{M}], \;\;\; k = 0, \dots, M-1
-% \]
-% partition $[-\pi, \pi]$, and so
-% \[
-% \expect R_1 \sin\fracpart{\Phi_1} = \int_{-\pi}^{\pi} \sin\fracpart{\phi} g(\phi) d\phi = \sum_{k=0}^{M-1} I_k = 0.
-% \]
-\end{IEEEproof}
-
 \begin{lemma}\label{lem:empiricprocforrho} Let $T_L(\lambda) = \expect G_L(\lambda)$.  We have
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) = o_P(1) + X_L,
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L) \big) =  X_L + o_P(1),
 \]
@@
 where
-\[
-q_1(\lambda) = \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big),
-\]
-\begin{equation}
-q_2(\lambda) = \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
-\end{equation}
+\begin{align}
+q_1(\lambda) &= \expect R_1\big( \cos(\lambda + \Phi_1) - \cos(\Phi_1) \big) \;\;\; \text{and} \nonumber \\
+q_2(\lambda) &= \expect R_1\big( \cos\fracpart{\lambda + \Phi_1} - \cos\fracpart{\Phi_1} \big). \label{eq:q2def}
+\end{align}
 We analyse $q_1(\lambda)$ first.  By a Taylor expansion of $\cos(\lambda + \Phi_1)$ about $\lambda = 0$,
@@
 \begin{IEEEproof}
-Observe that $X_L = C_L + D_L$ where
-\[
-C_L = \frac{1}{\sqrt{L}} \sum_{i \in P} \big( R_i \cos(\Phi_i) - h_1(0) \big),
-\]
-\[
-D_L = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ),
+Observe that $X_L = C_L^\prime + D_L^\prime$ where
+\[
+C_L^\prime = \frac{1}{\sqrt{L}} \sum_{i \in P} \big( R_i \cos(\Phi_i) - h_1(0) \big),
+\]
+\[
+D_L^\prime = \frac{1}{\sqrt{L}} \sum_{i \in D} (R_i \cos\fracpart{\Phi_i} - h_2(0) ),
 \]
 where $h_1$ and $h_2$ are defined in the statement of Theorem~\ref{thm:consistency}.
-From the standard central limit theorem the distribution of $C_L$ converges to the normal with zero mean and variance
+From the standard central limit theorem the distribution of $C_L^\prime$ converges to the normal with zero mean and variance
 \[
@@
 \]
-since $h_1(0) = 1$.  Similarly the distribution of $D_L$ converges to the normal with zero mean and variance
+since $h_1(0) = 1$.  Similarly the distribution of $D_L^\prime$ converges to the normal with zero mean and variance
 \[
@@
 \]
-The lemma follows since $C_L$ and $D_L$ are independent.
+The lemma follows since $C_L^\prime$ and $D_L^\prime$ are independent.
 \end{IEEEproof}

### 2012-12-29 paper.tex
more minor edits

@@
 %\]
-The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $SS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.  Equivalently
-\[
-\hat{\theta} = \arg\max_{\theta \in [-\pi,\pi)} \Re(Z(\theta))
-\]
-with no constraints required.
+The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $SS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
+Equivalently $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ with no constraints required.
 
@@
 \[
-q_1(\lambda) + j k_1(\lambda) = \expect\big( R_1 e^{j(\lambda + \Phi_1)} - R_1 e^{j\Phi_1}  \big).
-\]
-It follows that $q_1(\lambda) = o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$
+q_1(\lambda) + j k_1(\lambda) = \expect\big[ R_1 e^{j(\lambda + \Phi_1)} - R_1 e^{j\Phi_1}  \big].
+\]
+We have $q_1(\lambda) = \hat{\lambda}_L o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$.
 \end{lemma}
@@
 \]
-since the mean of $R_1e^{j\Phi_1}$ is 1.
-Thus $q_1(\lambda) = \cos(\lambda) - 1$ and $k_1(\lambda) = \sin(\lambda)$.
-Now $q_1(\hat{\lambda}_L) = o_P(1)$ since $\hat{\lambda}_L$ converges almost surely to zero as $L\rightarrow\infty$.
-By a Taylor expansion of $\sin(\lambda)$ about $\lambda = 0$ we obtain $k_1(\lambda) = \lambda \big( 1 + \zeta(\lambda) \big)$  where $\zeta(\lambda)$ is a continuous function with $\zeta(0) = 0$.
-Thus $\zeta(\hat{\lambda}_L) = o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$.
+since the mean of $R_1e^{j\Phi_1}$ is 1.  By a Taylor expansion about $\lambda = 0$ we obtain
+\[
+q_1(\lambda) + j k_1(\lambda) = \lambda\big( j + \zeta(\lambda) \big)
+\]
+where $\zeta(\lambda)$ is a continuous function with $\zeta(0) = 0$.
+Since $\hat{\lambda}_L$ converges almost surely to zero as $L\rightarrow\infty$ it follows that $\zeta(\hat{\lambda}_L) = o_P(1)$.
+Thus
+\[
+q_1(\hat{\lambda}_L) + j k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big( j + o_P(1) \big)
+\]
+and the lemma follow by taking real and imaginary parts.
 \end{proof}
@@
 \[
-q_2(\lambda) + j k_2(\lambda) = \expect\big( R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big).
-\]
-It follows that $q_2(\hat{\lambda}_L) = o_P(1)$ and $k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H + o_P(1) \big)$ where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
+q_2(\lambda) + j k_2(\lambda) = \expect\big[ R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big].
+\]
+We have $q_2(\hat{\lambda}_L) = \hat{\lambda}_L o_P(1)$ and $k_2(\hat{\lambda}_L) = \hat{\lambda}_L \big( H + o_P(1) \big)$ where $H$ is defined in the statement of Theorem~\ref{thm:normality}.
 \end{lemma}
@@
 \begin{align*}
-q_2(\lambda) + j k_2(\lambda) &= \big(e^{j\lambda} - 1 \big) h_2(0) \\
-&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big)
-\end{align*}
-since $\expect R_1 e^{j\fracpart{\Phi_1}} = h_2(0)$ as a result of Lemma~\ref{lem:expectImRfracpart} and where $\chi(\Phi_1, \lambda) = \sum_{k \in \ints}\chi_k(\Phi_1, \lambda)$.
+q_2(\lambda) + j k_2(\lambda) &= \expect\big[ R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big] \\
+&= \big(e^{j\lambda} - 1 \big) h_2(0) \\
+&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) \expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda) 
+\end{align*}
+since $\expect R_1 e^{j\fracpart{\Phi_1}} = h_2(0) + \expect R_1 \sin\fracpart{\Phi_1} = h_2(0)$ as a result of Lemma~\ref{lem:expectImRfracpart} and where $\chi(\Phi_1, \lambda) = \sum_{k \in \ints}\chi_k(\Phi_1, \lambda)$.
 Now
 \begin{align*}
- \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big) &= \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) \chi(\phi,\lambda) dr d\phi \\
+ \expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)  &= \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) \chi(\phi,\lambda) dr d\phi \\
 &= \sum_{k\in\ints}\int_{0}^{2\pi} g(\phi) \chi_k(\phi,\lambda) d\phi \\
 &= \sum_{k=0}^{M-1}\int_{\psi_k-\lambda}^{\psi_k} g(\phi)  d\phi \\
-&= \lambda \sum_{k=1}^{M-1} g(\phi_k) 
-\end{align*}
-where for each $k$, $\phi_k \in (\psi_k-\lambda, \psi_k)$.  The third line above follows from
+&= \lambda \sum_{k=0}^{M-1} g\big( \phi_k(\lambda) \big) 
+\end{align*}
+where $\phi_k(\lambda) \in (\psi_k - \lambda, \psi_k)$ for each $k$.
+The third line in the previous equation follows because the $\chi_k(\phi,\lambda)$ terms inside the integral are zero for all $\phi \in [0,2\pi]$ when $k \notin \{0,\dots,M-1\}$.
+We now have
+\begin{equation}\label{eq:q2k2form}
+q_2(\lambda) + j k_2(\lambda) = \big(e^{j\lambda} - 1 \big) h_2(0) + B(\lambda) \sum_{k=0}^{M-1} g\big(\phi_k(\lambda)\big) 
+\end{equation}
+where $B(\lambda) = \lambda e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)$.
+Since $\phi_k(\lambda) \rightarrow \psi_k$ as $\lambda \rightarrow 0$ and $g(\psi_k)$ is continuous at $\psi_k$ (by assumption in Theorem~\ref{thm:normality}) we have
+\[
+ \sum_{k=0}^{M-1} g\big(\phi_k(\hat{\lambda}_L)\big) =  \sum_{k=0}^{M-1} g(\psi_k) + o_P(1).
+\]
+Also $e^{j\hat{\lambda}_L} - 1 = \hat{\lambda}_L \big( j  + o_P(1) \big)$ by the argument in Lemma~\ref{lem:q1k1parts}.  Finally
+\[
+B(\hat{\lambda}_L) = \hat{\lambda}_L \big( o_P(1) - 2 j \sin(\tfrac{\pi}{M})\big)
+\]
+by Lemma~\ref{lem:Blambdaconv}.  Combining these results into~\eqref{eq:q2k2form} and taking real and imaginary parts we obtain $q_2(\hat{\lambda}_L) = \hat{\lambda}_L o_P(1)$ and 
+\begin{align*}
+q_2(\lambda) + &jk_2(\hat{\lambda}_L) \\
+&= \hat{\lambda}_L \left( jh_2(0) - 2j\sin(\tfrac{\pi}{M}) \sum_{k=0}^{M-1} g(\psi_k) + o_P(1) \right) \\
+&= \hat{\lambda}_L\big( jH + o_P(1) \big)
+\end{align*}
+and the lemma follows by taking real and imaginary parts.
+\end{proof}
+
+\begin{lemma}\label{lem:Blambdaconv}
+$B(\hat{\lambda}_L) = \hat{\lambda}_L \big( o_P(1) - 2 j \sin(\tfrac{\pi}{M})\big)$.
+\end{lemma}
+\begin{proof}
+A Taylor expansion about $\lambda = 0$ gives
+\[
+B(\lambda) = \big(  \big)\big( e^{-j\tfrac{2\pi}{M}} - 1 \big)
+\]
 \end{proof}

### 2012-12-29 edits/barryedits.tex
more minor edits
more edits

@@
 However the term involving $\cos\fracpart{\phi}$ isn't always differentiable so the mean value theorem doesn't always apply.
-It's obviouly Lipschitz continuous though, just plot it.
+It's obviously Lipschitz continuous though, just plot it.
 We can add a proof if you really think it's necessary.
-\item Page 12, column 1, ``for $i \in D$ such that $R_i > 0$.'' I think this is necesary, otherwise the surrounding statement is not always true.
+\item Page 12, column 1, ``for $i \in D$ such that $R_i > 0$.'' I think this is necessary, otherwise the surrounding statement is not always true.
 \item Lemma 11, I would also like there to be a nicer proof!  I haven't found one though.
-\item 
+\item You wrote
+\[
+R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda) = \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) \chi(\phi,\lambda) dr d\phi
+\]
+in your notes.  I fixed this to
+\[
+R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda) = \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) e^{j\fracpart{\phi}} \chi(\phi,\lambda) dr d\phi.
+\]
+Everything then works as expected.
 \end{enumerate}

### 2012-12-29 paper.tex
more edits

--- 2012-12-28 paper.tex
@@
 \]
-since the mean of $R_1e^{j\Phi_1}$ is 1.
-Thus $q_1(\lambda) = \cos(\lambda) - 1$ and $k_1(\lambda) = \sin(\lambda)$.
-Now $q_1(\hat{\lambda}_L) = o_P(1)$ since $\hat{\lambda}_L$ converges almost surely to zero as $L\rightarrow\infty$.
-By a Taylor expansion of $\sin(\lambda)$ about $\lambda = 0$ we obtain $k_1(\lambda) = \lambda \big( 1 + \zeta(\lambda) \big)$  where $\zeta(\lambda)$ is a continuous function with $\zeta(0) = 0$.
-Thus $\zeta(\hat{\lambda}_L) = o_P(1)$ and $k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big(1 + o_P(1) \big)$.
+since the mean of $R_1e^{j\Phi_1}$ is 1.  By a first order expansion about $\lambda = 0$ we obtain
+\[
+q_1(\lambda) + j k_1(\lambda) = \lambda\big( j + O(\lambda) \big)
+\]
+Since $\hat{\lambda}_L$ converges almost surely to zero as $L\rightarrow\infty$ it follows that $O(\hat{\lambda}_L) = o_P(1)$.  Thus
+\[
+q_1(\hat{\lambda}_L) + j k_1(\hat{\lambda}_L) = \hat{\lambda}_L\big( j + o_P(1) \big)
+\]
+and the lemma follow by taking real and imaginary parts.
 \end{proof}
@@
 Now
-\begin{align*}
-q_2(\lambda) + j k_2(\lambda) &= \big(e^{j\lambda} - 1 \big) h_2(0) \\
-&\hspace{0.5cm} + e^{j\lambda}\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big)
-\end{align*}
-since $\expect R_1 e^{j\fracpart{\Phi_1}} = h_2(0)$ as a result of Lemma~\ref{lem:expectImRfracpart} and where $\chi(\Phi_1, \lambda) = \sum_{k \in \ints}\chi_k(\Phi_1, \lambda)$.
-Now
-\begin{align*}
- \expect\big(R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)\big) &= \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) \chi(\phi,\lambda) dr d\phi \\
-&= \sum_{k\in\ints}\int_{0}^{2\pi} g(\phi) \chi_k(\phi,\lambda) d\phi \\
-&= \sum_{k=0}^{M-1}\int_{\psi_k-\lambda}^{\psi_k} g(\phi)  d\phi \\
-&= \lambda \sum_{k=1}^{M-1} g(\phi_k) 
-\end{align*}
-where for each $k$, $\phi_k \in (\psi_k-\lambda, \psi_k)$.  The third line above follows from
+\begin{align}
+q_2(\lambda) + j k_2(\lambda) &= \expect\big[ R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big] \nonumber \\
+&= \big(e^{j\lambda} - 1 \big) h_2(0)  + e^{j\lambda} B(\lambda) \label{eq:q2k2withB}
+\end{align}
+since $\expect R_1 e^{j\fracpart{\Phi_1}} = h_2(0) + \expect R_1 \sin\fracpart{\Phi_1} = h_2(0)$ as a result of Lemma~\ref{lem:expectImRfracpart} and where
+\begin{equation}\label{eq:Blambdadefn}
+B(\lambda) = \big( e^{-j\tfrac{2\pi}{M}} - 1 \big)\expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)
+\end{equation}
+and $\chi(\Phi_1, \lambda) = \sum_{k \in \ints}\chi_k(\Phi_1, \lambda)$.
+
+Now $e^{j\hat{\lambda}_L} = 1 + o_P(1)$ and $e^{j\hat{\lambda}_L} - 1 = \hat{\lambda}_L \big( j  + o_P(1) \big)$ by the argument in Lemma~\ref{lem:q1k1parts} and
+\[
+B(\hat{\lambda}_L) = -\hat{\lambda}_L\left(2j\sin(\tfrac{\pi}{M}) \sum_{k=0}^{M-1} g(\psi_k) + o_P(1)\right) 
+\]
+by Lemma~\ref{lem:Blambdaconv}.  Combining these results into~\eqref{eq:q2k2withB} we obtain
+\begin{align*}
+q_2(\lambda) + &jk_2(\hat{\lambda}_L) \\
+&= \hat{\lambda}_L \left( jh_2(0) - 2j\sin(\tfrac{\pi}{M}) \sum_{k=0}^{M-1} g(\psi_k) + o_P(1) \right) \\
+&= \hat{\lambda}_L\big( jH + o_P(1) \big)
+\end{align*}
+and the lemma follows by taking real and imaginary parts.
+\end{proof}
+
+\begin{lemma}\label{lem:Blambdaconv}
+With $B(\lambda)$ defined in~\eqref{eq:Blambdadefn} we have 
+\[
+B(\hat{\lambda}_L) = -\hat{\lambda}_L\left(2j\sin(\tfrac{\pi}{M}) \sum_{k=0}^{M-1} g(\psi_k) + o_P(1)\right) 
+\]
+\end{lemma}
+\begin{proof}
+Put $A(\lambda) = \expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)$.  Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$ we have
+\begin{align*}
+ A(\lambda)  &= \int_{0}^{2\pi}\int_{0}^{\infty} r f(r,\phi) e^{j\fracpart{\phi}}\chi(\phi,\lambda) dr d\phi \\
+&= \sum_{k\in\ints}\int_{0}^{2\pi} g(\phi) e^{j\fracpart{\phi}} \chi_k(\phi,\lambda) d\phi \\
+&= \sum_{k=0}^{M-1}\int_{\psi_k-\lambda}^{\psi_k} g(\phi)  e^{j\fracpart{\phi}} d\phi,
+\end{align*}
+the last line because the $\chi_k(\phi,\lambda)$ terms inside the integral are zero for all $\phi \in [0,2\pi]$ when $k \notin \{0,\dots,M-1\}$.
+Observe that $\fracpart{\phi} \rightarrow \tfrac{\pi}{M}$ as $\phi$ approaches $\psi_k$ from below.
+Because $g(\psi_k)$ is continuous at $\psi_k$ for each $k = 0, \dots, M-1$ (by assumption in Theorem~\ref{thm:normality}) we have
+\[
+\frac{1}{\lambda} A(\lambda) \rightarrow \sum_{k=0}^{M-1} g(\psi_k) e^{j\tfrac{\pi}{M}}
+\]
+as $\lambda$ approaches zero from above.
+We are only interested in the limit from above because we are working under the assumption that $0 \leq \lambda < \frac{\pi}{M}$ (see the proof of Lemma~\ref{lem:q2k2parts}).
+The analogous argument when $-\tfrac{\pi}{M} \leq \lambda < 0$ would involve limits as $\lambda$ approaches zero from below.
+Thus
+\[
+A(\hat{\lambda}_L) = \hat{\lambda}_L \left(  e^{j\tfrac{\pi}{M}} \sum_{k=0}^{M-1} g(\psi_k) + o_P(1) \right)
+\]
+and the lemma follows since $\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) e^{j\frac{\pi}{M}} = -2j\sin(\tfrac{\pi}{M})$ and $B(\lambda) = \big( e^{-j\tfrac{2\pi}{M}} - 1 \big)A(\lambda)$.
 \end{proof}
@@
 \begin{IEEEproof}
-Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$, we have
+Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$ we have
 \begin{align*}
@@
 \end{align}
-We analyse $q_1(\lambda)$ first.  By a Taylor expansion of $\cos(\lambda + \Phi_1)$ about $\lambda = 0$,
-\[
-\cos(\lambda + \Phi_1) - \cos(\Phi_1) = -\lambda \big( \sin(\Phi_1) + \zeta(\lambda) \big),
-\]
-where $\zeta$ is a continuous function with $\zeta(0) = 0$. So,
-\begin{align*}
-q_1(\lambda) &= -\expect R_1\lambda(\sin(\Phi_1) +  \zeta(\lambda)) \\
-&= -\lambda \big( \expect R_1\sin(\Phi_1) +  \zeta(\lambda) \expect R_1 \big) \\
-&= -\lambda \zeta(\lambda) \expect R_1,
-\end{align*}
-since $\expect R_1\sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}.  Now, 
-\begin{align*}
-\sqrt{L} q_1(\hat{\lambda}_L) = -\sqrt{L} \hat{\lambda}_L \zeta(\hat{\lambda}_L) \expect R_1 = o_P(1)
-\end{align*}
-because $\sqrt{L} \hat{\lambda}_L$ converges in distribution and $\zeta(\hat{\lambda}_L) = o_P(1)$.
-An argument analogous to that used in Lemma~\ref{lem:k2conv} shows that $\sqrt{L} q_2(\hat{\lambda}_L) = o_P(1)$.
-Thus,
-\[
-\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1) = o_P(1)
-\]
-as required.
+Lemma~\ref{lem:q1k1parts} shows that $q_1(\hat{\lambda}_L) = \hat{\lambda}_Lo_P(1)$ and Lemma~\ref{lem:q2k2parts} shows that $q_2(\hat{\lambda}_L) = \hat{\lambda}_L o_P(1)$ and so
+\[
+\sqrt{L}\big( T_L(\lambda) - G(0) \big) =\sqrt{L}\hat{\lambda}_L o_P(1) + o(1).
+\]
+The lemma follows since $\sqrt{L}\hat{\lambda}_L$ converges in distribution.
 \end{IEEEproof}

### 2012-12-29 paper.tex
added stuff

@@
 The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
-The solid lines are the MSEs predicted by Theorem~\ref{thm:normality}.
+The solid lines are the MSEs predicted by Theorem~\ref{thm:normality} (and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).
 The theorem accurately predicts the behaviour of the phase estimator when the SNR is sufficiently large.
@@
 However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
-As indicated in Figures~\ref{fig:plotampBPSK}~and~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.
+As indicated in Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.
 However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.
@@
 \begin{align*}
-\sqrt{L}\big( T_L(\lambda) - G(0) \big) = p q_1(\lambda) + d q_2(\lambda) + o(1),
+\sqrt{L}\big( T_L(\lambda) - G(0) \big) = \sqrt{L}p q_1(\lambda) + \sqrt{L} d q_2(\lambda) + o(1),
 \end{align*}

### 2013-01-07 paper.tex
added some of Vaughan's comments

@@
 \begin{IEEEkeywords}
-Coherent detection, noncoherent detection, phase shift keying, asymptotic statistics
+Detection, phase shift keying, asymptotic statistics
 \end{IEEEkeywords}
@@
 \end{lemma}
-\begin{proof}
+\begin{IEEEproof}
 We have
@@
 and the lemma follow by taking real and imaginary parts.
-\end{proof}
+\end{IEEEproof}
 
@@
 \end{lemma}
-\begin{proof}
+\begin{IEEEproof}
 Because $\hat{\lambda}_L \rightarrow 0$ almost surely as $L\rightarrow\infty$, it is only the behaviour of $q_2(\lambda)$ and $k_2(\lambda)$ around $\lambda = 0$ that is relevant.
@@
 and the lemma follows by taking real and imaginary parts.
-\end{proof}
+\end{IEEEproof}
 
@@
 \end{lemma}
-\begin{proof}
+\begin{IEEEproof}
 Put $A(\lambda) = \expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)$.  Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$ we have
@@
 and the lemma follows since $\big( e^{-j\tfrac{2\pi}{M}} - 1 \big) e^{j\frac{\pi}{M}} = -2j\sin(\tfrac{\pi}{M})$ and $B(\lambda) = \big( e^{-j\tfrac{2\pi}{M}} - 1 \big)A(\lambda)$.
-\end{proof}
+\end{IEEEproof}
 

### 2013-01-07 paper.tex
minor

--- 2012-12-29 paper.tex
@@
 \]
-The next two theorems describe the asymptotic properties of the least squares estimator.
+The next two theorems describe the asymptotic properties of the least squares estimator.  These are the central results and the chief original contributions of this paper.
 
@@
 Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\hat{\theta} - \theta_0}_\pi$ will not converge.
-One can check that the assumption holds when $w_1$ is circularly symmetric and normally distributed.
+One can check that this assumption holds when $w_1$ is circularly symmetric and normally distributed.
 We will not attempt to further classify those distributions for which the assumption holds here.
@@
 \end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$.  Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
+where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$ and where, as in Section~\ref{sec:least-squar-estim}, we consider $\hat{u}_i(\theta)$ equivalent modulo $2\pi$.
+Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
 \begin{align}
@@
 However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
-Similar properties have been used to analyse the behaviour of other estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.%\cite{McKilliam_LSU_polyest_part1_2012,McKilliam_LSU_polyest_part2_2012}.
+Similar properties have been used to analyse the behaviour of other estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.
+%\cite{McKilliam_LSU_polyest_part1_2012,McKilliam_LSU_polyest_part2_2012}.
 Define the function

### 2013-01-07 paper.tex
added more of Vaughans edits

@@
 \[
-Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}},
-\]
-and
-\[
-\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
-\] 
+Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}.
+\]
+%and
+%\[
+%\Re(Z(\theta)) = \rho_0 \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \rho_0 \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i}.
+%\] 
 Let 
@@
 However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
-Similar properties have been used to analyse the behaviour of other estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.
+Similar properties have been used by some of the present authors to analyse the behaviour of polynomial-phase estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.
 %\cite{McKilliam_LSU_polyest_part1_2012,McKilliam_LSU_polyest_part2_2012}.
@@
 \]
-where $o_P(1)$ denotes a number converging in probability to zero as $L \rightarrow \infty$, and $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+where $o_P(1)$ denotes a sequence of random variables converging in probability to zero as $L \rightarrow \infty$, and $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
 Lemma~\ref{lem:empiricprocc} shows that
@@
 We present the results of Monte-Carlo simulations with the least squares estimator.
-In all simulations the noise $w_1,\dots,w_L$ is independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+In all simulations the noise samples $w_1,\dots,w_L$ are independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
 Under these conditions the least squares estimator is also the maximum likelihood estimator.
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{2}, \tfrac{L}{8}, \tfrac{L}{32}, L$.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.
 When $\abs{P} \neq 0$ (i.e.

### 2013-01-07 paper.tex
added more of Vaughan's edits

@@
 \thanks{
-Supported under the Australian Government’s Australian Space Research Program.
+A preliminary version of this paper has been submitted to ICASSP'13~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Supported under the Australian Government’s Australian Space Research Program.
 Robby McKilliam, Andr\'{e} Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
@@
 The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
-The solid lines are the MSEs predicted by Theorem~\ref{thm:normality} (and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).
-The theorem accurately predicts the behaviour of the phase estimator when the SNR is sufficiently large.
+The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality} (and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).
+The prediction is made by dividing the asymptotic covariance matrix in Theorem~\ref{thm:normality} by $L$.
+The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large.
 As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-Theorem~\ref{thm:normality} does not model this behaviour.
+Theorem~\ref{thm:normality} does not model this behaviour in the sense that for any fixed $L$ there exists a sufficiently small value of $SNR$ such that Theorem~\ref{thm:normality} does not produce accurate predictions.
 
@@
 
-Now $e^{j\hat{\lambda}_L} = 1 + o_P(1)$ and $e^{j\hat{\lambda}_L} - 1 = \hat{\lambda}_L \big( j  + o_P(1) \big)$ by the argument in Lemma~\ref{lem:q1k1parts} and
+Now $e^{j\hat{\lambda}_L} = 1 + o_P(1)$ and $e^{j\hat{\lambda}_L} - 1 = \hat{\lambda}_L \big( j  + o_P(1) \big)$ by the argument in Lemma~\ref{lem:q1k1parts}.  Also
 \[

### 2013-01-08 paper.tex
more of Vaughan's edits

@@
 A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
-\[
+\begin{equation}\label{eq:viterbiviterbi}
 A = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
-\]
+\end{equation}
 Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
@@
 Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
+We show, under some reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
@@
 g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
+%g(\phi) = \frac{\cos\phi}{2\pi}e^{-\kappa} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\kappa\sin^2\phi} \big(1 + 2\kappa\cos^2\phi \big)
 \]
@@
 %and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  
-The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using these formula.
+The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using this formula.
 
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.
 When $\abs{P} \neq 0$ (i.e.
@@
 The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality} (and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).
-The prediction is made by dividing the asymptotic covariance matrix in Theorem~\ref{thm:normality} by $L$.
+The prediction is made by dividing the asymptotic covariance matrix by $L$.
 The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large.
 As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-Theorem~\ref{thm:normality} does not model this behaviour in the sense that for any fixed $L$ there exists a sufficiently small value of $SNR$ such that Theorem~\ref{thm:normality} does not produce accurate predictions.
+Theorem~\ref{thm:normality} does not model this behaviour in the sense that for any fixed $L$ there exist sufficiently small values of SNR for which Theorem~\ref{thm:normality} does not produce accurate predictions of the MSE.
+
+The figures display some interesing behaviour.  When the SNR is small \dots.  When the SNR is large \dots.
+
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the noncoherent phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
+This estimator requires selection of a function $F$ that transforms the amplitude of each sample prior to the final estimation step.
+% (see equation (2) in~\cite{ViterbiViterbi_phase_est_1983}).
+Viterbi and Viterbi propose several viable alternatives, from which we have chosen $F(x) = 1$.
+The Viterbi and Viterbi estimator is only applicable in the noncoherent setting, i.e.
+when $\abs{P} = 0$.
+The sample MSE of the least squares estimator (when $\abs{P} = 0$) and the Viterbi and Viterbi estimator is similar.
+The least squares estimator appears slightly more accurate for some values of SNR.
 
@@
 
+\section{When the SNR is small}
+
+In this section we study the behaviour of the least squares estimator when SNR goes to zero.
+We consider only the case when the noise samples $w_1,\dots,w_L$ are independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
+To simplify our notation denote the signal to noise ratio by $\nu = \frac{\rho_0^2}{2\sigma^2}$.
+In this case the function $g(\phi)$ takes to form
+\[
+g(\phi) = \frac{1}{2\sqrt{\pi\nu}} + \frac{\cos(\phi)}{2\pi} + o(1)
+\]
+as $\nu \rightarrow 0$.  Now
+\[
+A_1 = \int_{-\pi}^{\pi} g(\phi) \sin^2(\phi) d\phi = \frac{\pi}{2\sqrt{\pi\nu}} + O(1),
+\]
+\begin{align*}
+A_2 &= \int_{-\pi}^{\pi} g(\phi) \sin^2\fracpart{\phi} d\phi \\
+&= \frac{M}{2\sqrt{\pi\nu}} \int_{-\pi/M}^{\pi/M}\sin^2(\phi) d\phi + O(1) \\
+&= \frac{\pi - M\sin(\tfrac{\pi}{M})}{2\sqrt{\pi\nu}},
+\end{align*}
+\begin{align*}
+h_2(0) &= \int_{-\pi}^{\pi} g(\phi) \cos\fracpart{\phi} d\phi \\
+&= \frac{M}{2\sqrt{\pi\nu}} \int_{-\pi/M}^{\pi/M}\cos(\phi) d\phi + O(1) \\
+&= \frac{M\sin(\tfrac{\pi}{M})}{\sqrt{\pi\nu}}
+\end{align*}
+and so
+\begin{align*}
+H &= h_2(0) - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}) \\
+&= h_2(0) -  \frac{M\sin(\pi/M)}{\sqrt{\pi\nu}}\\
+&= O(1)
+\end{align*}
+Ok, you need to expand to one more term.
+\section{When the SNR is large}
+
 \section{Conclusion}
@@
 
-We showed, under some conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+We showed, under some reasonably general conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
 However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
@@
 Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
-This suggests the use of a weighed objective function, which would be an interesting topic for future research.
+This suggests the use of a weighted objective function, which would be an interesting topic for future research.
 

### 2013-01-09 paper.tex
Added new stuff

@@
 \[
+%\begin{equation}\label{eq:gfornormal}
 %g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
-%g(\phi) = \frac{\cos\phi}{2\pi}e^{-\kappa} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\kappa\sin^2\phi} \big(1 + 2\kappa\cos^2\phi \big)
-\]
-where $\kappa = \rho_0/\sigma$ and $\Psi(x) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{x} e^{-x^2/2} dx$
+%g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
+g(\phi) = \frac{\cos\phi}{2\pi}e^{-\kappa} + \frac{\Psi(\sqrt{2\kappa} \cos\phi)}{\sqrt{\pi\kappa}}  e^{-\kappa\sin^2\phi} \big(2 + \kappa\cos^2\phi \big)
+%\end{equation}
+\]
+where $\kappa = \tfrac{\rho_0^2}{2\sigma^2}$ and $\Psi(t) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{t} e^{-x^2/2} dx$
 is the cumulative density function of the standard normal.
@@
 Under these conditions the least squares estimator is also the maximum likelihood estimator.
-Simulations are run with $M=2,4,8$ (BPSK and QPSK and $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.
+Simulations are run with $M=2,4,8$ (BPSK, QPSK, $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.
 The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
@@
 The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
-The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality} (and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).
+The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality}.
+%(and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).
 The prediction is made by dividing the asymptotic covariance matrix by $L$.
@@
 Theorem~\ref{thm:normality} does not model this behaviour in the sense that for any fixed $L$ there exist sufficiently small values of SNR for which Theorem~\ref{thm:normality} does not produce accurate predictions of the MSE.
-
-The figures display some interesing behaviour.  When the SNR is small \dots.  When the SNR is large \dots.
+As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e.
+$\abs{P} = L$.
+
+% We can use our asymptotic theorem to predict the SNR at which the least squares estimator will be close to that of the unmodulated carrier estimate.
+The asymptotic variance on the unmodulated carrier estimator is given by setting $p = 1$ and $d=0$ in Theorem~\ref{thm:normality}.
+We obtain the variance to be $A_1$.
+Thus, if we desire the least squares estimator to have variance close to $A_1$, say $(1 + \delta)A_1$ for some small $\delta > 0$, then we require SNR at which
+% \[
+% \frac{pA_1 + dA_2}{(p + dH)^2} = (1 + \delta)A_1
+% \]
+% where $\delta$ is a positive
+%The figures display some interesing behaviour.  When the SNR is small \dots.  When the SNR is large \dots.
 
@@
 However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
-As indicated in Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the unmodulated carrier.
+%As indicated in Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the estimator when all symbols are pilots.
 However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.
@@
 
-\section{When the SNR is small}
-
-In this section we study the behaviour of the least squares estimator when SNR goes to zero.
+% \section{When the SNR is small}
+
+% In this section we study the behaviour of the least squares estimator when SNR goes to zero.
 We consider only the case when the noise samples $w_1,\dots,w_L$ are independent and identically distributed circularly symmetric and Gaussian with real and imaginary parts having variance $\sigma^2$.
-To simplify our notation denote the signal to noise ratio by $\nu = \frac{\rho_0^2}{2\sigma^2}$.
-In this case the function $g(\phi)$ takes to form
-\[
-g(\phi) = \frac{1}{2\sqrt{\pi\nu}} + \frac{\cos(\phi)}{2\pi} + o(1)
-\]
-as $\nu \rightarrow 0$.  Now
-\[
-A_1 = \int_{-\pi}^{\pi} g(\phi) \sin^2(\phi) d\phi = \frac{\pi}{2\sqrt{\pi\nu}} + O(1),
-\]
-\begin{align*}
-A_2 &= \int_{-\pi}^{\pi} g(\phi) \sin^2\fracpart{\phi} d\phi \\
-&= \frac{M}{2\sqrt{\pi\nu}} \int_{-\pi/M}^{\pi/M}\sin^2(\phi) d\phi + O(1) \\
-&= \frac{\pi - M\sin(\tfrac{\pi}{M})}{2\sqrt{\pi\nu}},
-\end{align*}
-\begin{align*}
-h_2(0) &= \int_{-\pi}^{\pi} g(\phi) \cos\fracpart{\phi} d\phi \\
-&= \frac{M}{2\sqrt{\pi\nu}} \int_{-\pi/M}^{\pi/M}\cos(\phi) d\phi + O(1) \\
-&= \frac{M\sin(\tfrac{\pi}{M})}{\sqrt{\pi\nu}}
-\end{align*}
-and so
-\begin{align*}
-H &= h_2(0) - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}) \\
-&= h_2(0) -  \frac{M\sin(\pi/M)}{\sqrt{\pi\nu}}\\
-&= O(1)
-\end{align*}
-Ok, you need to expand to one more term.
-\section{When the SNR is large}
+In this case the function $g(\theta)$ is given by~\eqref{eq:gfornormal} where $\kappa = \frac{\rho_0^2}{2\sigma^2}$ is the SNR.
+Observe the following expansions
+% \begin{align*}
+% \frac{\cos\phi}{2\pi}e^{-\kappa} &= \sum_{n=0}^\infty a_n \kappa^n\\
+% e^{-\kappa\sin^2\phi} &= \sum_{n=0}^\infty  b_n \kappa^n \\
+% \frac{\Psi(\sqrt{2\kappa} \cos\phi)}{\sqrt{\pi\kappa}}  &= \frac{1}{2\sqrt{\pi\kappa}} + \sum_{n=0}^\infty c_n \kappa^n
+% \end{align*}
+% where
+% \begin{align*}
+% a_n &= (-1)^n \frac{\cos\phi}{ 2 \pi n!} \\
+% b_n &= (-1)^n \frac{ \sin^{2n}(\phi)}{n!} \\
+% c_n &= (-1)^n \frac{\cos^{2n+1}(\phi)}{n! (2n+1)\pi }
+% \end{align*}
+% Thus
+% \[
+% \frac{\Psi(\sqrt{2\kappa} \cos\phi)}{\sqrt{\pi\kappa}}e^{-\kappa\sin^2\phi} = \frac{1}{2\sqrt{\pi\kappa}} \sum_{n=0}^\infty b_n \kappa^{n}  + \sum_{n=0}^\infty d_n \kappa^n,
+% \]
+% where
+% \[
+% d_n =  \sum_{s=0}^n b_{n-s} c_s = (-1)^n \sum_{s=0}^n \frac{\sin^{2(n-s)}(\phi)\cos^{2s+1}(\phi)}{s! (n-s)! (2s+1) \pi}.
+% \]
+% Combining these into~\eqref{eq:gfornormal} we obtain
+% \[
+% g(\theta) = \frac{1}{\sqrt{\kappa}}\sum_{n=0}^\infty \ell_{n} \kappa^{n} + \sum_{n=0}^\infty q_{n} \kappa^{n},
+% \]
+% where the coefficients
+% \begin{align*}
+% \ell_{n} &= \tfrac{1}{2\sqrt{\pi}}( 2 b_n + \cos^2(\phi) b_{n-1}) \\
+% q_{n} &= a_n + 2 d_n + \cos^2(\phi) d_{n-1}.
+% \end{align*}
+
+% Now
+% \[
+% A_1 = \int_{-\pi}^{\pi} g(\phi) \sin^2(\phi) d\phi = \frac{\pi}{2\sqrt{\pi\nu}} + O(1),
+% \]
+% \begin{align*}
+% A_2 &= \int_{-\pi}^{\pi} g(\phi) \sin^2\fracpart{\phi} d\phi \\
+% &= \frac{M}{2\sqrt{\pi\nu}} \int_{-\pi/M}^{\pi/M}\sin^2(\phi) d\phi + O(1) \\
+% &= \frac{\pi - M\sin(\tfrac{\pi}{M})}{2\sqrt{\pi\nu}},
+% \end{align*}
+% \begin{align*}
+% h_2(0) &= \int_{-\pi}^{\pi} g(\phi) \cos\fracpart{\phi} d\phi \\
+% &= \frac{M}{2\sqrt{\pi\nu}} \int_{-\pi/M}^{\pi/M}\cos(\phi) d\phi + O(1) \\
+% &= \frac{M\sin(\tfrac{\pi}{M})}{\sqrt{\pi\nu}}
+% \end{align*}
+% and so
+% \begin{align*}
+% H &= h_2(0) - 2 \sin(\tfrac{\pi}{M}) \sum_{k = 0}^{M-1} g(\tfrac{2\pi}{M}k + \tfrac{\pi}{M}) \\
+% &= h_2(0) -  \frac{M\sin(\pi/M)}{\sqrt{\pi\nu}}\\
+% &= O(1)
+% \end{align*}
+% Ok, you need to expand to one more term.
+
+%\section{When the SNR is large}
 

### 2013-01-15 paper.tex
some simpifications

@@
 We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator is not consistent, but converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
-Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.
+However, the amplitude estimator is not consistent and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
+%Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.
 The results of Monte Carlo simulations are provided and these agree with the theoretical results.   
@@
 
-We assume that time offset estimation has been performed and that $L$ noisy $M$-PSK symbols are observed by the receiver.  The received signal after matched filtering is,
+We assume that %time offset estimation has been performed and that 
+$L$ noisy $M$-PSK symbols are observed by the receiver,  %The received signal after matched filtering is,
 \begin{equation}\label{eq:sigmod}
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Swelden considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
+Both Mackenthun and Sweldens considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
@@
 The metric of performance is correspondingly the \emph{symbol error rate}, or \emph{bit error rate}.
-While estimating the symbols (or more precisely the transmitted bits) is ultimately the goal, we take the opposite point of view here.
+While estimating the data symbols is ultimately the goal, we take the opposite point of view here.
 Our aim is to estimate $a_0$, and we treat the unknown data symbols as nuisance parameters.
@@
 
-It is possible to generalise the results we present here to allow data symbols with varying constellation size, i.e.
+%It is possible to generalise the results we present here to allow data symbols with varying constellation size, i.e.
 varying $M$.
@@
 Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under some reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
+We show, under reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
@@
 \end{equation}
-where $L = \abs{P \cup D}$ is the total number of symbols transmitted, and to simplify our notation we have put 
+where $L = \abs{P \cup D}$ and to simplify our notation we have put 
 \[
@@
 
-To see this, let $a = \rho e^{j\theta}$ where $\rho$ is a nonnegative real.  Now,
+To see this, let $a = \rho e^{j\theta}$ where $\rho > 0$ is real.  Now,
 \begin{align}
@@
 \]
-be the phase difference between the received symbol $y_i$ and the hard decision resulting when $\theta = 0$, i.e. $\round{\angle{y_i}}$.  Then,
+%be the phase difference between the received symbol $y_i$ and the hard decision resulting when $\theta = 0$, i.e. $\round{\angle{y_i}}$.  
+Then,
 \begin{equation}\label{eq:uicombos}
@@
 If $\hat{\theta}$ is the least squares estimator of the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $d_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
+Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $\hat{d}_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
 
@@
 \]
-So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in a $O(1)$ arithmetic operations.
+So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in $O(1)$ arithmetic operations.
 Given $Y_k$, the value of $SS(f_k)$ can be computed in $O(1)$ operations using~\eqref{eq:SSfk}.

### 2013-01-15 paper.tex
added new trans sp version

@@
 \]
-Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
+% Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
 If $\expect\abs{W} = \expect Z$ is finite, then $W$ has zero mean because
-\begin{align*}
- \expect W &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
- &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
- &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
- \end{align*}
-If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
-\[
-f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
-\]
+% \begin{align*}
+%  \expect W &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
+%  &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
+%  &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
+%  \end{align*}
+% If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
+% \[
+% f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
+% \]
 %from which follows
@@
 \]
-The joint pdf of $R$ and $\Phi$ can be shown to be
-\begin{equation}\label{eq:pdfRPhi}
-f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
-\end{equation}
+% The joint pdf of $R$ and $\Phi$ can be shown to be
+% \begin{equation}\label{eq:pdfRPhi}
+% f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
+% \end{equation}
 Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi]$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$.
@@
 So,
-\begin{equation}\label{eq:expectRepartRphi}
+%\begin{equation}\label{eq:expectRepartRphi}
+\[
 \expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
-\end{equation}
+\]
+%\end{equation}
 where $\Re(\cdot)$ denotes the real part, and
@@
 
-The assumption that $\expect \sabs{w_1}^2 = \expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.
+%The assumption that $\expect \sabs{w_1}^2 = \expect \sabs{w_i}^2$ is finite implies that $R_i$ has finite variance since $\expect R_i^2 = 1 + \expect \abs{w_i}^2$.
 This is required in Theorem~\ref{thm:normality} so that the constants $A_1$, $A_2$, $B_1$ and $B_2$ exist.
@@
 Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\hat{\theta} - \theta_0}_\pi$ will not converge.
-One can check that this assumption holds when $w_1$ is circularly symmetric and normally distributed.
+One can check that this assumption holds when $w_i$ is circularly symmetric and normally distributed.
 We will not attempt to further classify those distributions for which the assumption holds here.
@@
 Theorem~\ref{thm:consistency} defines real numbers $p$ and $d$ to represent the proportion of pilot symbols and data symbols in the limit as $L$ goes to infinity.
-For Theorem~\ref{thm:normality} we need the slightly stronger condition that
-\[
-\frac{\abs{P}}{L} = p + o(L^{-1/2}) \qquad \text{and} \qquad \frac{\abs{D}}{L} = d + o(L^{-1/2}).
-\] 
+For Theorem~\ref{thm:normality} we need the slightly stronger condition that $\frac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\frac{\abs{D}}{L} = d + o(L^{-1/2})$.
 %where $o(L^{-1/2})$ denotes a number such that $\sqrt{L} o(L^{-1/2})$ goes to zero as $L\rightarrow \infty$.  
 This stronger condition is required to prove the asymptotic normality of $\sqrt{L}\hat{m}_L$.
-
-The next two sections give proofs of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.  The proofs make use of various lemmas, which are proved in the appendix.
+The next two sections give proofs of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+%The proofs make use of various lemmas, which are proved in the appendix.
 
@@
 Let 
-\begin{equation}\label{eq:GLdefn}
-G_L(\lambda) = \frac{1}{\rho_0 L} \Re(Z(\theta))
-\end{equation}
+\begin{align}
+%\begin{split}
+G_L(\lambda) &= \frac{1}{\rho_0 L} \Re(Z(\theta)) \label{eq:GLdefn} \\
+&= \frac{1}{L} \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i} \nonumber
+%\end{split}
+\end{align}
 and put $\hat{\lambda}_L = -\sfracpart{\hat{\theta} - \theta_0}_\pi = \sfracpart{\theta_0 - \hat{\theta}}_\pi$.
@@
 Lemma~\ref{lem:GLtoG0} in the appendix shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
-It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
-These are proved in Section~\ref{sec:proof-almost-sureappendix} of the appendix.
+%It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
+These are proved in %Section~\ref{sec:proof-almost-sureappendix} of
+%the appendix.
+
+\begin{lemma}\label{lem:convtoexpGlamL} 
+$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Since $G(x)$ is uniquely maximised at $x=0$,
+\[
+0 \leq G(0) - G(\hat{\lambda}_L),
+\]
+and since $\hat{\lambda}_L$ is the maximiser of $G_L(x)$,
+\[ 
+0 \leq G_L(\hat{\lambda}_L) - G_L(0).
+\]
+Thus,
+\begin{align*}
+0 &\leq G(0) - G(\hat{\lambda}_L) \\ 
+& \leq G(0) - G(\hat{\lambda}_L) + G_L(\hat{\lambda}_L) - G_L(0) \\
+& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
+&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
+\end{align*}
+and the last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:uniflawGGL} 
+$\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+This type of result is called a~\emph{uniform law of large numbers} and follows from standard arguments.
+See, for example van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
+A specific proof of this result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:GLtoG0}
+$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+By the triangle inequality,
+\[
+\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
+\]
+Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{IEEEproof}
 
@@
 \]
-as $L \rightarrow \infty$.  The lemmas that we have used are proved in Section~\ref{sec:proof-asympt-normappendix} of the appendix.
+as $L \rightarrow \infty$.  The lemmas that we have used are proved in the appendix.
 
@@
 The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large.
-As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$ and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
 Theorem~\ref{thm:normality} does not model this behaviour in the sense that for any fixed $L$ there exist sufficiently small values of SNR for which Theorem~\ref{thm:normality} does not produce accurate predictions of the MSE.
@@
 
-Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the variance of the amplitude estimator when $M=2,4,8$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
+shows the variance of the amplitude estimator when $M=2$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
 The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
@@
 This requires $G(0)$ to be known.
-In practice $G(0)$ may not be known at the receiver, so Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} serve to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
+In practice $G(0)$ may not be known at the receiver, so Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK}
+serves to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
 When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
@@
 
-\begin{figure}[p]
+\begin{figure}[tp]
 	\centering
@@
 
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-		\caption{Unbiased amplitude error versus SNR for QPSK.}
-		\label{fig:plotampQPSK}
-\end{figure}
-
-\begin{figure}[p]
-	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-		\caption{Unbiased amplitude error versus SNR for $8$-PSK.}
-		\label{fig:plotamp8PSK}
-\end{figure}
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+% 		\caption{Unbiased amplitude error versus SNR for QPSK.}
+% 		\label{fig:plotampQPSK}
+% \end{figure}
+
+% \begin{figure}[p]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+% 		\caption{Unbiased amplitude error versus SNR for $8$-PSK.}
+% 		\label{fig:plotamp8PSK}
+% \end{figure}
 
@@
 Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
-This suggests the use of a weighted objective function, which would be an interesting topic for future research.
+%This suggests the use of a weighted objective function, which would be an interesting topic for future research.
 
@@
 
-\subsection{Lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sureappendix}
-
-\begin{lemma}\label{lem:convtoexpGlamL} 
-$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Since $G(x)$ is uniquely maximised at $x=0$,
-\[
-0 \leq G(0) - G(\hat{\lambda}_L),
-\]
-and since $\hat{\lambda}_L$ is the maximiser of $G_L(x)$,
-\[ 
-0 \leq G_L(\hat{\lambda}_L) - G_L(0).
-\]
-Thus,
-\begin{align*}
-0 &\leq G(0) - G(\hat{\lambda}_L) \\ 
-& \leq G(0) - G(\hat{\lambda}_L) + G_L(\hat{\lambda}_L) - G_L(0) \\
-& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
-&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
-\end{align*}
-and the last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}.
-\end{IEEEproof}
+%\subsection{Lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sureappendix}
 
@@
 
-\begin{lemma}\label{lem:uniflawGGL} 
-$\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Put $T_L(\lambda) = \expect G_L(\lambda)$ and write
-\begin{align*}
-&\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
-&= \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda) + T_L(\lambda) - G(\lambda)} \\
-&\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)}.
-\end{align*}
-Now,
-\begin{align*}
-T_L(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
-&= o(1) h_1(\lambda) + o(1) h_2(\lambda)
-\end{align*}
-%where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  
-Since 
-\[
-\sabs{h_1(\lambda)} = \sabs{\expect R_1\cos(\lambda + \Phi_1)} \leq \expect R_1,
-\]
-and 
-\[
-\sabs{h_2(\lambda)} = \sabs{\expect R_1\cos\fracpart{\lambda + \Phi_1}} \leq \expect R_1
-\] 
-for all $\lambda \in [-\pi, \pi)$, it follows that 
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)} \leq o(1) \expect R_1 \rightarrow 0
-\]  
-as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
-\]
-almost surely as $L\rightarrow \infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:uniflawTGL} 
-Put $T_L(\lambda) = \expect G_L(\lambda)$.  Then
-\[
-\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
-\] 
-almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Put $D_L(\lambda) = G_L(\lambda) - T_L(\lambda)$ and let
-\[
-\lambda_n = \tfrac{2\pi}{N}(n-1) - \pi, \qquad n = 1, \dots, N
-\]
-be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n = [\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
-\begin{align*}
-\sup_{\lambda \in [-\pi, \pi)}&\sabs{D_L(\lambda)} \\
-&= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n) + D_L(\lambda_n)} \\
-&\leq U_L + V_L,
-\end{align*}
-where 
-\[
-U_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)} \;\;\; \text{and}
-\]
-\[
-V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
-\]
-Lemma~\ref{lem:supDLlambdan} shows that for any $N$ and $\epsilon > 0$, 
-\[
-\prob\left\{ \lim_{L\rightarrow\infty} U_L > \epsilon \right\} = 0,
-\]
-that is, $U_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
-\] 
-If we choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ then
-\begin{align*}
-\prob& \left\{ \lim_{L\rightarrow\infty} \sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} > 3\epsilon \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty}(U_L + V_L) > 3\epsilon \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty} (U_L + V_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
-&\leq \prob\left\{\lim_{L\rightarrow\infty} U_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
-&= 0.
-\end{align*}
-Thus $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:supDLlambdan}
-For any $N > 0$, $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$ where $U_L$ is defined in the proof of Lemma~\ref{lem:uniflawTGL}.
-\end{lemma}
-\begin{IEEEproof}
-Put
-\[
-Z_i(\lambda) = \begin{cases}
-R_i\cos(\lambda + \Phi_i) , & i \in P \\
-R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
-\end{cases}
-\]
-so that
-\[
-D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
-\]
-Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
-\[
-\abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda_n)\big)} \rightarrow 0
-\]
-almost surely as $L\rightarrow\infty$ by Kolmogorov's strong law of large numbers~\cite{Billingsley1979_probability_and_measure}.  Thus
-\[
-U_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
-\]
-almost surely to zero as $L \rightarrow \infty$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} = 0.
-\]
-\end{lemma}
-\begin{IEEEproof}
-Observe that
-\begin{align*}
-&\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
-&=  \abs{G_L(\lambda) - T_L(\lambda) - G_L(\lambda_n) + T_L(\lambda_n)} \\
-&\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
-&\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
-\end{align*}
-the last line following from Jensen's inequality.  Put
-\[
-C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) },
-\]
-so that
-\begin{align*}
-V_L &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
-&\leq  C_L +  \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect\abs{G_L(\lambda) - G_L(\lambda_n) } \\
-&\leq C_L + \expect C_L,
-\end{align*}
-where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
-\[
-\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
-\]
-Thus,
-\begin{align*}
-\prob&\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
-\end{align*}
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:CL} The following statements hold:
-\begin{enumerate}
-\item $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ for all positive integers $L$,
-\item for any $\epsilon > 0$, $\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0$.
-\end{enumerate}
-\end{lemma}
-\begin{IEEEproof}
-If $\lambda \in L_n$, then $\lambda = \lambda_n + \delta$ with $\delta < \tfrac{2\pi}{N}$, and from Lemma~\ref{lem:coslipshitz},
-\begin{align*}
-&\abs{\cos(\lambda + \Phi_i) - \cos(\lambda_n + \Phi_i)} \leq \tfrac{2\pi}{N}, \;\;\; \text{and} \\
-&\abs{\cos\sfracpart{\lambda + \Phi_i} - \cos\sfracpart{\lambda_n + \Phi_i}} \leq \tfrac{2\pi}{N}.
-\end{align*}
-Because these results do not depend on $n$,
-\[
-\sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \leq R_i \frac{2\pi}{N}
-\]
-for all $i = P \cup D$.  Also
-\begin{align*}
-C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i \in P \cup D} Z_i(\lambda) - Z_i(\lambda_n)} \\
-&\leq \frac{1}{L}\sum_{i \in P \cup D} \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
-&\leq \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i .
-\end{align*}
-Thus, 
-\[
-\expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i = \frac{2\pi}{N}\expect R_1
-\]
-and the first statement holds.  Now, 
-\[
-\frac{2\pi}{N L}\sum_{i \in P \cup D} R_i \rightarrow \frac{2\pi}{N}\expect R_1
-\] 
-almost surely as $L \rightarrow\infty$ by the strong law of large numbers, and so, for any $\epsilon > 0$,
-\begin{align*}
-\prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} \\
-&\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
-\end{align*}
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:coslipshitz}
-Let $x$ and $\delta$ be real numbers.  Then
-\begin{align*}
-&\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta}, \;\;\; \text{and} \\
-&\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
-\end{align*}
-\end{lemma}
-\begin{IEEEproof}
-Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipschitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
-\begin{align*}
-&\abs{\cos(y) - \cos(x)} \leq K\abs{x-y} = \abs{x-y}, \;\;\; \text{and} \\
-&\abs{\cos\fracpart{y} - \cos\fracpart{x}} \leq K\abs{x-y} = \abs{x-y}.
-\end{align*}
- The lemma follows by putting $y = x + \delta$.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:GLtoG0}
-$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-By the triangle inequality,
-\[
-\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
-\]
-Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
-\end{IEEEproof}
-
-\subsection{Lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-normappendix}
+% \begin{lemma}\label{lem:uniflawGGL} 
+% $\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+% \end{lemma}
+% \begin{IEEEproof}
+% Put $T_L(\lambda) = \expect G_L(\lambda)$ and write
+% \begin{align*}
+% &\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \\
+% &= \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda) + T_L(\lambda) - G(\lambda)} \\
+% &\leq \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} +  \sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)}.
+% \end{align*}
+% Now,
+% \begin{align*}
+% T_L(\lambda) - G(\lambda) &= \big(\tfrac{\abs{P}}{L} - p\big)h_1(\lambda) + \big(\tfrac{\abs{D}}{L} - d\big)h_2(\lambda) \\
+% &= o(1) h_1(\lambda) + o(1) h_2(\lambda)
+% \end{align*}
+% %where $o(1)$ denotes a number converging to zero as $L \rightarrow \infty$.  
+% Since 
+% \[
+% \sabs{h_1(\lambda)} = \sabs{\expect R_1\cos(\lambda + \Phi_1)} \leq \expect R_1,
+% \]
+% and 
+% \[
+% \sabs{h_2(\lambda)} = \sabs{\expect R_1\cos\fracpart{\lambda + \Phi_1}} \leq \expect R_1
+% \] 
+% for all $\lambda \in [-\pi, \pi)$, it follows that 
+% \[
+% \sup_{\lambda \in [-\pi, \pi)}\sabs{T_L(\lambda) - G(\lambda)} \leq o(1) \expect R_1 \rightarrow 0
+% \]  
+% as $L\rightarrow \infty$.  Lemma~\ref{lem:uniflawTGL} shows that 
+% \[
+% \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
+% \]
+% almost surely as $L\rightarrow \infty$.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:uniflawTGL} 
+% Put $T_L(\lambda) = \expect G_L(\lambda)$.  Then
+% \[
+% \sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - T_L(\lambda)} \rightarrow 0
+% \] 
+% almost surely as $L \rightarrow \infty$.
+% \end{lemma}
+% \begin{IEEEproof}
+% Put $D_L(\lambda) = G_L(\lambda) - T_L(\lambda)$ and let
+% \[
+% \lambda_n = \tfrac{2\pi}{N}(n-1) - \pi, \qquad n = 1, \dots, N
+% \]
+% be $N$ points uniformly spaced on the interval $[-\pi, \pi)$.  Let $L_n = [\lambda_n, \lambda_n + \tfrac{2\pi}{N})$ and observe that $L_1, \dots, L_N$ partition $[-\pi, \pi)$.  Now
+% \begin{align*}
+% \sup_{\lambda \in [-\pi, \pi)}&\sabs{D_L(\lambda)} \\
+% &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n) + D_L(\lambda_n)} \\
+% &\leq U_L + V_L,
+% \end{align*}
+% where 
+% \[
+% U_L = \sup_{n = 1,\dots,N}\sabs{D_L(\lambda_n)} \;\;\; \text{and}
+% \]
+% \[
+% V_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)}.
+% \]
+% Lemma~\ref{lem:supDLlambdan} shows that for any $N$ and $\epsilon > 0$, 
+% \[
+% \prob\left\{ \lim_{L\rightarrow\infty} U_L > \epsilon \right\} = 0,
+% \]
+% that is, $U_L \rightarrow 0$ almost surely as $L\rightarrow\infty$.  Lemma~\ref{lem:supsupDLlambda} shows that for any $\epsilon > 0$,
+% \[
+% \prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} = 0.
+% \] 
+% If we choose $N$ large enough that $4\pi\expect R_i < \epsilon N$ then
+% \begin{align*}
+% \prob& \left\{ \lim_{L\rightarrow\infty} \sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} > 3\epsilon \right\} \\
+% &\leq \prob\left\{\lim_{L\rightarrow\infty}(U_L + V_L) > 3\epsilon \right\} \\
+% &\leq \prob\left\{\lim_{L\rightarrow\infty} (U_L + V_L) > \epsilon + \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+% &\leq \prob\left\{\lim_{L\rightarrow\infty} U_L > \epsilon\right\} +  \prob\left\{ \lim_{L\rightarrow\infty} V_L > \epsilon + \tfrac{4\pi }{N}\expect R_i \right\} \\
+% &= 0.
+% \end{align*}
+% Thus $\sup_{\lambda \in [-\pi, \pi)}\sabs{D_L(\lambda)} \rightarrow 0$ almost surely as $L\rightarrow\infty$.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:supDLlambdan}
+% For any $N > 0$, $U_L \rightarrow 0$ almost surely as $L \rightarrow \infty$ where $U_L$ is defined in the proof of Lemma~\ref{lem:uniflawTGL}.
+% \end{lemma}
+% \begin{IEEEproof}
+% Put
+% \[
+% Z_i(\lambda) = \begin{cases}
+% R_i\cos(\lambda + \Phi_i) , & i \in P \\
+% R_i\cos\sfracpart{\lambda + \Phi_i} , & i \in D
+% \end{cases}
+% \]
+% so that
+% \[
+% D_L(\lambda) = G_L(\lambda) - T_L(\lambda) = \frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda) - \expect Z_i(\lambda) \big).
+% \]
+% Now $Z_1(\lambda_n), \dots,Z_L(\lambda_n)$ are independent with finite variance (because $\expect R_i^2$ is finite), so for each $n =1, \dots, N$,
+% \[
+% \abs{D_L(\lambda_n)} = \abs{\frac{1}{L}\sum_{i\in P \cup D} \big( Z_i(\lambda_n) - \expect Z_i(\lambda_n)\big)} \rightarrow 0
+% \]
+% almost surely as $L\rightarrow\infty$ by Kolmogorov's strong law of large numbers~\cite{Billingsley1979_probability_and_measure}.  Thus
+% \[
+% U_L = \sup_{n=1,\dots,N}\sabs{D_L(\lambda_n)} \leq \sum_{n=1}^N \sabs{D_L(\lambda_n)} \rightarrow 0
+% \]
+% almost surely to zero as $L \rightarrow \infty$.
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:supsupDLlambda} For any $\epsilon > 0$,
+% \[
+% \prob\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_i \right\} = 0.
+% \]
+% \end{lemma}
+% \begin{IEEEproof}
+% Observe that
+% \begin{align*}
+% &\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
+% &=  \abs{G_L(\lambda) - T_L(\lambda) - G_L(\lambda_n) + T_L(\lambda_n)} \\
+% &\leq  \abs{G_L(\lambda) - G_L(\lambda_n)} + \abs{\expect G_L(\lambda) - \expect G_L(\lambda_n)} \\
+% &\leq \abs{G_L(\lambda) - G_L(\lambda_n) } + \expect \abs{G_L(\lambda) - G_L(\lambda_n)},
+% \end{align*}
+% the last line following from Jensen's inequality.  Put
+% \[
+% C_L = \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \abs{G_L(\lambda) - G_L(\lambda_n) },
+% \]
+% so that
+% \begin{align*}
+% V_L &= \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n}\sabs{D_L(\lambda) - D_L(\lambda_n)} \\
+% &\leq  C_L +  \sup_{n = 1,\dots,N}\sup_{\lambda \in L_n} \expect\abs{G_L(\lambda) - G_L(\lambda_n) } \\
+% &\leq C_L + \expect C_L,
+% \end{align*}
+% where the last line follows because $\sup \expect\vert \dots\vert \leq \expect \sup \vert \dots \vert$.  Lemma~\ref{lem:CL} shows that $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ and also that
+% \[
+% \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+% \]
+% Thus,
+% \begin{align*}
+% \prob&\left\{ \lim_{L \rightarrow \infty} V_L > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
+% &\leq \prob\left\{ \lim_{L \rightarrow \infty} (C_L + \expect C_L) > \epsilon + \tfrac{4\pi}{N}\expect R_1 \right\} \\
+% &\leq \prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+% \end{align*}
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:CL} The following statements hold:
+% \begin{enumerate}
+% \item $\expect C_L \leq \tfrac{2\pi}{N}\expect R_1$ for all positive integers $L$,
+% \item for any $\epsilon > 0$, $\prob\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0$.
+% \end{enumerate}
+% \end{lemma}
+% \begin{IEEEproof}
+% If $\lambda \in L_n$, then $\lambda = \lambda_n + \delta$ with $\delta < \tfrac{2\pi}{N}$, and from Lemma~\ref{lem:coslipshitz},
+% \begin{align*}
+% &\abs{\cos(\lambda + \Phi_i) - \cos(\lambda_n + \Phi_i)} \leq \tfrac{2\pi}{N}, \;\;\; \text{and} \\
+% &\abs{\cos\sfracpart{\lambda + \Phi_i} - \cos\sfracpart{\lambda_n + \Phi_i}} \leq \tfrac{2\pi}{N}.
+% \end{align*}
+% Because these results do not depend on $n$,
+% \[
+% \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \leq R_i \frac{2\pi}{N}
+% \]
+% for all $i = P \cup D$.  Also
+% \begin{align*}
+% C_L &= \sup_{n=1,\dots,N}\sup_{\lambda \in L_n}\abs{\frac{1}{L}\sum_{i \in P \cup D} Z_i(\lambda) - Z_i(\lambda_n)} \\
+% &\leq \frac{1}{L}\sum_{i \in P \cup D} \sup_{n=1,\dots,N}\sup_{\lambda \in L_n} \abs{Z_i(\lambda) - Z_i(\lambda_n)} \\
+% &\leq \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i .
+% \end{align*}
+% Thus, 
+% \[
+% \expect C_L \leq \expect \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i = \frac{2\pi}{N}\expect R_1
+% \]
+% and the first statement holds.  Now, 
+% \[
+% \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i \rightarrow \frac{2\pi}{N}\expect R_1
+% \] 
+% almost surely as $L \rightarrow\infty$ by the strong law of large numbers, and so, for any $\epsilon > 0$,
+% \begin{align*}
+% \prob&\left\{ \lim_{L \rightarrow \infty} C_L > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} \\
+% &\leq \prob\left\{ \lim_{L \rightarrow \infty} \frac{2\pi}{N L}\sum_{i \in P \cup D} R_i > \epsilon + \tfrac{2\pi}{N}\expect R_1 \right\} = 0.
+% \end{align*}
+% \end{IEEEproof}
+
+% \begin{lemma}\label{lem:coslipshitz}
+% Let $x$ and $\delta$ be real numbers.  Then
+% \begin{align*}
+% &\abs{\cos(x + \delta) - \cos(x)} \leq \abs{\delta}, \;\;\; \text{and} \\
+% &\abs{\cos\fracpart{x + \delta} - \cos\fracpart{x}} \leq \abs{\delta}.
+% \end{align*}
+% \end{lemma}
+% \begin{IEEEproof}
+% Both $\cos(x)$ and $\cos\fracpart{x}$ are Lipschitz continuous functions from $\reals$ to $\reals$ with constant $K=1$.  That is, for any $x$ and $y$ in $\reals$,
+% \begin{align*}
+% &\abs{\cos(y) - \cos(x)} \leq K\abs{x-y} = \abs{x-y}, \;\;\; \text{and} \\
+% &\abs{\cos\fracpart{y} - \cos\fracpart{x}} \leq K\abs{x-y} = \abs{x-y}.
+% \end{align*}
+%  The lemma follows by putting $y = x + \delta$.
+% \end{IEEEproof}
+
+%\subsection{Lemmas required for the proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-normappendix}
 
@@
 \begin{equation}\label{eq:YLdef}
-Y_L(\lambda) = \sqrt{L}\big( G_L(\lambda) - T_L(\lambda) \big) - X_L.
+Y_L(\lambda) = \sqrt{L}\big( G_L(\lambda) - T_L(\lambda) \big) - X_L
 \end{equation}

### 2013-01-21 paper.tex
minor edit
minor

@@
 \]
-We again abuse notation by reusing $SS$, but this should not cause confusion as $SS(\rho,\theta)$ and $SS(\theta)$ are easily told apart by their inputs.
+%We again abuse notation by reusing $SS$, but this should not cause confusion as $SS(\rho,\theta)$ and $SS(\theta)$ are easily told apart by their inputs.  
 By definition the amplitude $\rho_0$ and its estimator $\hat{\rho}$ are positive.
@@
 This type of result is called a~\emph{uniform law of large numbers} and follows from standard arguments.
-See, for example van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
-A specific proof of this result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
+See, for example, van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
+A proof of this specific result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
 \end{IEEEproof}
@@
 Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.
-When $\abs{P} \neq 0$ (i.e.
-coherent detection) the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
+When $\abs{P} \neq 0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
 Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
@@
 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-		\caption{Unbiased amplitude error versus SNR for BPSK.}
+		\caption{Amplitude estimator variance (bias removed) versus SNR for BPSK.}
 		\label{fig:plotampBPSK}
@@
 \end{equation}
-is an empirical process indexed by $\lambda$, similar to $W_L$ from~\eqref{eq:WLdef}.  As with $W_L$ results from the literature on empirical processes can be used to show that  
+is an empirical process indexed by $\lambda$, similar to $W_L$ from~\eqref{eq:WLdef}.  As with $W_L$ from~\eqref{eq:WLdef} results from the literature on empirical processes can be used to show that  
 %Lemma~\ref{lem:ZYLempiricproc} 
@@
 &=  \big( p + o(L^{-1/2}) \big) \expect R_1\cos(\lambda + \Phi_1) \\
-&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1},
+&\hspace{2cm} + \big( d + o(L^{-1/2}) \big) \expect R_1\cos\fracpart{\lambda + \Phi_1}.
 \end{align*}
-and because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ (by assumption in Theorem~\ref{thm:normality}), we have
+Because $\tfrac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\tfrac{\abs{D}}{L} = d + o(L^{-1/2})$ by assumption in Theorem~\ref{thm:normality}, we have
 \begin{align*}

### 2013-07-03 paper.tex
added changes?

@@
 The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.
-A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{A}$ where $\angle$ denotes the complex argument and
+A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{C}$ where $\angle$ denotes the complex argument and
 \begin{equation}\label{eq:viterbiviterbi}
-A = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
+C = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
 \end{equation}
@@
 Consequently, our metric of performance will not be symbol or bit error rate, but $\sabs{\hat{a} - a_0}$.
-It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\sabs{\langle\hat{\theta} - \theta_0\rangle_\pi}$ and $\sabs{\hat{\rho} - \rho_0}$.
+It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\sabs{\langle\theta_0 - \hat{\theta}\rangle_\pi}$ and $\sabs{\hat{\rho} - \rho_0}$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
-It will become apparent why $\langle\hat{\theta} - \theta_0\rangle_\pi$ rather than $\hat{\theta} - \theta_0$ is the appropriate measure of error for the phase parameter.
+It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi$ rather than $\theta_0 - \hat{\theta}$ is the appropriate measure of error for the phase parameter.
+It is usual to write \emph{estimator} minus \emph{error}, i.e., $\hat{\theta} - \theta_0$ rather than $\theta_0 - \hat{\theta}$.
+However, it is notationally convenient to use $\langle\theta_0 - \hat{\theta}\rangle_\pi$ rather than $\langle\hat{\theta} - \theta_0\rangle_\pi$ here.
 
@@
 Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\hat{\theta} - \theta_0\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\hat{\theta} - \theta_0\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
+We show, under reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
@@
 \]
-% The joint pdf of $R$ and $\Phi$ can be shown to be
-% \begin{equation}\label{eq:pdfRPhi}
-% f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
-% \end{equation}
+The joint pdf of $R$ and $\Phi$ can be shown to be
+ \begin{equation}\label{eq:pdfRPhi}
+ f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
+ \end{equation}
 Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi]$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$.
@@
 \begin{enumerate}
-\item $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
 \item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
@@
 and assume that $\frac{\abs{P}}{L} = p + o(L^{-1/2})$ and $\frac{\abs{D}}{L} = d + o(L^{-1/2})$.
-Put $\hat{\lambda}_L = -\sfracpart{\hat{\theta} - \theta_0}_\pi = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
+Put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$ and $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$.
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
@@
 
-The theorems place conditions on $\sfracpart{\hat{\theta} - \theta_0}_\pi$ rather than directly on $\hat{\theta}  - \theta_0$.
+The theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ rather than directly on $\theta_0 - \hat{\theta}$.
 This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
-So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{-0.99\pi - 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert -0.99\pi - 0.99\pi\vert = 1.98\pi$.
+So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{0.99\pi + 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert 0.99\pi + 0.99\pi\vert = 1.98\pi$.
 
@@
 This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\hat{\theta} - \theta_0}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\hat{\theta} - \theta_0}_\pi$ will not converge.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
 One can check that this assumption holds when $w_i$ is circularly symmetric and normally distributed.
@@
 \]
-Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ because $G$ is continuous and $\hat{\lambda}_L \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ by Lemma~\ref{lem:convtoexpGlamL}.
 \end{IEEEproof}

### 2013-11-06 response.tex
added Matlab ST1 framework

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{mathbf-abbrevs}
\input{defs}

%\usepackage{xr}
%\externaldocument{paper2}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, I.\ Vaughan L.\ Clarkson and Barry Quinn  
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

\maketitle

\section*{Reviewer 1}

\begin{enumerate}
\item\textbf{Comment:} 
	This paper is concerned with carrier phase and amplitude estimation for PSK modulated signalling.
The considered least squares estimator seems to be only a minor extension of the one from [7]. Yet, more than 4 pages are dedicated to its derivation. Perhaps this section could be made more concise.
	\\\textbf{Response:}
	We have shortened Sec.\ II to make it more concise and improve readability.
\item\label{R1it2}\textbf{Comment:} 
	As stated by the authors themselves, the key results and main contributions of the paper involve the asymptotic properties of the least squares estimator. As far as I can see the derivation seems correct and complete, yet it is mathematically very involved and rather difficult to follow.
	\\\textbf{Response:}	
	For improved readability of the manuscript, we have now moved all of the lemmas into the appendix. We have also added a discussion of the implications and the practical relevance of the theorems to Sec.\ III (Sec.\ IV in the original manuscript).

	\comment{Stress practical relevance by showing ASRP PER results. Mention trials and patents. Point out that knowledge of the estimator statistics makes it possible to assess the estimator performance for a given number of symbols/pilot symbols at a certain SNR without having to run time consuming simulations. Also point out that the results are valid for a wider range of noise distributions, whilst CRBs need to be derived for a specific noise distribution\dots}
\item
A few minor remarks are:
\begin{enumerate}
\item\label{R1it3a}\textbf{Comment:} 
	Estimation from data symbols only is referred to as noncoherent, I rather consider a detector with non-data-aided synchronization as coherent (since it uses estimates of the synchronization parameters to detect the data).
	\\\textbf{Response:}
	We agree with the reviewer and apologise for causing confusion with the use of the terms \emph{coherent} and \emph{noncoherent} in the original manuscript. In the updated manuscript, we refer to an estimator that uses known pilot symbols as \emph{data-aided} and to one that relies on receiver decisions on the data symbols as \emph{decision-directed}.
\item\textbf{Comment:} 
	missing `,' after (5)
	\\\textbf{Response:} We thank the reviewer for spotting the missing comma and have added it.
\item\textbf{Comment:} 
	p.\ 19, at the bottom: sortindicIes (typo, delete 'i')
	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
\item\textbf{Comment:} 
	In Section VI, I would prefer a reference to the Lemma's in the Appendix at the start of the section.
	\\\textbf{Response:} We have added a reference to the Lemmas in the appendix at the beginning of the section. Please note that Sec.\ VI of the original manuscript is now Sec.\ III.
\item\textbf{Comment:} 
	Section VIII, ``Theorem 2 does not model this behaviour\dots'' can you please further explain why? + Comparison with the Cramer-Rao bound results from the existing literature would be interesting.
	\\\textbf{Response:}
	We have removed the aforementioned sentence from the manuscript as we feel it has caused more confusion than provided insight. We hope the following explanation clears up any confusion it may have caused.  Theorem 2 shows that for $L\rightarrow\infty$, $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ is normally distributed with zero mean and finite variance.  The support of $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ is the interval $[-\sqrt{L}\pi, \sqrt{L}\pi)$, i.e.\ $(-\infty, \infty)$ as $L\rightarrow\infty$.  For finite $L$, however, $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ has finite support and thus, cannot be exactly normally distributed.  Nevertheless, provided the variance of the noise is sufficiently small (i.e.\ the SNR sufficiently high), the distribution of $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ will be concentrated in a finite interval much smaller than $[-\sqrt{L}\pi, \sqrt{L}\pi)$ and is very close to a normal distribution.  On the other hand, as the SNR decreases, $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ approaches a uniform distribution on $[-\sqrt{L}\pi, \sqrt{L}\pi)$.  Consequently, the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$.  The argument for the case without pilots is analogous.

	\comment{Add CRBs to plots.} Please note that Sec.\ VIII of the original manuscript is now Sec.\ VI.
\end{enumerate}
\end{enumerate}

\section*{Reviewer 2}

\begin{enumerate}
\item
	General remarks
\begin{enumerate}
\item\textbf{Comment:} 
	Estimation of carrier phase and amplitude for $M$-ary PSK signals has been studied by the authors. Extending MackenthunÕs algorithm to pilots, a least square (LS) method is developed, where major parts of the manuscript are dedicated to the statistical analysis of the LS approach.

	Generally speaking it is to admit that both wording and writing follow the IEEE standards, although I would say that the list of References should be checked and updated in this respect.
	\\\textbf{Response:}
	We thank the reviewer for pointing us to the suggested references and have added \comment{some/most of} them. Specifically, we have included references [c, d, e]. \comment{Add the remaining ones?}
\end{enumerate}
\item
Some editorial comments
\begin{enumerate}
\item\textbf{Comment:} 
	Paragraph before (13): \dots can be computed from its predecessor \dots
	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
\item\textbf{Comment:} 
	Section III and VII: it would be helpful to provide $f(r, \phi)$
	\\\textbf{Response:}
	We thank the reviewer for pointing out the missing definition of $f(r, \phi)$ and have added it. Please note that in response to comment \ref{R4it2e} of reviewer 4, Sec.\ III has been slightly shortened and has been moved to the appendix.
\item\textbf{Comment:} 
	In the introductory section (paragraph before Eq.\ 3), the authors are talking about ``noncoherent detection where no pilot symbols exist''. In my mind this is somewhat confusing, because parameter estimation without knowing the underlying data is usually denoted as non-data aided (NDA) or decision-directed (DD), whereas when these data are known we have a data-aided (DA) situation [a]. On the other hand, coherent/noncoherent scenarios are given if the carrier phase is known/unknown to the receiver station.
	\\\textbf{Response:}
	Please refer to our response to comment \ref{R1it3a} of reviewer 1.
\end{enumerate}
\item
Major issues
\begin{enumerate}
\item\textbf{Comment:} 
	Figs.\ 1Ð3 illustrate the evolution of the phase MSE vs.\ SNR. Only in case that $|P| = 0$,
the LS results are compared to an alternative solution represented by the Viterbi and Viterbi (VV) algorithm. Why not doing the same with $|P| > 0$ by the following approach, which is definitely much simpler from the computational point of view (complexity in the order $O(L)$ instead of $O(L \log(L))$ required by LS) and the jitter performance will probably differ not very much from that of the LS variant (in Figs.\ 1Ð3 this can be seen for $|P| = 0$):
	\begin{itemize}
	\item
		If SNR $<$ SNRth, then the DA algorithm has to be used, else the VV algorithm must be applied.
	\item
		The SNRth is a threshold value known in advance, e.g., by the intersection of the Cramer-Rao lower bounds (CRLBs) for DA and NDA phase estimation [a, b].
	\item
		The SNR estimates might be obtained by a simple method based on second- and fourth-order moments [c].
	\item
		The same philosophy applies also to the estimation of amplitudes; for carrier- blind NDA results, the moment-based approach used for SNR estimation can be exploited.
	\end{itemize}
	\textbf{Response:}
	\comment{add response\dots}
\item\textbf{Comment:} 
	By detailed inspection of Fig.\ 4 showing the jitter variance for amplitude estimates, it is easily observed that $|P| < L$ performs better than $|P| = L$ in the medium-to-low SNR range. In my mind this is somewhat strange, since it should be the other way round (similar to the evolution of the phase curves in Figs.\ 1Ð3 and 5). Can you comment on this phenomenon ?
	\\\textbf{Response:}
	\comment{Replace variance plot by MSE plot.}
\item\textbf{Comment:} 
	In a much more powerful approach to the topic addressed in the current paper, the pilot sequence would be employed for a coarse estimation/correction of the carrier phase of the payload data. In the sequel, by embedding carrier phase recovery into an iterative (turbo) decoding scheme [d, e] or any other error correcting method [f, g], the jitter performance is close to the CRLB for DA estimation even at very small SNRs, which are mainly determined by the convergence region of the (turbo) decoder. By taking into account the fact that modern and powerful communication systems are not conceivable without any FEC scheme, it is not clear to me why I should choose your algorithm in such a context.
	\\\textbf{Response:}
	We have added a brief discussion of the need for sufficiently accurate coarse estimates to initialise an iterative code-aided synchronisation stage, citing [d, e].  We have also added a set of new simulation results, which show the packet error rate of an LDPC-coded satellite communication system to demonstrate the use and performance of the proposed estimator in a practical system.  \comment{Add details\dots}

	\comment{Add response related to [f, g]\dots}
\end{enumerate}
\end{enumerate}

\subsection*{Reviewer 2 references}
\begin{enumerate}\footnotesize
\renewcommand{\theenumi}{\alph{enumi}}
\renewcommand{\labelenumi}{[\theenumi]}
\addtolength{\itemsep}{-1.5ex}
\item
	Mengali, D'Andrea, \textit{Synchronization Techniques for Digital Receivers}, Plenum Press: New York, 1997.
\item
	Rice, Cowley, Moran, Rice, ``Cramer-Rao Lower Bounds for QAM Phase and Frequency Estimation'', \textit{IEEE Trans.\ Commun.}, Sept 2001.
\item
	Pauluzzi, Beaulieu, ``A Comparison of SNR Estimation Techniques for the AWGN Channel'', \textit{IEEE Trans.\ Commun.}, Oct.\ 2000.
\item
	Lottici, Luise, ``Embedding Carrier Phase Recovery into Iterative Decoding of Turbo-Coded Linear Modulations'', \textit{IEEE Trans.\ Commun.}, April 2004.
\item
	Noels et al., ``A Theoretical Framework for Soft-Information-Based Synchronization in Iterative (Turbo) Receivers'', \textit{Eurasip Journal Wireless Commun.\ and Networking}, April 2005.
\item
	D'Andrea, Mengali, Vitetta, ``Approximate ML Decoding of Coded PSK with no Explicit Carrier Phase Reference'', \textit{IEEE Trans.\ Commun.}, Feb./March/April 1994.
\item
	Vanelli, Salmi, Cioni, Corazza, Polydoros, ``A Performance Review of PSP for Joint Phase/Frequency and Data Estimation in Future Broadband Satellite Networks'', \textit{IEEE Journal on Selected Areas in Commun.}, Dec.\ 2001.
\end{enumerate}

\section*{Reviewer 3}

\begin{enumerate}
\item
	\textbf{Comment:} 
The paper studies the estimation of amplitude and phase due to the distortion introduced by the channel in communication systems with $M$-PSK signals. The estimator is derived from the observation of $L$ noisy samples taken at a rate of one sample per symbol. Part of the $L$ observed samples corresponds to pilot transmitted symbols, that are already known a priori by the receiver, while the rest of samples are associated with data symbols which are unknown at the receiver. The focus of the article is to determine the amplitude and phase without considering the estimation of the transmitted data like this problem is usually treated in the literature. The work modifies the Mackenthun algorithm to include samples corresponding to pilot symbols while the Mackenthun paper calculates a least squares estimator for a similar problem but without pilot symbols.

The major contribution of this paper is the description of the asymptotic properties of the least squares estimator for the phase and amplitude. Although it seems technically correct, the authors do not make clear the relevance of their work to signal processing applications and to timely problems. They limit themselves in presenting several lemmas and proofs. Therefore, I cannot recommend this manuscript for publication in the IEEE Transactions on Signal Processing.
	\\\textbf{Response:}
	We thank the reviewer for the comments. Please refer to our response to comment \ref{R1it2} of reviewer 1.
\end{enumerate}

\section*{Reviewer 4}

\begin{enumerate}
\item\textbf{Comment:}
	This work addresses the problem of carrier phase and amplitude estimation for $M$-ary PSK signals. Based on a modification of the Mackhentum's algorithm by inserting pilots, the main contribution of this manuscript is in fact the study of the asymptotic behavior of the least squares estimator for carrier phase and amplitude. Although the derivations seems correct, it is not clear to the reader why this method should be applicable in a communications context.  A cconsiderable part of the manuscript is dedicated to proofs of Lemmas and Theorems, which makes the paper rather difficult to read. There are several methods in the litterature that proposes the use of pilots to estimate the carrier phase followed by efficient decoding such as turbo decoding. How does the proposed method compare with the existing uncoherent methods that do not require estimation of the amplitudes? Even if the data are coded, what are the gains in terms of data recovery of the proposed method and traditional solutions? In my opinion, the authors fail to convincing on this aspect and this is an important point.
	\\\textbf{Response:}
	\comment{Stress practical relevance by showing ASRP PER results. In particular ST1, where the estimators are used to seed turbo sync. Mention trials and patents\dots}

	\comment{Not clear what the reviewer means by ``existing uncoherent methods that do not require estimation of the amplitudes''.}

\item
Some additional comments:
\begin{enumerate}
\item\textbf{Comment:} 
	The authors should better motivate the interest in using their proposed method in actual communication systems. The lack of motivation for addressing the problem raised in the manuscript is a weak aspect of this work.
	\\\textbf{Response:}
	\comment{Show PER ASRP results. Mention trials and patents\dots}
\item\textbf{Comment:} 
	The simulation results should bring comparisons with existing methods that use error correction and turbo decoding techniques, which have shown to perform reasonably well under adverse situations and low SNR levels.
	\\\textbf{Response:}
	\comment{Show PER ASRP ST1 turbo sync results. Comparison with other estimators?}
\item\textbf{Comment:} 
	In Fig.\ 4, the influence of $P$ on the performance is not clear to me and does not seem to be in accordance with other results in the manuscript.
	\\\textbf{Response:} 
	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the estimator MSE. This was done to validate the theoretical value of the asymptotic variance of $\hat\rho$ predicted by Theorem~2 against Monte-Carlo simulations. For consistency with the other figures and for easier interpretation, we have replaced the variance plot by a plot of the MSE of $\hat\rho$.
\item\textbf{Comment:} 
	It is important that the authors address an appropriate method for correcting the bias, as they mentioned in the text. A solution to provide the receiver with an estimate of $G(0)$ must be addressed in order to make this contribution applicable in a real communications context.
	\\\textbf{Response:} 
	Biased estimators are quite common in the communications literature and in fact, are often used in practice. Prime examples are state-of-the-art signal-to-noise ratio (SNR) estimators for $M$-PSK modulated signals such as those discussed by Pauluzzi and Beaulieu \cite{Pauluzzi2000}. With the exception of the data-aided SNR estimators, all of the estimators in \cite{Pauluzzi2000} are biased and particularly at low SNR, the bias can be large. To the best of the authors' knowledge, the value of the bias of these estimators has only been determined by computer simulation and has not been analysed theoretically. In contrast, the asymptotic analysis of the amplitude estimator $\hat\rho$ presented in Theorem~1 of our manuscript results in a closed-form expression for the bias and thus, lays the foundation for improving the estimator by addressing the bias. However, removal of the bias is beyond the scope of this manuscript.
\item\textbf{Comment:} \label{R4it2e}
	Some sectioms can be shortened to make the work more readable and concise. This is the case of Section II for instance. Section III is not necessary. It is too short and the main concepts are well known. A reference to a textbook would be sufficient. Section VII is also too short and is not justified.
	\\\textbf{Response:}
	We have shortened Sec.\ II to make it more concise and improve readability. Sec.\ VII has been removed from the revised manuscript.

	Furthermore, we agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that Sec.\ III specifically discusses random variables of the form $Re^{j\varPhi}=1{+}W$, where $W$ is circularly symmetric. The joint pdf of $R$ and $\varPhi$ is a crucial ingredient for the asymptotic analysis presented in the manuscript and therefore, we feel the section should be retained. However, for improved readability, Sec.\ III has been slightly shortened and has been moved to the appendix.
\end{enumerate}
\end{enumerate}

\bibstyle{IEEEtran}
\bibliography{bib}

\end{document}

### 2013-11-06 code/st1matlab/Acquisition/ST1ULacquisitionRefineNoiseVar.m

@@ -1,0 +1,58 @@
+% Refine noise variance estimate.
+%-------------------------------------------------------------
+% ST1ULacquisitionRefineNoiseVar.m 
+%
+% [SoftDec] = ST1ULacquisitionRefineNoiseVar(Rx, SoftDec, ModemParams, STAcqParams)
+% 
+% This function should be called after ST1ULacquisitionRefinePeriodoWrapper
+% and refines the noise variance estimate in code-aided fashion, i.e. the
+% soft symbols provided by the decoder are treated as pilot symbols.
+% 
+% Inputs:
+%   Rx              - Vector containing samples of received signal
+%   SoftDec         - ST1UL SoftDec struct
+%   ModemParams     - ST1UL ModemParams struct
+%   AcqParams       - Struct holding aquisition parameters
+%       .P              - position (indices) of pilot symbols
+%       .p              - pilot symbols
+%       .D              - position (indices) of data symbols
+%       .M              - M for M-PSK, i.e. size of constellation
+%       ToneInSymbols   - Vector containing the tone after matched 
+%                         filtering and downsampling to symbol rate.
+%
+% Outputs:
+%   SoftDec         - ST1UL SoftDec struct with refined noise variance
+%                     estimate
+%
+% Comments:
+% 
+%-------------------------------------------------------------
+% Author: A. Pollok, Institute for Telecommunications Research
+% Project: ASRP
+%-------------------------------------------------------------
+% Copyright 2013
+% Institute for Telecommunications Research
+% University of South Australia
+%-------------------------------------------------------------
+
+function [SoftDec] = ST1ULacquisitionRefineNoiseVar(Rx, SoftDec, ModemParams, STAcqParams)
+
+% get matched filter output
+ModemParams.ChanType    = 'chancomp';
+DataComp                = satchan(Rx, SoftDec, ModemParams, SoftDec, [],  STAcqParams);
+MFoutput                = DataComp/sqrt(SoftDec.SigPowEst); %#ok<*AGROW>
+MFoutput                = MFoutput(1:ModemParams.CodedSymLenR);
+
+% tone cancellation
+if STAcqParams.CancelTone
+    MFoutput = MFoutput - STAcqParams.ToneInSymbols;
+end
+
+% re-estimate Es/No, treating decoder soft symbols as pilots
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/buildmex.m

@@ -1,0 +1,28 @@
+% Utility to compile all MEX files for reference model
+%--------------------------------------------------------------------------
+% buildmex.m - This script enables compilation of the .cpp mex files used in
+% the reference model
+%    
+% buildmex
+%
+% Inputs
+%   None
+% Outputs
+%   None
+% Comments:
+% 
+%--------------------------------------------------------------------------
+% Author:  ITR
+% Project: ASRP
+%--------------------------------------------------------------------------
+% Copyright 2012 : 
+% Institute for Telecommunications Research
+% University of South Australia
+%--------------------------------------------------------------------------
+% 
+%--------------------------------------------------------------------------
+
+make_LDPCDec
+make_QAMModulator
+make_CoherentMackenthun
+make_TimingEstimator

### 2013-11-06 code/st1matlab/impl/software/library/phy/include/FinitePulse.h

@@ -1,0 +1,261 @@
+/* 
+ * File:   FinitePulse.h
+ * Author: Robby McKilliam
+ *
+ * Created on 23 January 2013, 1:10 PM
+ */
+
+#ifndef FINITEPULSE_H
+#define	FINITEPULSE_H
+
+#include <vector>
+#include <iostream>
+#include "Util.h"
+
+#ifndef VALTYPE
+#define VALTYPE double
+#endif
+#ifndef MSGTYPE
+#define MSGTYPE VALTYPE
+#endif
+
+using namespace std;
+
+namespace TimeOffset {
+
+class FinitePulse {
+public:
+    virtual ~FinitePulse() {
+    };
+    /** pulse is zero for t less than tmin */
+    virtual VALTYPE tmin() const = 0;
+    /** pulse is zero for t more than tmax */
+    virtual VALTYPE tmax() const = 0;
+    /** pulse period */
+    virtual VALTYPE T() const = 0;
+    /** The transmit pulse */
+    virtual VALTYPE pulse(VALTYPE t) const = 0;
+};
+
+/** Truncated sinc pulse goes out to the numzeros zero in both positive and negative directions */
+class TruncatedSincPulse : public FinitePulse {
+public:
+    
+    TruncatedSincPulse(VALTYPE T, unsigned int numzeros) : tmin_(-T*numzeros), tmax_(T*numzeros), T_(T), sqrtT(sqrt(T)) {};
+ TruncatedSincPulse(VALTYPE T, VALTYPE duration, bool dodgyflag) : tmin_(-T*(int)ceil(duration/T/2)), tmax_(T*(int)ceil(duration/T/2)), T_(T), sqrtT(sqrt(T)) {};
+    //TruncatedSincPulse() = delete; //no default constructor //c++11 only
+
+    /** The truncated sinc pulse */
+    inline VALTYPE pulse(VALTYPE t) const {
+        if (t > tmin_ && t < tmax_) return sinc(t / T_) / sqrtT;
+        else return 0.0;
+    }
+
+    inline VALTYPE tmin() const {
+        return tmin_;
+    }
+
+    inline VALTYPE tmax() const {
+        return tmax_;
+    }
+
+    inline VALTYPE T() const {
+        return T_;
+    }
+
+protected:
+    const VALTYPE tmin_;
+    const VALTYPE tmax_;
+    const VALTYPE T_;
+    const VALTYPE sqrtT;
+private:
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/include/LDPCDecV2.h

@@ -1,0 +1,72 @@
+#ifndef LDPCDEC_H
+#define LDPCDEC_H
+
+#define DECODER_SPA		0
+#define DECODER_MSA		1
+
+#ifndef MSGTYPE
+  #define MSGTYPE double
+#endif
+
+/// \brief		Main class for LDPC decoding
+/// \author		Gottfried Lechner (gottfried.lechner@unisa.edu.au)
+/// \version	3.0
+/// \date     Created      : March 2008
+/// \date     Last modified: April 2012
+///
+/// This class holds the LDPC decoder.
+class CLDPCDec
+{
+public:
+	CLDPCDec(void);
+	~CLDPCDec(void);
+
+	int						readDecoder(const char* filename);
+	
+  unsigned int  encodeLDGM(unsigned int *info, unsigned int *parity);
+  unsigned int  encodeRA(unsigned int *info, unsigned int *parity);
+	unsigned int	decode(MSGTYPE *Lch, MSGTYPE *Lapp, unsigned int maxit);
+
+	void					set_param_clearmsg(bool clearmsg);
+	void					set_param_method(unsigned int method);
+  void          set_param_corrvec(unsigned int len, MSGTYPE *corrvec);
+  
+	bool					dec_loaded();
+
+	unsigned int	getN();
+	unsigned int	getM();
+	unsigned int	getK();
+	unsigned int	getE();
+
+private:
+	unsigned int	m_N;
+	unsigned int	m_M;
+	unsigned int	m_E;
+	unsigned int	m_g;
+	unsigned int	*m_vardegree;
+	unsigned int	*m_chkdegree;
+	unsigned int	*m_interleaver;
+	MSGTYPE				*m_c2v;
+	MSGTYPE				*m_v2c;
+  MSGTYPE       *m_tmp;
+	int						*m_hard;
+	unsigned int	m_dcmax;
+	MSGTYPE 		  *m_mf, *m_mb;
+	unsigned int 	*m_sf, *m_sb;
+
+	bool					m_clearmsg;
+	unsigned int	m_method;
+	
+	bool					m_decloaded;
+
+  unsigned int  m_corrveclen;
+  MSGTYPE       *m_corrvec;
+  
+	unsigned int	decodeSPA(MSGTYPE *Lch, MSGTYPE *Lapp, unsigned int maxit);
+	unsigned int	decodeMSA(MSGTYPE *Lch, MSGTYPE *Lapp, unsigned int maxit);
+  
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/include/MainLoop.h

@@ -1,0 +1,91 @@
+/* 
+ * File:   MainLoop.h
+ * Author: mckillrg
+ *
+ * Created on 4 March 2013, 3:13 PM
+ */
+
+#ifndef MAINLOOP_H
+#define	MAINLOOP_H
+
+#include <complex>
+
+#ifndef VALTYPE
+#define VALTYPE double
+#endif
+#ifndef MSGTYPE
+#define MSGTYPE VALTYPE
+#endif
+
+using namespace std;
+
+namespace TimeOffset {
+
+#ifdef __ARM_NEON__
+  #warning "Compiling with NEON instruction set"
+  #include "arm_neon.h"
+  #include "neon_mathfun.h"
+    //some useful macros
+#define RMEMSTACKED(m) {std::real(rmem[m]), std::imag(rmem[m]), std::real(rmem[m+mstep]), std::imag(rmem[m+mstep])};
+#define RMEMFILL(v,m) v[0]=std::real(rmem[m]); v[1]=std::imag(rmem[m]); v[2]=std::real(rmem[m+mstep]); v[3]=std::imag(rmem[m+mstep]);
+#define PULSESTACKED(i) {pulsetable[i], pulsetable[i], pulsetable[i+istep], pulsetable[i+istep]};
+#define PULSEFILL(v,i) v[0]=pulsetable[i]; v[1]=pulsetable[i]; v[2]=pulsetable[i+istep]; v[3]=pulsetable[i+istep];
+     /** Runs the main filter loop, this is where most time is spent */
+    inline complex<VALTYPE> mainFilterLoop(int mfrom, int mto, int mstep, int ifrom, int istep, const complex<VALTYPE>* rmem, const VALTYPE* pulsetable) {
+
+      float32x4_t sum;
+      int numsteps = (mto - mfrom + 1)/mstep;
+      if( numsteps%2 == 0) { //load zeros if even
+	float32_t sump[4] = {0,0,0,0};
+	sum = vld1q_f32(sump);
+      } else { //load first one and increment counters if it's odd
+	complex<VALTYPE> first = rmem[mfrom] * pulsetable[ifrom];
+	float32_t sump[4] = {std::real(first),std::imag(first),0,0};
+	sum = vld1q_f32(sump);
+	mfrom += mstep;
+	ifrom += istep;
+      }
+      int istep2 = istep*2, mstep2 = mstep*2, mstep8 = mstep*8, istep8 = istep*8; //new counters
+      
+      float32_t ap[4], bp[4];
+      //main loop
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/include/MatlabHelper.h

@@ -1,0 +1,37 @@
+//
+//  MatlabHelper.h
+//  QAMModulator
+//
+//  Created by Gottfried Lechner on 12/04/12.
+//  Copyright (c) 2012 ITR, University of South Australia. All rights reserved.
+//
+
+#ifndef _CoherehentMackenthun_MatlabHelper_h
+#define _CoherehentMackenthun_MatlabHelper_h
+
+#ifndef MSGTYPE
+  #define MSGTYPE double
+#endif
+
+#include <complex>
+#include <vector>
+#include "mex.h"
+
+typedef std::complex<MSGTYPE> complexT;
+
+std::vector<MSGTYPE> GetFloatArray(const mxArray *src);
+std::vector<complexT> GetComplexArray(const mxArray *src);
+std::vector<int> GetIntArray(const mxArray *src);
+std::vector<int> GetIndexArray(const mxArray *src);
+mxArray* ToMatlabComplex(complexT src);
+mxArray* ToMatlabComplex(std::vector<complexT> src);
+void CopyMatlabComplexToVector(const mxArray *src, std::vector<complexT>& v);
+void CopyMatlabComplexToVector(const mxArray *src, std::vector<MSGTYPE>& v);
+void CopyMatlabComplexToVector(const mxArray *src, std::vector<int>& v);
+
+void GetFloatArray(const mxArray *src, MSGTYPE *dst, bool imag = false);
+void GetIntArray(const mxArray *src, unsigned int *dst);
+void SetFloatArray(MSGTYPE *src, mxArray *dst, bool imag = false);
+void SetIntArray(unsigned int *src, mxArray *dst);
+
+#endif

### 2013-11-06 code/st1matlab/impl/software/library/phy/include/QAMModulatorV2.h

@@ -1,0 +1,54 @@
+//
+//  QAMModulatorV2.h
+//  QAMModulator
+//
+//  Created by Gottfried Lechner on 09/02/13.
+//  Copyright (c) 2012 ITR, University of South Australia. All rights reserved.
+//
+
+#ifndef QAMModulator_QAMModulator_h
+#define QAMModulator_QAMModulator_h
+
+#ifndef MSGTYPE
+  #define MSGTYPE double
+#endif
+      
+#include <complex>
+#include <vector>
+typedef std::complex<MSGTYPE> complex;
+        
+class CQAMModulator
+{
+
+public:
+  CQAMModulator(void);
+  ~CQAMModulator(void);
+
+  void          init(unsigned int numSections,
+                     unsigned int numSymbols,
+                     std::vector<complex> Sym);
+    
+  void          set_ch(std::vector<complex> y, std::vector<MSGTYPE> sigma2);
+  void          set_ch(std::vector<complex> y, MSGTYPE sigma2);
+  
+  void          set_apri(std::vector<MSGTYPE> La);
+  
+  void          modulate(unsigned int *data, std::vector<complex>& y);
+  void          demodulate(std::vector<MSGTYPE>& Lout);
+  void          remodulate(std::vector<complex>& y, std::vector<MSGTYPE>& ResVar);
+  
+  unsigned int  getlength();
+  unsigned int  getbitspersymbol();
+  
+private:
+  unsigned int  m_numSections;
+  unsigned int  m_numSymbols;
+  unsigned int  m_bitsperSymbol;
+  unsigned int  m_apriset;
+  
+  std::vector<complex>  m_Sym;
+  std::vector<MSGTYPE>  m_La;
+  std::vector<MSGTYPE>  m_pch;
+};
+
+#endif

### 2013-11-06 code/st1matlab/impl/software/library/phy/include/fastlog.h

@@ -1,0 +1,144 @@
+/*=====================================================================*
+ *                   Copyright (C) 2011 Paul Mineiro                   *
+ * All rights reserved.                                                *
+ *                                                                     *
+ * Redistribution and use in source and binary forms, with             *
+ * or without modification, are permitted provided that the            *
+ * following conditions are met:                                       *
+ *                                                                     *
+ *     * Redistributions of source code must retain the                *
+ *     above copyright notice, this list of conditions and             *
+ *     the following disclaimer.                                       *
+ *                                                                     *
+ *     * Redistributions in binary form must reproduce the             *
+ *     above copyright notice, this list of conditions and             *
+ *     the following disclaimer in the documentation and/or            *
+ *     other materials provided with the distribution.                 *
+ *                                                                     *
+ *     * Neither the name of Paul Mineiro nor the names                *
+ *     of other contributors may be used to endorse or promote         *
+ *     products derived from this software without specific            *
+ *     prior written permission.                                       *
+ *                                                                     *
+ * THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND              *
+ * CONTRIBUTORS "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES,         *
+ * INCLUDING, BUT NOT LIMITED TO, THE IMPLIED WARRANTIES               *
+ * OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE             *
+ * ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER               *
+ * OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT,                 *
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/make_QAMModulator.m

@@ -1,0 +1,14 @@
+% compile with internal messages as double (default)
+% mex src/mexQAMModulator.cpp src/QAMModulator.cpp src/MatlabHelper.cpp -Iinclude -outdir bin
+% compile with internal messages as float
+% mex src/mexQAMModulator.cpp src/QAMModulator.cpp src/MatlabHelper.cpp -Iinclude -outdir bin -DMSGTYPE=float
+
+mex -output mexQAMModulator src/mexQAMModulatorV2.cpp src/QAMModulatorV2.cpp src/MatlabHelper.cpp -Iinclude -outdir bin -DMSGTYPE=float
+
+cd bin
+mexQAMModulator
+cd ..
+
+cd test
+test_QAMModulator;
+cd ..

### 2013-11-06 code/st1matlab/impl/software/library/phy/mexTimingEstimator_help.m

@@ -1,0 +1,36 @@
+% mexTimingEstimator   MATLAB wrapper for TimingEstimator class
+%
+%   This mex-file enables the use of the TimingEstimator class from MATLAB.
+%
+%   Construct a ST1 timing estimator 
+%       mexTimingEstimator(name,P,D,pilots,T,Ts,p,q,taumin,taumax,rolloff,duration,tol);
+%
+%   Construct a ST2 timing estimator 
+%       mexTimingEstimator(name,P,D,pilots,T,Ts,taumin,taumax,rolloff,duration);
+%   
+%   Run timing estimator
+%       tauhat = mexTimingEstimator(name,r);
+%
+%    Input parameters:
+%       name      - name of this estimator, a string. 
+%       P           - position (indices) of pilot symbols
+%       D           - position (indices) of data symbols       
+%       pilots      - complex pilot symbols
+%       T           - symbol period
+%       Ts         - sample period
+%       p, q       - relatively prime integers such that Ts=p*T/q.
+%                     For example for 4 times oversampling p = 1
+%                     and q = 4.
+%       taumin      - minimum time offset
+%       taumax      - maximum time offset
+%       rolloff     - rolloff for the root raised cosine transmit pulse 
+%       duration - duration of the transmit pulse
+%       tol         - [OPTIONAL] tolerance for time estimate. If omitted,
+%                     a default value of 1e-6 is used.
+%       r            - received complex signal
+%
+%   Output parameters:
+%       tauhat      - timing offset estimate
+%
+%   (C) 2012 Andre Pollok, ITR/UniSA
+%

### 2013-11-06 code/st1matlab/impl/software/library/phy/src/FrequencyEstimator.cpp

@@ -1,0 +1,727 @@
+//
+//  FrequencyEstimator.cpp
+//  FrequencyEstimator
+//
+//  Created by Andre Pollok on 07/02/2013.
+//  Copyright (c) 2013 ITR, University of South Australia. All rights reserved.
+//
+
+#include <iostream>
+# include <iomanip>
+#include <math.h>
+#include <stdlib.h>
+#include <limits>
+#include <complex>
+#include <vector>
+#include <algorithm>
+using namespace std;
+#include <fftw3.h>
+
+#ifdef __ARM_NEON__
+  #include "arm_neon.h"
+  #include "neon_mathfun.h"
+  #define applyPhasePolynomial applyPhasePolynomialNEON
+  // FFTW only available in single-precision (float) on ARM
+  #warning "Compiling with float version of FFTW..."
+  #define fftw_destroy_plan fftwf_destroy_plan
+  #define fftw_plan fftwf_plan
+  #define fftw_plan_dft_1d fftwf_plan_dft_1d
+  #define fftw_complex fftwf_complex
+  #define fftw_cleanup fftwf_cleanup
+  #define fftw_execute fftwf_execute
+#else
+  #define applyPhasePolynomial applyPhasePolynomialC
+#endif
+
+#ifndef FFTWFLOAT
+  #define FFTWFLOAT 0
+#endif
+#if FFTWFLOAT==1
+  #warning "Compiling with float version of FFTW..."
+  #define fftw_destroy_plan fftwf_destroy_plan
+  #define fftw_plan fftwf_plan
+  #define fftw_plan_dft_1d fftwf_plan_dft_1d
+  #define fftw_complex fftwf_complex
+  #define fftw_cleanup fftwf_cleanup
+  #define fftw_execute fftwf_execute
+#endif
+
+#include "FrequencyEstimator.h"
+
+#ifndef M_PI
+#define M_PI (3.141592653589793)
+#endif 
+#ifndef M_PI_4
+#define M_PI_4 (M_PI/4.0)
+#endif 
+
+//%%% FrequencyEstimator %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+/// \brief  Standard constructor.
+///
+FrequencyEstimator::FrequencyEstimator()
+{
+}
+
+//-----------------------------------------------------------------------------
+
+/// \brief  Standard destructor.
+///
+FrequencyEstimator::~FrequencyEstimator(void)
+{  
+}
+
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/src/IndexedRealSDR.cpp

@@ -1,0 +1,15 @@
+#include "IndexedRealSDR.h"
+
+/// \brief	Standard constructor.
+///
+IndexedReal::IndexedReal( CMACK_MSGTYPE value, int index ) : v(value), i(index)
+{
+}
+
+/// Overload the < operator for sorting.
+bool IndexedReal::operator<(const IndexedReal& other) const 
+{
+  return v < other.v;
+}
+
+

### 2013-11-06 code/st1matlab/impl/software/library/phy/src/LDPCDecV2.cpp

@@ -1,0 +1,514 @@
+#include "LDPCDecV2.h"
+#include <math.h>
+#include <stdlib.h>
+#include <stdio.h>
+
+#ifdef FAST_MATH
+  #include  "fastexp.h"
+  #include  "fastlog.h"
+
+  #define   LOG(x)  fasterlog(x)
+  #define   EXP(x)  fasterexp(x)
+#else
+  #define   LOG(x)  log(x)
+  #define   EXP(x)  exp(x)
+#endif
+
+#ifdef _WIN32
+  #define log2(x) 			(log(x)/log((MSGTYPE)2))
+#endif
+
+/// \brief	defines the maximum absolute value of L-values used
+///
+#define MAXLLR	100
+
+/// \brief	returns minimum of a and b
+///
+#define MIN(a,b)			(((a) < (b)) ? (a) : (b))
+
+/// \brief	returns maximum of a and b
+///
+#define MAX(a,b)			(((a) > (b)) ? (a) : (b))
+
+/// \brief	clips a at magnitude b
+///
+#define CLIP(a,b)			MAX(MIN(a,b), -b)
+
+/// \brief	return sign of a
+///
+#define SIGN(a)				(((a)<0) ? (-1) : (1))
+
+/// \brief	return hard decision of a
+///
+#define HD(a)				  (signbit(a))
+
+//#define FABS(a)				(((a)<0) ? (-a) : (a))
+#define FABS(a)				copysignf(a,1.0f)
+
+/// \brief	Standard constructor.
+///
+CLDPCDec::CLDPCDec(void)
+{
+	m_N						= 0;
+	m_M						= 0;
+	m_E						= 0;
+	m_g						= 0;
+	m_vardegree		= NULL;
+	m_chkdegree		= NULL;
+	m_interleaver = NULL;
+	m_v2c					= NULL;
+	m_c2v					= NULL;
+  m_tmp         = NULL;
+	m_hard				= NULL;
+  
+  m_mf          = NULL;
+  m_mb          = NULL;
+  m_sf          = NULL;
+  m_sb          = NULL;
+  m_dcmax       = 0;
+
+	m_clearmsg		= true;
+	m_method			= DECODER_SPA;
+	
+	m_decloaded		= false;
+  
+  m_corrveclen  = 0;
+  m_corrvec     = NULL;
+}
+
+/// \brief	Standard destructor.
+///
+CLDPCDec::~CLDPCDec(void)
+{
+	freeMem();
+}
+
+/// \brief  Allocate memory.
+///
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/src/mexCoherentMackenthun.cpp

@@ -1,0 +1,143 @@
+#include "mex.h"
+#include "CoherentMackenthun.h"
+#include "MatlabHelper.h"
+#include <map>
+#include <string>
+#include <sstream> 
+
+#define STRINGIZE(x) #x
+#define STRINGIZE_VALUE_OF(x) STRINGIZE(x)
+
+void Cleanup();
+
+void DisplayHelp();
+
+static std::map<std::string,PhaseEstimator*> cmap; //memory for estimators
+static std::vector<complex> y; //memory for input data
+static std::vector<complex> p; //memory for pilots
+
+#ifdef _WIN32
+#define log2(x) 			(log(x)/log((MSGTYPE)2.0))
+#define round(x)			(floor((x)+0.5))
+#endif
+
+
+/// \brief	Frees allocated memory.
+///
+void Cleanup()
+{
+  //delete all the estimators in the map.
+  std::map<std::string,PhaseEstimator*>::iterator itr;
+  for(itr = cmap.begin(); itr!=cmap.end(); ++itr) delete itr->second;
+  cmap.clear();
+  y.clear();
+  p.clear();
+}
+
+/// \brief  Displays help text
+///
+void DisplayHelp()
+{
+  mexPrintf("\nmexCoherentMackenthun MATLAB wrapper the coherent phase estimator\n");
+  mexPrintf("\n(C) 2009-2012 Robby McKilliam, ITR/UniSA\n");
+  mexPrintf("Compiled: %s %s\n\n", __DATE__, __TIME__);
+  mexPrintf("Messages are represented as %s.\n\n", STRINGIZE_VALUE_OF(MSGTYPE));
+}
+
+/// \brief  Main function for MATLAB calls.
+///
+void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[])
+{
+	
+  mexAtExit(Cleanup);	
+	
+  //construct complex gain estimator
+  if (nrhs == 5){
+    std::string name = mxArrayToString(prhs[0]);
+    std::vector<int> D = GetIndexArray(prhs[1]); //mexPrintf("D = "); for(unsigned int i = 0; i < D.size(); i++) mexPrintf("%i, ", D[i]);
+    std::vector<int> P = GetIndexArray(prhs[2]); //mexPrintf("P = "); for(unsigned int i = 0; i < P.size(); i++) mexPrintf("%i, ", P[i]);
+    std::vector<complex> pilots = GetComplexArray(prhs[3]);
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/src/mexFrequencyEstimator.cpp

@@ -1,0 +1,123 @@
+//
+//  mexFrequencyEstimator.cpp
+//  mexFrequencyEstimator
+//
+//  Created by Andre Pollok on 07/02/2013.
+//  Copyright (c) 2013 ITR, University of South Australia. All rights reserved.
+//
+
+#include "mex.h"
+#include "MatlabHelper.h"
+#include "FrequencyEstimator.h"
+#include <map>
+#include <string>
+#include <sstream> 
+
+#define STRINGIZE(x) #x
+#define STRINGIZE_VALUE_OF(x) STRINGIZE(x)
+
+void Cleanup();
+
+void DisplayHelp();
+
+static std::map<std::string,FrequencyEstimator*> freqEstMap; // memory for estimators
+static std::vector<complexT> rxSamp, txHatSamp; // memory for received and estimated transmit samples
+static std::vector<VALTYPE> tSamp; // memory for sample times
+
+
+/// \brief	Frees allocated memory.
+///
+void Cleanup()
+{
+  //delete all the estimators in the map.
+  std::map<std::string,FrequencyEstimator*>::iterator itr;
+  for(itr = freqEstMap.begin(); itr!=freqEstMap.end(); ++itr) delete itr->second;
+  freqEstMap.clear();
+  rxSamp.clear();
+  txHatSamp.clear();
+  tSamp.clear();
+}
+
+/// \brief  Displays help text
+///
+void DisplayHelp()
+{
+  mexPrintf("\nmexFrequencyEstimator MATLAB wrapper the coherent phase estimator\n");
+  mexPrintf("\n(C) 2013 Andre Pollok, ITR/UniSA\n");
+  mexPrintf("Compiled: %s %s\n\n", __DATE__, __TIME__);
+  mexPrintf("Messages are represented as %s.\n\n", STRINGIZE_VALUE_OF(MSGTYPE));
+}
+
+/// \brief  Main function for MATLAB calls.
+///
+void mexFunction(int nlhs, mxArray *plhs[], int nrhs, const mxArray *prhs[])
+{
+	
+  mexAtExit(Cleanup);	
+	
+  // construct least-squares frequency estimator
+  if (nrhs == 4){
+    std::string name = mxArrayToString(prhs[0]);
+    const unsigned int Nfft = (const unsigned int)mxGetScalar(prhs[1]);
+    const VALTYPE Ts = (const VALTYPE)mxGetScalar(prhs[2]);
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/src/mexLDPCDec.cpp

@@ -1,0 +1,246 @@
+#include "mex.h"
+#include "LDPCDec.h"
+#include <string.h>
+#include "MatlabHelper.h"
+
+#define STRINGIZE(x) #x
+#define STRINGIZE_VALUE_OF(x) STRINGIZE(x)
+
+void Cleanup();
+
+void DisplayHelp();
+
+/// \brief	defines the maximum number of LDPC instances
+///
+#define	MAXLDPC	200
+
+static CLDPCDec		**pLDPC	= NULL;
+
+/// \brief	Frees allocated memory.
+///
+void Cleanup()
+{
+	unsigned int	i;
+
+	for(i=0;i<MAXLDPC;i++)
+		delete(pLDPC[i]);
+	free(pLDPC);
+	pLDPC = NULL;
+}
+
+/// \brief  Displays help text
+///
+void DisplayHelp()
+{
+  mexPrintf("\nmexLDPCDec   MATLAB wrapper for LDPC class\n");
+  mexPrintf("This mex-file enables the use of the LDPC class from MATLAB\n");
+  mexPrintf("\n(C) 2009-2012 Gottfried Lechner, ITR/UniSA\n");
+  mexPrintf("Compiled: %s %s\n\n", __DATE__, __TIME__);
+  mexPrintf("Messages are represented as %s.\n\n", STRINGIZE_VALUE_OF(MSGTYPE));
+  mexPrintf("Use mexLDPCDec(id, command)\n\n");
+}
+
+/// \brief  Main function for MATLAB calls.
+/// \param	nlhs		number of left-hand-side arguments
+/// \param	plhs		array of left-hand-side arguments
+/// \param	nrhs		number of right-hand-side arguments
+/// \param	prhs		array of right-hand-side arguments
+///
+///	The first argument is the ID of the LDPC instance (starting from 0).
+/// This allows to use many LDPC codes in parallel. The maximum number of
+/// parallel codes is defined as MAXLDPC (default=100).
+///
+/// Depending on the second argument, different tasks are performed and the
+/// remaining arguments have a different meaning.
+/// Possible commands (second argument) and further arguments:
+/// - mexLDPCDec(ID, 'readdecoder', filename)
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/src/mexTimingEstimator.cpp

@@ -1,0 +1,134 @@
+//
+//  mexTimingEstimator.cpp
+//  TimingEstimator
+//
+//  Created by Andre Pollok on 31/05/2012.
+//  Modifed by Robby McKilliam 16/2/2012.
+//  Copyright (c) 2012 ITR, University of South Australia. All rights reserved.
+//
+
+#include "mex.h"
+#include <iostream>
+#include <complex>
+#include "MatlabHelper.h"
+#include "FinitePulse.h"
+#include "TimeOffsetEstimator.h"
+#include "FastAlgorithm.h"
+#include "TabulatedFastAlgorithm.h"
+#include "OerderMeyerMassey.h"
+#include <math.h>
+#include <string>
+#include <map>
+
+#ifndef VALTYPE
+#define VALTYPE double
+#endif
+#ifndef MSGTYPE
+#define MSGTYPE VALTYPE
+#endif
+
+using namespace std;
+using namespace TimeOffset;
+
+#define STRINGIZE(x) #x
+#define STRINGIZE_VALUE_OF(x) STRINGIZE(x)
+
+static std::map<std::string,TimeOffsetEstimator*> TimeOffsetEstimatorMap; // memory for timing estimators
+
+void Cleanup();
+void DisplayHelp();
+
+void Cleanup()
+{
+  //delete all the estimators in the map.
+  std::map<std::string,TimeOffsetEstimator*>::iterator itr2;
+  for(itr2 = TimeOffsetEstimatorMap.begin(); itr2!=TimeOffsetEstimatorMap.end(); ++itr2) delete itr2->second;
+  TimeOffsetEstimatorMap.clear();
+}
+
+/// \brief  Displays help text
+///
+void DisplayHelp()
+{
+  mexPrintf("\nmexTimeOffsetEstimator   MATLAB wrapper for TimeOffsetEstimator class\n");
+  mexPrintf("This mex-file enables the use of the TimeOffsetEstimator class from MATLAB\n");
+  mexPrintf("\n(C) 2012 Andre Pollok, ITR/UniSA\n");
+  mexPrintf("Compiled: %s %s\n\n", __DATE__, __TIME__);
+  mexPrintf("Floating point numbers are represented as %s.\n\n", STRINGIZE_VALUE_OF(VALTYPE));  
+  mexPrintf("Please refer to mexTimeOffsetEstimator_help.m for detailed help.\n\n");
+}
+
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/coherentmackenthun.m

@@ -1,0 +1,63 @@
+function [ahat, Qhat] = coherentmackenthun(rx, acqparams)
+% This function provides channel parameter offset estimates for PSK 
+% signals. The function operates on an oversampled received
+% signal and produces estimates for carrier frequency offset,
+% frequecy rate timing offset, complex signal amplitude and noise
+% variance.  Also returns a matched filtered output at symbol rate.
+%
+% INPUTS:
+%   rx              - received signal at symbol rate
+%   acqparam        - struct holding aquisition parameters
+%   	.P              - position (indices) of pilot symbols
+%     .p              - pilot symbols
+%     .D              - position (indices) of data symbols
+%     .M             - M for M-PSK, i.e size of constellation
+%
+% OUTPUTS:
+%  ahat             - complex gain estimate
+% Qhat             - value of objective function at estimator ahat (optional)
+%
+% (c) 2012 ITR-UniSA
+% Authors: Robby McKilliam
+% Created: 31 March 2012
+
+    D = acqparams.D;
+    P = acqparams.P;
+    p = acqparams.p(:);
+    M = acqparams.M;
+    L = length(D) + length(P);
+    
+    %transpose stuff if required.
+    y = rx(:);
+    
+    %some repeatedly used variables
+    w = 2*pi/M;
+    eta = exp(1i*w);
+    nu = eta - 1.0;
+    
+    %setup structures
+    phi = angle(y(D));
+    u = w*round(phi/w);
+    z = [phi - u, (1:length(D))'] ;
+    g = y(D).*exp(-1i*u);
+    %disp([ g(1), g(end) ]);
+    
+    %sort z and get permutation
+    s = sortrows(z);
+    sigma = @(k) s(mod(k-1, length(D)) + 1, 2);
+    
+    %now run the search loop
+    Y = sum(y(P).*conj(p)) + sum(g);
+    ahat = Y / L;
+    Qhat = abs(Y)^2 / L;
+    for( k = 1:(M+1)*length(D) )
+        Y = Y + nu*g(sigma(k));
+        g(sigma(k)) = eta*g(sigma(k));
+        Q = abs(Y)^2 / L;
+        if( Q > Qhat )
+            Qhat = Q;
+            ahat = Y / L;
+        end
+    end
+    
+end

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/coherentmackenthunwithdoppler.m

@@ -1,0 +1,153 @@
+function [ahat,fhat,frhat] = coherentmackenthunwithdoppler(rx, acqparams)
+% Runs the phase estimator a number of times at different Doppler
+% and Doppler rate.  Returns the estimated gain, amplitude, Doppler
+% and Doppler rate.
+    
+% INPUTS:
+%   rx              - received signal at symbol rate
+%   acqparam        - struct holding aquisition parameters
+%       .P              - position (indices) of pilot symbols
+%       .D              - position (indices) of data symbols
+%     .p              - pilot symbols
+%     .M             - M for M-PSK, i.e size of constellation
+%     .T              - symbol period
+%     .fr_max      - maximum Doppler rate
+%     .fr_min       - minimum Doppler (defaults to zero)
+%     .f_max      - maximum Doppler
+%     .f_min       - minimum Doppler
+%
+% OUTPUTS:
+%  ahat             - complex gain estimate
+%  frhat           - Doppler/frequency rate estimate 
+%
+% (c) 2012 ITR-UniSA
+% Authors: Robby McKilliam, Andre Pollok
+% Created: 21 May 2012, Robby McKilliam
+% Modified: 25 May 2012, Andre Pollok
+
+    M = acqparams.M;
+    D = acqparams.D;
+    P = acqparams.P;
+    p = acqparams.p(:);
+    L = length(D) + length(P);
+    T = acqparams.T; % symbol period
+    max_index = max([P,D]);
+    inddif = max_index - min([P,D]);
+
+    fr_min = acqparams.fr_min; % CAUTION: 2nd derivative of phase produces a factor 2
+    fr_max = acqparams.fr_max; % CAUTION: 2nd derivative of phase produces a factor 2
+    f_min = acqparams.f_min;
+    f_max = acqparams.f_max;
+    
+    SEARCH_OVERSAMPLE = 2.0; %adjusts how fine the search grid is
+                           %(smaller is faster, larger is more accurate)
+    
+    %number of times Mackenthun's estimator will run.  This could
+    %be tuned a bit better.
+    fr_trials = ceil(SEARCH_OVERSAMPLE*M*(T*inddif)^2*(fr_max - ...
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/coherentmackenthunwithonlydoppler.m

@@ -1,0 +1,123 @@
+function [ahat,fhat] = coherentmackenthunwithonlydoppler(rx, acqparams)
+% Runs the phase estimator a number of times at different Doppler
+% and Doppler rate.  Returns the estimated gain, amplitude, Doppler
+% and Doppler rate.
+    
+% INPUTS:
+%   rx              - received signal at symbol rate
+%   acqparam        - struct holding aquisition parameters
+%       .P              - position (indices) of pilot symbols
+%       .D              - position (indices) of data symbols
+%     .p              - pilot symbols
+%     .M             - M for M-PSK, i.e size of constellation
+%     .T              - symbol period
+%     .f_max      - maximum Doppler
+%     .f_min       - minimum Doppler
+%
+% OUTPUTS:
+%  ahat             - complex gain estimate
+%
+% (c) 2012 ITR-UniSA
+% Authors: Robby McKilliam, Andre Pollok
+% Created: 21 May 2012, Robby McKilliam
+% Modified: 25 May 2012, Andre Pollok
+
+    M = acqparams.M;
+    D = acqparams.D;
+    P = acqparams.P;
+    p = acqparams.p(:);
+    L = length(D) + length(P);
+    T = acqparams.T; % symbol period
+    max_index = max([P,D]);
+    inddif = max_index - min([P,D]);
+
+    f_min = acqparams.f_min;
+    f_max = acqparams.f_max;
+    
+    SEARCH_OVERSAMPLE = 2.0; %adjusts how fine the search grid is
+                           %(smaller is faster, larger is more accurate)
+    
+    f_trials = ceil(SEARCH_OVERSAMPLE*M*(T*inddif)*(f_max - f_min));
+    fstep = (f_max - f_min)/f_trials;
+    if( isnan(fstep) ) 
+        fstep = 1;
+    end
+    
+    %symbol times vector
+    ts = T*[P,D];
+     
+    %some repeatedly used variables
+    w = 2*pi/M;
+    eta = exp(1i*w);
+    nu = eta - 1.0;
+    
+    bestQ = 0;
+    fhatc = f_min;
+    ahatc = 0;
+    for f = f_min:fstep:f_max
+            
+            %compute the Doppler rate shifted recieved signal.
+            y = zeros(1,max_index);
+            y([P,D]) = rx([P,D]) .* exp(-2i*pi*(f*ts));
+            
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/mexTimingEstimator_test.m

@@ -1,0 +1,130 @@
+addpath ../bin
+addpath ..
+addpath ./timeoffsettests/oldtimingestimator
+
+
+%local variables
+T = 1/10e3; %symbol period, 10KHz symbol rate
+p = 1;
+q = 4;
+OS = q/p;
+Ts = T/OS;
+taumin = 0;
+taumax = 30e-3; %maximum 10 millisconds time of flight 
+M = 4; %QPSK
+
+% root raised cosine
+rolloff = 0.5;
+Nsym = 10; % filter order in symbols
+duration = Nsym*T; % duration of transmit pulse shape
+g = @(t) rrcpulse(t,rolloff,T,OS) / sqrt(Ts/OS) .* (abs(t) <= duration/2);
+
+L = 100; %number of symbols
+absP = 50; %number of pilots
+P = 1:absP; %pilot indices
+D = absP+1:L; %data indices
+
+a0 = 1;
+
+%symbols
+tau0 = 5.111e-3;
+sigma2 = 0.00000001/Ts;
+N = ceil((T*L+taumax)/Ts);
+
+%struct with the modulation parameters and estimator parameters    
+acqparams.T = T; %symbol period, 10KHz symbol rate
+acqparams.OS =  OS;
+acqparams.taumin = taumin;
+acqparams.taumax = taumax; %maximum 10 millisconds time of flight 
+acqparams.M = M; %QPSK
+acqparams.pshape = g; %transmission pulse
+acqparams.P = P; %pilot indices
+acqparams.D = D; %data indices
+acqparams.g = g;
+acqparams.Tpshape = duration;
+
+%generate random psk symbols
+s = exp(2*pi*1i*(randi(M,[1,L])-1)/M);
+%set the pilots
+pilots = s(P);
+acqparams.p = pilots;
+acqparams.pilots = pilots;
+
+%generate recieved signal.
+recieved = @(t) a0*transmitted(t - tau0, s, acqparams);
+rx = recieved(Ts*(1:N)) + sqrt(sigma2)/sqrt(2)*(randn(1,N) + ...
+                                                  1i*randn(1,N));
+
+%figure(3); plot(Ts*(1:N), rx); hold on;
+%plot(Ts*(1:N), recieved(Ts*(1:N)), 'r');
+
+
+%%%%%%%%%%%%% Construction and deletion tests %%%%%%%%%%%%%%%%
+iters = 50;
+
+for n = 1:iters
+
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/readldpccode.m

@@ -1,0 +1,25 @@
+function code = readldpccode(filename)
+%readldpccode
+%
+%   reads LDPC struct from a file
+%
+%   code = readldpccode(filename)
+%
+%   code     ... ldpc code struct
+%   filename ... file to read from
+%
+%   (c) 2006 Gottfried Lechner
+
+fid = fopen(filename,'r');
+
+N = fscanf(fid, '%i', 1);
+M = fscanf(fid, '%i', 1);
+E = fscanf(fid, '%i', 1);
+
+code.vardegree   = fscanf(fid, '%i', N)';
+code.chkdegree   = fscanf(fid, '%i', M)';
+code.interleaver = fscanf(fid, '%i', E)';
+
+code.interleaver = code.interleaver + 1;
+
+fclose(fid);

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/test_QAMModulator.m

@@ -1,0 +1,284 @@
+function success = test_QAMModulator()
+
+success = false;
+
+addpath ../bin
+
+rng('default');
+rng(0);
+
+% display the module under test
+fullpath = which('mexQAMModulator');
+if isempty(fullpath)
+  disp('could not find mexQAMModulator');
+  return;
+end
+fprintf('testing %s\n', fullpath);
+
+% use QAM constellation with 4 symbols
+Constellation = [-1-1i -1+1i 1-1i 1+1i]/sqrt(2);
+bps           = log2(length(Constellation));
+tolerance     = 1e-6;
+N             = 1e3;
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% MODULATE TEST
+%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% initialise modulator
+mexQAMModulator(0, 'init', 1, double(Constellation));
+
+% test every symbol individually
+for s=1:length(Constellation)
+  bits  = de2bi(s-1, bps, 'left-msb');
+  x     = mexQAMModulator(0, 'modulate', bits);
+  if abs(x-Constellation(s))>tolerance
+    disp('incorrect results when testing modulator using doubles');
+    return;
+  end
+end
+
+% initialise longer modulator
+mexQAMModulator(0, 'init', N, double(Constellation));
+
+% generate random data
+bits  = randi(2,1,N*bps)-1;
+
+% modulate in Matlab
+idx   = (2.^(0:(bps-1))) * flipud(reshape(bits, bps, N)) + 1;
+yref  = Constellation(idx);
+
+% modulate
+x     = mexQAMModulator(0, 'modulate', double(bits));
+if length(x)~=N
+  disp('modulator returned wrong length using doubles');
+  return;
+end
+if ~strcmp(class(x),'double')
+  disp('modulator returned wrong type using doubles');
+  return;  
+end
+if max(abs(x-yref))>tolerance
+  disp('incorrect results when running modulator using doubles');
+  return;    
+end
+
+
+
+% initialise modulator
+mexQAMModulator(0, 'init', 1, single(Constellation));
+
+% test every symbol individually
+for s=1:length(Constellation)
+  bits  = de2bi(s-1, bps, 'left-msb');
+  x     = mexQAMModulator(0, 'modulate', bits);
+  if abs(x-Constellation(s))>tolerance
+    disp('incorrect results when testing modulator using singles');
+    return;
+  end
+end
+
+% initialise longer modulator
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/timeoffsettests/NaiveTest.cpp

@@ -1,0 +1,118 @@
+/* 
+ * File:   TimeOffsetEstimatorTest.cpp
+ * Author: Robby McKilliam
+ *
+ * Created on 24/01/2013, 1:32:09 PM
+ */
+
+#include <cstdlib>
+#include <iostream>
+
+#include "mex.h"
+#include "TimeOffsetEstimator.h"
+#include "FinitePulse.h"
+
+using namespace TimeOffset;
+
+static VALTYPE tol = 1e-6;
+static VALTYPE T = 1.0; //symbol period
+static unsigned int p = 1;
+static unsigned int q = 5;
+static unsigned int c = 5; //new c, its will divide T
+static VALTYPE Ts = p * T / q; //sample period
+static VALTYPE taumin = 10.0;
+static VALTYPE taumax = 30.0;
+static VALTYPE step = 5.0;
+static unsigned int L = 10;
+static unsigned int N = (int) ceil((T * L + taumax) / Ts);
+static int numzeros = 2;
+static vector<int> P;
+static vector<int> D;
+static vector<complex<VALTYPE> > pilots;
+static vector<complex<VALTYPE> > r;
+static unsigned int M = 4; //QPSK
+static int numpilots = 4;
+
+template <class Pulse> void NaiveDotr(Naive<Pulse>& est) {
+    std::cout << "NaiveDotr ... ";
+    bool pass = true;
+    pass &= std::abs(est.dotr(taumin) - complex<VALTYPE>(1.3499373963348864, 4.563483625803526)) < tol;
+    pass &= std::abs(est.dotr(taumin+step) - complex<VALTYPE>(1.6496241036561894, -4.463905699456121)) < tol;
+    pass &= std::abs(est.dotr(taumin+2*step) - complex<VALTYPE>(-3.9931090337278663, 2.5889754772421516)) < tol;
+    pass &= std::abs(est.dotr(taumin+3*step) - complex<VALTYPE>(4.748483513451538, 0.31562335065586683)) < tol;
+    pass &= std::abs(est.dotr(taumin+4*step) - complex<VALTYPE>(-3.6153254669352815, -3.0946947418331097)) < tol;
+    if (pass) std::cout << "PASS" << std::endl;
+    else std::cout << "FAIL" << std::endl;
+}
+
+template <class Pulse> void NaiveY(Naive<Pulse>& est) {
+    std::cout << "NaiveY ... ";
+    bool pass = true;
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/timeoffsettests/OerderMeyerMasseyTest.cpp

@@ -1,0 +1,215 @@
+/* 
+ * File:   OerderMeyerMasseyTest.cpp
+ * Author: Robby McKilliam
+ *
+ * Created on 20/02/2013, 4:55:35 PM
+ */
+
+#include "mex.h"
+#include <cstdlib>
+#include <iostream>
+#include <stdlib.h>
+#include <complex>
+#include "FinitePulse.h"
+#include "Util.h"
+#include "OerderMeyerMassey.h"
+#include "FastAlgorithm.h"
+
+#ifndef VALTYPE
+#define VALTYPE VALTYPE
+#endif
+#ifndef MSGTYPE
+#define MSGTYPE VALTYPE
+#endif
+
+using namespace std;
+using namespace TimeOffset;
+
+
+static VALTYPE tol = 1e-2;
+static VALTYPE T = 1.0; //symbol period
+static unsigned int p = 3;
+static unsigned int q = 17;
+static unsigned int c = 5; //new c, its will divide T
+static VALTYPE Ts = p * T / q; //sample period
+static VALTYPE Delta = T/c;
+static VALTYPE taumin = 10.0;
+static VALTYPE taumax = 30.0;
+static unsigned int L = 100;
+static unsigned int N = (int) ceil((T * L + taumax) / Ts);
+static int numzeros = 2;
+static vector<int> P;
+static vector<int> D;
+static vector<complex<VALTYPE> > pilots;
+static vector<complex<VALTYPE> > s;
+static vector<complex<VALTYPE> > r;
+static unsigned int M = 4; //QPSK
+static int numpilots = 30;
+
+    //transmitted signal
+complex<VALTYPE> x(vector<complex<VALTYPE> >&s, TruncatedRootRaisedCosine& tx, VALTYPE T, int L, VALTYPE t) {
+      int mini = max(1, (int)ceil((t - tx.tmax())/T));
+      int maxi = min(L, (int)floor((t - tx.tmin())/T));
+      complex<VALTYPE> sum(0,0);
+      for(int i = mini; i <= maxi; i++) sum += s[i-1] * tx.pulse(t - i*T);
+      return sum;
+};
+
+void testwithstore() {
+  //generator.seed(10);
+  const VALTYPE benchtime = 10.0; //number of seconds to run each benchmark for.
+  const VALTYPE tol = 1e-2;
+
+  const unsigned int M = 4;
+  const unsigned int c = 4;
+  const VALTYPE T = 1.0; //symbol period
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/timeoffsettests/benchmark.cpp

@@ -1,0 +1,102 @@
+/* 
+ * File:   montecarlo.cpp
+ * Author: Robby McKilliam
+ */
+
+#include <stdlib.h>
+#include <iostream>
+#include <vector>
+#include <math.h>
+#include "Util.h"
+#include "FinitePulse.h"
+#include "TimeOffsetEstimator.h"
+#include "Recursive.h"
+#include "Direct.h"
+#include "OerderMeyerMassey.h"
+#include <random>
+#include <ctime>
+
+using namespace std;
+using namespace TimeOffset;
+
+int main(int argc, char** argv) {
+
+  default_random_engine generator; 
+  const double benchtime = 20.0; //number of seconds to run each benchmark for.
+
+  const double iters = 4000;
+  const unsigned int M = 4;
+  const unsigned int c = 4;
+  const double T = 1.0; //symbol period
+  const unsigned int p = 1;
+  const unsigned int q = 4;
+  const double Ts = p*T/q; //sample period
+  const double taumin = 40.0;
+  const double taumax = 50.0;
+  const double tau0 = 46.1;
+  const complex<double> a0 = polar<double>(1,0.1*pi);
+  //const vector<int> Ls = {11,15,20,30,40,50,60,80,100,120,140,160,180,200,250,300,400,500,600,800,1000,1200,1500,2000,2600,3500};
+  const vector<int> Ls = {3990};
+ 
+
+  //SNRs
+  double SNRdB = 10.0;
+  double SNR = pow(10.0, SNRdB/10.0);
+  normal_distribution<double> noise(0.0,T/Ts/SNR/2);
+  uniform_int_distribution<int> unifM(1, M); //for generating M-PSK symbols
+
+  for(auto L : Ls) {
+
+    const unsigned int numpilots = L/10; //about 10% pilots 
+    //symbol position setup
+    vector<int> P;
+    for(int i = 0; i < numpilots; i++) P.push_back(i); //pilots at the front
+    vector<int> D;
+    for(int i = numpilots; i < L; i++) D.push_back(i); //data at the back
+    vector<complex<double>> s;
+    for (int m = 1; m <= L; m++) s.push_back(polar<double>(1.0, 2 * pi * unifM(generator) / M));
+    vector<complex<double>> pilots;
+    for (int m = 0; m < numpilots; m++) pilots.push_back(s[m]);
+
+    //the transmit pulse
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/timeoffsettests/oldtimingestimator/MatlabHelper.cpp

@@ -1,0 +1,260 @@
+//
+//  MatlabHelper.cpp
+//  CoherentMackenthun
+//
+//  Created by Gottfried Lechner on 12/04/12.
+//  Modified by Robby McKilliam on 5/04/12
+//  Copyright (c) 2012 ITR, University of South Australia. All rights reserved.
+//
+
+#include "MatlabHelper.h"
+
+#ifdef _WIN32
+#define round(x)			(floor((x)+0.5))
+#endif
+
+/// \brief  Copy a matlab float array into a vector of MSGTYPE (i.e. doubles or floats).  Ignores imaginary part.
+///
+std::vector<MSGTYPE> GetFloatArray(const mxArray *src)
+{
+  unsigned int N = mxGetNumberOfElements(src);
+  std::vector<MSGTYPE> ret(N);
+  
+  double  *dr;
+  float   *fr;
+  
+  switch(mxGetClassID(src))
+  {
+    case mxDOUBLE_CLASS:
+      dr = (double*) mxGetData(src);
+      for(unsigned int n = 0; n < N; n++) ret[n] = (MSGTYPE)dr[n];
+      break;
+    case mxSINGLE_CLASS:
+      fr = (float*) mxGetData(src);
+      for(unsigned int n = 0; n < N; n++) ret[n] = (MSGTYPE)fr[n];
+      break;
+  }
+  return ret;
+}
+
+//Copies the values in a matlab mxArray* into a std::vector<complex>.  Will resize the std::vector
+//if neccessary. 
+void CopyMatlabComplexToVector(const mxArray *src, std::vector<complex>& v){
+  unsigned int N = mxGetNumberOfElements(src);
+  if( v.size() != N ) v.resize(N, complex(0,0));
+  double* r = (double*) mxGetData(src);
+  double* i = (double*) mxGetImagData(src);
+  if(i != NULL) for(unsigned int n = 0; n < N; n++) v[n] = complex(r[n],i[n]); 
+  else for(unsigned int n = 0; n < N; n++) v[n] = complex(r[n],0.0);
+}
+
+//Copies the real part of a matlab mxArray* into a std::vector<MSGTYPE>.  Will resize the std::vector
+//if neccessary. 
+void CopyMatlabComplexToVector(const mxArray *src, std::vector<MSGTYPE>& v){
+  unsigned int N = mxGetNumberOfElements(src);
+  if( v.size() != N ) v.resize(N, 0.0);
+  double* r = (double*) mxGetData(src); 
[…]

### 2013-11-06 code/st1matlab/impl/software/library/phy/test/timeoffsettests/oldtimingestimator/mexOldTimeEstimator.cpp

@@ -1,0 +1,265 @@
+//
+//  mexTimingEstimator.cpp
+//  TimingEstimator
+//
+//  Created by Andre Pollok on 31/05/2012.
+//  Copyright (c) 2012 ITR, University of South Australia. All rights reserved.
+//
+
+#include "mex.h"
+#include <iostream>
+#include <complex>
+#include "MatlabHelper.h"
+#include "TimingEstimator.h"
+#include <math.h>
+#include <string>
+#include <map>
+
+using namespace std;
+
+#define STRINGIZE(x) #x
+#define STRINGIZE_VALUE_OF(x) STRINGIZE(x)
+
+static std::map<std::string,PulseShape*> PulseShapeMap; // memory for pulse shapes
+static std::map<std::string,TimingEstimator*> TimingEstimatorMap; // memory for timing estimators
+
+void Cleanup();
+void DisplayHelp();
+
+void Cleanup()
+{
+  //delete all the pulse shapes in the map.
+  std::map<std::string,PulseShape*>::iterator itr;
+  for(itr = PulseShapeMap.begin(); itr!=PulseShapeMap.end(); ++itr) delete itr->second;
+  PulseShapeMap.clear();
+
+  //delete all the estimators in the map.
+  std::map<std::string,TimingEstimator*>::iterator itr2;
+  for(itr2 = TimingEstimatorMap.begin(); itr2!=TimingEstimatorMap.end(); ++itr2) delete itr2->second;
+  TimingEstimatorMap.clear();
+}
+
+/// \brief  Displays help text
+///
+void DisplayHelp()
+{
+  mexPrintf("\nmexTimingEstimator   MATLAB wrapper for TimingEstimator class\n");
+  mexPrintf("This mex-file enables the use of the TimingEstimator class from MATLAB\n");
+  mexPrintf("\n(C) 2012 Andre Pollok, ITR/UniSA\n");
+  mexPrintf("Compiled: %s %s\n\n", __DATE__, __TIME__);
+  mexPrintf("Floating point numbers are represented as %s.\n\n", STRINGIZE_VALUE_OF(VALTYPE));  
+  mexPrintf("Please refer to mexTimingEstimator_help.m for detailed help.\n\n");
+}
+
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/RemovePilots.m

@@ -1,0 +1,40 @@
+% This function simply removes an index of pilots from the data
+%--------------------------------------------------------------------------
+% RemovePilots.m - This function simply removes an index of pilots 
+% from the data. It also check if pilots are expected in data portion 
+% of packet and if not, is transparent and returns output identical to
+% input.
+% 
+% [ dataout] = RemovePilots( datain,userparam,modparam )
+% Inputs:
+%   datain      - Data sequence including pilots
+%   userparam   - User Information containing pilot index
+%   modparam    - Modem Information (e.g. AIS mode etc...)
+% Outputs:
+%   dataout     - Data sequence with pilots removed
+% Comments:
+%      
+%--------------------------------------------------------------------------
+% Author:   M. Lavenant, ITR
+% Project : ASRP
+%--------------------------------------------------------------------------
+% Copyright 2011 : 
+% Institute for Telecommunications Research
+% University of South Australia
+%--------------------------------------------------------------------------
+% Contains ITR Background IP
+%--------------------------------------------------------------------------
+function [ dataout] = RemovePilots( datain,~,modparam )
+
+dataout=datain;
+if modparam.AISFormat
+    if modparam.N_pilots >1 %more than just default start train/flag
+        startoffset=modparam.TrainL+modparam.FlagL;
+        dataout( modparam.pilotindex(startoffset+1:end)-startoffset)=[];
+    end
+else
+    if modparam.N_pilots >0
+        dataout( modparam.pilotindex)=[];
+    end
+end
+

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/ST2DLacquisition_init.m

@@ -1,0 +1,135 @@
+function ST2DLacquisition_init(acqparams)
+% This function initialises the ST2DL acquisition module and must be called
+% prior to running ST2DLacquisition.m. Specifically, this function
+% constructs objects for the frequency, time and phase offset estimators.
+%
+% Comment: It is possible to only use a subset of pilot and data symbols in
+%          the timing and phase estimators. For example, this can be useful
+%          to meet constraints on the computational complexity.
+%
+%   ST2DLacquisition_init(acqparams)
+%
+% INPUTS:
+%   acqparams       - struct holding aquisition parameters
+%       .T              - symbol period
+%       .OS             - oversampling factor
+%       .P              - position (indices) of pilot symbols
+%       .p              - pilot symbols
+%       .D              - position (indices) of data symbols
+%       .PTime          - position (indices) of pilot symbols used by
+%                         timing estimator (equal to or a subset of .P)
+%       .pTime          - pilot symbols used by timing estimator (equal to 
+%                         or a subset of .p)
+%       .DTime          - position (indices) of data symbols used by timing
+%                         estimator (equal to or a subset of .D)
+%       .PPhase         - position (indices) of pilot symbols used by phase
+%                         estimator (equal to or a subset of .P)
+%       .pPhase         - pilot symbols used by phase estimator (equal to
+%                         or a subset of .p)
+%       .DPhase         - position (indices) of data symbols used by phase
+%                         estimator (equal to or a subset of .D)
+%       .M              - M for M-PSK, i.e. size of constellation
+%       .rolloff        - roll-off factor of root raised cosine pulse
+%       .Tpshape        - duration of the transmit pulse shape in seconds
+%       .timingEstimatorName
+%                       - string identifying the timing estimator
+%       .taumin         - minimum time offset
+%       .taumax         - maximum time offset
+%       .phaseEstimatorName
+%                       - string identifying the phase estimator
+%       .fr_max         - maximum Doppler rate
+%       .fr_min         - minimum Doppler rate
+%       .f_max          - range of residual Doppler after initial frequency
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/EsNoest_coder.cpp

@@ -1,0 +1,145 @@
+/*
+ * EsNoest_coder.cpp
+ *
+ * Code generation for function 'EsNoest_coder'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "EsNoest_coder.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "abs.h"
+#include "mrdivide.h"
+#include "std.h"
+#include "mean.h"
+#include "exp.h"
+#include "mod.h"
+#include "round.h"
+#include "angle.h"
+#include "ST2DL_Rx_rtwutil.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+
+/*
+ * function [EsNohat] = EsNoest_coder(MFoutput,M,D,P,p)
+ */
+real_T EsNoest_coder(const creal_T MFoutput[3990], real_T M, const real_T D[3960],
+                     const real_T P[30], const creal_T p[30])
+{
+  static creal_T symbhat[3990];
+  int32_T i9;
+  static creal_T b_MFoutput[3960];
+  real_T symbsectorhat[3960];
+  real_T b_symbsectorhat[3960];
+  real_T c_symbsectorhat[3960];
+  static const creal_T dc2 = { 0.0, 6.2831853071795862 };
+
+  creal_T y;
+  creal_T dcv0[3960];
+  real_T symbhat_re;
+  real_T symbhat_im;
+
+  /*  Joint pilot and data based Es/No estimator. */
+  /* ------------------------------------------------------------- */
+  /*  EsNoest.m  */
+  /*  */
+  /*  [EsNohat] = EsNoest(MFoutput,AcqParams) */
+  /*  */
+  /*  This function etimates the Es/No from the complex matched filter (MF) */
+  /*  output. First, hard decisions for the data symbols are obtained from the */
+  /*  MF output. After inserting the known pilot symbols, the symbol stream is */
+  /*  used to strip the modulation of the MF output. The Es/No is estimated */
+  /*  from the resulting signal. */
+  /*   */
+  /*  Inputs: */
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/EsNoest_coder.h

@@ -1,0 +1,34 @@
+/*
+ * EsNoest_coder.h
+ *
+ * Code generation for function 'EsNoest_coder'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __ESNOEST_CODER_H__
+#define __ESNOEST_CODER_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern real_T EsNoest_coder(const creal_T MFoutput[3990], real_T M, const real_T D[3960], const real_T P[30], const creal_T p[30]);
+#endif
+/* End of code generation (EsNoest_coder.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/ST2DL_Rx_emxutil.h

@@ -1,0 +1,46 @@
+/*
+ * ST2DL_Rx_emxutil.h
+ *
+ * Code generation for function 'ST2DL_Rx_emxutil'
+ *
+ * C source code generated on: Fri Nov 23 11:11:00 2012
+ *
+ */
+
+#ifndef __ST2DL_RX_EMXUTIL_H__
+#define __ST2DL_RX_EMXUTIL_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void b_emxInit_creal_T(emxArray_creal_T **pEmxArray, int32_T numDimensions);
+extern void b_emxInit_int32_T(emxArray_int32_T **pEmxArray, int32_T numDimensions);
+extern void b_emxInit_real_T(emxArray_real_T **pEmxArray, int32_T numDimensions);
+extern void c_emxInit_creal_T(emxArray_creal_T **pEmxArray, int32_T numDimensions);
+extern void emxEnsureCapacity(emxArray__common *emxArray, int32_T oldNumel, int32_T elementSize);
+extern void emxFree_boolean_T(emxArray_boolean_T **pEmxArray);
+extern void emxFree_creal_T(emxArray_creal_T **pEmxArray);
+extern void emxFree_int32_T(emxArray_int32_T **pEmxArray);
+extern void emxFree_real_T(emxArray_real_T **pEmxArray);
+extern void emxInit_boolean_T(emxArray_boolean_T **pEmxArray, int32_T numDimensions);
+extern void emxInit_creal_T(emxArray_creal_T **pEmxArray, int32_T numDimensions);
+extern void emxInit_int32_T(emxArray_int32_T **pEmxArray, int32_T numDimensions);
+extern void emxInit_real_T(emxArray_real_T **pEmxArray, int32_T numDimensions);
+#endif
+/* End of code generation (ST2DL_Rx_emxutil.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/ST2DL_Rx_initialize.cpp

@@ -1,0 +1,34 @@
+/*
+ * ST2DL_Rx_initialize.cpp
+ *
+ * Code generation for function 'ST2DL_Rx_initialize'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "EsNoest_coder.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "ST2DL_Rx_initialize.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+void ST2DL_Rx_initialize(void)
+{
+  rt_InitInfAndNaN(8U);
+}
+
+/* End of code generation (ST2DL_Rx_initialize.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/ST2DL_Rx_terminate.cpp

@@ -1,0 +1,34 @@
+/*
+ * ST2DL_Rx_terminate.cpp
+ *
+ * Code generation for function 'ST2DL_Rx_terminate'
+ *
+ * C source code generated on: Fri Nov 23 11:11:00 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "EsNoest_coder.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "ST2DL_Rx_terminate.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+void ST2DL_Rx_terminate(void)
+{
+  /* (no terminate code required) */
+}
+
+/* End of code generation (ST2DL_Rx_terminate.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/abs.cpp

@@ -1,0 +1,74 @@
+/*
+ * abs.cpp
+ *
+ * Code generation for function 'abs'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "EsNoest_coder.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "abs.h"
+#include "ST2DL_Rx_emxutil.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+
+/*
+ *
+ */
+real_T b_abs(const creal_T x)
+{
+  real_T y;
+  real_T a;
+  a = fabs(x.re);
+  y = fabs(x.im);
+  if (a < y) {
+    a /= y;
+    y *= sqrt(a * a + 1.0);
+  } else if (a > y) {
+    y /= a;
+    y = sqrt(y * y + 1.0) * a;
+  } else if (rtIsNaN(y)) {
+  } else {
+    y = a * 1.4142135623730951;
+  }
+
+  return y;
+}
+
+/*
+ *
+ */
+void c_abs(const emxArray_real_T *x, emxArray_real_T *y)
+{
+  uint32_T uv2[2];
+  int32_T k;
+  for (k = 0; k < 2; k++) {
+    uv2[k] = (uint32_T)x->size[k];
+  }
+
+  k = y->size[0] * y->size[1];
+  y->size[0] = 1;
+  y->size[1] = (int32_T)uv2[1];
+  emxEnsureCapacity((emxArray__common *)y, k, (int32_T)sizeof(real_T));
+  for (k = 0; k <= x->size[1] - 1; k++) {
+    y->data[k] = fabs(x->data[k]);
+  }
+}
+
+/* End of code generation (abs.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/dcblockST2DL.h

@@ -1,0 +1,34 @@
+/*
+ * dcblockST2DL.h
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __DCBLOCKST2DL_H__
+#define __DCBLOCKST2DL_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void dcblockST2DL(const creal_T RxSamp[16000], creal_T RxSampFilt[16000]);
+#endif
+/* End of code generation (dcblockST2DL.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/exp.cpp

@@ -1,0 +1,58 @@
+/*
+ * exp.cpp
+ *
+ * Code generation for function 'exp'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "EsNoest_coder.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "exp.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+static void c_exp(creal_T x[3960]);
+
+/* Function Definitions */
+
+/*
+ *
+ */
+static void c_exp(creal_T x[3960])
+{
+  int32_T k;
+  real_T r;
+  real_T x_im;
+  real_T b_x_im;
+  for (k = 0; k < 3960; k++) {
+    r = exp(x[k].re / 2.0);
+    x_im = x[k].im;
+    b_x_im = x[k].im;
+    x[k].re = r * (r * cos(x_im));
+    x[k].im = r * (r * sin(b_x_im));
+  }
+}
+
+/*
+ *
+ */
+void b_exp(const creal_T x[3960], creal_T b_x[3960])
+{
+  memcpy((void *)&b_x[0], (void *)&x[0], 3960U * sizeof(creal_T));
+  c_exp(b_x);
+}
+
+/* End of code generation (exp.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/exp.h

@@ -1,0 +1,34 @@
+/*
+ * exp.h
+ *
+ * Code generation for function 'exp'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __EXP_H__
+#define __EXP_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void b_exp(const creal_T x[3960], creal_T b_x[3960]);
+#endif
+/* End of code generation (exp.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/freqest_periodogram_coder.h

@@ -1,0 +1,35 @@
+/*
+ * freqest_periodogram_coder.h
+ *
+ * Code generation for function 'freqest_periodogram_coder'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __FREQEST_PERIODOGRAM_CODER_H__
+#define __FREQEST_PERIODOGRAM_CODER_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern creal_T eml_div(const creal_T x, real_T y);
+extern void freqest_periodogram_coder(const creal_T RxSamp[16000], const real_T tSamp[16000], real_T T, real_T OS, real_T Lsin, real_T taumin, real_T taumax, real_T *fhat, real_T *phihat);
+#endif
+/* End of code generation (freqest_periodogram_coder.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/ipermute.h

@@ -1,0 +1,34 @@
+/*
+ * ipermute.h
+ *
+ * Code generation for function 'ipermute'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __IPERMUTE_H__
+#define __IPERMUTE_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void ipermute(const emxArray_creal_T *b, emxArray_creal_T *a);
+#endif
+/* End of code generation (ipermute.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/mod.h

@@ -1,0 +1,34 @@
+/*
+ * mod.h
+ *
+ * Code generation for function 'mod'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __MOD_H__
+#define __MOD_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void b_mod(const real_T x[3960], real_T y, real_T r[3960]);
+#endif
+/* End of code generation (mod.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/rdivide.h

@@ -1,0 +1,34 @@
+/*
+ * rdivide.h
+ *
+ * Code generation for function 'rdivide'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __RDIVIDE_H__
+#define __RDIVIDE_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void rdivide(const emxArray_real_T *x, const emxArray_real_T *y, emxArray_real_T *z);
+#endif
+/* End of code generation (rdivide.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/round.h

@@ -1,0 +1,34 @@
+/*
+ * round.h
+ *
+ * Code generation for function 'round'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __ROUND_H__
+#define __ROUND_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void b_round(real_T x[3960]);
+#endif
+/* End of code generation (round.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/rtGetInf.cpp

@@ -1,0 +1,139 @@
+/*
+ * rtGetInf.cpp
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Fri Nov 23 11:11:00 2012
+ *
+ */
+
+/*
+ * Abstract:
+ *       MATLAB for code generation function to initialize non-finite, Inf and MinusInf
+ */
+#include "rtGetInf.h"
+#define NumBitsPerChar	8U
+
+/* Function: rtGetInf ==================================================
+ * Abstract:
+ * Initialize rtInf needed by the generated code.
+ * Inf is initialized as non-signaling. Assumes IEEE.
+ */
+real_T rtGetInf(void)
+{
+  size_t bitsPerReal = sizeof(real_T) * (NumBitsPerChar);
+  real_T inf = 0.0;
+  if (bitsPerReal == 32U) {
+    inf = rtGetInfF();
+  } else {
+    uint16_T one = 1U;
+    enum {
+      LittleEndian,
+      BigEndian
+    } machByteOrder = (*((uint8_T *) &one) == 1U) ? LittleEndian : BigEndian;
+    switch (machByteOrder) {
+     case LittleEndian:
+      {
+        union {
+          LittleEndianIEEEDouble bitVal;
+          real_T fltVal;
+        } tmpVal;
+
+        tmpVal.bitVal.words.wordH = 0x7FF00000U;
+        tmpVal.bitVal.words.wordL = 0x00000000U;
+        inf = tmpVal.fltVal;
+        break;
+      }
+
+     case BigEndian:
+      {
+        union {
+          BigEndianIEEEDouble bitVal;
+          real_T fltVal;
+        } tmpVal;
+
+        tmpVal.bitVal.words.wordH = 0x7FF00000U;
+        tmpVal.bitVal.words.wordL = 0x00000000U;
+        inf = tmpVal.fltVal;
+        break;
+      }
+    }
+  }
+
+  return inf;
+}
+
+/* Function: rtGetInfF ==================================================
+ * Abstract:
+ * Initialize rtInfF needed by the generated code.
+ * Inf is initialized as non-signaling. Assumes IEEE.
+ */
+real32_T rtGetInfF(void)
+{
+  IEEESingle infF;
+  infF.wordL.wordLuint = 0x7F800000U;
+  return infF.wordL.wordLreal;
+}
+
+/* Function: rtGetMinusInf ==================================================
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/rtGetNaN.cpp

@@ -1,0 +1,96 @@
+/*
+ * rtGetNaN.cpp
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Fri Nov 23 11:11:00 2012
+ *
+ */
+
+/*
+ * Abstract:
+ *       MATLAB for code generation function to initialize non-finite, NaN
+ */
+#include "rtGetNaN.h"
+#define NumBitsPerChar	8U
+
+/* Function: rtGetNaN ==================================================
+ * Abstract:
+ * Initialize rtNaN needed by the generated code.
+ * NaN is initialized as non-signaling. Assumes IEEE.
+ */
+real_T rtGetNaN(void)
+{
+  size_t bitsPerReal = sizeof(real_T) * (NumBitsPerChar);
+  real_T nan = 0.0;
+  if (bitsPerReal == 32U) {
+    nan = rtGetNaNF();
+  } else {
+  uint16_T one = 1U;
+  enum {
+    LittleEndian,
+    BigEndian
+  } machByteOrder = (*((uint8_T *) &one) == 1U) ? LittleEndian : BigEndian;
+  switch (machByteOrder) {
+   case LittleEndian:
+    {
+      union {
+        LittleEndianIEEEDouble bitVal;
+        real_T fltVal;
+      } tmpVal;
+
+      tmpVal.bitVal.words.wordH = 0xFFF80000U;
+      tmpVal.bitVal.words.wordL = 0x00000000U;
+      nan = tmpVal.fltVal;
+      break;
+    }
+
+   case BigEndian:
+    {
+        union {
+          BigEndianIEEEDouble bitVal;
+          real_T fltVal;
+        } tmpVal;
+
+        tmpVal.bitVal.words.wordH = 0x7FFFFFFFU;
+        tmpVal.bitVal.words.wordL = 0xFFFFFFFFU;
+        nan = tmpVal.fltVal;
+        break;
+      }
+    }
+  }
+
+  return nan;
+}
+
+/* Function: rtGetNaNF ==================================================
+ * Abstract:
+ * Initialize rtNaNF needed by the generated code.
+ * NaN is initialized as non-signaling. Assumes IEEE.
+ */
+real32_T rtGetNaNF(void)
+{
+  IEEESingle nanF = { { 0 } };
+  uint16_T one = 1U;
+  enum {
+    LittleEndian,
+    BigEndian
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/ST2DL_Rx/std.h

@@ -1,0 +1,34 @@
+/*
+ * std.h
+ *
+ * Code generation for function 'std'
+ *
+ * C source code generated on: Fri Nov 23 11:10:59 2012
+ *
+ */
+
+#ifndef __STD_H__
+#define __STD_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "ST2DL_Rx_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern creal_T b_std(const creal_T varargin_1[3990]);
+#endif
+/* End of code generation (std.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/abs.h

@@ -1,0 +1,34 @@
+/*
+ * abs.h
+ *
+ * Code generation for function 'abs'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+#ifndef __ABS_H__
+#define __ABS_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "dcblockST2DL_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void b_abs(const emxArray_real_T *x, emxArray_real_T *y);
+#endif
+/* End of code generation (abs.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/dcblockST2DL.cpp

@@ -1,0 +1,86 @@
+/*
+ * dcblockST2DL.cpp
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+void dcblockST2DL(const creal_T RxSamp[16000], creal_T RxSampFilt[16000])
+{
+  creal_T dbuffer[2];
+  int32_T j;
+  int32_T k;
+
+  /*  DC blocking filter */
+  /* ------------------------------------------------------------- */
+  /*  dcblock.m  */
+  /*  */
+  /*  [RxSampFilt] = dcblockST2DL(RxSamp) */
+  /*  */
+  /*  This digital filter blocks the DC component in the received signal. It is */
+  /*  primarily intended to remove the complex exponential at DC that is */
+  /*  transmitted in the ST2 downlink waveform for frequency estimation */
+  /*  purposes. Once the frequency has been estimated, the discrete tone should */
+  /*  be removed in order to avoid bit error rate degradations. The */
+  /*  implementation uses an order-one IIR filter. */
+  /*   */
+  /*  Inputs: */
+  /*    RxSamp      - Vector containing samples of received signal */
+  /*  */
+  /*  Outputs: */
+  /*    RxSampFilt  - Vector containing filtered received samples */
+  /*  */
+  /*  Comments: For details see [J. de Freitas, "The DC Blocking Filter", Jan. */
+  /*            2007, available at */
+  /*            http://www.mathworks.com.au/matlabcentral/fileexchange/13792-the-dc-blocking-filter]  */
+  /*   */
+  /* ------------------------------------------------------------- */
+  /*  Author: A. Pollok, Institute for Telecommunications Research */
+  /*  Project: ASRP */
+  /* ------------------------------------------------------------- */
+  /*  Copyright 2012  */
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/dcblockST2DL_rtwutil.cpp

@@ -1,0 +1,74 @@
+/*
+ * dcblockST2DL_rtwutil.cpp
+ *
+ * Code generation for function 'dcblockST2DL_rtwutil'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "dcblockST2DL_rtwutil.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+real_T rt_powd_snf(real_T u0, real_T u1)
+{
+  real_T y;
+  real_T d0;
+  real_T d1;
+  if (rtIsNaN(u0) || rtIsNaN(u1)) {
+    y = rtNaN;
+  } else {
+    d0 = fabs(u0);
+    d1 = fabs(u1);
+    if (rtIsInf(u1)) {
+      if (d0 == 1.0) {
+        y = rtNaN;
+      } else if (d0 > 1.0) {
+        if (u1 > 0.0) {
+          y = rtInf;
+        } else {
+          y = 0.0;
+        }
+      } else if (u1 > 0.0) {
+        y = 0.0;
+      } else {
+        y = rtInf;
+      }
+    } else if (d1 == 0.0) {
+      y = 1.0;
+    } else if (d1 == 1.0) {
+      if (u1 > 0.0) {
+        y = u0;
+      } else {
+        y = 1.0 / u0;
+      }
+    } else if (u1 == 2.0) {
+      y = u0 * u0;
+    } else if ((u1 == 0.5) && (u0 >= 0.0)) {
+      y = sqrt(u0);
+    } else if ((u0 < 0.0) && (u1 > floor(u1))) {
+      y = rtNaN;
+    } else {
+      y = pow(u0, u1);
+    }
+  }
+
+  return y;
+}
+
+/* End of code generation (dcblockST2DL_rtwutil.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/dcblockST2DL_terminate.cpp

@@ -1,0 +1,33 @@
+/*
+ * dcblockST2DL_terminate.cpp
+ *
+ * Code generation for function 'dcblockST2DL_terminate'
+ *
+ * C source code generated on: Thu Nov  8 13:52:46 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "dcblockST2DL_terminate.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+void dcblockST2DL_terminate(void)
+{
+  /* (no terminate code required) */
+}
+
+/* End of code generation (dcblockST2DL_terminate.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/dcblockST2DL_terminate.h

@@ -1,0 +1,34 @@
+/*
+ * dcblockST2DL_terminate.h
+ *
+ * Code generation for function 'dcblockST2DL_terminate'
+ *
+ * C source code generated on: Thu Nov  8 13:52:46 2012
+ *
+ */
+
+#ifndef __DCBLOCKST2DL_TERMINATE_H__
+#define __DCBLOCKST2DL_TERMINATE_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "dcblockST2DL_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void dcblockST2DL_terminate(void);
+#endif
+/* End of code generation (dcblockST2DL_terminate.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/dcblockST2DL_types.h

@@ -1,0 +1,71 @@
+/*
+ * dcblockST2DL_types.h
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+#ifndef __DCBLOCKST2DL_TYPES_H__
+#define __DCBLOCKST2DL_TYPES_H__
+
+/* Type Definitions */
+#ifndef struct_emxArray__common
+#define struct_emxArray__common
+typedef struct emxArray__common
+{
+    void *data;
+    int32_T *size;
+    int32_T allocatedSize;
+    int32_T numDimensions;
+    boolean_T canFreeData;
+} emxArray__common;
+#endif
+#ifndef struct_emxArray_boolean_T
+#define struct_emxArray_boolean_T
+typedef struct emxArray_boolean_T
+{
+    boolean_T *data;
+    int32_T *size;
+    int32_T allocatedSize;
+    int32_T numDimensions;
+    boolean_T canFreeData;
+} emxArray_boolean_T;
+#endif
+#ifndef struct_emxArray_creal_T
+#define struct_emxArray_creal_T
+typedef struct emxArray_creal_T
+{
+    creal_T *data;
+    int32_T *size;
+    int32_T allocatedSize;
+    int32_T numDimensions;
+    boolean_T canFreeData;
+} emxArray_creal_T;
+#endif
+#ifndef struct_emxArray_int32_T
+#define struct_emxArray_int32_T
+typedef struct emxArray_int32_T
+{
+    int32_T *data;
+    int32_T *size;
+    int32_T allocatedSize;
+    int32_T numDimensions;
+    boolean_T canFreeData;
+} emxArray_int32_T;
+#endif
+#ifndef struct_emxArray_real_T
+#define struct_emxArray_real_T
+typedef struct emxArray_real_T
+{
+    real_T *data;
+    int32_T *size;
+    int32_T allocatedSize;
+    int32_T numDimensions;
+    boolean_T canFreeData;
+} emxArray_real_T;
+#endif
+
+#endif
+/* End of code generation (dcblockST2DL_types.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/freqest_periodogram_coder.cpp

@@ -1,0 +1,801 @@
+/*
+ * freqest_periodogram_coder.cpp
+ *
+ * Code generation for function 'freqest_periodogram_coder'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "dcblockST2DL_emxutil.h"
+#include "sum.h"
+#include "power.h"
+#include "ipermute.h"
+#include "dcblockST2DL_rtwutil.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+static void eml_fft(const emxArray_creal_T *x, uint32_T n, emxArray_creal_T *y);
+static real_T rt_atan2d_snf(real_T u0, real_T u1);
+
+/* Function Definitions */
+static void eml_fft(const emxArray_creal_T *x, uint32_T n, emxArray_creal_T *y)
+{
+  int32_T iDelta2;
+  int32_T nd2;
+  uint32_T ju;
+  int32_T nRows;
+  emxArray_real_T *costab1q;
+  int32_T nRowsM1;
+  int32_T ixDelta;
+  int32_T nRowsD2;
+  int32_T nRowsD4;
+  int32_T lastChan;
+  real_T e;
+  int32_T k;
+  emxArray_real_T *costab;
+  emxArray_real_T *sintab;
+  int32_T b_n;
+  int32_T n2;
+  int32_T ix;
+  int32_T chanStart;
+  int32_T i;
+  uint32_T c_n;
+  boolean_T tst;
+  real_T temp_re;
+  real_T temp_im;
+  int32_T iheight;
+  int32_T ihi;
+  real_T twid_im;
+  iDelta2 = y->size[0];
+  y->size[0] = (int32_T)n;
+  emxEnsureCapacity((emxArray__common *)y, iDelta2, (int32_T)sizeof(creal_T));
+  if (n > (uint32_T)x->size[0]) {
+    nd2 = y->size[0];
+    iDelta2 = y->size[0];
+    y->size[0] = nd2;
+    emxEnsureCapacity((emxArray__common *)y, iDelta2, (int32_T)sizeof(creal_T));
+    nd2--;
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/freqest_periodogram_coder.h

@@ -1,0 +1,34 @@
+/*
+ * freqest_periodogram_coder.h
+ *
+ * Code generation for function 'freqest_periodogram_coder'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+#ifndef __FREQEST_PERIODOGRAM_CODER_H__
+#define __FREQEST_PERIODOGRAM_CODER_H__
+/* Include files */
+#include <math.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <string.h>
+#include "rt_defines.h"
+#include "rt_nonfinite.h"
+
+#include "rtwtypes.h"
+#include "dcblockST2DL_types.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+extern void freqest_periodogram_coder(const creal_T RxSamp[16000], const real_T tSamp[16000], real_T T, real_T OS, real_T Lsin, real_T taumin, real_T taumax, real_T *fhat, real_T *phihat);
+#endif
+/* End of code generation (freqest_periodogram_coder.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/ipermute.cpp

@@ -1,0 +1,46 @@
+/*
+ * ipermute.cpp
+ *
+ * Code generation for function 'ipermute'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "ipermute.h"
+#include "dcblockST2DL_emxutil.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+void ipermute(const emxArray_creal_T *b, emxArray_creal_T *a)
+{
+  int32_T i5;
+  int32_T unnamed_idx_1;
+  i5 = a->size[0] * a->size[1];
+  a->size[0] = 1;
+  emxEnsureCapacity((emxArray__common *)a, i5, (int32_T)sizeof(creal_T));
+  unnamed_idx_1 = b->size[0];
+  i5 = a->size[0] * a->size[1];
+  a->size[1] = unnamed_idx_1;
+  emxEnsureCapacity((emxArray__common *)a, i5, (int32_T)sizeof(creal_T));
+  unnamed_idx_1 = b->size[0] - 1;
+  for (i5 = 0; i5 <= unnamed_idx_1; i5++) {
+    a->data[i5] = b->data[i5];
+  }
+}
+
+/* End of code generation (ipermute.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/mrdivide.cpp

@@ -1,0 +1,33 @@
+/*
+ * mrdivide.cpp
+ *
+ * Code generation for function 'mrdivide'
+ *
+ * C source code generated on: Thu Nov  8 13:52:45 2012
+ *
+ */
+
+/* Include files */
+#include "rt_nonfinite.h"
+#include "dcblockST2DL.h"
+#include "freqest_periodogram_coder.h"
+#include "mf_coder.h"
+#include "mrdivide.h"
+
+/* Type Definitions */
+
+/* Named Constants */
+
+/* Variable Declarations */
+
+/* Variable Definitions */
+
+/* Function Declarations */
+
+/* Function Definitions */
+real_T mrdivide(real_T A, real_T B)
+{
+  return A / B;
+}
+
+/* End of code generation (mrdivide.cpp) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/rtGetInf.cpp

@@ -1,0 +1,139 @@
+/*
+ * rtGetInf.cpp
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Thu Nov  8 13:52:46 2012
+ *
+ */
+
+/*
+ * Abstract:
+ *       MATLAB for code generation function to initialize non-finite, Inf and MinusInf
+ */
+#include "rtGetInf.h"
+#define NumBitsPerChar	8U
+
+/* Function: rtGetInf ==================================================
+ * Abstract:
+ * Initialize rtInf needed by the generated code.
+ * Inf is initialized as non-signaling. Assumes IEEE.
+ */
+real_T rtGetInf(void)
+{
+  size_t bitsPerReal = sizeof(real_T) * (NumBitsPerChar);
+  real_T inf = 0.0;
+  if (bitsPerReal == 32U) {
+    inf = rtGetInfF();
+  } else {
+    uint16_T one = 1U;
+    enum {
+      LittleEndian,
+      BigEndian
+    } machByteOrder = (*((uint8_T *) &one) == 1U) ? LittleEndian : BigEndian;
+    switch (machByteOrder) {
+     case LittleEndian:
+      {
+        union {
+          LittleEndianIEEEDouble bitVal;
+          real_T fltVal;
+        } tmpVal;
+
+        tmpVal.bitVal.words.wordH = 0x7FF00000U;
+        tmpVal.bitVal.words.wordL = 0x00000000U;
+        inf = tmpVal.fltVal;
+        break;
+      }
+
+     case BigEndian:
+      {
+        union {
+          BigEndianIEEEDouble bitVal;
+          real_T fltVal;
+        } tmpVal;
+
+        tmpVal.bitVal.words.wordH = 0x7FF00000U;
+        tmpVal.bitVal.words.wordL = 0x00000000U;
+        inf = tmpVal.fltVal;
+        break;
+      }
+    }
+  }
+
+  return inf;
+}
+
+/* Function: rtGetInfF ==================================================
+ * Abstract:
+ * Initialize rtInfF needed by the generated code.
+ * Inf is initialized as non-signaling. Assumes IEEE.
+ */
+real32_T rtGetInfF(void)
+{
+  IEEESingle infF;
+  infF.wordL.wordLuint = 0x7F800000U;
+  return infF.wordL.wordLreal;
+}
+
+/* Function: rtGetMinusInf ==================================================
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/rt_defines.h

@@ -1,0 +1,24 @@
+/*
+ * rt_defines.h
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Thu Nov  8 13:52:46 2012
+ *
+ */
+
+#ifndef __RT_DEFINES_H__
+#define __RT_DEFINES_H__
+
+#include <stdlib.h>
+
+#define RT_PI	3.14159265358979323846
+#define RT_PIF	3.1415927F
+#define RT_LN_10	2.30258509299404568402
+#define RT_LN_10F	2.3025851F
+#define RT_LOG10E	0.43429448190325182765
+#define RT_LOG10EF	0.43429449F
+#define RT_E	2.7182818284590452354
+#define RT_EF	2.7182817F
+#endif
+/* End of code generation (rt_defines.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/codegen/lib/dcblockST2DL/rt_nonfinite.h

@@ -1,0 +1,53 @@
+/*
+ * rt_nonfinite.h
+ *
+ * Code generation for function 'dcblockST2DL'
+ *
+ * C source code generated on: Thu Nov  8 13:52:46 2012
+ *
+ */
+
+#ifndef __RT_NONFINITE_H__
+#define __RT_NONFINITE_H__
+
+#if defined(_MSC_VER) && (_MSC_VER <= 1200)
+#include <float.h>
+#endif
+#include <stddef.h>
+#include "rtwtypes.h"
+
+extern real_T rtInf;
+extern real_T rtMinusInf;
+extern real_T rtNaN;
+extern real32_T rtInfF;
+extern real32_T rtMinusInfF;
+extern real32_T rtNaNF;
+extern void rt_InitInfAndNaN(size_t realSize);
+extern boolean_T rtIsInf(real_T value);
+extern boolean_T rtIsInfF(real32_T value);
+extern boolean_T rtIsNaN(real_T value);
+extern boolean_T rtIsNaNF(real32_T value);
+
+typedef struct {
+  struct {
+    uint32_T wordH;
+    uint32_T wordL;
+  } words;
+} BigEndianIEEEDouble;
+
+typedef struct {
+  struct {
+    uint32_T wordL;
+    uint32_T wordH;
+  } words;
+} LittleEndianIEEEDouble;
+
+typedef struct {
+  union {
+    real32_T wordLreal;
+    uint32_T wordLuint;
+  } wordL;
+} IEEESingle;
+
+#endif
+/* End of code generation (rt_nonfinite.h) */

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/dcblockingfilter/dcblockST1UL.m

@@ -1,0 +1,46 @@
+% DC blocking filter
+%-------------------------------------------------------------
+% dcblock.m 
+%
+% [RxSampFilt] = dcblockST1UL(RxSamp)
+%
+% This digital filter blocks the DC component in the received signal. It is
+% primarily intended to remove the complex exponential at DC that is
+% transmitted in the ST1 uplink waveform for frequency estimation purposes.
+% Once the frequency has been estimated, the discrete tone should be
+% removed in order to avoid bit error rate degradations. The implementation
+% uses an order-one IIR filter.
+% 
+% Inputs:
+%   RxSamp      - Vector containing samples of received signal
+%
+% Outputs:
+%   RxSampFilt  - Vector containing filtered received samples
+%
+% Comments: For details see [J. de Freitas, "The DC Blocking Filter", Jan.
+%           2007, available at
+%           http://www.mathworks.com.au/matlabcentral/fileexchange/13792-the-dc-blocking-filter] 
+% 
+%-------------------------------------------------------------
+% Author: A. Pollok, Institute for Telecommunications Research
+% Project: ASRP
+%-------------------------------------------------------------
+% Copyright 2012 
+% Institute for Telecommunications Research
+% University of South Australia
+%-------------------------------------------------------------
+
+function [RxSampFilt] = dcblockST1UL(RxSamp)
+    
+% ST1 UL: coefficents of order-one IIR filter
+
+p = 0.999501949180559; % coeff for cut-on frequency at 3Hz
+    % fc = 3 / (OS/(2*Ts)); % normalised cut-on frequency
+    % p = dcblock_MatlabCentral(fc);
+b = [1 -1];
+a = [1 -p];
+    
+% filter frequency corrected samples
+RxSampFilt = filter(b,a,RxSamp);
+
+end

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/frequencyestimator/freqest_periodogram.m

@@ -1,0 +1,114 @@
+% Frequency estimation from periodogram.
+%-------------------------------------------------------------
+% freqest_periodogram.m 
+%
+% [fhat,cgainhat] = freqest_periodogram(RxSamp,tSamp,AcqParams)
+%
+% This function etimates the frequency of a noisy complex exponential by
+% maximising the periodogram. First, a coarse estimate is obtained by
+% computing the periodogram at the discrete frequencies of an FFT. This
+% coarse estimate then serves as the initialisation for the second
+% estimator stage, which climbs the objective function using Newton's
+% method. As proposed by Quinn et al. [1], Newton's method is applied to
+% the logarithm of the periodogram rather than the periodogram itself. This
+% approach eliminates the need for zero padding to increase the resolution
+% of the FFT-based initial estimate.
+%
+%   [1] B. Quinn et al., "Maximizing the Periodogram", in Proc. IEEE Global
+%       Global Telecommunications Conference, Nov. 2008 
+% 
+% Inputs:
+%   RxSamp      - Vector containing samples of received signal
+%   tSamp       - Vector of sample times in seconds
+%   AcqParams   - Struct holding aquisition parameters
+%       .T          - symbol period
+%       .OS         - oversampling factor
+%       .taumin     - minimum time offset
+%       .Nfft       - FFT length. If set to [] the length of RxSamp rounded
+%                     up to the next higher power of 2 will be used.
+%
+% Outputs:
+%   fhat        - Frequency estimate [Hz]
+%   cgainhat    - complex gain estimate
+%
+% Comments:
+% 
+%-------------------------------------------------------------
+% Author: A. Pollok, Institute for Telecommunications Research
+% Project: ASRP
+%-------------------------------------------------------------
+% Copyright 2012 
+% Institute for Telecommunications Research
+% University of South Australia
+%-------------------------------------------------------------
+
+function [fhat,cgainhat] = freqest_periodogram(RxSamp,tSamp,AcqParams)
+
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/frequencyestimator/test/freqest_periodogram_test.m

@@ -1,0 +1,79 @@
+% Test script for freqest_periodogram.m.
+
+%% system parameters
+
+% local variables
+T = 1/10e3; %symbol period, 10KHz symbol rate
+OS = 4;
+Ts = T/OS;
+taumin = 0;
+taumax = 10e-3; %maximum 10 millisconds time of flight 
+M = 4; %QPSK
+g = @(t) sinc(t/T)/sqrt(T); %transmission pulse
+L = 1000; % number of symbols
+P = []; % pilot indices
+D = 1:L; % data indices
+
+% complex exponential for frequency estimation
+fsin = (1+0.1)/(2*T); % frequency shift relative to band centre
+signal_to_sine_ratio_dB = 10;
+amplsin = sqrt( 1 / (10^(signal_to_sine_ratio_dB/10)*T) );
+fr_max = 0; % maximum Doppler rate [Hz/s]
+
+%% channel parameters
+a0 = 1;
+tau0 = 5.111e-3;
+f0 = -2364.15; % frequency offset [Hz]
+fr0 = 0; % Doppler rate [Hz/s]
+
+snr_dbs = 3;
+sigma2 = 10.^(-snr_dbs/10)/Ts;
+N = ceil((T*L+taumax)/Ts);
+ts = (0:N-1)*Ts;
+
+%struct with the modulation parameters and estimator parameters    
+acqparams.T = T; %symbol period, 10KHz symbol rate
+acqparams.OS =  OS;
+acqparams.taumin = taumin;
+acqparams.taumax = taumax; %maximum 10 millisconds time of flight 
+acqparams.M = M; %QPSK
+acqparams.pshape = g; %transmission pulse
+acqparams.P = P; %pilot indices
+acqparams.D = D; %data indices
+acqparams.g = g;
+
+acqparams.fsin = fsin;
+acqparams.amplsin = amplsin;
+acqparams.fr_max = fr_max;
+
+%generate random psk symbols
+s = exp(2i*pi*(randi(M,[1,L])-1)/M);
+% set the pilots
+acqparams.p = s(P);
+
+%generate recieved signal.
+received = @(t) a0*transmitted(t, s, acqparams);
+% complex exponential
+cexp = @(t) amplsin*exp(2i*pi*fsin*t) .* (t>=T) .* (t<L*T);
+% apply frequency offset and Doppler rate
+rx = (received(ts - tau0) + cexp(ts - tau0)) .* exp(2i*pi * (f0*ts + (fr0/2)*ts.^2));
+% add noise
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/matchedfilter/mf_coder.m

@@ -1,0 +1,47 @@
+function mfout = mf_coder(rx, tauhat, rolloff, q,T,P,D,RrcNsym)
+%Run a matched filter on the received samples, returns recieved
+%symbols.
+%    
+%    mfout = mf(rx, tauhat, acqparams)
+%
+% INPUTS:
+%   rx              - sampled received signal
+%   tauhat        - time offset estimate
+%   acqparam        - struct holding aquisition parameters
+%       .T              - symbol period
+%       .L              - number of symbols (assumed consequtive)
+%       .OS             - oversampling factor
+%       .pshape         - function representing transmission pulse shape
+%
+% OUTPUTS:
+%   mfout        - matched filtered signal at symbol rate
+    
+% (c) 2012 ITR-UniSA
+% Authors: Robby McKilliam
+% Created: 4 April 2012
+    
+    %g = acqparam.pshape;
+    p = 1;
+    %q = acqparam.OS;
+    %T = acqparam.T;
+    Ts = p*T/q;
+    mini = min(min(P),min(D)); 
+    maxi = max(max(P),max(D));
+    N = length(rx);
+    L = maxi;
+    
+    ns = (p*q*mini - p*N : p*q*maxi - p)*(-1)*Ts/p - tauhat;
+    %lvec = conj( g(-ns*Ts/p - tauhat) );
+    lvec = conj( rrcpulse(ns,rolloff,T,q) .*...
+    (abs(ns) <= RrcNsym*q*Ts));
+    zvec = zeros(1,p*N - p + 1) + 1j*zeros(1,p*N - p + 1); 
+    zvec(1:p:p*N-p+1) = rx; %zerofilled r
+    h = conv_fft(lvec, zvec);
+
+    %plot(real(h));
+    %length(h)
+    %length(q*(mini:maxi))
+    
+    mfout = h(q*(mini:maxi) - q*mini + 1)*Ts;
+    
+end

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/polynomialPhaseTransform.m

@@ -1,0 +1,24 @@
+function [phiShifted,freqShifted,frateShifted] = polynomialPhaseTransform(phi,freq,frate,tau)
+
+% This function applies a polynomial phase transform to change the time
+% reference point (t=0). The user-provided coefficients describe the
+% second-order polynomial phase signal
+%
+%       phi + 2*pi*freq*t + 2*pi*(frate/2)*t^2,
+%
+% which are used to calculate the coefficients of the polynomial
+%
+%       phiShifted + 2*pi*freqShifted*(t-tau) + 2*pi*(frateShifted/2)*(t-tau)^2.
+%
+% Relative to the original phase polynomial, the time origin of the new
+% polynomial is shifted by tau.
+%
+% For details, see Andre Pollok's ASRP workbook 007, page 39.
+%
+% (c) 2012 ITR-UniSA
+% Authors: Andre Pollok
+% Created: 29 July 2012
+
+frateShifted = frate;
+freqShifted  = freq + 2*(frate/2)*tau;
+phiShifted   = phi  + 2*pi*freq*tau + 2*pi*(frate/2)*tau^2;

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Acquisition/pskacquisition_init.m

@@ -1,0 +1,124 @@
+function pskacquisition_init(acqparams)
+% This function initialises the acquisition module and must be called prior
+% to running pskacquisition.m. Specifically, this function constructs
+% objects for the time and phase offset estimators.
+%
+% Comment: It is possible to only use a subset of pilot and data symbols in
+%          the timing and phase estimators. For example, this can be useful
+%          to meet constraints on the computational complexity.
+%
+%   pskacquisition_init(acqparams)
+%
+% INPUTS:
+%   acqparams       - struct holding aquisition parameters
+%       .T              - symbol period
+%       .OS             - oversampling factor
+%       .P              - position (indices) of pilot symbols
+%       .p              - pilot symbols
+%       .D              - position (indices) of data symbols
+%       .PTime          - position (indices) of pilot symbols used by
+%                         timing estimator (equal to or a subset of .P)
+%       .pTime          - pilot symbols used by timing estimator (equal to 
+%                         or a subset of .p)
+%       .DTime          - position (indices) of data symbols used by timing
+%                         estimator (equal to or a subset of .D)
+%       .PPhase         - position (indices) of pilot symbols used by phase
+%                         estimator (equal to or a subset of .P)
+%       .pPhase         - pilot symbols used by phase estimator (equal to
+%                         or a subset of .p)
+%       .DPhase         - position (indices) of data symbols used by phase
+%                         estimator (equal to or a subset of .D)
+%       .M              - M for M-PSK, i.e. size of constellation
+%       .rolloff        - roll-off factor of root raised cosine pulse
+%       .Tpshape        - duration of the transmit pulse shape in seconds
+%       .timingEstimatorName
+%                       - string identifying the timing estimator
+%       .taumin         - minimum time offset
+%       .taumax         - maximum time offset
+%       .phaseEstimatorName
+%                       - string identifying the phase estimator
+%       .fr_max         - maximum Doppler rate
+%       .fr_min         - minimum Doppler rate
+%       .f_max          - range of residual Doppler after initial frequency
+%                         estimation
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Codecs/Files/readsparsepegmatrix.m

@@ -1,0 +1,51 @@
+% Reads parity-check matrix from PEG file format in compressed matrix format
+%---------------------------------------------------------------------
+% readsparsepegmatrix.m - reads parity-check matrix from PEG file format
+% in compressed matrix format
+% 
+%   c = readsparsepegmatrix(filename)
+%
+% Inputs:
+%   filename    - Filename of the PEG file
+% Outputs:
+%    c          - LDPC code in compressed matrix format
+% Comments:
+%      
+%---------------------------------------------------------------------
+% Author:   G. Lechner, ITR
+% Project : ASRP
+%---------------------------------------------------------------------
+% Copyright 2011 : 
+% Institute for Telecommunications Research
+% University of South Australia
+%---------------------------------------------------------------------
+%
+%---------------------------------------------------------------------
+
+function c = readsparsepegmatrix(filename)
+
+
+fid = fopen(filename, 'r');
+
+N   = fscanf(fid, '%i',1);
+M   = fscanf(fid, '%i',1);
+dc  = fscanf(fid, '%i',1);
+
+sparsepeg = zeros(M,dc);
+for m=1:M
+  for d=1:dc
+    sparsepeg(m,d) = fscanf(fid, '%i',1);
+  end
+end
+fclose(fid);
+
+[dv,x]= hist(nonzeros(sparsepeg(:)),1:N);
+dvmax = max(dv);
+c     = zeros(N,dvmax);
+
+for n=1:N
+  idx = find(sum((sparsepeg==n),2))';
+  idx = [idx zeros(1,dvmax-size(idx,2))];
+
+  c(n,:) = idx;
+end

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Codecs/PEG/BigGirth.cpp

@@ -1,0 +1,361 @@
+#include <stdio.h>
+#include <stdlib.h>
+#include <iostream>
+#include <fstream>
+#include "BigGirth.h"
+
+#define MAX(a,b)	((a)>(b) ? (a) : (b))
+#define MIN(a,b)	((a)<(b) ? (a) : (b))
+
+using namespace std;
+
+long int Random::uniform(long int  a, long int b) // [a, b-1]
+{
+  return (long int)((double)rand()/(RAND_MAX+1.0)*(b-a)+a);
+}
+
+Node::Node(void)
+{
+	connection = NULL;
+}
+
+Node::~Node(void)
+{
+	delete connection;
+}
+
+void Node::Init(int deg, int max)
+{
+	d						= deg;
+	dmax				= max;
+	connection	= new int[dmax];
+}
+
+void Node::Init(int deg)
+{
+	Init(deg, deg);
+}
+
+bool Node::active()
+{
+	return d<dmax;
+}
+
+
+// ============================================================================
+
+BigGirth::BigGirth(void)
+{
+}
+
+BigGirth::BigGirth(long int rows, long int cols, int *symbolDegSequence, int *checkDegSequence, char *filename, long int sglConcent, long int tgtGirth)
+{
+	long int n, m, d, index, localDepth=2*cols;
+	int iter;
+
+	EXPAND_DEPTH			= MAX((tgtGirth-4)/2, 0);
+	myrandom					= new Random();
+	M									= rows;
+	N									= cols;
+	(*this).filename	= filename;
+	localGirth				= new long int[N];
+	varnodes					= new Node[N];
+	chknodes					= new Node[M];
+
+	for(n=0;n<N;n++)
+		varnodes[n].Init(symbolDegSequence[n]);
+	
+	for(m=0;m<M;m++)
+	{
+		if(sglConcent==0)
+			chknodes[m].Init(0, checkDegSequence[m]);
+		else
+			chknodes[m].Init(0, 2*checkDegSequence[m]);
+	}
+      
+	for(n=0;n<N;n++)
+	{
+		index	= -1;
+		for(m=0;m<M;m++)
+		{
+			if(((chknodes[m].d<d) || (index<0)) && chknodes[m].active())
+			{
+				d			= chknodes[m].d;
+				index	= m;
+			}
+		}
+		varnodes[n].connection[0] = index;
+
+		iter = 0; 
+		ITER:
+		localGirth[n] = 2*N;
[…]

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/Codecs/readldpccode.m

@@ -1,0 +1,37 @@
+% Reads LDPC structure from a file
+%---------------------------------------------------------------------
+% readldpccode.m - Reads LDPC structure from a file
+% 
+% code = readldpccode(filename)
+%
+% Inputs:
+%   filename    - File to read from
+% Outputs:
+%    code       - LDPC code structure
+% Comments:
+%      
+%---------------------------------------------------------------------
+% Author:   G.Lechner,ITR
+% Project : ASRP
+%---------------------------------------------------------------------
+% Copyright 2006 : 
+% Institute for Telecommunications Research
+% University of South Australia
+%---------------------------------------------------------------------
+% Contains ITR Background IP
+%---------------------------------------------------------------------
+function code = readldpccode(filename)
+
+fid = fopen(filename,'r');
+
+N = fscanf(fid, '%i', 1);
+M = fscanf(fid, '%i', 1);
+E = fscanf(fid, '%i', 1);
+
+code.vardegree   = fscanf(fid, '%i', N)';
+code.chkdegree   = fscanf(fid, '%i', M)';
+code.interleaver = fscanf(fid, '%i', E)';
+
+code.interleaver = code.interleaver + 1;
+
+fclose(fid);

### 2013-11-06 code/st1matlab/impl/software/ref-model/mlibrary/pulseshapefilter.m

@@ -1,0 +1,18 @@
+function TxSymbols = pulseshapefilter(Syms, ModemParams)
+
+%-------------------------------------------------------------%
+%   Pulse Shape Filter (RRC)
+%-------------------------------------------------------------%
+TxSymbols = upfirdn(Syms,ModemParams.TxRrcCoefs,ModemParams.P,1);
+
+
+% now the tail of the filter is included to match the hardware
+% implementation (5.2.2013 - Andre, Gottfried)
+%--------------------------------------------------------%
+%----------- Truncate Rrc Filter end Tail----------------%
+% Ltail     = length(TxSymbols)-(length(Syms)*ModemParams.P);
+% Lh        = length(ModemParams.TxRrcCoefs);
+% Lend      = Ltail-(Lh-1)/2;
+% TxSymbols(end-Lend+1:end)=[];%chop end tail only
+
+end

### 2013-11-06 code/st1matlab/plotFER.m

@@ -1,0 +1,66 @@
+
+% load('results/st1_5p_da_20130909T114133.mat')
+% figure
+% semilogy(EbNodB_vec,numFrameErr./numFramesSimulated,'rx-','DisplayName','data-aided, 5 pilots')
+% hold on, grid on
+% load('results/st1_10p_da_20130909T114110.mat')
+% semilogy(EbNodB_vec,numFrameErr./numFramesSimulated,'rx-','DisplayName','data-aided, 10 pilots')
+% load('results/st1_15p_da_20130909T114048.mat')
+% semilogy(EbNodB_vec,numFrameErr./numFramesSimulated,'rx-','DisplayName','data-aided, 15 pilots')
+% load('results/st1_5p_Mackenthun_20130909T110106.mat')
+% semilogy(EbNodB_vec,numFrameErr./numFramesSimulated,'go-','DisplayName','Mackenthun, 5 pilots')
+% load('results/st1_10p_Mackenthun_20130909T110215.mat')
+% semilogy(EbNodB_vec,numFrameErr./numFramesSimulated,'go-','DisplayName','Mackenthun, 10 pilots')
+% load('results/st1_15p_Mackenthun_20130909T110746.mat')
+% semilogy(EbNodB_vec,numFrameErr./numFramesSimulated,'go-','DisplayName','Mackenthun, 15 pilots')
+% load('results/st1_5p_pc_20130909T151318.mat')
+% semilogy(EbNodB_vec,numFrameErr./numFramesSimulated,'ks-','DisplayName','perfect channel')
+% 
+% xlabel('Eb/No [dB]')
+% ylabel('FER')
+
+%%
+clear
+
+% dirname = 'resultsOLD/'; fileCountExpected = 10;
+% dirname = 'results/'; fileCountExpected = 10;
+dirname = 'resultsCOMBINED/'; fileCountExpected = 20;
+
+% basefilename = {'st1_5p_da_2013', 'st1_10p_da_2013', 'st1_15p_da_2013', ...
+%   'st1_5p_Mackenthun_2013', 'st1_10p_Mackenthun_2013', 'st1_15p_Mackenthun_2013', ...
+%   'st1_5p_pc_2013'};
+
+basefilename = {'st1_5p_da_job', 'st1_10p_da_job', 'st1_15p_da_job', ...
+  'st1_5p_Mackenthun_job', 'st1_10p_Mackenthun_job', 'st1_15p_Mackenthun_job', ...
[…]

### 2013-11-06 code/st1matlab/run_on_cluster.sh

@@ -1,0 +1,30 @@
+#!/bin/bash
+filename=$1
+for file in $filename 
+do
+  if [ -f "$file" ]
+  then
+    tmp=${file/run_}
+    wrapper=${tmp/.sh/_wrap.sh}
+    matfile=${tmp/.sh} 
+    if [ -f "$wrapper" ]
+    then
+      echo "Skipping $wrapper as file already exists."
+    else
+      echo "Processing $file"
+      echo "    Generating $wrapper..."
+      echo "#!/bin/bash" >> $wrapper
+      echo ./$file /usr/local/MATLAB/MATLAB_Compiler_Runtime/v80 \"$matfile\" >> $wrapper
+      chmod u+x $wrapper
+      echo "    Executing qsub $wrapper"
+      qrun `qsub -d /home/polloka/st1matlab/standalone/ $wrapper` 
+    fi
+  else
+    echo "$file does not exist."
+  fi
+done
+sleep 2s
+echo
+echo "QSTAT:"
+echo
+qstat

### 2013-11-06 code/st1matlab/setpath.m

@@ -1,0 +1,41 @@
+% used to enable acces to sub-directory files
+%--------------------------------------------------------------------------
+% setpath.m - This script sets the paths for various modem modules.
+% 
+% setpath
+%
+% Inputs:
+%       None
+% Outputs:
+%       None
+% Comments:
+%       This  script is called prior to each set of new bursts
+%--------------------------------------------------------------------------
+% Author:  ITR
+% Project: ASRP
+%--------------------------------------------------------------------------
+% Copyright 2011 : 
+% Institute for Telecommunications Research
+% University of South Australia
+%--------------------------------------------------------------------------
+% Contains ITR Background IP
+%--------------------------------------------------------------------------
+
+rDIR = '.';
+
+% addpath([rDIR '/SNR']);
+% addpath([rDIR '/branches/st1demo_Flight/postproc']);
+% addpath([rDIR '/branches/st1demo_Flight/postprocNTS']);
+% addpath([rDIR '/branches/st1demo_Flight/visTools']);
+addpath(rDIR);
+addpath([rDIR '/.']);
+addpath([rDIR '/Acquisition']);
+addpath([rDIR '/impl/software/library/phy/bin']);
+addpath([rDIR '/impl/software/ref-model/mlibrary']);
+addpath([rDIR '/impl/software/ref-model/mlibrary/Acquisition']);
+addpath([rDIR '/impl/software/ref-model/mlibrary/Acquisition/frequencyestimator']);
+addpath([rDIR '/impl/software/ref-model/mlibrary/Acquisition/matchedfilter']);
+addpath([rDIR '/impl/software/ref-model/mlibrary/Acquisition/noiseestimator']);
+addpath([rDIR '/impl/software/ref-model/mlibrary/Acquisition/dcblockingfilter']);
+addpath([rDIR '/impl/software/ref-model/mlibrary/Codecs/']);
+clear rDIR

### 2013-11-14 response.tex
moved paper back to old version and started modifying introduction.  Need to add that we've implemented it in asrp

@@
 	\\\textbf{Response:}
-	We have shortened Sec.\ II to make it more concise and improve readability.
+%	We have shortened Sec.\ II to make it more concise and improve readability.
 \item\label{R1it2}\textbf{Comment:} 
@@
 	\\\textbf{Response:}	
-	For improved readability of the manuscript, we have now moved all of the lemmas into the appendix.
+%	For improved readability of the manuscript, we have now moved all of the lemmas into the appendix.
 We have also added a discussion of the implications and the practical relevance of the theorems to Sec.\ III (Sec.\ IV in the original manuscript).
 
-	\comment{Stress practical relevance by showing ASRP PER results.
+%	\comment{Stress practical relevance by showing ASRP PER results.
 Mention trials and patents.
@@
 	\\\textbf{Response:}
-	We agree with the reviewer and apologise for causing confusion with the use of the terms \emph{coherent} and \emph{noncoherent} in the original manuscript.
+%	We agree with the reviewer and apologise for causing confusion with the use of the terms \emph{coherent} and \emph{noncoherent} in the original manuscript.
 In the updated manuscript, we refer to an estimator that uses known pilot symbols as \emph{data-aided} and to one that relies on receiver decisions on the data symbols as \emph{decision-directed}.
@@
 	missing `,' after (5)
-	\\\textbf{Response:} We thank the reviewer for spotting the missing comma and have added it.
+	\\\textbf{Response:} 
+%We thank the reviewer for spotting the missing comma and have added it.
 \item\textbf{Comment:} 
 	p.\ 19, at the bottom: sortindicIes (typo, delete 'i')
-	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
+	\\\textbf{Response:} % Thank you for pointing out this typographical error. We have corrected it.
 \item\textbf{Comment:} 
 	In Section VI, I would prefer a reference to the Lemma's in the Appendix at the start of the section.
-	\\\textbf{Response:} We have added a reference to the Lemmas in the appendix at the beginning of the section. Please note that Sec.\ VI of the original manuscript is now Sec.\ III.
+	\\\textbf{Response:} %We have added a reference to the Lemmas in the appendix at the beginning of the section. Please note that Sec.\ VI of the original manuscript is now Sec.\ III.
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-	We have removed the aforementioned sentence from the manuscript as we feel it has caused more confusion than provided insight.
+%	We have removed the aforementioned sentence from the manuscript as we feel it has caused more confusion than provided insight.
 We hope the following explanation clears up any confusion it may have caused.
@@
 
-	\comment{Add CRBs to plots.} Please note that Sec.\ VIII of the original manuscript is now Sec.\ VI.
+%	\comment{Add CRBs to plots.} Please note that Sec.\ VIII of the original manuscript is now Sec.\ VI.
 \end{enumerate}
@@
 	\\\textbf{Response:}
-	We thank the reviewer for pointing us to the suggested references and have added \comment{some/most of} them. Specifically, we have included references [c, d, e]. \comment{Add the remaining ones?}
+%	We thank the reviewer for pointing us to the suggested references and have added \comment{some/most of} them. Specifically, we have included references [c, d, e]. \comment{Add the remaining ones?}
 \end{enumerate}
@@
 	Paragraph before (13): \dots can be computed from its predecessor \dots
-	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
+	\\\textbf{Response:} %Thank you for pointing out this typographical error. We have corrected it.
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-	We thank the reviewer for pointing out the missing definition of $f(r, \phi)$ and have added it.
+%	We thank the reviewer for pointing out the missing definition of $f(r, \phi)$ and have added it.
 Please note that in response to comment \ref{R4it2e} of reviewer 4, Sec.\ III has been slightly shortened and has been moved to the appendix.
@@
 	\\\textbf{Response:}
-	Please refer to our response to comment \ref{R1it3a} of reviewer 1.
+%	Please refer to our response to comment \ref{R1it3a} of reviewer 1.
 \end{enumerate}
@@
 \item\textbf{Comment:} 
-	Figs.\ 1Ð3 illustrate the evolution of the phase MSE vs.\ SNR. Only in case that $|P| = 0$,
+	Figs.1 to 3 illustrate the evolution of the phase MSE vs.\ SNR. Only in case that $|P| = 0$,
 the LS results are compared to an alternative solution represented by the Viterbi and Viterbi (VV) algorithm.
@@
 	\textbf{Response:}
-	\comment{add response\dots}
+%	\comment{add response\dots}
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-	\comment{Replace variance plot by MSE plot.}
+%	\comment{Replace variance plot by MSE plot.}
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-	We have added a brief discussion of the need for sufficiently accurate coarse estimates to initialise an iterative code-aided synchronisation stage, citing [d, e].
+%	We have added a brief discussion of the need for sufficiently accurate coarse estimates to initialise an iterative code-aided synchronisation stage, citing [d, e].
 We have also added a set of new simulation results, which show the packet error rate of an LDPC-coded satellite communication system to demonstrate the use and performance of the proposed estimator in a practical system.
@@
 
-	\comment{Add response related to [f, g]\dots}
+%	\comment{Add response related to [f, g]\dots}
 \end{enumerate}
@@
 	\\\textbf{Response:}
-	We thank the reviewer for the comments. Please refer to our response to comment \ref{R1it2} of reviewer 1.
+%	We thank the reviewer for the comments. Please refer to our response to comment \ref{R1it2} of reviewer 1.
 \end{enumerate}
@@
 	\\\textbf{Response:}
-	\comment{Stress practical relevance by showing ASRP PER results. In particular ST1, where the estimators are used to seed turbo sync. Mention trials and patents\dots}
+%	\comment{Stress practical relevance by showing ASRP PER results. In particular ST1, where the estimators are used to seed turbo sync. Mention trials and patents\dots}
 
-	\comment{Not clear what the reviewer means by ``existing uncoherent methods that do not require estimation of the amplitudes''.}
+%	\comment{Not clear what the reviewer means by ``existing uncoherent methods that do not require estimation of the amplitudes''.}
 
@@
 	\\\textbf{Response:}
-	\comment{Show PER ASRP results. Mention trials and patents\dots}
+%	\comment{Show PER ASRP results. Mention trials and patents\dots}
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-	\comment{Show PER ASRP ST1 turbo sync results. Comparison with other estimators?}
+%	\comment{Show PER ASRP ST1 turbo sync results. Comparison with other estimators?}
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:} 
-	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the estimator MSE.
+%	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the estimator MSE.
 This was done to validate the theoretical value of the asymptotic variance of $\hat\rho$ predicted by Theorem~2 against Monte-Carlo simulations.
@@
 	\\\textbf{Response:} 
-	Biased estimators are quite common in the communications literature and in fact, are often used in practice.
+%	Biased estimators are quite common in the communications literature and in fact, are often used in practice.
 Prime examples are state-of-the-art signal-to-noise ratio (SNR) estimators for $M$-PSK modulated signals such as those discussed by Pauluzzi and Beaulieu \cite{Pauluzzi2000}.
@@
 	\\\textbf{Response:}
-	We have shortened Sec.\ II to make it more concise and improve readability. Sec.\ VII has been removed from the revised manuscript.
+%	We have shortened Sec.\ II to make it more concise and improve readability. Sec.\ VII has been removed from the revised manuscript.
 
-	Furthermore, we agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that Sec.\ III specifically discusses random variables of the form $Re^{j\varPhi}=1{+}W$, where $W$ is circularly symmetric.
+%	Furthermore, we agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that Sec.\ III specifically discusses random variables of the form $Re^{j\varPhi}=1{+}W$, where $W$ is circularly symmetric.
 The joint pdf of $R$ and $\varPhi$ is a crucial ingredient for the asymptotic analysis presented in the manuscript and therefore, we feel the section should be retained.

### 2013-11-14 paper.tex
moved paper back to old version and started modifying introduction.  Need to add that we've implemented it in asrp

@@
 These effects must be compensated for at the receiver.
-In this paper we assume that the time offset has been previously handled, and we focus on estimating the phase shift and attenuation.
+In this paper we assume that the time offset has been previously handled and we focus on estimating the phase shift and attenuation.
 We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK) and $M$-ary phase shift keying ($M$-PSK).
@@
 where $\abs{x}$ denotes the magnitude of the complex number $x$.
-The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is additive white and Gaussian.
+The least squares estimator is also the maximum likelihood estimator under the assumption that the noise sequence $\{w_i, i \in \ints\}$ is white and Gaussian.
 However, as we show, the estimator works well under less stringent assumptions.
@@
 
-The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$).
+% The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} (also know as \emph{blind}, \emph{non-data aided} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
 In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
-A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+A popular approach is estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
 The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.
 A function $F: \reals \mapsto \reals$ is chosen and the estimator of the carrier phase $\theta_0$ is taken to be $\tfrac{1}{M}\angle{C}$ where $\angle$ denotes the complex argument and
-\begin{equation}\label{eq:viterbiviterbi}
-C = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
-\end{equation}
-Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
+% \begin{equation}\label{eq:viterbiviterbi}
+% C = \frac{1}{L}\sum_{i \in P \cup D} F(\abs{y_i}) \big(\tfrac{y_i}{\abs{y_i}}\big)^M.
+% \end{equation}
+% Various choices for $F$ are suggested in~\cite{ViterbiViterbi_phase_est_1983} and a statistical analysis is presented.
 A caveat of this estimator is that it is not obvious how pilot symbols should be included.
 %Techniques are available but they are somewhat ad hoc~\cite{Cowley_ref_sym_carr_1998}.
-This problem does not occur with the least square estimator.
+% This problem does not occur with the least square estimator.
 
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Sweldens considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
+Both Mackenthun and Sweldens considered what is called \emph{noncoherent detection} (also know as \emph{blind}, \emph{non-data aided} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
+We show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
 Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
 
-In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.
+Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least square estimator of carrier amplitude and phase obtained by minimising~\eqref{eq:SSdefn}.
+In the literature, analytic studies of non-data aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
+The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994}, or to analyse bound that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
+To the authors knowledge, all of the existing bounds are applicable only in the fully non-data aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
+Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.
+
+For the case of carrier amplitude the situations is even more complicated.
+A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that all known non-data aided estimators for carrier amplitude exhibit bias.
+The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
+Neverthess, Cramer-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
+Unfortunately, these bounds may not be meaningful since known non-data aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
+
+In this paper we present an alternative approach to the study of estimators for carrier phase and amplitude.
+Rather than compute bounds, we directly study the least squares estimators of carrier phase and amplitude asymptotically as the number of transmitted symbols $L\to \infty$.
+We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent (asymptotically unbiased) and asymptotically normally distributed.
+However, the least squares amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation $M$.
+To our knowledge, this is the first time an analytic description of the bias or variance of non-data aided amplitude estimator has been obtained.
+Our expressions for the asymptotic variance and bias are simple to compute.
+%Our results are indirectly supported by the simulation study of Pauluzzi~\cite{Pauluzzi2000}.
+We present our our set of simulations supporting our analytic results in Section~\ref{}.
+
+%Two criticisms might be directed at our asymptotic approach when compared to the Cramer-Rao bounds already developed in the literature.
+Perhaps the most obvious is that the results apply asymptotically as $L\to\infty$.
+The simulation results we present in Section~\ref{} confirm that our asymptotic results accurately model the behaviour of the least squares estimator for modest numbers of symbols, for example $L=256$
+
+One criticism that might be directed at our asymptotic approach when compared to the Cramer-Rao bounds already developed in the literature is that our specific results apply only to the least square estimator.
+One response to this criticism is that, because the least squares estimator can be rapidly computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian, then it is likely to be a popular practical choice.
+Perhaps an even more convinving response is that, while the specific Theorems~\ref{thm:consistency} and~\ref{thm:normality} that we given apply only to the least square estimtor, the techniques that we develop to derive these results can likely be deployed for the analysis of other estimators.
+For example, it would be reasonably straigthforward to extend our results to allow data and pilot symbols with varying constellation size, i.e.
+varying $M$.
+For example, one might give more importance to certain data symbols and use BPSK ($M=2$) for these, but QPSK ($M=4$) for other less important symbols.
+This is related to what is called \emph{unequal error protection} in the literature~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010} and is commonly used in practical systems.
+Perhaps a more involved extension would be the inclusion of carrier frequency as a parameter to be estimated.
+The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, would potentially allow this.
+
+% In the literature it has been common to assume that the data symbols $\{d_i, i \in D\}$ are of primary interest and that the complex amplitude $a_0$ is a nuisance parameter.
 The metric of performance is correspondingly the \emph{symbol error rate}, or \emph{bit error rate}.

### 2013-11-14 paperandrenew.tex
added andre's version for ref

--- 2013-07-03 paper.tex
@@
 
+\usepackage{color}
+\newcommand{\comment}[1]{\textcolor{red}{#1}}
+
 \newcommand{\sgn}{\operatorname{sgn}}
@@
 
+% use italics version of capital Greek letters
+\renewcommand{\Phi}{\varPhi}
+\renewcommand{\Theta}{\varTheta}
+
 %opening
-\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
+\title{Asymptotic statistics of carrier phase and amplitude estimation for PSK using pilots and data}
 \author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, I. Vaughan L. Clarkson and Barry Quinn  
 \thanks{
-A preliminary version of this paper has been submitted to ICASSP'13~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Supported under the Australian Government’s Australian Space Research Program.
+A preliminary version of this paper was presented at ICASSP'13~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.  Supported under the Australian Government's Australian Space Research Program.
 Robby McKilliam, Andr\'{e} Pollok and Bill Cowley are with the Institute for Telecommunications Research, The University of South Australia, SA, 5095.
@@
 However, the amplitude estimator is not consistent and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
-%Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.
-The results of Monte Carlo simulations are provided and these agree with the theoretical results.   
+%Our theoretical results can also be applied to the case where no pilot symbols exist.
+The results of Monte-Carlo simulations are provided and these agree with the theoretical results.   
 \end{abstract}
@@
 s_i = \begin{cases}
-p_i & i \in P \\
-d_i & i \in D,
+p_i, & i \in P \\
+d_i, & i \in D,
 \end{cases}
@@
 where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.
-The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$  where $\emptyset$ is the empty set, and we let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
+The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$, where $\emptyset$ is the empty set, and we let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
 
@@
 
-The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} where no pilot symbols exist ($P = \emptyset$).
-In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
-A popular approach is the so called \emph{non-data aided}, sometimes also called \emph{non-decision directed}, estimator based on the paper of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+\comment{
+Doesn't Bill's paper \cite{Cowley_ref_sym_carr_1998} consider the case with pilots?
+}
+
+\comment{
+The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers \emph{non-data aided} estimation, where no pilot symbols exist ($P = \emptyset$).
+In this setting \emph{differential encoding} and \emph{differential (i.e.
+noncoherent) detection} are often used to handle the resulting $M$-fold ambiguity of the phase estimate.
+A popular non-data aided estimator was proposed by Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.%
+}
 The idea is to `strip' the modulation from the received signal by taking $y_i / \abs{y_i}$ to the power of $M$.
@@
 %Techniques are available but they are somewhat ad hoc~\cite{Cowley_ref_sym_carr_1998}.
-This problem does not occur with the least square estimator.
+This problem does not occur with the least squares estimator.
 
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Sweldens considered only the noncoherent setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
-Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
+Both Mackenthun and Sweldens considered only the non-data aided setting, but we show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
+Our model includes the non-data aided case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
 
@@
 The estimator $\hat{a}$ is also used in the computation of decoder metrics for modern decoders, and for interference cancellation in multiuser systems.
-Consequently, our metric of performance will not be symbol or bit error rate, but $\sabs{\hat{a} - a_0}$.
-It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\sabs{\langle\theta_0 - \hat{\theta}\rangle_\pi}$ and $\sabs{\hat{\rho} - \rho_0}$.
+\comment{Furthermore, iterative receiver synchronisation methods that are aided by the error correcting code, also known as \emph{turbo synchronisation} \cite{Lottici2004,Noels2005}, %\cite{Herzet2007,Herzet2007_TurboSyncFramework}
+typically converge correctly only if initialised at some $a$ sufficiently close to $a_0$.
+Obtaining an accurate estimator $\hat{a}$ of $a_0$ is therefore a crucial ingredient for an effective turbo synchronisation receiver.} Consequently, our primary metric of performance will not be symbol or bit error rate, but the estimation error.
+\comment{However, we will also present packet error rate results for an LDPC-coded satellite communication system to demonstrate the use of the proposed estimator in a practical system\dots}
+%It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$ where $\hat{\rho}$ is a positive real number, then we consider $\sabs{\langle\theta_0 - \hat{\theta}\rangle_\pi}$ and $\sabs{\hat{\rho} - \rho_0}$.
 The function $\fracpart{\cdot}_\pi$ denotes its argument taken `modulo $2\pi$' into the interval $[-\pi, \pi)$.
 It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi$ rather than $\theta_0 - \hat{\theta}$ is the appropriate measure of error for the phase parameter.
-It is usual to write \emph{estimator} minus \emph{error}, i.e., $\hat{\theta} - \theta_0$ rather than $\theta_0 - \hat{\theta}$.
+It is usual to write \emph{estimator} minus \emph{truth}, i.e., $\hat{\theta} - \theta_0$ rather than $\theta_0 - \hat{\theta}$.
 However, it is notationally convenient to use $\langle\theta_0 - \hat{\theta}\rangle_\pi$ rather than $\langle\hat{\theta} - \theta_0\rangle_\pi$ here.
 
-%It is possible to generalise the results we present here to allow data symbols with varying constellation size, i.e.
-varying $M$.
+%It is possible to generalise the results we present here to allow data symbols with varying constellation size, i.e.\ varying $M$.
 For example, one might give more importance to certain data symbols and use BPSK ($M=2$) for these, but QPSK ($M=4$) for other less important symbols.
@@
 The paper is organised in the following way.
-Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when both pilot symbols and data symbols exist.
-Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
-Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
+Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the case, when both pilot symbols and data symbols exist.
+Section~\ref{sec:stat-prop-least} states two theorems that describe the asymptotic statistical properties of the least squares estimator of carrier phase and amplitude.
+%Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
 We show, under reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
-The asymptotic bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but the asymptotic bias is significant when the SNR is small.
+The asymptotic bias of $\hat{\rho}$ is small when the signal to noise ratio (SNR) is large, but significant when the SNR is small.
 Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.  
-In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
-In this case, our expressions for the asymptotic distribution can be simplified.
-Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.
-These simulations agree with the derived asymptotic properties.
+%In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.  In this case, our expressions for the asymptotic distribution can be simplified.  
+In Section~\ref{sec:simulations} we present the results of Monte-Carlo simulations.  These simulations agree with the derived asymptotic properties.
 
@@
 
+\comment{
 \section{Mackenthun's algorithm with pilots}\label{sec:least-squar-estim}
-
-In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, so we modify the algorithm to include the pilot symbols.
+}
+
+%In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.  
+Mackenthun specifically considered the non-data aided setting and assumed differential encoding and detection to deal with the $M$-fold phase ambiguity \cite{Mackenthun1994}.
+In contrast, we consider non-differential encoding with coherent detection and to this end, extend Mackenthun's algorithm to include pilot symbols.
+
 For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 
-%In Section~\ref{sec:line-time-algor} we will show that a full sort is not neccesary, and that the least squares estimator can be implemented in $O(L)$ operations.
-
-Define the sum of squares function
+%In Section~\ref{sec:line-time-algor} we will show that a full sort is not necessary, and that the least squares estimator can be implemented in $O(L)$ operations.
+
+Our starting point is the sum of squares function defined in \eqref{eq:SSdefn}, which can be rewritten as
+%Define the sum of squares function
 \begin{align}
@@
 \end{align}
-where $*$ denotes complex conjugate.  The minimiser of $SS$ with respect to $a$ as a function of $\{d_i, i \in D\}$ is
+where $*$ denotes complex conjugation.  The minimiser of $SS$ with respect to $a$ as a function of $\{d_i, i \in D\}$ is
 \begin{equation}\label{eq:hata}
-\hat{a}(\{d_i, i \in D\}) = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y
+\hat{a}(\{d_i, i \in D\}) = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y(\{d_i, i \in D\}),
 \end{equation}
 where $L = \abs{P \cup D}$ and to simplify our notation we have put 
-\[
-Y = \sum_{i \in P \cup D} y_i s_i^* = \sum_{i \in P } y_i p_i^* + \sum_{i \in D } y_i d_i^*.
-\]  
-Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to suppress the argument $(\{ d_i, i \in D\})$ for notational brevity.
-Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain $SS$ minimised with respect to $a$,
+\begin{align}\label{eq:Y}
+	Y(\{d_i, i \in D\}) = \sum_{i \in P \cup D} y_i s_i^* 
+		%= \sum_{i \in P } y_i p_i^* + \sum_{i \in D } y_i d_i^*
+		= B + \sum_{i \in D } y_i d_i^*,
+\end{align}
+where $B = \sum_{i \in P} y_i p_i^*$ is independent of the data symbols.
+For notational brevity, we also write $Y$ instead of $Y(\{ d_i, i \in D\})$.
+Substituting $\hat{a}$ from \eqref{eq:hata} for $a$ into~\eqref{eq:SS} we obtain $SS$ minimised with respect to $a$,
 \begin{equation}\label{eq:SSdatasymbols}
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y}^2,
+	SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y(\{d_i, i \in D\})}^2,
 \end{equation}
 where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ does not depend on the $d_i$.
-The least squares estimators of the data symbols are the minimisers of~\eqref{eq:SSdatasymbols}.
-Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-It turns out that there are at most $M\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
-%When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
-
-To see this, let $a = \rho e^{j\theta}$ where $\rho > 0$ is real.  Now,
-\begin{align}
-SS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
-&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:SSallparams}
-\end{align}
-We have slightly abused notation here by reusing $SS$.
-This should not cause confusion as $SS(a, \{d_i, i \in D\})$, $SS(\rho, \theta, \{d_i, i \in D\})$, and $SS(\{d_i, i \in D\})$ are easily told apart by their arguments.
-For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
-\begin{equation}\label{eq:hatdfinxtheta}
-\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)} \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
-\end{equation}
+The least squares estimators of the data symbols, denoted $\{\hat{d}_i,i\in D\}$, are the minimisers of~\eqref{eq:SSdatasymbols} or, equivalently, the maximisers of $\abs{Y}^2$.
+%\begin{align}
+%	Q(\{d_i, i \in D\}) = \abs{Y(\{d_i, i \in D\})}^2.
+%\end{align}
+Observe that given candidate values for the data symbols, we can compute the corresponding $\abs{Y}^2$ in $O(L)$ arithmetic operations.
+The na\"{\i}ve approach is to compute the objective function $\abs{Y}^2$ for all possible $M^\abs{D}$ data sequences and select the sequence that yields the maximum value.
+Using Mackenthun's algorithm, we are able to avoid this computationally expensive exhaustive search, which would require $O(LM^L)$ arithmetic operations.
+
+First consider the case without pilot symbols, i.e.\ $P=\emptyset$.
+In this case, $\abs{Y}^2$ is identical to the objective function considered by Mackenthun \cite[Eq.\ (2)]{Mackenthun1994}.
+This equivalence stems from the fact that the least squares estimators $\hat{d}_i$ are also the maximum likelihood estimators under the additive white Gaussian noise assumption of \cite{Mackenthun1994}.
+Consequently, Mackenthun's algorithm as presented in~\cite{Mackenthun1994} or~\cite{Sweldens2001} can be applied without modification to obtain the $\hat{d}_i$.
+As shown in~\cite{Sweldens2001,Mackenthun1994}, at most $M\abs{D}$ candidate sequences need to be trialled.
+In fact, due to an $M$-fold phase ambiguity in the absence of pilot symbols, the objective function $\abs{Y}^2$ is periodic in $\abs{D}$ and thus, only $\abs{D}$ of the $M\abs{D}$ candidate sequences need to be considered.
+We do not explicitly state the algorithm at this point, but instead describe it below for the more general setting with pilots.
+
+In the presence of pilot symbols, all $M\abs{D}$ candidate sequences are relevant and we rely on the pilots to resolve the phase ambiguity.
+In the following we describe how these sequences can be enumerated.
+Let $\hat{b}_i$ denote the hard decision on the received data symbol $y_i$, $i \in D$, i.e.
+\[
+	\hat{b}_i = e^{j\hat{u}_i}, \qquad \text{where} \qquad \hat{u}_i = \round{\angle y_i},
+\]
 where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
@@
 This is not the case here.
-If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer then,
+If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer\footnote{The direction of rounding for half-integers is not important so long as it is consistent.
+We have chosen to round up half-integers here.} then,
 \[
@@
 \]
-Note that $\hat{d}_i(\theta)$ does not depend on $\rho$.
-As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
-With this in mind,
-\[
-\hat{u}_i(\theta) = \round{\angle{y_i} - \theta }
-\]
-which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $2\pi$.
-
-We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Let $b_i = \hat{d}_i(0)$ and let 
-\[
-z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}
-\]
-%be the phase difference between the received symbol $y_i$ and the hard decision resulting when $\theta = 0$, i.e. $\round{\angle{y_i}}$.  
-Then,
-\begin{equation}\label{eq:uicombos}
-\hat{d}_i(\theta) = 
-\begin{cases}
-b_i, &  0 \leq \theta < z_i + \frac{\pi}{M} \\
-b_i e^{-j2\pi/M}, & z_i + \frac{\pi}{M} \leq \theta < z_i + \frac{3\pi}{M} \\ 
-\vdots & \\
-b_i e^{-j2\pi k /M}, & z_i + \frac{\pi(2k - 1)}{M} \leq \theta < z_i + \frac{\pi(2k + 1)}{M}  \\ 
-\vdots & \\
-b_i e^{-j2\pi} = b_i, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
-\end{cases}
+
+Let $S$ denote the set of all $M\abs{D}$ candidate sequences.  The first sequence in $S$ is 
+\[
+f_0 = \{ \hat{b}_i, i \in D \}.
+\]  
+Now let
+\[
+	z_i = \angle(y_i \hat{b}_i^*) = \angle{y_i} - \hat{u}_i = \angle{y_i} - \round{\angle{y_i}}
+\]
+and let $\sigma$ denote the permutation of the indices in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
+\begin{equation}\label{eq:sigmasortind}
+	z_{\sigma(i)} \leq z_{\sigma(k)},
 \end{equation}
-
-Let 
-\[
-f(\theta) = \{ \hat{d}_i(\theta), i \in D \}
-\]
-be a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
-Observe that $f(\theta)$ is piecewise continuous.
-The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains constant are determined by the values of $\{z_i, i \in D\}$.
-Let
-\[
-S = \{ f(\theta) \mid \theta \in [0, 2 \pi) \}
-\]
-be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$.
-If $\hat{\theta}$ is the least squares estimator of the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $\hat{d}_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
-
-The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indices in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
-\begin{equation}\label{eq:sigmasortind}
-z_{\sigma(i)} \leq z_{\sigma(k)}
-\end{equation}
-whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}-1\}$.
-It is convenient to define the indices into $\sigma$ to be taken modulo $\abs{D}$, that is, if $m$ is an integer not from $\{0, 1, \dots, \abs{D}-1\}$ then we define $\sigma(m) = \sigma(k)$ where $k \equiv m \mod \abs{D}$ and $k \in  \{0, 1, \dots, \abs{D}-1\}$.
-The first sequence in $S$ is
-\[
-f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
-\]  
+whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}{-}1\}$.
+
 The next sequence $f_1$ is given by replacing the element $b_{\sigma(0)}$ in $f_0$ with $b_{\sigma(0)}e^{-j2\pi/M}$.
 Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.
-Using this notation,
-\[
-f_1 = f_0 e_{\sigma(0)}.
-\] 
-The next sequence in $S$ is correspondingly 
-\[
-f_2 = f_0 e_{\sigma(0)} e_{\sigma(1)} = f_1 e_{\sigma(1)},
-\]
-and the $k$th sequence is
+Using this notation, $f_1 = f_0 e_{\sigma(0)}$ and more generally, the $k$th sequence in $S$ is
 \begin{equation}\label{eq:fkrec}
@@
 
-We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,M\abs{D}-1\}$.
-Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
+Let $Y_k = Y( f_k )$.
+We want to find the $f_k \in S$ corresponding to the maximiser $\{\hat{d}_i,i\in D\}$ of $\abs{Y_k}^2$.
+A na\"{\i}ve approach would be to compute $\abs{Y_k}^2$ for each $k \in \{0,1,\dots,M\abs{D}{-}1\}$.
+Computing $\abs{Y_k}^2$ for any particular $k$ requires $O(L)$ arithmetic operations.
 So, this na\"{\i}ve approach would require $O(L M \abs{D}) = O(L^2)$ operations in total.
-Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.
-
-Let,
-\begin{equation}\label{eq:SSfk}
-SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
-\end{equation}
-where, 
+Following Mackenthun~\cite{Mackenthun1994}, we show how to recursively compute $\abs{Y_k}^2$.
+
+Using \eqref{eq:Y}, $Y_k$ can be rewritten as
 \begin{align*}
-Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
-&= B + \sum_{i \in D}g_{ki},
+	Y_k = Y( f_k ) 
+		&= B  + \sum_{i \in D} y_i f_{ki}^*
+		= B + \sum_{i \in D}g_{ki},
 \end{align*}
-where $B = \sum_{i \in P} y_i p_i^*$ is independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
+where $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
 Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
@@
 \]
-So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in $O(1)$ arithmetic operations.
-Given $Y_k$, the value of $SS(f_k)$ can be computed in $O(1)$ operations using~\eqref{eq:SSfk}.
-Let $\hat{k} = \arg\min SS(f_k)$.
-The least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata},
+So, each $Y_k$ can be computed from its predecessor $Y_{k-1}$ in $O(1)$ arithmetic operations.  
+%Given $Y_k$, the value of
+%\begin{equation*}
+%	SS(f_k) = A - \frac{1}{L}\abs{Y(f_k)}^2
+%		= A - \frac{1}{L}\abs{Y_k}^2
+%\end{equation*}
+%can be computed in $O(1)$ operations.  Let $\hat{k} = \arg\min_k SS(f_k)$.  
+Letting $\hat{k} = \arg\max_k \abs{Y_k}^2$, the least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata},
 \begin{equation}\label{eq:ahatYhat}
@@
 Line~\ref{alg_sortindices} contains the function $\operatorname{sortindices}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
-The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
+The $\operatorname{sortindices}$ function requires sorting $\abs{D}$ elements.
 This requires $O(L \log L)$ operations.
-The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.
+The $\operatorname{sortindices}$ function is the primary bottleneck in this algorithm when $L$ is large.
 The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
-%In the next sections we will show how to the sortinces function can be avoided.
+%In the next sections we will show how the sortindices function can be avoided.
 This leads to an algorithm that requires only $O(L)$ operations.
+
+Recalling that $f_{\hat{k}}$ is the sequence of least squares estimators $\{\hat{d}_i, i\in D\}$, we can rewrite $\hat{a}$ in \eqref{eq:ahatYhat} as
+\begin{align*}
+	\hat{a} = \frac{1}{L} Y_{\hat{k}}
+		= \frac{1}{L} Y(\{\hat{d}_i,i\in D\})
+		= \frac{1}{L} \sum_{i \in P} y_i p_i^*  + \frac{1}{L} \sum_{i \in D} y_i \hat{d}_{i}^*
+\end{align*}
+to illustrates that the first and second terms of the estimator operate in \emph{data-aided} and \emph{decision-directed} mode, respectively.
+If no pilots are present, $\abs{P}=0$ and the estimator is purely decision-directed.
+Conversely, not utilising any data symbols for estimation purposes, i.e.\ $\abs{D}=0$, corresponds to a fully data-aided estimator.
 
@@
 $\hat{a} = \frac{1}{L} Y$ \nllabel{alg_hata1} \;
-$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
+$\hat{Q} = \abs{Y}^2$ \nllabel{alg_Q} \;
 $\eta = e^{j2\pi/M} - 1$ \;
@@
 $g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
-$Q = \frac{1}{L}\abs{Y}^2$\;
+$Q = \abs{Y}^2$ \;
 \If{$Q > \hat{Q}$}{
@@
 
-\section{Circularly symmetric complex random variables}\label{sec:circ-symm-compl}
-
-Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
-A complex random variable $W$ is said to be \emph{circularly symmetric} if its phase $\angle{W}$ is independent of its magnitude $\abs{W}$ and if the distribution of $\angle{W}$ is uniform on $[0,2\pi)$.
-That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
-If the probability density function (pdf) of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
-\[
-f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
-\]
-% Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
-If $\expect\abs{W} = \expect Z$ is finite, then $W$ has zero mean because
-% \begin{align*}
-%  \expect W &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
-%  &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
-%  &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
-%  \end{align*}
-% If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
-% \[
-% f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
-% \]
-%from which follows
-%\begin{equation}\label{eq:fzandfZYwithz}
-%f_Z(z) = 2\pi z f_{X,Y}(z,0) = 2\pi z f_{X,Y}(0,z)
+%Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
+It turns out that there are at most $M\abs{D}$ candidate values \comment{$\leftarrow$ sequences?} of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
+%When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
+%To see this, let $a = \rho e^{j\theta}$ where $\rho > 0$ is real.  Now,
+%\begin{align}
+%SS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
+%&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:SSallparams}
+%\end{align}
+%We have slightly abused notation here by reusing $SS$.
+This should not cause confusion as $SS(a, \{d_i, i \in D\})$, $SS(\rho, \theta, \{d_i, i \in D\})$, and $SS(\{d_i, i \in D\})$ are easily told apart by their arguments.
+For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
+%\begin{equation}\label{eq:hatdfinxtheta}
+%\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)}, \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)},
 %\end{equation}
-%for all $z > 0$.
-
-We will have particular use of complex random variables of the form $1 + W$ where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying, 
-\[
-R e^{j\Phi} = 1 + W.
-\]
-The joint pdf of $R$ and $\Phi$ can be shown to be
- \begin{equation}\label{eq:pdfRPhi}
- f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
- \end{equation}
-Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi]$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$.
-The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.
-So,
-%\begin{equation}\label{eq:expectRepartRphi}
-\[
-\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
-\]
+%where $\angle(\cdot)$ denotes the complex argument (or phase), and $\round{\cdot}$ rounds its argument to the nearest multiple of $\frac{2\pi}{M}$.
+A word of caution, the notation $\round{\cdot}$ is often used to denote rounding to the nearest \emph{integer}.
+This is not the case here.
+If the function $\operatorname{round}(\cdot)$ takes its argument to the nearest integer\footnote{The direction of rounding for half-integers is not important so long as it is consistent.
+We have chosen to round up half-integers here.} then,
+%\[
+%\round{x} = \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right).
+%\] 
+%Note that $\hat{d}_i(\theta)$ does not depend on $\rho$.
+As defined, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
+With this in mind,
+%\[
+%\hat{u}_i(\theta) = \round{\angle{y_i} - \theta },
+%\]
+%which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $2\pi$.
+%
+%We only require to consider $\theta$ in the interval $[0, 2\pi)$.  Consider how $\hat{d}_i(\theta)$ changes as $\theta$ varies from $0$ to $2\pi$.  Let $b_i = \hat{d}_i(0)$ and let 
+%\[
+%z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}.
+%\]
+%%be the phase difference between the received symbol $y_i$ and the hard decision resulting when $\theta = 0$, i.e.\ $\round{\angle{y_i}}$.  
+%Then,
+%\begin{equation}\label{eq:uicombos}
+%\hat{d}_i(\theta) = 
+%\begin{cases}
+%b_i, &  0 \leq \theta < z_i + \frac{\pi}{M} \\
+%b_i e^{-j2\pi/M}, & z_i + \frac{\pi}{M} \leq \theta < z_i + \frac{3\pi}{M} \\ 
+%\vdots & \\
+%b_i e^{-j2\pi k /M}, & z_i + \frac{\pi(2k - 1)}{M} \leq \theta < z_i + \frac{\pi(2k + 1)}{M}  \\ 
+%\vdots & \\
+%b_i e^{-j2\pi} = b_i, &  z_i + \frac{\pi(2M - 1)}{M} \leq \theta < 2\pi. \\
+%\end{cases}
 %\end{equation}
-where $\Re(\cdot)$ denotes the real part, and
-\begin{equation}\label{eq:expectImpartRphi}
-\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
-\end{equation}
-where $\Im(\cdot)$ denotes the imaginary part.
-
-% \begin{lemma}\label{lem:h1minedcircsym}
-% Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
-Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+%
+%Let 
+%\[
+%f(\theta) = \{ \hat{d}_i(\theta), i \in D \}
+%\]
+%be a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
+Observe that $f(\theta)$ is piecewise continuous \comment{$\leftarrow$ constant?}.
+The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains constant are determined by the values of $\{z_i, i \in D\}$.
 Let
-% \[
-% h_1(x) = \expect R \cos(x + \Phi).
-% \]
-% Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
-% \end{lemma}
-% % Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
-That is, the pdf of $Z e^{j \Theta}$ in rectangular coordinates is given by $z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$ and $x$ and $y$ are the real and imaginary parts of $Z e^{j \Theta}$, and this pdf is non increasing with $z$.
-For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
-
-% % By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if
-% % \[
-% % \abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
-% % \]
-% % We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
-% \begin{IEEEproof}
-% BLERG
-% \end{IEEEproof}
+%\[
+%S = \{ f(\theta) \mid \theta \in [0, 2 \pi) \}
+%\]
+%be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$.
+If $\hat{\theta}$ is the least squares estimator of the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
+Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $\hat{d}_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
+%
+%The sequences in $S$ can be enumerated as follows.  Let $\sigma$ denote the permutation of the indices in $D$ such that $z_{\sigma(i)}$ are in ascending order, that is,
+%\begin{equation}\label{eq:sigmasortind}
+%z_{\sigma(i)} \leq z_{\sigma(k)},
+%\end{equation}
+%whenever $i < k $ where $i, k \in \{0, 1, \dots, \abs{D}{-}1\}$.
+It is convenient to define the indices into $\sigma$ to be taken modulo $\abs{D}$, that is, if $m$ is an integer not from $\{0, 1, \dots, \abs{D}{-}1\}$, then we define $\sigma(m) = \sigma(k)$ where $k \equiv m \mod \abs{D}$ and $k \in  \{0, 1, \dots, \abs{D}{-}1\}$.
+The first sequence in $S$ is
+%\[
+%f_0 = f(0) = \{ \hat{d}_i(0), i \in D \} = \{ b_i, i \in D \}.
+%\]  
+%The next sequence $f_1$ is given by replacing the element $b_{\sigma(0)}$ in $f_0$ with $b_{\sigma(0)}e^{-j2\pi/M}$.
+Given a sequence $x$ we use $x e_i$ to denote $x$ with the $i$th element replaced by $x_i e^{-j2\pi/M}$.
+Using this notation,
+%\[
+%f_1 = f_0 e_{\sigma(0)}.
+%\] 
+%The next sequence in $S$ is correspondingly 
+%\[
+%f_2 = f_0 e_{\sigma(0)} e_{\sigma(1)} = f_1 e_{\sigma(1)},
+%\]
+%and the $k$th sequence is
+%\begin{equation}\label{eq:fkrec}
+%f_{k+1} = f_{k} e_{\sigma(k)}.
+%\end{equation}
+%In this way, all $M\abs{D}$ sequences in $S$ can be recursively enumerated.
+%
+%We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
+A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,M\abs{D}{-}1\}$.
+Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
+So, this na\"{\i}ve approach would require $O(L M \abs{D}) = O(L^2)$ operations in total.
+Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.
+%
+%Let,
+%\begin{equation}\label{eq:SSfk}
+%SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
+%\end{equation}
+%where, 
+%\begin{align*}
+%Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
+%&= B + \sum_{i \in D}g_{ki},
+%\end{align*}
+%where $B = \sum_{i \in P} y_i p_i^*$ is independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
+Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
+%\[
+%g_{k+1} = g_{k} e_{\sigma(k)}^*,
+%\]
+%where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/M}$.  Now,
+%\[
+%Y_0 = B + \sum_{i \in D} g_{0i}
+%\] 
+%can be computed in $O(L)$ operations, and
+%\begin{align*}
+%Y_1 &= B + \sum_{i \in D} g_{1i} \\
+%&= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
+%&= Y_0 + \eta g_{0\sigma(0)},
+%\end{align*}
+%where $\eta = e^{j2\pi/M} - 1$.  In general,
+%\[
+%Y_{k+1} = Y_k + \eta g_{k\sigma(k)}.
+%\]
+%So, each $Y_k$ can be computed from its predecessor $Y_{k-1}$ in $O(1)$ arithmetic operations.
+Given $Y_k$, the value of $SS(f_k)$ can be computed in $O(1)$ operations using~\eqref{eq:SSfk}.
+Let $\hat{k} = \arg\min_k SS(f_k)$.
+The least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata},
+%\begin{equation}\label{eq:ahatYhat}
+%\hat{a} = \frac{1}{L} Y_{\hat{k}}.
+%\end{equation}
+%Pseudocode is given in Algorithm~\ref{alg:loglinear}.
+Line~\ref{alg_sortindices} contains the function $\operatorname{sortindices}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
+The $\operatorname{sortindices}$ function requires sorting $\abs{D}$ elements.
+This requires $O(L \log L)$ operations.
+The $\operatorname{sortindices}$ function is the primary bottleneck in this algorithm when $L$ is large.
+The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
+%In the next sections we will show how the sortindices function can be avoided.
+This leads to an algorithm that requires only $O(L)$ operations.
+%
+%\begin{algorithm}[t] \label{alg:loglinear}
+%\SetAlCapFnt{\small}
+%\SetAlTitleFnt{}
+%\caption{Mackenthun's algorithm with pilot symbols}
+%\DontPrintSemicolon
+%\KwIn{$\{y_i, i \in P \cup D \}$}
+%\For{$i \in D$ \nllabel{alg_loop_setup}}{
+%$\phi = \angle{y_i}$ \;
+%$u = \round{\phi} $ \;
+%$z_i = \phi -  u $ \;
+%$g_i = y_i e^{-j u}$ \nllabel{alg_gset} \;
+%}
+%$Y = \sum_{i \in P} y_i p_i^* + \sum_{i \in D} g_i $ \nllabel{alg_Y}\;
+%$\hat{a} = \frac{1}{L} Y$ \nllabel{alg_hata1} \;
+%$\hat{Q} = \frac{1}{L}\abs{Y}^2$ \nllabel{alg_Q} \;
+%$\eta = e^{j2\pi/M} - 1$ \;
+%$\sigma = \operatorname{sortindices}(z)$ \nllabel{alg_sortindices} \;
+%\For{$k= 0$ \emph{\textbf{to}} $M\abs{D}-1$ \nllabel{alg_loop_search}}{
+%$Y = Y + \eta g_{\sigma(k)}$ \;
+%$g_{\sigma(k)} = (\eta + 1) g_{\sigma(k)} $\;
+%$Q = \frac{1}{L}\abs{Y}^2$\;
+%\If{$Q > \hat{Q}$}{
+% 	$\hat{Q} = Q$ \;
+% 	$\hat{a} =  \frac{1}{L} Y$ \nllabel{alg_hata2} \;
+% }
+%}
+%\Return{$\hat{a}$ \nllabel{alg_return}}
+%\end{algorithm}
 
@@
 In this section we describe the asymptotic properties of the least squares estimator.
-In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
-\[
-\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right),
-\]
-where $\operatorname{round}(\cdot)$ takes its argument to the nearest integer.
+\comment{It will be informative to consider the carrier phase and amplitude estimators separately, that is, if $\hat{a} = \hat{\rho}e^{j\hat{\theta}}$, where $\hat{\rho}$ is a positive real number, then we consider the estimation errors $\sabs{\langle\theta_0 - \hat{\theta}\rangle_\pi}$ and $\sabs{\hat{\rho} - \rho_0}$.
+The function $\sfracpart{x}_\pi$ denotes $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
+\[
+\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right).
+\]
+In what follows we similarly use $\sfracpart{x}$ to denote $x$ taken `modulo $\tfrac{2\pi}{M}$' into the interval $\left[-\tfrac{\pi}{M}, \tfrac{\pi}{M}\right)$, that is
+\[
+\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right) = x - \round{x}.
+\]
+}
+
+\comment{
+It will become apparent why $\langle\theta_0 - \hat{\theta}\rangle_\pi$ rather than $\theta_0 - \hat{\theta}$ is the appropriate measure of error for the phase parameter.
+It is usual to write \emph{estimator} minus \emph{truth}, i.e., $\hat{\theta} - \theta_0$ rather than $\theta_0 - \hat{\theta}$.
+However, it is notationally convenient to use $\langle\theta_0 - \hat{\theta}\rangle_\pi$ rather than $\langle\hat{\theta} - \theta_0\rangle_\pi$ here.
+}
+
+%In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
+%\[
+%\fracpart{x}_\pi = x - 2\pi\operatorname{round}\left(\frac{x}{2\pi}\right),
+%\]
+%where $\operatorname{round}(\cdot)$ takes its argument to the nearest integer.
 The direction of rounding for half-integers is not important so long as it is consistent.
@@
 Similarly we use $\sfracpart{x}$ to denote $x$ taken `modulo $\tfrac{2\pi}{M}$' into the interval $\left[-\tfrac{\pi}{M}, \tfrac{\pi}{M}\right)$, that is
-\[
-\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right) = x - \round{x}.
-\]
+%\[
+%\fracpart{x} = x - \tfrac{2\pi}{M}\operatorname{round}\left(\tfrac{M}{2\pi}x\right) = x - \round{x}.
+%\]
+
 The next two theorems describe the asymptotic properties of the least squares estimator.  These are the central results and the chief original contributions of this paper.
@@
 \] 
-Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying
+Let $R_i \geq 0$ and $\Phi_i \in [0,2\pi)$ be real random variables satisfying (see Appendix~\ref{apdx:circ-symm-compl})
 \begin{equation}\label{eq:RiandPhii}
-R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
+R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i}
 \end{equation}
@@
 \[
-G(x) = p h_1(x) + d h_2(x) \qquad \text{where}
+G(x) = p h_1(x) + d h_2(x), \qquad \text{where}
 \]
@@
 \]
-If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$ then
+If $p > 0$ and if $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\pi,\pi)$, then
 \begin{enumerate}
 \item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
+\label{thm:consistency:part1}
 \item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
+\label{thm:consistency:part2}
 \end{enumerate}
@@
 
+\comment{CAUTION: Theorem~\ref{thm:consistency} does not include the case without pilots, i.e.\ $p=0$, which was discussed in our ICASSP paper \cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+In this case, $G(x)$ is uniquely maximised at $x = 0$ over the interval $[-\frac{\pi}{M},\frac{\pi}{M})$ such that $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ in part 1 of the theorem needs to be replaced by $\sfracpart{\theta_0 - \hat{\theta}}$.}
+
 \begin{theorem}\label{thm:normality} (Asymptotic normality)
-Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$, let
+Under the same conditions as Theorem~\ref{thm:consistency}, let $f(r,\phi)$ be the joint probability density function of $R_1$ and $\Phi_1$ (see Appendix~\ref{apdx:circ-symm-compl}), let
 \[
@@
 %If $g(\phi)$ is continuous at $\phi = \tfrac{2\pi}{M}k+\tfrac{\pi}{M}$ for each $k = 0, 1, \dots M-1$, then
-If the function $g$ is continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$, then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
+If the function $g$ is continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M{-}1$, then the distribution of $(\sqrt{L}\hat{\lambda}_L, \sqrt{L}\hat{m}_L)$ converges to the bivariate normal with zero mean and covariance matrix
 \[
@@
 \]
-\[
-A_1 = \expect R_1^2\sin^2(\Phi_1), \qquad A_2 = \expect R_1^2\sin^2\fracpart{\Phi_1},
-\]
-\[
-B_1 = \expect R_1^2 \cos^2(\Phi_1) - 1, \;\;\; B_2 = \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
-\]
+\begin{alignat*}{5}
+A_1 &= \expect R_1^2\sin^2(\Phi_1), &A_2 &= \expect R_1^2\sin^2\fracpart{\Phi_1},
+\\
+B_1 &= \expect R_1^2 \cos^2(\Phi_1) - 1,\qquad  &B_2 &= \expect R_1^2 \cos^2\fracpart{\Phi_1} - h_2^2(0).
+\end{alignat*}
 \end{theorem}
+
+\comment{
+Part~\ref{thm:consistency:part1} of Theorem~\ref{thm:consistency} shows that the phase estimation error $\langle\theta_0 - \hat{\theta}\rangle_\pi$ converges almost surely to zero as $L\rightarrow \infty$, i.e.\ the estimator $\hat{\theta}$ is strongly consistent.
+In contrast, part~\ref{thm:consistency:part2} of the theorem shows that $\hat{\rho}$ is not a consistent estimator of the true amplitude $\rho_0$ and converges to $\rho_0 G(0)$.
+The value $G(0)$ is a function of the noise distribution and $G(0)$ is close to $1$ when the variance of the noise is small, but grows without bound as the noise variance increases.
+That is, the asymptotic bias is small when the SNR is large, but significant when the SNR is small.
+We will further investigate the bias in Sec.~\ref{sec:simulations}.
+Theorem~\ref{thm:normality} shows that $\sqrt{L}(\hat{\rho} - \rho_0 G(0))$ (i.e.\ the amplitude estimation error after bias removal) and $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ are asymptotically Gaussian distributed and statistically independent as $L\rightarrow\infty$.
+}
+
+\comment{
+We would like to stress that many practical estimators such as the SNR estimators in~\cite{Pauluzzi2000} are biased and particularly at low SNR, the bias can be large.
+To the best of the authors' knowledge, the value of the bias of these estimators has only been determined by computer simulation and no theoretical analysis exists.
+In contrast, we have presented a closed-form expression for the asymptotic bias of $\hat\rho$, thereby laying the foundation for an improved estimator.
+However, removal of the bias requires knowledge of $G(0)$, which may not be available at the receiver.
+Estimation of $G(0)$ is beyond the scope of this paper.
+}
 
@@
 
-The proof of Theorem~\ref{thm:consistency} is in Section~\ref{sec:proof-almost-sure} and the proof of Theorem~\ref{thm:normality} is in Section~\ref{sec:proof-asympt-norm}.
-Before giving the proofs we discuss the assumptions made by these theorems.
+Before giving the proofs of Theorems~\ref{thm:consistency} and \ref{thm:normality} in Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm}, we discuss the assumptions made by these theorems.
 The assumption that $w_1, \dots, w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
@@
 
-Theorem~\ref{thm:normality} requires the function $g$ to be continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M-1$.
+Theorem~\ref{thm:normality} requires the function $g$ to be continuous at $\tfrac{2\pi}{M}k + \tfrac{\pi}{M}$ for each $k = 0, \dots, M{-}1$.
 This places mild restrictions on the distribution of the noise $w_i$.
-For example, the requirements are satisfied if the joint pdf of the real and imaginary parts of $w_i$ is continuous, since in this case $f(r,\phi)$ is continuous.
-Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$ it follows that $g$ has period $2\pi$ and is even on $[-\pi, \pi]$.
+For example, the requirements are satisfied if the joint probability density function (pdf) of the real and imaginary parts of $w_i$ is continuous, since in this case $f(r,\phi)$ is continuous.
+Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi)$ with respect to $\phi$ (see Appendix~\ref{apdx:circ-symm-compl}) it follows that $g$ has period $2\pi$ and is even on $[-\pi, \pi)$.
 
@@
 A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
-Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
+This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$, then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$, then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised, then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
 One can check that this assumption holds when $w_i$ is circularly symmetric and normally distributed.
@@
 
+\comment{
+Let $a = \rho e^{j\theta}$.  Slightly abusing notation by reusing $SS$, let $SS(\rho, \theta, \{d_i, i \in D\}) = SS(\rho e^{j\theta}, \{d_i, i \in D\})$. From \eqref{eq:SS},
+\begin{align}
+	SS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
+	&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:SSallparams}
+\end{align}
+%This should not cause confusion as $SS(a, \{d_i, i \in D\})$, $SS(\rho, \theta, \{d_i, i \in D\})$, and $SS(\{d_i, i \in D\})$ are easily told apart by their arguments.  
+For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ minimises $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
+\begin{equation}\label{eq:hatdfinxtheta}
+\hat{d}_i(\theta) = e^{j\hat{u}_i(\theta)}, \qquad \text{where} \qquad \hat{u}_i(\theta) = \round{\angle( e^{-j\theta}y_i)}.
+\end{equation}
+As defined here, $\hat{u}_i(\theta)$ is not strictly inside the set $\{0, \tfrac{2\pi}{M}, \dots, \tfrac{2\pi(M-1)}{M}\}$, but this is not of consequence, as we intend its value to be considered equivalent modulo $2\pi$.
+With this in mind,
+\[
+\hat{u}_i(\theta) = \round{\angle{y_i} - \theta },
+\]
+which is equivalent to the definition from~\eqref{eq:hatdfinxtheta} modulo $2\pi$ and hence, we only require to consider $\theta$ in the interval $[0, 2\pi)$.
+}
+
+\comment{
+Note from \eqref{eq:hatdfinxtheta} that $\hat{d}_i(\theta)$ does not depend on $\rho$.
+Thus, if $\hat{\theta}$ is the least squares estimator of the phase, then the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponds to the least squares estimators of the data symbols that minimise~\eqref{eq:SSdatasymbols}, i.e.\ $\{ \hat{d}_i=\hat{d}_i(\hat{\theta}), i \in D \}$.
+}
+
 Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ minimised with respect to the data symbols,
@@
 where
-\[
-Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
-\]
-and $Z^*(\theta)$ is the conjugate of $Z(\theta)$.
+\begin{align}
+Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta).
+	\label{eq:Z}
+\end{align}
 Differentiating with respect to $\rho$ and setting the resulting expression to zero gives the least squares estimator of $\rho_0$ as a function of $\theta$,
@@
 By definition the amplitude $\rho_0$ and its estimator $\hat{\rho}$ are positive.
-However, $\hat{\rho}(\theta) = \Re(Z(\theta))$ may take negative values for some $\theta \in [-\pi,\pi)$.
+However, $\hat{\rho}(\theta) = \tfrac{1}{L}\Re(Z(\theta))$ may take negative values for some $\theta \in [-\pi,\pi)$.
 %In this case we would replace $\theta$ with $\fracpart{\theta + \phi}$ and $\hat{\rho}(\theta)$ with $-\hat{\rho}(\theta)$ and observe that
@@
 %\]
-The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $SS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
+The least squares estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $SS(\theta)$ under the constraint $\hat{\rho}(\theta) = \tfrac{1}{L}\Re(Z(\theta)) > 0$.
 Equivalently $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ with no constraints required.
@@
 \end{align*}
-where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$ and where, as in Section~\ref{sec:least-squar-estim}, we consider $\hat{u}_i(\theta)$ equivalent modulo $2\pi$.
+where we put $\lambda = \fracpart{\theta_0 - \theta}_\pi$ and where, as in before, we consider $\hat{u}_i(\theta)$ equivalent modulo $2\pi$.
 Because $\hat{d}_i^*(\theta) = e^{-j\hat{u}_i(\theta)}$, it follows that, when $i \in D$,
@@
 \]
-Now,
+Now, \eqref{eq:Z} can be rewritten as
 \[
@@
 %\begin{split}
-G_L(\lambda) &= \frac{1}{\rho_0 L} \Re(Z(\theta)) \label{eq:GLdefn} \\
+G_L(\lambda) &%= \frac{\hat{\rho}(\theta)}{\rho_0}
+	= \frac{1}{\rho_0 L} \Re(Z(\theta)) \label{eq:GLdefn} \\
 &= \frac{1}{L} \sum_{i \in P} R_i \cos(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \cos\sfracpart{\lambda + \Phi_i} \nonumber
@@
 We will show that $\hat{\lambda}_L$ converges almost surely to zero as $L \rightarrow \infty$.
-The proof of part 1 of Theorem~\ref{thm:consistency} follows from this.
+The proof of part~\ref{thm:consistency:part1} of Theorem~\ref{thm:consistency} follows from this.
 
@@
 \]
-As is customary, let $\Omega$ be the sample space on which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ on which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
-Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
-Let $A'$ be the subset of the sample space on which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
+As is customary, let $\varOmega$ be the sample space on which the random variables $\{w_i\}$ are defined.
+Let $\varLambda$ be the subset of the sample space $\varOmega$ on which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
+Lemma~\ref{lem:convtoexpGlamL} in Appendix~\ref{apdx:proof-almost-sure} shows that $\prob\{\varLambda\} = 1$.
+Let $\varLambda'$ be the subset of the sample space on which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
 Because $G(x)$ is uniquely maximised at $x=0$, it follows that $G(\hat{\lambda}_L) \rightarrow G(0)$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
-So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
-Part 1 of Theorem~\ref{thm:consistency} follows.
-
-It remains to prove part 2 of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ},
+So $\varLambda \subseteq \varLambda'$ and therefore $\prob\{\varLambda'\} \geq \prob\{\varLambda\} = 1$.
+Part~\ref{thm:consistency:part1} of Theorem~\ref{thm:consistency} follows.
+
+It remains to prove part~\ref{thm:consistency:part2} of the theorem regarding the convergence of the amplitude estimator $\hat{\rho}$.  From~\eqref{eq:hatrhoZ} and \eqref{eq:GLdefn},
 \begin{equation}\label{eq:rhoGLZ}
-\hat{\rho} = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
+\hat{\rho} = \hat{\rho}(\hat{\theta}) = \frac{1}{L}\Re(Z(\hat{\theta})) = \rho_0 G_L(\hat{\lambda}_L).
 \end{equation}  
-Lemma~\ref{lem:GLtoG0} in the appendix shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+Lemma~\ref{lem:GLtoG0} in Appendix~\ref{apdx:proof-almost-sure} shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
 %It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
@@
 
-\begin{lemma}\label{lem:convtoexpGlamL} 
-$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Since $G(x)$ is uniquely maximised at $x=0$,
-\[
-0 \leq G(0) - G(\hat{\lambda}_L),
-\]
-and since $\hat{\lambda}_L$ is the maximiser of $G_L(x)$,
-\[ 
-0 \leq G_L(\hat{\lambda}_L) - G_L(0).
-\]
-Thus,
-\begin{align*}
-0 &\leq G(0) - G(\hat{\lambda}_L) \\ 
-& \leq G(0) - G(\hat{\lambda}_L) + G_L(\hat{\lambda}_L) - G_L(0) \\
-& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
-&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
-\end{align*}
-and the last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:uniflawGGL} 
-$\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-This type of result is called a~\emph{uniform law of large numbers} and follows from standard arguments.
-See, for example, van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
-A proof of this specific result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:GLtoG0}
-$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-By the triangle inequality,
-\[
-\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
-\]
-Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ by Lemma~\ref{lem:convtoexpGlamL}.
-\end{IEEEproof}
-
 \section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
 
+The proof of Theorem~\ref{thm:normality} hinges on Lemmas \ref{lem:diffatlambdaL}--\ref{lem:XL}, which can be found in Appendix~\ref{apex:proof-asymptotic-normality}.
 We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
@@
 Define the function
-\begin{equation}\label{eq:RLdef}
-R_L(\lambda) = \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.
-\end{equation}
+\begin{align}\label{eq:RLdef}
+R_L(\lambda) &= \frac{1}{\rho_0 L} \Im(Z(\theta))
+	\\
+	&= \frac{1}{L} \sum_{i \in P} R_i \sin(\lambda + \Phi_i) + \frac{1}{L} \sum_{i \in D} R_i \sin\sfracpart{\lambda + \Phi_i}.\notag\end{align}
 Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$, and so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
@@
 \]
-where $o_P(1)$ denotes a sequence of random variables converging in probability to zero as $L \rightarrow \infty$, and $p,d$ and $H$ are defined in the statement of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
+where $o_P(1)$ denotes a sequence of random variables converging in probability to zero as $L \rightarrow \infty$, and $p,d$ and $H$ are defined in the statements of Theorems~\ref{thm:consistency}~and~\ref{thm:normality}.
 Lemma~\ref{lem:empiricprocc} shows that
@@
 \]
-It follows from the three equations above that,
+It follows from the three equations above that
 \[
@@
 \]
-and rearranging gives,
+and rearranging gives
 \[
@@
 \]
-Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$ where $A_1$ and $A_2$ are defined in the statement of Theorem~\ref{thm:normality}.
+Lemma~\ref{lem:convdistGLdash} shows that the distribution of $\sqrt{L}R_L(0)$ converges to the normal with zero mean and variance $pA_1 + dA_2$, where $A_1$ and $A_2$ are defined in the statement of Theorem~\ref{thm:normality}.
 It follows that the distribution of $\sqrt{L}\hat{\lambda}_L$ converges to the normal with zero mean and variance
@@
 It follows that $\sqrt{L} \hat{m}_L =  \rho_0 X_L + o_P(1)$.
-Lemma~\ref{lem:XL} shows that the distribution of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$ where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
+Lemma~\ref{lem:XL} shows that the distribution of $X_L$ converges to the normal with zero mean and variance $p B_1 + d B_2$ as $L\rightarrow\infty$, where $B_1$ and $B_2$ are defined in the statement of Theorem~\ref{thm:normality}.
 Thus, the distribution of $\sqrt{L} \hat{m}_L$ converges to the normal with zero mean and variance $\rho_0^2(p B_1 + d B_2)$ as required.
@@
 \]
-as $L \rightarrow \infty$.  The lemmas that we have used are proved in the appendix.
-
-\section{The Gaussian noise case}\label{sec:gaussian-noise-case}
-
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
-\[
-\frac{1}{2\pi\sigma^2}e^{-\frac{1}{2\sigma^2}(x^2 + y^2)}.
-\]
-Theorems~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
-%and the joint density function of the real and imaginary parts of $Z$ is
+as $L \rightarrow \infty$.
+
+%\section{The Gaussian noise case \comment{-- remove or expand?}}\label{sec:gaussian-noise-case}
+%
+%Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
 %\[
-%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
+%\frac{1}{2\pi\sigma^2}e^{-\frac{1}{2\sigma^2}(x^2 + y^2)}.
 %\]
-%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
-It can be shown that
-\[
-%\begin{equation}\label{eq:gfornormal}
-%g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-%g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
-g(\phi) = \frac{\cos\phi}{2\pi}e^{-\kappa} + \frac{\Psi(\sqrt{2\kappa} \cos\phi)}{\sqrt{\pi\kappa}}  e^{-\kappa\sin^2\phi} \big(2 + \kappa\cos^2\phi \big)
-%\end{equation}
-\]
-where $\kappa = \tfrac{\rho_0^2}{2\sigma^2}$ and $\Psi(t) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{t} e^{-x^2/2} dx$
-is the cumulative density function of the standard normal.
+%Theorems~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+%%and the joint density function of the real and imaginary parts of $Z$ is
+%%\[
+%%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
+%%\] 
+%%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
+%It can be shown that
 %\[
-%\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos\phi}{\sqrt{2}}\right)
+%%\begin{equation}\label{eq:gfornormal}
+%%g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+%%g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
+%g(\phi) = \frac{\cos\phi}{2\pi}e^{-\kappa} + \frac{\Psi(\sqrt{2\kappa} \cos\phi)}{\sqrt{\pi\kappa}}  e^{-\kappa\sin^2\phi} \big(2 + \kappa\cos^2\phi \big)
+%%\end{equation}
 %\]
-%and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  
-The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using this formula.
+%where $\kappa = \tfrac{\rho_0^2}{2\sigma^2}$ and $\Psi(t) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{t} e^{-x^2/2} dx$
+%is the cumulative density function of the standard normal.
+%%\[
+%%\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos\phi}{\sqrt{2}}\right)
+%%\]
+%%and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  
+%The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using this formula.
 
@@
 Under these conditions the least squares estimator is also the maximum likelihood estimator.
-Simulations are run with $M=2,4,8$ (BPSK, QPSK, $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.
+Simulations are run with $M=2$, $4$, $8$ (BPSK, QPSK, $8$-PSK) and with signal to noise ratio $\text{SNR} = \tfrac{\rho_0^2}{2\sigma^2}$ between \unit[-20]{dB} and \unit[20]{dB} in steps of \unit[1]{dB}.
 The amplitude $\rho_0=1$ and $\theta_0$ is uniformly distributed on $[-\pi, \pi)$.
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.
+Figs.~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2$, $4$, $8$ with $L=4096$ and for varying proportions of pilot symbols $\abs{P} = 0$, $\tfrac{L}{32}$, $\tfrac{L}{8}$, $\tfrac{L}{2}$, $L$.
 When $\abs{P} \neq 0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
 Otherwise, when $\abs{P}=0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}^2$ as in~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least square estimator.
-The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality}.
+The dots, squares, circles and crosses are the results of Monte-Carlo simulations with the least squares estimator.
+\comment{The Cram\'er-Rao bounds (CRB) for the fully data-aided case with $\abs{P}=L$ are shown as a reference.
+Note that the MSE curves for the least squares estimator for $\abs{P}=L$ coincide with the CRBs.}  The solid lines are the estimator MSEs predicted by Theorem~\ref{thm:normality}.
 %(and by Theorem~2 from~\cite{McKilliam_leastsqPSKnoncoICASSP_2012} when $\abs{P} = 0$).
@@
 The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large.
-As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-Theorem~\ref{thm:normality} does not model this behaviour in the sense that for any fixed $L$ there exist sufficiently small values of SNR for which Theorem~\ref{thm:normality} does not produce accurate predictions of the MSE.
-As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e.
-$\abs{P} = L$.
+As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P}>0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+As the SNR increases the variance of the estimators converges to that of the estimator where all symbols are pilots, i.e.\ $\abs{P} = L$.
 
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the noncoherent phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
+Figs.~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non-data aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
 This estimator requires selection of a function $F$ that transforms the amplitude of each sample prior to the final estimation step.
@@
 Viterbi and Viterbi propose several viable alternatives, from which we have chosen $F(x) = 1$.
-The Viterbi and Viterbi estimator is only applicable in the noncoherent setting, i.e.
-when $\abs{P} = 0$.
+The Viterbi and Viterbi estimator is only applicable in the non-data aided setting, i.e.\ when $\abs{P} = 0$.
 The sample MSE of the least squares estimator (when $\abs{P} = 0$) and the Viterbi and Viterbi estimator is similar.
@@
 
-Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
-shows the variance of the amplitude estimator when $M=2$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+\comment{Replace variance plot by MSE plot and rewrite/shorten para.}  Fig.~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
+shows the variance of the amplitude estimator when $M=2$ and with $L=32$, $256$, $2048$ and when the number of pilots symbols is $\abs{P} = 0$, $\tfrac{L}{2}$, $L$.
 The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
@@
 This requires $G(0)$ to be known.
-In practice $G(0)$ may not be known at the receiver, so Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK}
+In practice $G(0)$ may not be known at the receiver, so Fig.~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK}
 serves to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
@@
 However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
-%As indicated in Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the estimator when all symbols are pilots.
+%As indicated in Figs.~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} the variance of the least squares amplitude estimator is smaller than that of the estimator when all symbols are pilots.
 However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.
 
-Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
+\comment{Add a plot of the theoretical bias $\rho_0 (G(0)-1)$ and the simulated bias $\tfrac{1}{T}\sum_{i=1}^T(\hat\rho_i-\rho_0)$ as a function of SNR for different values of $L$?}
+
+Fig.~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32$, $256$, $2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8}$, $L$.
 The figure depicts an interesting phenomenon.
@@
 
+\comment{Add packet error rate results for an ASRP ST1 uplink:
+\begin{itemize}
+\item
+	single user, bursts of 246 QPSK symbols, preamble of 14 pilots, symbol rate 1024Hz, burst duration 250ms, oversampling factor 32
+\item
+	iterative refinement of estimates using LDPC decoder output (i.e.\ turbo sync)
+\item
+	genie-aided synchronisation of all other channel parameters? Our current ST1 framework doesn't have this functionality.
+\item
+	vary number of pilots?
+\item
+	performance comparison with Viterbi Viterbi initialisation when $\abs{P}=0$ and purely data-aided initialisation when $\abs{P}>0$
+\end{itemize}
+}
+
 \begin{figure}[p]
@@
 		\includegraphics[width=\linewidth]{code/data/plotM2-2.mps}
-		\caption{Phase error versus SNR for BPSK with $L=4096$.}
+		\caption{Phase error versus SNR for BPSK with $L=4096$. \comment{Add Viterbi Viterbi curve and include data-aided CRB.}}
 		\label{fig:plotphaseBPSK}
@@
 		\includegraphics[width=\linewidth]{code/data/plotM4-2.mps}
-		\caption{Phase error versus SNR for QPSK with $L=4096$.}
+		\caption{Phase error versus SNR for QPSK with $L=4096$. \comment{Include data-aided CRB.}}
 		\label{fig:plotphaseQPSK}
@@
 		\includegraphics[width=\linewidth]{code/data/plotM8-2.mps}
-		\caption{Phase error versus SNR for $8$-PSK with $L=4096$.}
+		\caption{Phase error versus SNR for $8$-PSK with $L=4096$. \comment{Include data-aided CRB.}}
 		\label{fig:plotphase8PSK}
@@
 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-		\caption{Amplitude estimator variance (bias removed) versus SNR for BPSK.}
+		\caption{Amplitude estimator variance (bias removed) versus SNR for BPSK. \comment{Change to MSE.}}
 		\label{fig:plotampBPSK}
@@
 		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
-		\caption{Phase error versus SNR for QPSK.}
+		\caption{Phase error versus SNR for QPSK. \comment{Include data-aided CRBs.}}
 		\label{fig:plotphaseQPSKmultL}
@@
 
-Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
-Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
+Monte-Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
+Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Fig.~\ref{fig:plotphaseQPSKmultL}).
 %This suggests the use of a weighted objective function, which would be an interesting topic for future research.
@@
 \normalsize
-\appendix
+\appendices
+
+\section{Circularly symmetric complex random variables}\label{apdx:circ-symm-compl}
+
+%Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
+A complex random variable $W$ is said to be \emph{circularly symmetric} if its phase $\angle{W}$ is independent of its magnitude $\abs{W}$ and if the distribution of $\angle{W}$ is uniform on $[0,2\pi)$.
+That is, if $Z \geq 0$ and $\Theta \in [0,2\pi)$ are real random variables such that $Ze^{j\Theta} = W$, then $\Theta$ is uniformly distributed on $[0,2\pi)$ and is independent of $Z$.
+If the pdf of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
+
+Let $Z \geq 0$ and $\Theta \in [0,2\pi)$ be real random variables and let $\Theta$ be uniformly distributed on $[0,2\pi)$ and independent of $Z$.
+Then the complex random variable $W=Ze^{j\Theta}$ is said to be \emph{circularly symmetric}.
+If the pdf of $Z$ is $f_Z(z)$, then the joint pdf of $\Theta$ and $Z$ is
+\[
+f_{Z,\Theta}(z,\theta) = \frac{1}{2\pi}f_Z(z).
+\]
+% Observe that for any real number $\phi$, the pdf of $W$ and $e^{j\phi}W$ are the same, that is, the pdf is invariant to phase rotation.
+If $\expect\abs{W} = \expect Z$ is finite, then $W$ has zero mean because
+% \begin{align*}
+%  \expect W &= \int_{0}^{2\pi} \int_{0}^\infty z e^{j\theta} f_{Z,\Theta}(z,\theta) dz d\theta \\
+%  &= \frac{1}{2\pi} \int_{0}^{2\pi} e^{j\theta} \int_{0}^\infty z f_Z(z) dz d\theta \\
+%  &= \frac{1}{2\pi}\expect Z \int_{0}^{2\pi} e^{j\theta} d\theta = 0.
+%  \end{align*}
+% If $X$ and $Y$ are real random variables equal to the real and imaginary parts of $W = X + jY$ then the joint pdf of $X$ and $Y$ is
+% \[
+% f_{X,Y}(x,y) = \frac{f_Z(\sqrt{x^2 + y^2})}{2\pi \sqrt{x^2 + y^2}}.
+% \]
+%from which follows
+%\begin{equation}\label{eq:fzandfZYwithz}
+%f_Z(z) = 2\pi z f_{X,Y}(z,0) = 2\pi z f_{X,Y}(0,z)
+%\end{equation}
+%for all $z > 0$.
+
+We will have particular use of complex random variables of the form $1 + W$, where $W$ is circularly symmetric.  Let $R \geq 0$ and $\Phi \in [0,2\pi)$ be real random variables satisfying
+\[
+R e^{j\Phi} = 1 + W.
+\]
+The joint pdf of $R$ and $\Phi$ can be shown to be
+ \begin{equation}\label{eq:pdfRPhi}
+ f(r,\phi) = \frac{r f_Z(\sqrt{r^2 - 2r\cos\phi + 1})}{2\pi\sqrt{r^2 - 2r\cos\phi + 1}}.
+ \end{equation}
+Since $\cos\phi$ has period $2\pi$ and is even on $[-\pi,\pi)$ it follows that $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi)$ with respect to $\phi$.
+The mean of $R e^{j\Phi}$ is equal to one because the mean of $W$ is zero.
+So,
+\begin{align}%\label{eq:expectRepartRphi}
+\expect \Re(R e^{j\Phi}) = \expect R \cos(\Phi) = 1,
+\notag\\
+\label{eq:expectImpartRphi}
+\expect \Im(R e^{j\Phi}) = \expect R \sin(\Phi) = 0,
+\end{align}
+where $\Re(\cdot)$ and $\Im(\cdot)$ denote the real and imaginary parts.
+
+% \begin{lemma}\label{lem:h1minedcircsym}
+% Let $X = Z e^{j\Theta}$ be a circularly symmetric complex random variable with pdf $\tfrac{1}{2\pi}f_Z(z)$.
+Let $R > 0$ and $\Phi \in [0, 2\pi)$ be random variables satisfying $R e^{j\Phi} = 1 + X$.
+Let
+% \[
+% h_1(x) = \expect R \cos(x + \Phi).
+% \]
+% Then $h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$.
+% \end{lemma}
+% % Before we begin the proof note that the requirement for $z^{-1}f_{Z}(z)$ to be non increasing implies that the probability density function of $Z e^{j \Theta}$ decreases as we move away from the origin.
+That is, the pdf of $Z e^{j \Theta}$ in rectangular coordinates is given by $z^{-1}f_{Z}(z)$, where $z = \sqrt{x^2 + y^2}$ and $x$ and $y$ are the real and imaginary parts of $Z e^{j \Theta}$, and this pdf is non increasing with $z$.
+For example, the zero mean complex Gaussian distribution with independent real and imaginary parts satisfies this requirement.
+
+% % By the phrase ``$h_1(x)$ is uniquely maximised at $0$ over the interval $[-\pi,\pi]$'' it is meant that for any $\delta > 0$ there exists an $\epsilon > 0$ such that for those $x \in [-\pi, \pi]$, if
+% % \[
+% % \abs{x} > \delta \qquad \text{then} \qquad  h_1(0) - h_2(x) > \epsilon.
+% % \]
+% % We will have further use of this definition in Section~\ref{sec:stat-prop-least}.  We are now ready to prove Lemma~\ref{lem:h1minedcircsym}.
+% \begin{IEEEproof}
+% BLERG
+% \end{IEEEproof}
 
 %\subsection{Lemmas required for the proof of almost sure convergence (Theorem~\ref{thm:consistency}) } \label{sec:proof-almost-sureappendix}
+
+\section{Lemmas for the proof of Theorem~\ref{thm:consistency}} \label{apdx:proof-almost-sure}
+
+\begin{lemma}\label{lem:convtoexpGlamL} 
+$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Since $G(x)$ is uniquely maximised at $x=0$,
+\[
+0 \leq G(0) - G(\hat{\lambda}_L),
+\]
+and since $\hat{\lambda}_L$ is the maximiser of $G_L(x)$,
+\[ 
+0 \leq G_L(\hat{\lambda}_L) - G_L(0).
+\]
+Thus,
+\begin{align*}
+0 &\leq G(0) - G(\hat{\lambda}_L) \\ 
+& \leq G(0) - G(\hat{\lambda}_L) + G_L(\hat{\lambda}_L) - G_L(0) \\
+& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
+&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)}.
+\end{align*}
+%The last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}. 
+\comment{The last line converges almost surely to zero.
+This type of result is called a~\emph{uniform law of large numbers} and follows from standard arguments.
+See, for example, van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
+A proof of this specific result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
+}
+\end{IEEEproof}
+
+%\begin{lemma}\label{lem:uniflawGGL} 
+%$\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+%\end{lemma}
+%\begin{IEEEproof}
+%This type of result is called a~\emph{uniform law of large numbers} and follows from standard arguments.
+See, for example, van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
+A proof of this specific result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
+%\end{IEEEproof}
+
+\begin{lemma}\label{lem:GLtoG0}
+$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+By the triangle inequality,
+\[
+\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
+\]
+Now by Lemma~\ref{lem:convtoexpGlamL}, $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ \comment{almost surely?} as $L \rightarrow \infty$ and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+%Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ by Lemma~\ref{lem:convtoexpGlamL}.
+\end{IEEEproof}
+
+\section{Lemmas for the proof of Theorem~\ref{thm:normality}} \label{apex:proof-asymptotic-normality}
 
@@
 \[
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
+SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2
 \]
@@
 \begin{IEEEproof}
-Put $A(\lambda) = \expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)$.  Recalling that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$ we have
+Put $A(\lambda) = \expect R_1 e^{j\fracpart{\Phi_1}} \chi(\Phi_1,\lambda)$.  Recalling from \eqref{eq:pdfRPhi} that $f(r,\phi)$ is the joint pdf of $R_1$ and $\Phi_1$ we have
 \begin{align*}
@@
 \]
-From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
+From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\,\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi} in Appendix~\ref{apdx:circ-symm-compl}, and variance
 \[

### 2013-11-16 paper.tex
added to intro

@@
 We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent and asymptotically normally distributed.
-However, the amplitude estimator is not consistent and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
+However, the amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
+This appears to be the first time that the statistical properties of a non-data aided estimator for carrier amplitude have been analysed theoretically.
 %Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.  
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Sweldens considered what is called \emph{noncoherent detection} (also know as \emph{blind}, \emph{non-data aided} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
+Both Mackenthun and Sweldens considered what is called \emph{non-data adied detection} (also know as \emph{blind}, \emph{noncoherent} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
 We show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
-Our model includes the noncoherent case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-
-Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least square estimator of carrier amplitude and phase obtained by minimising~\eqref{eq:SSdefn}.
-In the literature, analytic studies of non-data aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
-The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994}, or to analyse bound that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
+Our model includes the non-data aided case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
+
+Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:SSdefn}.
+In the literature, analytic study of non-data aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
+The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
 To the authors knowledge, all of the existing bounds are applicable only in the fully non-data aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
@@
 For the case of carrier amplitude the situations is even more complicated.
-A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that all known non-data aided estimators for carrier amplitude exhibit bias.
+A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non-data aided estimators for carrier amplitude exhibit bias.
 The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
@@
 Rather than compute bounds, we directly study the least squares estimators of carrier phase and amplitude asymptotically as the number of transmitted symbols $L\to \infty$.
-We show, under reasonably mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent (asymptotically unbiased) and asymptotically normally distributed.
+We show, under comparatively mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent (asymptotically unbiased) and asymptotically normally distributed.
 However, the least squares amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation $M$.
-To our knowledge, this is the first time an analytic description of the bias or variance of non-data aided amplitude estimator has been obtained.
+To our knowledge, this is the first time an analytic description of the bias or variance of a non-data aided amplitude estimator has been obtained.
 Our expressions for the asymptotic variance and bias are simple to compute.
@@
 One criticism that might be directed at our asymptotic approach when compared to the Cramer-Rao bounds already developed in the literature is that our specific results apply only to the least square estimator.
-One response to this criticism is that, because the least squares estimator can be rapidly computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian, then it is likely to be a popular practical choice.
-Perhaps an even more convinving response is that, while the specific Theorems~\ref{thm:consistency} and~\ref{thm:normality} that we given apply only to the least square estimtor, the techniques that we develop to derive these results can likely be deployed for the analysis of other estimators.
-For example, it would be reasonably straigthforward to extend our results to allow data and pilot symbols with varying constellation size, i.e.
-varying $M$.
-For example, one might give more importance to certain data symbols and use BPSK ($M=2$) for these, but QPSK ($M=4$) for other less important symbols.
-This is related to what is called \emph{unequal error protection} in the literature~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010} and is commonly used in practical systems.
-Perhaps a more involved extension would be the inclusion of carrier frequency as a parameter to be estimated.
-The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009}, would potentially allow this.
+One response to this criticism is that the least squares estimator should be a popular practical choice because it can be efficiently computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian.
+Perhaps an even more convinving response is that, while the specific results given here apply only to the least square estimtor, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
+For example, it is common in practical systems to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.
+This is related to what is called \emph{unequal error protection} in the literature~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.
+It would be reasonably straigthforward to extend our results to allow data and pilot symbols with varying constellation size.
+A more involved extension would be the inclusion of carrier frequency as a parameter to be estimated.
+The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010}, would potentially allow this.
+
+More ambitiously, our techniques might allow analysis of non-data aided channel estimators for modern communications systems such as mulitple anntannae systems~\cite{4138046}, and systems employing orthogonal frequency division multiplexing.
+We see the results presented here as a starting point for an potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.
+
+That this paper contains novel theoretical developments should not suggest to the reader that our results are not of immediate practical interest.
+The design of our estimator was motivated by a project at the University of South Australia's Institute for Telecommunications Research to develop a narrowband multiuser communication system to collect sensor network data using a satellite in orbit.
+This estimator has been implemented on a prototype that was successfully trialled from April to July of 2013 with signals from multiple low power ground-based transmitters being successfully received and decoded by both aircraft and satellites in orbit~\cite{ASRPpromovideo}.
+Our theoretical analysis aided design of this system.
+Based on our Theorem~\ref{thm:normality} in Section~\ref{sec:stat-prop-least} the performance of the least squares carrier phase and amplitude estimators can be accurately predicted without the need for time consuming Monte-Carlo simulations.
+This provides easy to use and rigourously justified methods for choosing systems parameters such as the number and placement of pilot symbols and also provides useful information to other reciever components, such as channel tracking systems.
+
+Recently, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
+These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
+These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
+A key problem is that the EM algorithm converges correctly only if initialised at some channel estimate sufficiently close to the true channel parameters.
+Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
+For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe in the paper is a good candidate as an initialiser for a turbo synchronisation proceedure.
+%However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by applying the EM algorithm will not justify the extra computational cost.
 
@@
 The paper is organised in the following way.
-Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the coherent case, when both pilot symbols and data symbols exist.
+Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the case when both pilot symbols and data symbols exist.
 Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
-Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase $\hat{\theta}$ and amplitude $\hat{\rho}$.
-We show, under reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
+Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase and amplitude.
+%We show, under reasonably general assumptions about the distribution of the noise $w_1,\dots,w_L$, that $\langle\theta_0 - \hat{\theta}\rangle_\pi$ converges almost surely to zero and that $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ is asymptotically normally distributed as $L\rightarrow \infty$.
 However, $\hat{\rho}$ is not a consistent estimator of the amplitude $\rho_0$.
@@
 Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} provide proofs of the theorems stated in Section~\ref{sec:stat-prop-least}.
-In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
+%In Section~\ref{sec:gaussian-noise-case} we consider the special case when the noise is Gaussian.
 In this case, our expressions for the asymptotic distribution can be simplified.
@@
 In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the noncoherent setting, so we modify the algorithm to include the pilot symbols.
+Mackenthun specifically considered the non-data aided setting, so we modify the algorithm to include pilot symbols.
 For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the noncoherent phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non-data aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
 This estimator requires selection of a function $F$ that transforms the amplitude of each sample prior to the final estimation step.
@@
 Viterbi and Viterbi propose several viable alternatives, from which we have chosen $F(x) = 1$.
-The Viterbi and Viterbi estimator is only applicable in the noncoherent setting, i.e.
+The Viterbi and Viterbi estimator is only applicable in the non-data aided setting, i.e.
 when $\abs{P} = 0$.
@@
 %When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
-This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
+% This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
 For example, rather than minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
-\[
-SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
-\]
-where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
+% \[
+% SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
+% \]
+% where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
 Computing the $\hat{a}$ that minimises $SS_\beta$ can be achieved with only a minor modification to algorithm~\ref{alg:loglinear}.

### 2013-11-16 paper.tex
added minor

@@
 
-More ambitiously, our techniques might allow analysis of non-data aided channel estimators for modern communications systems such as mulitple anntannae systems~\cite{4138046}, and systems employing orthogonal frequency division multiplexing.
-We see the results presented here as a starting point for an potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.
+More ambitiously, our techniques might allow analysis of non-data aided channel estimators for modern communications systems such as mulitple anntannae systems, and systems employing orthogonal frequency division multiplexing~\cite{4138046,Heath99exploitinginput}.
+We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.
 

### 2013-11-16 response.tex
added to paper

@@
 
-%\usepackage{xr}
-%\externaldocument{paper2}
+\usepackage{zref-xr,zref-user}
+\zexternaldocument*{paper}
 
@@
 	\\\textbf{Response:}
+We have shorteed Section~\zref{sec:least-squar-estim} but have left the description of the algorithm intact.
+The primary reason for this is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.
+To qualify this for the reader we have included the following sentences at the beginning of Section~\zref{sec:least-squar-estim}.
+
+``
+In this section we derive Mackentun's algorithm to compute the least squares estimators of the carrier phase and amplitude~\cite{Mackenthun1994}.
+Mackenthun specifically considered the non-data aided setting, so we modify the algorithm to include pilot symbols.
+Although the modification is straightforward, much of the notation that we will use for analysing the estimator in Section~\ref{sec:stat-prop-least} is naturally introduced during the description of the algorithm.
+For this reason, and so that our paper is self contained, we describe the algorithm in full.
+''
+
 %	We have shortened Sec.\ II to make it more concise and improve readability.
@@
 	\\\textbf{Response:}	
-%	For improved readability of the manuscript, we have now moved all of the lemmas into the appendix.
-We have also added a discussion of the implications and the practical relevance of the theorems to Sec.\ III (Sec.\ IV in the original manuscript).
+	For improved readability of the manuscript, we have now moved all of the lemmas into the appendix.
+We have also added a discussion of the implications and the practical relevance of the theorems after the statement of Theorems~\zref{thm:consistency} and~\zref{thm:normality} in Section.~\zref{sec:stat-prop-least}.
 
@@
 	\\\textbf{Response:}
-%	We agree with the reviewer and apologise for causing confusion with the use of the terms \emph{coherent} and \emph{noncoherent} in the original manuscript.
-In the updated manuscript, we refer to an estimator that uses known pilot symbols as \emph{data-aided} and to one that relies on receiver decisions on the data symbols as \emph{decision-directed}.
+	We agree with the reviewer and apologise for causing confusion with the use of the terms \emph{coherent} and \emph{noncoherent} in the original manuscript.
+In the updated manuscript, we refer to an estimator that does not use pilot symbols as \emph{non data-aided}.
+%In the updated manuscript, we refer to an estimator that uses known pilot symbols as \emph{data-aided} and to one that relies on receiver decisions on the data symbols as \emph{decision-directed}.
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:} 
-%We thank the reviewer for spotting the missing comma and have added it.
+We thank the reviewer for spotting the missing comma and have added it.
 \item\textbf{Comment:} 
 	p.\ 19, at the bottom: sortindicIes (typo, delete 'i')
-	\\\textbf{Response:} % Thank you for pointing out this typographical error. We have corrected it.
+	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
 \item\textbf{Comment:} 
 	In Section VI, I would prefer a reference to the Lemma's in the Appendix at the start of the section.
-	\\\textbf{Response:} %We have added a reference to the Lemmas in the appendix at the beginning of the section. Please note that Sec.\ VI of the original manuscript is now Sec.\ III.
+	\\\textbf{Response:} We have added a reference to the Lemmas in the appendix at the beginning of the section. %Please note that Sec.\ VI of the original manuscript is now Sec.\ III.
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-%	We have removed the aforementioned sentence from the manuscript as we feel it has caused more confusion than provided insight.
+	We have removed the aforementioned sentence from the manuscript as we feel it has caused more confusion than provided insight.
 We hope the following explanation clears up any confusion it may have caused.
@@
 
-%	\comment{Add CRBs to plots.} Please note that Sec.\ VIII of the original manuscript is now Sec.\ VI.
+	\comment{Add CRBs to plots.}
 \end{enumerate}
@@
 	Estimation of carrier phase and amplitude for $M$-ary PSK signals has been studied by the authors.
-Extending MackenthunÕs algorithm to pilots, a least square (LS) method is developed, where major parts of the manuscript are dedicated to the statistical analysis of the LS approach.
-
+Extending Mackenthun's algorithm to pilots, a least square (LS) method is developed, where major parts of the manuscript are dedicated to the statistical analysis of the LS approach.
 Generally speaking it is to admit that both wording and writing follow the IEEE standards, although I would say that the list of References should be checked and updated in this respect.
 	\\\textbf{Response:}
-%	We thank the reviewer for pointing us to the suggested references and have added \comment{some/most of} them. Specifically, we have included references [c, d, e]. \comment{Add the remaining ones?}
+	We thank the reviewer for pointing us to the suggested references and have added them.
 \end{enumerate}
@@
 	Paragraph before (13): \dots can be computed from its predecessor \dots
-	\\\textbf{Response:} %Thank you for pointing out this typographical error. We have corrected it.
+	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-%	We thank the reviewer for pointing out the missing definition of $f(r, \phi)$ and have added it.
-Please note that in response to comment \ref{R4it2e} of reviewer 4, Sec.\ III has been slightly shortened and has been moved to the appendix.
+	We thank the reviewer for pointing out the missing definition of $f(r, \phi)$ and have added it.
+%Please note that in response to comment \ref{R4it2e} of reviewer 4, Sec.\ III has been slightly shortened and has been moved to the appendix.
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-%	Please refer to our response to comment \ref{R1it3a} of reviewer 1.
+	Please refer to our response to comment \ref{R1it3a} of reviewer 1.
 \end{enumerate}
@@
 	\\\textbf{Response:}
-%	We have added a brief discussion of the need for sufficiently accurate coarse estimates to initialise an iterative code-aided synchronisation stage, citing [d, e].
-We have also added a set of new simulation results, which show the packet error rate of an LDPC-coded satellite communication system to demonstrate the use and performance of the proposed estimator in a practical system.
+	We have added the following paragraph to the introduction describing the need for sufficiently accurate coarse estimates to initialise an iterative code-aided synchronisation stage. %citing [d, e].
+
+``In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
+These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
+These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
+A key problem is that the EM algorithm converges correctly only if initialised at some channel estimate sufficiently close to the true channel parameters.
+Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
+For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.''
+
+%We have also added a set of new simulation results, which show the packet error rate of an LDPC-coded satellite communication system to demonstrate the use and performance of the proposed estimator in a practical system.
 \comment{Add details\dots}
@@
 	\\\textbf{Response:}
-%	We thank the reviewer for the comments. Please refer to our response to comment \ref{R1it2} of reviewer 1.
+We thank their reviewer for the comments. We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our paper clear.
 \end{enumerate}
@@
 	\\\textbf{Response:}
-%	We have shortened Sec.\ II to make it more concise and improve readability. Sec.\ VII has been removed from the revised manuscript.
+%	We have shortened Sec.\ II to make it more concise and improve readability. 
+\comment{BLERG}
 
+The original Sec.\ VII has been removed from the revised manuscript.
+
+\comment{BLERG}
 %	Furthermore, we agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that Sec.\ III specifically discusses random variables of the form $Re^{j\varPhi}=1{+}W$, where $W$ is circularly symmetric.

### 2013-11-16 paper.tex
added to paper

@@
 However, the amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
-This appears to be the first time that the statistical properties of a non-data aided estimator for carrier amplitude have been analysed theoretically.
+This appears to be the first time that the statistical properties of a non data-aided estimator for carrier amplitude have been analysed theoretically.
 %Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.  
@@
 Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the receiver and must also be estimated.
-Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
+Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990,580211,974597,Mengali_andre_synchro_book}.
 A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
@@
 
-% The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} (also know as \emph{blind}, \emph{non-data aided} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
+% The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} (also know as \emph{blind}, \emph{non data-aided} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
 In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
@@
 We show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
-Our model includes the non-data aided case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
+Our model includes the non data-aided case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
 
 Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:SSdefn}.
-In the literature, analytic study of non-data aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
+In the literature, analytic study of non data-aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
 The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
-To the authors knowledge, all of the existing bounds are applicable only in the fully non-data aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
+To the authors knowledge, all of the existing bounds are applicable only in the fully non data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
 Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.
@@
 For the case of carrier amplitude the situations is even more complicated.
-A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non-data aided estimators for carrier amplitude exhibit bias.
+A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non data-aided estimators for carrier amplitude exhibit bias.
 The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
 Neverthess, Cramer-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
-Unfortunately, these bounds may not be meaningful since known non-data aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
+Unfortunately, these bounds may not be meaningful since known non data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
 
@@
 However, the least squares amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation $M$.
-To our knowledge, this is the first time an analytic description of the bias or variance of a non-data aided amplitude estimator has been obtained.
+To our knowledge, this is the first time an analytic description of the bias or variance of a non data-aided amplitude estimator has been obtained.
 Our expressions for the asymptotic variance and bias are simple to compute.
@@
 One response to this criticism is that the least squares estimator should be a popular practical choice because it can be efficiently computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian.
-Perhaps an even more convinving response is that, while the specific results given here apply only to the least square estimtor, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
-For example, it is common in practical systems to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.
+Perhaps an even more convinving response is that, while the specific results given here apply only to the least square estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
+For example, in practical systems it is sometimes desirable to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.
 This is related to what is called \emph{unequal error protection} in the literature~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.
@@
 
-More ambitiously, our techniques might allow analysis of non-data aided channel estimators for modern communications systems such as mulitple anntannae systems, and systems employing orthogonal frequency division multiplexing~\cite{4138046,Heath99exploitinginput}.
+More ambitiously, our techniques might allow analysis of non data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,Heath99exploitinginput,inthft:204506}.
 We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.
 
-That this paper contains novel theoretical developments should not suggest to the reader that our results are not of immediate practical interest.
+That this paper contains novel theoretical developments should does not imply that our results are of no immediate practical interest.
 The design of our estimator was motivated by a project at the University of South Australia's Institute for Telecommunications Research to develop a narrowband multiuser communication system to collect sensor network data using a satellite in orbit.
@@
 
-Recently, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
+In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
 These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
@@
 Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe in the paper is a good candidate as an initialiser for a turbo synchronisation proceedure.
+For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.
 %However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by applying the EM algorithm will not justify the extra computational cost.
@@
 
-In this section we derive Mackentun's algorithm to compute the least squares estimator of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the non-data aided setting, so we modify the algorithm to include pilot symbols.
+In this section we derive Mackenthun's algorithm to compute the least squares estimators of the carrier phase and amplitude~\cite{Mackenthun1994}.
+Mackenthun specifically considered the non data-aided setting, so we modify the algorithm to include pilot symbols.
+Although the modification is straightforward, much of the notation that we will use for analysing the estimator in Section~\ref{sec:stat-prop-least} is naturally introduced during the description of the algorithm.
+For this reason, and so that our paper is self contained, we describe the algorithm in full.
 For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 \begin{equation}\label{eq:hata}
-\hat{a}(\{d_i, i \in D\}) = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y
+\hat{a}(\{d_i, i \in D\}) = \frac{1}{L} \sum_{i \in P \cup D} y_i s_i^* = \frac{1}{L} Y,
 \end{equation}
@@
 \]
-So, each $Y_k$ can be computed from it predecessor $Y_{k-1}$ in $O(1)$ arithmetic operations.
+So, each $Y_k$ can be computed from its predecessor $Y_{k-1}$ in $O(1)$ arithmetic operations.
 Given $Y_k$, the value of $SS(f_k)$ can be computed in $O(1)$ operations using~\eqref{eq:SSfk}.
@@
 Line~\ref{alg_sortindices} contains the function $\operatorname{sortindices}$ that, given $z = \{z_i, i \in D\}$, returns the permutation $\sigma$ as described in~\eqref{eq:sigmasortind}.
-The $\operatorname{sortindicies}$ function requires sorting $\abs{D}$ elements.
+The $\operatorname{sortindices}$ function requires sorting $\abs{D}$ elements.
 This requires $O(L \log L)$ operations.
-The $\operatorname{sortindicies}$ function is the primary bottleneck in this algorithm when $L$ is large.
+The $\operatorname{sortindices}$ function is the primary bottleneck in this algorithm when $L$ is large.
 The loops on lines~\ref{alg_loop_setup} and~\ref{alg_loop_search} and the operations on lines~\ref{alg_Y} to lines~\ref{alg_Q} all require $O(L)$ or less operations.
@@
 \begin{enumerate}
-\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$,
-\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$.
+\item $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow 0$ almost surely as $L \rightarrow \infty$, \label{thm:consistency:part1}
+\item $\hat{\rho} \rightarrow \rho_0 G(0)$ almost surely as $L \rightarrow \infty$. \label{thm:consistency:part2}
 \end{enumerate}
@@
 The proof of Theorem~\ref{thm:consistency} is in Section~\ref{sec:proof-almost-sure} and the proof of Theorem~\ref{thm:normality} is in Section~\ref{sec:proof-asympt-norm}.
-Before giving the proofs we discuss the assumptions made by these theorems.
+Before giving the proofs we discuss some of the consequences and the assumptions made by these theorems.
+First observe that the theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ rather than directly on $\theta_0 - \hat{\theta}$.
+This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
+So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{0.99\pi + 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert 0.99\pi + 0.99\pi\vert = 1.98\pi$.
+In the literature it is usual to write \emph{estimator} minus \emph{error}, i.e., $\hat{\theta} - \theta_0$ rather than $\theta_0 - \hat{\theta}$.
+However, it will be notationally convenient in Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} to use $\langle\theta_0 - \hat{\theta}\rangle_\pi$.
+The theorems require minor modification when the number of pilots symbols $\abs{P} = 0$.
+Theorem statements for this special case appear in \cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+
+Part~\ref{thm:consistency:part1} of Theorem~\ref{thm:consistency} shows that the phase estimator error $\langle\theta_0 - \hat{\theta}\rangle_\pi$ converges almost surely to zero as $L\rightarrow \infty$, i.e., the estimator $\hat{\theta}$ is strongly consistent.
+In contrast, part~\ref{thm:consistency:part2} of the theorem shows that $\hat{\rho}$ is not a consistent estimator of the true amplitude $\rho_0$ and converges to $\rho_0 G(0)$.
+The value $G(0)$ is a function of the noise distribution and $G(0)$ is close to $1$ when the variance of the noise is small, but grows without bound as the noise variance increases.
+That is, the asymptotic bias is small when the SNR is large, but significant when the SNR is small.
+We will further investigate the bias in Sec.~\ref{sec:simulations}.
+Theorem~\ref{thm:normality} shows that $\sqrt{L}(\hat{\rho} - \rho_0 G(0))$ (i.e.\ the amplitude estimator error with bias removed) and $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ are asymptotically normally distributed and statistically independent as $L\rightarrow\infty$.
+
+We would like to stress that many practical estimators such as the SNR estimators in~\cite{Pauluzzi2000} are biased and particularly at low SNR, the bias can be large.
+To the best of the authors' knowledge, the value of the bias of these estimators has only been determined by computer simulation and no theoretical analysis exists.
+In contrast, we have presented a closed-form expression for the asymptotic bias of $\hat\rho$, thereby laying the foundation for an improved estimator.
+Removal of the bias appears to require knowledge of $G(0)$, which may not be available at the receiver.
+Estimation of $G(0)$ is beyond the scope of this paper.
+
 The assumption that $w_1, \dots, w_L$ are circularly symmetric can be relaxed, but this comes at the expense of making the theorem statements more complicated.
@@
 We will also use the fact that $R_i$ has finite variance to simplify the proof of Theorem~\ref{thm:consistency} by use of Kolmogorov's strong law of large numbers~\cite{Billingsley1979_probability_and_measure}.
-
-The theorems place conditions on $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ rather than directly on $\theta_0 - \hat{\theta}$.
-This makes sense because the phases $\theta_0$ and $\theta_0 + 2\pi k$ are equivalent for any integer $k$.
-So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{0.99\pi + 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert 0.99\pi + 0.99\pi\vert = 1.98\pi$.
 
@@
 
+The proof of Theorem~\ref{thm:consistency} hinges on Lemmas \ref{lem:convtoexpGlamL} to~\ref{lem:GLtoG0}, which can be found in the appendix.
 Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ minimised with respect to the data symbols,
@@
 
-\begin{lemma}\label{lem:convtoexpGlamL} 
-$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-Since $G(x)$ is uniquely maximised at $x=0$,
-\[
-0 \leq G(0) - G(\hat{\lambda}_L),
-\]
-and since $\hat{\lambda}_L$ is the maximiser of $G_L(x)$,
-\[ 
-0 \leq G_L(\hat{\lambda}_L) - G_L(0).
-\]
-Thus,
-\begin{align*}
-0 &\leq G(0) - G(\hat{\lambda}_L) \\ 
-& \leq G(0) - G(\hat{\lambda}_L) + G_L(\hat{\lambda}_L) - G_L(0) \\
-& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
-&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
-\end{align*}
-and the last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:uniflawGGL} 
-$\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-This type of result is called a~\emph{uniform law of large numbers} and follows from standard arguments.
-See, for example, van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
-A proof of this specific result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
-\end{IEEEproof}
-
-\begin{lemma}\label{lem:GLtoG0}
-$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
-\end{lemma}
-\begin{IEEEproof}
-By the triangle inequality,
-\[
-\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
-\]
-Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ by Lemma~\ref{lem:convtoexpGlamL}.
-\end{IEEEproof}
-
 \section{Proof of asymptotic normality (Theorem~\ref{thm:normality}) } \label{sec:proof-asympt-norm}
 
+The proof of Theorem~\ref{thm:normality} hinges on Lemmas \ref{lem:diffatlambdaL} to~\ref{lem:XL}, which can be found in the appendix.
 We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
@@
 
-\section{The Gaussian noise case}\label{sec:gaussian-noise-case}
-
-Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.  The joint density function of the real and imaginary parts is
-\[
-\frac{1}{2\pi\sigma^2}e^{-\frac{1}{2\sigma^2}(x^2 + y^2)}.
-\]
-Theorems~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
-%and the joint density function of the real and imaginary parts of $Z$ is
-%\[
-%p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
-%\] 
-%where $\kappa = \tfrac{\rho_0}{\sigma}$.  
-It can be shown that
-\[
-%\begin{equation}\label{eq:gfornormal}
-%g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
-%g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
-g(\phi) = \frac{\cos\phi}{2\pi}e^{-\kappa} + \frac{\Psi(\sqrt{2\kappa} \cos\phi)}{\sqrt{\pi\kappa}}  e^{-\kappa\sin^2\phi} \big(2 + \kappa\cos^2\phi \big)
-%\end{equation}
-\]
-where $\kappa = \tfrac{\rho_0^2}{2\sigma^2}$ and $\Psi(t) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{t} e^{-x^2/2} dx$
-is the cumulative density function of the standard normal.
-%\[
-%\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos\phi}{\sqrt{2}}\right)
-%\]
-%and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  
-The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using this formula.
+% \section{The Gaussian noise case}\label{sec:gaussian-noise-case}
+
+% Let the noise sequence $\{w_i\}$ be complex Gaussian with independent real and imaginary parts having zero mean and variance $\sigma^2$.
+The joint density function of the real and imaginary parts is
+% \[
+% \frac{1}{2\pi\sigma^2}e^{-\frac{1}{2\sigma^2}(x^2 + y^2)}.
+% \]
+% Theorems~\ref{thm:consistency}~and~\ref{thm:normality} hold, and since the distribution of $w_1$ is circularly symmetric, the distribution of $R_1e^{j\Phi_1}$ is identical to the distribution of $1 + \frac{1}{\rho_0} w_1$.
+% %and the joint density function of the real and imaginary parts of $Z$ is
+% %\[
+% %p_Z(x,y) = \rho_0^2 p( \rho_0 (x - 1), \rho_0 y ) = \frac{\kappa^2}{2\pi}e^{-\kappa^2(x ^2 - 2x + 1 + y^2) }
+% %\] 
+% %where $\kappa = \tfrac{\rho_0}{\sigma}$.  
+% It can be shown that
+% \[
+% %\begin{equation}\label{eq:gfornormal}
+% %g(\phi) = \frac{\cos^2(\phi)}{4\sqrt{\pi}}\left( \frac{e^{-\kappa^2} }{\sqrt{\pi}} + \kappa \Psi(\phi)  e^{-\kappa^2\sin^2(\phi)}\cos(\phi) \right)
+% %g(\phi) = \frac{\cos\phi}{2\pi}e^{-\frac{1}{2}\kappa^2} + \frac{\Psi(\kappa \cos\phi)}{\kappa\sqrt{2\pi}}  e^{-\frac{1}{2}\kappa^2\sin^2\phi} \big(1 + \kappa^2\cos^2\phi \big)
+% g(\phi) = \frac{\cos\phi}{2\pi}e^{-\kappa} + \frac{\Psi(\sqrt{2\kappa} \cos\phi)}{\sqrt{\pi\kappa}}  e^{-\kappa\sin^2\phi} \big(2 + \kappa\cos^2\phi \big)
+% %\end{equation}
+% \]
+% where $\kappa = \tfrac{\rho_0^2}{2\sigma^2}$ and $\Psi(t) = \frac{1}{\sqrt{2\pi}} \int_{-\infty}^{t} e^{-x^2/2} dx$
+% is the cumulative density function of the standard normal.
+% %\[
+% %\Psi(\phi) = \frac{1}{2} + \frac{1}{2} \erf\left(\frac{\kappa \cos\phi}{\sqrt{2}}\right)
+% %\]
+% %and $\erf(x) = \frac{2}{\sqrt{\pi}}\int_0^{x} e^{-t^2} dt$ is the error function.  
+% The value of $A_1, A_2, B_1$ and $B_2$ can be efficiently computed by numerical integration using this formula.
 
@@
 As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-Theorem~\ref{thm:normality} does not model this behaviour in the sense that for any fixed $L$ there exist sufficiently small values of SNR for which Theorem~\ref{thm:normality} does not produce accurate predictions of the MSE.
 As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e.
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non-data aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non data-aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
 This estimator requires selection of a function $F$ that transforms the amplitude of each sample prior to the final estimation step.
@@
 Viterbi and Viterbi propose several viable alternatives, from which we have chosen $F(x) = 1$.
-The Viterbi and Viterbi estimator is only applicable in the non-data aided setting, i.e.
+The Viterbi and Viterbi estimator is only applicable in the non data-aided setting, i.e.
 when $\abs{P} = 0$.
@@
 % \end{IEEEproof}
+
+\begin{lemma}\label{lem:convtoexpGlamL} 
+$G(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+Since $G(x)$ is uniquely maximised at $x=0$,
+\[
+0 \leq G(0) - G(\hat{\lambda}_L),
+\]
+and since $\hat{\lambda}_L$ is the maximiser of $G_L(x)$,
+\[ 
+0 \leq G_L(\hat{\lambda}_L) - G_L(0).
+\]
+Thus,
+\begin{align*}
+0 &\leq G(0) - G(\hat{\lambda}_L) \\ 
+& \leq G(0) - G(\hat{\lambda}_L) + G_L(\hat{\lambda}_L) - G_L(0) \\
+& \leq |G(0) - G_L(0)| + |G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)| \\
+&\leq 2\sup_{\lambda \in [-\pi, \pi)} \sabs{G_L(\lambda) - G(\lambda)},
+\end{align*}
+and the last line converges almost surely to zero by Lemma~\ref{lem:uniflawGGL}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:uniflawGGL} 
+$\sup_{\lambda \in [-\pi, \pi)}\sabs{G_L(\lambda) - G(\lambda)} \rightarrow 0$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+This type of result is called a~\emph{uniform law of large numbers} and follows from standard arguments.
+See, for example, van~de~Geer~\cite[Ch.~3]{van2009empirical} or Pollard~\citep[Ch.~2]{Pollard_conv_stat_proc_1984}.
+A proof of this specific result is given in~\cite[Lemma~2]{McKilliam_leastsqPSKpilotsdata_arxiv}.
+\end{IEEEproof}
+
+\begin{lemma}\label{lem:GLtoG0}
+$G_L(\hat{\lambda}_L) \rightarrow G(0)$ almost surely as $L \rightarrow \infty$.
+\end{lemma}
+\begin{IEEEproof}
+By the triangle inequality,
+\[
+\sabs{G_L(\hat{\lambda}_L) - G(0)} \leq \sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} + \sabs{G(\hat{\lambda}_L) - G(0)}.
+\]
+Now $\sabs{G_L(\hat{\lambda}_L) - G(\hat{\lambda}_L)} \rightarrow 0$ as $L \rightarrow \infty$ as a result of Lemma~\ref{lem:uniflawGGL}, and $\sabs{G(\hat{\lambda}_L) - G(0)} \rightarrow 0$ almost surely as $L \rightarrow \infty$ by Lemma~\ref{lem:convtoexpGlamL}.
+\end{IEEEproof}
 

### 2013-11-18 response.tex
modified the mse amplitude estimator plot

@@
 We have also added a discussion of the implications and the practical relevance of the theorems after the statement of Theorems~\zref{thm:consistency} and~\zref{thm:normality} in Section.~\zref{sec:stat-prop-least}.
+We have also expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
 
@@
 	\\\textbf{Response:}
-%	\comment{Replace variance plot by MSE plot.}
+In the original manuscript we plotted the \emph{variance} of the amplitude estimator, not the mean square error.
+That is, the solid lines where the theoretical variance of $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$ and each point from the Monte-Carlo simulations was computed as the unbiased error $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+We chose this approach in order to highlight the correctness of our asymptotic theory.
+This was decribed in the 4th paragraph of the old Section~8 containing the simulations section.
+
+We do agree with the reviewer this this plot by itself initially appears misleading.
+To aleviate in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the bias plus variance squared.
+The following paragraph has also be placed in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
+
+``Figure~\zref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
+shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+The solid lines show the mean square error predicted by the bias $\rho_0\big(1 - G(0)\big)$ squared plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\zref{thm:normality}.
+That is, the solid lines are given by
+\[
+\rho_0^2\big(1 - G(0)\big)^2 + \rho_0^2\frac{p B_1 + d B_2}{L}. 
+\]
+When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
+However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+Figure~\zref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
+''
+
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
-We thank their reviewer for the comments. We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our paper clear.
+We thank their reviewer for the comments. We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
 \end{enumerate}

### 2013-11-18 paper.tex
modified the mse amplitude estimator plot

@@
 In this paper we present an alternative approach to the study of estimators for carrier phase and amplitude.
-Rather than compute bounds, we directly study the least squares estimators of carrier phase and amplitude asymptotically as the number of transmitted symbols $L\to \infty$.
+Rather than compute bounds, we directly study the least squares estimator asymptotically as the number of transmitted symbols $L\to \infty$.
 We show, under comparatively mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent (asymptotically unbiased) and asymptotically normally distributed.
@@
 For example, in practical systems it is sometimes desirable to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.
-This is related to what is called \emph{unequal error protection} in the literature~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.
+This is related to what is called \emph{unequal error protection}~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.
 It would be reasonably straigthforward to extend our results to allow data and pilot symbols with varying constellation size.
@@
 
-That this paper contains novel theoretical developments should does not imply that our results are of no immediate practical interest.
+That this paper contains novel theoretical developments does not imply that our results are of no immediate practical interest.
 The design of our estimator was motivated by a project at the University of South Australia's Institute for Telecommunications Research to develop a narrowband multiuser communication system to collect sensor network data using a satellite in orbit.
@@
 Our theoretical analysis aided design of this system.
-Based on our Theorem~\ref{thm:normality} in Section~\ref{sec:stat-prop-least} the performance of the least squares carrier phase and amplitude estimators can be accurately predicted without the need for time consuming Monte-Carlo simulations.
-This provides easy to use and rigourously justified methods for choosing systems parameters such as the number and placement of pilot symbols and also provides useful information to other reciever components, such as channel tracking systems.
+Based on our Theorem~\ref{thm:normality} in Section~\ref{sec:stat-prop-least} the performance of the least squares estimator can be accurately predicted without the need for time consuming Monte-Carlo simulations.
+This provides easy to use and rigourously justified methods for choosing systems parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.
 
@@
 
-Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
-shows the variance of the amplitude estimator when $M=2$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+% Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
+% shows the variance of the amplitude estimator when $M=2$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
 The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
@@
 In practice $G(0)$ may not be known at the receiver, so Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK}
-serves to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
+% serves to validate the correctness of our asymptotic theory, rather than to suggest the practical performance of the amplitude estimator.
 When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
@@
 However, due to the bias, the MSE of the least squares amplitude estimator is \emph{not} smaller than that of the unmodulated carrier.
+
+Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
+shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+The solid lines show the mean square error predicted by the bias $\rho_0\big(1 - G(0)\big)$ squared plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\ref{thm:normality}.
+That is, the solid lines are given by
+\[
+\rho_0^2\big(1 - G(0)\big)^2 + \rho_0^2\frac{p B_1 + d B_2}{L}. 
+\]
+When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
+However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+Figure~\ref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
 
@@
 % SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
-% \]
+% \]  
 % where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
@@
 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-		\caption{Amplitude estimator variance (bias removed) versus SNR for BPSK.}
+		\caption{Amplitude estimator versus SNR for BPSK with $L=2048$}
 		\label{fig:plotampBPSK}

### 2013-11-18 response.tex
added data

@@
 	\\\textbf{Response:}
-We have shorteed Section~\zref{sec:least-squar-estim} but have left the description of the algorithm intact.
+\comment{look at shortening sec 2}
+
+We have shortened Section~\zref{sec:least-squar-estim} but have left the description of the algorithm intact.
 The primary reason for this is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.
@@
 	\textbf{Response:}
-%	\comment{add response\dots}
+The non data-aided estimation of SNR is a significant difficulty when implementing this approach.
+As observed by simulation in [c] known non data-aided estimators of SNR appear to be biased.
+The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
+This makes it difficult to estimate SNR so as to know when to use a DA algorithm with only the available pilots and when to use the VV algorithm.
+Analysis of the bias of the chosen non data-aided SNR estimator is required in order to justify this approach.
+To date the bias and variance of non data-aided estimators of SNR (or equivalently carrier amplitude) appear to have been measured only by Monte-Carlo simulation [c].
+In the abstract of our new paper we point out that our of paper
+
+``
+appears to be the first time that the statistical properties of a non data-aided estimator for carrier amplitude have been analysed theoretically.
+''
+
+and in the new introduction we affirm that
+
+``
+while the specific results given here apply only to the least square estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
+''
+
+Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non data-aided SNR estimator in order to implement the reviewers proposal.
+We confirm however, that analysing and implementing the reviewers suggested technique is well beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
+
 \item\textbf{Comment:} 
@@
 We chose this approach in order to highlight the correctness of our asymptotic theory.
-This was decribed in the 4th paragraph of the old Section~8 containing the simulations section.
+This was described in the 4th paragraph of the old Section~8 containing the simulations section.
 
 We do agree with the reviewer this this plot by itself initially appears misleading.
-To aleviate in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the bias plus variance squared.
+To alleviate in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the bias plus variance squared.
 The following paragraph has also be placed in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
@@
 Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.''
+For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.''
 
@@
 	This work addresses the problem of carrier phase and amplitude estimation for $M$-ary PSK signals.
-Based on a modification of the Mackhentum's algorithm by inserting pilots, the main contribution of this manuscript is in fact the study of the asymptotic behavior of the least squares estimator for carrier phase and amplitude.
+Based on a modification of the Mackhentun's algorithm by inserting pilots, the main contribution of this manuscript is in fact the study of the asymptotic behavior of the least squares estimator for carrier phase and amplitude.
 Although the derivations seems correct, it is not clear to the reader why this method should be applicable in a communications context.
@@
 	\\\textbf{Response:}
-%	\comment{Show PER ASRP results. Mention trials and patents\dots}
+We thank their reviewer for the comments. We have expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
+% We have added the following paragraph to the introduction
+
+% ``In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
+These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
+% These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
+A key problem is that the EM algorithm converges correctly only if initialised at some channel estimate sufficiently close to the true channel parameters.
+Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
+For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.''
+
+% In view of this, a direct comparison with a turbo synchronisation proceedure is not meaningful.
+A meaninful simulation that could be implemented is to compare the results of turbo synchronisation with different initialisers.
+For example one could initialise the proceedure using only the pilots and a data-aided algorithm and also with our least squares algorithm and compare the bit error rates.
+It can be reasonably expected that the initialiser that is a more accurate estimator of carrier phase and amplitude with also be a better initialiser for a turbo synchronisation proceedure and will thus have a smaller bit error rate.
+So, for example, we expect our least square estimator to be a beter initialiser at moderate and high SNR and a data-aided approach to be better at low SNR provided there are sufficiently many pilots symbols.
+
+% For this reason that we do not expect simulations of this sort to yield substantially more insight than the mean square error plots that we have presented in the paper.
+The primary purpose of our paper is the statistical analysis of the least squares estimator and we feel that
+
 %	\comment{Show PER ASRP ST1 turbo sync results. Comparison with other estimators?}

### 2013-11-18 paper.tex
added data

@@
 
-We would like to stress that many practical estimators such as the SNR estimators in~\cite{Pauluzzi2000} are biased and particularly at low SNR, the bias can be large.
-To the best of the authors' knowledge, the value of the bias of these estimators has only been determined by computer simulation and no theoretical analysis exists.
-In contrast, we have presented a closed-form expression for the asymptotic bias of $\hat\rho$, thereby laying the foundation for an improved estimator.
+We would like to stress that many practical estimators such as the SNR estimators in~\cite{Pauluzzi2000} are biased and, particularly at low SNR, the bias can be large.
+To the best of the authors' knowledge, the value of the bias of these estimators has only been approximated by computer simulation and no theoretical analysis exists.
+In contrast, we have presented a closed-form expression for the asymptotic bias of $\hat\rho$, thereby laying a foundation for an improved estimator.
 Removal of the bias appears to require knowledge of $G(0)$, which may not be available at the receiver.
@@
 A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
-%When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
-% This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
+%Thus, when the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
+ This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
 For example, rather than minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
-% \[
-% SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
-% \]  
-% where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
+ \[
+ SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
+ \] 
+ where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
 Computing the $\hat{a}$ that minimises $SS_\beta$ can be achieved with only a minor modification to algorithm~\ref{alg:loglinear}.

### 2013-11-19 response.tex
more additions to response/paper

--- 2013-11-18 response.tex
@@
 
-\section*{Reviewer 1}
+\section*{Reviewer 1}\label{sec:reviewer-1}
 
 \begin{enumerate}
-\item\textbf{Comment:} 
+\item\textbf{Comment:} \label{comment:shortersec2}
 	This paper is concerned with carrier phase and amplitude estimation for PSK modulated signalling.
@@
 	\\\textbf{Response:}
-We have shorteed Section~\zref{sec:least-squar-estim} but have left the description of the algorithm intact.
+%\comment{look at shortening sec 2}
+
+We have left the description of the algorithm intact.
 The primary reason for this is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.
@@
 
-``
-In this section we derive Mackentun's algorithm to compute the least squares estimators of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the non-data aided setting, so we modify the algorithm to include pilot symbols.
-Although the modification is straightforward, much of the notation that we will use for analysing the estimator in Section~\ref{sec:stat-prop-least} is naturally introduced during the description of the algorithm.
-For this reason, and so that our paper is self contained, we describe the algorithm in full.
-''
+``In this section we derive Mackenthun's algorithm to compute the least squares estimators of the carrier phase and amplitude~\cite{Mackenthun1994}.
+Mackenthun specifically considered the non data-aided setting, so we modify the algorithm to include pilot symbols.
+Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\ref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
+For this reason, and so that our paper is self contained, we describe the algorithm in full.''
 
@@
 	In the introductory section (paragraph before Eq.\ 3), the authors are talking about ``noncoherent detection where no pilot symbols exist''.
-In my mind this is somewhat confusing, because parameter estimation without knowing the underlying data is usually denoted as non-data aided (NDA) or decision-directed (DD), whereas when these data are known we have a data-aided (DA) situation [a].
+In my mind this is somewhat confusing, because parameter estimation without knowing the underlying data is usually denoted as non-data aided (NDA) or decision-directed (DD), whereas when these data are known we have a data-aided (DA) situation\cite{Mengali_andre_synchro_book}.
 On the other hand, coherent/noncoherent scenarios are given if the carrier phase is known/unknown to the receiver station.
@@
 the LS results are compared to an alternative solution represented by the Viterbi and Viterbi (VV) algorithm.
-Why not doing the same with $|P| > 0$ by the following approach, which is definitely much simpler from the computational point of view (complexity in the order $O(L)$ instead of $O(L \log(L))$ required by LS) and the jitter performance will probably differ not very much from that of the LS variant (in Figs.\ 1Ð3 this can be seen for $|P| = 0$):
+Why not doing the same with $|P| > 0$ by the following approach, which is definitely much simpler from the computational point of view (complexity in the order $O(L)$ instead of $O(L \log(L))$ required by LS) and the jitter performance will probably differ not very much from that of the LS variant (in Figs.~1 to 3 this can be seen for $|P| = 0$):
 	\begin{itemize}
@@
 	\item
-		The SNRth is a threshold value known in advance, e.g., by the intersection of the Cramer-Rao lower bounds (CRLBs) for DA and NDA phase estimation [a, b].
+		The SNRth is a threshold value known in advance, e.g., by the intersection of the Cramer-Rao lower bounds (CRLBs) for DA and NDA phase estimation\cite{Mengali_andre_synchro_book,950345}.
 	\item
-		The SNR estimates might be obtained by a simple method based on second- and fourth-order moments [c].
+		The SNR estimates might be obtained by a simple method based on second- and fourth-order moments~\cite{Pauluzzi2000}.
 	\item
@@
 	\textbf{Response:}
-%	\comment{add response\dots}
+The non data-aided estimation of SNR is a significant difficulty when implementing this approach.
+As observed by simulation in~\cite{Pauluzzi2000} known non data-aided estimators of SNR appear to be biased.
+The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
+This makes it difficult to estimate SNR so as to know when to use a DA algorithm with only the available pilots and when to use the VV algorithm.
+Analysis of the bias of the chosen non data-aided SNR estimator is required in order to justify this approach.
+To date the bias and variance of non data-aided estimators of SNR (or equivalently carrier amplitude) appear to have been measured only by Monte-Carlo simulation~\cite{Pauluzzi2000}.
+In the abstract of our new paper we point out that our of paper
+
+``
+appears to be the first time that the statistical properties of a non data-aided estimator for carrier amplitude have been analysed theoretically.
+''
+
+and in the new introduction we affirm that
+
+``
+while the specific results given here apply only to the least square estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
+''
+
+Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non data-aided SNR estimator in order to implement the reviewers proposal.
+We confirm however, that analysing and implementing the reviewers suggested technique is well beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
+
 \item\textbf{Comment:} 
@@
 	In a much more powerful approach to the topic addressed in the current paper, the pilot sequence would be employed for a coarse estimation/correction of the carrier phase of the payload data.
-In the sequel, by embedding carrier phase recovery into an iterative (turbo) decoding scheme [d, e] or any other error correcting method [f, g], the jitter performance is close to the CRLB for DA estimation even at very small SNRs, which are mainly determined by the convergence region of the (turbo) decoder.
+In the sequel, by embedding carrier phase recovery into an iterative (turbo) decoding scheme\cite{Lottici2004,Noels2005} or any other error correcting method~\cite{580211,974597}, the jitter performance is close to the CRLB for DA estimation even at very small SNRs, which are mainly determined by the convergence region of the (turbo) decoder.
 By taking into account the fact that modern and powerful communication systems are not conceivable without any FEC scheme, it is not clear to me why I should choose your algorithm in such a context.
@@
 Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.''
+For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.''
+
+\comment{BLERG, add LDPC turbo sync sims}
 
@@
 \end{enumerate}
-\end{enumerate}
-
-\subsection*{Reviewer 2 references}
-\begin{enumerate}\footnotesize
-\renewcommand{\theenumi}{\alph{enumi}}
-\renewcommand{\labelenumi}{[\theenumi]}
-\addtolength{\itemsep}{-1.5ex}
-\item
-	Mengali, D'Andrea, \textit{Synchronization Techniques for Digital Receivers}, Plenum Press: New York, 1997.
-\item
-	Rice, Cowley, Moran, Rice, ``Cramer-Rao Lower Bounds for QAM Phase and Frequency Estimation'', \textit{IEEE Trans.\ Commun.}, Sept 2001.
-\item
-	Pauluzzi, Beaulieu, ``A Comparison of SNR Estimation Techniques for the AWGN Channel'', \textit{IEEE Trans.\ Commun.}, Oct.\ 2000.
-\item
-	Lottici, Luise, ``Embedding Carrier Phase Recovery into Iterative Decoding of Turbo-Coded Linear Modulations'', \textit{IEEE Trans.\ Commun.}, April 2004.
-\item
-	Noels et al., ``A Theoretical Framework for Soft-Information-Based Synchronization in Iterative (Turbo) Receivers'', \textit{Eurasip Journal Wireless Commun.\ and Networking}, April 2005.
-\item
-	D'Andrea, Mengali, Vitetta, ``Approximate ML Decoding of Coded PSK with no Explicit Carrier Phase Reference'', \textit{IEEE Trans.\ Commun.}, Feb./March/April 1994.
-\item
-	Vanelli, Salmi, Cioni, Corazza, Polydoros, ``A Performance Review of PSP for Joint Phase/Frequency and Data Estimation in Future Broadband Satellite Networks'', \textit{IEEE Journal on Selected Areas in Commun.}, Dec.\ 2001.
 \end{enumerate}
@@
 	\\\textbf{Response:}
+We thank their reviewer for their comments. We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
+
 %	\comment{Stress practical relevance by showing ASRP PER results. In particular ST1, where the estimators are used to seed turbo sync. Mention trials and patents\dots}
@@
 	\\\textbf{Response:}
-%	\comment{Show PER ASRP results. Mention trials and patents\dots}
+We have expanded the introduction of the paper to better motivate the work.  In particular, the following paragraph has been added:
+
+``That this paper contains novel theoretical developments does not imply that our results are of no immediate practical interest.
+The design of our estimator was motivated by a project at the University of South Australia's Institute for Telecommunications Research to develop a narrowband multiuser communication system to collect sensor network data using a satellite in orbit.
+This estimator has been implemented on a prototype that was successfully trialled from April to July of 2013 with signals from multiple low power ground-based transmitters being successfully received and decoded by both aircraft and satellites in orbit~\cite{ASRPpromovideo}.
+Our theoretical analysis aided design of this system.
+Based on our Theorem~\zref{thm:normality} in Section~\zref{sec:stat-prop-least} the performance of the least squares estimator can be accurately predicted without the need for time consuming Monte-Carlo simulations.
+This provides easy to use and rigourously justified methods for choosing systems parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.''
+
 \item\textbf{Comment:} 
@@
 	\\\textbf{Response:}
+We have added the following paragraph to the introduction
+
+``In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
+These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
+ These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
+A key problem is that the EM algorithm converges correctly only if initialised at some channel estimate sufficiently close to the true channel parameters.
+Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
+For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.''
+
+\comment{Run turbo sync simulations and write something on this}
+
+%It can be reasonably expected that the initialiser that is a more accurate estimator of carrier phase and amplitude with also be a better initialiser for a turbo synchronisation proceedure and will thus have a smaller bit error rate.
+So, for example, we expect our least square estimator to be a beter initialiser at moderate and high SNR and a data-aided approach to be better at low SNR provided there are sufficiently many pilots symbols.
+
+% For this reason that we do not expect simulations of this sort to yield substantially more insight than the mean square error plots that we have presented in the paper.
+The primary purpose of our paper is the statistical analysis of the least squares estimator and we feel that
+
 %	\comment{Show PER ASRP ST1 turbo sync results. Comparison with other estimators?}
@@
 	\\\textbf{Response:} 
-%	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the estimator MSE.
+	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the estimator MSE.
 This was done to validate the theoretical value of the asymptotic variance of $\hat\rho$ predicted by Theorem~2 against Monte-Carlo simulations.
@@
 	\\\textbf{Response:} 
-%	Biased estimators are quite common in the communications literature and in fact, are often used in practice.
+	Biased estimators are quite common in the communications literature and in fact, are often used in practice.
 Prime examples are state-of-the-art signal-to-noise ratio (SNR) estimators for $M$-PSK modulated signals such as those discussed by Pauluzzi and Beaulieu \cite{Pauluzzi2000}.
@@
 	\\\textbf{Response:}
-%	We have shortened Sec.\ II to make it more concise and improve readability. 
-\comment{BLERG}
 
-The original Sec.\ VII has been removed from the revised manuscript.
+The original Sec.~VII describing the Gaussian noise case has been removed from the revised manuscript.
 
-\comment{BLERG}
-%	Furthermore, we agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that Sec.\ III specifically discusses random variables of the form $Re^{j\varPhi}=1{+}W$, where $W$ is circularly symmetric.
-The joint pdf of $R$ and $\varPhi$ is a crucial ingredient for the asymptotic analysis presented in the manuscript and therefore, we feel the section should be retained.
-However, for improved readability, Sec.\ III has been slightly shortened and has been moved to the appendix.
+We agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that section specifically discussed random variables of the form $Re^{j\Phi}=1{+}W$, where $W$ is circularly symmetric.
+The joint pdf of $R$ and $\Phi$ is a crucial ingredient for the asymptotic analysis presented in the manuscript and therefore, we feel that this material must be retained.
+However, we have modified the way that this material is introduced.
+It is now introduced as needed when the asymptotic properties of the least squares estimator are derived.
+
+Regarding Section~II we refer the reader to our response to comment~\eqref{comment:shortersec2} of reviewer~\ref{sec:reviewer-1}.
+
+% Section~II describes our modification of Mackethun's algorithm to compute the least squares estimator.
+Although the modification is simple we have left this section intact since much of the notation required for the statistical analysis of the least squares estimator is most naturally introduced during the description of the algorithm.
+To qualify this for the reader we have included the following sentence at the start of Section~\zref{sec:least-squar-estim}.
+
+% ``In this section we derive Mackenthun's algorithm to compute the least squares estimators of the carrier phase and amplitude~\cite{Mackenthun1994}.
+Mackenthun specifically considered the non data-aided setting, so we modify the algorithm to include pilot symbols.
+Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\zref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
+For this reason, and so that our paper is self contained, we describe the algorithm in full.''
+
+%For example, equations~\eqref{eq:SSdatasymbols},~\eqref{eq:SSallparams} and~\eqref{eq:hatdfinxtheta} are all used during the
+
 \end{enumerate}

### 2013-11-19 paper.tex
more additions to response/paper

--- 2013-11-18 paper.tex
@@
 
-More ambitiously, our techniques might allow analysis of non data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,Heath99exploitinginput,inthft:204506}.
+More ambitiously, our techniques might allow analysis of non data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,inthft:204506}.
 We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.
@@
 
-In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
+In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_framework_turbo_sync_2007}.
 These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
@@
 Section~\ref{sec:least-squar-estim} extends Mackenthun's algorithm for the case when both pilot symbols and data symbols exist.
-Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
+%Section~\ref{sec:circ-symm-compl} describes properties of complex random variables that we need.
 Section~\ref{sec:stat-prop-least} states two theorems that describe the statistical properties of the least squares estimator of carrier phase and amplitude.
@@
 Mackenthun specifically considered the non data-aided setting, so we modify the algorithm to include pilot symbols.
-Although the modification is straightforward, much of the notation that we will use for analysing the estimator in Section~\ref{sec:stat-prop-least} is naturally introduced during the description of the algorithm.
+Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\ref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
 For this reason, and so that our paper is self contained, we describe the algorithm in full.
@@
 where, 
-\begin{align*}
-Y_k = Y( f_k ) &= \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* \\
-&= B + \sum_{i \in D}g_{ki},
-\end{align*}
+\[
+Y_k = Y( f_k ) = \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* = B + \sum_{i \in D}g_{ki},
+\]
 where $B = \sum_{i \in P} y_i p_i^*$ is independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
@@
 The least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata},
-\begin{equation}\label{eq:ahatYhat}
+\[
 \hat{a} = \frac{1}{L} Y_{\hat{k}}.
-\end{equation}
+\]
 Pseudocode is given in Algorithm~\ref{alg:loglinear}.
@@
 
-\section{Circularly symmetric complex random variables}\label{sec:circ-symm-compl}
-
-Before describing the statistical properties of the least squares estimator, we first require some properties of complex valued random variables.
+\section{Statistical properties of the least squares estimator}\label{sec:stat-prop-least}
+
+In this section we describe the asymptotic properties of the least squares estimator.
+We first require some properties of complex valued random variables.
 A complex random variable $W$ is said to be \emph{circularly symmetric} if its phase $\angle{W}$ is independent of its magnitude $\abs{W}$ and if the distribution of $\angle{W}$ is uniform on $[0,2\pi)$.
@@
 
-\section{Statistical properties of the least squares estimator}\label{sec:stat-prop-least}
-
-In this section we describe the asymptotic properties of the least squares estimator.
 In what follows we use $\sfracpart{x}_\pi$ to denote $x$ taken `modulo $2\pi$' into the interval $[-\pi, \pi)$, that is
@@
 However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
-Similar properties have been used by some of the present authors to analyse the behaviour of polynomial-phase estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.
+%Similar properties have been used by some of the present authors to analyse the behaviour of polynomial-phase estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.
 %\cite{McKilliam_LSU_polyest_part1_2012,McKilliam_LSU_polyest_part2_2012}.
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non data-aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983} described by~\eqref{eq:viterbiviterbi}.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non data-aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
 This estimator requires selection of a function $F$ that transforms the amplitude of each sample prior to the final estimation step.
@@
 A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
-%When the SNR is small it is better to ignore the data symbols and use only the pilot symbols.
-% This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
+This behaviour suggests that using the data symbols is counterproductive when the SNR is low.
+
+%Thus, when the SNR is small it is better to ignore the data symbols and use only the pilot symbols.   
+ %This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
 For example, rather than minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
@@
  % SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
-% \]  
+ % \] 
  % where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.

### 2014-01-29 response.tex
added to response

@@
 Mackenthun specifically considered the non data-aided setting, so we modify the algorithm to include pilot symbols.
-Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\ref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
+Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\zref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
 For this reason, and so that our paper is self contained, we describe the algorithm in full.''
@@
 	For improved readability of the manuscript, we have now moved all of the lemmas into the appendix.
-We have also added a discussion of the implications and the practical relevance of the theorems after the statement of Theorems~\zref{thm:consistency} and~\zref{thm:normality} in Section.~\zref{sec:stat-prop-least}.
+We have also added a discussion of the implications and the practical relevance of the theorems after the statement of Theorems~\zref{thm:consistency} and~\zref{thm:normality} in Section~\zref{sec:stat-prop-least}.
 We have also expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
@@
 	\\\textbf{Response:}
-	We agree with the reviewer and apologise for causing confusion with the use of the terms \emph{coherent} and \emph{noncoherent} in the original manuscript.
-In the updated manuscript, we refer to an estimator that does not use pilot symbols as \emph{non data-aided}.
-%In the updated manuscript, we refer to an estimator that uses known pilot symbols as \emph{data-aided} and to one that relies on receiver decisions on the data symbols as \emph{decision-directed}.
+	We agree with the reviewer.  In the updated manuscript, we refer to an estimator that does not use pilot symbols as \emph{non-data-aided}.
 \item\textbf{Comment:} 
 	missing `,' after (5)
-	\\\textbf{Response:} 
-We thank the reviewer for spotting the missing comma and have added it.
+	\\\textbf{Response:} Fixed.
 \item\textbf{Comment:} 
 	p.\ 19, at the bottom: sortindicIes (typo, delete 'i')
-	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
+	\\\textbf{Response:} Fixed.
 \item\textbf{Comment:} 
@@
 Theorem 2 shows that for $L\rightarrow\infty$, $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ is normally distributed with zero mean and finite variance.
-The support of $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ is the interval $[-\sqrt{L}\pi, \sqrt{L}\pi)$, i.e.\ $(-\infty, \infty)$ as $L\rightarrow\infty$.
+The support of $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ is the interval $[-\sqrt{L}\pi, \sqrt{L}\pi)$, i.e., $(-\infty, \infty)$ as $L\rightarrow\infty$.
 For finite $L$, however, $\sqrt{L}\sfracpart{\theta_0-\hat{\theta}}_\pi$ has finite support and thus, cannot be exactly normally distributed.
@@
 	Paragraph before (13): \dots can be computed from its predecessor \dots
-	\\\textbf{Response:} Thank you for pointing out this typographical error. We have corrected it.
+	\\\textbf{Response:} This typographical error has been corrected.
 \item\textbf{Comment:} 
@@
 	\textbf{Response:}
-The non data-aided estimation of SNR is a significant difficulty when implementing this approach.
-As observed by simulation in~\cite{Pauluzzi2000} known non data-aided estimators of SNR appear to be biased.
+The non-data-aided estimation of SNR is a significant difficulty when implementing the approach suggested by the reviewer.
+As observed by simulation in~\cite{Pauluzzi2000} known non-data-aided estimators of SNR appear to be biased.
 The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
 This makes it difficult to estimate SNR so as to know when to use a DA algorithm with only the available pilots and when to use the VV algorithm.
-Analysis of the bias of the chosen non data-aided SNR estimator is required in order to justify this approach.
-To date the bias and variance of non data-aided estimators of SNR (or equivalently carrier amplitude) appear to have been measured only by Monte-Carlo simulation~\cite{Pauluzzi2000}.
-In the abstract of our new paper we point out that our of paper
+Analysis of the bias of the chosen non-data-aided SNR estimator is required in order to justify this approach.
+To date the bias and variance of non-data-aided estimators of SNR (or equivalently carrier amplitude) appear to have been measured only by Monte-Carlo simulation~\cite{Pauluzzi2000}.
+In the abstract of our new paper we point out that
 
-``
-appears to be the first time that the statistical properties of a non data-aided estimator for carrier amplitude have been analysed theoretically.
-''
+``This appears to be the first time that the statistical properties of a non-data-aided estimator for carrier amplitude have been analysed theoretically.''
 
@@
 
-``
-while the specific results given here apply only to the least square estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
-''
+``while the specific results given here apply only to the least square estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.''
 
-Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non data-aided SNR estimator in order to implement the reviewers proposal.
-We confirm however, that analysing and implementing the reviewers suggested technique is well beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
+Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non-data-aided SNR estimator in order to implement the reviewers proposal.
+We confirm, however, that analysing and implementing the reviewers suggested technique is well beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
 
-\item\textbf{Comment:} 
+\item\textbf{Comment:} \label{rev2:biasmse}
 	By detailed inspection of Fig.\ 4 showing the jitter variance for amplitude estimates, it is easily observed that $|P| < L$ performs better than $|P| = L$ in the medium-to-low SNR range.
@@
 We do agree with the reviewer this this plot by itself initially appears misleading.
-To alleviate in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the bias plus variance squared.
-The following paragraph has also be placed in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
+To alleviate this confusion in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the bias plus variance squared.
+The following paragraph has also been included in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
 
@@
 When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
-However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+However, $G(0)$ grows without bound as the variance of the noise increases so the bias is significant when SNR is small.
 Figure~\zref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
@@
 
-\item\textbf{Comment:} 
+\item\textbf{Comment:} \label{rev2:turbosync}
 	In a much more powerful approach to the topic addressed in the current paper, the pilot sequence would be employed for a coarse estimation/correction of the carrier phase of the payload data.
@@
 These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
-A key problem is that the EM algorithm converges correctly only if initialised at some channel estimate sufficiently close to the true channel parameters.
+A key problem is that the EM algorithm converges correctly only if initialised with a channel estimate sufficiently close to the true channel parameters.
 Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.''
+The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.''
 
@@
 	\\\textbf{Response:}
-We thank their reviewer for the comments. We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
+We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
+Also see our response to comment~\ref{rev4:shouldbettermotivate} of reviewer~4.
 \end{enumerate}
@@
 \begin{enumerate}
-\item\textbf{Comment:} 
+\item\textbf{Comment:} \label{rev4:shouldbettermotivate}
 	The authors should better motivate the interest in using their proposed method in actual communication systems.
@@
 	\\\textbf{Response:}
-We have added the following paragraph to the introduction
-
-``In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
-These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
- These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
-A key problem is that the EM algorithm converges correctly only if initialised at some channel estimate sufficiently close to the true channel parameters.
-Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.''
-
-\comment{Run turbo sync simulations and write something on this}
+We have included simulations with turbo synchronisation proceedures.  See our response to comment~\ref{rev2:turbosync} of reviewer~2.
 
@@
 	\\\textbf{Response:} 
-	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the estimator MSE.
-This was done to validate the theoretical value of the asymptotic variance of $\hat\rho$ predicted by Theorem~2 against Monte-Carlo simulations.
+Please see response~\ref{rev2:biasmse} to reviewer~2.
+%	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the %estimator MSE.
+This was done to validate the theoretical value of the asymptotic variance of $\hat\rho$ predicted by Theorem~2 against Monte-Carlo %simulations.
 For consistency with the other figures and for easier interpretation, we have replaced the variance plot by a plot of the MSE of $\hat\rho$.
@@
 	\\\textbf{Response:} 
-	Biased estimators are quite common in the communications literature and in fact, are often used in practice.
-Prime examples are state-of-the-art signal-to-noise ratio (SNR) estimators for $M$-PSK modulated signals such as those discussed by Pauluzzi and Beaulieu \cite{Pauluzzi2000}.
+	Biased estimators are quite common in the communications literature and are often used in practice.
+Prime examples are state-of-the-art signal-to-noise ratio (SNR) estimators for $M$-PSK modulated signals such as those discussed by Pauluzzi~and~Beaulieu~\cite{Pauluzzi2000}.
 With the exception of the data-aided SNR estimators, all of the estimators in \cite{Pauluzzi2000} are biased and particularly at low SNR, the bias can be large.
@@
 In contrast, the asymptotic analysis of the amplitude estimator $\hat\rho$ presented in Theorem~1 of our manuscript results in a closed-form expression for the bias and thus, lays the foundation for improving the estimator by addressing the bias.
-However, removal of the bias is beyond the scope of this manuscript.
+However, we confirm that the development of methods for removal of the bias is beyond the scope of this manuscript.
 \item\textbf{Comment:} \label{R4it2e}
@@
 
-The original Sec.~VII describing the Gaussian noise case has been removed from the revised manuscript.
+The original Section~VII describing the Gaussian noise case has been removed from the revised manuscript.
 
-We agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that section specifically discussed random variables of the form $Re^{j\Phi}=1{+}W$, where $W$ is circularly symmetric.
+We agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that section specifically discussed random variables of the form $Re^{j\Phi}=1+W$, where $W$ is circularly symmetric.
 The joint pdf of $R$ and $\Phi$ is a crucial ingredient for the asymptotic analysis presented in the manuscript and therefore, we feel that this material must be retained.
@@
 Regarding Section~II we refer the reader to our response to comment~\eqref{comment:shortersec2} of reviewer~\ref{sec:reviewer-1}.
-
-% Section~II describes our modification of Mackethun's algorithm to compute the least squares estimator.
-Although the modification is simple we have left this section intact since much of the notation required for the statistical analysis of the least squares estimator is most naturally introduced during the description of the algorithm.
-To qualify this for the reader we have included the following sentence at the start of Section~\zref{sec:least-squar-estim}.
-
-% ``In this section we derive Mackenthun's algorithm to compute the least squares estimators of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the non data-aided setting, so we modify the algorithm to include pilot symbols.
-Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\zref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
-For this reason, and so that our paper is self contained, we describe the algorithm in full.''
-
-%For example, equations~\eqref{eq:SSdatasymbols},~\eqref{eq:SSallparams} and~\eqref{eq:hatdfinxtheta} are all used during the
 

### 2014-01-29 paper.tex
added to response

@@
 where $P$ is the set of indices describing the position of the pilot symbols $p_i$, and $D$ is a set of indices describing the position of the data symbols $d_i$.
-The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$  where $\emptyset$ is the empty set, and we let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
+The sets $P$ and $D$ are disjoint, i.e., $P \cap D = \emptyset$  where $\emptyset$ is the empty set.
+We let $L = \abs{P \cup D}$ be the total number of symbols transmitted.
 
@@
 \end{equation}
-where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and amplitude $\rho_0$ (by definition $\rho_0$ is a positive real number).
+where $w_i$ is noise and $a_0 = \rho_0 e^{j\theta_0}$ is a complex number representing both carrier phase $\theta_0$ and amplitude $\rho_0$.
+By definition $\rho_0$ is a positive real number.
 Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
@@
 A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
-\begin{equation}\label{eq:SSdefn}
+\begin{equation}\label{eq:LSdefn}
 \begin{split}
-SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2  \\
+LS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2  \\
 &= \sum_{i \in P} \abs{ y_i - a p_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2,
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Sweldens considered what is called \emph{non-data adied detection} (also know as \emph{blind}, \emph{noncoherent} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
+Both Mackenthun and Sweldens considered what is called \emph{non-data-aided detection} (also know as \emph{blind}, \emph{noncoherent} or \emph{non-decision-directed detection}) where no pilot symbols exist ($P = \emptyset$).
 We show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
-Our model includes the non data-aided case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
-
-Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:SSdefn}.
-In the literature, analytic study of non data-aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
+Our model includes the non-data-aided case by setting the number of pilot symbols to zero, that is, putting $P = \emptyset$.
+
+Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:LSdefn}.
+In the literature, analytic study of non-data-aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
 The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
-To the authors knowledge, all of the existing bounds are applicable only in the fully non data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
+To the authors knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
 Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.
@@
 For the case of carrier amplitude the situations is even more complicated.
-A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non data-aided estimators for carrier amplitude exhibit bias.
+A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non-data-aided estimators for carrier amplitude exhibit bias.
 The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
 Neverthess, Cramer-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
-Unfortunately, these bounds may not be meaningful since known non data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
+Unfortunately, these bounds may not be meaningful since known non-data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
 
@@
 We show, under comparatively mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent (asymptotically unbiased) and asymptotically normally distributed.
-However, the least squares amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation $M$.
-To our knowledge, this is the first time an analytic description of the bias or variance of a non data-aided amplitude estimator has been obtained.
+However, the least squares amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution, and the size of the constellation $M$.
+To our knowledge, this is the first time an analytic description of the bias or variance of a non-data-aided amplitude estimator has been obtained.
 Our expressions for the asymptotic variance and bias are simple to compute.
@@
 One response to this criticism is that the least squares estimator should be a popular practical choice because it can be efficiently computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian.
-Perhaps an even more convinving response is that, while the specific results given here apply only to the least square estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
+Perhaps an even more convinving response is that, while the specific results given here apply only to the least squares estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
 For example, in practical systems it is sometimes desirable to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.
@@
 The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010}, would potentially allow this.
-
-More ambitiously, our techniques might allow analysis of non data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,inthft:204506}.
+More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,inthft:204506}.
 We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.
@@
 These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
-A key problem is that the EM algorithm converges correctly only if initialised at some channel estimate sufficiently close to the true channel parameters.
+A key problem is that the EM algorithm converges correctly only if initialised with a channel estimate sufficiently close to the true channel parameters.
 Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-For the purpose of carrier phase and amplitude estimation, the least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.
+The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.
 %However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by applying the EM algorithm will not justify the extra computational cost.
@@
 These simulations agree with the derived asymptotic properties.
+Simulations showing the effectivness of our estimator as an initialiser for a turbo synchonisation proceedure are also presented.
 
@@
 In this section we derive Mackenthun's algorithm to compute the least squares estimators of the carrier phase and amplitude~\cite{Mackenthun1994}.
-Mackenthun specifically considered the non data-aided setting, so we modify the algorithm to include pilot symbols.
+Mackenthun specifically considered the non-data-aided setting, so we modify the algorithm to include pilot symbols.
 Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\ref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
-For this reason, and so that our paper is self contained, we describe the algorithm in full.
-For the purpose of analysing computational complexity, we will assume that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
+For this reason, and so that the paper is self contained, we describe the algorithm in full.
+For the purpose of analysing computational complexity it is assumed that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
 In this case Mackentun's algorithm requires $O(L \log L)$ arithmetic operations.
@@
 \begin{align}
-SS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
-&= \sum_{i \in P \cup D} \big( \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^* \big), \label{eq:SS}
+LS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2 \nonumber \\
+&= \sum_{i \in P \cup D} \big( \abs{y_i}^2 - a s_i y_i^* - a^* s_i^* y_i + aa^* \big), \label{eq:LS}
 \end{align}
-where $*$ denotes complex conjugate.  The minimiser of $SS$ with respect to $a$ as a function of $\{d_i, i \in D\}$ is
+where $*$ denotes complex conjugate.  The minimiser of $LS$ with respect to $a$ as a function of $\{d_i, i \in D\}$ is
 \begin{equation}\label{eq:hata}
@@
 \end{equation}
-where $L = \abs{P \cup D}$ and to simplify our notation we have put 
+where $L = \abs{P \cup D}$ and to simplify notation we have put 
 \[
@@
 Note that $Y$ is a function of the unknown data symbols $\{ d_i, i \in D\}$ and we could write $Y(\{ d_i, i \in D\})$, but have chosen to suppress the argument $(\{ d_i, i \in D\})$ for notational brevity.
-Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:SS} we obtain $SS$ minimised with respect to $a$,
-\begin{equation}\label{eq:SSdatasymbols}
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y}^2,
+Substituting $\frac{1}{L}Y$ for $a$ into~\eqref{eq:LS} we obtain $LS$ minimised with respect to $a$,
+\begin{equation}\label{eq:LSdatasymbols}
+LS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y}^2,
 \end{equation}
-where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ does not depend on the $d_i$.
-The least squares estimators of the data symbols are the minimisers of~\eqref{eq:SSdatasymbols}.
-Observe that given candidate values for the data symbols, we can compute the corresponding $SS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
+where $A = \sum_{i \in P \cup D}\abs{y_i}^2$ does not depend on the data symbols.
+The least squares estimators of the data symbols are the minimisers of~\eqref{eq:LSdatasymbols}.
+Observe that given candidate values for the data symbols, we can compute the corresponding $LS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
 It turns out that there are at most $M\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
@@
 
-To see this, let $a = \rho e^{j\theta}$ where $\rho > 0$ is real.  Now,
+To see this, let $a = \rho e^{j\theta}$ where $\rho > 0$ is real.  Now $LS$ from~\eqref{eq:LS} is to
 \begin{align}
-SS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
-&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:SSallparams}
+LS(\rho, \theta, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - \rho e^{j\theta} s_i }^2  \nonumber \\
+&= \sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} d_i }^2. \label{eq:LSallparams}
 \end{align}
-We have slightly abused notation here by reusing $SS$.
-This should not cause confusion as $SS(a, \{d_i, i \in D\})$, $SS(\rho, \theta, \{d_i, i \in D\})$, and $SS(\{d_i, i \in D\})$ are easily told apart by their arguments.
+We have slightly abused notation here by reusing $LS$.
+This should not cause confusion as $LS(a, \{d_i, i \in D\})$, $LS(\rho, \theta, \{d_i, i \in D\})$, and $LS(\{d_i, i \in D\})$ are easily told apart by their arguments.
 For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
@@
 \[
-z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}
+z_i = \angle{y_i} - \hat{u}_i(0) = \angle{y_i} - \round{\angle{y_i}}.
 \]
@@
 b_i, &  0 \leq \theta < z_i + \frac{\pi}{M} \\
-b_i e^{-j2\pi/M}, & z_i + \frac{\pi}{M} \leq \theta < z_i + \frac{3\pi}{M} \\ 
+b_i e^{-j2\pi/M}, & \frac{\pi}{M} \leq \theta - z_i < + \frac{3\pi}{M} \\ 
 \vdots & \\
-b_i e^{-j2\pi k /M}, & z_i + \frac{\pi(2k - 1)}{M} \leq \theta < z_i + \frac{\pi(2k + 1)}{M}  \\ 
+b_i e^{-j2\pi k /M}, & \frac{\pi(2k - 1)}{M} \leq \theta - z_i < \frac{\pi(2k + 1)}{M}  \\ 
 \vdots & \\
@@
 be a function mapping the interval $[0, 2\pi)$ to a sequence of $M$-PSK symbols indexed by the elements of $D$.
-Observe that $f(\theta)$ is piecewise continuous.
-The subintervals of $[0, 2\pi)$ over which $f(\theta)$ remains constant are determined by the values of $\{z_i, i \in D\}$.
+Observe that $f$ is piecewise continuous.
+The subintervals of $[0, 2\pi)$ over which $f$ remains constant are determined by the values of $\{z_i, i \in D\}$.
 Let
@@
 be the set of all sequences $f(\theta)$ as $\theta$ varies from $0$ to $2\pi$.
-If $\hat{\theta}$ is the least squares estimator of the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:SSdatasymbols}.
-Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$, because there are $M$ distinct values of $\hat{d}_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
+If $\hat{\theta}$ is the least squares estimator of the phase then $S$ contains the sequence $\{ \hat{d}_i(\hat{\theta}), i \in D \}$ corresponding to the least squares estimator of the data symbols, i.e., $S$ contains the minimiser of~\eqref{eq:LSdatasymbols}.
+Observe from~\eqref{eq:uicombos} that there are at most $M\abs{D}$ sequences in $S$ because there are $M$ distinct values of $\hat{d}_i(\theta)$ for each $i \in D$ as $\theta$ varies from $0$ to $2\pi$.
 
@@
 
-We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:SSdatasymbols}).
-A na\"{\i}ve approach would be to compute $SS(f_k)$ for each $k \in \{0,1,\dots,M\abs{D}-1\}$.
-Computing $SS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
+We want to find the $f_k \in S$ corresponding to the minimiser of~(\ref{eq:LSdatasymbols}).
+A na\"{\i}ve approach would be to compute $LS(f_k)$ for each $k \in \{0,1,\dots,M\abs{D}-1\}$.
+Computing $LS(f_k)$ for any particular $k$ requires $O(L)$ arithmetic operations.
 So, this na\"{\i}ve approach would require $O(L M \abs{D}) = O(L^2)$ operations in total.
-Following Mackenthun~\cite{Mackenthun1994}, we show how $SS(f_k)$ can be computed recursively.
+Following Mackenthun~\cite{Mackenthun1994}, we show how $LS(f_k)$ can be computed recursively.
 
 Let,
-\begin{equation}\label{eq:SSfk}
-SS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
+\begin{equation}\label{eq:LSfk}
+LS(f_k) = A - \frac{1}{L}\abs{Y_k}^2,
 \end{equation}
@@
 So, each $Y_k$ can be computed from its predecessor $Y_{k-1}$ in $O(1)$ arithmetic operations.
-Given $Y_k$, the value of $SS(f_k)$ can be computed in $O(1)$ operations using~\eqref{eq:SSfk}.
-Let $\hat{k} = \arg\min SS(f_k)$.
+Given $Y_k$, the value of $LS(f_k)$ can be computed in $O(1)$ operations using~\eqref{eq:LSfk}.
+Let $\hat{k} = \arg\min LS(f_k)$.
 The least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata},
@@
 So, for example, we expect the phases $0.99\pi$ and $-0.99\pi$ to be close together, the difference between them being $\vert\sfracpart{0.99\pi + 0.99\pi}_\pi\vert = 0.02\pi$, and not $\vert 0.99\pi + 0.99\pi\vert = 1.98\pi$.
-In the literature it is usual to write \emph{estimator} minus \emph{error}, i.e., $\hat{\theta} - \theta_0$ rather than $\theta_0 - \hat{\theta}$.
+%In the literature it is usual to write \emph{estimator} minus \emph{error}, i.e., $\hat{\theta} - \theta_0$ rather than $\theta_0 - \hat{\theta}$.
 However, it will be notationally convenient in Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} to use $\langle\theta_0 - \hat{\theta}\rangle_\pi$.
-The theorems require minor modification when the number of pilots symbols $\abs{P} = 0$.
-Theorem statements for this special case appear in \cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+The theorems require minor modification when the number of pilots symbols $\abs{P} = 0$.  Theorem statements for this special case appear in \cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
 
@@
 The value $G(0)$ is a function of the noise distribution and $G(0)$ is close to $1$ when the variance of the noise is small, but grows without bound as the noise variance increases.
-That is, the asymptotic bias is small when the SNR is large, but significant when the SNR is small.
-We will further investigate the bias in Sec.~\ref{sec:simulations}.
-Theorem~\ref{thm:normality} shows that $\sqrt{L}(\hat{\rho} - \rho_0 G(0))$ (i.e.\ the amplitude estimator error with bias removed) and $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ are asymptotically normally distributed and statistically independent as $L\rightarrow\infty$.
+Thus, the asymptotic bias is small when the SNR is large, but significant when the SNR is small.
+We will further investigate the bias in Section~\ref{sec:simulations}.
+Theorem~\ref{thm:normality} shows that $\sqrt{L}(\hat{\rho} - \rho_0 G(0))$ (the amplitude estimator error with bias removed) and $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ are asymptotically normally distributed and statistically independent as $L\rightarrow\infty$.
 
 We would like to stress that many practical estimators such as the SNR estimators in~\cite{Pauluzzi2000} are biased and, particularly at low SNR, the bias can be large.
-To the best of the authors' knowledge, the value of the bias of these estimators has only been approximated by computer simulation and no theoretical analysis exists.
+To the authors' knowledge, the value of the bias of these estimators has only been approximated by computer simulation and no theoretical analysis exists~\cite{Pauluzzi2000}.
 In contrast, we have presented a closed-form expression for the asymptotic bias of $\hat\rho$, thereby laying a foundation for an improved estimator.
@@
 This places mild restrictions on the distribution of the noise $w_i$.
-For example, the requirements are satisfied if the joint pdf of the real and imaginary parts of $w_i$ is continuous, since in this case $f(r,\phi)$ is continuous.
-Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$ it follows that $g$ has period $2\pi$ and is even on $[-\pi, \pi]$.
+For example, the requirements are satisfied if the joint pdf of the real and imaginary parts of $w_i$ is continuous because then
+$f(r,\phi)$ is continuous.  Because $f(r,\phi)$ has period $2\pi$ and is even on $[-\pi,\pi]$ with respect to $\phi$ it follows that $g$ has period $2\pi$ and is even on $[-\pi, \pi]$.
 
@@
 A key assumption in Theorem~\ref{thm:consistency} is that $G(x)$ is uniquely maximised at $x = 0$ for $x \in [-\pi, \pi)$.
-This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence of numbers from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
+This assumption asserts that $G(x) \leq G(0)$ for all $x \in [-\pi, \pi)$ and that if $\{x_i\}$ is a sequence from $[-\pi,\pi)$ such that $G(x_i) \rightarrow G(0)$ as $i \rightarrow \infty$ then $x_i \rightarrow 0$ as $i \rightarrow \infty$.
 Although we will not prove it here, this assumption is not only sufficient, but also necessary, for if $G(x)$ is uniquely maximised at some $x \neq 0$ then $\sfracpart{\theta_0 - \hat{\theta}}_\pi \rightarrow x$ almost surely as $L\rightarrow\infty$, while if $G(x)$ is not uniquely maximised then $\sfracpart{\theta_0 - \hat{\theta}}_\pi$ will not converge.
-One can check that this assumption holds when $w_i$ is circularly symmetric and normally distributed.
+One can check that this assumption holds in the important case when $w_i$ is circularly symmetric and normally distributed.
 We will not attempt to further classify those distributions for which the assumption holds here.
@@
 The proof of Theorem~\ref{thm:consistency} hinges on Lemmas \ref{lem:convtoexpGlamL} to~\ref{lem:GLtoG0}, which can be found in the appendix.
-Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:SSallparams} we obtain $SS$ minimised with respect to the data symbols,
+Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:LSallparams} we obtain $LS$ minimised with respect to the data symbols,
  \begin{align*}
-SS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
+LS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
 &= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
@@
 \begin{equation}\label{eq:hatrhoZ}
-\hat{\rho}(\theta) = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)),
+\hat{\rho}(\theta) = \frac{Z(\theta) + Z^*(\theta)}{2L} = \frac{1}{L}\Re(Z(\theta)).
 \end{equation}
-where $\Re(\cdot)$ denotes the real part.  %Since $\hat{\rho}(\theta)$ is nonnegative it follows that $\Re(Z(\theta))$ is nonnegative for all $\theta \in [-\pi, \pi)$.  
-Substituting this expression into $SS(\rho, \theta)$ gives $SS$ minimised with respect to $\rho$ and the data symbols,
-\[
-SS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
-\]
-%We again abuse notation by reusing $SS$, but this should not cause confusion as $SS(\rho,\theta)$ and $SS(\theta)$ are easily told apart by their inputs.  
+%where $\Re(\cdot)$ denotes the real part.  %Since $\hat{\rho}(\theta)$ is nonnegative it follows that $\Re(Z(\theta))$ is nonnegative for all $\theta \in [-\pi, \pi)$.  
+Substituting this expression into $LS(\rho, \theta)$ gives $LS$ minimised with respect to $\rho$ and the data symbols,
+\[
+LS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
+\]
+%We again abuse notation by reusing $LS$, but this should not cause confusion as $LS(\rho,\theta)$ and $LS(\theta)$ are easily told apart by their inputs.  
 By definition the amplitude $\rho_0$ and its estimator $\hat{\rho}$ are positive.
@@
 %\[
-%SS(\rho,\theta) = SS(-\rho, \fracpart{\theta + \pi})
+%LS(\rho,\theta) = LS(-\rho, \fracpart{\theta + \pi})
 %\]
-The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $SS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
+The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $LS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
 Equivalently $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ with no constraints required.
@@
 \end{align}
-and put $\hat{\lambda}_L = -\sfracpart{\hat{\theta} - \theta_0}_\pi = \sfracpart{\theta_0 - \hat{\theta}}_\pi$.
+and put $\hat{\lambda}_L = \sfracpart{\theta_0 - \hat{\theta}}_\pi$.
 Since $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ it follows that $\hat{\lambda}_L$ is the maximiser of $G_L(\lambda)$.
@@
 \]
+%This is kind of disjointed without the ULL included.  Unfortunately there is no space.
 As is customary, let $\Omega$ be the sample space on which the random variables $\{w_i\}$ are defined.
@@
 The proof is complicated by the fact that $G_L$ is not differentiable everywhere due to the function $\fracpart{\cdot}$ not being differentiable at multiples of $\tfrac{\pi}{M}$.
-This prevents the use of ``standard approaches'' to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989,van2009empirical}.
-However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist, and is equal to zero, at $\hat{\lambda}_L$.
+This prevents the use of standard approaches to proving normality that are based on the mean value theorem~\cite{vonMises_diff_stats_1947,vanDerVart1971_asymptotic_stats,Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989,van2009empirical}.
+However, Lemma~\ref{lem:diffatlambdaL} shows that the derivative $G_L^\prime$ does exist and is equal to zero at $\hat{\lambda}_L$.
 %Similar properties have been used by some of the present authors to analyse the behaviour of polynomial-phase estimators~\cite{McKilliam_LSU_polyest_part_arxiv_2012}.
@@
 \end{equation}
-Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$, and so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.
-Let $Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)$ and write
+Whenever $G_L(\lambda)$ is differentiable $G_L^\prime(\lambda) = R_L(\lambda)$ and so $R_L(\hat{\lambda}_L) = G_L^\prime(\hat{\lambda}_L) = 0$ by Lemma~\ref{lem:diffatlambdaL}.  Define the function 
+\[
+Q_L(\lambda) = \expect R_L(\lambda) - \expect R_L(0)
+\]
+and write
 \begin{align*}
@@
 \[
-0 = o_P(1) + \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big)
+0 = \sqrt{L}R_L(0) + \sqrt{L} \hat{\lambda}_L \big( p + Hd  + o_P(1) \big) + o_P(1)
 \]
@@
 \[
-\sqrt{L} \hat{\lambda}_L = o_P(1) - \frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)}.
+\sqrt{L} \hat{\lambda}_L =  -\frac{\sqrt{L}R_L(0)}{p + Hd  + o_P(1)} + o_P(1).
 \]
@@
 \[
-\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = o_P(1) + X_L,
+\sqrt{L}\big( G_L(\hat{\lambda}_L) - T_L(\hat{\lambda}_L)  \big) = X_L + o_P(1),
 \]
@@
 \]
-as $L \rightarrow \infty$.  The lemmas that we have used are proved in the appendix.
+as $L \rightarrow \infty$.  This completes the proof of Theorem~\ref{thm:normality}.
 
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non data-aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the sample MSE of the non-data-aided phase estimator of Viterbi and Viterbi~\cite{ViterbiViterbi_phase_est_1983}.
 This estimator requires selection of a function $F$ that transforms the amplitude of each sample prior to the final estimation step.
@@
 Viterbi and Viterbi propose several viable alternatives, from which we have chosen $F(x) = 1$.
-The Viterbi and Viterbi estimator is only applicable in the non data-aided setting, i.e.
+The Viterbi and Viterbi estimator is only applicable in the non-data-aided setting, i.e.
 when $\abs{P} = 0$.
@@
 When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
-However, $G(0)$ grows without bound as the variance of the noise increases, so the bias is significant when SNR is small.
+However, $G(0)$ grows without bound as the variance of the noise increases so the bias is significant when SNR is small.
 Figure~\ref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
@@
  %This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
-For example, rather than minimise~\eqref{eq:SSdefn} we could instead minimise a weighted version of it,
- % \[
- % SS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
+For example, rather than minimise~\eqref{eq:LSdefn} we could instead minimise a weighted version of it,
+ % \[
+ % LS_{\beta}(a, \{d_i, i \in D\}) = \sum_{i \in P} \abs{ y_i - a s_i }^2 + \beta \sum_{i \in D} \abs{ y_i - a d_i }^2,
  % \] 
  % where the weight $\beta$ would be small when SNR is small and near $1$ when SNR is large.
-Computing the $\hat{a}$ that minimises $SS_\beta$ can be achieved with only a minor modification to algorithm~\ref{alg:loglinear}.
+Computing the $\hat{a}$ that minimises $LS_\beta$ can be achieved with only a minor modification to algorithm~\ref{alg:loglinear}.
 Line~\ref{alg_gset} is modified to $g_i = \beta y_i e^{-j u}$ and lines~\ref{alg_hata1}~and~\ref{alg_hata2} are modified to $\hat{a} =  \frac{1}{\abs{P} + \beta\abs{D}} Y$.
@@
 We focused on $M$-ary phase shift keying constellations.
-The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994}, and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
-
-We showed, under some reasonably general conditions on the distribution of the noise, that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
+The least squares estimator can be computed in $O(L\log L)$ operations using a modification of an algorithm due to Mackenthun~\cite{Mackenthun1994} and is the maximum likelihood estimator in the case that the noise is additive white and Gaussian.
+
+Under reasonably general conditions on the distribution of the noise, it was shown that the phase estimator $\hat{\theta}$ is strongly consistent and asymptotically normally distributed.
 However, the amplitude estimator $\hat{\rho}_0$ is biased, and converges to $G(0)\rho_0$.
@@
 %This suggests the use of a weighted objective function, which would be an interesting topic for future research.
+The simulations also indicate our estimator to be an effective initialiser for turbo synchronisation proceedures.
 
@@
 \[
-SS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
-\]
-defined in~\eqref{eq:SSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
+LS(\{d_i, i \in D\}) = A - \frac{1}{L}\sabs{Y(\{d_i, i \in D\})}^2,
+\]
+defined in~\eqref{eq:LSdatasymbols}. The proof now proceeds by contradiction.  Assume that 
 \begin{equation}\label{eq:fraclambassumption}
@@
 \[
-SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
-\]
-violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  First observe that,
+LS(\{r_i, i \in D\}) < LS(\{\hat{d}_i, i \in D\}),
+\]
+violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $LS$.  First observe that,
 \begin{align*}
@@
 \begin{align*}
-SS&(\{r_i, i \in D\}) \\
+LS&(\{r_i, i \in D\}) \\
 &= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
@@
 &= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
-&= SS(\{\hat{d}_i, i \in D\}) - C,
+&= LS(\{\hat{d}_i, i \in D\}) - C,
 \end{align*}
@@
 \[
-SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\}),
-\] 
-violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $SS$.  So~\eqref{eq:fraclambassumption} is false.
+LS(\{r_i, i \in D\}) < LS(\{\hat{d}_i, i \in D\}),
+\] 
+violating the fact that $\{\hat{d}_i, i \in D\}$ are minimisers of $LS$.  So~\eqref{eq:fraclambassumption} is false.
 
@@
 Recalling the notation $e_k^*$ defined in Section~\ref{sec:least-squar-estim}, put $r_i = \hat{d}_i e_k^*$.
-Now an analogous argument can be used to show that $SS(\{r_i, i \in D\}) < SS(\{\hat{d}_i, i \in D\})$ again.
+Now an analogous argument can be used to show that $LS(\{r_i, i \in D\}) < LS(\{\hat{d}_i, i \in D\})$ again.
 \end{IEEEproof}
@@
 An analogous argument follows when $-\tfrac{\pi}{M} < \lambda < 0$.
-To keep our notation clean put
+To keep notation clean put
 \[
@@
 \]
-From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$ as a result of~\eqref{eq:expectImpartRphi}, and variance
+From the standard central limit theorem the distribution of $C_L$ converges to the normal with mean $\sqrt{p}\expect R_1 \sin(\Phi_1) = 0$, as a result of~\eqref{eq:expectImpartRphi}, and variance
 \[

### 2014-01-30 code/PEG/main.cpp
added scripts for generating RA codes.  Added metapost plot for turbo sync sims

@@ -1,0 +1,236 @@
+/************************************************************************/
+/*                                                                      */
+/*        Free software: Progressive edge-growth (PEG) algorithm        */
+/*        Created by Xiaoyu Hu                                          */
+/*                   Evangelos Eletheriou                               */
+/*                   Dieter Arnold                                      */
+/*        IBM Research, Zurich Research Lab., Switzerland               */
+/*                                                                      */
+/*        The C++ sources files have been compiled using xlC compiler   */
+/*        at IBM RS/6000 running AIX. For other compilers and platforms,*/
+/*        minor changes might be needed.                                */
+/*                                                                      */
+/*        Bug reporting to: xhu@zurich.ibm.com                          */
+/************************************************************************/
+
+/************************************************************************/
+/*                                                                      */
+/*        Modifications made by Gottfried Lechner                       */
+/*                                                                      */
+/*        - fixed a memory leak that lead to a crash for long codes     */
+/*        - added the option to supply the check node distribution      */
+/*        - changes such that codes with M>N can be constructed         */
+/*        - restructuring of the code                                   */
+/*        - changed parameter checking to ignore case                   */
+/*                                                                      */
+/*        Bug reporting to: gottfried.lechner@unisa.edu.au              */
+/************************************************************************/
+
+#include <stdlib.h>
+#include <stdio.h>
+#include <string.h>
+#include <iostream>
+#include <iomanip>
+#include <fstream>
+#include <math.h>
+
+#include "BigGirth.h"
+
+using namespace std;
+
+#define EPS  1e-6
+
+int main(int argc, char * argv[]){
+  long int	i, j, m, mc, N, M, n, ev, ec;
[…]

### 2014-02-03 paper.tex
added scripts for generating RA codes.  Added metapost plot for turbo sync sims

@@
 In this case, our expressions for the asymptotic distribution can be simplified.
-Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.
-These simulations agree with the derived asymptotic properties.
+Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.  These simulations agree with the derived asymptotic properties.  
 Simulations showing the effectivness of our estimator as an initialiser for a turbo synchonisation proceedure are also presented.
@@
 
-Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
+%Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.
 The figure depicts an interesting phenomenon.
@@
 
+% \begin{figure}[tp]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
+% 		\caption{Phase error versus SNR for QPSK.}
+% 		\label{fig:plotphaseQPSKmultL}
+% \end{figure}
+
 \begin{figure}[tp]
 	\centering
-		\includegraphics[width=\linewidth]{code/data/plotM4-3.mps}
-		\caption{Phase error versus SNR for QPSK.}
+		\includegraphics[width=\linewidth]{code/data/plotturbo-1.mps}
+		\caption{Probability of bit error $P_B$ with turbo synchronisation.}
 		\label{fig:plotphaseQPSKmultL}
@@
 Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
-Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
+%Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
 %This suggests the use of a weighted objective function, which would be an interesting topic for future research.

### 2014-02-04 paper.tex
finished of QPSK turbo sync code.

--- 2014-01-29 paper.tex
@@
 In this paper we assume that the time offset has been previously handled and we focus on estimating the phase shift and attenuation.
-We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK) and $M$-ary phase shift keying ($M$-PSK).
+We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-ary phase shift keying ($M$-PSK).
 In this case, the transmitted symbols take the form,
@@
 Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:LSdefn}.
-In the literature, analytic study of non-data-aided estimators of carrier phase have primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
+In the literature, analytic study of non-data-aided estimators of carrier phase has primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
 The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
@@
 The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.
-%However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by applying the EM algorithm will not justify the extra computational cost.
+However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by turbo synchronisation will not justify the extra computational cost.
+
+Turbo synchrnoisation can require a large number of EM iterations.
+For applications with low power receivers, such as sensor networks, the complexity of turbo sychronisation might be too great and computationally simple channel estimators are required.
+For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.
+In this case it is necessary that the number of pilots symbols is sufficiently large for channel estimates to be sufficiently accurate.
+If more accuracy is required, then more pilots symbols must be used.
+Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols, but instead using a subset of the data symbols.
+For computational complexity reasons it may not be desirable to use all of the existing data symbols for channel estimation.
+This is easily possible with our estimator by choosing the indices in $D$ to correspond with the subset of the data symbols that are to be used.
 
@@
 In this case, our expressions for the asymptotic distribution can be simplified.
-Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.
-These simulations agree with the derived asymptotic properties.
-Simulations showing the effectivness of our estimator as an initialiser for a turbo synchonisation proceedure are also presented.
+Section~\ref{sec:simulations} presents the results of Monte-Carlo simulations.  These simulations agree with the derived asymptotic properties.  
+%Simulations showing the effectivness of our estimator as an initialiser for a turbo synchonisation proceedure are also presented.
 
@@
 \end{equation} 
-Lemma~\ref{lem:GLtoG0} in the appendix shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$, and $\hat{\rho}$ consequently converges almost surely to $\rho_0 G(0)$ as required.
+Lemma~\ref{lem:GLtoG0} in the appendix shows that $G_L(\hat{\lambda}_L)$ converges almost surely to $G(0)$ as $L\rightarrow\infty$.
+Consequently, $\hat{\rho}$ converges almost surely to $\rho_0 G(0)$ as required.
 %It remains to prove Lemmas~\ref{lem:convtoexpGlamL} and~\ref{lem:GLtoG0}.
@@
 
+% \begin{figure}[tp]
+% 	\centering
+% 		\includegraphics[width=\linewidth]{code/data/plotturbo-1.mps}
+% 		\caption{Probability of bit error $P_B$ with turbo synchronisation.}
+% 		\label{fig:plotphaseQPSKmultL}
+% \end{figure}
+
 % \section{When the SNR is small}
@@
 Monte Carlo simulations were used to assess the performance of the least squares estimator and also to validate our asymptotic theory.
-Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
+%Interestingly, when the SNR is small, it is counterproductive to use the data symbols to estimate the phase (Figure~\ref{fig:plotphaseQPSKmultL}).
 %This suggests the use of a weighted objective function, which would be an interesting topic for future research.
-The simulations also indicate our estimator to be an effective initialiser for turbo synchronisation proceedures.
+%The simulations also indicate our estimator to be an effective initialiser for turbo synchronisation proceedures.
 

### 2014-02-04 code/turbosyncsim.cpp

@@ -47,7 +47,6 @@ void runperfectchannelsim(const function<CodedConstellation*()> codedconstf, con
   	double snrdB = snrdbs[snrind]; 
   	double var = amplitude*amplitude*pow(10, -snrdB/10); //variance of real and imaginary parts of noise
   	default_random_engine gen((long)clock()); //random number generator, seed with clock 
-  	uniform_int_distribution<int> unifM(0, M-1); //for generating M-PSK symbols
   	uniform_real_distribution<double> unifphase(-pi,pi); //generator for true timeoffset 
   	normal_distribution<double> gn(0.0,sqrt(var));
 
@@ -62,11 +61,11 @@ void runperfectchannelsim(const function<CodedConstellation*()> codedconstf, con
   	int itrcount = 0; //count total number of iterations
   	while( numerrs < toerrs ) {
   	  const complex<double> a0 = polar<double>(amplitude, unifphase(gen)); //amplitude random phase
-  	  for(int i = 0; i < K; i++) info[i] = (unsigned int) unifM(gen); //generate some random bits
+  	  for(int i = 0; i < K; i++) info[i] = rand() % 2; //generate some random bits
   	  const vector<complexd>& cw = codec->encode(info); //encode the bits, result goes in cw
-  	  for(int i = 0; i < L; i++) s[D[i]] = cw[i]; //convert to M-psk symbols
+  	  for(int i = 0; i < L; i++) s[D[i]] = cw[i]; //fill data symbols
   	  for(int i = 0; i < L; i++) y[i] = a0*s[i] + complexd(gn(gen), gn(gen)); //compute transmitted signal
-  	  const std::vector<unsigned int>& rbits = invdec.decode(y, a0, var); //decode recieved signal
+  	  const std::vector<unsigned int>& rbits = invdec.decode(y, a0, 2*var); //decode recieved signal (pass in variance of complex noise!)
   	  for(int i = 0; i < K; i++) numerrs += (rbits[i] == info[i]) ? 0 : 1;
   	  itrcount++;
   	}
@@ -91,7 +90,7 @@ void runperfectchannelsim(const function<CodedConstellation*()> codedconstf, con
 }
 
 //run simulation with given initial channel estimator
[…]

### 2014-02-04 paper.tex
added bpsk crbs. qpsk is next.

@@
 For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.
-In this case it is necessary that the number of pilots symbols is sufficiently large for channel estimates to be sufficiently accurate.
+In this case the number of pilots symbols must be sufficiently large for channel estimates to be sufficiently accurate.
 If more accuracy is required, then more pilots symbols must be used.
-Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols, but instead using a subset of the data symbols.
-For computational complexity reasons it may not be desirable to use all of the existing data symbols for channel estimation.
+Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols.
+Instead, the pilots symbols and a subset of the data symbols can be jointly used for channel estimation.
+%For computational complexity reasons it may not be desirable to use all of the existing data symbols for channel estimation.
 This is easily possible with our estimator by choosing the indices in $D$ to correspond with the subset of the data symbols that are to be used.

### 2014-02-04 response.tex
minor edits to response

@@
 
-We have left the description of the algorithm intact.
-The primary reason for this is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.
-To qualify this for the reader we have included the following sentences at the beginning of Section~\zref{sec:least-squar-estim}.
+In the new manuscript only two pages are dedicated to the description of the algorithm.
+An important reason for describing the algorithm is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.
+For example, the proof of Lemma~\zref{lem:diffatlambdaL} makes use of the notation $e_i$ and the function $LS(\{d_i, i \in D\})$ introduced during the description of the algorithm.
+To qualify this for the reader we have included the following sentences at the beginning of Section~\zref{sec:least-squar-estim}:
 

### 2014-02-05 response.tex
added edits to response. All that's left is to tidy the proof of Lemma 4, and it's useful for me to do this anyway

@@
 
-In the new manuscript only two pages are dedicated to the description of the algorithm.
+The description of the algorithm has been left intact.
 An important reason for describing the algorithm is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.
@@
 
-	\comment{Add CRBs to plots.}
+We have included CRBs for figures in the simulations section.  The follow paragraph has been included:
+
+``Figures~\zref{fig:plotphaseBPSK} and~\zref{fig:plotphaseQPSK} display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.
+To the authors knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).
+These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=\emptyset$.''
+
 \end{enumerate}
@@
 In the original manuscript we plotted the \emph{variance} of the amplitude estimator, not the mean square error.
-That is, the solid lines where the theoretical variance of $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$ and each point from the Monte-Carlo simulations was computed as the unbiased error $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
+That is, the solid lines displayed the asymptotic variance of $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$ and each point from the Monte-Carlo simulations was computed as the unbiased error $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.
 We chose this approach in order to highlight the correctness of our asymptotic theory.
@@
 We do agree with the reviewer this this plot by itself initially appears misleading.
-To alleviate this confusion in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the bias plus variance squared.
-The following paragraph has also been included in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
+To alleviate this confusion in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the asymptotic squared bias plus variance.
+The following paragraph has also included in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
 
@@
 shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
-The solid lines show the mean square error predicted by the bias $\rho_0\big(1 - G(0)\big)$ squared plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\zref{thm:normality}.
+The solid lines show the mean square error predicted by the squared bias $\rho_0^2\big(1 - G(0)\big)^2$ plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\zref{thm:normality}.
 That is, the solid lines are given by
@@
 	\\\textbf{Response:}
-	We have added the following paragraph to the introduction describing the need for sufficiently accurate coarse estimates to initialise an iterative code-aided synchronisation stage. %citing [d, e].
+	We have added the following paragraphs to the introduction to place our results in context with exiting turbo synchronisation literature.
 
-``In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_timingturbosynch_2007,Herzet_framework_turbo_sync_2007}.
-These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
+``In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_framework_turbo_sync_2007}.
+These methods often go by the name of~\emph{turbo synchronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
 These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
@@
 Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.''
+The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.
+However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by turbo synchronisation will not justify the extra computational cost.
 
-\comment{BLERG, add LDPC turbo sync sims}
+Turbo synchronisation can require a large number of EM iterations.
+For applications with low power receivers, such as sensor networks, the complexity of turbo synchronisation might be too great and computationally simple channel estimators are required.
+For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.
+In this case the number of pilots symbols must be sufficiently large for channel estimates to be sufficiently accurate.
+If more accuracy is required, then more pilots symbols must be used.
+Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols.
+Instead, the pilots symbols and a subset of the data symbols can be jointly used for channel estimation.''
 
-%We have also added a set of new simulation results, which show the packet error rate of an LDPC-coded satellite communication system to demonstrate the use and performance of the proposed estimator in a practical system.
-\comment{Add details\dots}
-
-%	\comment{Add response related to [f, g]\dots}
 \end{enumerate}
@@
 	\\\textbf{Response:}
-We thank their reviewer for their comments. We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
+We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
 
@@
 	\\\textbf{Response:}
-We have included simulations with turbo synchronisation proceedures.  See our response to comment~\ref{rev2:turbosync} of reviewer~2.
+See our response to comment~\ref{rev2:turbosync} of reviewer~2.
 
@@
 In contrast, the asymptotic analysis of the amplitude estimator $\hat\rho$ presented in Theorem~1 of our manuscript results in a closed-form expression for the bias and thus, lays the foundation for improving the estimator by addressing the bias.
-However, we confirm that the development of methods for removal of the bias is beyond the scope of this manuscript.
+The development of methods for removal of the bias is non-trivial and constitutes an entirely new avenue of research.
+It is beyond the scope of this manuscript.
+
 \item\textbf{Comment:} \label{R4it2e}
@@
 	\\\textbf{Response:}
-
-The original Section~VII describing the Gaussian noise case has been removed from the revised manuscript.
-
-We agree with the reviewer that circularly symmetric random variables are well known, but would like to point out that section specifically discussed random variables of the form $Re^{j\Phi}=1+W$, where $W$ is circularly symmetric.
+The original Section~7 describing the Gaussian noise case has been removed from the revised manuscript.
+The original Section~3 has been removed.
+Necessary results about circular random variables are now introduced as needed.
+%We agree with the reviewer that circularly symmetric random variables (described in Section~3 of the original manuscript) are well known, but point out that  section specifically discussed random variables of the form $Re^{j\Phi}=1+W$, where $W$ is circularly symmetric.
 The joint pdf of $R$ and $\Phi$ is a crucial ingredient for the asymptotic analysis presented in the manuscript and therefore, we feel that this material must be retained.
@@
 It is now introduced as needed when the asymptotic properties of the least squares estimator are derived.
-
-Regarding Section~II we refer the reader to our response to comment~\eqref{comment:shortersec2} of reviewer~\ref{sec:reviewer-1}.
+Regarding Section~2 we refer the reviewer to our response to comment~\eqref{comment:shortersec2} of reviewer~1.
 

### 2014-02-05 paper.tex
added edits to response. All that's left is to tidy the proof of Lemma 4, and it's useful for me to do this anyway

@@
 Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:LSdefn}.
-In the literature, analytic study of non-data-aided estimators of carrier phase has primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
-The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
+In the literature, analytic study of non-data-aided estimators of carrier phase has primarily involved the derivation of Cram\'{e}r-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
+The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
 To the authors knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
@@
 The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
-Neverthess, Cramer-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
+Neverthess, Cram\'{e}r-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
 Unfortunately, these bounds may not be meaningful since known non-data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
@@
 
-%Two criticisms might be directed at our asymptotic approach when compared to the Cramer-Rao bounds already developed in the literature.
+%Two criticisms might be directed at our asymptotic approach when compared to the Cram\'{e}r-Rao bounds already developed in the literature.
 Perhaps the most obvious is that the results apply asymptotically as $L\to\infty$.
@@
 
-One criticism that might be directed at our asymptotic approach when compared to the Cramer-Rao bounds already developed in the literature is that our specific results apply only to the least square estimator.
+One criticism that might be directed at our asymptotic approach when compared to the Cram\'{e}r-Rao bounds already developed in the literature is that our specific results apply only to the least square estimator.
 One response to this criticism is that the least squares estimator should be a popular practical choice because it can be efficiently computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian.
@@
 More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,inthft:204506}.
-We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.
+We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cram\'{e}r-Rao lower bounds.
 
@@
 In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_framework_turbo_sync_2007}.
-These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
+These methods often go by the name of~\emph{turbo synchronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.
 These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.
@@
 Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.
-The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.
+The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.
 However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by turbo synchronisation will not justify the extra computational cost.
 
-Turbo synchrnoisation can require a large number of EM iterations.
-For applications with low power receivers, such as sensor networks, the complexity of turbo sychronisation might be too great and computationally simple channel estimators are required.
+Turbo synchronisation can require a large number of EM iterations.
+For applications with low power receivers, such as sensor networks, the complexity of turbo synchronisation might be too great and computationally simple channel estimators are required.
 For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.
@@
 
+Figures~\ref{fig:plotphaseBPSK} and~\ref{fig:plotphaseQPSK} display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.
+To the authors knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).
+These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=\emptyset$.
+
 % Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
@@
 shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
-The solid lines show the mean square error predicted by the bias $\rho_0\big(1 - G(0)\big)$ squared plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\ref{thm:normality}.
+The solid lines show the mean square error predicted by the squared bias $\rho_0^2\big(1 - G(0)\big)^2$ plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\ref{thm:normality}.
 That is, the solid lines are given by
@@
 A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
-This behaviour suggests that using the data symbols is counterproductive when the SNR is low.
+This behaviour suggests that using the data symbols is counterproductive when the SNR is low and the proportion of pilot symbols is sufficiently large.
 

### 2014-02-05 response.tex
minor

@@
 An important reason for describing the algorithm is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.
-For example, the proof of Lemma~\zref{lem:diffatlambdaL} makes use of the notation $e_i$ and the function $LS(\{d_i, i \in D\})$ introduced during the description of the algorithm.
+For example, the proof of Lemma~\zref{lem:fracpartlambdahatnotpi} makes use of the notation $e_i$ and the function $LS(\{d_i, i \in D\})$ introduced during the description of the algorithm.
 To qualify this for the reader we have included the following sentences at the beginning of Section~\zref{sec:least-squar-estim}:
@@
 	\\\textbf{Response:}	
-	For improved readability of the manuscript, we have now moved all of the lemmas into the appendix.
+	For improved readability of the manuscript, we have moved all of the lemmas into the appendix.
 We have also added a discussion of the implications and the practical relevance of the theorems after the statement of Theorems~\zref{thm:consistency} and~\zref{thm:normality} in Section~\zref{sec:stat-prop-least}.
-We have also expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
+The introduction has been expanded to make the practical and theoretical importance of our of results clear.
 
@@
 
-We have included CRBs for figures in the simulations section.  The follow paragraph has been included:
+We have included CRBs for figures in the simulations section.  The following paragraph has been included:
 
@@
 Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non-data-aided SNR estimator in order to implement the reviewers proposal.
-We confirm, however, that analysing and implementing the reviewers suggested technique is well beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
+We confirm, however, that analysing and implementing the reviewers suggested technique is beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
 

### 2014-02-11 paper.tex
minor
added Andre's comments

@@
 Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990,580211,974597,Mengali_andre_synchro_book}.
-A practical approach is the least squares estimator, that is, the minimisers of the sum of squares function
+A practical approach is the least squares estimator, that is, the minimiser of the sum of squares function
 \begin{equation}\label{eq:LSdefn}
@@
 Sweldens~\cite{Sweldens2001} rediscovered Mackenthun's algorithm in 2001.
-Both Mackenthun and Sweldens considered what is called \emph{non-data-aided detection} (also know as \emph{blind}, \emph{noncoherent} or \emph{non-decision-directed detection}) where no pilot symbols exist ($P = \emptyset$).
+Both Mackenthun and Sweldens considered what is called \emph{non-data-aided detection} (also know as \emph{blind}, or \emph{non-decision-directed detection}) where no pilot symbols exist ($P = \emptyset$).
 We show in Section~\ref{sec:least-squar-estim}~that Mackenthun's algorithm can be modified to include pilot symbols.
@@
 The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
-To the authors knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
+To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
 Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.
@@
 Based on our Theorem~\ref{thm:normality} in Section~\ref{sec:stat-prop-least} the performance of the least squares estimator can be accurately predicted without the need for time consuming Monte-Carlo simulations.
-This provides easy to use and rigourously justified methods for choosing systems parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.
+This provides easy to use and rigourously justified methods for choosing system parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.
 
@@
 Although the modification is straightforward, much of the notation that will be used for analysing the estimator in Section~\ref{sec:stat-prop-least} is most naturally introduced during the description of the algorithm.
-For this reason, and so that the paper is self contained, we describe the algorithm in full.
+For this reason, and so that the paper is self-contained, we describe the algorithm in full.
 For the purpose of analysing computational complexity it is assumed that the number of data symbols $\abs{D}$ is proportional to the total number of symbols $L$, so that, for example, $O(L) = O(\abs{D})$.
@@
 Figures~\ref{fig:plotphaseBPSK} and~\ref{fig:plotphaseQPSK} display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.
-To the authors knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).
-These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=\emptyset$.
+To the authors' knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).
+These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=0$.
 
@@
 \] 
-is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ with $R_i > 0$.
-Let $q_i$ be the smallest number from the interval $[-\tfrac{\pi}{M}, 0]$ such that
-\[
- \sin(q_i) > -\frac{\rho_0 \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
+is differentiable everywhere except when $\fracpart{\lambda + \Phi_i} = -\tfrac{\pi}{M}$ for any $i \in D$ with $R_i > 0$.  Let $q_i \in [-\tfrac{\pi}{M}, 0]$ satisfy
+\[
+ \sin(q_i) = -\frac{\rho_0 \sabs{\eta}^2 R_i}{4 L \hat{\rho} \sin(\pi/M)}
 \]

### 2014-02-13 response.tex

@@
 ``Figures~\zref{fig:plotphaseBPSK} and~\zref{fig:plotphaseQPSK} display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.
-To the authors knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).
+To the authors' knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).
 These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=\emptyset$.''
@@
 
-Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non-data-aided SNR estimator in order to implement the reviewers proposal.
-We confirm, however, that analysing and implementing the reviewers suggested technique is beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
+Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non-data-aided SNR estimator in order to implement the reviewer's proposal.
+We confirm, however, that analysing and implementing the reviewer's suggested technique is beyond the scope of the present paper which places primary interest in analysing the least squares estimator.
 
@@
 
-We do agree with the reviewer this this plot by itself initially appears misleading.
+We do agree with the reviewer that this plot by itself initially appears misleading.
 To alleviate this confusion in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the asymptotic squared bias plus variance.
-The following paragraph has also included in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
+The following paragraph has also been included in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.
 
@@
 	\\\textbf{Response:}
-We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our of results clear.
+We have greatly expanded the introduction of the manuscript to make the practical and theoretical importance of our results clear.
 Also see our response to comment~\ref{rev4:shouldbettermotivate} of reviewer~4.
@@
 Based on our Theorem~\zref{thm:normality} in Section~\zref{sec:stat-prop-least} the performance of the least squares estimator can be accurately predicted without the need for time consuming Monte-Carlo simulations.
-This provides easy to use and rigourously justified methods for choosing systems parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.''
+This provides easy to use and rigourously justified methods for choosing system parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.''
 

### 2014-02-13 paper.tex
added some comments at end re cleaning up the wild formula for 8psk computation
added Andre's comments regarding `hard decision' notation

@@
 \] 
+It is common to call $\hat{d}_i(\theta)$ a `hard decision'  made for the $i$th data symbol under the hypothesis that the channel phase is $\theta$.
 Note that $\hat{d}_i(\theta)$ does not depend on $\rho$.
@@
 The least squares estimator appears slightly more accurate for some values of SNR.
-
-Figures~\ref{fig:plotphaseBPSK} and~\ref{fig:plotphaseQPSK} display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.
-To the authors' knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.
 These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=0$.
@@
 
+% The function below is derived by Rice, Cowley, Moran and Rice~\cite{950345}.
+It is involved in the computation of CRBs for 8PSK (and QAM).
+The function $\Lambda : \complex \mapsto \reals$ is defined as
+% \[
+% \Lambda(x) = -\frac{A(x)^2}{\sigma^2D(x)^2} + \frac{B(x)}{\sigma^2 D(x)} - \frac{C(x)}{D(x)}
+% \]
+% where
+% \[
+% A(x) = \sum_{a\in Q}\big( s(ax)\Im(ax) + s(a^*x)\Im(a^*x) \big)
+% \]
+% \[
+% B(x) = \sum_{a\in Q}\big( c(ax)\Im(ax)^2 + c(a^*x)\Im(a^*x)^2 \big)
+% \]
+% \[
+% C(x) = \sum_{a\in Q}\big( s(ax)\Re(ax) + s(a^*x)\Re(a^*x) \big)
+% \]
+% \[
+% D(x) = \sum_{a\in Q}\big( c(ax) + c(a^*x) \big)
+% \]
+% and where
+% \[
+% s(x) = \sinh(\Re{x}/\sigma^2)
+% \]
+% and 
+% \[
+% c(x) = \cosh(\Re{x}/\sigma^2)
+% \]
+
 \end{document}

### 2014-02-13 code/crb.scala

@@ -40,8 +40,8 @@ val SNRdBs = -20.0 to 20.0 by 0.5 //snrs we will use
 
 //list of CRBS we will compute with names
 val crbfs = List[((Double)=>CramerRaoBound, String)](
-//( (v : Double) => new BPSKCRB(v), "BPSKCRB" ),
-//( (v : Double) => new QPSKCRB(v), "QPSKCRB" ),
+( (v : Double) => new BPSKCRB(v), "BPSKCRB" ),
+( (v : Double) => new QPSKCRB(v), "QPSKCRB" ),
 )
 
 val starttime = (new java.util.Date).getTime

### 2014-02-14 paper.tex
minor

@@
 The least squares estimator appears slightly more accurate for some values of SNR.
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996,950345}.
 These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=0$.

### 2014-05-12 response2.tex
added 2nd round of responses

\documentclass{article}

\usepackage[pdftex]{graphicx}
%\usepackage{thumbpdf}
%\usepackage[naturalnames]{hyperref}
\usepackage{xr}
\externaldocument{../async}

\usepackage{bm, booktabs}
\usepackage{amsmath}%
\usepackage{amsfonts}%
\usepackage{amssymb}%
%\usepackage{amsthm}
\usepackage{mathrsfs}
%\usepackage{graphicx}
%\usepackage{setspace}
\usepackage{cite}
\usepackage{units}
	%nice looking units
%\usepackage{times}
\usepackage[normalem]{ulem}
%\usepackage{algorithmic}
%\usepackage[figure, vlined, linesnumbered]{algorithm2e}
\usepackage[vlined, linesnumbered]{algorithm2e}
%\usepackage{algorithm}

%%%%% set up the bibliography style
\bibliographystyle{../IEEEbst}
%\usepackage[square,comma,numbers]{natbib}

\newcommand{\round}[1]{\lfloor #1 \rceil}
\newcommand{\floor}[1]{\lfloor #1 \rfloor}
\newcommand{\ceil}[1]{\lceil #1 \rceil}
\newcommand{\reals}{{\mathbb R}}
\newcommand{\ints}{{\mathbb Z}}
\newcommand{\integers}{{\mathbb Z}}
\newcommand{\uy}{\underline{\bm{Y}}}
\newcommand{\uey}{\underline{Y}}
\newcommand{\sign}[1]{\mathtt{sign}(#1)}
\newcommand{\Qbf}{{\mathbf Q}}
\newcommand{\Bbf}{{\mathbf B}}
\newcommand{\Ibf}{{\mathbf I}}
\newcommand{\ybf}{{\mathbf y}}
\newcommand{\xbf}{{\mathbf x}}
\newcommand{\zbf}{{\mathbf z}}
\newcommand{\ebf}{{\mathbf e}}
\newcommand{\vbf}{{\mathbf v}}
\newcommand{\wbf}{{\mathbf w}}
\newcommand{\kbf}{{\mathbf k}}
\newcommand{\sbf}{{\mathbf s}}
\newcommand{\tbf}{{\mathbf t}}
\newcommand{\pbf}{{\mathbf p}}
\newcommand{\Pibf}{{\mathbf \Pi}}
\newcommand{\onebf}{{\mathbf 1}}
\newcommand{\fbf}{{\mathbf f}}
\newcommand{\ubf}{{\mathbf u}}

\newcommand{\NP}{\operatorname{NearestPt}}
\newcommand{\NS}{\operatorname{NearestSet}}
\newcommand{\bres}{\operatorname{Bres}}
\newcommand{\vol}{\operatorname{vol}}
\newcommand{\vor}{\operatorname{Vor}}
\newcommand{\coef}{\operatorname{Coef}}
\newcommand{\Int}{\operatorname{Int}}

\newtheorem{theorem}{Theorem}

\begin{document}

\section{Comments from reviewer 1}

\begin{enumerate}

\item \textbf{Comment:} Consistently write ``non-data-aided'' throughout the whole text \\
\textbf{Response:}

\item \textbf{Comment:} Page (P) 42, right column (RC), line (L) 50: Aside from this extension of \dots \\
\textbf{Response:}

\item \textbf{Comment:} P 43, LC, L 44: \dots convincing response \dots \\
\textbf{Response:}

\item \textbf{Comment:} P 43, RC, L 2+4: \dots as multiple antenna systems \dots must withstand rapidly \dots \\
\textbf{Response:}

\item \textbf{Comment:} P 43, RC, L 21+28: \dots aided the design of \dots to other receiver components \dots \\
\textbf{Response:}

\item \textbf{Comment:} Conference papers: location missing \\
\textbf{Response:}

\item \textbf{Comment:} Some journal papers: month of publication missing \\
\textbf{Response:}

\item \textbf{Comment:} [8] \dots IEEE Commun. Letters \dots \\
\textbf{Response:}

\item \textbf{Comment:} [31] N. Noels, V. Lottici, \dots \\
\textbf{Response:}

\end{enumerate}

\section{Comments from reviewer 2}

\begin{enumerate}

\item \textbf{Comment:} In the abstract: non data-aided --> change to: non-data-aided \\
\textbf{Response:}

\item \textbf{Comment:} In the abstract: non data-aided --> change to: non-data-aided \\
\textbf{Response:}

\item \textbf{Comment:} In the abstract: non data-aided --> change to: non-data-aided \\
\textbf{Response:}

- In the introductory section:
     * "we assume that the time-offset has been previously handled" --> perhaps some motivation for this assumption and the consequences with respect to the system model would be in place
     * "such as binary phase shift keying (BPSK),...." --> mentioning BPSK and QPSK apart from general M-PSK is pure overhead in this stage, I would drop it
     * "the data symbols are not known to the receiver and must also be estimated" --> I would drop the second part of the sentence because (e.g. Viterbi and Vterbi estimator) non-data-aided estimators typically do not actually estimate the unknown data symbols
     * p. 3: "bounds that are valid asymptoticially": typo in asymptotically
     * p. 3: "all of the exiting bounds are applicable only in the fully non-data-aided case or in the completely data aided case..." --> bounds for the common scenario where there is a mixture of pilot and data symbols can be found in "N. Noels, H. Steendam, M. Moeneclaey en H. Bruneel, "Carrier Phase and Frequency Estimation for Pilot Symbol Assisted Transmission: Bounds and Algorithms, IEEE TSP, Vol. 53, No. 12, p. 4578-4587, Dec. 2005. "
     * p. 3: "the situations is" --> drop 's' at the end of situations
     * p. 3: in continuous text I prefer the use of "approached infinity" rather than "--> \infty"
     * p. 3: last line: typo in "convincing"
     * p. 4: "systems such as mulitple anntannae" --> should be "multiple antennae"
     * p. 4: "the other reciever components" --> should be "receiver components"
     * p. 5 (and further throughout the text): "pilots symbols" (with 's') and "pilot symbols" (without 's') are both used --> use only 1, I prefer "pilot symbols"
     * I think it would be a good idea to specify the main notational conventions at the end of the introductory section
- section II:
     * between (5) and (6): start new paragraph
     * before (6): drop the 'to'
     * before (7): change "for given \theta" to "for given \theta and \rho"
     * what is the use of keeping the constants 'A' and 'B' in the expressions --> can they not be dropped?
     * I would prefer to see comma's between two subindices; so, 'k,i' rather than 'ki'
     * below (11) 'g_ik' should be 'g_ki' (change order of the subscripts)
- section III:
     * I would prefer to see brackets around the functions of which the mean is computed (use of the operator \mathbb E)
     * p. 12: "the amplitude estimation error with bias removed" --> bit strange formulation
- section IV:
     * p. 14: "Equivalently \hat\theta is the maximiser of ..." --> could you please provide some further explanation (is the maximum of \Re(Z(\theta)) never negative? and why)
     * p. 15: the use of 'A' as notation for a sample space is confusing since 'A' has already been used as a constant
- section VI:
     * in general, I find the numerical results section rather short
     * although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text
     * p. 18, first sentence: "when L is sufficiently large" --> in these figures the value of L is fixed, should this be "when the SNR is sufficiently large"
     * a qualitative explanation for the fact the theory and simulations diverge at low SNR would be appreciated
     * p.18, last sentence of first paragraph: non-data aided --> should be "non-data-aided"
     * although the most important contribution of the paper is with respect to the biased amplitude estimate, only 1 figure is presented for amplitude estimation, namely for L=2048 and BPSK --> this seems not well balanced
     * a separate plot of the two terms that contribute to Fig. 4 would be welcome

\end{enumerate}

\section{Comments from reviewer 3}

\bibliography{../../../bib/bib}

\end{document}

### 2014-05-13 response2.tex
added minor

@@
 
+\newcommand{\fracpart}[1]{\left\langle #1 \right\rangle}
+\newcommand{\sfracpart}[1]{\langle #1 \rangle}
 \newcommand{\round}[1]{\lfloor #1 \rceil}
@@
 
-\item \textbf{Comment:} In the abstract: non data-aided --> change to: non-data-aided \\
+\item \textbf{Comment:} In the abstract: non data-aided $\to$ change to: non-data-aided \\
 \textbf{Response:}
 
-\item \textbf{Comment:} In the abstract: non data-aided --> change to: non-data-aided \\
+\item \textbf{Comment:} ``we assume that the time-offset has been previously handled'' $\to$ perhaps some motivation for this assumption and the consequences with respect to the system model would be in place \\
 \textbf{Response:}
 
-\item \textbf{Comment:} In the abstract: non data-aided --> change to: non-data-aided \\
+\item \textbf{Comment:} ``such as binary phase shift keying (BPSK), \dots'' $\to$ mentioning BPSK and QPSK apart from general M-PSK is pure overhead in this stage, I would drop it \\
 \textbf{Response:}
 
-- In the introductory section:
-     * "we assume that the time-offset has been previously handled" --> perhaps some motivation for this assumption and the consequences with respect to the system model would be in place
-     * "such as binary phase shift keying (BPSK),...." --> mentioning BPSK and QPSK apart from general M-PSK is pure overhead in this stage, I would drop it
-     * "the data symbols are not known to the receiver and must also be estimated" --> I would drop the second part of the sentence because (e.g.
-Viterbi and Vterbi estimator) non-data-aided estimators typically do not actually estimate the unknown data symbols
-     * p. 3: "bounds that are valid asymptoticially": typo in asymptotically
-     * p.
-3: "all of the exiting bounds are applicable only in the fully non-data-aided case or in the completely data aided case..." --> bounds for the common scenario where there is a mixture of pilot and data symbols can be found in "N.
+\item \textbf{Comment:} ``the data symbols are not known to the receiver and must also be estimated'' $\to$ I would drop the second part of the sentence because (e.g.
+Viterbi and Vterbi estimator) non-data-aided estimators typically do not actually estimate the unknown data symbols \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 3: ``bounds that are valid asymptoticially'': typo in asymptotically \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p.
+3: ``all of the exiting bounds are applicable only in the fully non-data-aided case or in the completely data aided case \dots'' --> bounds for the common scenario where there is a mixture of pilot and data symbols can be found in "N.
 Noels, H.
@@
 Moeneclaey en H.
-Bruneel, "Carrier Phase and Frequency Estimation for Pilot Symbol Assisted Transmission: Bounds and Algorithms, IEEE TSP, Vol.
+Bruneel, Carrier Phase and Frequency Estimation for Pilot Symbol Assisted Transmission: Bounds and Algorithms, IEEE TSP, Vol.
 53, No.
@@
 2005.
-"
-     * p. 3: "the situations is" --> drop 's' at the end of situations
-     * p. 3: in continuous text I prefer the use of "approached infinity" rather than "--> \infty"
-     * p. 3: last line: typo in "convincing"
-     * p. 4: "systems such as mulitple anntannae" --> should be "multiple antennae"
-     * p. 4: "the other reciever components" --> should be "receiver components"
-     * p. 5 (and further throughout the text): "pilots symbols" (with 's') and "pilot symbols" (without 's') are both used --> use only 1, I prefer "pilot symbols"
-     * I think it would be a good idea to specify the main notational conventions at the end of the introductory section
-- section II:
-     * between (5) and (6): start new paragraph
-     * before (6): drop the 'to'
-     * before (7): change "for given \theta" to "for given \theta and \rho"
-     * what is the use of keeping the constants 'A' and 'B' in the expressions --> can they not be dropped?
-     * I would prefer to see comma's between two subindices; so, 'k,i' rather than 'ki'
-     * below (11) 'g_ik' should be 'g_ki' (change order of the subscripts)
-- section III:
-     * I would prefer to see brackets around the functions of which the mean is computed (use of the operator \mathbb E)
-     * p. 12: "the amplitude estimation error with bias removed" --> bit strange formulation
-- section IV:
-     * p. 14: "Equivalently \hat\theta is the maximiser of ..." --> could you please provide some further explanation (is the maximum of \Re(Z(\theta)) never negative? and why)
-     * p. 15: the use of 'A' as notation for a sample space is confusing since 'A' has already been used as a constant
-- section VI:
-     * in general, I find the numerical results section rather short
-     * although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text
-     * p. 18, first sentence: "when L is sufficiently large" --> in these figures the value of L is fixed, should this be "when the SNR is sufficiently large"
-     * a qualitative explanation for the fact the theory and simulations diverge at low SNR would be appreciated
-     * p.18, last sentence of first paragraph: non-data aided --> should be "non-data-aided"
-     * although the most important contribution of the paper is with respect to the biased amplitude estimate, only 1 figure is presented for amplitude estimation, namely for L=2048 and BPSK --> this seems not well balanced
-     * a separate plot of the two terms that contribute to Fig. 4 would be welcome
+\\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 3: ``the situations is'' --> drop `s' at the end of situations \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 3: in continuous text I prefer the use of ``approached infinity'' rather than ``$\to \infty$'' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 3: last line: typo in ``convincing'' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 4: ``systems such as mulitple anntannae'' $\to$ should be ``multiple antennae'' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 4: ``the other reciever components'' $\to$ should be ``receiver components'' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 5 (and further throughout the text): ``pilots symbols" (with `s') and ``pilot symbols'' (without `s') are both used $\to$ use only 1, I prefer ``pilot symbols'' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} I think it would be a good idea to specify the main notational conventions at the end of the introductory section \\
+\textbf{Response:}
+
+\item \textbf{Comment:} between (5) and (6): start new paragraph \\
+\textbf{Response:}
+
+\item \textbf{Comment:} before (6): drop the 'to' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} before (7): change ``for given $\theta$'' to ``for given $\theta$ and $\rho$'' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} what is the use of keeping the constants `A' and `B' in the expressions $\to$ can they not be dropped? \\
+\textbf{Response:}
+
+\item \textbf{Comment:} I would prefer to see comma's between two subindices; so, `$k,i$' rather than`'$ki$' \\
+\textbf{Response:}
+
+\item \textbf{Comment:} below (11) '$g_{ik}$' should be '$g_{ki}$' (change order of the subscripts) \\
+\textbf{Response:}
+
+\item \textbf{Comment:} I would prefer to see brackets around the functions of which the mean is computed (use of the operator $\mathbb E$) \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 12: ``the amplitude estimation error with bias removed'' $\to$ bit strange formulation \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p.
+14: "Equivalently $\hat\theta$ is the maximiser of \dots" $\to$ could you please provide some further explanation (is the maximum of $\Re(Z(\theta)$) never negative?
+and why) \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 15: the use of `A' as notation for a sample space is confusing since `A' has already been used as a constant \\
+\textbf{Response:}
+
+\item \textbf{Comment:} in general, I find the numerical results section rather short \\
+\textbf{Response:}
+
+\item \textbf{Comment:} although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text  \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 18, first sentence: ``when $L$ is sufficiently large'' $\to$ in these figures the value of $L$ is fixed, should this be ``when the SNR is sufficiently large''  \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p. 18, first sentence: a qualitative explanation for the fact the theory and simulations diverge at low SNR would be appreciated  \\
+\textbf{Response:}
+
+\item \textbf{Comment:} p.18, last sentence of first paragraph: ``non-data aided'' $\to$ should be ``non-data-aided''  \\
+\textbf{Response:}
+
+\item \textbf{Comment:}  although the most important contribution of the paper is with respect to the biased amplitude estimate, only 1 figure is presented for amplitude estimation, namely for L=2048 and BPSK --> this seems not well balanced  \\
+\textbf{Response:}
+
+\item \textbf{Comment:}  a separate plot of the two terms that contribute to Fig. 4 would be welcome \\
+\textbf{Response:}
 
@@
 
+\begin{enumerate}
+
+\item \textbf{Comment:}  Eq. (14) would be easier to follow with the intermediate step inserted $\dots = (a_0s_i + w_i)/(a_0s_i) = \dots$ \\
+\textbf{Response:}
+
+\item \textbf{Comment:}  There are many formulas (starting from page 11) with the notation $\fracpart{x}$ meaning ``$x$ taken modulo $2\pi/M$''.
+These angle brackets don't strike the eye and can be easily confused with ordinary parentheses.
+To make them much more visible, the authors could use the same approach as for ``$x$ taken modulo $2\pi$'' - by using corresponding subscript, i.e.
+use $\fracpart{x}_{2\pi/M}$ instead of $\fracpart{x}$.
+\\
+\textbf{Response:}
+
+ \end{enumerate}
+
+\section{Comments from reviewer 4}
+
+\begin{enumerate}
+
+\item \textbf{Comment:}  The variable  ``a'' in the equation 2 is not defined in the text \\
+\textbf{Response:}
+
+\item \textbf{Comment:}  There are typografical errors in the words  ``multiple antenna'' in line 2, page 2, section 1 \\
+\textbf{Response:}
+
+\item \textbf{Comment:}  And also in the word ``receiver" in line 28 of the same page.
+ \\
+\textbf{Response:}
+
+\end{enumerate}
+
 \bibliography{../../../bib/bib}

### 2014-05-26 paper.tex
added minor

@@
 However, the amplitude estimator is asymptotically biased and converges to a positive real number that is a function of the true carrier amplitude, the noise distribution and the size of the constellation.
-This appears to be the first time that the statistical properties of a non data-aided estimator for carrier amplitude have been analysed theoretically.
+This appears to be the first time that the statistical properties of a non-data-aided estimator for carrier amplitude have been analysed theoretically.
 %Our theoretical results can also be applied to the case where no pilot symbols exist, i.e., noncoherent detection.  
@@
 In this paper we assume that the time offset has been previously handled and we focus on estimating the phase shift and attenuation.
-We consider signalling constellations that have symbols evenly distributed on the complex unit circle such as binary phase shift keying (BPSK), quaternary phase shift keying (QPSK), and $M$-ary phase shift keying ($M$-PSK).
-In this case, the transmitted symbols take the form,
+Methods for estimating the time offset are given in~\cite{Massey1972optimumframe,Oerder_synch_square_circstat_2008,McKilliam_time_offset_pilots_data_2013}.
+The transmitted symbols are assumed to come from an $M$-ary phase shift keying ($M$-PSK) constellation.
+That is, the transmitted symbols take the form,
 \[
@@
 Our aim is to estimate $a_0$ from the noisy symbols $\{ y_i, i \in P \cup D \}$.
-Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the receiver and must also be estimated.
+Complicating matters is that the data symbols $\{d_i, i \in D\}$ are not known to the receiver.
 Estimation problems of this type have undergone extensive prior study~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990,580211,974597,Mengali_andre_synchro_book}.
@@
 
-% The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} (also know as \emph{blind}, \emph{non data-aided} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
+% The existing literature~\cite{Mackenthun1994,Cowley_ref_sym_carr_1998,ViterbiViterbi_phase_est_1983,Sweldens2001,Wilson1989,Makrakis1990,Liu1991} mostly considers what is called \emph{noncoherent detection} (also know as \emph{blind}, \emph{non-data-aided} or \emph{non-decision directed detection}) where no pilot symbols exist ($P = \emptyset$).
 In the noncoherent setting \emph{differential encoding} is often used, and for this reason the estimation problem has been called \emph{multiple symbol differential detection}.
@@
 
-Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:LSdefn}.
+Aside from this extension of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:LSdefn}.
 In the literature, analytic study of non-data-aided estimators of carrier phase has primarily involved the derivation of Cram\'{e}r-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
-The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
-To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
+The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptotically as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
+%To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
 Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.
 
-For the case of carrier amplitude the situations is even more complicated.
+For the case of carrier amplitude the situation is more complicated.
 A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non-data-aided estimators for carrier amplitude exhibit bias.
 The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
-Neverthess, Cram\'{e}r-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
+Neverthess, Cram\'{e}r-Rao bounds have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
 Unfortunately, these bounds may not be meaningful since known non-data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
@@
 One response to this criticism is that the least squares estimator should be a popular practical choice because it can be efficiently computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian.
-Perhaps an even more convinving response is that, while the specific results given here apply only to the least squares estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
+Perhaps an even more convincing response is that, while the specific results given here apply only to the least squares estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.
 For example, in practical systems it is sometimes desirable to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.
@@
 The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010}, would potentially allow this.
-More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,inthft:204506}.
+More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple antenna systems, systems employing orthogonal frequency division multiplexing, and systems that must withstand rapidly time varying channels~\cite{4138046,inthft:204506}.
 We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cram\'{e}r-Rao lower bounds.
@@
 This estimator has been implemented on a prototype that was successfully trialled from April to July of 2013 with signals from multiple low power ground-based transmitters being successfully received and decoded by both aircraft and satellites in orbit~\cite{ASRPpromovideo}.
-Our theoretical analysis aided design of this system.
+Our theoretical analysis aided the design of this system.
 Based on our Theorem~\ref{thm:normality} in Section~\ref{sec:stat-prop-least} the performance of the least squares estimator can be accurately predicted without the need for time consuming Monte-Carlo simulations.
-This provides easy to use and rigourously justified methods for choosing system parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.
+This provides easy to use and rigourously justified methods for choosing system parameters, such as the number and placement of pilot symbols, and also provides useful information to other receiver components, such as channel tracking systems.
 

### 2014-05-26 response2.tex
added minor

@@
 \usepackage{xr}
-\externaldocument{../async}
+\externaldocument{paper}
 
@@
 \item \textbf{Comment:} Consistently write ``non-data-aided'' throughout the whole text \\
-\textbf{Response:}
+\textbf{Response:} Done.
 
 \item \textbf{Comment:} Page (P) 42, right column (RC), line (L) 50: Aside from this extension of \dots \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
 \item \textbf{Comment:} P 43, LC, L 44: \dots convincing response \dots \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
-\item \textbf{Comment:} P 43, RC, L 2+4: \dots as multiple antenna systems \dots must withstand rapidly \dots \\
-\textbf{Response:}
+\item \label{eq:rev1:multantenna} \textbf{Comment:} P 43, RC, L 2+4: \dots as multiple antenna systems \dots must withstand rapidly \dots \\
+\textbf{Response:} Fixed.
 
 \item \textbf{Comment:} P 43, RC, L 21+28: \dots aided the design of \dots to other receiver components \dots \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
@@
 \item \textbf{Comment:} In the abstract: non data-aided $\to$ change to: non-data-aided \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
 \item \textbf{Comment:} ``we assume that the time-offset has been previously handled'' $\to$ perhaps some motivation for this assumption and the consequences with respect to the system model would be in place \\
-\textbf{Response:}
+\textbf{Response:}  There exist numerous estimators for the time offset in the literature.  We have added the sentence
+
+``Methods for estimating the time offset are given in~\cite{Massey1972optimumframe,Oerder_synch_square_circstat_2008,McKilliam_time_offset_pilots_data_2013}.''
+
+to inform the reader of some of these methods.
+The interested reader will find further discussion about time offset estimation within these references.
+The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
 
 \item \textbf{Comment:} ``such as binary phase shift keying (BPSK), \dots'' $\to$ mentioning BPSK and QPSK apart from general M-PSK is pure overhead in this stage, I would drop it \\
-\textbf{Response:}
+\textbf{Response:} Done.
 
@@
 Viterbi and Vterbi estimator) non-data-aided estimators typically do not actually estimate the unknown data symbols \\
-\textbf{Response:}
+\textbf{Response:} Done.
 
 \item \textbf{Comment:} p. 3: ``bounds that are valid asymptoticially'': typo in asymptotically \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
@@
 \\
-\textbf{Response:}
+\textbf{Response:} BLERG. I've removed this paragraph from the introduction but will have to update the code and simulations.
 
 \item \textbf{Comment:} p. 3: ``the situations is'' --> drop `s' at the end of situations \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
@@
 \item \textbf{Comment:} p. 3: last line: typo in ``convincing'' \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
 \item \textbf{Comment:} p. 4: ``systems such as mulitple anntannae'' $\to$ should be ``multiple antennae'' \\
-\textbf{Response:}
+\textbf{Response:} Reviewer 1 comment~\eqref{eq:rev1:multantenna} wanted ``antenna'' and I have used this spelling.
 
 \item \textbf{Comment:} p. 4: ``the other reciever components'' $\to$ should be ``receiver components'' \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
@@
 
-\bibliography{../../../bib/bib}
+\input{paper.bbl}
 

### 2014-05-26 paper.tex
added further to comments

@@
 In this paper we present an alternative approach to the study of estimators for carrier phase and amplitude.
-Rather than compute bounds, we directly study the least squares estimator asymptotically as the number of transmitted symbols $L\to \infty$.
+Rather than compute bounds, we directly study the least squares estimator asymptotically as the number of transmitted symbols $L$ approaches $\infty$.
 We show, under comparatively mild conditions on the distribution of the noise, that the least squares estimator of carrier phase is strongly consistent (asymptotically unbiased) and asymptotically normally distributed.
@@
 For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.
-In this case the number of pilots symbols must be sufficiently large for channel estimates to be sufficiently accurate.
-If more accuracy is required, then more pilots symbols must be used.
+In this case the number of pilot symbols must be sufficiently large for channel estimates to be sufficiently accurate.
+If more accuracy is required, then more pilot symbols must be used.
 Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols.
-Instead, the pilots symbols and a subset of the data symbols can be jointly used for channel estimation.
+Instead, the pilot symbols and a subset of the data symbols can be jointly used for channel estimation.
 %For computational complexity reasons it may not be desirable to use all of the existing data symbols for channel estimation.
@@
 Observe that given candidate values for the data symbols, we can compute the corresponding $LS(\{d_i, i \in D\})$ in $O(L)$ arithmetic operations.
-It turns out that there are at most $M\abs{D}$ candidate values of the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
+It turns out that there are at most $M\abs{D}$ candidate values for the least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}.
 %When the number of data symbols is not small, this set is substantially smaller than the entire set of possible transmitted symbols, which is of size $M^{\abs{D}}$.
 
-To see this, let $a = \rho e^{j\theta}$ where $\rho > 0$ is real.  Now $LS$ from~\eqref{eq:LS} is to
+To see this, let $a = \rho e^{j\theta}$ where $\rho > 0$ is real.  Now $LS$ from~\eqref{eq:LS} is
 \begin{align}
@@
 This should not cause confusion as $LS(a, \{d_i, i \in D\})$, $LS(\rho, \theta, \{d_i, i \in D\})$, and $LS(\{d_i, i \in D\})$ are easily told apart by their arguments.
-For given $\theta$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
+For given $\theta$ and $\rho$, the least squares estimator of the $i$th data symbol $d_i$ is given by minimising $\abs{ y_i - \rho e^{j\theta} d_i }^2$, that is,
 \begin{equation}\label{eq:hatdfinxtheta}
@@
 \[
-Y_k = Y( f_k ) = \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{ki}^* = B + \sum_{i \in D}g_{ki},
-\]
-where $B = \sum_{i \in P} y_i p_i^*$ is independent of the data symbols, and $f_{ki}$ denotes the $i$th symbol in $f_k$, and for convenience, we put $g_{ki}  = y_i f_{ki}^*$.
-Letting $g_{k}$ be the sequence $\{g_{ik}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
+Y_k = Y( f_k ) = \sum_{i \in P} y_i p_i^*  + \sum_{i \in D} y_i f_{k,i}^* = B + \sum_{i \in D}g_{k,i},
+\]
+where $B = \sum_{i \in P} y_i p_i^*$ is independent of the data symbols, and $f_{k,i}$ denotes the $i$th symbol in $f_k$ and, for convenience, we put $g_{k,i}  = y_i f_{k,i}^*$.
+Letting $g_{k}$ be the sequence $\{g_{k,i}, i \in D\}$ we have, from~\eqref{eq:fkrec}, that $g_k$ satisfies the recursive equation
 \[
@@
 \]
-where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k \sigma(k)}e^{j2\pi/M}$.  Now,
-\[
-Y_0 = B + \sum_{i \in D} g_{0i}
+where $g_{k} e_{\sigma(k)}^*$ indicates the sequence $g_k$ with the $\sigma(k)$th element replaced by $g_{k,\sigma(k)}e^{j2\pi/M}$.  Now,
+\[
+Y_0 = B + \sum_{i \in D} g_{0,i}
 \] 
@@
 \begin{align*}
-Y_1 &= B + \sum_{i \in D} g_{1i} \\
-&= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
-&= Y_0 + \eta g_{0\sigma(0)},
+Y_1 &= B + \sum_{i \in D} g_{1,i} \\
+&= B +  (e^{j2\pi/M} - 1)g_{0,\sigma(0)} + \sum_{i \in D} g_{0,i} \\
+&= Y_0 + \eta g_{0,\sigma(0)},
 \end{align*}
@@
 \[
-Y_{k+1} = Y_k + \eta g_{k\sigma(k)}.
+Y_{k+1} = Y_k + \eta g_{k,\sigma(k)}.
 \]
@@
 Let $\hat{k} = \arg\min LS(f_k)$.
-The least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata},
+The least squares estimator of $a_0$ is then computed according to~\eqref{eq:hata}, that is,
 \[
@@
 \begin{equation}\label{eq:RiandPhii}
-R_ie^{j\Phi_i} = 1 + \frac{w_i}{a_0 s_i} ,
+R_ie^{j\Phi_i} =  \frac{a_0s_i + w_i}{a_0s_i}  = 1 + \frac{w_i}{a_0 s_i} ,
 \end{equation}
@@
 However, it will be notationally convenient in Sections~\ref{sec:proof-almost-sure} and~\ref{sec:proof-asympt-norm} to use $\langle\theta_0 - \hat{\theta}\rangle_\pi$.
-The theorems require minor modification when the number of pilots symbols $\abs{P} = 0$.  Theorem statements for this special case appear in \cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+The theorems require minor modification when the number of pilot symbols $\abs{P} = 0$.  Theorem statements for this special case appear in \cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
 
@@
 We will further investigate the bias in Section~\ref{sec:simulations}.
-Theorem~\ref{thm:normality} shows that $\sqrt{L}(\hat{\rho} - \rho_0 G(0))$ (the amplitude estimator error with bias removed) and $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ are asymptotically normally distributed and statistically independent as $L\rightarrow\infty$.
+Theorem~\ref{thm:normality} shows that $\sqrt{L}(\hat{\rho} - \rho_0 G(0))$ and $\sqrt{L}\langle\theta_0 - \hat{\theta}\rangle_\pi$ are asymptotically normally distributed and statistically independent as $L\rightarrow\infty$.
 
@@
 As is customary, let $\Omega$ be the sample space on which the random variables $\{w_i\}$ are defined.
-Let $A$ be the subset of the sample space $\Omega$ on which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
-Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{A\} = 1$.
-Let $A'$ be the subset of the sample space on which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
+Let $\Sigma$ be the subset of the sample space $\Omega$ on which $G(\hat{\lambda}_L) \rightarrow G(0)$ as $L\rightarrow\infty$.
+Lemma~\ref{lem:convtoexpGlamL} shows that $\prob\{\Sigma\} = 1$.
+Let $\Sigma'$ be the subset of the sample space on which $\hat{\lambda}_L \rightarrow 0$ as $L\rightarrow \infty$.
 Because $G(x)$ is uniquely maximised at $x=0$, it follows that $G(\hat{\lambda}_L) \rightarrow G(0)$ only if $\hat{\lambda}_L \rightarrow 0$ as $L \rightarrow\infty$.
-So $A \subseteq A'$ and therefore $\prob\{A'\} \geq \prob\{A\} = 1$.
+So $\Sigma \subseteq \Sigma'$ and therefore $\prob\{\Sigma'\} \geq \prob\{\Sigma\} = 1$.
 Part 1 of Theorem~\ref{thm:consistency} follows.
@@
 
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilots symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} show the sample mean square error (MSE) of the phase estimator when $M=2,4,8$ with $L=4096$ and for varying proportions of pilot symbols $\abs{P} = 0, \tfrac{L}{32}, \tfrac{L}{8}, \tfrac{L}{2}, L$.
 When $\abs{P} \neq 0$ the mean square error is computed as $\tfrac{1}{T}\sum_{i=1}^T\sfracpart{\hat{\theta}_i - \theta_0}_\pi^2$.
@@
 Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996,950345}.
-These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=0$.
+These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilot symbols, i.e., $\abs{P}=0$.
 
 % Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
-% shows the variance of the amplitude estimator when $M=2$ and with $L=32, 256, 2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+% shows the variance of the amplitude estimator when $M=2$ and with $L=32, 256, 2048$ and when the number of pilot symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
 The solid lines are the variance predicted by Theorem~\ref{thm:normality}.
@@
 Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
-shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilot symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
 The solid lines show the mean square error predicted by the squared bias $\rho_0^2\big(1 - G(0)\big)^2$ plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\ref{thm:normality}.
@@
 The figure depicts an interesting phenomenon.
-When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.
-When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.
+When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilot symbols is the same as when $L=\abs{P} = 256$.
+When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilot symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilot symbols.
 A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.
@@
 %Thus, when the SNR is small it is better to ignore the data symbols and use only the pilot symbols.   
- %This behaviour suggests modifying the objective function to give the pilots symbols more importance when the SNR is low.
+ %This behaviour suggests modifying the objective function to give the pilot symbols more importance when the SNR is low.
 For example, rather than minimise~\eqref{eq:LSdefn} we could instead minimise a weighted version of it,

### 2014-05-26 response2.tex
added further to comments

@@
 \newcommand{\ceil}[1]{\lceil #1 \rceil}
+\newcommand{\abs}[1]{{\left\vert #1 \right\vert}}
+\newcommand{\sabs}[1]{{\vert #1 \vert}}
 \newcommand{\reals}{{\mathbb R}}
@@
 \item \textbf{Comment:} p. 3: in continuous text I prefer the use of ``approached infinity'' rather than ``$\to \infty$'' \\
-\textbf{Response:}
+\textbf{Response:} This change has been made wherever it occurs in the introduction, but not throughout the whole text. This is a personal preference of the authors.
 
@@
 \item \textbf{Comment:} p. 5 (and further throughout the text): ``pilots symbols" (with `s') and ``pilot symbols'' (without `s') are both used $\to$ use only 1, I prefer ``pilot symbols'' \\
-\textbf{Response:}
+\textbf{Response:} All occurences of pilots symbols have been changed to pilot symbols.
 
-\item \textbf{Comment:} I think it would be a good idea to specify the main notational conventions at the end of the introductory section \\
-\textbf{Response:}
+\item \label{rev2:commentnotationsection} \textbf{Comment:} I think it would be a good idea to specify the main notational conventions at the end of the introductory section \\
+\textbf{Response:} It is the authors preference to introduce notation as it is used in the text rather than in a dedicated notation section.
+%BLERG.
+This need to be done.
+It's connected with comment~\ref{rev2:commentE}.
 
 \item \textbf{Comment:} between (5) and (6): start new paragraph \\
-\textbf{Response:}
+\textbf{Response:} The paragraph ends at ``\dots least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}'' and the new paragraph begins with ``To see this, let $a = \rho e^{j\theta}$ \dots''
 
 \item \textbf{Comment:} before (6): drop the 'to' \\
-\textbf{Response:}
+\textbf{Response:} Done.
 
 \item \textbf{Comment:} before (7): change ``for given $\theta$'' to ``for given $\theta$ and $\rho$'' \\
-\textbf{Response:}
+\textbf{Response:} Done.
 
 \item \textbf{Comment:} what is the use of keeping the constants `A' and `B' in the expressions $\to$ can they not be dropped? \\
-\textbf{Response:}
+\textbf{Response:} The constants are retained so that statements of equality, such as
+\[
+LS(\{d_i, i \in D\}) = A - \frac{1}{L}\abs{Y}^2
+\]
+are valid.
+For the purpose of minimisation the constants have no effect.
+However, in this case, we feel that introducing notation to describe ``equivalent for the purpose of minimisation'' is more cumbersome than simply retaining constants in the few cases that they occur.
 
 \item \textbf{Comment:} I would prefer to see comma's between two subindices; so, `$k,i$' rather than`'$ki$' \\
-\textbf{Response:}
+\textbf{Response:} Done.
 
 \item \textbf{Comment:} below (11) '$g_{ik}$' should be '$g_{ki}$' (change order of the subscripts) \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
-\item \textbf{Comment:} I would prefer to see brackets around the functions of which the mean is computed (use of the operator $\mathbb E$) \\
-\textbf{Response:}
+\item \label{rev2:commentE} \textbf{Comment:} I would prefer to see brackets around the functions of which the mean is computed (use of the operator $\mathbb E$) \\
+\textbf{Response:} BLERG.  Connected with comment~\ref{rev2:commentnotationsection}.
 
 \item \textbf{Comment:} p. 12: ``the amplitude estimation error with bias removed'' $\to$ bit strange formulation \\
-\textbf{Response:}
+\textbf{Response:} This phrase has been removed.
 
@@
 \item \textbf{Comment:} p. 15: the use of `A' as notation for a sample space is confusing since `A' has already been used as a constant \\
-\textbf{Response:}
+\textbf{Response:} $A$ has been replaced by $\Sigma$ and $A^\prime$ by $\Sigma^\prime$.
 
@@
 \item \textbf{Comment:}  Eq. (14) would be easier to follow with the intermediate step inserted $\dots = (a_0s_i + w_i)/(a_0s_i) = \dots$ \\
-\textbf{Response:}
+\textbf{Response:} Done.
 
@@
 \item \textbf{Comment:}  The variable  ``a'' in the equation 2 is not defined in the text \\
-\textbf{Response:}
+\textbf{Response:} In~\eqref{eq:LSdefn} the variable $a$ is an argument of the function $LS$, that is,
+\[
+\begin{split}
+LS(a, &\{d_i, i \in D\}) = \sum_{i \in P \cup D} \abs{ y_i - a s_i }^2  \\
+&= \sum_{i \in P} \abs{ y_i - a p_i }^2 + \sum_{i \in D} \abs{ y_i - a d_i }^2.
+\end{split}
+\]
 
 \item \textbf{Comment:}  There are typografical errors in the words  ``multiple antenna'' in line 2, page 2, section 1 \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
@@
  \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 

### 2014-05-27 paper.tex
added more the bib and responses

@@
 The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptotically as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
-%To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
+%To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data-aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
 Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.
@@
 
-The proof of Theorem~\ref{thm:consistency} hinges on Lemmas \ref{lem:convtoexpGlamL} to~\ref{lem:GLtoG0}, which can be found in the appendix.
+The proof of Theorem~\ref{thm:consistency} relies on Lemmas \ref{lem:convtoexpGlamL} to~\ref{lem:GLtoG0}, which can be found in the appendix.
 Substituting $\{ \hat{d}_i(\theta), i \in D \}$ from~\eqref{eq:hatdfinxtheta} into~\eqref{eq:LSallparams} we obtain $LS$ minimised with respect to the data symbols,
@@
 
-The proof of Theorem~\ref{thm:normality} hinges on Lemmas \ref{lem:diffatlambdaL} to~\ref{lem:XL}, which can be found in the appendix.
+The proof of Theorem~\ref{thm:normality} relies on Lemmas \ref{lem:diffatlambdaL} to~\ref{lem:XL}, which can be found in the appendix.
 We first prove the asymptotic normality of $\sqrt{L} \hat{\lambda}_L$.
@@
 The prediction is made by dividing the asymptotic covariance matrix by $L$.
-The theorem accurately predicts the behaviour of the phase estimator when $L$ is sufficiently large.
+The theorem accurately predicts the behaviour of the phase estimator when SNR is sufficiently large.
 As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
@@
 The least squares estimator appears slightly more accurate for some values of SNR.
-Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996,950345}.
-These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilot symbols, i.e., $\abs{P}=0$.
+Figures~\ref{fig:plotphaseBPSK},~\ref{fig:plotphaseQPSK}~and~\ref{fig:plotphase8PSK} also display the non-data-aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996,950345}.
+These CRBs are derived for the non-data-aided setting and are only valid for estimators that do not make use of pilot symbols, i.e., $\abs{P}=0$.
 
@@
 		\includegraphics[width=\linewidth]{code/data/plotM2-1.mps}
-		\caption{Amplitude estimator versus SNR for BPSK with $L=2048$}
+		\caption{Amplitude estimator versus SNR for BPSK with $L=4096$.}
 		\label{fig:plotampBPSK}
@@
 
-% \begin{figure}[p]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
-% 		\caption{Unbiased amplitude error versus SNR for QPSK.}
-% 		\label{fig:plotampQPSK}
-% \end{figure}
-
-% \begin{figure}[p]
-% 	\centering
-% 		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
-% 		\caption{Unbiased amplitude error versus SNR for $8$-PSK.}
-% 		\label{fig:plotamp8PSK}
-% \end{figure}
-
-%\begin{figure}[tp]
-%	\centering
-%		\includegraphics[width=\linewidth]{code/data/plotM2-3.mps}
-%		\caption{Phase error versus SNR for BPSK.}
-%		\label{fig:plotphase}
-%\end{figure}
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM4-1.mps}
+		\caption{Amplitude error versus SNR for QPSK with $L=4096$.}
+		\label{fig:plotampQPSK}
+\end{figure}
+
+\begin{figure}[p]
+	\centering
+		\includegraphics[width=\linewidth]{code/data/plotM8-1.mps}
+		\caption{Amplitude error versus SNR for $8$-PSK with $L=4096$.}
+		\label{fig:plotamp8PSK}
+\end{figure}
 

### 2014-05-27 response2.tex
added more the bib and responses

@@
 \item \textbf{Comment:} Conference papers: location missing \\
-\textbf{Response:}
+\textbf{Response:} The location of conferences has been added.
 
 \item \textbf{Comment:} Some journal papers: month of publication missing \\
-\textbf{Response:}
+\textbf{Response:} The month has been added whenever it was available in the original publication.
 
 \item \textbf{Comment:} [8] \dots IEEE Commun. Letters \dots \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
 \item \textbf{Comment:} [31] N. Noels, V. Lottici, \dots \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
@@
 
-\item \textbf{Comment:} p. 3: ``the situations is'' --> drop `s' at the end of situations \\
+\item \textbf{Comment:} p. 3: ``the situations is'' $\to$ drop `s' at the end of situations \\
 \textbf{Response:} Fixed.
@@
 \item \textbf{Comment:} in general, I find the numerical results section rather short \\
-\textbf{Response:}
+\textbf{Response:} We have greatly expanded the simulation section. BLERG: this includes \dots ref to below.
 
@@
 \item \textbf{Comment:} p. 18, first sentence: ``when $L$ is sufficiently large'' $\to$ in these figures the value of $L$ is fixed, should this be ``when the SNR is sufficiently large''  \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
 \item \textbf{Comment:} p. 18, first sentence: a qualitative explanation for the fact the theory and simulations diverge at low SNR would be appreciated  \\
-\textbf{Response:}
+\textbf{Response:} BLERG: Added a reference to Tazras crb paper.
 
 \item \textbf{Comment:} p.18, last sentence of first paragraph: ``non-data aided'' $\to$ should be ``non-data-aided''  \\
-\textbf{Response:}
+\textbf{Response:} Fixed.
 
-\item \textbf{Comment:}  although the most important contribution of the paper is with respect to the biased amplitude estimate, only 1 figure is presented for amplitude estimation, namely for L=2048 and BPSK --> this seems not well balanced  \\
-\textbf{Response:}
+\item \textbf{Comment:}  although the most important contribution of the paper is with respect to the biased amplitude estimate, only 1 figure is presented for amplitude estimation, namely for $L=2048$ and BPSK $\to$ this seems not well balanced  \\
+\textbf{Response:} Figures have been added showing the amplitude estimation error for BSPK, QPSK, and 8-PSK.
 

### 2014-05-27 paper.tex
added more to responses. Getting there.

@@
 In the literature, analytic study of non-data-aided estimators of carrier phase has primarily involved the derivation of Cram\'{e}r-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.
-The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptotically as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.
+The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptotically as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Noels_crbs_with_pilots_2005,Delmas_exact_crb_psk_2008}.
 %To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data-aided case, that is, when either no pilot symbols exist, or when no data symbols exist.
@@
 A more involved extension would be the inclusion of carrier frequency as a parameter to be estimated.
-The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010}, would potentially allow this.
+The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010,McKilliam_pps_unwrapping_tsp_2014}, would potentially allow this.
 More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple antenna systems, systems employing orthogonal frequency division multiplexing, and systems that must withstand rapidly time varying channels~\cite{4138046,inthft:204506}.
@@
 \end{equation}
-where $\Im(\cdot)$ denotes the imaginary part.
+where $\Im(\cdot)$ denotes the imaginary part, and $\expect$ denotes the expected value.
+Whenever there is no chance of ambiguity we write simply $\expect X$ rather than $\expect(X)$ to denote the expected value of the random variable $X$.
+For two random variables $X$ and $Y$ the notation $\expect X Y$ will always mean the expected value of the product $XY$, that is, $\expect(XY)$ and not $\expect(X)Y$.
 

### 2014-05-27 response2.tex
added more to responses. Getting there.

@@
 \newcommand{\ubf}{{\mathbf u}}
+\newcommand{\expect}{{\mathbb E}}
 
@@
 \\
-\textbf{Response:} BLERG. I've removed this paragraph from the introduction but will have to update the code and simulations.
+\textbf{Response:} We have added this reference and have removed the sentence begining ``To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case \dots''.
+Note that the model considered in the paper by Noels~et.~al.
+treats the amplitude as fixed and known and includes a frequency offset term.
+Thus, the bounds are not immediately applicable in our setting and we are not able to meaningfully compare these bounds with our simulation results.
+For this reason with have kept with the bound described in~\cite{950345} as in the previous version of the manuscript.
 
@@
 \item \textbf{Comment:} p. 5 (and further throughout the text): ``pilots symbols" (with `s') and ``pilot symbols'' (without `s') are both used $\to$ use only 1, I prefer ``pilot symbols'' \\
-\textbf{Response:} All occurences of pilots symbols have been changed to pilot symbols.
+\textbf{Response:} All occurences of ``pilots symbols'' have been changed to ``pilot symbols''.
 
 \item \label{rev2:commentnotationsection} \textbf{Comment:} I think it would be a good idea to specify the main notational conventions at the end of the introductory section \\
-\textbf{Response:} It is the authors preference to introduce notation as it is used in the text rather than in a dedicated notation section.
+\textbf{Response:} We request not to make this change.
+It is the authors' preference to introduce notation as it is used in the text rather than in a dedicated notation section.
 %BLERG.
@@
 \item \label{rev2:commentE} \textbf{Comment:} I would prefer to see brackets around the functions of which the mean is computed (use of the operator $\mathbb E$) \\
-\textbf{Response:} BLERG.  Connected with comment~\ref{rev2:commentnotationsection}.
+\textbf{Response:} We prefer not to use brackets whenever there is no chance of ambiguity.  To clarify this for the reader the following sentences have been added to Section~\ref{sec:stat-prop-least}:
+
+``Whenever there is no chance of ambiguity we write simply $\expect X$ rather than $\expect(X)$ to denote the expected value of the random variable $X$.
+For two random variables $X$ and $Y$ the notation $\expect X Y$ will always mean the expected value of the product $XY$, that is, $\expect(XY)$ and not $\expect(X)Y$.''
+
+The bracketless notation is heavily employed in the statistical literature on empirical processes~\cite{Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989,van2009empirical}.
+The notation would become heavily cluttered with brackets otherwise.
+Our proofs in the appendix rely upon this literature.
 
@@
 \item \textbf{Comment:}  although the most important contribution of the paper is with respect to the biased amplitude estimate, only 1 figure is presented for amplitude estimation, namely for $L=2048$ and BPSK $\to$ this seems not well balanced  \\
-\textbf{Response:} Figures have been added showing the amplitude estimation error for BSPK, QPSK, and 8-PSK.
+\textbf{Response:} Figures have been added showing the amplitude estimation error for BSPK, QPSK, and 8-PSK. BLERG: Add to text!
 
@@
 \\
-\textbf{Response:}
+\textbf{Response:} We request not to make this change.
+There are two reasons.
+The first is that this notation leads to situations where there are subscripts withing superscripts.
+For example, in the proof of Theorem 1 we would have expressions of the form
+\[
+\rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}_{2\pi/M}}
+\]
+and in the appendix we would obtain expressions such as
+\[
+2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}_{2\pi/M}}\right)
+\]
+and
+\[
+q_2(\lambda) + j k_2(\lambda) = \expect\big[ R_1 e^{j\fracpart{\lambda + \Phi_1}_{2\pi/M}} - R_1 e^{j\fracpart{\Phi_1}_{2\pi/M}}  \big].
+\]
+We feel that these are more difficult to parse that the existing simpler
+\[
+\rho_0 R_i e^{j\sfracpart{\lambda + \Phi_i}},
+\]
+\[
+2 \hat{\rho} \rho_0 R_k \Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right),
+\]
+and
+\[
+q_2(\lambda) + j k_2(\lambda) = \expect\big[ R_1 e^{j\fracpart{\lambda + \Phi_1}} - R_1 e^{j\fracpart{\Phi_1}}  \big].
+\]
+
+The second reason is that some of the authors have made use of the notation $\fracpart{\cdot}$ is previous work~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010,McKilliam_pps_unwrapping_tsp_2014} and there is a desire for continuity in the notation between these papers.
 

### 2014-05-27 paper.tex
added new data

@@
 $\abs{P} = L$.
+As expected, the figures indicate the estimator to become more accurate as the number of pilot symbols is increased.
+An interesting property distriguishing the case when no pilot symbols exist is the slope of the MSE curve at low SNR.
+When a finite proportion of pilot symbols exist the slope at low SNR is similar to that when all symbols are pilots.
+The slope is steeper in the case that no pilot symbols exist.
 
@@
 Viterbi and Viterbi propose several viable alternatives, from which we have chosen $F(x) = 1$.
-The Viterbi and Viterbi estimator is only applicable in the non-data-aided setting, i.e.
-when $\abs{P} = 0$.
+The Viterbi and Viterbi estimator is only applicable in the non-data-aided setting, i.e., when $\abs{P} = 0$.
 The sample MSE of the least squares estimator (when $\abs{P} = 0$) and the Viterbi and Viterbi estimator is similar.
@@
 
-Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
-shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilot symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
+Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} show the mean square error of the amplitude estimator when $M=2,4,8$ and $L=4096$ and when the number of pilot symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.
 The solid lines show the mean square error predicted by the squared bias $\rho_0^2\big(1 - G(0)\big)^2$ plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\ref{thm:normality}.
@@
 \]
-When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
+When the SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.
 However, $G(0)$ grows without bound as the variance of the noise increases so the bias is significant when SNR is small.
-Figure~\ref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
+Figure~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
 

### 2014-05-27 response2.tex
added new data

@@
 \textbf{Response:} We have added this reference and have removed the sentence begining ``To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case \dots''.
-Note that the model considered in the paper by Noels~et.~al.
-treats the amplitude as fixed and known and includes a frequency offset term.
+Note that the model considered in the paper by Noels~et.~al.~\cite{Noels_crbs_with_pilots_2005} treats the amplitude as fixed and known and includes a frequency offset term.
 Thus, the bounds are not immediately applicable in our setting and we are not able to meaningfully compare these bounds with our simulation results.
-For this reason with have kept with the bound described in~\cite{950345} as in the previous version of the manuscript.
+%For this reason with have kept with the bound described in~\cite{950345} as in the previous version of the manuscript.
 
@@
 \item \textbf{Comment:} in general, I find the numerical results section rather short \\
-\textbf{Response:} We have greatly expanded the simulation section. BLERG: this includes \dots ref to below.
+\textbf{Response:} We have greatly expanded the simulation section as detailed in the responses that follow.
 
 \item \textbf{Comment:} although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text  \\
-\textbf{Response:}
+\textbf{Response:}  We have included the follows sentences to Section~\ref{sec:simulations}:
+
+``As expected, the figures indicate the estimator to become more accurate as the number of pilot symbols is increased.
+An interesting property distriguishing the case when no pilot symbols exist is the slope of the MSE curve at low SNR.
+When a finite proportion of pilot symbols exist the slope at low SNR is similar to that when all symbols are pilots.
+The slope is steeper in the case that no pilot symbols exist.''
 
@@
 \item \textbf{Comment:}  although the most important contribution of the paper is with respect to the biased amplitude estimate, only 1 figure is presented for amplitude estimation, namely for $L=2048$ and BPSK $\to$ this seems not well balanced  \\
-\textbf{Response:} Figures have been added showing the amplitude estimation error for BSPK, QPSK, and 8-PSK. BLERG: Add to text!
+\textbf{Response:} Figures have been added showing the amplitude estimation error for BSPK, QPSK, and 8-PSK.
 
 \item \textbf{Comment:}  a separate plot of the two terms that contribute to Fig. 4 would be welcome \\
+
 \textbf{Response:}

### 2014-05-28 response2.tex
more edits

@@
 \item \textbf{Comment:} ``we assume that the time-offset has been previously handled'' $\to$ perhaps some motivation for this assumption and the consequences with respect to the system model would be in place \\
-\textbf{Response:}  There exist numerous estimators for the time offset in the literature.  We have added the sentence
+\textbf{Response:}  The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990} and is unlikely to come as a surprise to the intended audience.
+There exist numerous estimators for the time offset in the literature.
+We have added the sentence
 
@@
 
-to inform the reader of some of these methods.
-The interested reader will find further discussion about time offset estimation within these references.
-The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.
+to inform the reader of some of these methods.  The reader interested in the motivation and consequences of this assumption can consult this literature.
 
@@
 \item \textbf{Comment:} p.
-3: ``all of the exiting bounds are applicable only in the fully non-data-aided case or in the completely data aided case \dots'' --> bounds for the common scenario where there is a mixture of pilot and data symbols can be found in "N.
+3: ``all of the exiting bounds are applicable only in the fully non-data-aided case or in the completely data-aided case \dots'' $\to$ bounds for the common scenario where there is a mixture of pilot and data symbols can be found in ``N.
 Noels, H.
@@
 4578-4587, Dec.
-2005.
-\\
-\textbf{Response:} We have added this reference and have removed the sentence begining ``To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case \dots''.
+2005.'' \\
+\textbf{Response:} We have added this reference and have removed the sentences begining ``To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case \dots''.
 Note that the model considered in the paper by Noels~et.~al.~\cite{Noels_crbs_with_pilots_2005} treats the amplitude as fixed and known and includes a frequency offset term.
@@
 \item \textbf{Comment:} p. 4: ``systems such as mulitple anntannae'' $\to$ should be ``multiple antennae'' \\
-\textbf{Response:} Reviewer 1 comment~\eqref{eq:rev1:multantenna} wanted ``antenna'' and I have used this spelling.
+\textbf{Response:} Reviewer 1 comment~\eqref{eq:rev1:multantenna} wanted ``antenna'' and we have used this spelling.
 
@@
 \item \label{rev2:commentnotationsection} \textbf{Comment:} I think it would be a good idea to specify the main notational conventions at the end of the introductory section \\
-\textbf{Response:} We request not to make this change.
+\textbf{Response:} We respectfully request not to make this change.
 It is the authors' preference to introduce notation as it is used in the text rather than in a dedicated notation section.
@@
 \item \textbf{Comment:} between (5) and (6): start new paragraph \\
-\textbf{Response:} The paragraph ends at ``\dots least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}'' and the new paragraph begins with ``To see this, let $a = \rho e^{j\theta}$ \dots''
+\textbf{Response:} Done.
+The paragraph ends at ``\dots least squares estimator of the data symbols~\cite{Sweldens2001,Mackenthun1994}'' and the new paragraph begins with ``To see this, let $a = \rho e^{j\theta}$ \dots''
 
@@
 The bracketless notation is heavily employed in the statistical literature on empirical processes~\cite{Pollard_new_ways_clts_1986,Pollard_conv_stat_proc_1984,Pollard_asymp_empi_proc_1989,van2009empirical}.
-The notation would become heavily cluttered with brackets otherwise.
+The notation would become heavily cluttered otherwise.
 Our proofs in the appendix rely upon this literature.
@@
 \item \textbf{Comment:} p.
-14: "Equivalently $\hat\theta$ is the maximiser of \dots" $\to$ could you please provide some further explanation (is the maximum of $\Re(Z(\theta)$) never negative?
+14: ``Equivalently $\hat\theta$ is the maximiser of \dots'' $\to$ could you please provide some further explanation (is the maximum of $\Re(Z(\theta))$ never negative?
 and why) \\
-\textbf{Response:}
+\textbf{Response:} The least squares phase estimator is the minimiser of
+\[
+LS(\theta) = A - \frac{1}{L}\Re(Z(\theta))^2.
+\]
+Because the least squares amplitude estimator $\hat{\rho} = \Re(Z(\hat{\theta}))$ is positive by definition, we are only interested in those values of $\theta$ that correspond with a positive amplitude estimator, that is, such that
+\[
+\hat{\rho}(\theta) = \Re(Z(\theta)) > 0.
+\]
+Thus, the least squares phase estimator $\hat{\theta}$ must be such that
+\[
+\hat{\rho} = \Re(Z(\hat{\theta})) > 0.
+\]
+This is explained in the text by the sentences:
+
+``By definition the amplitude $\rho_0$ and its estimator $\hat{\rho}$ are positive.
+However, $\hat{\rho}(\theta) = \Re(Z(\theta))$ may take negative values for some $\theta \in [-\pi,\pi)$.
+The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $LS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
+Equivalently $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ with no constraints required.''
 

### 2014-05-28 code/clt.scala

@@ -45,14 +45,14 @@ for( L <- Ls; M <- Ms ) {
 
     val filep = new java.io.FileWriter("data/" + cltname + "p")
     val filea = new java.io.FileWriter("data/" + cltname + "a")
-    val fileb = new java.io.FileWriter("data/" + cltname + "b")
     val filemsea = new java.io.FileWriter("data/" + cltname + "msea")
+    val fileb = new java.io.FileWriter("data/" + cltname + "b")
       (cltlist, SNRdBs).zipped.foreach{ (mse, snr) =>
-	val (vp,va,b2,msea) = mse
+	val (vp,va,msea,b2) = mse
         filea.write(snr.toString.replace('E', 'e') + "\t" + va.toString.replace('E', 'e')  + "\n")
 	filep.write(snr.toString.replace('E', 'e') + "\t" + vp.toString.replace('E', 'e')  + "\n")
-        fileb.write(snr.toString.replace('E', 'e') + "\t" + b2.toString.replace('E', 'e')  + "\n")
         filemsea.write(snr.toString.replace('E', 'e') + "\t" + msea.toString.replace('E', 'e')  + "\n")
+        fileb.write(snr.toString.replace('E', 'e') + "\t" + b2.toString.replace('E', 'e')  + "\n")
       }
     filea.close; filep.close; fileb.close; filemsea.close //close all the files we wrote to
     println(" finished")

### 2014-05-28 paper.tex
more edits
more response edits

@@
 The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010,McKilliam_pps_unwrapping_tsp_2014}, would potentially allow this.
-More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple antenna systems, systems employing orthogonal frequency division multiplexing, and systems that must withstand rapidly time varying channels~\cite{4138046,inthft:204506}.
+More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as multiple antenna systems, systems employing orthogonal frequency division multiplexing, and systems that must withstand rapidly time varying channels~\cite{4138046,inthft:204506}.
 We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cram\'{e}r-Rao lower bounds.
@@
 %\]
-The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $LS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
+The least squares estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $LS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
 Equivalently $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ with no constraints required.
@@
 As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
+Similar conclusions are made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.
 As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e.
 $\abs{P} = L$.
-As expected, the figures indicate the estimator to become more accurate as the number of pilot symbols is increased.
-An interesting property distriguishing the case when no pilot symbols exist is the slope of the MSE curve at low SNR.
-When a finite proportion of pilot symbols exist the slope at low SNR is similar to that when all symbols are pilots.
-The slope is steeper in the case that no pilot symbols exist.
+As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
+One interesting observation is that, at low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.
 
@@
 However, $G(0)$ grows without bound as the variance of the noise increases so the bias is significant when SNR is small.
-Figure~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
+Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
 

### 2014-05-28 response2.tex
more response edits

@@
 2005.'' \\
-\textbf{Response:} We have added this reference and have removed the sentences begining ``To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case \dots''.
+\textbf{Response:} We have added this reference and have removed the sentences beginning ``To the authors' knowledge, all of the existing bounds are applicable only in the fully non-data-aided case \dots''.
 Note that the model considered in the paper by Noels~et.~al.~\cite{Noels_crbs_with_pilots_2005} treats the amplitude as fixed and known and includes a frequency offset term.
@@
 \item \textbf{Comment:} p. 4: ``systems such as mulitple anntannae'' $\to$ should be ``multiple antennae'' \\
-\textbf{Response:} Reviewer 1 comment~\eqref{eq:rev1:multantenna} wanted ``antenna'' and we have used this spelling.
+\textbf{Response:} Reviewer 1 comment~\eqref{eq:rev1:multantenna} wanted ``antenna'' and we have used this spelling. ``Multiple'' is now spelt correctly.
 
@@
 \item \textbf{Comment:} p. 5 (and further throughout the text): ``pilots symbols" (with `s') and ``pilot symbols'' (without `s') are both used $\to$ use only 1, I prefer ``pilot symbols'' \\
-\textbf{Response:} All occurences of ``pilots symbols'' have been changed to ``pilot symbols''.
+\textbf{Response:} All occurrences of ``pilots symbols'' have been changed to ``pilot symbols''.
 
@@
 
-\item \textbf{Comment:} before (6): drop the 'to' \\
+\item \textbf{Comment:} before (6): drop the `to' \\
 \textbf{Response:} Done.
@@
 
-\item \textbf{Comment:} I would prefer to see comma's between two subindices; so, `$k,i$' rather than`'$ki$' \\
+\item \textbf{Comment:} I would prefer to see comma's between two subindices; so, `$k,i$' rather than`$ki$' \\
 \textbf{Response:} Done.
@@
 \]
-Because the least squares amplitude estimator $\hat{\rho} = \Re(Z(\hat{\theta}))$ is positive by definition, we are only interested in those values of $\theta$ that correspond with a positive amplitude estimator, that is, such that
+Because the least squares amplitude estimator $\hat{\rho} = \Re(Z(\hat{\theta}))$ is positive by definition, we are only interested in those values of phase $\theta$ that correspond with a positive amplitude estimator, that is, such that
 \[
@@
 However, $\hat{\rho}(\theta) = \Re(Z(\theta))$ may take negative values for some $\theta \in [-\pi,\pi)$.
-The least square estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $LS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
+The least squares estimator $\hat{\theta}$ of $\theta_0$ is the minimiser of $LS(\theta)$ under the constraint $\hat{\rho}(\theta) = \Re(Z(\theta)) > 0$.
 Equivalently $\hat{\theta}$ is the maximiser of $\Re(Z(\theta))$ with no constraints required.''
@@
 \item \textbf{Comment:} in general, I find the numerical results section rather short \\
-\textbf{Response:} We have greatly expanded the simulation section as detailed in the responses that follow.
+\textbf{Response:} We have expanded the simulation section as detailed in the responses that follow.
 
@@
 
-``As expected, the figures indicate the estimator to become more accurate as the number of pilot symbols is increased.
-An interesting property distriguishing the case when no pilot symbols exist is the slope of the MSE curve at low SNR.
-When a finite proportion of pilot symbols exist the slope at low SNR is similar to that when all symbols are pilots.
-The slope is steeper in the case that no pilot symbols exist.''
+``As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
+One interesting observation is that, at low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.''
 
@@
 \item \textbf{Comment:} p. 18, first sentence: a qualitative explanation for the fact the theory and simulations diverge at low SNR would be appreciated  \\
-\textbf{Response:} BLERG: Added a reference to Tazras crb paper.
+\textbf{Response:} The following text has been included in Section~\ref{sec:simulations}:
+
+``As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
+Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
+Similar conclusions are made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.''
 
@@
 
-\textbf{Response:}
+\textbf{Response:} Figure~\ref{fig:biasvarcomp} in this document displays such a plot.  The number of symbols is set to $L=4096$.  The solid line shows that asymptotic squared bias 
+\[
+\rho_0^2\big(1 - G(0)\big)^2
+\]
+and the dashed line in the asymptotic variance
+\[
+\rho_0^2\frac{p B_1 + d B_2}{L}.
+\]
+The number of pilots consider is $\abs{P}=0$ and $\abs{P}=L/2$.
+The obvious relationship is observed: the bias dominates the variance at low SNR and the variance dominates the bias at high SNR.
+We believe that this plot is not particularly informative and have not included it in the manuscript.
 
 \end{enumerate}
+
+\begin{figure}[tp] 
+	\centering  
+		\includegraphics[]{code/data/biasvarplot-1.mps}
+		\caption{Plot of asymptotic bias squared (dashed) and variance (solid) of the amplitude estimator for QPSK with $L=4096$.}
+		\label{fig:biasvarcomp}
+\end{figure}
 
@@
 \\
-\textbf{Response:} We request not to make this change.
-There are two reasons.
-The first is that this notation leads to situations where there are subscripts withing superscripts.
-For example, in the proof of Theorem 1 we would have expressions of the form
+\textbf{Response:} We respectfully request not to make this change.  There are two reasons.  The first is aesthetic.  For example, in the proof of Theorem 1 we would have expressions of the form
 \[
@@
 
+{
+\small
 \input{paper.bbl}
+}
 

### 2014-06-04 paper.tex
minor

@@
 The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.
-Neverthess, Cram\'{e}r-Rao bounds have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
+Nevertheless, Cram\'{e}r-Rao bounds have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.
 Unfortunately, these bounds may not be meaningful since known non-data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.
@@
 This is related to what is called \emph{unequal error protection}~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.
-It would be reasonably straigthforward to extend our results to allow data and pilot symbols with varying constellation size.
+It would be reasonably straightforward to extend our results to allow data and pilot symbols with varying constellation size.
 A more involved extension would be the inclusion of carrier frequency as a parameter to be estimated.
@@
 where $\Im(\cdot)$ denotes the imaginary part, and $\expect$ denotes the expected value.
-Whenever there is no chance of ambiguity we write simply $\expect X$ rather than $\expect(X)$ to denote the expected value of the random variable $X$.
+Whenever there is no chance of ambiguity we write simply $\expect X$ rather than $\expect(X)$ to denote the expected value of a random variable $X$.
 For two random variables $X$ and $Y$ the notation $\expect X Y$ will always mean the expected value of the product $XY$, that is, $\expect(XY)$ and not $\expect(X)Y$.
@@
 Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
-Similar conclusions are made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.
+Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.
 As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e.
@@
 As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
-One interesting observation is that, at low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.
+One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.
+This indicates that pilot symbols are particularly important when the SNR is low.
 

### 2014-06-04 response2.tex
minor

@@
 \item \textbf{Comment:} ``the data symbols are not known to the receiver and must also be estimated'' $\to$ I would drop the second part of the sentence because (e.g.
-Viterbi and Vterbi estimator) non-data-aided estimators typically do not actually estimate the unknown data symbols \\
+Viterbi and Viterbi estimator) non-data-aided estimators typically do not actually estimate the unknown data symbols \\
 \textbf{Response:} Done.
@@
 \item \textbf{Comment:} p. 4: ``systems such as mulitple anntannae'' $\to$ should be ``multiple antennae'' \\
-\textbf{Response:} Reviewer 1 comment~\eqref{eq:rev1:multantenna} wanted ``antenna'' and we have used this spelling. ``Multiple'' is now spelt correctly.
+\textbf{Response:} Reviewer 1 comment~\eqref{eq:rev1:multantenna} wanted ``antenna'' and we have used this spelling. ``Multiple'' is now spelled correctly.
 
@@
 
-``Whenever there is no chance of ambiguity we write simply $\expect X$ rather than $\expect(X)$ to denote the expected value of the random variable $X$.
+``Whenever there is no chance of ambiguity we write simply $\expect X$ rather than $\expect(X)$ to denote the expected value of a random variable $X$.
 For two random variables $X$ and $Y$ the notation $\expect X Y$ will always mean the expected value of the product $XY$, that is, $\expect(XY)$ and not $\expect(X)Y$.''
@@
 \item \textbf{Comment:} although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text  \\
-\textbf{Response:}  We have included the follows sentences to Section~\ref{sec:simulations}:
+\textbf{Response:}  We have included the following sentences to Section~\ref{sec:simulations}:
 
 ``As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
-One interesting observation is that, at low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.''
+One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.
+This indicates that pilot symbols are particularly important when the SNR is low.''
 
@@
 Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
-Similar conclusions are made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.''
+Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.''
 
@@
 \]
-and the dashed line in the asymptotic variance
+and the dashed line is the asymptotic variance
 \[
@@
 \]
-The number of pilots consider is $\abs{P}=0$ and $\abs{P}=L/2$.
+The number of pilots considered is $\abs{P}=0$ and $\abs{P}=L/2$.
 The obvious relationship is observed: the bias dominates the variance at low SNR and the variance dominates the bias at high SNR.
-We believe that this plot is not particularly informative and have not included it in the manuscript.
+We believe that this plot is not particularly informative because the same conclusion can be drawn from Figures~\ref{fig:plotampBPSK},~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK}.
+For this reason we have not included this plot in the manuscript.
 
@@
 
-\item \textbf{Comment:}  There are typografical errors in the words  ``multiple antenna'' in line 2, page 2, section 1 \\
+\item \textbf{Comment:}  There are typographical errors in the words  ``multiple antenna'' in line 2, page 2, section 1 \\
 \textbf{Response:} Fixed.

### 2014-06-04 paper.tex
minor edits

@@
 As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
+Our asymptotic theory asserts the normality of the phase estimator as $L\to\infty$ and therefore cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
 Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.
@@
 $\abs{P} = L$.
-As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
+The phase estimator is more accurate for smaller constellation sizes, i.e.
+smaller $M$.
+The value of SNR at which the accuracy of the estimator departs from that when all symbols are pilots increases with $M$.
+The point of departure is approximately \unit[0]{dB}, \unit[6]{dB}, and \unit[13]{dB} for BPSK, QPSK and $8$-PSK respectively.
+%As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
 One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.

### 2014-06-04 response2.tex
minor edits

@@
 \item \textbf{Comment:} ``we assume that the time-offset has been previously handled'' $\to$ perhaps some motivation for this assumption and the consequences with respect to the system model would be in place \\
-\textbf{Response:}  The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990} and is unlikely to come as a surprise to the intended audience.
-There exist numerous estimators for the time offset in the literature.
-We have added the sentence
+\textbf{Response:}  The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990}.% and is unlikely to come as a surprise to the intended audience.
+There exist numerous estimators for the time offset in the literature.  We have added the sentence
 
@@
 \item \textbf{Comment:} although different figures are shown for M=2,4,8 (phase estimation), the effect of the constellation size is not addressed in the text  \\
-\textbf{Response:}  We have included the following sentences to Section~\ref{sec:simulations}:
+\textbf{Response:}  The following text is included in Section~\ref{sec:simulations}:
 
-``As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
+``As the SNR increases the variance of the estimators converge to that of the estimator where all symbols are pilots, i.e.
+$\abs{P} = L$.
+The phase estimator is more accurate for smaller constellation sizes, i.e.
+smaller $M$.
+The value of SNR at which the accuracy of the estimator departs from that when all symbols are pilots increases with $M$.
+The point of departure is approximately \unit[0]{dB}, \unit[6]{dB}, and \unit[13]{dB} for BPSK, QPSK and $8$-PSK respectively.''
+
+% ``As expected, the figures indicate that the accuracy of the estimator increases with the number of pilot symbols.
 One interesting observation is that, for low SNR, the slope of the MSE curve when no pilot symbols exist is steeper than the slope when pilot symbols do exist.
@@
 ``As the SNR decreases the variance of the phase estimator approaches that of the uniform distribution on $[-\pi, \pi)$ when $\abs{P} \neq 0$, and the uniform distribution on $[-\tfrac{\pi}{M}, \tfrac{\pi}{M})$ when $\abs{P}=0$~\cite{McKilliam_leastsqPSKnoncoICASSP_2012}.
-Our asymptotic theory cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
-Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}.''
+Our asymptotic theory asserts the normality of the phase estimator as $L\to\infty$ and therefore cannot be expected to model the behaviour of the estimator for fixed $L$ as SNR goes to zero.
+Similar observations have been made in the literature~\cite{ViterbiViterbi_phase_est_1983,6365847}''
 

### 2014-06-19 response2.tex
minr?

--- 2014-05-28 response2.tex
@@
 \item \textbf{Comment:} ``we assume that the time-offset has been previously handled'' $\to$ perhaps some motivation for this assumption and the consequences with respect to the system model would be in place \\
-\textbf{Response:}  The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990} and is unlikely to come as a surprise to the intended audience.
-There exist numerous estimators for the time offset in the literature.
-We have added the sentence
+\textbf{Response:}  %The assumption of an accurate time offset estimator is common in this literature~\cite{ViterbiViterbi_phase_est_1983,Cowley_ref_sym_carr_1998,Wilson1989,Makrakis1990,Liu1991,Mackenthun1994,Sweldens2001,McKilliamLinearTimeBlockPSK2009,Divsalar1990} and is unlikely to come as a surprise to the intended audience.
+There exist numerous estimators for the time offset in the literature.  We have added the sentence
 

### 2014-07-03 responses/proofcorrection.tex
minor

\documentclass[a4paper,10pt]{article}
%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

\usepackage{../mathbf-abbrevs}
\input{../defs}

\usepackage{zref-xr,zref-user}
\zexternaldocument*{paper}

\usepackage[left=2cm,right=2cm,top=2cm,bottom=2cm]{geometry}

\usepackage{amsmath,amsfonts,amssymb, amsthm, bm}

%\usepackage[square,comma,numbers,sort&compress]{natbib}

\usepackage{color}
\newcommand{\comment}[1]{\textcolor{red}{#1}}

\newcommand{\sgn}{\operatorname{sgn}}
\newcommand{\sinc}{\operatorname{sinc}}
\newcommand{\rect}[1]{\operatorname{rect}\left(#1\right)}

%opening
\title{Carrier phase and amplitude estimation for phase shift keying using pilots and data}
\author{Robby McKilliam, Andr\'{e} Pollok, Bill Cowley, I.\ Vaughan L.\ Clarkson and Barry Quinn  
%\thanks{}
}

\renewcommand{\theenumii}{(\alph{enumii})}
\renewcommand{\labelenumii}{\theenumii}

\begin{document}

%\maketitle

\begin{enumerate}
\item Page 1, column 2.  The formating of the first equation is wrong.  Space between $p_i$ and $i$ and $d_i$ and $i$ must be included.  It should look like:
\[
s_i = \begin{cases}
p_i & i \in P \\
d_i & i \in D,
\end{cases}
\]

% \item Page 4, bottom of column 1.  The equation beginning $Y_1 = B + \sum_{i \in D} g_{1i}$ is broken up.  This need to all be in one, i.e., like:
% \begin{align*}
% Y_1 &= B + \sum_{i \in D} g_{1i} \\
% &= B +  (e^{j2\pi/M} - 1)g_{0\sigma(0)} + \sum_{i \in D} g_{0i} \\
% &= Y_0 + \eta g_{0\sigma(0)},
% \end{align*}

\item On page 7 the section entitled ``Proof of asymptotic normality (Theorem~2)'' does not have a section number.  It should be V (i.e. the roman numeral for 5).  The sections that follow entitled ``Simulations'' and ``Conclusion'' and correspondingly numbered incorrecty.  The ``Simulations'' section should be section VI (i.e. 6) and the ``Conclusion'' section should be VII (i.e. 7).

\end{enumerate}

\end{document}

### 2014-07-03 responses/proofcorrection.tex
added more changes.  Past the sectioning error on page 7 now.

@@
 
+\item Page 4, column 2.  The phrase ``the loops on lines 1 and II'' the II should be 11 (the number eleven rather than two capital I's).  That is, ``the loops on lines 1 and 11''.
+
+\item Page 5, column 1.
+The phrase ``to denote $x$ taken `~modulo $2\pi$' into the interval \dots''.
+There is a space between the apostrophe and the word modulo.
+That is `~modulo $2\pi$' shoud be `modulo $2\pi$'.
+
+\item Page 6, column 1, start of section 4.
+The first equation starting $LS(\rho, \theta) = \dots$ has the word `where' placed in the middle of the column.
+The word `where' should be left justified rather than centered.
+It should look like:
+ \begin{align*}
+LS(\rho, \theta) &=\sum_{i \in P} \abs{ y_i - \rho e^{j\theta} p_i }^2 + \sum_{i \in D} \abs{ y_i - \rho e^{j\theta} \hat{d_i}(\theta) }^2 \\
+&= A - \rho Z(\theta) - \rho Z^*(\theta) + L \rho^2,
+\end{align*}
+where
+\[
+Z(\theta)  = \sum_{i \in P} y_i e^{-j\theta} p_i^* + \sum_{i \in D} y_i e^{-j\theta} \hat{d}_i^*(\theta),
+\]
+
+\item Page 6 column 2 second last equation starting $Z(\theta) = \dots$.
+The equation ends with a full stop.
+The full stop is put into the superscript of the equations, but it should not be.
+It should look like:
+\[
+Z(\theta) = \rho_0 \sum_{i \in P} R_i e^{j(\lambda + \Phi_i)} + \rho_0  \sum_{i \in D} R_i e^{j\sfracpart{\lambda + \Phi_i}}.
+\]
+
 \item On page 7 the section entitled ``Proof of asymptotic normality (Theorem~2)'' does not have a section number.
@@
 
+\item
+
 \end{enumerate}

### 2014-07-04 responses/proofcorrection.tex
minor

@@
 \begin{enumerate}
-\item Page 1, column 2.  The formating of the first equation is wrong.  Space between $p_i$ and $i$ and $d_i$ and $i$ must be included.  It should look like:
+\item Page 1, column 2.  The formatting of the first equation is wrong.  Space between $p_i$ and $i$, and $d_i$ and $i$ need to be included.  It should look like:
 \[
@@
 There is a space between the apostrophe and the word modulo.
-That is `~modulo $2\pi$' shoud be `modulo $2\pi$'.
+That is `~modulo $2\pi$' should be `modulo $2\pi$'.
 
-\item Page 6, column 1, start of section 4.
+\item \label{com:leftjustwhere} Page 6, column 1, start of section 4.
 The first equation starting $LS(\rho, \theta) = \dots$ has the word `where' placed in the middle of the column.
@@
 The equation ends with a full stop.
-The full stop is put into the superscript of the equations, but it should not be.
+The full stop is put into the superscript of the equation, but it should not be.
 It should look like:
@@
 the roman numeral for 5).
-The sections that follow entitled ``Simulations'' and ``Conclusion'' and correspondingly numbered incorrecty.
+The sections that follow entitled ``Simulations'' and ``Conclusion'' are correspondingly numbered incorrectly.
 The ``Simulations'' section should be section VI (i.e.
@@
 
-\item
+\item Page 10, bottom of first column, equation beginning $LS(\{r_i, i \in D\})$.  This equation would preferably not be split over multiple columns, i.e., it should all be in a single column like:
+\begin{align*}
+LS&(\{r_i, i \in D\}) \\
+&= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
+&= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
+&= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
+&= LS(\{\hat{d}_i, i \in D\}) - C,
+\end{align*}
+% If you can make it fit then:
+% \begin{align*} 
+% LS(\{r_i, i \in D\}) &= A - \frac{1}{L} \sabs{ Y(\{ r_i, i \in D\}) }^2 \\
+% &= A - \frac{1}{L} \sabs{ \hat{Y} + \eta y_k\hat{d}_k^* }^2 \\
+% &= A - \frac{1}{L}\sabs{\hat{Y}}^2 - \frac{2}{L}\Re\left(\eta \hat{Y}^* y_k \hat{d}_k^* \right) -  \frac{1}{L}\sabs{ \eta y_k}^2\\
+% &= LS(\{\hat{d}_i, i \in D\}) - C,
+% \end{align*}
+% would be even better.
+
+\item Page 10 column 2, between first and second equation.  The word `where' is centered rather than left justified.  This is similar to my comment~\ref{com:leftjustwhere}.
+
+\item Page 10 column 2, half way down.  The word `and' is centered rather than left justified.  This is similar to the above comment and my comment~\ref{com:leftjustwhere}.  It should look like:
+\begin{align*}
+\eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}} &= (e^{-j2\pi/M} - 1) e^{j \pi/M} e^{jv} \\
+&= (e^{-j\pi/M} - e^{j\pi/M}) e^{jv} \\
+&= -2 j \sin(\tfrac{\pi}{M}) e^{j v },
+\end{align*}
+and
+\[
+\Re\left( \eta e^{j\sfracpart{\hat{\lambda}_L + \Phi_k}}\right) = 2 \sin(\tfrac{\pi}{M}) \sin(v).
+\]
+
+\item Page 11, top of column 2, between first and second equations.  The word `where' is centered when it should be left justified.
 
